### 2018-08-01 paper/abstract.tex
initial files

In this paper, new contributions to testing Communicating Sequential
Processes (CSP) are presented, with focus on the generation of complete,
finite test suites. A test suite is complete if it can uncover every
conformance violation of the system under test with respect to a reference
model. Both reference models and implementation behaviours are represented as
CSP processes. As conformance relation, we consider trace equivalence and
trace refinement, as well as failures equivalence and failures refinement.
Complete black-box test suites here rely on the fact that the SUT's true
behaviour is represented by a member of a fault-domain, that is, a collection
of CSP processes that may or may not conform to the reference model. We
define fault domains by bounding the number of excessive states occurring in
a fault domain member's representation as a normalised transition graph, when
comparing it to the number of states present in the graph of the reference
model. This notion of fault domains is quite close to the way they are
defined for finite state machines, and these fault domains guarantee the
existence of {\it finite} complete test suites.

### 2018-08-01 paper/casestudies.tex
initial files

\section{Applications}
\label{sec:case}

% =======================================================================
\subsection{Testing for Trace Equivalence}

% =======================================================================
\subsection{Testing for Trace Refinement}

% =======================================================================
\subsection{Testing for Failures Equivalence}

% =======================================================================
\subsection{Testing for Failures Refinement}

% ======================================================================

### 2018-08-01 paper/conclusion.tex
initial files

\section{Conclusion}
\label{sec:conc}
% ==============================================================================

% ==============================================================================

### 2018-08-01 paper/finitecomplete-failures.tex
initial files

% ==========================================================================
\section{Finite Complete Test Suites for CSP Failures Refinment}
\label{sec:finitecompletefails}
% ==========================================================================

% ==========================================================================

### 2018-08-01 paper/introduction.tex
initial files

% =========================================================================

\section{Introduction}
\label{sec:intro}

% =================================================================================

\subsubsection*{Motivation}

Model-based testing (MBT) is an  active research field that is currently
evaluated and integrated into industrial verification processes by many
companies. This holds particularly for the embedded and cyber-physical
systems domain. While MBT is applied in different flavours, we consider the
most effective variant to be the one where test cases and concrete test data,
as well as checkers for the expected results~(\emph{test oracles}), are
automatically generated from a reference model:~it guarantees the maximal
return of investment for the time and effort invested into creating the test
model. The test suites generated in this way, however, usually have different
test strength, depending on the generation algorithms applied.

For the safety-critical domain, test suites with guaranteed fault coverage
are of particular interest. For black-box testing, guarantees can be given
only if certain hypotheses are satisfied. These hypotheses are usually
specified by a \emph{fault domain}:~a set of models that may or may not
conform to the SUT. The so-called \emph{complete} test strategies guarantee
to uncover every conformance violation of the SUT with respect to a reference
model, provided that the true SUT behaviour is captured by a member of the
fault domain.

Generation methods for complete test suites have been developed for various
modelling formalisms. In this paper, we use \emph{Communicating Sequential
Processes (CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}; this is a mature
process-algebraic approach that has been shown to be well-suited for the
description of reactive control systems in many publications over almost five
decades. Industrial success has also been reported.

% ==================================================================================

\subsubsection*{Contributions}

This paper complements work published by two of the authors
in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
collections of processes refining a  ``most general'' fault domain member.
With this concept of fault domains, complete test suites may be finite or
infinite. While this gives important insight into the theory of complete test
suites, we are particularly interested in finite suites when it comes to
their practical application.

Therefore, we present a complementary approach to the definition of CSP fault
domains in this paper. To this end, we observe that every finite-state CSP
process can be semantically represented as a finite normalised transition
graph, whose edges are labelled by the events the process engages in, and
whose nodes are labelled by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals  express the
degree of nondeterminism that is present in a given CSP process state that is
in one-one-correspondence to a node of the normalised transition graph.
Inspired by the way that fault-domains are specified for finite state
machines (FSMs), we define them here as the set of CSP processes whose
normalised transition graphs do not exceed the size of the reference model's
graph by more than a give constant.

Our main contributions in this paper are as follows.
%
\begin{enumerate}
\item It is proven that for fault domains of the described type, complete
    test suite generation methods can be given for verifying (1)~trace
    equivalence, (2)~trace refinement, (3)~failures equivalence, and
    (4)~failures refinement.

\item We prove that finite complete test suites can be generated in all
    four cases, when using the fault domains based on the size of the
    members' normalised transition graphs.

\item We present test suite generation techniques for each of the four
    conformance relations by translating algorithms originally elaborated
    for the FSM domain into the CSP world. This translation preserves the
    completeness properties that have previously been established for the
    FSM domain by other authors.
\end{enumerate}
%
The possibility to translate complete test suites between different
formalisms (here FSMs and CSP processes) has been investigated before by two
of the authors~\cite{Huang2017}. \fxwarning{alcc: I miss a paragraph that
discusses what is difficult. So far, it sounds like just application of known
results, which is not true.}

% ==================================================================================

\subsubsection*{Overview}

@todo

% ==================================================================================

### 2018-08-01 paper/preliminaries.tex
initial files

% =========================================================================
\section{Preliminaries}
\label{section:preliminaries}
% =========================================================================

In this section, we present the background material relevant to our work.

% =========================================================================
\subsection{Complete Test Suites}
\label{sec:fsmfm}
% =========================================================================

We use the term \emph{signature} to denote a collection of comparable models
represented in an arbitrary formalism. In this article, signatures represent
sets of finite state machines over fixed input and output alphabets, or CSP
processes with finite state, represented by their normalised transition
graphs (see Section~\ref{sec:ntg}).

Given a signature $Sig$  of models, a  \emph{fault model} ${\cal F} =
(M,\le,Dom)$ specifies a \emph{reference model} $M\in Sig$, a
\emph{conformance relation} $\le\ \subseteq Sig\times Sig$ between models,
and a \emph{fault domain} $Dom\subseteq Sig$. This terminology
follows~\cite{gotzhein_fault_1996}, where fault models have been originally
introduced in the context of testing of finite state machines. Note that
fault domains may contain both models conforming to the reference model and
models violating the conformance relation. Note further that the reference
model $M$ is not necessarily a member of the fault domain, although a model
of the SUT behaviour is. For example, $M$ could be nondeterministic, while
only deterministic implementation behaviours might be considered in the fault
domain.
%By $F(Sig,\le)$ we denote the set of all fault models ${\cal F}$
%defined for
%signature $Sig$ and conformance relation $\le$.

Let $\tc(Sig)$ denote the set of all \emph{test cases} applicable to elements
of $Sig$. This abstract notion of test case requires only the existence of a
relation $\pass\subseteq Sig\times \tc(Sig)$. For $(M,U)\in\pass$, the infix
notation $M\ \pass\ U$ is used, and interpreted as {\it `model $M$ passes the
test case $U$'}. If $(M,U)\not\in\pass$ holds, this is abbreviated by $M\
\tfail\ U$. Our specific notion of test cases for CSP models is elaborated in
Section~\ref{sec:csptc}.

A \emph{test suite} $\TS \subseteq \tc(Sig)$ denotes  a set of test cases. A
model $M$ \emph{passes the test suite} $\TS$, also written as $M\ \pass\
\TS$, if, and only if, $M\ \pass\ U$ for all $U\in \TS$. A test suite $\TS$
is called \emph{complete} for fault model ${\cal F} = (M,\le,Dom)$, if, and
only if, the following properties hold.
\begin{enumerate}
\item If a member $M'$ of the fault domain  conforms to the reference model $M$,
it passes the test suite, that is,
$$
\forall M'\in Dom: M'\le M \Rightarrow M'\ \pass\ \TS
$$
This property is usually called \emph{soundness} of the test suite.

\item If a member of the fault domain passes the test suite, it conforms to the reference model, that is,
$$
\forall M'\in Dom: M'\ \pass\ \TS \Rightarrow M'\le M
$$
This property is usually called \emph{exhaustiveness}.
\end{enumerate}
A test suite $\TS$ is \emph{finite} if it contains finitely many test cases and every test
case $U\in\TS$ is finite in the sense that it terminates after a finite number of steps.
It is trivial to see that, if $\TS$ is complete  for   ${\cal F} = (M,\le,Dom)$
and $Dom'\subseteq Dom$, then $\TS$ is also complete for ${\cal F}' = (M,\le,Dom')$.

% =========================================================================
\subsection{CSP and Refinement}

% =========================================================================

\subsubsection*{Communicating Sequential Processes} @todo
\fxwarning{alcc: I can make this small contribution.}

% =========================================================================
\subsubsection*{Normalised Transition Graphs for CSP Processes}
\label{sec:ntg}

As shown in~\cite{Roscoe:1994:CME:197600}, any finite-state CSP process $P$
can be represented by a \emph{normalised transition graph}
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$, if, and only if, $(n,e)$ is in
the domain of $t$, that is, there is a transition from $n$ with label $e$.
Normalisation of $G(P)$ is reflected by the fact that $t$ is a function.

A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if there
is a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
$s$. The set of traces of $P$ is denoted by $\trc(P)$. If $s\in\trc(P)$, then
the process corresponding to $P$ after having executed $s$ is denoted by
$P/s$. Since $G(P)$ is normalised, there is a unique node reached by applying
the events from $s$ one by one, starting in $\ii n$. Therefore, $G(P)/s$  is
also well defined.

By $[n]^0$ we denote the \emph{fan-out} of $n$:~the set of events occurring
as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
We also use this notation for CSP processes:~$[P]^0$ is the set
of events $P$ may engage into, in other words, the initials of $P$ after the
empty trace of events, that is, $initials(P/\langle\rangle)$ as defined
in~\cite{Roscoe2010}.

The total function $r$ maps each node $n$ to its \emph{refulsals} $r(n) =
\refs(n)$. Each element of $r(n)$ is a set of events that the CSP process $P$
might refuse to engage into, when in a process state corresponding to $n$.
The number of refusal sets in $\refs(P/s)$ specifies the degree of
nondeterminism that is present in process state $P/s$: the more refusal sets
contained in  $\refs(P/s)$, the more nondeterministic is the behaviour in
state $P/s$. If $P/s$ is deterministic, its refusals coincide with the set of
subsets of $\Sigma - [P/s]^0$, including the empty set.

For finite CSP processes, since the refusals of each process state are
subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}, $\refs(P/s)$ can be
re-constructed by knowing the set of \emph{maximal refusals}
$\maxrefs(P/s)\subseteq\refs(P/s)$. More formally, the maximal refusals
$\maxrefs(P/s)$ are defined as
$$
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
$$
Conversely, with the maximal refusals at hand, we can reconstruct the refusals by subset-closure:
$$
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
$$
To see that this approach works only for finite CSP processes, consider the
example where $\Sigma$ is infinite. In this case,
$\maxrefs(STOP/\langle\rangle)$ is empty, and so we cannot use this set to
calculate the refusals of $STOP$, that is, $\refs(STOP/\langle\rangle)$ as
defined above. As with refusals, we also use the transition graph-oriented
notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
associated with graph state $n$: if $n$ is the state reached in the
transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
= \maxrefs(P/s)$.

Well-formed normalised transition graphs must not refuse an event of the
fan-out of a state in {\it every} refusal applicable in this state; more
formally,
\begin{equation}
\label{eq:wellformedg}
\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
\exists R\in \maxrefs(n): e\not\in R
\end{equation}
%The total function $a$ maps each node to its set of \emph{minimal acceptances}:
%if $n\in N$ corresponds to a deterministic process state of $P$, $ac(n)$ contains a single acceptance $A\subseteq \Sigma$, and every $e\in A$ is in one-one-correspondence with a transition $t(n,e)$. If $n$ corresponds to a nondeterministic process state, $ac(n)$ contains at least two acceptances $A_1, A_2, \dots, A_k$. This reflects the fact that in a nondeterministic state, $P$ must accept all events of one acceptance
%$A_i, i \in \{ 1,\dots,k\}$, but may refuse all events $e$ from
%$A_j \setminus A_i, j\neq i$.
%
%Each well-defined transition graph $G(P)$ fulfils the following condition. The union of all minimal acceptances in each node corresponds to the set of events labelling its outgoing transitions.
%\begin{equation}
%\label{eq:wellformedg}
%\forall n\in N: (n,e)\in\dom~t \Leftrightarrow e\in\bigcup ac(n)
%\end{equation}
%In this condition, $\dom~t$ denotes the domain of function $t$.
By construction, normalised transition graphs reflect the \emph{failures
semantics} of finite-state CSP processes:~the traces $s$ of a process are the
sequences of transition associated with paths through the graph,
starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
r(G(P)/s)$ represent the failures $\failure(P)$ of $P$.

% .....................................................................................
 \begin{figure}
 %%\hspace*{-40mm}
 \begin{center}
\includegraphics[width=\textwidth]{q0.pdf}
\end{center}
%%\vspace*{-10mm}
\caption{Normalised transition graph of CSP process $P$ from Example~\ref{ex:a}.}
 \label{fig:tga}
 \end{figure}
% .......................................................................................

\begin{example}{ex:a}
Consider the CSP process $P$ defined below, and that the set of events
$\Sigma$ is $\{a,b,c\}$.
\begin{eqnarray*}
P & = & a \then (Q\intchoice R)
\\
Q & = & a \then P \extchoice c \then P
\\
R & = & b \then P \extchoice c \then R
\end{eqnarray*}
Its transition graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state
$P/\langle\rangle$ is represented there as Node\_0, with $\{ b,c\}$ as the
only maximal refusal, since $a$ can never be refused, and no other events are
accepted. Having engaged into $a$, the transition emanating from Node\_0
leads to Node\_2 representing  the process state $P/a = Q\intchoice R$. The
internal choice operator induces several refusal sets derived from $Q$ and
$R$. Since these processes accept their initial events in external choice,
$Q\intchoice R$ induces two maximal refusal sets $\{b\}$ and
$\{a\}$. Note that event $c$ can never be refused, since it is not a member
of any maximal refusal.

Having engaged into $c$, the next process state is represented by Node\_1.
Due to normalisation, there is only a single transition satisfying
$t(\text{Node\_2},c) = \text{Node\_1}$. This transition, however, can have
been caused by either $Q$ or $R$ engaging into $c$, so Node\_1 corresponds to
process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
two maximal refusals labelling Node\_1.

Similar considerations explain the other nodes and transitions in
Fig.~\ref{fig:tga}.

Note that the node names are generated by the FDR tool (see next paragraph).
The node numbering is generated by FDR during the normalisation procedure.
Therefore, the node numbers do not reflect the distance from the initial node
Node\_0.
%Node\_2 is a direct successor to the initial node Node\_0, while Node\_1 is a direct successor of Node\_2.
\end{example}

% =========================================================================

\subsubsection*{Tool Considerations}
The FDR tool~\cite{fdr} supports model checking and semantic analyses of CSP
processes. It provides an API~\cite{fdrmanual} that can be used to construct
normalised transition graphs for CSP processes.

The FDR graph nodes are labelled by \emph{minimal acceptances} instead of
maximal refusals as described above. Since such a minimal acceptance set is
the complement of a maximal refusal, the function $r$ mapping states to their
refusals can be implemented by creating the complements of all minimal
acceptances and then building all subsets of these complements. For practical
applications, the subset closure is never constructed in an explicit way;
instead, sets are checked with respect to containment in a maximal refusal.

%
%The maximal refusals in each process state $P/s$
% are the complements of
%the minimal acceptances of the node $n$
%corresponding to $P/s$. As a consequences, all failures
%of $P$ are represented by some $(s,R)$, where $s$ is an initialised path through the transition graph and $R\subseteq (\Sigma-A)$ for some minimal acceptance $A\in ac(n)$,
%such that $n$ is the node corresponding to $P/s$.

### 2018-08-01 additional-material/fdr/t.cpp
initial

@@ -1,0 +1,29 @@
+#include <fdr/fdr.h>
+#include <iostream>
+
+int main(int argc, char** argv)
+{
+    FDR::library_init(&argc, &argv);
+
+    try
+    {
+        FDR::Session session;
+        session.load_file("phils8.csp");
+        for (const std::shared_ptr<FDR::Assertions::Assertion>& assertion
+                : session.assertions())
+        {
+            assertion->execute(nullptr);
+            std::cout << assertion->to_string() << " "
+                << (assertion->passed() ? "Passed" : "Failed");
+        }
+    }
+    catch (const FDR::Error& error)
+    {
+        std::cout << error.what() << std::endl;
+    }
+
+    FDR::library_exit();
+
+    return 0;
+}
+

### 2018-08-02 paper/introduction.tex
Lemma about refinement and hitting sets added

@@
 
-This paper complements work published by two of the authors
+This paper presents complete black-box test suites for divergence-free\footnote{The assumption of divergence freedom is usually applied in black-box testing, since it cannot distinguish between divergence and deadlock.} CSP processes interpreted
+both in the trace   and the failures semantics.  
+Our results complement work published by two of the authors
 in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
@@
 
-Therefore, we present a complementary approach to the definition of CSP fault
-domains in this paper. To this end, we observe that every finite-state CSP
+Therefore,  a complementary approach to the definition of CSP fault
+domains is presented in this paper. To this end, we observe that every finite-state CSP
 process can be semantically represented as a finite normalised transition
@@
 Inspired by the way that fault-domains are specified for finite state
-machines (FSMs), we define them here as the set of CSP processes whose
+machines, we define them here as the set of CSP processes whose
 normalised transition graphs do not exceed the size of the reference model's
@@
 
-Our main contributions in this paper are as follows.
-%
-\begin{enumerate}
-\item It is proven that for fault domains of the described type, complete
+Our main contributions in this paper consist in the proof that 
+for fault domains of the described type, finite, complete
     test suite generation methods can be given for verifying (1)~trace
@@
 
-\item We prove that finite complete test suites can be generated in all
-    four cases, when using the fault domains based on the size of the
-    members' normalised transition graphs.
+The existence of -- possibly infinite -- complete test suites has been established for process algebras in general and  for CSP in particular by several authors, 
+e.g.~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
+To our best knowledge, however, this article is the first to present {\it finite}, complete test suites associated with this class of fault domains.
 
-\item We present test suite generation techniques for each of the four
-    conformance relations by translating algorithms originally elaborated
-    for the FSM domain into the CSP world. This translation preserves the
-    completeness properties that have previously been established for the
-    FSM domain by other authors.
-\end{enumerate}
-%
-The possibility to translate complete test suites between different
-formalisms (here FSMs and CSP processes) has been investigated before by two
-of the authors~\cite{Huang2017}. \fxwarning{alcc: I miss a paragraph that
-discusses what is difficult. So far, it sounds like just application of known
-results, which is not true.}
+It should be noted that the presented results are of a theoretical nature: despite being finite, the resulting test suite size will still be too large to be applied in practice to real-world problems.
+We discuss a number of promising options how the test suite size can be reduced to become practically applicable.
 
@@
 \subsubsection*{Overview}
+In Section~\ref{section:preliminaries}, 
+we present the background material relevant to our work.
 

### 2018-08-02 paper/preliminaries.tex
Lemma about refinement and hitting sets added

@@
 % =========================================================================
-
-In this section, we present the background material relevant to our work.
-
-% =========================================================================
-\subsection{Complete Test Suites}
-\label{sec:fsmfm}
-% =========================================================================
-
-We use the term \emph{signature} to denote a collection of comparable models
-represented in an arbitrary formalism. In this article, signatures represent
-sets of finite state machines over fixed input and output alphabets, or CSP
-processes with finite state, represented by their normalised transition
-graphs (see Section~\ref{sec:ntg}).
-
-Given a signature $Sig$  of models, a  \emph{fault model} ${\cal F} =
-(M,\le,Dom)$ specifies a \emph{reference model} $M\in Sig$, a
-\emph{conformance relation} $\le\ \subseteq Sig\times Sig$ between models,
-and a \emph{fault domain} $Dom\subseteq Sig$. This terminology
-follows~\cite{gotzhein_fault_1996}, where fault models have been originally
-introduced in the context of testing of finite state machines. Note that
-fault domains may contain both models conforming to the reference model and
-models violating the conformance relation. Note further that the reference
-model $M$ is not necessarily a member of the fault domain, although a model
-of the SUT behaviour is. For example, $M$ could be nondeterministic, while
-only deterministic implementation behaviours might be considered in the fault
-domain.
-%By $F(Sig,\le)$ we denote the set of all fault models ${\cal F}$
-%defined for
-%signature $Sig$ and conformance relation $\le$.
-
-Let $\tc(Sig)$ denote the set of all \emph{test cases} applicable to elements
-of $Sig$. This abstract notion of test case requires only the existence of a
-relation $\pass\subseteq Sig\times \tc(Sig)$. For $(M,U)\in\pass$, the infix
-notation $M\ \pass\ U$ is used, and interpreted as {\it `model $M$ passes the
-test case $U$'}. If $(M,U)\not\in\pass$ holds, this is abbreviated by $M\
-\tfail\ U$. Our specific notion of test cases for CSP models is elaborated in
-Section~\ref{sec:csptc}.
-
-A \emph{test suite} $\TS \subseteq \tc(Sig)$ denotes  a set of test cases. A
-model $M$ \emph{passes the test suite} $\TS$, also written as $M\ \pass\
-\TS$, if, and only if, $M\ \pass\ U$ for all $U\in \TS$. A test suite $\TS$
-is called \emph{complete} for fault model ${\cal F} = (M,\le,Dom)$, if, and
-only if, the following properties hold.
-\begin{enumerate}
-\item If a member $M'$ of the fault domain  conforms to the reference model $M$,
-it passes the test suite, that is,
-$$
-\forall M'\in Dom: M'\le M \Rightarrow M'\ \pass\ \TS
-$$
-This property is usually called \emph{soundness} of the test suite.
-
-\item If a member of the fault domain passes the test suite, it conforms to the reference model, that is,
-$$
-\forall M'\in Dom: M'\ \pass\ \TS \Rightarrow M'\le M
-$$
-This property is usually called \emph{exhaustiveness}.
-\end{enumerate}
-A test suite $\TS$ is \emph{finite} if it contains finitely many test cases and every test
-case $U\in\TS$ is finite in the sense that it terminates after a finite number of steps.
-It is trivial to see that, if $\TS$ is complete  for   ${\cal F} = (M,\le,Dom)$
-and $Dom'\subseteq Dom$, then $\TS$ is also complete for ${\cal F}' = (M,\le,Dom')$.
 
@@
 
+Refinement relations between finite-state CSP processes $P, Q$ 
+can be be expressed by means of their normalised transition graphs in the following way.
+
+\begin{lemma}
+\label{lemma:tgtrcref}
+\begin{eqnarray}
+P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
+\label{eq:trcrefa}
+\\
+\label{eq:failrefa}
+P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {}
+\\ & & \forall s\in\trc(G(Q)), R_Q\in\maxrefs(G(Q)/s): 
+\\ & & \tabd
+\exists R_P\in\maxrefs(G(P)/s): R_Q\subseteq R_P
+\\
+\label{eq:failrefb}
+P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {}
+\\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s):
+\\ & & \tabd
+\exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
+\end{eqnarray}
+\end{lemma}
+
 % =========================================================================
@@
 %such that $n$ is the node corresponding to $P/s$.
+
+% =========================================================================
+\subsection{Test Cases and Complete Test Suites}
+\label{sec:cspcompletedef}
+% =========================================================================
+
+@todo
+
+% =========================================================================
+\subsection{Minimal Hitting Sets}
+\label{sec:hit}
+% =========================================================================
+
+When testing for failures refinement, it will become apparent that the main idea of the underlying test strategy can be based on solving a \emph{hitting set problem}.
+Given a collection of finite sets $C = \{ A_1,\dots,A_n\}$ that are each subsets of a universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set satisfying
+\begin{equation}
+\label{eq:hit}
+\forall A\in C: H\cap C \neq\varnothing.
+\end{equation}
+
+A \emph{minimal hitting set} is a hitting set which cannot be further reduced without losing 
+the characteristinc property (\ref{eq:hit}). The problem of determining minimal hitting sets is known to be NP-hard~\cite{Book1975-BOOKRM}, but we will see below that it reduces 
+the effort of testing for failures refinement from a factor of $2^\Sigma$ to a factor that equals the number of minimal hitting sets.
+
+For testing in the failures model, the following lemmas about hitting sets are required.
+
+\begin{lemma}
+\label{lemma:hseta}
+Let $P, Q$ be two finite-state CSP processes satisfying $P\lessdet_T Q$. 
+For each $s\in\trc(P)$,
+let $\text{MINHIT}(s)$ denote the
+collection of all minimal hitting sets of $\minaccs(P/s)$. 
+Then the following statements are equivalent.
+\begin{enumerate}
+\item $P\lessdet_F Q$
+\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{MINHIT}(s)$, $H$ is
+a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
+\end{enumerate}
+\end{lemma}
+\begin{proof}
+For showing ``$1 \Rightarrow 2$'', assume that $P\lessdet_F Q$ and suppose that
+$s\in\trc(P)\cap \trc(Q)$. Then Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), implies that 
+\[
+\forall A_Q\in\minaccs(G(Q)/s):
+\exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
+\]
+Therefore, every minimal acceptance set $A_Q\in\minaccs(G(Q)/s)$ contains a
+minimal acceptance set from  \minaccs(G(P)/s). As a consequence, any hitting set
+$H \in  \text{MINHIT}(s)$ also has a non-empty intersection with every 
+$A_Q\in\minaccs(G(Q)/s)$.
+
+To prove ``$2 \Rightarrow 1$'',
+assume $P\lessdet_T Q$, but $P\not\lessdet_F Q$. According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that 
+\[
+\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q (*)
+\]
+Define 
+\[
+\overline H = \bigcup_{A_P\in\minaccs(G(P)/s)} \big( A_P \setminus A_Q \big).
+\]
+Since $A_P \setminus A_Q \neq\varnothing$ because of (*), $\overline H$ is a hitting set
+of $\minaccs(G(P)/s)$ which has a non-empty intersection with $A_Q$. Minimising 
+$\overline H$ induces the existence
+of a minimal hitting set $H\in \text{MINHIT}(s)$ which is {\it not} a hitting set of 
+of $\minaccs(G(Q)/s)$. This completes the proof of the lemma.
+\xbox
+\end{proof}
+
+%\begin{lemma}
+%\label{lemma:hsetb}
+%
+%\end{lemma}
+%\begin{proof}
+%asd
+%\end{proof}
+%

### 2018-08-03 paper/finitecomplete-failures.tex
update

@@
 % ==========================================================================
-\section{Finite Complete Test Suites for CSP Failures Refinment}
+\section{Finite Complete Test Suites for CSP Failures Refinement}
 \label{sec:finitecompletefails}
@@
 
+Given a finite-state CSP process $P$ and its normalised transition graph 
+\[
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+\]
+suppose that $V\subseteq\Sigma^*$ is a 
+prefix-closed set  of sequences of events. By $t(\ii n,V)$ we denote the set
+\[
+t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
+\] 
+of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$.
+
+\begin{lemma}
+\label{lemma:extendV}
+Let $P$ be a CSP process with normalised transition graph $G(P)$, 
+such that all states in $N$ are reachable.
+Let $V\subseteq\Sigma^*$ be a finite prefix-closed set of
+sequences of events. Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is,
+$|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended by an event of $\Sigma$.
+Then $G(P)$ reaches at least $(k+1)$ nodes under $V\cup V.\Sigma$.
+\end{lemma}
+\begin{proof}
+Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable, there exists
+a trace $s$ such that $G(P)/s = n'$.
+Decompose $s = s_1.e.s_2$ with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and $G(P)/s_1.e \not\in t(\ii n,V)$.
+Such a decomposition always exists, because $V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
+Note, however, that it is not necessarily the case that $s_1\in V$.
+
+Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such that
+$\ol n = G(P)/u = G(P)/s_1$. Since $s = s_1.e.s_2$ is a trace of $P$ and $n = G(P)/s_1$,
+$(\ol u,e)$ is in the domain of $t$, therefore,  
+$n = G(P)/u.e = G(P)/s_1.e \not\in t(\ii n,V)$ is a well-defined node
+of $N$ which is not contained in $t(\ii n,V)$. Since $u.e\in V\cup V.\Sigma$,
+$G(P)$ reaches at least the additional node $n$ under $V\cup V.\Sigma$. This completes the proof.
+\xbox
+\end{proof}
+
 % ==========================================================================
+\subsection{Test Cases for Verifying CSP Failures Refinement}
+
+For a given reference process $P$ and for each $n\ge 0$, 
+we define a CSP test process as follows. 
+\begin{eqnarray}
+U_F(n) & = & U_F(n,\varepsilon)
+\\
+U_F(n,s) & = & \big( e:(\Sigma - [P/s]^0) \then \dag\then \Skip \big)
+\label{eq:ufa}
+\\ & & \extchoice \nonumber
+\\ & & ([P/s]^0 = \varnothing)    \&   \big( \omega \then \Skip \big)
+\label{eq:ufb}
+\\ & & \extchoice \nonumber
+\\ & & (\#s < n) \& \big( e:[P/s]^0 \then U_F(n,s.e) \big)
+\label{eq:ufc}
+\\ & & \extchoice \nonumber
+\\ & & (\#s = n) \& \big( \sqcap_{H\in\text{MINHIT}(P/s)} ( e:H \then \omega \then\Skip   )  \big)
+\label{eq:ufd}
+\end{eqnarray}
+
+A test is performed by running $U_F(n)$ concurrently with any SUT process $Q$
+that operates on the same alphabet as $P$, synchronising over
+alphabet $\Sigma$. Therefore, a test execution is any trace of the concurrent process
+\[
+Q\parallel[\Sigma] U_F(n).
+\]
+It is assumed that the events $\dag$ and $\omega$ denoting FAIL and PASS of the test execution, respectively, are events outside $\Sigma$.
+Since we assume that $Q$ is free of livelocks, it is guaranteed that each test execution terminates after some $s\in\trc(P)$ with length $(n+1)$ at the latest. 
+The test is \emph{passed} by the SUT (we write $Q\ \pass\ U(n)$) if and only if {\it every} execution of $Q\parallel[\Sigma] U_F(n)$ terminates with PASS event $\omega$.
+This can also be  expressed by means of a trace equivalence relation:
+\[
+Q\ \pass\ U(n) \equiv  (Q\parallel[\Sigma] U_F(n)) \hide \Sigma =_T (\omega\then\Skip)
+\]
+This type of pass relation is often called \emph{must test}, because every test execution must end with the PASS event $\omega$ \cite{Hennessy:1988:ATP:50497}.
+
+Intuitively speaking, $U_F(n)$ is able to perform any trace $s$ of $P$, 
+up to a length of $n$. If, after having already run through $s\in\trc(P)$ with $\#s < n$,
+an event is accepted by the SUT which is outside the initials of $P/s$, the test 
+immediately terminates with FAIL-event $\dag$. This is handled by the first operand of the
+external choice in sub-process $U(n,s)$.
+
+If $P/s$ is the STOP-process, this is revealed by its initials being empty. In this case, the test may terminate 
+successfully (second operand of the external choice in $U(n,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
+Therefore,
+if the SUT would still accept an event in a state where $P/s$ is supposed to have stopped, there exists a test execution which terminates with FAIL by choosing the first
+operand of the external choice.
+
+If the length of $s$ is still less than $n$, the test accepts any event from the initials
+$[P/s]^0$ and continues recursively as sub-process $U(n,s.e)$ (third operand of the external choice).
+A test of this type is called \emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts its consecutive behaviour to the event selected by the SUT.
+
+After having  run successfully 
+through a trace of length $n$, the test changes its behaviour:
+instead of offering {\it all} legal events from $[P/s]^0$ to the SUT, 
+it nondeterministically chooses 
+a minimal hitting set of $\minaccs(P/s)$ and only offers the events contained in this set.
+If the SUT refuses to engage into any of these events, this reveals a violation of the
+failures refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept
+at least one event of each minimal hitting set in $\text{MINHIT}(P/s)$. Therefore, the test
+only terminates with success $\omega$, if such an event is accepted by the SUT.
+
+After this informal explanation of tests representing adaptive test cases, we are ready to prove the main theorem of this paper.
+
+\begin{theorem}
+Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
+whose normalised transition graph $G(P)$ has $n$ states. Define fault domain ${\cal D}$ as
+the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
+has at most $m$ states with $m \ge n$. 
+Then the test suite 
+\[
+\TS_F = \{ U_F(k)~|~0 \le k \le nm  \}
+\]
+is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
+\end{theorem}
+\begin{proof}
+To prove soundness of $\TS_F$, suppose that $Q\in{\cal D} \wedge P\lessdet_F Q$. Then
+$\trc(Q)\subseteq \trc(P)$, and Lemma~\ref{lemma:hseta} implies that 
+for all traces $s$ of $Q$, every $H\in\text{MINHIT}(P/s)$ is a 
+    hitting set for $\minaccs(Q/s)$. As a consequence, when running in parallel
+with $Q$, any adaptive test $U_F(n)$ will always enter the branches 
+(\ref{eq:ufb}), (\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction
+for $U_F(n,s)$. Branch (\ref{eq:ufb}) always leads to a PASS verdict, branch (\ref{eq:ufc})
+always to test continuation without a verdict. For the last branch, we note that 
+any selected minimal hitting set $H\in\text{MINHIT}(P/s)$ has a non-empty intersection with
+each of the minimal acceptances of $Q/s$.
+As a consequence, $Q/s$ will never block when offered events from $H$, and the test terminates with PASS event $\omega$.
+Note that this argument requires that $Q$ is free of livelocks, because otherwise the PASS-events might not become visible, due to unbounded sequences of hidden events performed by $Q$.
+Note further, that this proof did not refer in any way to the size of $Q$'s normalised transition graph, so the test suite is sound for {\it all} non-divergent CSP process refining $P$.
+
+To prove exhaustiveness, consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
+\begin{enumerate}
+\item $\trc(Q)\not\subseteq \trc(P)$
+\item There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
+of $\minaccs(Q/s)$, such that 
+(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
+\[
+\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
+\] 
+\end{enumerate}
+It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U(k))$ with $k\le nm$ ends with the FAIL event $\dag$.
+\xbox
+\end{proof}
+
+% ==========================================================================

### 2018-08-03 paper/preliminaries.tex
update

@@
 \fxwarning{alcc: I can make this small contribution.}
+
+Throughout this paper, the alphabet of CSP processes $P$ is denoted by $\Sigma$. 
+Since we are only considering ``classical'' CSP processes, the alphabet is always finite.
 
@@
 
-By $[n]^0$ we denote the \emph{fan-out} of $n$:~the set of events occurring
+By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
 as labels in any outgoing transitions.
@@
 $\maxrefs(P/s)$ are defined as
-$$
+\begin{equation}
 \maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
-$$
+\end{equation}
 Conversely, with the maximal refusals at hand, we can reconstruct the refusals by subset-closure:
-$$
+\begin{equation}
 \refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
-$$
+\end{equation}
 To see that this approach works only for finite CSP processes, consider the
@@
 \end{equation}
-%The total function $a$ maps each node to its set of \emph{minimal acceptances}:
-%if $n\in N$ corresponds to a deterministic process state of $P$, $ac(n)$ contains a single acceptance $A\subseteq \Sigma$, and every $e\in A$ is in one-one-correspondence with a transition $t(n,e)$.
-If $n$ corresponds to a nondeterministic process state, $ac(n)$ contains at least two acceptances $A_1, A_2, \dots, A_k$.
-This reflects the fact that in a nondeterministic state, $P$ must accept all events of one acceptance
-%$A_i, i \in \{ 1,\dots,k\}$, but may refuse all events $e$ from
-%$A_j \setminus A_i, j\neq i$.
-%
-%Each well-defined transition graph $G(P)$ fulfils the following condition. The union of all minimal acceptances in each node corresponds to the set of events labelling its outgoing transitions.
-%\begin{equation}
-%\label{eq:wellformedg}
-%\forall n\in N: (n,e)\in\dom~t \Leftrightarrow e\in\bigcup ac(n)
-%\end{equation}
-%In this condition, $\dom~t$ denotes the domain of function $t$.
+
 By construction, normalised transition graphs reflect the \emph{failures
@@
 r(G(P)/s)$ represent the failures $\failure(P)$ of $P$.
+
+When investigating  tests for failures refinement, 
+the notion of \emph{acceptances}~\cite{Hennessy:1988:ATP:50497} 
+which is dual to refusals will also be useful. An acceptance set of $P/s$ is a subset
+of the initials $[P/s]^0$, i.e. a subset of events labelling  outgoing transitions
+of $G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance equals
+$[P/s]^0$, because $P/s$ will never refuse any of the events contained in this set. If
+$P/s$ is nondeterministic, it will internally choose one of it 
+\emph{minimal acceptance sets} $A$ and never refuse any event in $A$, while refusing the
+events in $\Sigma-A$. The acceptances of 
+ $P/s$ are denoted by $\accs(P/s)$, the minimal acceptances by $\minaccs(P/s)$. 
+ They fulfil the following rules.
+\begin{eqnarray}
+A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
+\label{eq:accref}
+\\
+\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
+\label{eq:accinitials}
+\\
+ X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
+ \label{eq:accsubset}
+\end{eqnarray} 
+Note that (\ref{eq:accref}) can be regarded as a definition of minimal acceptances by means
+of maximal refusals. Then (\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences
+of this definition and  the fact that refusals are subset-closed.
+
+Every node of a transition graph can be labelled with with its minimal acceptances, and this information is equivalent to 
+that contained in its maximal refusals.
 
@@
 \label{eq:failrefa}
-P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {}
-\\ & & \forall s\in\trc(G(Q)), R_Q\in\maxrefs(G(Q)/s): 
+P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
+\\ & & \forall s\in\trc(G(Q)), R_Q\in\maxrefs(G(Q)/s):  \nonumber
 \\ & & \tabd
@@
 \label{eq:failrefb}
-P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {}
-\\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s):
+P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
+\\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
 \\ & & \tabd
@@
 \end{lemma}
+
+For proving our main theorems, it is necessary to consider the \emph{product} of 
+normalised transition graphs. We need this only for the investigation of corresponding traces in reference processes and processes under test. Therefore, the 
+labelling of nodes with maximal refusals or minimal acceptances   will be disregard in the 
+product construction.
+Consider two normalised transition graphs 
+\[
+G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
+\]
+over the same alphabet $\Sigma$. Then their product is defined by
+\begin{eqnarray}
+G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
+\\
+\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma~|   \nonumber
+\\ & & \tabc
+(n_1,e)\in\dom~t_1\wedge 
+(n_2,e) \in\dom~t_2    \}
+\\
+t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
+\end{eqnarray}
 
@@
 instead, sets are checked with respect to containment in a maximal refusal.
+
+@todo
 
@@
 For each $s\in\trc(P)$,
-let $\text{MINHIT}(s)$ denote the
+let $\text{MINHIT}(P/s)$ denote the
 collection of all minimal hitting sets of $\minaccs(P/s)$. 
@@
 \item $P\lessdet_F Q$
-\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{MINHIT}(s)$, $H$ is
+\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{MINHIT}(P/s)$, $H$ is
 a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
@@
 For showing ``$1 \Rightarrow 2$'', assume that $P\lessdet_F Q$ and suppose that
-$s\in\trc(P)\cap \trc(Q)$. Then Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), implies that 
+$s\in\trc(P)\cap \trc(Q)$. Then Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), 
+states that 
 \[
@@
 \]
-Therefore, every minimal acceptance set $A_Q\in\minaccs(G(Q)/s)$ contains a
-minimal acceptance set from  \minaccs(G(P)/s). As a consequence, any hitting set
-$H \in  \text{MINHIT}(s)$ also has a non-empty intersection with every 
-$A_Q\in\minaccs(G(Q)/s)$.
+Therefore,
+$H \in  \text{MINHIT}(P/s)$ not only implies $H\cap A_P\neq\varnothing$ for all minimal
+acceptances $A_P$, but also
+$H\cap A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq A_Q$ for at least one $A_P$.
+As a consequence, each $H \in  \text{MINHIT}(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
 
@@
 \[
-\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q (*)
+\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q 
+\qquad (*)
 \]
@@
 \[
-\overline H = \bigcup_{A_P\in\minaccs(G(P)/s)} \big( A_P \setminus A_Q \big).
+\overline H = \bigcup\{ A_P \setminus A_Q~|~A_P\in\minaccs(G(P)/s) \}.
 \]
-Since $A_P \setminus A_Q \neq\varnothing$ because of (*), $\overline H$ is a hitting set
-of $\minaccs(G(P)/s)$ which has a non-empty intersection with $A_Q$. Minimising 
+Since $A_P \setminus A_Q \neq\varnothing$ for all $A_P$ because of (*), $\overline H$ is a hitting set
+of $\minaccs(G(P)/s)$ which has an  empty intersection with $A_Q$. Minimising 
 $\overline H$ induces the existence
-of a minimal hitting set $H\in \text{MINHIT}(s)$ which is {\it not} a hitting set of 
+of a minimal hitting set $H\in \text{MINHIT}(P/s)$ which is {\it not} a hitting set of 
 of $\minaccs(G(Q)/s)$. This completes the proof of the lemma.

### 2018-08-04 paper/finitecomplete-failures.tex
First complete version of Theorem 1 (complete finite test suite for failures refinement)

@@
 To prove exhaustiveness, consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
-\begin{enumerate}
-\item $\trc(Q)\not\subseteq \trc(P)$
-\item There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
+\begin{description}
+\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
+\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
 of $\minaccs(Q/s)$, such that 
 (see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
-\[
+\begin{equation}
+\label{eq:accsnotcontained}
 \forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
-\] 
-\end{enumerate}
+\end{equation} 
+\end{description}
 It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U(k))$ with $k\le nm$ ends with the FAIL event $\dag$.
+
+For the first case, consider a  trace $s.e\in\trc(Q)$ such that $s\in\trc(P)$, but 
+$s.e\not\in\trc(P)$.
+Then $s$ is also a trace of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}.
+From the construction of $G$ described there, we know that $G$ has at most $nm$ reachable states, because $G(P)$ has $n$ states, and $G(Q)$ has at most $m$ states.
+Suppose that $G/s = (n_P,n_Q)$.
+Applying Lemma~\ref{lemma:extendV} implies that this state can be reached by a trace $u\in\trc(G)$ of length $\#u\le nm-1$.
+Now the construction of the transition function of $G$ implies that $u$ is also a trace of $P$ and $Q$.
+Since test $U_F(nm)$ accepts all traces of $P$ up to length $nm$, $u$ is also a trace of this test, and, by construction, $U_F(nm)/u = U_F(nm,u)$.
+Since $s.e\not\in\trc(P)$,
+$e$ is an element of $\Sigma-[P/u]^0$.
+Therefore, $U_F(nm,u)$ can only execute its first branch (\ref{eq:ufa}), so that the test FAILS with event $\dag$.
+Again, the assumption of non-divergence of Q is needed for this conclusion.
+
+For the Case~2, we note that trace $s$ is again a trace of the product graph $G$. Therefore, $G/s$ can again be reached by a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length
+$\#u\le nm-1$. Consider test $U_F(\# u)$ which satisfies $U_F(\# u)/u = U_F(\#u,u)$. This
+test may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd}) only: 
+Assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has at least one non-empty minimal acceptance, so the guard condition $([P/s]^0 = \varnothing)$ of branch
+ (\ref{eq:ufb}) evaluates to $\isf$ for $U_F(\#u,u)$. Moreover, 
+ the guard condition $(\#s < n)$
+for branch (\ref{eq:ufc}) evaluates to $\isf$ for $U_F(\#u,u)$, too. 
+If branch (\ref{eq:ufa}) is executed, the test always fails. If branch (\ref{eq:ufd})
+is executed, the test fails for the execution where a minimal hitting set 
+$H\in\text{MINHIT}(P/u)$ is chosen by $U_F(\#u,u)$ which has an empty intersection
+with the minimal acceptance $A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an $H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there exists 
+a test execution   where $Q/u$ selects acceptance $A_Q$ and  $U_F(\#u,u)$ 
+selects $H$. This execution deadlocks in 
+process state $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot 
+produce the PASS event $\omega$; this  means that the test fails. This concludes the proof.
 \xbox

### 2018-08-06 paper/finitecomplete-failures.tex
Small updates in U_F and explanations

@@
 \\
-U_F(n,s) & = & \big( e:(\Sigma - [P/s]^0) \then \dag\then \Skip \big)
+U_F(n,s) & = & \big( e:(\Sigma - [P/s]^0) \then \dag\then \Stop \big)
 \label{eq:ufa}
 \\ & & \extchoice \nonumber
-\\ & & ([P/s]^0 = \varnothing)    \&   \big( \omega \then \Skip \big)
+\\ & & ([P/s]^0 = \varnothing)    \&   \big( \omega \then \Stop \big)
 \label{eq:ufb}
@@
 \\ & & \extchoice \nonumber
-\\ & & (\#s = n) \& \big( \sqcap_{H\in\text{MINHIT}(P/s)} ( e:H \then \omega \then\Skip   )  \big)
+\\ & & (\#s = n) \& \big( \sqcap_{H\in\text{MINHIT}(P/s)} ( e:H \then \omega \then\Stop   )  \big)
 \label{eq:ufd}
@@
 \[
-Q\ \pass\ U(n) \equiv  (Q\parallel[\Sigma] U_F(n)) \hide \Sigma =_T (\omega\then\Skip)
+Q\ \pass\ U(n) \equiv (\omega\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(n)) \hide \Sigma  
 \]
-This type of pass relation is often called \emph{must test}, because every test execution must end with the PASS event $\omega$ \cite{Hennessy:1988:ATP:50497}.
+This type of pass relation is often called \emph{must test}, because every test execution must end with the PASS event $\omega$ \cite{Hennessy:1988:ATP:50497}. Note that it
+is necessary to use the failures refinement relation in this condition, and not the trace refinement relation: process $(Q\parallel[\Sigma] U_F(n)) \hide \Sigma$ may have  the same visible traces $\varepsilon, \omega$ as the ``Test Passed Process'' $(\omega\then\Stop)$.
+However,
+the former may nondeterministically refuse $\omega$, due to a deadlock occurring when
+a faulty SUT processes executes concurrently with $U_F(n,s)$ executing branch (\ref{eq:ufd}). This is explained further in the next paragraphs.
 

### 2018-08-06 paper/finitecomplete-failures.tex
Improvements suggested by Wen-ling

--- 2018-08-03 paper/finitecomplete-failures.tex
@@
 sequences of events. Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is,
-$|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended by an event of $\Sigma$.
+$|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, 
+extended by any event of $\Sigma$.
 Then $G(P)$ reaches at least $(k+1)$ nodes under $V\cup V.\Sigma$.
@@
 Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such that
-$\ol n = G(P)/u = G(P)/s_1$. Since $s = s_1.e.s_2$ is a trace of $P$ and $n = G(P)/s_1$,
-$(\ol u,e)$ is in the domain of $t$, therefore,  
+$\ol n = G(P)/u = G(P)/s_1$. Since $s = s_1.e.s_2$ is a trace of $P$ and 
+$\ol n = G(P)/s_1$,
+$(\ol n,e)$ is in the domain of $t$. Therefore,  
 $n = G(P)/u.e = G(P)/s_1.e \not\in t(\ii n,V)$ is a well-defined node
@@
 
-For a given reference process $P$ and for each $n\ge 0$, 
+For a given reference process $P$ and for each integer $p\ge 0$, 
 we define a CSP test process as follows. 
 \begin{eqnarray}
-U_F(n) & = & U_F(n,\varepsilon)
+U_F(p) & = & U_F(p,\varepsilon)
 \\
-U_F(n,s) & = & \big( e:(\Sigma - [P/s]^0) \then \dag\then \Skip \big)
+U_F(p,s) & = & \big( e:(\Sigma - [P/s]^0) \then \efail\then \Stop \big)
 \label{eq:ufa}
 \\ & & \extchoice \nonumber
-\\ & & ([P/s]^0 = \varnothing)    \&   \big( \omega \then \Skip \big)
+\\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
 \label{eq:ufb}
 \\ & & \extchoice \nonumber
-\\ & & (\#s < n) \& \big( e:[P/s]^0 \then U_F(n,s.e) \big)
+\\ & & (\#s < p) \& \big( e:[P/s]^0 \then U_F(p,s.e) \big)
 \label{eq:ufc}
 \\ & & \extchoice \nonumber
-\\ & & (\#s = n) \& \big( \sqcap_{H\in\text{MINHIT}(P/s)} ( e:H \then \omega \then\Skip   )  \big)
+\\ & & (\#s = p) \& \big( \sqcap_{H\in\text{minHit}(P/s)} ( e:H \then \epass \then\Stop   )  \big)
 \label{eq:ufd}
@@
 
-A test is performed by running $U_F(n)$ concurrently with any SUT process $Q$
+A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$
 that operates on the same alphabet as $P$, synchronising over
@@
 \[
-Q\parallel[\Sigma] U_F(n).
+Q\parallel[\Sigma] U_F(p).
 \]
-It is assumed that the events $\dag$ and $\omega$ denoting FAIL and PASS of the test execution, respectively, are events outside $\Sigma$.
-Since we assume that $Q$ is free of livelocks, it is guaranteed that each test execution terminates after some $s\in\trc(P)$ with length $(n+1)$ at the latest. 
-The test is \emph{passed} by the SUT (we write $Q\ \pass\ U(n)$) if and only if {\it every} execution of $Q\parallel[\Sigma] U_F(n)$ terminates with PASS event $\omega$.
-This can also be  expressed by means of a trace equivalence relation:
+It is assumed that the events $\efail$ and $\epass$ denoting FAIL and PASS of the test execution, respectively, are events outside $\Sigma$.
+Since we assume that $Q$ is free of livelocks, it is guaranteed that each test execution terminates after some $s\in\trc(P)$ with length $(p+1)$ at the latest. 
+The test is \emph{passed} by the SUT (we write $Q\ \pass\ U_F(p)$) if and only if {\it every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with PASS event $\epass$.
+This can also be  expressed by means of a failures refinement relation:
 \[
-Q\ \pass\ U(n) \equiv  (Q\parallel[\Sigma] U_F(n)) \hide \Sigma =_T (\omega\then\Skip)
+Q\ \pass\ U_F(p) \equiv (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma  
 \]
-This type of pass relation is often called \emph{must test}, because every test execution must end with the PASS event $\omega$ \cite{Hennessy:1988:ATP:50497}.
+This type of pass relation is often called \emph{must test}, because every test execution must end with the $\epass$-event \cite{Hennessy:1988:ATP:50497}. Note that it
+is necessary to use the failures refinement relation in this condition, and not the trace refinement relation: process $(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same visible traces $\varepsilon, \epass$ as the ``Test Passed Process'' $(\epass\then\Stop)$.
+However,
+the former may nondeterministically refuse $\epass$, due to a deadlock occurring when
+a faulty SUT process executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
+because $\#s = p$. This is explained further in the next paragraphs.
 
-Intuitively speaking, $U_F(n)$ is able to perform any trace $s$ of $P$, 
-up to a length of $n$. If, after having already run through $s\in\trc(P)$ with $\#s < n$,
+Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, 
+up to a length of $p$. 
+If, after having already run through $s\in\trc(P)$ with $\#s < p$,
 an event is accepted by the SUT which is outside the initials of $P/s$, the test 
-immediately terminates with FAIL-event $\dag$. This is handled by the first operand of the
-external choice in sub-process $U(n,s)$.
+immediately terminates with FAIL-event $\efail$. This is handled by the first operand of the
+external choice in sub-process $U_F(p,s)$.
 
 If $P/s$ is the STOP-process, this is revealed by its initials being empty. In this case, the test may terminate 
-successfully (second operand of the external choice in $U(n,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
+successfully (second operand of the external choice in $UF(p,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
 Therefore,
@@
 
-If the length of $s$ is still less than $n$, the test accepts any event from the initials
-$[P/s]^0$ and continues recursively as sub-process $U(n,s.e)$ (third operand of the external choice).
+If the length of $s$ is still less than $p$, the test accepts any event from the initials
+$[P/s]^0$ and continues recursively as sub-process $U(p,s.e)$ (third operand of the external choice).
 A test of this type is called \emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts its consecutive behaviour to the event selected by the SUT.
@@
 After having  run successfully 
-through a trace of length $n$, the test changes its behaviour:
+through a trace of length $p$, the test changes its behaviour:
 instead of offering {\it all} legal events from $[P/s]^0$ to the SUT, 
@@
 failures refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept
-at least one event of each minimal hitting set in $\text{MINHIT}(P/s)$. Therefore, the test
-only terminates with success $\omega$, if such an event is accepted by the SUT.
+at least one event of each minimal hitting set in $\text{minHit}(P/s)$. Therefore, the test
+only terminates with success $\epass$, if such an event is accepted by the SUT.
 
@@
 Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
-whose normalised transition graph $G(P)$ has $n$ states. Define fault domain ${\cal D}$ as
+whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
 the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
-has at most $m$ states with $m \ge n$. 
+has at most $q$ states with $q \ge p$. 
 Then the test suite 
 \[
-\TS_F = \{ U_F(k)~|~0 \le k \le nm  \}
+\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
 \]
@@
 $\trc(Q)\subseteq \trc(P)$, and Lemma~\ref{lemma:hseta} implies that 
-for all traces $s$ of $Q$, every $H\in\text{MINHIT}(P/s)$ is a 
+for all traces $s$ of $Q$, every $H\in\text{minHit}(P/s)$ is a 
     hitting set for $\minaccs(Q/s)$. As a consequence, when running in parallel
-with $Q$, any adaptive test $U_F(n)$ will always enter the branches 
+with $Q$, any adaptive test $U_F(p)$ will always enter the branches 
 (\ref{eq:ufb}), (\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction
-for $U_F(n,s)$. Branch (\ref{eq:ufb}) always leads to a PASS verdict, branch (\ref{eq:ufc})
+for $U_F(p,s)$. Branch (\ref{eq:ufb}) always leads to a PASS verdict, branch (\ref{eq:ufc})
 always to test continuation without a verdict. For the last branch, we note that 
-any selected minimal hitting set $H\in\text{MINHIT}(P/s)$ has a non-empty intersection with
+any selected minimal hitting set $H\in\text{minHit}(P/s)$ has a non-empty intersection with
 each of the minimal acceptances of $Q/s$.
-As a consequence, $Q/s$ will never block when offered events from $H$, and the test terminates with PASS event $\omega$.
+As a consequence, $Q/s$ will never block when offered events from $H$, and the test terminates with PASS event $\epass$.
 Note that this argument requires that $Q$ is free of livelocks, because otherwise the PASS-events might not become visible, due to unbounded sequences of hidden events performed by $Q$.
@@
 To prove exhaustiveness, consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
-\begin{enumerate}
-\item $\trc(Q)\not\subseteq \trc(P)$
-\item There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
+\begin{description}
+\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
+\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
 of $\minaccs(Q/s)$, such that 
 (see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
-\[
+\begin{equation}
+\label{eq:accsnotcontained}
 \forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
-\] 
-\end{enumerate}
-It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U(k))$ with $k\le nm$ ends with the FAIL event $\dag$.
+\end{equation} 
+\end{description}
+It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U(k))$ with $k < pq$ ends with the FAIL event $\efail$.
+
+For the first case, consider a  trace $s.e\in\trc(Q)$ such that $s\in\trc(P)$, but 
+$s.e\not\in\trc(P)$.
+Then $s$ is also a trace of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}.
+From the construction of $G$ described there, we know that $G$ has at most $pq$ reachable states, because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states.
+Suppose that $G/s = (n_P,n_Q)$.
+Applying Lemma~\ref{lemma:extendV} implies that this state can be reached by a trace $u\in\trc(G)$ of length $\#u < pq$.
+Now the construction of the transition function of $G$ implies that $u$ is also a trace of $P$ and $Q$.
+Since test $U_F(pq-1)$ accepts all traces of $P$ up to length $pq$, $u$ is also a trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$.
+Since $s.e\not\in\trc(P)$,
+$e$ is an element of $\Sigma-[P/u]^0$.
+Therefore, in at least one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa}) with this event $e$, so that the test fails with event $\efail$.
+Again, the assumption of non-divergence of Q is needed for this conclusion.
+
+For the Case~2, we note that trace $s$ is again a trace of the product graph $G$. Therefore, $G/s$ can again be reached by a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length
+$\#u < pq$. Consider test $U_F(\# u)$ which satisfies $U_F(\# u)/u = U_F(\#u,u)$, because
+it always performs branch (\ref{eq:ufc}) until the trace $u$ has been completely processed. $U_F(\#u,u)$
+  may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd}) only: 
+Assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has at least one non-empty minimal acceptance, so the guard condition $([P/s]^0 = \varnothing)$ of branch
+ (\ref{eq:ufb}) evaluates to $\isf$ for $U_F(\#u,u)$. Moreover, 
+ the guard condition $(\#s < p)$
+for branch (\ref{eq:ufc}) evaluates to $\isf$ for $U_F(\#u,u)$, too. 
+If branch (\ref{eq:ufa}) is executed, the test always fails. If branch (\ref{eq:ufd})
+is executed, the test fails for the execution where a minimal hitting set 
+$H\in\text{minHit}(P/u)$ is chosen by $U_F(\#u,u)$ which has an empty intersection
+with the minimal acceptance $A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an $H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there exists 
+a test execution   where $Q/u$ selects acceptance $A_Q$ and  $U_F(\#u,u)$ 
+selects $H$. This execution deadlocks in 
+process state $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot 
+produce the pass event $\epass$; this  means that the test fails. This concludes the proof.
 \xbox

### 2018-08-06 paper/preliminaries.tex
Improvements suggested by Wen-ling

@@
 
-\begin{example}{ex:a}
+\begin{example}\label{ex:a}
 Consider the CSP process $P$ defined below, and that the set of events
@@
 When testing for failures refinement, it will become apparent that the main idea of the underlying test strategy can be based on solving a \emph{hitting set problem}.
-Given a collection of finite sets $C = \{ A_1,\dots,A_n\}$ that are each subsets of a universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set satisfying
+Given a collection of finite sets $C = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a
+subset of a universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set satisfying
 \begin{equation}
 \label{eq:hit}
-\forall A\in C: H\cap C \neq\varnothing.
+\forall A\in C: H\cap A \neq\varnothing.
 \end{equation}
@@
 A \emph{minimal hitting set} is a hitting set which cannot be further reduced without losing 
-the characteristinc property (\ref{eq:hit}). The problem of determining minimal hitting sets is known to be NP-hard~\cite{Book1975-BOOKRM}, but we will see below that it reduces 
+the characteristic property (\ref{eq:hit}). The problem of determining minimal hitting sets is known to be NP-hard~\cite{Book1975-BOOKRM}, but we will see below that it reduces 
 the effort of testing for failures refinement from a factor of $2^\Sigma$ to a factor that equals the number of minimal hitting sets.
@@
 For each $s\in\trc(P)$,
-let $\text{MINHIT}(P/s)$ denote the
+let $\text{minHit}(P/s)$ denote the
 collection of all minimal hitting sets of $\minaccs(P/s)$. 
@@
 \item $P\lessdet_F Q$
-\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{MINHIT}(P/s)$, $H$ is
+\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{minHit}(P/s)$, $H$ is
 a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
@@
 Therefore,
-$H \in  \text{MINHIT}(P/s)$ not only implies $H\cap A_P\neq\varnothing$ for all minimal
+$H \in  \text{minHit}(P/s)$ not only implies $H\cap A_P\neq\varnothing$ for all minimal
 acceptances $A_P$, but also
 $H\cap A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq A_Q$ for at least one $A_P$.
-As a consequence, each $H \in  \text{MINHIT}(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
+As a consequence, each $H \in  \text{minHit}(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
 
@@
 \]
+Let $A$ be such a set $A_Q$ fulfilling (*).
 Define 
 \[
-\overline H = \bigcup\{ A_P \setminus A_Q~|~A_P\in\minaccs(G(P)/s) \}.
+\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
 \]
-Since $A_P \setminus A_Q \neq\varnothing$ for all $A_P$ because of (*), $\overline H$ is a hitting set
+Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*), $\overline H$ is a hitting set
 of $\minaccs(G(P)/s)$ which has an  empty intersection with $A_Q$. Minimising 
 $\overline H$ induces the existence
-of a minimal hitting set $H\in \text{MINHIT}(P/s)$ which is {\it not} a hitting set of 
-of $\minaccs(G(Q)/s)$. This completes the proof of the lemma.
+of a minimal hitting set $H\in \text{minHit}(P/s)$ which is {\it not} a hitting set  
+of $\minaccs(G(Q)/s)$, a contradiction to Assumption~2. This completes the proof of the lemma.
 \xbox

### 2018-08-07 paper/finitecomplete-trace.tex
initial files
Theorem on trace refinement testing added

--- 2018-08-01 paper/finitecomplete-failures.tex
@@
 % ==========================================================================
-\section{Finite Complete Test Suites for CSP Failures Refinment}
-\label{sec:finitecompletefails}
+\section{Finite Complete Test Suites for CSP Trace Refinement}
+\label{sec:finitecomplete}
 % ==========================================================================
 
+For establishing trace refinement, the following class of adaptive test cases will be used; again, they are defined for integers $p \ge 0$.
+
+\begin{eqnarray}
+U_T(p) & = & U_T(p,\varepsilon)
+\\
+U_T(p,s) & = & \big( e:(\Sigma - [P/s]^0) \then \efail\then \Stop \big)
+\label{eq:ufa}
+\\ & & \extchoice \nonumber
+\\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
+\label{eq:ufb}
+\\ & & \extchoice \nonumber
+\\ & & (\#s < p) \& \big( e:[P/s]^0 \then U_T(p,s.e) \big)
+\label{eq:ufc}
+\\ & & \extchoice \nonumber
+\\ & & (\#s = p) \& \big( \epass\then \Stop  \big)
+\label{eq:ufd}
+\end{eqnarray}
+
+The difference between adaptive tests $U_T(p)$ for trace refinement and
+$U_F(p)$ for failures refinement consists in the fact that the former do not ``probe''
+the SUT with respect to minimal sets of events to be accepted without blocking.
+
 % ==========================================================================
+The existence of complete, finite test suites is expressed in analogy to Theorem~\ref{th:failurestest}. A noteworthy difference consists in the fact that
+the complete suite for trace refinement just needs the single adaptive test case 
+$U_T(pq-1)$, while the complete failures test suite required the execution of 
+$\{ U_F(0),\dots,U_F(pq-1)\}$. The reason for this is that $U_T(pq-1)$ identifies trace
+errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for erroneous
+acceptances at the end of each trace of length $(pq -1)$.
+
+\begin{theorem}\label{th:tracetest}
+Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
+whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
+the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
+has at most $q$ states with $q \ge p$. 
+Then the test suite 
+\[
+\TS_T = \{ U_T(pq-1)   \}
+\]
+is complete with respect to ${\cal F} = (P,\lessdet_T,{\cal D})$.
+\xbox
+\end{theorem}
+
+We skip the proof of Theorem~\ref{th:tracetest}, since it is just a simplified version of 
+the proof of Theorem~\ref{th:failurestest}.

### 2018-08-07 paper/introduction.tex
Theorem on trace refinement testing added

@@
 normalised transition graphs do not exceed the size of the reference model's
-graph by more than a give constant.
+graph by more than a given constant.
 
-Our main contributions in this paper consist in the proof that 
+The main contribution of this paper consists in the proof that 
 for fault domains of the described type, finite, complete
-    test suite generation methods can be given for verifying (1)~trace
-    equivalence, (2)~trace refinement, (3)~failures equivalence, and
-    (4)~failures refinement.
-
+    test suite generation methods can be given for testing against the conformance relations 
+    trace refinement and failures refinement. 
 The existence of -- possibly infinite -- complete test suites has been established for process algebras in general and  for CSP in particular by several authors, 
 e.g.~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
-To our best knowledge, however, this article is the first to present {\it finite}, complete test suites associated with this class of fault domains.
+To our best knowledge, however, this article is the first to present {\it finite}, complete test suites associated with this class of fault domains and conformance relations.
 

### 2018-08-07 paper/preliminaries.tex
Theorem on trace refinement testing added

@@
 might refuse to engage into, when in a process state corresponding to $n$.
-The number of refusal sets in $\refs(P/s)$ specifies the degree of
-nondeterminism that is present in process state $P/s$: the more refusal sets
-contained in  $\refs(P/s)$, the more nondeterministic is the behaviour in
-state $P/s$. If $P/s$ is deterministic, its refusals coincide with the set of
+ If $P/s$ is deterministic, its refusals coincide with the set of
 subsets of $\Sigma - [P/s]^0$, including the empty set.
@@
 \end{equation}
+The cardinality of $\maxrefs(P/s)$ reflects the degree of
+nondeterminism that is present in process state $P/s$: the more maximal refusal sets
+contained in  $\maxrefs(P/s)$, the more nondeterministic is the behaviour in
+state $P/s$. Deterministic process states $P/s$ have exactly the one maximal refusal
+$\Sigma-[P/s]^0$.
+
 To see that this approach works only for finite CSP processes, consider the
 example where $\Sigma$ is infinite. In this case,
-$\maxrefs(STOP/\langle\rangle)$ is empty, and so we cannot use this set to
-calculate the refusals of $STOP$, that is, $\refs(STOP/\langle\rangle)$ as
+$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
+calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
 defined above. As with refusals, we also use the transition graph-oriented
@@
 Well-formed normalised transition graphs must not refuse an event of the
-fan-out of a state in {\it every} refusal applicable in this state; more
+initials of a state in {\it every} refusal applicable in this state; more
 formally,
@@
 
-Every node of a transition graph can be labelled with with its minimal acceptances, and this information is equivalent to 
+Every node of a transition graph can alternatively 
+be labelled  with its minimal acceptances, and this information is equivalent to 
 that contained in its maximal refusals.
@@
 
-Note that the node names are generated by the FDR tool (see next paragraph).
-The node numbering is generated by FDR during the normalisation procedure.
+Note that the node names including their number suffixes 
+are generated by the FDR tool (see next paragraph).
+The numbering is generated   during the normalisation procedure.
 Therefore, the node numbers do not reflect the distance from the initial node
 Node\_0.
-%Node\_2 is a direct successor to the initial node Node\_0, while Node\_1 is a direct successor of Node\_2.
 \end{example}

### 2018-08-08 paper/casestudies.tex
Theorem on trace refinement testing added
small improvements

@@
 \label{sec:case}
-
-% =======================================================================
-\subsection{Testing for Trace Equivalence}
 
@@
 % =======================================================================
-\subsection{Testing for Failures Equivalence}
-
-% =======================================================================
 \subsection{Testing for Failures Refinement}
 
+For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, 
+it is advisable to avoid an enumeration
+of traces $s$ the reference process has run through. Instead, we calculate
+the following auxiliary functions from $P$'s transition graph.
+\begin{eqnarray*}
+\text{initials} & : & N \fun \power (\Sigma) 
+\\
+\text{minHit} & : & N \fun \power\power(\Sigma)
+\end{eqnarray*}
+In a state $n = G(P)/s$, the set $\text{initials(n)}$ equals the events labelling
+outgoing edges of $n$, so  $\text{initials(n)} = [P/s]^0$. Function $\text{minHit}$
+maps $n$ to the set of all minimal hitting sets associated with the minimal acceptances
+of $n$, so $\text{minHit}(n) = \text{minHit}(P/s)$. Then, using the transition function
+of $t$ in addition to the two auxiliary functions, $U_F(p)$ can be re-written
+as the failures-equivalent CSP process
+\begin{eqnarray}
+U_F^1(p) & = & U_F^1(p,0,\ii n)
+\\
+U_F^1(p,k,n) & = & \big( e:(\Sigma - \text{initials}(n)) \then \efail\then \Stop \big)
+\label{eq:xufa}
+\\ & & \extchoice \nonumber
+\\ & & (\text{initials}(n) = \varnothing)    \&   \big( \epass \then \Stop \big)
+\label{eq:xufb}
+\\ & & \extchoice \nonumber
+\\ & & (k < p) \& \big( e:\text{initials}(n) \then U_F^1(p,(k+1),t(n,e)) \big)
+\label{eq:xufc}
+\\ & & \extchoice \nonumber
+\\ & & (k = p) \& \big( \sqcap_{H\in\text{minHit}(n)} ( e:H \then \epass \then\Stop   )  \big)
+\label{eq:xufd}
+\end{eqnarray}
+
+\begin{example}
+\label{ex:uf1tests}
+
+\end{example}
+
 % ======================================================================

### 2018-08-08 paper/preliminaries.tex
small improvements

@@
 Its transition graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state
-$P/\langle\rangle$ is represented there as Node\_0, with $\{ b,c\}$ as the
-only maximal refusal, since $a$ can never be refused, and no other events are
+$P/\langle\rangle$ is represented there as Node\_0, with $\{ a\}$ as the
+only minimal acceptance, since $a$ can never be refused, and no other events are
 accepted. Having engaged into $a$, the transition emanating from Node\_0
 leads to Node\_2 representing  the process state $P/a = Q\intchoice R$. The
-internal choice operator induces several refusal sets derived from $Q$ and
+internal choice operator induces several minimal acceptances derived from $Q$ and
 $R$. Since these processes accept their initial events in external choice,
-$Q\intchoice R$ induces two maximal refusal sets $\{b\}$ and
-$\{a\}$. Note that event $c$ can never be refused, since it is not a member
-of any maximal refusal.
+$Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$ and
+$\{b,c\}$. Note that event $c$ can never be refused, since it is contained in each minimal acceptance set.
 
@@
 process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
-two maximal refusals labelling Node\_1.
-
+two minimal acceptance sets labelling Node\_1.
 Similar considerations explain the other nodes and transitions in
@@
 Node\_0.
+\xbox
 \end{example}
@@
 \end{eqnarray}
+\xbox
 \end{lemma}
@@
 normalised transition graphs. We need this only for the investigation of corresponding traces in reference processes and processes under test. Therefore, the 
-labelling of nodes with maximal refusals or minimal acceptances   will be disregard in the 
+labelling of nodes with maximal refusals or minimal acceptances   will be disregarded in the 
 product construction.

### 2018-08-08 paper/finitecomplete-failures.tex
Theorem on trace refinement testing added
Fixes after suggestions by Wen-ling

@@
 If $P/s$ is the STOP-process, this is revealed by its initials being empty. In this case, the test may terminate 
-successfully (second operand of the external choice in $UF(p,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
+successfully (second operand of the external choice in $U_F(p,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
 Therefore,
@@
 If the length of $s$ is still less than $p$, the test accepts any event from the initials
-$[P/s]^0$ and continues recursively as sub-process $U(p,s.e)$ (third operand of the external choice).
+$[P/s]^0$ and continues recursively as sub-process $U_F(p,s.e)$ (third operand of the external choice).
 A test of this type is called \emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts its consecutive behaviour to the event selected by the SUT.
@@
 
-\begin{theorem}
+\begin{theorem}\label{th:failurestest}
 Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
@@
 \end{description}
-It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U(k))$ with $k < pq$ ends with the FAIL event $\efail$.
+It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the FAIL event $\efail$ or without giving any verdict.
+The latter case is also interpreted   as FAIL, since then
+$(\epass\then\Stop)$ is no longer refined by the test execution in the failures model.
 
@@
 Now the construction of the transition function of $G$ implies that $u$ is also a trace of $P$ and $Q$.
-Since test $U_F(pq-1)$ accepts all traces of $P$ up to length $pq$, $u$ is also a trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$.
+Since test $U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$.
 Since $s.e\not\in\trc(P)$,

### 2018-08-11 paper/casestudies.tex
Example for failures-testing added

@@
 
+% .....................................................................................
+ \begin{figure}
+ %%\hspace*{-40mm}
+ \begin{center}
+\includegraphics[width=\textwidth]{z_acc.pdf}
+\end{center}
+%%\vspace*{-10mm}
+\caption{Normalised transition graph of faulty implementation $Z$  from Example~\ref{ex:uf1tests}.}
+ \label{fig:tgZ}
+ \end{figure}
+% .......................................................................................
+
 \begin{example}
 \label{ex:uf1tests}
+Consider the following erroneous implementation $Z$ of process $P$ from Example~\ref{ex:a}.
+\begin{eqnarray*}
+Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
+\\
+Q_1 & = & a\then Z \extchoice c\then Z
+\\
+R_1(r_{max},k) & = & (k < r_{max}) \& \big( c\then R_1(r_{max},k+1)\extchoice b\then Z  \big)
+\\ & & \extchoice
+\\ & & (k = r_{max}) \& \big( c\then R_1(r_{max},r_{max})\intchoice b\then Z  \big)
+\end{eqnarray*}
+It is easy to see (and can be checked with FDR4) that $Z$ is trace-equivalent to $P$. While $k < r_{max}$, process $Z$ also accepts the same sets of events as $P$. If, however,
+sub-process $R_1(r_{max},k)$ runs through several recursions until condition 
+$k = r_{max}$ is fulfilled, the process uses internal choice instead of external choice, and so this process state   no longer refines   the corresponding process state $R$ of the reference process $P$ in the failures model.
+For $r_{max} = 3$,   the
+normalised transition graph of $Z$ is displayed in Fig.~\ref{fig:tgZ}.
 
+Running the test $U_F^1(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p = 4$ states
+and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test depth to be used 
+according to Theorem~\ref{th:failurestest}), tests $U_F^1(0),\dots, U_F^1(3)$ are passed by $Z$, but $Z$ fails $U_F^1(4)$, because after  execution trace
+\[
+s = a.c.c.c, \qquad\text{(note that $G(Z)/s = \text{Node\_3}$ according to Fig.~\ref{fig:tgZ})},
+\]
+it may be the case that   $Z$  
+only accepts $\{b\}$ due to internal choice and $U_F^1(4)$ 
+-- also due to internal choice -- 
+only accepts
+the minimal hitting set $\{ c \}$ in union with the event $a\in (\Sigma - [P/s]^0)$. 
+As
+a consequence, $(Z\parallel[\Sigma] U_F^1(4))/s$ deadlocks for this execution, 
+and the pass-event cannot be
+produced. Another failing execution would be the situation where $Z/s$ choses to
+accept only $\{c \}$, while $U_F^1(4))/s$ choses to accept only $\{a,b\}$.
+
+Therefore, 
+\[
+(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F^1(4)),
+\]
+and the test fails.
+\xbox
 \end{example}

### 2018-08-12 paper/introduction.tex
small changes to introduction

--- 2018-08-02 paper/introduction.tex
@@
 evaluated and integrated into industrial verification processes by many
-companies. This holds particularly for the embedded and cyber-physical
-systems domain. While MBT is applied in different flavours, we consider the
-most effective variant to be the one where test cases and concrete test data,
-as well as checkers for the expected results~(\emph{test oracles}), are
-automatically generated from a reference model:~it guarantees the maximal
-return of investment for the time and effort invested into creating the test
-model. The test suites generated in this way, however, usually have different
-test strength, depending on the generation algorithms applied.
+companies worldwide. This holds particularly for the embedded and
+cyber-physical systems domains, where critical systems require rigorous
+testing.
+
+While MBT is applied in different flavours, we consider the most effective
+variant to be the one where test cases and concrete test data, as well as
+checkers for the expected results~(\emph{test oracles}), are automatically
+generated from a reference model. This guarantees maximal return of the
+investment of time and effort to create the test model. The test suites
+generated in this way, however, usually have different test strength,
+depending on the generation algorithms applied.
 
@@
 modelling formalisms. In this paper, we use \emph{Communicating Sequential
-Processes (CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}; this is a mature
+Processes (CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
 process-algebraic approach that has been shown to be well-suited for the
@@
 
-This paper presents complete black-box test suites for divergence-free\footnote{The assumption of divergence freedom is usually applied in black-box testing, since it cannot distinguish between divergence and deadlock.} CSP processes interpreted
-both in the trace   and the failures semantics.  
-Our results complement work published by two of the authors
+This paper presents complete black-box test suites for CSP processes that are
+divergence-free\footnote{The assumption of divergence freedom is usually
+applied in black-box testing, since it cannot distinguish between divergence
+and deadlock.} and interpreted both in the trace and the failures semantics.
+Our results complement work by two of the authors
 in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
 collections of processes refining a  ``most general'' fault domain member.
-With this concept of fault domains, complete test suites may be finite or
-infinite. While this gives important insight into the theory of complete test
-suites, we are particularly interested in finite suites when it comes to
-their practical application.
+With that concept, complete test suites may be finite or infinite. Yhis gives
+important insight into the theory of fault-domain testing for CSP, but we are
+particularly interested in finite suites when it comes to practical
+application. While~\cite{DBLP:conf/pts/CavalcantiS17} requires additional
+criteria to select tests from infinite test suites, here, we further restrict
+fault domains using a graph representation of processes used in model
+checking to obtain test suites that are finite.
 
-Therefore,  a complementary approach to the definition of CSP fault
-domains is presented in this paper. To this end, we observe that every finite-state CSP
-process can be semantically represented as a finite normalised transition
-graph, whose edges are labelled by the events the process engages in, and
-whose nodes are labelled by minimal acceptances or, alternatively, maximal
+Our complementary approach to the definition of CSP fault domains is
+presented in this paper. We observe that every finite-state CSP process can
+be semantically represented as a finite normalised transition graph, whose
+edges are labelled by the events the process engages in, and whose nodes are
+labelled by minimal acceptances or, alternatively, maximal
 refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
-degree of nondeterminism that is present in a given CSP process state that is
-in one-one-correspondence to a node of the normalised transition graph.
-Inspired by the way that fault-domains are specified for finite state
-machines, we define them here as the set of CSP processes whose
-normalised transition graphs do not exceed the size of the reference model's
-graph by more than a give constant.
+degree of nondeterminism present in a given process state that is in
+one-one-correspondence to a node of the normalised transition graph. Inspired
+by the way that fault-domains are specified for finite state machines, we
+define them as the set of CSP processes whose normalised transition graphs do
+not exceed the size of the reference model's graph by more than a given
+constant.
 
-Our main contributions in this paper consist in the proof that 
-for fault domains of the described type, finite, complete
-    test suite generation methods can be given for verifying (1)~trace
-    equivalence, (2)~trace refinement, (3)~failures equivalence, and
-    (4)~failures refinement.
+The main contribution of this paper is the proof that for fault domains of
+the described type, finite, complete test suite generation methods can be
+given for testing against trace and failures refinement and equivalence. The
+existence of -- possibly infinite -- complete test suites has been
+established for process algebras, and for CSP in particular, by several
+authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
+To the best of our knowledge, however, this article is the first to present
+{\it finite}, complete test suites associated with this class of fault
+domains and conformance relations.
 
-The existence of -- possibly infinite -- complete test suites has been established for process algebras in general and  for CSP in particular by several authors, 
-e.g.~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
-To our best knowledge, however, this article is the first to present {\it finite}, complete test suites associated with this class of fault domains.
-
-It should be noted that the presented results are of a theoretical nature: despite being finite, the resulting test suite size will still be too large to be applied in practice to real-world problems.
-We discuss a number of promising options how the test suite size can be reduced to become practically applicable.
+It should be noted that the presented results are of a theoretical
+nature:~despite being finite, the resulting test suite size will still be too
+large to be applied in practice to real-world problems. We discuss a number
+of promising options how the test suite size can be reduced to become
+practically applicable.
 

### 2018-08-15 paper/preliminaries.tex
small changes to Section 2, but a couple of notes that need attention

@@
 
+\begin{itemize}
+  \item traces and failures model
+  \item refinement and equivalence
+\end{itemize}
+
+The FDR tool~\cite{fdr} supports model checking and semantic analyses of CSP
+processes.
+
 % =========================================================================
@@
 partial \emph{transition function} $t$ maps a node $n$ and an event
-$e\in\Sigma$ to its successor node $t(n,e)$, if, and only if, $(n,e)$ is in
-the domain of $t$, that is, there is a transition from $n$ with label $e$.
-Normalisation of $G(P)$ is reflected by the fact that $t$ is a function.
+$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
+$t$, then there is a transition from $n$ with label $e$. Normalisation of
+$G(P)$ is reflected by the fact that $t$ is a function.
 
@@
 the events from $s$ one by one, starting in $\ii n$. Therefore, $G(P)/s$  is
-also well defined.
-
-By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
-as labels in any outgoing transitions.
+also well defined. By $[n]^0$ we denote the \emph{initials} of $n$:~the set
+of events occurring as labels in any outgoing transitions.
 $$
@@
 $$
-We also use this notation for CSP processes:~$[P]^0$ is the set
-of events $P$ may engage into, in other words, the initials of $P$ after the
-empty trace of events, that is, $initials(P/\langle\rangle)$ as defined
-in~\cite{Roscoe2010}.
+We also use this notation for CSP processes:~$[P]^0$ is the set of events $P$
+may engage into, in other words, the initials of $P$ after the empty trace of
+events, that is, $initials(P/\langle\rangle)$ as defined
+in~\cite{Roscoe2010}, for example.
 
@@
 subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}, $\refs(P/s)$ can be
-re-constructed by knowing the set of \emph{maximal refusals}
-$\maxrefs(P/s)\subseteq\refs(P/s)$. More formally, the maximal refusals
-$\maxrefs(P/s)$ are defined as
+constructed from the set of \emph{maximal refusals}
+$\maxrefs(P/s)\subseteq\refs(P/s)$. More formally, $\maxrefs(P/s)$ is defined
+as follows.
+%
 \begin{equation}
@@
 \end{equation}
-Conversely, with the maximal refusals at hand, we can reconstruct the refusals by subset-closure:
+%
+Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
+reconstruct the refusals $\refs(P/s)$ by subset-closure as follows.
+%
 \begin{equation}
@@
 \end{equation}
+%
 The cardinality of $\maxrefs(P/s)$ reflects the degree of
@@
 
-To see that this approach works only for finite CSP processes, consider the
-example where $\Sigma$ is infinite. In this case,
+To see that this approach works only for finite CSP processes, we consider
+the example where $\Sigma$ is infinite. In this case,
 $\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
@@
 formally,
+%
 \begin{equation}
@@
 \end{equation}
+%
+By construction, normalised transition graphs reflect the \emph{failures
+semantics} of finite-state CSP processes:~the traces $s$ of a process are
+defined by the sequences of transitions associated with paths through its
+graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
+r(G(P)/s)$ represent the failures of $P$.
 
-By construction, normalised transition graphs reflect the \emph{failures
-semantics} of finite-state CSP processes:~the traces $s$ of a process are the
-sequences of transition associated with paths through the graph,
-starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
-r(G(P)/s)$ represent the failures $\failure(P)$ of $P$.
-
-When investigating  tests for failures refinement, 
-the notion of \emph{acceptances}~\cite{Hennessy:1988:ATP:50497} 
-which is dual to refusals will also be useful. An acceptance set of $P/s$ is a subset
-of the initials $[P/s]^0$, i.e. a subset of events labelling  outgoing transitions
-of $G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance equals
-$[P/s]^0$, because $P/s$ will never refuse any of the events contained in this set. If
-$P/s$ is nondeterministic, it will internally choose one of it 
-\emph{minimal acceptance sets} $A$ and never refuse any event in $A$, while refusing the
-events in $\Sigma-A$. The acceptances of 
- $P/s$ are denoted by $\accs(P/s)$, the minimal acceptances by $\minaccs(P/s)$. 
- They fulfil the following rules.
+When investigating  tests for failures refinement, the notion of
+\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
+is also useful. An acceptance set of $P/s$ is a subset of the initials
+$[P/s]^0$, i.e., a subset of events labelling  outgoing transitions of
+$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
+equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
+this set. If $P/s$ is nondeterministic, it internally chooses one of its
+\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
+refusing the events in $\Sigma-A$. The acceptances of $P/s$ are denoted by
+$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
+the following properties.
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-Note that (\ref{eq:accref}) can be regarded as a definition of minimal acceptances by means
-of maximal refusals. Then (\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences
-of this definition and  the fact that refusals are subset-closed.
+%
+Note that (\ref{eq:accref}) can be regarded as a definition of minimal
+acceptances by means of maximal refusals. In this case,
+(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
+definition and the fact that refusals are subset-closed. \fixme{alcc: are you
+sure? (4) does not talk about Acc. Is this important?}
 
-Every node of a transition graph can alternatively 
-be labelled  with its minimal acceptances, and this information is equivalent to 
-that contained in its maximal refusals.
+Every node of a normalised transition graph can alternatively be labelled
+with their minimal acceptances, and this information is equivalent to that
+contained in the maximal refusals. This is the representation used by FDR.
 
 % .....................................................................................
- \begin{figure}
+ \begin{figure}[t]
    %%\hspace*{-40mm}
@@
 \begin{example}\label{ex:a}
-Consider the CSP process $P$ defined below, and that the set of events
-$\Sigma$ is $\{a,b,c\}$.
+Consider the CSP process $P$ defined below, and $\Sigma = \{a,b,c\}$.
 \begin{eqnarray*}
@@
 
-Note that the node names including their number suffixes 
-are generated by the FDR tool (see next paragraph).
-The numbering is generated   during the normalisation procedure.
-Therefore, the node numbers do not reflect the distance from the initial node
-Node\_0.
+Note that the node names including their number suffixes are generated by the
+FDR tool. The numbering is generated during the normalisation procedure. So,
+the node numbers do not reflect the distance from the initial node Node\_0.
 \xbox
@@
 
-Refinement relations between finite-state CSP processes $P, Q$ 
-can be be expressed by means of their normalised transition graphs in the following way.
-
+Refinement relations between finite-state CSP processes $P, Q$ can be be
+expressed by means of their normalised transition graphs in the following
+way. \fixme{alcc: trc was defined for P, not G(P) and this lemma is used
+later for trc(P). So, we need to be clearer as to what we mean.}
+%
 \begin{lemma}
@@
 \end{lemma}
-
-For proving our main theorems, it is necessary to consider the \emph{product} of 
-normalised transition graphs. We need this only for the investigation of corresponding traces in reference processes and processes under test. Therefore, the 
-labelling of nodes with maximal refusals or minimal acceptances   will be disregarded in the 
-product construction.
-Consider two normalised transition graphs 
+%
+For proving our main theorems, it is necessary to consider the \emph{product}
+of normalised transition graphs. We need this only for the investigation of
+corresponding traces in reference processes and processes for SUTs.
+Therefore, the labelling of nodes with maximal refusals or minimal
+acceptances are disregarded in the product construction. Consider two
+normalised transition graphs
 \[
@@
 \]
-over the same alphabet $\Sigma$. Then their product is defined by
+over the same alphabet $\Sigma$. Their product is defined by
 \begin{eqnarray}
@@
 \subsubsection*{Tool Considerations}
-The FDR tool~\cite{fdr} supports model checking and semantic analyses of CSP
-processes. It provides an API~\cite{fdrmanual} that can be used to construct
-normalised transition graphs for CSP processes.
-
-The FDR graph nodes are labelled by \emph{minimal acceptances} instead of
-maximal refusals as described above. Since such a minimal acceptance set is
-the complement of a maximal refusal, the function $r$ mapping states to their
-refusals can be implemented by creating the complements of all minimal
-acceptances and then building all subsets of these complements. For practical
-applications, the subset closure is never constructed in an explicit way;
-instead, sets are checked with respect to containment in a maximal refusal.
+FDR provides an API~\cite{fdrmanual} that can be used to construct normalised
+transition graphs for CSP processes. The FDR graph nodes are labelled by
+\emph{minimal acceptances} instead of maximal refusals as described above.
+Since such a minimal acceptance set is the complement of a maximal refusal,
+the function $r$ mapping states to their refusals can be implemented by
+taking the complements of all minimal acceptances and then building all their
+subsets. For practical applications, the subset closure is never constructed
+in an explicit way; instead, sets are checked with respect to containment in
+a maximal refusal.
 
@@
 
-%
 %The maximal refusals in each process state $P/s$
@@
 
-When testing for failures refinement, it will become apparent that the main idea of the underlying test strategy can be based on solving a \emph{hitting set problem}.
-Given a collection of finite sets $C = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a
-subset of a universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set satisfying
+The main idea of the underlying test strategy for failures refinement can be
+based on solving a \emph{hitting set problem}. Given a finite collection of
+finite sets $C = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a subset of a
+universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set
+satisfying the following property.
+%
 \begin{equation}
@@
 \end{equation}
+%
+A \emph{minimal hitting set} is a hitting set that cannot be further reduced
+without losing the characteristic property (\ref{eq:hit}). The problem of
+determining minimal hitting sets is known to be
+NP-hard~\cite{Book1975-BOOKRM}, but we will see below that it reduces the
+effort of testing for failures refinement from a factor of $2^\Sigma$ to a
+factor that equals the number of minimal hitting sets.
 
-A \emph{minimal hitting set} is a hitting set which cannot be further reduced without losing 
-the characteristic property (\ref{eq:hit}). The problem of determining minimal hitting sets is known to be NP-hard~\cite{Book1975-BOOKRM}, but we will see below that it reduces 
-the effort of testing for failures refinement from a factor of $2^\Sigma$ to a factor that equals the number of minimal hitting sets.
-
-For testing in the failures model, the following lemmas about hitting sets are required.
+For testing, the following lemma about hitting sets are required.
 
@@
 \begin{enumerate}
-\item $P\lessdet_F Q$
+\item $P\lessdet_F Q$ \fixme{Note that if traces refinement is already
+    given, this is conf.}
 \item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{minHit}(P/s)$, $H$ is
@@
 \begin{proof}
-For showing ``$1 \Rightarrow 2$'', assume that $P\lessdet_F Q$ and suppose that
-$s\in\trc(P)\cap \trc(Q)$. Then Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), 
-states that 
+For showing ``$1 \Rightarrow 2$'', assume that $P\lessdet_F Q$ and suppose
+that $s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref},
+(\ref{eq:failrefb}), states that
 \[
@@
 \]
-Therefore,
-$H \in  \text{minHit}(P/s)$ not only implies $H\cap A_P\neq\varnothing$ for all minimal
-acceptances $A_P$, but also
-$H\cap A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq A_Q$ for at least one $A_P$.
-As a consequence, each $H \in  \text{minHit}(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
+Therefore, $H \in  \text{minHit}(P/s)$ not only implies $H\cap
+A_P\neq\varnothing$ for all minimal acceptances $A_P$, but also $H\cap
+A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq
+A_Q$ for at least one $A_P$. As a consequence, each $H \in
+\text{minHit}(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
 
-To prove ``$2 \Rightarrow 1$'',
-assume $P\lessdet_T Q$, but $P\not\lessdet_F Q$. According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that 
+To prove ``$2 \Rightarrow 1$'', assume $P\lessdet_T Q$, but $P\not\lessdet_F
+Q$. According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
+exists $s\in\trc(P)\cap \trc(Q)$ such that
 \[
@@
 \]
-Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*), $\overline H$ is a hitting set
-of $\minaccs(G(P)/s)$ which has an  empty intersection with $A_Q$. Minimising 
-$\overline H$ induces the existence
-of a minimal hitting set $H\in \text{minHit}(P/s)$ which is {\it not} a hitting set  
-of $\minaccs(G(Q)/s)$, a contradiction to Assumption~2. This completes the proof of the lemma.
-\xbox
+Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
+$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
+intersection with $A_Q$. \fixme{alcc: minaccs is used for P above.}
+Minimising $\overline H$ induces the existence of a minimal hitting set $H\in
+\text{minHit}(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
+contradiction to Assumption~2. This completes the proof of the lemma. \xbox
 \end{proof}
+%
+This result is used in the sequel in Section~\ref{}.
 

### 2018-08-16 paper/finitecomplete-failures.tex
small changes to Section 3, but a couple of notes that need attention

@@
 \begin{lemma}
-\label{lemma:extendV}
-Let $P$ be a CSP process with normalised transition graph $G(P)$, 
-such that all states in $N$ are reachable.
-Let $V\subseteq\Sigma^*$ be a finite prefix-closed set of
-sequences of events. Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is,
-$|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, 
-extended by any event of $\Sigma$.
-Then $G(P)$ reaches at least $(k+1)$ nodes under $V\cup V.\Sigma$.
+\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
+graph $G(P)$, such that all states in $N$ are reachable. \fixme{alcc: can a
+normalised graph for a process have states that are unreachable?} Let
+$V\subseteq\Sigma^*$ be a finite prefix-closed set of sequences of events.
+Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is, $|t(\ii
+n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended
+by any event of $\Sigma$. Then $G(P)$ reaches at least $(k+1)$ nodes under
+$V\cup V.\Sigma$.
 \end{lemma}
 \begin{proof}
-Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable, there exists
-a trace $s$ such that $G(P)/s = n'$.
-Decompose $s = s_1.e.s_2$ with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and $G(P)/s_1.e \not\in t(\ii n,V)$.
-Such a decomposition always exists, because $V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
+Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
+there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
+with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
+$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
+$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
 Note, however, that it is not necessarily the case that $s_1\in V$.
 
-Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such that
-$\ol n = G(P)/u = G(P)/s_1$. Since $s = s_1.e.s_2$ is a trace of $P$ and 
-$\ol n = G(P)/s_1$,
-$(\ol n,e)$ is in the domain of $t$. Therefore,  
-$n = G(P)/u.e = G(P)/s_1.e \not\in t(\ii n,V)$ is a well-defined node
-of $N$ which is not contained in $t(\ii n,V)$. Since $u.e\in V\cup V.\Sigma$,
-$G(P)$ reaches at least the additional node $n$ under $V\cup V.\Sigma$. This completes the proof.
-\xbox
+Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such
+that $G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
+$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
+= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
+n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
+node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
 \end{proof}
+\fixme{alcc: Explain the structure of the section? Can the above result be in
+a section, rather than the introduction of the section?}
 
@@
 
-For a given reference process $P$ and for each integer $p\ge 0$, 
-we define a CSP test process as follows. 
+For a given reference process $P$ and for each integer $p\ge 0$, we define a
+CSP test process for failures refinement as shown below.
+%
 \begin{eqnarray}
@@
 \\
-U_F(p,s) & = & \big( e:(\Sigma - [P/s]^0) \then \efail\then \Stop \big)
+U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
 \label{eq:ufa}
@@
 \\ & & \extchoice \nonumber
-\\ & & (\#s < p) \& \big( e:[P/s]^0 \then U_F(p,s.e) \big)
+\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
 \label{eq:ufc}
 \\ & & \extchoice \nonumber
-\\ & & (\#s = p) \& \big( \sqcap_{H\in\text{minHit}(P/s)} ( e:H \then \epass \then\Stop   )  \big)
+\\ & & (\#s = p) \& \big( \sqcap_{H\in\text{minHit}(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
 \label{eq:ufd}
 \end{eqnarray}
-
+%
 A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$
-that operates on the same alphabet as $P$, synchronising over
-alphabet $\Sigma$. Therefore, a test execution is any trace of the concurrent process
+\fixme{alcc: add an example?} that operates on the same alphabet as $P$,
+synchronising over alphabet $\Sigma$. Therefore, a test execution is any
+trace of the concurrent process
 \[
@@
 \]
-It is assumed that the events $\efail$ and $\epass$ denoting FAIL and PASS of the test execution, respectively, are events outside $\Sigma$.
-Since we assume that $Q$ is free of livelocks, it is guaranteed that each test execution terminates after some $s\in\trc(P)$ with length $(p+1)$ at the latest. 
-The test is \emph{passed} by the SUT (we write $Q\ \pass\ U_F(p)$) if and only if {\it every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with PASS event $\epass$.
-This can also be  expressed by means of a failures refinement relation:
+It is assumed that the events $\efail$ and $\epass$, denoting FAIL and PASS
+of the test execution, are events outside $\Sigma$. Since we assume that $Q$
+is free of livelocks, it is guaranteed that each test execution terminates
+after some $s\in\trc(P)$ with length $(p+1)$ at the latest. The test is
+\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
+every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with PASS event
+$\epass$. This can also be  expressed by means of a failures refinement.
 \[
@@
 \]
-This type of pass relation is often called \emph{must test}, because every test execution must end with the $\epass$-event \cite{Hennessy:1988:ATP:50497}. Note that it
-is necessary to use the failures refinement relation in this condition, and not the trace refinement relation: process $(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same visible traces $\varepsilon, \epass$ as the ``Test Passed Process'' $(\epass\then\Stop)$.
-However,
-the former may nondeterministically refuse $\epass$, due to a deadlock occurring when
-a faulty SUT process executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
+This type of pass relation is often called \emph{must test}, because every
+test execution must end with the $\epass$
+event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
+failures-refinement relation in this condition, and not the trace-refinement
+relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
+visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
+Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
+refuse $\epass$, due to a deadlock occurring when a faulty SUT process
+executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
 because $\#s = p$. This is explained further in the next paragraphs.
 
-Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, 
-up to a length of $p$. 
-If, after having already run through $s\in\trc(P)$ with $\#s < p$,
-an event is accepted by the SUT which is outside the initials of $P/s$, the test 
-immediately terminates with FAIL-event $\efail$. This is handled by the first operand of the
-external choice in sub-process $U_F(p,s)$.
+Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
+a length $p$. If, after having already run through $s\in\trc(P)$ with $\#s <
+p$, an event is accepted by the SUT that is outside the initials of $P/s$,
+the test immediately terminates with FAIL-event $\efail$. This is handled by
+the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,s)$
+defined above.
 
-If $P/s$ is the STOP-process, this is revealed by its initials being empty. In this case, the test may terminate 
-successfully (second operand of the external choice in $U_F(p,s)$). Note that at the same time, any (now illegal) event of the alphabet is also accepted by the test.
-Therefore,
-if the SUT would still accept an event in a state where $P/s$ is supposed to have stopped, there exists a test execution which terminates with FAIL by choosing the first
-operand of the external choice.
+If $P/s$ is the $STOP$ process, this is revealed by its initials being empty.
+In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
+the external choice in $U_F(p,s)$). Note that at the same time, any (illegal)
+event of the alphabet is also accepted by the test in branch~(\ref{eq:ufa}).
+So, if the SUT accepts an event in a state where $P/s$ is supposed to have
+stopped, there exists a test execution that terminates with FAIL by choosing
+the first branch of the external choice.
 
-If the length of $s$ is still less than $p$, the test accepts any event from the initials
-$[P/s]^0$ and continues recursively as sub-process $U_F(p,s.e)$ (third operand of the external choice).
-A test of this type is called \emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts its consecutive behaviour to the event selected by the SUT.
+If the length of $s$ is still less than $p$, the test accepts any event from
+the initials $[P/s]^0$ and continues recursively as $U_F(p,s.e)$ in
+branch~(\ref{eq:ufc}). A test of this type is called \emph{adaptive}, because
+it accepts any legal behaviour of the SUT and adapts its consecutive
+behaviour to the event selected by the SUT.
 
@@
 \begin{proof}
-To prove soundness of $\TS_F$, suppose that $Q\in{\cal D} \wedge P\lessdet_F Q$. Then
-$\trc(Q)\subseteq \trc(P)$, and Lemma~\ref{lemma:hseta} implies that 
-for all traces $s$ of $Q$, every $H\in\text{minHit}(P/s)$ is a 
-    hitting set for $\minaccs(Q/s)$. As a consequence, when running in parallel
-with $Q$, any adaptive test $U_F(p)$ will always enter the branches 
-(\ref{eq:ufb}), (\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction
-for $U_F(p,s)$. Branch (\ref{eq:ufb}) always leads to a PASS verdict, branch (\ref{eq:ufc})
-always to test continuation without a verdict. For the last branch, we note that 
-any selected minimal hitting set $H\in\text{minHit}(P/s)$ has a non-empty intersection with
-each of the minimal acceptances of $Q/s$.
-As a consequence, $Q/s$ will never block when offered events from $H$, and the test terminates with PASS event $\epass$.
-Note that this argument requires that $Q$ is free of livelocks, because otherwise the PASS-events might not become visible, due to unbounded sequences of hidden events performed by $Q$.
-Note further, that this proof did not refer in any way to the size of $Q$'s normalised transition graph, so the test suite is sound for {\it all} non-divergent CSP process refining $P$.
+To prove soundness of $\TS_F$, suppose that $Q\in{\cal D} \wedge P\lessdet_F
+Q$. In this case, $\trc(Q)\subseteq \trc(P)$, and Lemma~\ref{lemma:hseta}
+implies that for all traces $s$ of $Q$, every $H$ in $\text{minHit}(P/s)$ is
+a hitting set for $\minaccs(Q/s)$. So, when running in parallel with $Q$, any
+adaptive test $U_F(p)$ will always enter the branches (\ref{eq:ufb}),
+(\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction for
+$U_F(p,s)$. Branch (\ref{eq:ufb}) leads always to a PASS verdict, and branch
+(\ref{eq:ufc}) to test continuation without a verdict. For the last branch,
+we note that any selected minimal hitting set $H\in\text{minHit}(P/s)$ has a
+non-empty intersection with each of the minimal acceptances of $Q/s$. As a
+consequence, $Q/s$ never blocks when offered events from $H$, and the test
+terminates with PASS event $\epass$. Note that this argument requires that
+$Q$ is free of livelocks, because otherwise the PASS-events might not become
+visible, due to unbounded sequences of hidden events performed by $Q$. Note
+further, that this proof did not refer in any way to the size of $Q$'s
+normalised transition graph, so the test suite is sound for {\it all}
+non-divergent CSP process refining $P$. \fixme{alcc: Soundness does not
+depend on the fault domain. I would have it as a separate theorem.}
 
-To prove exhaustiveness, consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
+To prove exhaustiveness, consider a process $Q\in{\cal D}$ with
+$P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
 \begin{description}
@@
 \end{description}
-It has to be shown for each of the two possibilities that at least one test execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the FAIL event $\efail$ or without giving any verdict.
-The latter case is also interpreted   as FAIL, since then
-$(\epass\then\Stop)$ is no longer refined by the test execution in the failures model.
+It has to be shown for each of the two possibilities that at least one test
+execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
+FAIL event $\efail$ or without giving any verdict. The latter case is also
+interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
+failures-refined by the test execution.
 
-For the first case, consider a  trace $s.e\in\trc(Q)$ such that $s\in\trc(P)$, but 
-$s.e\not\in\trc(P)$.
-Then $s$ is also a trace of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}.
-From the construction of $G$ described there, we know that $G$ has at most $pq$ reachable states, because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states.
-Suppose that $G/s = (n_P,n_Q)$.
-Applying Lemma~\ref{lemma:extendV} implies that this state can be reached by a trace $u\in\trc(G)$ of length $\#u < pq$.
-Now the construction of the transition function of $G$ implies that $u$ is also a trace of $P$ and $Q$.
-Since test $U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$.
-Since $s.e\not\in\trc(P)$,
-$e$ is an element of $\Sigma-[P/u]^0$.
-Therefore, in at least one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa}) with this event $e$, so that the test fails with event $\efail$.
-Again, the assumption of non-divergence of Q is needed for this conclusion.
+For the first case, consider a  trace $s.e\in\trc(Q)$ such that
+$s\in\trc(P)$, but $s.e\not\in\trc(P)$. Such a trace always exists because
+$\varepsilon$ is a trace of every process. In this case, $s$ is also a trace
+of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}.
+From the construction of $G$ described there, we know that $G$ has at most
+$pq$ reachable states, because $G(P)$ has $p$ states, and $G(Q)$ has at most
+$q$ states. Suppose that $G/s = (n_P,n_Q)$. Applying
+Lemma~\ref{lemma:extendV} implies that this state can be reached by a trace
+$u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
+function of $G$ implies that $u$ is also a trace of $P$ and $Q$. Since test
+$U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a
+trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$. Since
+$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0$. Therefore, in at
+least one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa})
+with this event $e$, so that the test fails with event $\efail$. Again, the
+assumption of non-divergence of Q is needed for this conclusion. \fixme{alcc:
+I didn't see how Lemma 3 is being applied here. Also, given the definition of
+the test as a process, I'm not sure why you need to make the argument using
+graph product.}
 
-For the Case~2, we note that trace $s$ is again a trace of the product graph $G$. Therefore, $G/s$ can again be reached by a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length
-$\#u < pq$. Consider test $U_F(\# u)$ which satisfies $U_F(\# u)/u = U_F(\#u,u)$, because
-it always performs branch (\ref{eq:ufc}) until the trace $u$ has been completely processed. $U_F(\#u,u)$
-  may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd}) only: 
-Assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has at least one non-empty minimal acceptance, so the guard condition $([P/s]^0 = \varnothing)$ of branch
- (\ref{eq:ufb}) evaluates to $\isf$ for $U_F(\#u,u)$. Moreover, 
- the guard condition $(\#s < p)$
-for branch (\ref{eq:ufc}) evaluates to $\isf$ for $U_F(\#u,u)$, too. 
-If branch (\ref{eq:ufa}) is executed, the test always fails. If branch (\ref{eq:ufd})
-is executed, the test fails for the execution where a minimal hitting set 
-$H\in\text{minHit}(P/u)$ is chosen by $U_F(\#u,u)$ which has an empty intersection
-with the minimal acceptance $A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an $H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there exists 
-a test execution   where $Q/u$ selects acceptance $A_Q$ and  $U_F(\#u,u)$ 
-selects $H$. This execution deadlocks in 
-process state $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot 
-produce the pass event $\epass$; this  means that the test fails. This concludes the proof.
-\xbox
+For the Case~2, we note that trace $s$ is again a trace of the product graph
+$G$. Therefore, $G/s$ can again be reached by a trace
+$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#
+u)$, which satisfies $U_F(\# u)/u = U_F(\#u,u)$, because it always performs
+branch (\ref{eq:ufc}) until the trace $u$ has been completely processed.
+$U_F(\#u,u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
+only:~assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has
+at least one non-empty minimal acceptance, so the guard condition $([P/s]^0 =
+\varnothing)$ of branch (\ref{eq:ufb}) evaluates to $\isf$ for $U_F(\#u,u)$.
+Moreover, the guard condition $(\#s < p)$ for branch (\ref{eq:ufc}) evaluates
+to $\isf$ for $U_F(\#u,u)$, too. If branch (\ref{eq:ufa}) is executed, the
+test always fails. If branch (\ref{eq:ufd}) is executed, the test fails for
+the execution where a minimal hitting set $H\in\text{minHit}(P/u)$ is chosen
+by $U_F(\#u,u)$ that has an empty intersection with the minimal acceptance
+$A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an
+$H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there
+exists a test execution   where $Q/u$ selects acceptance $A_Q$ and
+$U_F(\#u,u)$ selects $H$. This execution deadlocks in process state
+$(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the pass event
+$\epass$; this  means that the test fails. This concludes the proof. \xbox
 \end{proof}

### 2018-08-16 paper/finitecomplete-trace.tex
small improvements
small changes to Section 4

@@
 
-For establishing trace refinement, the following class of adaptive test cases will be used; again, they are defined for integers $p \ge 0$.
-
+For establishing trace refinement, the following class of adaptive test cases
+will be used; again, they are defined for integers $p \ge 0$.
+%
 \begin{eqnarray}
@@
 U_T(p,s) & = & \big( e:(\Sigma - [P/s]^0) \then \efail\then \Stop \big)
-\label{eq:ufa}
+\label{eq:uta}
 \\ & & \extchoice \nonumber
 \\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
-\label{eq:ufb}
+\label{eq:utb}
 \\ & & \extchoice \nonumber
 \\ & & (\#s < p) \& \big( e:[P/s]^0 \then U_T(p,s.e) \big)
-\label{eq:ufc}
+\label{eq:utc}
 \\ & & \extchoice \nonumber
 \\ & & (\#s = p) \& \big( \epass\then \Stop  \big)
-\label{eq:ufd}
+\label{eq:utd}
 \end{eqnarray}
-
+%
 The difference between adaptive tests $U_T(p)$ for trace refinement and
-$U_F(p)$ for failures refinement consists in the fact that the former do not ``probe''
-the SUT with respect to minimal sets of events to be accepted without blocking.
+$U_F(p)$ for failures refinement consists in the fact that the former do not
+``probe'' the SUT with respect to minimal sets of events to be accepted
+without blocking.
 
 % ==========================================================================
-The existence of complete, finite test suites is expressed in analogy to Theorem~\ref{th:failurestest}. A noteworthy difference consists in the fact that
-the complete suite for trace refinement just needs the single adaptive test case 
-$U_T(pq-1)$, while the complete failures test suite required the execution of 
-$\{ U_F(0),\dots,U_F(pq-1)\}$. The reason for this is that $U_T(pq-1)$ identifies trace
-errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for erroneous
-acceptances at the end of each trace of length $(pq -1)$.
+The existence of complete, finite test suites is expressed in analogy to
+Theorem~\ref{th:failurestest}. A noteworthy difference is that the complete
+suite for trace refinement just needs the single adaptive test case
+$U_T(pq-1)$, while failures refinement required the execution of $\{
+U_F(0),\dots,U_F(pq-1)\}$. The reason for this is that $U_T(pq-1)$ identifies
+trace errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes
+for erroneous acceptances at the end of each trace of length $(pq -1)$.
 
 \begin{theorem}\label{th:tracetest}
-Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ 
-whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
-the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
-has at most $q$ states with $q \ge p$. 
-Then the test suite 
+Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
+normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
+D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
+whose transition graph has at most $q$ states with $q \ge p$. Then the test
+suite
 \[
@@
 \end{theorem}
-
-We skip the proof of Theorem~\ref{th:tracetest}, since it is just a simplified version of 
-the proof of Theorem~\ref{th:failurestest}.
+%
+We skip the proof of Theorem~\ref{th:tracetest}, since it is just a
+simplified version of the proof of Theorem~\ref{th:failurestest}.

### 2018-08-16 paper/casestudies.tex
small changes to Section 5, and rewording to avoid reference to refinement of states

--- 2018-08-01 paper/casestudies.tex
@@
+% =======================================================================
 \section{Applications}
 \label{sec:case}
-
 % =======================================================================
-\subsection{Testing for Trace Equivalence}
 
@@
 % =======================================================================
-\subsection{Testing for Failures Equivalence}
-
-% =======================================================================
 \subsection{Testing for Failures Refinement}
 
+For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, it is
+advisable to avoid an enumeration of traces $s$ of the reference process.
+Instead, we calculate the following auxiliary functions from $P$'s transition
+graph.
+%
+\begin{eqnarray*}
+\text{initials} & : & N \fun \power (\Sigma)
+\\
+\text{minHit} & : & N \fun \power\power(\Sigma)
+\end{eqnarray*}
+%
+In a state $n = G(P)/s$, the set $\text{initials(n)}$ equals the events
+labelling outgoing edges of $n$, so  $\text{initials(n)} = [P/s]^0$. Function
+$\text{minHit}$ maps $n$ to the set of all minimal hitting sets associated
+with the minimal acceptances of $n$, so $\text{minHit}(n) =
+\text{minHit}(P/s)$. Using the transition function $t$ and the above
+functions, $U_F(p)$ can be re-written as the failures-equivalent CSP process
+below.
+%
+\begin{eqnarray}
+U_F^1(p) & = & U_F^1(p,0,\ii n)
+\\
+U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{initials}(n)) @ e \then \efail\then \Stop \big)
+\label{eq:xufa}
+\\ & & \extchoice \nonumber
+\\ & & (\text{initials}(n) = \varnothing)    \&   \big( \epass \then \Stop \big)
+\label{eq:xufb}
+\\ & & \extchoice \nonumber
+\\ & & (k < p) \& \big(\Extchoice e:\text{initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
+\label{eq:xufc}
+\\ & & \extchoice \nonumber
+\\ & & (k = p) \& \big( \sqcap_{H\in\text{minHit}(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
+\label{eq:xufd}
+\end{eqnarray}
+
+% .....................................................................................
+ \begin{figure}
+ %%\hspace*{-40mm}
+ \begin{center}
+\includegraphics[width=\textwidth]{z_acc.pdf}
+\end{center}
+%%\vspace*{-10mm}
+\caption{Normalised transition graph of faulty implementation $Z$  from Example~\ref{ex:uf1tests}.}
+ \label{fig:tgZ}
+ \end{figure}
+% .......................................................................................
+
+\begin{example}
+\label{ex:uf1tests} Consider the following erroneous implementation $Z$ of
+process $P$ from Example~\ref{ex:a} from the point of view of failures
+refinement.
+\begin{eqnarray*}
+Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
+\\
+Q_1 & = & a\then Z \extchoice c\then Z
+\\
+R_1(r_{max},k) & = & (k < r_{max}) \& \big( c\then R_1(r_{max},k+1)\extchoice b\then Z  \big)
+\\ & & \extchoice
+\\ & & (k = r_{max}) \& \big( c\then R_1(r_{max},r_{max})\intchoice b\then Z  \big)
+\end{eqnarray*}
+It is easy to see (and can be checked with FDR) that $Z$ is trace-equivalent
+to $P$. While $k < r_{max}$, $Z$ also accepts the same sets of events as $P$.
+When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$ is
+fulfilled, however, $R_1(r_{max},k)$ makes an internal choice, instead of
+offering an external choice, and refinement does not hold.
+% alcc: I don't think there is a notion of refinement for state?
+%so this process state no
+%longer refines the corresponding process state $R$ of the reference process
+%$P$ in the failures model.
+Fig.~\ref{fig:tgZ} shows the normalised transition graph of $Z$ for $r_{max}
+= 3$.
+
+Running the test $U_F^1(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p =
+4$ states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test
+depth to be used according to Theorem~\ref{th:failurestest}), tests
+$U_F^1(0),\dots, U_F^1(3)$ are passed by $Z$, but $Z$ fails $U_F^1(4)$,
+because after  execution trace
+\[
+s = a.c.c.c, \qquad\text{(note that $G(Z)/s = \text{Node\_3}$ according to Fig.~\ref{fig:tgZ})},
+\]
+it may be the case that $Z$ accepts only $\{b\}$ due to the internal choice
+and $U_F^1(4)$ -- also due to internal choice -- accepts only the minimal
+hitting set $\{ c \}$ and the event $a\in (\Sigma - [P/s]^0)$. So,
+$(Z\parallel[\Sigma] U_F^1(4))/s$ deadlocks, and the $pass$ event cannot be
+produced. Another failing execution arises if $Z/s$ chooses to accept only
+$\{c \}$, while $U_F^1(4))/s$ choses to accept only $\{a,b\}$. Therefore,
+\[
+(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F^1(4)),
+\]
+and the test fails.
+\xbox
+\end{example}
+
 % ======================================================================

### 2018-08-19 paper/finitecomplete-failures.tex
Improvements according to Ana's questions and suggestions.

@@
 % ==========================================================================
-
-Given a finite-state CSP process $P$ and its normalised transition graph
-\[
-G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
-\]
-suppose that $V\subseteq\Sigma^*$ is a
-prefix-closed set  of sequences of events. By $t(\ii n,V)$ we denote the set
-\[
-t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
-\]
-of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$.
-
-\begin{lemma}
-\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
-graph $G(P)$, such that all states in $N$ are reachable. \fixme{alcc: can a
-normalised graph for a process have states that are unreachable?} Let
-$V\subseteq\Sigma^*$ be a finite prefix-closed set of sequences of events.
-Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is, $|t(\ii
-n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended
-by any event of $\Sigma$. Then $G(P)$ reaches at least $(k+1)$ nodes under
-$V\cup V.\Sigma$.
-\end{lemma}
-\begin{proof}
-Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
-there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
-with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
-$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
-$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
-Note, however, that it is not necessarily the case that $s_1\in V$.
-
-Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such
-that $G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
-$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
-= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
-n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
-node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
-\end{proof}
-\fixme{alcc: Explain the structure of the section? Can the above result be in
-a section, rather than the introduction of the section?}
 
@@
 A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$
-\fixme{alcc: add an example?} that operates on the same alphabet as $P$,
+\fixme{alcc: add an example?}\fxnote{jp: I think that the following explanations, together with examples in Section~\ref{sec:case}, suffice.} that operates on the same alphabet as $P$,
 synchronising over alphabet $\Sigma$. Therefore, a test execution is any
@@
 
-After this informal explanation of tests representing adaptive test cases, we are ready to prove the main theorem of this paper.
+% =========================================================================
+\subsection{A Finite Complete Test Suite}
+
+After the informal explanation of tests representing adaptive test cases, we are ready to state the main theorem of this paper.
 
@@
 \end{theorem}
+
+The proof of the theorem follows from the following lemmas. The first states that test
+suite $\TS_F$ is sound, the second states that the suite is also exhaustive.
+
+\begin{lemma}\label{lemma:mainfsound}
+Test suite $\TS_F$ generated from CSP process $P$, as specified in Theorem~\ref{th:failurestest}, is passed by
+every CSP process $Q$ satisfying $P\lessdet_F Q$.
+\end{lemma}
 \begin{proof}
-To prove soundness of $\TS_F$, suppose that $Q\in{\cal D} \wedge P\lessdet_F
-Q$. In this case, $\trc(Q)\subseteq \trc(P)$, and Lemma~\ref{lemma:hseta}
-implies that for all traces $s$ of $Q$, every $H$ in $\text{minHit}(P/s)$ is
-a hitting set for $\minaccs(Q/s)$. So, when running in parallel with $Q$, any
+Suppose that $P\lessdet_F Q$. Then 
+$\trc(Q)\subseteq \trc(P)$. Therefore, when running in parallel with $Q$, any
 adaptive test $U_F(p)$ will always enter the branches (\ref{eq:ufb}),
 (\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction for
-$U_F(p,s)$. Branch (\ref{eq:ufb}) leads always to a PASS verdict, and branch
+$U_F(p,s)$. Branch~(\ref{eq:ufa}) can never be entered in the prallel execution of $Q$ and
+$U_F(p)$, because $[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.
+
+Moreover,  Lemma~\ref{lemma:hseta}
+implies that for all traces $s\in\trc(Q)$, every $H$ in $\text{minHit}(P/s)$ is
+a hitting set for $\minaccs(Q/s)$. Branch (\ref{eq:ufb}) leads always to a PASS verdict, and branch
 (\ref{eq:ufc}) to test continuation without a verdict. For the last branch,
@@
 $Q$ is free of livelocks, because otherwise the PASS-events might not become
-visible, due to unbounded sequences of hidden events performed by $Q$. Note
-further, that this proof did not refer in any way to the size of $Q$'s
-normalised transition graph, so the test suite is sound for {\it all}
-non-divergent CSP process refining $P$. \fixme{alcc: Soundness does not
+visible, due to unbounded sequences of hidden events performed by $Q$. 
+%Note
+%further, that this proof did not refer in any way to the size of $Q$'s
+%normalised transition graph, so the test suite is sound for {\it all}
+%non-divergent CSP process refining $P$. 
+\fixme{alcc: Soundness does not
 depend on the fault domain. I would have it as a separate theorem.}
+\fxnote{jp: this is a good idea, I have split the proof of the main theorem in 2 lemmas.}
+\xbox
+\end{proof}
 
-To prove exhaustiveness, consider a process $Q\in{\cal D}$ with
+\begin{lemma}\label{lemma:mainfexhaustive}
+Test suite $\TS_F$ specified im Theorem~\ref{th:failurestest} is exhaustive for the fault 
+model specified there.
+\end{lemma}
+\begin{proof}
+Consider a process $Q\in{\cal D}$ with
 $P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
@@
 
-For the first case, consider a  trace $s.e\in\trc(Q)$ such that
+For Case~1, consider a  trace $s.e\in\trc(Q)$ such that
 $s\in\trc(P)$, but $s.e\not\in\trc(P)$. Such a trace always exists because
@@
 of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}.
-From the construction of $G$ described there, we know that $G$ has at most
+Suppose that $G/s = (n_P,n_Q)$.
+The length of $s$ is not known, but 
+from the construction of $G$,  we know that $G$ has at most
 $pq$ reachable states, because $G(P)$ has $p$ states, and $G(Q)$ has at most
-$q$ states. Suppose that $G/s = (n_P,n_Q)$. Applying
-Lemma~\ref{lemma:extendV} implies that this state can be reached by a trace
+$q$ states.  Applying
+Lemma~\ref{lemma:reachproduc} implies that $(n_P,n_Q)$ can be reached by a trace
 $u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
@@
 graph product.}
+\fxnote{jp: I do hope that this is now clearer from the new structure of the lemmas}
 
-For the Case~2, we note that trace $s$ is again a trace of the product graph
-$G$. Therefore, $G/s$ can again be reached by a trace
+For   Case~2, we note that trace $s$ is again a trace of the product graph
+$G$, but we do not know its length. Again, by applying Lemma~\ref{lemma:reachproduc},
+we know that the state   $G/s$ can   be reached by a trace
 $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#

### 2018-08-19 paper/introduction.tex
Improvements according to Ana's questions and suggestions.

@@
 
-This paper presents complete black-box test suites for CSP processes that are
+This paper presents complete black-box test suites for software and systems modelled using the {\it Communicating Sequential Processes (CSP)} process 
+algebra~\cite{Hoare:1985:CSP:3921,Roscoe:1997:TPC:550448}. This process algebra has been widely used for the specification and verification of communicating 
+concurrent control systems, many of these applications have been described in the monograph~\cite{Roscoe2010} and the references 
+contained there.
+
+The complete testing theory presented here applies to 
+CSP processes that are
 divergence-free\footnote{The assumption of divergence freedom is usually
@@
 and deadlock.} and interpreted both in the trace and the failures semantics.
-Our results complement work by two of the authors
+Our results complement work published 
 in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
 collections of processes refining a  ``most general'' fault domain member.
-With that concept, complete test suites may be finite or infinite. Yhis gives
+With that concept, complete test suites may be finite or infinite. This gives
 important insight into the theory of fault-domain testing for CSP, but we are
-particularly interested in finite suites when it comes to practical
+particularly interested in {\it finite} suites when it comes to practical
 application. While~\cite{DBLP:conf/pts/CavalcantiS17} requires additional
@@
 one-one-correspondence to a node of the normalised transition graph. Inspired
-by the way that fault-domains are specified for finite state machines, we
+by the way that fault-domains are specified for finite state machines (FSMs), we
 define them as the set of CSP processes whose normalised transition graphs do
@@
 {\it finite}, complete test suites associated with this class of fault
-domains and conformance relations.
+domains and conformance relations. Moreover, our result is not a simple transcription of
+existing knowledge  about finite, complete test suites for finite state machines:
+the capabilities of CSP to express nondeterminism in a more fine-grained way than it is possible for FSMs requires a more complex approach to testing systems for conformity in the failures-model than it is required for model-based testing against nondeterministic FSMs, as published, for example, in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}.
 

### 2018-08-19 paper/preliminaries.tex
Improvements according to Ana's questions and suggestions.

@@
 $t$, then there is a transition from $n$ with label $e$. Normalisation of
-$G(P)$ is reflected by the fact that $t$ is a function.
+$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
+described in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable 
+by a sequence of edges labelled by $e_1\dots e_k$ and connecting states 
+$\ii n,n_1,\dots,n_{k-1},n$, such that
+\[
+n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
+n= t(n_{k-1},e_k).
+\]
 
-A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if there
+A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and only if 
+there
 is a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
-$s$. The set of traces of $P$ is denoted by $\trc(P)$. If $s\in\trc(P)$, then
+$s$. The set of traces of $P$ is denoted by $\trc(P)$. We use the analgous notation
+$\trc(G(P))$ for the set of finite, initialised paths through $G(P)$, each path represented by its finite sequence of edge labels.
+If $s\in\trc(P)$, then
 the process corresponding to $P$ after having executed $s$ is denoted by
@@
 sure? (4) does not talk about Acc. Is this important?}
+\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
+removed it.}
 
-Every node of a normalised transition graph can alternatively be labelled
+Exploiting formulas (\ref{eq:accref},\ref{eq:accinitials},\ref{eq:accsubset}),
+every node of a normalised transition graph can alternatively be labelled
 with their minimal acceptances, and this information is equivalent to that
-contained in the maximal refusals. This is the representation used by FDR.
+contained in the maximal refusals. Since process states $P/s$ are equivalently expressed
+by states $G(P)/s$ of $P$'s normalised transition graph, we also write
+$\minaccs(G(P)/s)$ and $\maxrefs(G(P)/s)$ to denote the minimal acceptences and maximal refusals, respectively, labelling graph state $G(P)/s$ and note 
+that $\minaccs(G(P)/s) = \minaccs(P/s)$ and $\maxrefs(G(P)/s) = \maxrefs(P/s)$.
 
@@
 later for trc(P). So, we need to be clearer as to what we mean.}
+\fxnote{jp: I have now introduced notation $\trc(G(P))$ above, so I think that the
+statement of the lemma is now sufficiently clear.}
 %
@@
 %
+
+Given a finite-state CSP process $P$ and its normalised transition graph
+\[
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+\]
+suppose that $V\subseteq\Sigma^*$ is a
+prefix-closed set  of sequences of events. By $t(\ii n,V)$ we denote the set
+\[
+t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
+\]
+of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$.
+The following lemma specifies a simple construction method for such sets $V$ reaching 
+{\it every} node of $N$.
+
+\begin{lemma}
+\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
+graph $G(P)$. \fixme{alcc: can a
+normalised graph for a process have states that are unreachable?}
+\fxnote{jp: You are rights: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
+Let
+$V\subseteq\Sigma^*$ be a finite prefix-closed set of sequences of events.
+Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is, $|t(\ii
+n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended
+by any event of $\Sigma$. Then $G(P)$ reaches at least $(k+1)$ nodes under
+$V\cup V.\Sigma$.
+\end{lemma}
+\begin{proof}
+Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
+there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
+with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
+$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
+$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
+Note, however, that it is not necessarily the case that $s_1\in V$.
+
+Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such
+that $G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
+$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
+= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
+n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
+node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
+\end{proof}
+\fixme{alcc: Explain the structure of the section? Can the above result be in
+a section, rather than the introduction of the section?}
+\fxnote{jp: I have moved this lemma now to the section on preliminaries, this should make the main section on the finite complete test suite easier to read.}
+
 For proving our main theorems, it is necessary to consider the \emph{product}
@@
 \end{eqnarray}
+
+The following lemma will be used in the proof of our main theorem. 
+\begin{lemma}\label{lemma:reachproduc}
+If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
+$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace 
+%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
+of maximal length $(pq-1)$.
+\end{lemma}
+\begin{proof}
+The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
+reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
+$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
+all of its reachable states (there are at most $pq$ of them) under 
+$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
+$V'$ is $(pq-1)$.
+\xbox
+\end{proof}
 
@@
     given, this is conf.}
+    \fxnote{jp: sorry, I do not understand what you are saying here}
 \item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{minHit}(P/s)$, $H$ is
@@
 intersection with $A_Q$. \fixme{alcc: minaccs is used for P above.}
+\fxnote{jp: Above Example 1, I have now introduced the notation $\minaccs(G(P)/s)$ and
+$\maxrefs(G(P)/s)$.}
 Minimising $\overline H$ induces the existence of a minimal hitting set $H\in
@@
 %
-This result is used in the sequel in Section~\ref{}.
+This result is used in the sequel in Section~\ref{sec:finitecompletefails}.
 

### 2018-08-19 paper/finitecomplete-failures.tex
checked the fixes

@@
 A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$
-\fixme{alcc: add an example?}\fxnote{jp: I think that the following explanations, together with examples in Section~\ref{sec:case}, suffice.} that operates on the same alphabet as $P$,
+%\fixme{alcc: add an example?}\fxnote{jp: I think that the following explanations, together with examples in Section~\ref{sec:case}, suffice.} that operates on the same alphabet as $P$,
 synchronising over alphabet $\Sigma$. Therefore, a test execution is any
@@
 \begin{proof}
-Suppose that $P\lessdet_F Q$. Then 
-$\trc(Q)\subseteq \trc(P)$. Therefore, when running in parallel with $Q$, any
-adaptive test $U_F(p)$ will always enter the branches (\ref{eq:ufb}),
-(\ref{eq:ufc}),  or (\ref{eq:ufd}) of the external choice construction for
-$U_F(p,s)$. Branch~(\ref{eq:ufa}) can never be entered in the prallel execution of $Q$ and
-$U_F(p)$, because $[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.
+Suppose that $P\lessdet_F Q$. In this case, in particular, $\trc(Q)\subseteq
+\trc(P)$. Therefore, when running in parallel with $Q$, any adaptive test
+$U_F(p)$ always enters the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or
+(\ref{eq:ufd}) of the external choice construction for $U_F(p,s)$.
+Branch~(\ref{eq:ufa}) can never be entered in the parallel execution of $Q$
+and $U_F(p)$, because $[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.
 
@@
 %non-divergent CSP process refining $P$.
-\fixme{alcc: Soundness does not
-depend on the fault domain. I would have it as a separate theorem.}
-\fxnote{jp: this is a good idea, I have split the proof of the main theorem in 2 lemmas.}
+%\fixme{alcc: Soundness does not
+%depend on the fault domain. I would have it as a separate theorem.}
+%\fxnote{jp: this is a good idea, I have split the proof of the main theorem in 2 lemmas.}
 \xbox
@@
 
-For Case~1, consider a  trace $s.e\in\trc(Q)$ such that
-$s\in\trc(P)$, but $s.e\not\in\trc(P)$. Such a trace always exists because
-$\varepsilon$ is a trace of every process. In this case, $s$ is also a trace
-of the product graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}.
-Suppose that $G/s = (n_P,n_Q)$.
-The length of $s$ is not known, but 
-from the construction of $G$,  we know that $G$ has at most
-$pq$ reachable states, because $G(P)$ has $p$ states, and $G(Q)$ has at most
-$q$ states.  Applying
-Lemma~\ref{lemma:reachproduc} implies that $(n_P,n_Q)$ can be reached by a trace
+For Case~1, consider a  trace $s.e\in\trc(Q)$ such that $s\in\trc(P)$, but
+$s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is a
+trace of every process. In this case, $s$ is also a trace of the product
+graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}. Suppose that
+$G/s = (n_P,n_Q)$. The length of $s$ is not known, but from the construction
+of $G$,  we know that $G$ has at most $pq$ reachable states, because $G(P)$
+has $p$ states, and $G(Q)$ has at most $q$ states. By
+Lemma~\ref{lemma:reachproduc}, $(n_P,n_Q)$ can be reached by a trace
 $u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
@@
 trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$. Since
-$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0$. Therefore, in at
-least one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa})
-with this event $e$, so that the test fails with event $\efail$. Again, the
-assumption of non-divergence of Q is needed for this conclusion. \fixme{alcc:
-I didn't see how Lemma 3 is being applied here. Also, given the definition of
-the test as a process, I'm not sure why you need to make the argument using
-graph product.}
-\fxnote{jp: I do hope that this is now clearer from the new structure of the lemmas}
+$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0$. So, in at least
+one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa}) with
+this event $e$, so that the test fails. Again, the
+assumption of non-divergence of Q is needed for this conclusion. %\fixme{alcc:
+%I didn't see how Lemma 3 is being applied here. Also, given the definition of
+%the test as a process, I'm not sure why you need to make the argument using
+%graph product.}
+%\fxnote{jp: I do hope that this is now clearer from the new structure of the lemmas}
 

### 2018-08-19 paper/introduction.tex
checked the fixes

@@
 
-This paper presents complete black-box test suites for software and systems modelled using the {\it Communicating Sequential Processes (CSP)} process 
-algebra~\cite{Hoare:1985:CSP:3921,Roscoe:1997:TPC:550448}. This process algebra has been widely used for the specification and verification of communicating 
-concurrent control systems, many of these applications have been described in the monograph~\cite{Roscoe2010} and the references 
-contained there.
+This paper presents complete black-box test suites for software and systems
+modelled using the {\it Communicating Sequential Processes (CSP)} process
+algebra~\cite{Hoare:1985:CSP:3921,Roscoe:1997:TPC:550448}. This process
+algebra has been widely used for the specification and verification of
+communicating concurrent control systems. Many of these applications have
+been described in~\cite{Roscoe2010} and in the references there.
 
-The complete testing theory presented here applies to 
-CSP processes that are
+The complete testing theory presented here applies to CSP processes that are
 divergence-free\footnote{The assumption of divergence freedom is usually
@@
 and deadlock.} and interpreted both in the trace and the failures semantics.
-Our results complement work published 
-in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
-collections of processes refining a  ``most general'' fault domain member.
-With that concept, complete test suites may be finite or infinite. This gives
-important insight into the theory of fault-domain testing for CSP, but we are
-particularly interested in {\it finite} suites when it comes to practical
-application. While~\cite{DBLP:conf/pts/CavalcantiS17} requires additional
-criteria to select tests from infinite test suites, here, we further restrict
-fault domains using a graph representation of processes used in model
-checking to obtain test suites that are finite.
+Our results complement work published in~\cite{DBLP:conf/pts/CavalcantiS17}.
+There, fault domains are specified as collections of processes refining a
+``most general'' fault domain member. With that concept, complete test suites
+may be finite or infinite. This gives important insight into the theory of
+fault-domain testing for CSP, but we are particularly interested in {\it
+finite} suites when it comes to practical application.
+While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
+select tests from still infinite test suites, here, we further restrict fault
+domains using a graph representation of processes used in model checking to
+obtain test suites that are finite.
 
@@
 presented in this paper. We observe that every finite-state CSP process can
-be semantically represented as a finite normalised transition graph, whose
-edges are labelled by the events the process engages in, and whose nodes are
-labelled by minimal acceptances or, alternatively, maximal
+be represented as a finite normalised transition graph, whose edges are
+labelled by the events the process engages in, and whose nodes are labelled
+by minimal acceptances or, alternatively, maximal
 refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
@@
 one-one-correspondence to a node of the normalised transition graph. Inspired
-by the way that fault-domains are specified for finite state machines (FSMs), we
-define them as the set of CSP processes whose normalised transition graphs do
-not exceed the size of the reference model's graph by more than a given
+by the way that fault-domains are specified for finite state machines (FSMs),
+we define them as the set of CSP processes whose normalised transition graphs
+do not exceed the size of the reference model's graph by more than a given
 constant.
@@
 {\it finite}, complete test suites associated with this class of fault
-domains and conformance relations. Moreover, our result is not a simple transcription of
-existing knowledge  about finite, complete test suites for finite state machines:
-the capabilities of CSP to express nondeterminism in a more fine-grained way than it is possible for FSMs requires a more complex approach to testing systems for conformity in the failures-model than it is required for model-based testing against nondeterministic FSMs, as published, for example, in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}.
+domains and conformance relations.
+
+Moreover, our result is not a simple transcription of existing knowledge
+about finite, complete test suites for finite state machines: the
+capabilities of CSP to express nondeterminism in a more fine-grained way than
+it is possible for FSMs requires a more complex approach to testing systems
+for conformity in the failures-model than it is required for model-based
+testing against nondeterministic FSMs, as published, for example,
+in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}.
 

### 2018-08-19 paper/preliminaries.tex
checked the fixes

@@
 $G(P)$ is reflected by the fact that $t$ is a function. The graph construction
-described in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable 
+in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable
 by a sequence of edges labelled by $e_1\dots e_k$ and connecting states
@@
 \]
-
+%
 A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and only if
@@
 %
-Note that (\ref{eq:accref}) can be regarded as a definition of minimal
-acceptances by means of maximal refusals. In this case,
-(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
-definition and the fact that refusals are subset-closed. \fixme{alcc: are you
-sure? (4) does not talk about Acc. Is this important?}
-\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
-removed it.}
-
-Exploiting formulas (\ref{eq:accref},\ref{eq:accinitials},\ref{eq:accsubset}),
-every node of a normalised transition graph can alternatively be labelled
-with their minimal acceptances, and this information is equivalent to that
-contained in the maximal refusals. Since process states $P/s$ are equivalently expressed
-by states $G(P)/s$ of $P$'s normalised transition graph, we also write
-$\minaccs(G(P)/s)$ and $\maxrefs(G(P)/s)$ to denote the minimal acceptences and maximal refusals, respectively, labelling graph state $G(P)/s$ and note 
-that $\minaccs(G(P)/s) = \minaccs(P/s)$ and $\maxrefs(G(P)/s) = \maxrefs(P/s)$.
+%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
+%acceptances by means of maximal refusals. In this case,
+%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
+%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
+%sure? (4) does not talk about Acc. Is this important?}
+%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
+%removed it.}
+%
+Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
+(\ref{eq:accsubset}), every node of a normalised transition graph can
+alternatively be labelled with their minimal acceptances, and this
+information is equivalent to that contained in the maximal refusals. Since
+process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
+normalised transition graph, we also write $\minaccs(G(P)/s)$ and
+$\maxrefs(G(P)/s)$ to denote the minimal acceptences and maximal refusals,
+labelling graph state $G(P)/s$ and note that $\minaccs(G(P)/s) =
+\minaccs(P/s)$ and $\maxrefs(G(P)/s) = \maxrefs(P/s)$.
 
@@
 expressed by means of their normalised transition graphs in the following
-way. \fixme{alcc: trc was defined for P, not G(P) and this lemma is used
-later for trc(P). So, we need to be clearer as to what we mean.}
-\fxnote{jp: I have now introduced notation $\trc(G(P))$ above, so I think that the
-statement of the lemma is now sufficiently clear.}
+way. %\fixme{alcc: trc was defined for P, not G(P) and this lemma is used
+%later for trc(P). So, we need to be clearer as to what we mean.}
+%\fxnote{jp: I have now introduced notation $\trc(G(P))$ above, so I think that the
+%statement of the lemma is now sufficiently clear.}
 %
@@
 \end{lemma}
-%
+\fixme{alcc: I'll write an explanation of these results.}
 
@@
 \]
-of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$.
-The following lemma specifies a simple construction method for such sets $V$ reaching 
-{\it every} node of $N$.
+of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$. The
+lemma below specifies a construction method for such sets $V$ reaching {\it
+every} node of $N$.
 
@@
 \label{lemma:extendV} Let $P$ be a CSP process with normalised transition
-graph $G(P)$. \fixme{alcc: can a
-normalised graph for a process have states that are unreachable?}
-\fxnote{jp: You are rights: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
+graph $G(P)$. %\fixme{alcc: can a
+%normalised graph for a process have states that are unreachable?}
+%\fxnote{jp: You are rights: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
 Let
@@
 \end{proof}
-\fixme{alcc: Explain the structure of the section? Can the above result be in
-a section, rather than the introduction of the section?}
-\fxnote{jp: I have moved this lemma now to the section on preliminaries, this should make the main section on the finite complete test suite easier to read.}
+%\fixme{alcc: Explain the structure of the section? Can the above result be in
+%a section, rather than the introduction of the section?}
+%\fxnote{jp: I have moved this lemma now to the section on preliminaries, this should make the main section on the finite complete test suite easier to read.}
 
@@
 $\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
-intersection with $A_Q$. \fixme{alcc: minaccs is used for P above.}
-\fxnote{jp: Above Example 1, I have now introduced the notation $\minaccs(G(P)/s)$ and
-$\maxrefs(G(P)/s)$.}
+intersection with $A_Q$. %\fixme{alcc: minaccs is used for P above.}
+%\fxnote{jp: Above Example 1, I have now introduced the notation $\minaccs(G(P)/s)$ and
+%$\maxrefs(G(P)/s)$.}
 Minimising $\overline H$ induces the existence of a minimal hitting set $H\in

### 2018-08-24 paper/preliminaries.tex
section 2.1 and additional explanations

@@
 
+In this section we present the background material for our work.
+
 % =========================================================================
@@
 
+First, we introduce the CSP notation and refinement notions. We then discuss
+a graph-based semantics for finite-state CSP processes.
+
 % =========================================================================
+\subsubsection*{CSP}
+\label{section:CSP}
 
-\subsubsection*{Communicating Sequential Processes} @todo
-\fxwarning{alcc: I can make this small contribution.}
+CSP is a process algebra for refinement. Using CSP, we model both systems and
+their components using processes. There are characterised by their patterns
+of interactions, modelled by synchronous, instantaneous, and atomic events.
 
-Throughout this paper, the alphabet of CSP processes $P$ is denoted by $\Sigma$.
-Since we are only considering ``classical'' CSP processes, the alphabet is always finite.
+A prefixing operator $e \then P$ defines a process that is ready to engage in
+the event $e$, pending agreement of its environment to synchronising. After
+$e$ occurs, the process behaves as defined by $P$. The environment can be
+other processes, in parallel, or the environment of a system as a whole.
 
-\begin{itemize}
-  \item traces and failures model
-  \item refinement and equivalence
-\end{itemize}
+Two forms of choice support branching behaviour. An external choice $P
+\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
+initial events of $P$ and $Q$. Once synchronisation takes place, the process
+that offered this event is chosen and the other is discarded. In an internal
+choice $P \intchoice Q$, the environment does not have an opportunity to
+interfere:~the choice is made by the process itself.
 
-The FDR tool~\cite{fdr} supports model checking and semantic analyses of CSP
-processes.
+\begin{example}
+  We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
+  ready to engage in the event $a$, and then makes a choice to
+  behave as $Q$ or $R$.
+  \begin{eqnarray*}
+  P & = & a \then (Q\intchoice R)
+  \\
+  Q & = & a \then P \extchoice c \then P
+  \\
+  R & = & b \then P \extchoice c \then R
+  \end{eqnarray*}
+  $Q$, for instance, offers to the environment the choice to engage in $a$
+  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
+  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
+  chosen, the recursion is to $R$.
+  \xbox
+\end{example}
+%
+Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
+external and internal choice operators allow us to define a choice over a
+collection of processes $P(i)$.
+
+There are several parallelism operators. A widely used form of parallelism $P
+\parallel[cs] Q$ defines a process in which the behaviour is characterised by
+those of $P$ and $Q$ in parallel, synchronising on the events in the set
+$cs$. Other forms of parallelism can be defined using this operator.
+
+Interactions that are not supposed to be visible to the environment can be
+hidden. The operator $P \ H$ defines a process that behaves as $P$, with the
+interactions modelled by events in the set $H$ hidden. Frequently, hiding in
+used in conjunction with parallelism, so that communications in a network of
+parallel processes used just to coordinate them are affected
+by the environment.
+
+A rich collection of process operators allow us to define networks of
+parallel processes in a concise and elegant way, and reason about safety,
+liveness, and divergences.  A comprehensive account of the notation is given
+in~\cite{Roscoe2010}.
+
+A distinctive feature of CSP is its treatment of refinement~(as opposed to
+bisimulation), which is convenient for reasoning about program correctness,
+due to its treatment of nondeterminism and divergence.  A variety of semantic
+models capture different notions of refinement. The simplest model
+characterises a process by its possible traces:~sequence of events in which
+it can engage.  We say that a process $P$ is trace-refined by another process
+$Q$, written $P \lessdet_T Q$, if the set $traces(Q)$ of traces of $Q$ is
+contained in the set $traces(P)$ of traces of $P$.
+
+In fact, in every model, subset containment is used to define refinement. The
+model we focus here first is the failures model, which captures both
+sequences of interactions and deadlock behaviour. A failure of a process $P$
+is a pair $(t,X)$ containing a trace $t$ of $P$ and a refusal:~a set $X$ of
+events in which $P$ may refuse to engage, having engaged in the events of
+$t$. The failures model of a process $P$ records all its failures in a set
+$failures(P)$.
+
+Failures refinement, $\lessdet_F$, is defined by set containment, but can
+also be expressed in terms of traces refinement and another conformance
+relation called $conf$ and defined below. The set $Ref(P,t)$ used in this
+definition, and also specified below, contains the refusals of $P$ after the
+trace $t$.
+%
+\begin{displaymath}
+  Q\ conf\ P \defs \forall t: traces(P) \cap traces(Q) @ Ref(Q,t)
+  \subseteq Ref(P,t)
+  \\ %
+  \mbox{where } Ref(P,t) \defs \{~X | (t,X) \in failures(P)~\}
+\end{displaymath}
+%
+Conformance according to $conf$ requires that no new deadlocks are introduced
+in an implementation. The theorem below, proved in~\cite{CG07}, establishes
+that failures refinement amounts to traces refinement and $conf$.
+%
+\begin{theorem}
+  $(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)$
+  \label{thm:fref-tref-conf}
+\end{theorem}
+%
+Process equivalence in each model is given simply by mutual refinement.
+
+Throughout this paper, the alphabet of the processes, that is the set of
+events that are in scope, is denoted by $\Sigma$. Since we are only
+considering ``classical'' CSP processes, the alphabet is always finite. The
+FDR tool~\cite{fdr} supports model checking and semantic analyses of
+finite-state CSP processes.
 
@@
 %
-A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and only if
-there
-is a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
-$s$. The set of traces of $P$ is denoted by $\trc(P)$. We use the analgous notation
-$\trc(G(P))$ for the set of finite, initialised paths through $G(P)$, each path represented by its finite sequence of edge labels.
-If $s\in\trc(P)$, then
-the process corresponding to $P$ after having executed $s$ is denoted by
-$P/s$. Since $G(P)$ is normalised, there is a unique node reached by applying
-the events from $s$ one by one, starting in $\ii n$. Therefore, $G(P)/s$  is
-also well defined. By $[n]^0$ we denote the \emph{initials} of $n$:~the set
-of events occurring as labels in any outgoing transitions.
+A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and
+only if there is a path through $G(P)$ starting  at $\ii n$ whose edge labels
+coincide with $s$. The set of traces of $P$ is denoted by $\trc(P)$. We use
+the analgous notation $\trc(G(P))$ for the set of finite, initialised paths
+through $G(P)$, each path represented by its finite sequence of edge labels.
+If $s\in\trc(P)$, then the process corresponding to $P$ after having executed
+$s$ is denoted by $P/s$. Since $G(P)$ is normalised, there is a unique node
+reached by applying the events from $s$ one by one, starting in $\ii n$.
+Therefore, $G(P)/s$  is also well defined. By $[n]^0$ we denote the
+\emph{initials} of $n$:~the set of events occurring as labels in any outgoing
+transitions.
 $$
@@
 \begin{example}\label{ex:a}
-Consider the CSP process $P$ defined below, and $\Sigma = \{a,b,c\}$.
-\begin{eqnarray*}
-P & = & a \then (Q\intchoice R)
-\\
-Q & = & a \then P \extchoice c \then P
-\\
-R & = & b \then P \extchoice c \then R
-\end{eqnarray*}
-Its transition graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state
-$P/\langle\rangle$ is represented there as Node\_0, with $\{ a\}$ as the
-only minimal acceptance, since $a$ can never be refused, and no other events are
-accepted. Having engaged into $a$, the transition emanating from Node\_0
-leads to Node\_2 representing  the process state $P/a = Q\intchoice R$. The
-internal choice operator induces several minimal acceptances derived from $Q$ and
-$R$. Since these processes accept their initial events in external choice,
-$Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$ and
-$\{b,c\}$. Note that event $c$ can never be refused, since it is contained in each minimal acceptance set.
+Consider the CSP process $P$ in Example~\ref{example:CSP}. Its transition
+graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state $P/\langle\rangle$
+is represented there as Node\_0, with $\{ a\}$ as the only minimal
+acceptance, since $a$ can never be refused, and no other events are accepted.
+Having engaged into $a$, the transition emanating from Node\_0 leads to
+Node\_2 representing  the process state $P/a = Q\intchoice R$. The internal
+choice operator induces several minimal acceptances derived from $Q$ and $R$.
+Since these processes accept their initial events in external choice,
+$Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$ and $\{b,c\}$. Note
+that event $c$ can never be refused, since it is contained in each minimal
+acceptance set.
 
@@
 \end{lemma}
-\fixme{alcc: I'll write an explanation of these results.}
+%
+Traces refinement is, as already explained, characterised by subset inclusion
+of sets of traces, here defined in terms of the normalised graph of a
+process. In~(\ref{eq:failrefa}), we have a result similar to
+Theorem~\ref{thm:fref-tref-conf}, but rely on the fact that refusal sets are
+subset closed. In other words, if a process may refuse to engage in all
+events in a set $X$, then it can refuse to engage in all events of subsets of
+$X$. So, in considering $conf$, it is enough to consider maximal refusals.
+Finally, (\ref{eq:failrefb}) captures the same result using acceptances,
+rather than refusals, which as already observed, can be calculated by taking
+the complement of refusals. So, acceptances of $Q$ are required to have a
+corresponding larger set of acceptances in $P$.
 

### 2018-08-26 paper/finitecomplete-failures.tex
First part of updates for complete version

@@
 \[
-Q\ \pass\ U_F(p) \equiv (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
+Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
 \]
@@
 % =========================================================================
-\subsection{A Finite Complete Test Suite}
+\subsection{Fault Models and Complete Test Suites}
+
+A \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a reference process
+$P$, a conformance relation $\sqsubseteq \in \{\lessdet_T, \lessdet_F\}$, and a fault domain 
+${\cal D}$ which is a set of CSP processes over $P$'s alphabet that may or may not conform to $P$.
+
+A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
+if and only of the following conditions are fulfilled.
+\begin{description}
+\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
+\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$, 
+then $Q$ fails at least one test in $\TS$.
+\end{description}
+
+% =========================================================================
+\subsection{A Finite Complete Test Suite for Failures Refinement}
 
@@
 
-The proof of the theorem follows from the following lemmas. The first states that test
+The proof of the theorem follows from the two lemmas below. The first states that test
 suite $\TS_F$ is sound, the second states that the suite is also exhaustive.
@@
 \begin{proof}
-Suppose that $P\lessdet_F Q$. In this case, in particular, $\trc(Q)\subseteq
-\trc(P)$. Therefore, when running in parallel with $Q$, any adaptive test
-$U_F(p)$ always enters the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or
+Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\ Q$ 
+according to (\ref{eq:failconf}). 
+Since   $\trc(Q)\subseteq
+\trc(P)$,   any adaptive test
+$U_F(p)$ running in parallel with $Q$ will always enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or
 (\ref{eq:ufd}) of the external choice construction for $U_F(p,s)$.
@@
 Moreover,  Lemma~\ref{lemma:hseta}
-implies that for all traces $s\in\trc(Q)$, every $H$ in $\text{minHit}(P/s)$ is
-a hitting set for $\minaccs(Q/s)$. Branch (\ref{eq:ufb}) leads always to a PASS verdict, and branch
+implies that for all traces $s\in\trc(Q)\cap\trc(P)$, every $H$ in $\text{minHit}(P/s)$ is
+a hitting set for $\minaccs(Q/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,s)$ 
+leads always to a PASS verdict, and branch
 (\ref{eq:ufc}) to test continuation without a verdict. For the last branch,
@@
 Consider a process $Q\in{\cal D}$ with
-$P\not\lessdet_F Q$. This non-conformance can be caused in two possible ways.
+$P\not\lessdet_F Q$, According to (\ref{eq:failconf}),  
+this non-conformance can be caused in two possible ways corresponding to the cases $P\not\lessdet_T Q$ and 
+$\neg(P\ conf\ Q)$, respectively:
 \begin{description}

### 2018-08-26 paper/preliminaries.tex
First part of updates for complete version

@@
 % =========================================================================
-
-In this section we present the background material for our work.
 
@@
 % =========================================================================
-\subsubsection*{CSP}
-\label{section:CSP}
+%\subsubsection*{CSP}
+%\label{section:CSP}
 
-CSP is a process algebra for refinement. Using CSP, we model both systems and
+\subsubsection*{Communicating Sequential Processes (CSP)} is a process algebra 
+supporting system development by refinement. Using CSP, we model both systems and
 their components using processes. There are characterised by their patterns
@@
 
-\begin{example}
+\begin{example}\label{example:CSP}
   We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
@@
 Interactions that are not supposed to be visible to the environment can be
-hidden. The operator $P \ H$ defines a process that behaves as $P$, with the
-interactions modelled by events in the set $H$ hidden. Frequently, hiding in
-used in conjunction with parallelism, so that communications in a network of
-parallel processes used just to coordinate them are affected
-by the environment.
+hidden. The operator $P \hide H$ defines a process that behaves as $P$, with the
+interactions modelled by events in the set $H$ hidden. Frequently, hiding is used
+in conjunction with parallelism: it may be desirable to make interal actions of each process in a network invisible, while events happening at their interfaces remain observable.
 
-A rich collection of process operators allow us to define networks of
+%Frequently, hiding is
+%used in conjunction with parallelism, so that communications in a network of
+%parallel processes used just to coordinate them are affected
+%by the environment.
+
+A rich collection of process operators allows us to define networks of
 parallel processes in a concise and elegant way, and reason about safety,
@@
 models capture different notions of refinement. The simplest model
-characterises a process by its possible traces:~sequence of events in which
-it can engage.  We say that a process $P$ is trace-refined by another process
-$Q$, written $P \lessdet_T Q$, if the set $traces(Q)$ of traces of $Q$ is
-contained in the set $traces(P)$ of traces of $P$.
+characterises a process by its possible \emph{traces}; the set $\trc(P)$ denotes
+ the sequences of (non-hidden) 
+events in which
+$P$ can engage.  We say that a process \emph{$P$ is trace-refined by another process
+$Q$}, written $P \lessdet_T Q$, if $\trc(Q)\subseteq \trc(P)$.
 
-In fact, in every model, subset containment is used to define refinement. The
+In fact, in every semantic model, subset containment is used to define refinement. The
 model we focus here first is the failures model, which captures both
 sequences of interactions and deadlock behaviour. A failure of a process $P$
-is a pair $(t,X)$ containing a trace $t$ of $P$ and a refusal:~a set $X$ of
-events in which $P$ may refuse to engage, having engaged in the events of
-$t$. The failures model of a process $P$ records all its failures in a set
-$failures(P)$.
+is a pair $(s,X)$ containing a trace $s$ of $P$ and a refusal:~a set $X$ of
+events in which $P$ may refuse to engage, after having performed the events of
+$s$. The failures model of a process $P$ records all its failures in a set
+$\fails(P)$. 
+Using notation $P/s$ to denote the process $P$ after having engaged into trace $s$,
+the set 
+$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the \emph{refusals} 
+of $P$ after the
+trace $t$. 
+Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure of $P$ and $Y\subseteq X$,
+then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$ follows.
 
-Failures refinement, $\lessdet_F$, is defined by set containment, but can
-also be expressed in terms of traces refinement and another conformance
-relation called $conf$ and defined below. The set $Ref(P,t)$ used in this
-definition, and also specified below, contains the refusals of $P$ after the
-trace $t$.
+Failures refinement, $P \lessdet_F Q$, is defined by set containment 
+$\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed, this implies
+$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. As a consequence, failures 
+refinement implies trace refinement. Therefore, using the additional conformance relation
+\begin{equation}\label{eq:conf}
+  Q\ conf\ P \defs \forall s\in \trc(P) \cap \trc(Q): \refs(Q/s)
+  \subseteq \refs(P/s),
+\end{equation}
+failures refinement can be expressed by $\lessdet_T$ and $conf$.
+\begin{equation}\label{thm:fref-tref-conf}
+(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
+\end{equation}
+
 %
-\begin{displaymath}
-  Q\ conf\ P \defs \forall t: traces(P) \cap traces(Q) @ Ref(Q,t)
-  \subseteq Ref(P,t)
-  \\ %
-  \mbox{where } Ref(P,t) \defs \{~X | (t,X) \in failures(P)~\}
-\end{displaymath}
 %
-Conformance according to $conf$ requires that no new deadlocks are introduced
-in an implementation. The theorem below, proved in~\cite{CG07}, establishes
-that failures refinement amounts to traces refinement and $conf$.
 %
-\begin{theorem}
-  $(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)$
-  \label{thm:fref-tref-conf}
-\end{theorem}
-%
-Process equivalence in each model is given simply by mutual refinement.
+%but can
+%also be expressed in terms of traces refinement and another conformance
+%relation called $conf$ and defined below. The set 
+%$\refs(P,t) \defs \{~X | (t,X) \in \fails(P)~\}$ used in this
+%definition contains the refusals of $P$ after the
+%trace $t$.
+%%
+%\begin{displaymath}
+%  Q\ conf\ P \defs \forall t: \trc(P) \cap \trc(Q) @ \refs(Q,t)
+%  \subseteq \refs(P,t)
+%\end{displaymath}
+%%
+%Conformance according to $conf$ requires that no new deadlocks are introduced
+%in an implementation. The theorem below, proved in~\cite{CG07}, establishes
+%that failures refinement amounts to traces refinement and $conf$.
+%%
+%\begin{theorem}
+%  $(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)$
+%  \label{thm:fref-tref-conf}
+%\end{theorem}
+%%
+%Process equivalence in each model is given simply by mutual refinement.
 
 Throughout this paper, the alphabet of the processes, that is the set of
-events that are in scope, is denoted by $\Sigma$. Since we are only
-considering ``classical'' CSP processes, the alphabet is always finite. The
+events that are in scope, is denoted by $\Sigma$ and supposed to be finite. 
+%Since we are only
+%considering ``classical'' CSP processes, the alphabet is always finite. 
+The
 FDR tool~\cite{fdr} supports model checking and semantic analyses of
@@
 
-% =========================================================================
-\subsubsection*{Normalised Transition Graphs for CSP Processes}
-\label{sec:ntg}
-
-As shown in~\cite{Roscoe:1994:CME:197600}, any finite-state CSP process $P$
-can be represented by a \emph{normalised transition graph}
-$$
-G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
-$$
-with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
-partial \emph{transition function} $t$ maps a node $n$ and an event
-$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
-$t$, then there is a transition from $n$ with label $e$. Normalisation of
-$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
-in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable
-by a sequence of edges labelled by $e_1\dots e_k$ and connecting states
-$\ii n,n_1,\dots,n_{k-1},n$, such that
-\[
-n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
-n= t(n_{k-1},e_k).
-\]
-%
-A finite sequence of events $s\in\Sigma^*$ is a \emph{trace} of $P$, if and
-only if there is a path through $G(P)$ starting  at $\ii n$ whose edge labels
-coincide with $s$. The set of traces of $P$ is denoted by $\trc(P)$. We use
-the analgous notation $\trc(G(P))$ for the set of finite, initialised paths
-through $G(P)$, each path represented by its finite sequence of edge labels.
-If $s\in\trc(P)$, then the process corresponding to $P$ after having executed
-$s$ is denoted by $P/s$. Since $G(P)$ is normalised, there is a unique node
-reached by applying the events from $s$ one by one, starting in $\ii n$.
-Therefore, $G(P)/s$  is also well defined. By $[n]^0$ we denote the
-\emph{initials} of $n$:~the set of events occurring as labels in any outgoing
-transitions.
-$$
-[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
-$$
-We also use this notation for CSP processes:~$[P]^0$ is the set of events $P$
-may engage into, in other words, the initials of $P$ after the empty trace of
-events, that is, $initials(P/\langle\rangle)$ as defined
-in~\cite{Roscoe2010}, for example.
-
-The total function $r$ maps each node $n$ to its \emph{refulsals} $r(n) =
-\refs(n)$. Each element of $r(n)$ is a set of events that the CSP process $P$
-might refuse to engage into, when in a process state corresponding to $n$.
- If $P/s$ is deterministic, its refusals coincide with the set of
-subsets of $\Sigma - [P/s]^0$, including the empty set.
-
 For finite CSP processes, since the refusals of each process state are
-subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}, $\refs(P/s)$ can be
-constructed from the set of \emph{maximal refusals}
-$\maxrefs(P/s)\subseteq\refs(P/s)$. More formally, $\maxrefs(P/s)$ is defined
-as follows.
+subset-closed, $\refs(P/s)$ can be
+constructed from the set of \emph{maximal refusals}; these are defined by 
 %
@@
 Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
-reconstruct the refusals $\refs(P/s)$ by subset-closure as follows.
+reconstruct the refusals $\refs(P/s)$ by  
 %
@@
 state $P/s$. Deterministic process states $P/s$ have exactly the one maximal refusal
-$\Sigma-[P/s]^0$.
+$\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of $P/s$, i.e.,   the events that $P/s$ may engage into.
 
-To see that this approach works only for finite CSP processes, we consider
-the example where $\Sigma$ is infinite. In this case,
-$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
-calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
-defined above. As with refusals, we also use the transition graph-oriented
-notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
-associated with graph state $n$: if $n$ is the state reached in the
-transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
-= \maxrefs(P/s)$.
+% =========================================================================
+\subsubsection*{Normalised Transition Graphs for CSP Processes}
+\label{sec:ntg}
 
-Well-formed normalised transition graphs must not refuse an event of the
-initials of a state in {\it every} refusal applicable in this state; more
-formally,
+As shown in~\cite{Roscoe:1994:CME:197600}, the failures semantics of 
+any finite-state CSP process $P$
+can be represented by a \emph{normalised transition graph}
+$$
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+$$
+with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
+partial \emph{transition function} $t$ maps a node $n$ and an event
+$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
+$t$, then there is a transition, that is, an outgoing edge, from $n$ with label $e$, leading to node $t(n,e)$. Normalisation of
+$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
+in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable
+sequences of edges labelled by $e_1\dots e_k$ and connecting states
+$\ii n,n_1,\dots,n_{k-1},n$, such that
+\[
+n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
+n= t(n_{k-1},e_k).
+\]
 %
-\begin{equation}
-\label{eq:wellformedg}
-\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
-\exists R\in \maxrefs(n): e\not\in R
+By construction, $s\in\Sigma^*$ is a trace of $P$, if and only if 
+there is a path through $G(P)$ starting  at $\ii n$ whose edge labels
+coincide with $s$. In analogy to $\trc(P)$, we use
+notation $\trc(G(P))$ for the set of finite, initialised paths
+through $G(P)$, each path represented by its finite sequence of edge labels, and we
+note that $\trc(P) = \trc(G(P))$. 
+Since $G(P)$ is normalised, there is a unique node
+reached by applying the events from $s$ one by one, starting in $\ii n$.
+Therefore, $G(P)/s$  is also well defined. By $[n]^0$ we denote the
+\emph{initials} of $n$:~the set of events occurring as labels in any outgoing
+transitions.
+$$
+[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
+$$
+The graph construction from~\cite{Roscoe:1994:CME:197600} guarantees that
+$[G(P)/s]^0 = [P/s]^0$ for all traces of $P$.
+
+The total function $r$ maps each node $n$ to a set of non-empty subsets of $\Sigma$. 
+The graph construction guarantees that $r(G(P)/s)$ represents the maximal refusals
+of $P/s$ for all 
+$s\in\trc(P)$. As a consequence,  
+\begin{equation}\label{eq:refsAndR}
+(s,X)\in\fails(P) \iff s\in\trc(G(P)) \wedge \exists R\in r(G(P)/s): X\subseteq R,
 \end{equation}
+so $G(P)$ allows us to re-construct the failures semantics of $P$.
+
 %
-By construction, normalised transition graphs reflect the \emph{failures
-semantics} of finite-state CSP processes:~the traces $s$ of a process are
-defined by the sequences of transitions associated with paths through its
-graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
-r(G(P)/s)$ represent the failures of $P$.
+%
+%To see that this approach works only for finite CSP processes, we consider
+%the example where $\Sigma$ is infinite. In this case,
+%$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
+%calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
+%defined above. As with refusals, we also use the transition graph-oriented
+%notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
+%associated with graph state $n$: if $n$ is the state reached in the
+%transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
+%= \maxrefs(P/s)$.
+
+%Well-formed normalised transition graphs must not refuse an event of the
+%initials of a state in {\it every} refusal applicable in this state; more
+%formally,
+%%
+%\begin{equation}
+%\label{eq:wellformedg}
+%\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
+%\exists R\in r(n): e\not\in R
+%\end{equation}
+%%
+%
+%
+%By construction, normalised transition graphs reflect the \emph{failures
+%semantics} of finite-state CSP processes:~the traces $s$ of a process are
+%defined by the sequences of transitions associated with paths through its
+%graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
+%r(G(P)/s)$ represent the failures of $P$.
+
+% =========================================================================
+\subsubsection*{Acceptances}
+\label{sec:accs}
 
@@
 \emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
-is also useful. An acceptance set of $P/s$ is a subset of the initials
-$[P/s]^0$, i.e., a subset of events labelling  outgoing transitions of
+is also useful. An acceptance set of $P/s$ is a subset of its initials
+$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
 $G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
@@
 normalised transition graph, we also write $\minaccs(G(P)/s)$ and
-$\maxrefs(G(P)/s)$ to denote the minimal acceptences and maximal refusals,
-labelling graph state $G(P)/s$ and note that $\minaccs(G(P)/s) =
-\minaccs(P/s)$ and $\maxrefs(G(P)/s) = \maxrefs(P/s)$.
+note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply 
+\begin{equation}\label{eq:maxrefsminaccs}
+\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
+\end{equation}
 
@@
 Consider the CSP process $P$ in Example~\ref{example:CSP}. Its transition
-graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state $P/\langle\rangle$
+graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state $P/\varepsilon$ 
+($\varepsilon$ denotes the empty trace)
 is represented there as Node\_0, with $\{ a\}$ as the only minimal
@@
 
-Refinement relations between finite-state CSP processes $P, Q$ can be be
-expressed by means of their normalised transition graphs in the following
-way. %\fixme{alcc: trc was defined for P, not G(P) and this lemma is used
-%later for trc(P). So, we need to be clearer as to what we mean.}
-%\fxnote{jp: I have now introduced notation $\trc(G(P))$ above, so I think that the
-%statement of the lemma is now sufficiently clear.}
-%
+Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
+expressed by means of their normalised transition graphs 
+$$G(P) = ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))$$ 
+and 
+$$G(Q) = ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))$$
+in the following
+way. 
 \begin{lemma}
@@
   \\
+  \label{eq:failconf}
+  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge P\ conf\ Q
+  \\
   \label{eq:failrefa}
-  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
-  \\ & & \forall s\in\trc(G(Q)), R_Q\in\maxrefs(G(Q)/s):  \nonumber
+  P\ conf\ Q & \Leftrightarrow & \nonumber
+  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
   \\ & & \tabd
-  \exists R_P\in\maxrefs(G(P)/s): R_Q\subseteq R_P
+  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
   \\
   \label{eq:failrefb}
-  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
-  \\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
+   & \Leftrightarrow & 
+    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
    \\ & & \tabd
@@
 \end{lemma}
+%\begin{lemma}
+%  \label{lemma:tgtrcref}
+%  \begin{eqnarray}
+%  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
+%  \label{eq:trcrefa}
+%  \\
+%  \label{eq:failrefa}
+%  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
+%  \\ & & \forall s\in\trc(G(Q)), R_Q\in r_Q(G(Q)/s):  \nonumber
+%  \\ & & \tabd
+%  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
+%  \\
+%  \label{eq:failrefb}
+%  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
+%  \\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
+%  \\ & & \tabd
+%  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
+%  \end{eqnarray}
+%  \xbox
+%\end{lemma}
 %
-Traces refinement is, as already explained, characterised by subset inclusion
-of sets of traces, here defined in terms of the normalised graph of a
-process. In~(\ref{eq:failrefa}), we have a result similar to
-Theorem~\ref{thm:fref-tref-conf}, but rely on the fact that refusal sets are
-subset closed. In other words, if a process may refuse to engage in all
-events in a set $X$, then it can refuse to engage in all events of subsets of
-$X$. So, in considering $conf$, it is enough to consider maximal refusals.
-Finally, (\ref{eq:failrefb}) captures the same result using acceptances,
-rather than refusals, which as already observed, can be calculated by taking
-the complement of refusals. So, acceptances of $Q$ are required to have a
-corresponding larger set of acceptances in $P$.
+
+Formula (\ref{eq:trcrefa}) reflects trace refinement in terms of graph traces.  Formula~(\ref{eq:failrefa}) states how the $conf$ conformance relation can be expressed
+by means of the maximal refusal functions of the graphs involved, and (\ref{eq:failrefb})
+states the same in terms of the minimal acceptances that can be derived from the 
+maximal refusal functions by means of (\ref{eq:maxrefsminaccs}).
+
+%, we have a result similar to
+%Theorem~\ref{thm:fref-tref-conf}, but rely on the fact that refusal sets are
+%subset closed. In other words, if a process may refuse to engage in all
+%events in a set $X$, then it can refuse to engage in all events of subsets of
+%$X$. So, in considering $conf$, it is enough to consider maximal refusals.
+%Finally, (\ref{eq:failrefb}) captures the same result using acceptances,
+%rather than refusals, which as already observed, can be calculated by taking
+%the complement of refusals. So, acceptances of $Q$ are required to have a
+%corresponding smaller set of acceptances in $P$.
 
@@
 
-\subsubsection*{Tool Considerations}
-FDR provides an API~\cite{fdrmanual} that can be used to construct normalised
-transition graphs for CSP processes. The FDR graph nodes are labelled by
-\emph{minimal acceptances} instead of maximal refusals as described above.
-Since such a minimal acceptance set is the complement of a maximal refusal,
-the function $r$ mapping states to their refusals can be implemented by
-taking the complements of all minimal acceptances and then building all their
-subsets. For practical applications, the subset closure is never constructed
-in an explicit way; instead, sets are checked with respect to containment in
-a maximal refusal.
-
-@todo
+%\subsubsection*{Tool Considerations}
+%FDR provides an API~\cite{fdrmanual} that can be used to construct normalised
+%transition graphs for CSP processes. The FDR graph nodes are labelled by
+%\emph{minimal acceptances} instead of maximal refusals as described above.
+%Since such a minimal acceptance set is the complement of a maximal refusal,
+%the function $r$ mapping states to their refusals can be implemented by
+%taking the complements of all minimal acceptances and then building all their
+%subsets. For practical applications, the subset closure is never constructed
+%in an explicit way; instead, sets are checked with respect to containment in
+%a maximal refusal.
+%
+%@todo
 
@@
 
-% =========================================================================
-\subsection{Test Cases and Complete Test Suites}
-\label{sec:cspcompletedef}
-% =========================================================================
-
-@todo
+%% =========================================================================
+%\subsection{Test Cases and Complete Test Suites}
+%\label{sec:cspcompletedef}
+%% =========================================================================
+%
+%@todo
 
@@
 \label{lemma:hseta}
-Let $P, Q$ be two finite-state CSP processes satisfying $P\lessdet_T Q$.
+Let $P, Q$ be two finite-state CSP processes.
+%% satisfying $P\lessdet_T Q$.
 For each $s\in\trc(P)$,
@@
 \begin{enumerate}
-\item $P\lessdet_F Q$ \fixme{Note that if traces refinement is already
-    given, this is conf.}
-    \fxnote{jp: sorry, I do not understand what you are saying here}
+\item $P\ conf\ Q$
+
 \item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{minHit}(P/s)$, $H$ is
@@
 \begin{proof}
-For showing ``$1 \Rightarrow 2$'', assume that $P\lessdet_F Q$ and suppose
-that $s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref},
+For showing ``$1 \Rightarrow 2$'', assume   $P\ conf\ Q$ and  
+  $s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref},
 (\ref{eq:failrefb}), states that
@@
 
-To prove ``$2 \Rightarrow 1$'', assume $P\lessdet_T Q$, but $P\not\lessdet_F
-Q$. According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
+To prove ``$2 \Rightarrow 1$'', assume that that Statement~2 of the lemma holds but 
+that  $P\ conf\ Q$ does {\it not} hold. 
+According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
 exists $s\in\trc(P)\cap \trc(Q)$ such that
@@
 \]
-Let $A$ be such a set $A_Q$ fulfilling (*).
+Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
 Define
@@
 $\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
-intersection with $A_Q$. %\fixme{alcc: minaccs is used for P above.}
-%\fxnote{jp: Above Example 1, I have now introduced the notation $\minaccs(G(P)/s)$ and
-%$\maxrefs(G(P)/s)$.}
-Minimising $\overline H$ induces the existence of a minimal hitting set $H\in
+intersection with $A$. 
+Minimising $\overline H$ yields   a minimal hitting set $H\in
 \text{minHit}(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a

### 2018-08-27 paper/abstract.tex
small changes to introduction
Abstract updated to a shorter version

@@
-In this paper, new contributions to testing Communicating Sequential
-Processes (CSP) are presented, with focus on the generation of complete,
-finite test suites. A test suite is complete if it can uncover every
-conformance violation of the system under test with respect to a reference
-model. Both reference models and implementation behaviours are represented as
-CSP processes. As conformance relation, we consider trace equivalence and
-trace refinement, as well as failures equivalence and failures refinement.
-Complete black-box test suites here rely on the fact that the SUT's true
-behaviour is represented by a member of a fault-domain, that is, a collection
-of CSP processes that may or may not conform to the reference model. We
-define fault domains by bounding the number of excessive states occurring in
-a fault domain member's representation as a normalised transition graph, when
-comparing it to the number of states present in the graph of the reference
-model. This notion of fault domains is quite close to the way they are
-defined for finite state machines, and these fault domains guarantee the
-existence of {\it finite} complete test suites.
+In this paper, two new contributions for model-based testing  
+with   Communicating Sequential Processes (CSP) are presented. 
+For a given CSP process representing the reference model, test suites 
+checking the conformance relations 
+trace refinement and  failures refinement are constructed, and their finiteness
+and completeness (i.e.
+capability to uncover conformity violations) is proven.
+While complete test suites for CSP processes have been previously investigated by several authors, a sufficient condition for their finiteness is presented here for the first time.
+Moreover, an optimal technique for
+testing the admissibility of an implementation's nondeterministic behaviour is described.

### 2018-08-27 paper/finitecomplete-failures.tex
updates

@@
 \subsection{Test Cases for Verifying CSP Failures Refinement}
+
+In the domain of process algebras, test cases are typically represented 
+by processes interacting concurrently with the SUT 
+process~\cite{Hennessy:1988:ATP:50497}.
+They synchronise with the SUT over its visible events and use some additional events outside the SUT's alphabet to express whether the test execution  passed or failed.
+%%%, or if no verdict could be obtained.
 
@@
 %
-A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$
-%\fixme{alcc: add an example?}\fxnote{jp: I think that the following explanations, together with examples in Section~\ref{sec:case}, suffice.} that operates on the same alphabet as $P$,
+A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$,
 synchronising over alphabet $\Sigma$. Therefore, a test execution is any
@@
 because $\#s = p$. This is explained further in the next paragraphs.
+
+Observe that the number of possible executions of $Q\parallel[\Sigma] U_F(p)$ is finite,
+because the number of traces $s$ with maximal length $(p+1)$ is finite and the 
+sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and $\text{minHit}(P/s)$ are finite.
 

### 2018-08-27 paper/introduction.tex
updates

@@
 
-While MBT is applied in different flavours, we consider the most effective
-variant to be the one where test cases and concrete test data, as well as
-checkers for the expected results~(\emph{test oracles}), are automatically
-generated from a reference model. This guarantees maximal return of the
-investment of time and effort to create the test model. The test suites
-generated in this way, however, usually have different test strength,
-depending on the generation algorithms applied.
-
-For the safety-critical domain, test suites with guaranteed fault coverage
+In the safety-critical domains, test suites with guaranteed fault coverage
 are of particular interest. For black-box testing, guarantees can be given
@@
 specified by a \emph{fault domain}:~a set of models that may or may not
-conform to the SUT. The so-called \emph{complete} test strategies guarantee
-to uncover every conformance violation of the SUT with respect to a reference
-model, provided that the true SUT behaviour is captured by a member of the
+conform to a given reference model. 
+The so-called \emph{complete} test strategies guarantee
+to accept every  system under test (SUT) conforming to the reference model, while  uncovering every conformance violation, provided that the true SUT behaviour is captured by a member of the
 fault domain.
@@
 description of reactive control systems in many publications over almost five
-decades. Industrial success has also been reported.
+decades. Many of these applications have
+been described in~\cite{Roscoe2010} and in the references there.
+Industrial success has also been reported, see, for example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.
 
@@
 This paper presents complete black-box test suites for software and systems
-modelled using the {\it Communicating Sequential Processes (CSP)} process
-algebra~\cite{Hoare:1985:CSP:3921,Roscoe:1997:TPC:550448}. This process
-algebra has been widely used for the specification and verification of
-communicating concurrent control systems. Many of these applications have
-been described in~\cite{Roscoe2010} and in the references there.
-
-The complete testing theory presented here applies to CSP processes that are
-divergence-free\footnote{The assumption of divergence freedom is usually
+modelled using CSP.
+They can be generated for divergence-free\footnote{The assumption of divergence freedom is usually
 applied in black-box testing, since it cannot distinguish between divergence
-and deadlock.} and interpreted both in the trace and the failures semantics.
+and deadlock.} finite-state CSP processes with finite alphabets,   
+interpreted both in the trace and the failures semantics.
 Our results complement work published in~\cite{DBLP:conf/pts/CavalcantiS17}.
@@
 select tests from still infinite test suites, here, we further restrict fault
-domains using a graph representation of processes used in model checking to
+domains using a graph representation of processes (originally elaborated
+for the purpose of model checking) to
 obtain test suites that are finite.
@@
 Our complementary approach to the definition of CSP fault domains is
-presented in this paper. We observe that every finite-state CSP process can
+presented in this paper. We observe that the failures semantics of 
+every finite-state CSP process can
 be represented as a finite normalised transition graph, whose edges are
@@
 by the way that fault-domains are specified for finite state machines (FSMs),
-we define them as the set of CSP processes whose normalised transition graphs
-do not exceed the size of the reference model's graph by more than a given
+we define them here as the set of CSP processes whose normalised transition graphs
+do not exceed the order (i.e.~the number of nodes) 
+of the reference model's graph by more than a given
 constant.
 
-The main contribution of this paper is the proof that for fault domains of
-the described type, finite, complete test suite generation methods can be
-given for testing against trace and failures refinement and equivalence. The
+The main contribution of this paper consists in the construction of two test suites
+verifying the conformance relations \emph{trace refinement} and  \emph{failures refinement}, respectively, for any given non-divergent, finite-state, finite alphabet 
+CSP reference process. We prove their completeness with respect to fault domains
+of the described type.  The
 existence of -- possibly infinite -- complete test suites has been
@@
 testing against nondeterministic FSMs, as published, for example,
-in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}.
+in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the approach to testing the admissibility of an SUT's nondeterministic behaviour is optimal in the
+sense that it cannot be established with fewer test cases.
 
-It should be noted that the presented results are of a theoretical
-nature:~despite being finite, the resulting test suite size will still be too
-large to be applied in practice to real-world problems. We discuss a number
-of promising options how the test suite size can be reduced to become
-practically applicable.
+%It should be noted that the presented results are of a theoretical
+%nature:~despite being finite, the resulting test suite size will still be too
+%large to be applied in practice to real-world problems. We discuss a number
+%of promising options how the test suite size can be reduced to become
+%practically applicable.
 
@@
 we present the background material relevant to our work.
+In Section~\ref{sec:finitecompletefails}, finite complete test suites for verifying failures refinement are presented.
+Test suites checking trace refinement may be regarded as a simplified version of the former class, they are presented in Section~\ref{sec:finitecomplete}.
+A sample test suite is presented in Section~\ref{sec:case}.
+The results obtained are discussed in Section~\ref{sec:conc}.
 
-@todo
+References to related work are given throughout the paper where appropriate.
 

### 2018-08-27 paper/preliminaries.tex
updates

@@
 
-First, we introduce the CSP notation and refinement notions. We then discuss
-a graph-based semantics for finite-state CSP processes.
+%First, we introduce the CSP notation and refinement notions. We then discuss
+%a graph-based semantics for finite-state CSP processes.
 
@@
 supporting system development by refinement. Using CSP, we model both systems and
-their components using processes. There are characterised by their patterns
+their components using processes. They are characterised by their patterns
 of interactions, modelled by synchronous, instantaneous, and atomic events.
@@
   We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
-  ready to engage in the event $a$, and then makes a choice to
-  behave as $Q$ or $R$.
+  ready to engage in the event $a$, and then makes an internal  choice to
+  behave like either $Q$ or $R$.
   \begin{eqnarray*}
@@
 In fact, in every semantic model, subset containment is used to define refinement. The
-model we focus here first is the failures model, which captures both
-sequences of interactions and deadlock behaviour. A failure of a process $P$
-is a pair $(s,X)$ containing a trace $s$ of $P$ and a refusal:~a set $X$ of
+model we focus on first is the failures model, which captures both
+sequences of interactions and deadlock behaviour. A \emph{failure} of a process $P$
+is a pair $(s,X)$ containing a trace $s$ of $P$ and a \emph{refusal}:~a set $X$ of
 events in which $P$ may refuse to engage, after having performed the events of
@@
 the set 
-$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the \emph{refusals} 
+$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the  refusals
 of $P$ after the
-trace $t$. 
+trace $s$. 
 Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure of $P$ and $Y\subseteq X$,
@@
 $(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. As a consequence, failures 
-refinement implies trace refinement. Therefore, using the additional conformance relation
+refinement implies trace refinement. Therefore, using the auxiliary 
+conformance relation
 \begin{equation}\label{eq:conf}
@@
 \end{equation}
-
-%
-%
-%
-%but can
-%also be expressed in terms of traces refinement and another conformance
-%relation called $conf$ and defined below. The set 
-%$\refs(P,t) \defs \{~X | (t,X) \in \fails(P)~\}$ used in this
-%definition contains the refusals of $P$ after the
-%trace $t$.
-%%
-%\begin{displaymath}
-%  Q\ conf\ P \defs \forall t: \trc(P) \cap \trc(Q) @ \refs(Q,t)
-%  \subseteq \refs(P,t)
-%\end{displaymath}
-%%
-%Conformance according to $conf$ requires that no new deadlocks are introduced
-%in an implementation. The theorem below, proved in~\cite{CG07}, establishes
-%that failures refinement amounts to traces refinement and $conf$.
-%%
-%\begin{theorem}
-%  $(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)$
-%  \label{thm:fref-tref-conf}
-%\end{theorem}
-%%
-%Process equivalence in each model is given simply by mutual refinement.
 
@@
 $G(P)$ is reflected by the fact that $t$ is a function. The graph construction
-in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable
+in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable by
 sequences of edges labelled by $e_1\dots e_k$ and connecting states
@@
 
-The total function $r$ maps each node $n$ to a set of non-empty subsets of $\Sigma$. 
+The total function $r$ maps each node $n$ to a non-empty set of  (possibly empty) subsets of $\Sigma$. 
 The graph construction guarantees that $r(G(P)/s)$ represents the maximal refusals
@@
 \end{lemma}
-%\begin{lemma}
-%  \label{lemma:tgtrcref}
-%  \begin{eqnarray}
-%  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
-%  \label{eq:trcrefa}
-%  \\
-%  \label{eq:failrefa}
-%  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
-%  \\ & & \forall s\in\trc(G(Q)), R_Q\in r_Q(G(Q)/s):  \nonumber
-%  \\ & & \tabd
-%  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
-%  \\
-%  \label{eq:failrefb}
-%  P \lessdet_F Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P)) \wedge {} \nonumber
-%  \\ & & \forall s\in\trc(G(Q)), A_Q\in\minaccs(G(Q)/s): \nonumber
-%  \\ & & \tabd
-%  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
-%  \end{eqnarray}
-%  \xbox
-%\end{lemma}
-%
 
@@
 
-%, we have a result similar to
-%Theorem~\ref{thm:fref-tref-conf}, but rely on the fact that refusal sets are
-%subset closed. In other words, if a process may refuse to engage in all
-%events in a set $X$, then it can refuse to engage in all events of subsets of
-%$X$. So, in considering $conf$, it is enough to consider maximal refusals.
-%Finally, (\ref{eq:failrefb}) captures the same result using acceptances,
-%rather than refusals, which as already observed, can be calculated by taking
-%the complement of refusals. So, acceptances of $Q$ are required to have a
-%corresponding smaller set of acceptances in $P$.
-
+% =========================================================================
+\subsubsection*{Reachability Under Sets of Traces}
+\label{sec:V}
 Given a finite-state CSP process $P$ and its normalised transition graph
@@
 \end{proof}
-%\fixme{alcc: Explain the structure of the section? Can the above result be in
-%a section, rather than the introduction of the section?}
-%\fxnote{jp: I have moved this lemma now to the section on preliminaries, this should make the main section on the finite complete test suite easier to read.}
+
+% =========================================================================
+\subsubsection*{Graph Products}
+\label{sec:GP}
 
@@
 \end{proof}
-
-% =========================================================================
-
-%\subsubsection*{Tool Considerations}
-%FDR provides an API~\cite{fdrmanual} that can be used to construct normalised
-%transition graphs for CSP processes. The FDR graph nodes are labelled by
-%\emph{minimal acceptances} instead of maximal refusals as described above.
-%Since such a minimal acceptance set is the complement of a maximal refusal,
-%the function $r$ mapping states to their refusals can be implemented by
-%taking the complements of all minimal acceptances and then building all their
-%subsets. For practical applications, the subset closure is never constructed
-%in an explicit way; instead, sets are checked with respect to containment in
-%a maximal refusal.
-%
-%@todo
-
-%The maximal refusals in each process state $P/s$
-% are the complements of
-%the minimal acceptances of the node $n$
-%corresponding to $P/s$. As a consequences, all failures
-%of $P$ are represented by some $(s,R)$, where $s$ is an initialised path through the transition graph and $R\subseteq (\Sigma-A)$ for some minimal acceptance $A\in ac(n)$,
-%such that $n$ is the node corresponding to $P/s$.
-
-%% =========================================================================
-%\subsection{Test Cases and Complete Test Suites}
-%\label{sec:cspcompletedef}
-%% =========================================================================
-%
-%@todo
 
@@
 %
-This result is used in the sequel in Section~\ref{sec:finitecompletefails}.
-
-%\begin{lemma}
-%\label{lemma:hsetb}
-%
-%\end{lemma}
-%\begin{proof}
-%asd
-%\end{proof}
-%

### 2018-08-28 paper/casestudies.tex
updates
re-worked all definitions and proofs

@@
 % =======================================================================
-\section{Applications}
+\section{Testing for Failures Refinement -- an Example}
 \label{sec:case}
 % =======================================================================
-
-% =======================================================================
-\subsection{Testing for Trace Refinement}
-
-% =======================================================================
-\subsection{Testing for Failures Refinement}
 
@@
 \begin{eqnarray*}
-\text{initials} & : & N \fun \power (\Sigma)
+\text{\it initials} & : & N \fun \power (\Sigma)
 \\
-\text{minHit} & : & N \fun \power\power(\Sigma)
+\minhits & : & N \fun \power\power(\Sigma)
 \end{eqnarray*}
 %
-In a state $n = G(P)/s$, the set $\text{initials(n)}$ equals the events
-labelling outgoing edges of $n$, so  $\text{initials(n)} = [P/s]^0$. Function
-$\text{minHit}$ maps $n$ to the set of all minimal hitting sets associated
-with the minimal acceptances of $n$, so $\text{minHit}(n) =
-\text{minHit}(P/s)$. Using the transition function $t$ and the above
+In a state $n = G(P)/s$, the set $\text{\it initials}(n)$ equals the events
+labelling outgoing edges of $n$, so  $\text{\it initials}(n) = [P/s]^0$. Function
+$\minhits$ maps $n$ to the set of all minimal hitting sets associated
+with the minimal acceptances of $n$, so $\minhits(n) =
+\minhits(P/s)$. Using the transition function $t$ and the above
 functions, $U_F(p)$ can be re-written as the failures-equivalent CSP process
@@
 \\
-U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{initials}(n)) @ e \then \efail\then \Stop \big)
+U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{\it initials}(n)) @ e \then \efail\then \Stop \big)
 \label{eq:xufa}
 \\ & & \extchoice \nonumber
-\\ & & (\text{initials}(n) = \varnothing)    \&   \big( \epass \then \Stop \big)
+\\ & & (\minhits(n) = \varnothing)    \&   \big( \epass \then \Stop \big)
 \label{eq:xufb}
 \\ & & \extchoice \nonumber
-\\ & & (k < p) \& \big(\Extchoice e:\text{initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
+\\ & & (k < p) \& \big(\Extchoice e:\text{\it initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
 \label{eq:xufc}
 \\ & & \extchoice \nonumber
-\\ & & (k = p) \& \big( \sqcap_{H\in\text{minHit}(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
+\\ & & (k = p) \& \big( \sqcap_{H\in\minhits(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
 \label{eq:xufd}
@@
 produced. Another failing execution arises if $Z/s$ chooses to accept only
-$\{c \}$, while $U_F^1(4))/s$ choses to accept only $\{a,b\}$. Therefore,
+$\{c \}$, while $U_F^1(4)/s$ choses to accept only $\{a,b\}$. Therefore,
 \[

### 2018-08-28 paper/finitecomplete-failures.tex
re-worked all definitions and proofs

@@
 \\ & & \extchoice \nonumber
-\\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
+\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
 \label{eq:ufb}
@@
 \\ & & \extchoice \nonumber
-\\ & & (\#s = p) \& \big( \sqcap_{H\in\text{minHit}(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
+\\ & & (\#s = p ) \& \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
 \label{eq:ufd}
@@
 is free of livelocks, it is guaranteed that each test execution terminates
-after some $s\in\trc(P)$ with length $(p+1)$ at the latest. The test is
+after some $s\in\trc(Q)$ with length $(p+1)$ and consecutive $\epass$ or $\efail$ event at the latest. The test is
 \emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
-every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with PASS event
+every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
 $\epass$. This can also be  expressed by means of a failures refinement.
@@
 because the number of traces $s$ with maximal length $(p+1)$ is finite and the 
-sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and $\text{minHit}(P/s)$ are finite.
+sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and $\minhits(P/s)$ are finite. Further recall
+that $\minhits(P/s)$ may be empty, so that operand (\ref{eq:ufd})
+of the external choice construct  defaults to $\Stop$. In this case, however, 
+$\minaccs(P/s) = \{\varnothing\}$ applies, 
+so the test will execute branch (\ref{eq:ufb}) and
+terminate with $\epass$.
 
 Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
-a length $p$. If, after having already run through $s\in\trc(P)$ with $\#s <
-p$, an event is accepted by the SUT that is outside the initials of $P/s$,
+a length $p$. If, after having already run through $s\in\trc(P)$ with 
+$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$,
 the test immediately terminates with FAIL-event $\efail$. This is handled by
@@
 
-If $P/s$ is the $STOP$ process, this is revealed by its initials being empty.
+If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice, 
+this is revealed by $\varnothing$ being  its only minimal acceptance. 
 In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
-the external choice in $U_F(p,s)$). Note that at the same time, any (illegal)
-event of the alphabet is also accepted by the test in branch~(\ref{eq:ufa}).
-So, if the SUT accepts an event in a state where $P/s$ is supposed to have
-stopped, there exists a test execution that terminates with FAIL by choosing
-the first branch of the external choice.
+the external choice in $U_F(p,s)$).
 
@@
 After having  run successfully
-through a trace of length $p$, the test changes its behaviour:
+through a trace of length $p$, and if $\minaccs(P/s)\neq\{\varnothing\}$ so that
+als $\minhits(P/s)\neq\varnothing$,
+the test changes its behaviour:
 instead of offering {\it all} legal events from $[P/s]^0$ to the SUT,
@@
 failures refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept
-at least one event of each minimal hitting set in $\text{minHit}(P/s)$. Therefore, the test
-only terminates with success $\epass$, if such an event is accepted by the SUT.
+at least one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
+only terminates with  $\epass$, if such an event is accepted by the SUT.
 
@@
 
-A \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a reference process
+A CSP 
+\emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a reference process
 $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T, \lessdet_F\}$, and a fault domain 
@@
 \begin{proof}
+{\bf Step~1.}
 Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\ Q$ 
@@
 
+{\bf Step~2.}
 Moreover,  Lemma~\ref{lemma:hseta}
-implies that for all traces $s\in\trc(Q)\cap\trc(P)$, every $H$ in $\text{minHit}(P/s)$ is
+implies that for all traces $s\in\trc(Q)\cap\trc(P)$, every $H$ in $\minhits(P/s)$ is
 a hitting set for $\minaccs(Q/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,s)$ 
@@
 (\ref{eq:ufc}) to test continuation without a verdict. For the last branch,
-we note that any selected minimal hitting set $H\in\text{minHit}(P/s)$ has a
+we note that any selected minimal hitting set $H\in\minhits(P/s)$ has a
 non-empty intersection with each of the minimal acceptances of $Q/s$. As a
@@
 terminates with PASS event $\epass$. Note that this argument requires that
-$Q$ is free of livelocks, because otherwise the PASS-events might not become
+$Q$ is free of livelocks, because otherwise the $\epass$-events might not become
 visible, due to unbounded sequences of hidden events performed by $Q$.
-%Note
-%further, that this proof did not refer in any way to the size of $Q$'s
-%normalised transition graph, so the test suite is sound for {\it all}
-%non-divergent CSP process refining $P$.
-%\fixme{alcc: Soundness does not
-%depend on the fault domain. I would have it as a separate theorem.}
-%\fxnote{jp: this is a good idea, I have split the proof of the main theorem in 2 lemmas.}
 \xbox
@@
 \begin{lemma}\label{lemma:mainfexhaustive}
-Test suite $\TS_F$ specified im Theorem~\ref{th:failurestest} is exhaustive for the fault
+Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive for the fault
 model specified there.
@@
 execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
-FAIL event $\efail$ or without giving any verdict. The latter case is also
+   $\efail$-event or without giving any verdict. The latter case is also
 interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
@@
 assumption of non-divergence of Q is needed for this conclusion. %\fixme{alcc:
-%I didn't see how Lemma 3 is being applied here. Also, given the definition of
-%the test as a process, I'm not sure why you need to make the argument using
-%graph product.}
-%\fxnote{jp: I do hope that this is now clearer from the new structure of the lemmas}
 
@@
 only:~assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has
-at least one non-empty minimal acceptance, so the guard condition $([P/s]^0 =
-\varnothing)$ of branch (\ref{eq:ufb}) evaluates to $\isf$ for $U_F(\#u,u)$.
+at least one non-empty minimal acceptance, so the guard condition 
+$(\minaccs(P/s) = \{ \varnothing \})$ 
+of branch (\ref{eq:ufb}) evaluates to $\isf$ for $U_F(\#u,u)$.
 Moreover, the guard condition $(\#s < p)$ for branch (\ref{eq:ufc}) evaluates
@@
 test always fails. If branch (\ref{eq:ufd}) is executed, the test fails for
-the execution where a minimal hitting set $H\in\text{minHit}(P/u)$ is chosen
+the execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen
 by $U_F(\#u,u)$ that has an empty intersection with the minimal acceptance
@@
 $U_F(\#u,u)$ selects $H$. This execution deadlocks in process state
-$(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the pass event
-$\epass$; this  means that the test fails. This concludes the proof. \xbox
+$(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event; 
+this  means that the test fails and concludes the proof. \xbox
 \end{proof}

### 2018-08-28 paper/finitecomplete-trace.tex
small changes to Section 5, and rewording to avoid reference to refinement of states
re-worked all definitions and proofs

@@
 For establishing trace refinement, the following class of adaptive test cases
-will be used; again, they are defined for integers $p \ge 0$.
+will be used for a given reference process $P$ and  integers $p \ge 0$.
 %
@@
 \\
-U_T(p,s) & = & \big( e:(\Sigma - [P/s]^0) \then \efail\then \Stop \big)
+U_T(p,s) & = & \big(\Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
 \label{eq:uta}
 \\ & & \extchoice \nonumber
-\\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
+\\ & & (\minaccs(P/s) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
 \label{eq:utb}
 \\ & & \extchoice \nonumber
-\\ & & (\#s < p) \& \big( e:[P/s]^0 \then U_T(p,s.e) \big)
+\\ & & (\#s < p) \& \big(\Extchoice e:[P/s]^0 @ e \then U_T(p,s.e) \big)
 \label{eq:utc}
@@
 \end{theorem}
-%
-We skip the proof of Theorem~\ref{th:tracetest}, since it is just a
-simplified version of the proof of Theorem~\ref{th:failurestest}.
+\begin{proof}
+The theorem follows directly  
+from Step~1 in the proof of Lemma~\ref{lemma:mainfsound} and
+Case~1 in the proof of Lemma~\ref{lemma:mainfexhaustive}.
+\xbox
+\end{proof}

### 2018-08-28 paper/preliminaries.tex
re-worked all definitions and proofs

@@
 external and internal choice operators allow us to define a choice over a
-collection of processes $P(i)$.
+collection of processes $P(i)$. By convention, these choices evaluate to the process $\Stop$ 
+which does not engage into any event, if the index set $I$ is empty.
 
@@
 $s$. The failures model of a process $P$ records all its failures in a set
-$\fails(P)$. 
+$\fails(P)$. The semantic rules specify for each operator, how the failures of 
+the resulting process can be calculated from the failures of each operand. For example, 
+$\fails(P\intchoice Q) = \fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for the comprehensive list of these rules.
+
 Using notation $P/s$ to denote the process $P$ after having engaged into trace $s$,
@@
 %
-The cardinality of $\maxrefs(P/s)$ reflects the degree of
-nondeterminism that is present in process state $P/s$: the more maximal refusal sets
-contained in  $\maxrefs(P/s)$, the more nondeterministic is the behaviour in
-state $P/s$. Deterministic process states $P/s$ have exactly the one maximal refusal
+Deterministic process states $P/s$ have exactly the one maximal refusal
 $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of $P/s$, i.e.,   the events that $P/s$ may engage into.
+The maximal refusals in combination with the initials of a process express its degree of 
+nondeterminism.
+\begin{example}
+\label{ex:nondetdegree}
+Process $P = (\Stop \intchoice Q)$ has maximal refusals 
+$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ 
+refuses to engage into any event, and this carries over to $P$, when applying the internal 
+choice operator. However, $P$ is distinguished from $\Stop$ by its initials: 
+$[P]^0 = [ \Stop \intchoice Q]^0 = [Q]^0$, so $P$ may engage nondeterministically in
+any initial event of $Q$, but can also refuse everything, due to internal selection of the $\Stop$ process.
+
+Assuming that $\Sigma = \{a,b,c,d\}$, process
+$Q = (e:\{a,b\} \then \Stop) \intchoice (e:\{c,d\} \then \Stop)$ has maximal refusals
+$\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials $[Q]^0=\Sigma$, so, in contrast to
+$P$, nondeterminism is reflected
+here by the presence of two maximal refusals.
+\xbox
+\end{example}
 
@@
 \emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
-refusing the events in $\Sigma-A$. The acceptances of $P/s$ are denoted by
+{\it possibly} refusing the events from $[P/s]^0 - A$ and 
+{\it always} refusing those from $\Sigma - [P/s]^0$.  
+The acceptances of $P/s$ are denoted by
 $\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
@@
 A \emph{minimal hitting set} is a hitting set that cannot be further reduced
-without losing the characteristic property (\ref{eq:hit}). The problem of
-determining minimal hitting sets is known to be
-NP-hard~\cite{Book1975-BOOKRM}, but we will see below that it reduces the
-effort of testing for failures refinement from a factor of $2^\Sigma$ to a
+without losing the characteristic property (\ref{eq:hit}).
+By $\minhits(C)$ we denote the collection of minimal hitting sets for $C$.
+For the pathological case where $C$ contains an empty set, $\minhits(C)$ is also empty.
+ The problem of
+determining minimal hitting sets is known to be %%% \cite{Book1975-BOOKRM,5533149}
+NP-hard~\cite{5533149}, but we will see below that it reduces the
+effort of testing for failures refinement from a factor of $2^{|\Sigma|}$ to a
 factor that equals the number of minimal hitting sets.
 
-For testing, the following lemma about hitting sets are required.
-
+The following lemma establishes the fact that the $conf$ relation specified in (\ref{eq:conf}) can be chracterised by means of minimal acceptances and their
+minimal hitting sets.
 \begin{lemma}
@@
 For each $s\in\trc(P)$,
-let $\text{minHit}(P/s)$ denote the
+let $\minhits(P/s)$ denote the
 collection of all minimal hitting sets of $\minaccs(P/s)$.
@@
 
-\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \text{minHit}(P/s)$, $H$ is
+\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \minhits(P/s)$, $H$ is
 a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
@@
 \]
-Therefore, $H \in  \text{minHit}(P/s)$ not only implies $H\cap
+Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap
 A_P\neq\varnothing$ for all minimal acceptances $A_P$, but also $H\cap
@@
 A_Q$ for at least one $A_P$. As a consequence, each $H \in
-\text{minHit}(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
+\minhits(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
 
@@
 Minimising $\overline H$ yields   a minimal hitting set $H\in
-\text{minHit}(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
+\minhits(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
 contradiction to Assumption~2. This completes the proof of the lemma. \xbox
@@
 %
+
+It is important to point out that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
+Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and this carries over to $Q\intchoice \Stop$.
+From (\ref{eq:failrefb}) we conclude that
+$\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{ \varnothing \}$. These considerations
+clarify that $\minhits(P/s)$ is empty if and only if  $\minaccs(P) = \{ \varnothing \}$.
+We note that the proof of Lemma~\ref{lemma:hseta} also covers the situations where
+$\minaccs(P/s) = \{ \varnothing \}$ and therefore  $\minhits(P/s) = \varnothing$.

### 2018-08-28 paper/finitecomplete-failures.tex
Added complexity considerations

@@
 % ==========================================================================
+\subsection{Complexity Considerations}
+\label{sec:complexity}
+
+As can be seen from the specification of the test cases $U_F(k)$, the number of 
+executions ending in a $\epass$-event corresponds to the number $\ell$ of traces $s$ 
+of $P$ with length less or equal $p$, multiplied by the number $h$ of minimal hitting sets in
+$\minhits(P/s)$. The first factor 
+has worst-case upper bound $\ell\le |\Sigma|^p$.
+Given a set $\minaccs(P/s) = \{ A_1,\dots, A_a \}$ of   minimal acceptances,
+the cardinality $h$ of $\minhits(P/s)$ is maximal for the case where all $A_i$ are disjoint.
+In this case,  $h=|\minhits(P/s)| = \prod_{i=1}^a|A_i|$. This yields a worst-case upper bound $h\le 2^{\frac{|\Sigma|}{2}}$. This is better than the upper bound $(2^{|\Sigma|}-1)$
+resulting from the tests suggested in~\cite{Hennessy:1988:ATP:50497}, where {\it every}
+non-empty 
+subset of $\Sigma$ whose events cannot be completely refused in a given process state
+has to be checked during test executions. This improvement becomes more significant for process states where the number of minimal acceptances is small. In the best case, where
+ $\minaccs(P/s) = \{ A_1  \}$, we just have $|A_1|$ minimal hitting sets (each consisting of a single element) to test, while the tests from~\cite{Hennessy:1988:ATP:50497} 
+ require $(2^{|A_1|}-1)$ sets to be checked. From Lemma~\ref{lemma:hseta} we know that 
+ the worst-case bound $h\le 2^{\frac{|\Sigma|}{2}}$ cannot be further improved, since 
+the full collection of minimal hitting sets needs to be checked in order to verify
+the $conf$ relation.
+
+Summarising, the number of executions involved in one adaptive test $U_F(k)$ is bounded
+by $|\Sigma|^k\cdot 2^{\frac{|\Sigma|}{2}}$. According to Theorem~\ref{th:failurestest},
+we need to execute these test for $k = 0,\dots,pq$; this results in a worst-case bound
+of (we use the formula for the sum of the geometric progression here)
+\[
+    2^{\frac{|\Sigma|}{2}}\cdot\big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),
+    \quad\text{or, asymptotically,}\quad O(2^{(\frac{|\Sigma|}{2} + pq)}).
+\]
+
+It is discussed in Section~\ref{sec:conc} that the term $pq$ occurring in this bound can be significantly reduced by using more complex test cases that need not execute {\it all} 
+traces of $P$ with length less or equal to $pq$.
+
+% ==========================================================================

### 2018-08-28 paper/finitecomplete-failures.tex
Complexity considerations improved

--- 2018-08-28 paper/finitecomplete-failures.tex
@@
 % ==========================================================================
+\subsection{Complexity Considerations}
+\label{sec:complexity}
+
+As can be seen from the specification of the test cases $U_F(k)$, the number of 
+executions ending in a $\epass$-event corresponds to the number $\ell$ of traces $s$ 
+of $P$ with length equal to\footnote{In this estimation, we disregard 
+the case where the test
+terminates earlier due to entering branch (\ref{eq:ufb}).} $k$, 
+multiplied by the number $h$ of minimal hitting sets in
+$\minhits(P/s)$. The first factor 
+has worst-case upper bound $\ell\le |\Sigma|^k$.
+
+Given a set $\minaccs(P/s) = \{ A_1,\dots, A_a \}$ of   minimal acceptances,
+the cardinality $h$ of $\minhits(P/s)$ is maximal for the case where all $A_i$ are disjoint.
+In this case,  $h=|\minhits(P/s)| = \prod_{i=1}^a|A_i|$. To find an upper bound for $h$, 
+we need to solve the constrained optimisation problem
+\[
+\text{maximise}\ \prod_{i=1}^a|A_i| \ \text{under constraint}\ \sum_{i=1}^a|A_i| = |\Sigma|
+\wedge \bigwedge_{i=1}^a |A_i| > 0
+\]
+For fixed $a$, this can be solved using Lagrange multipliers, but there is no 
+closed expression for the maximum value to be achieved when the $|A_i|$ 
+and $a$ are considered as 
+unknown variables. Therefore, we only note that $h=|\minhits(P/s)|$ is significantly smaller
+than $2^{|\Sigma|}$, because for every $H\in  \minhits(P/s)$, all of its true subsets and true supersets are {\it not} contained in $\minhits(P/s)$. This leaves us with an estimate
+for the worst-case upper bound of $h\cdot |\Sigma|^k$ with $h\ll 2^{|\Sigma|}$.
+According to Theorem~\ref{th:failurestest},
+we need to execute these test for $k = 0,\dots,(pq-1)$; this results in a worst-case bound
+of (we use the formula for the sum of the geometric progression here)
+\[
+h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
+ \text{or, asymptotically,}\  O(h\cdot2^{(pq-1)}), 
+ \  \text{with}\ h\ll 2^{|\Sigma|}.
+\]
+
+From Lemma~\ref{lemma:hseta} we know that 
+ the worst-case bound $h\ll 2^{|\Sigma|}$ cannot be further improved, since 
+the full collection of minimal hitting sets needs to be checked in order to verify
+the $conf$ relation.
+
+The authors of~\cite{Hennessy:1988:ATP:50497} and \cite{DBLP:conf/icfem/CavalcantiG07} 
+suggest to test {\it every}
+non-empty 
+subset of $\Sigma$ whose events cannot be completely refused in a given process state
+of the reference model; this leads to a worst-case estimate of $2^{|\Sigma|}$
+for the number of different sets to be offered to the SUT in the last step of 
+the test execution, so our approach reduces the number of test executions in comparison
+ to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07}.
+
+It is discussed in Section~\ref{sec:conc} that the term $pq$ occurring in this bound can be significantly reduced by using more complex test cases that need not execute {\it all} 
+traces of $P$ with length less or equal to $pq$.
+
+% ==========================================================================

### 2018-08-28 paper/introduction.tex
Complexity considerations improved

@@
 established for process algebras, and for CSP in particular, by several
-authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
+authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
 To the best of our knowledge, however, this article is the first to present

### 2018-08-28 paper/preliminaries.tex
Complexity considerations improved

@@
 expressed by means of their normalised transition graphs 
-$$G(P) = ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))$$ 
-and 
-$$G(Q) = ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))$$
+\begin{eqnarray*}
+G(P) & = & ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))
+\\
+G(Q) & = & ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))
+\end{eqnarray*}
+%
 in the following

### 2018-08-29 paper/conclusion.tex
Improvements according to Ana's questions and suggestions.
updates
Discussion of adaptive test cases added

@@
-\section{Conclusion}
+\section{Discussion}
 \label{sec:conc}
 % ==============================================================================
+\subsubsection*{Discussion of Fault Domains}
 
 % ==============================================================================
+\subsubsection*{Discussion of Adaptive Test Cases}
+The tests suggested 
+in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were \emph{preset} in the
+sense that the trace to be executed was pre-defined for each test. As a consequence,
+the authors of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive}
+as a third test result, applicable to the situations where the intended trace
+of the execution was blocked, due to legal, but nondeterministic behaviour of the
+SUT. In contrast to that, 
+our test cases
+specified in Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
+adaptive. This has the advantage that test executions $(Q\parallel[\Sigma] U_F(p))$
+for failures refinement
+will never block before the final step specified by branch (\ref{eq:ufd}). As
+a consequence, we do not need inconclusive test results. It should be noted, however,
+that it is necessary for our test verdicts to recognise also deadlocks in the final
+test step and interpret 
+them as FAIL, as described in Section~\ref{sec:finitecompletefails}. In practice,
+this is realised by adding a timeout event to the testing environment which indicates
+deadlock situations. For real-time systems, this is an accepted technique, because
+the SUT has to respond within a pre-defined latency interval, otherwise its behaviour 
+is considered to be blocked and regarded as a failure. Our tests executions 
+$(Q\parallel[\Sigma] U_T(p))$ for trace refinement never block at all. The adaptive       
+behaviour of our test cases, however, induce the obligation to check that {\it all}
+possible executions have been performed before the test can be considered as passed.
+Typically, it is therefore assumed that a 
+\emph{complete testing assumption}~\cite{hierons_testing_2004} holds which means that
+every possible behaviour of the SUT has been performed after a finite number of
+test executions. 
+In practice, this is realised by executing each test several times, 
+recording the traces that have been performed, and using hardware or software coverage 
+analysers to determine whether all possible test execution behaviours of the SUT
+have been observed. Therefore, the application of adaptive test cases comes at the price
+of having to apply some grey-box testing techniques enabling us to decide whether
+all SUT behaviours have been observed.
+
+% ==============================================================================
+\subsubsection*{Implications for CSP Model Checking}
+
+% ==============================================================================
+\subsubsection*{Conclusion}
+
+% ==============================================================================
+
+\paragraph{Acknowledgements}
+The authors would lime to thank Bill Roscoe and Thomas Gibson-Robinson for their advice on using the FDR4 model checker and for very helpful discussions concerning the potential implications of this paper in the field of model-checking.

### 2018-08-29 paper/conclusion.tex
Completed the discussion - only the discussion part about fault domains is still missing.

@@
 \subsubsection*{Discussion of Fault Domains}
+
+@todo
+
+% ==============================================================================
+\subsubsection*{Discussion of Further Reductions of the Test Effort}
+It is known from complete testing strategies for finite state machines that 
+the upper bound $pq$ for the lengths of the 
+traces used in our tests to investigate the SUT behaviour can be reduced. It is also
+known from FSM testing that it is not necessary to test {\it all} traces up to this maximal length. Notable complete 
+strategies supporting this fact have been presented, for example,
+ in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}. From~\cite{Huang2017} it
+ is known that complete 
+ FSM testing theories can be translated to other formalisms, such as 
+Extended Finite State Machines, Kripke Structures, or CSP, resulting in likewise complete
+test strategies for the latter. We intend to study translations of several promising FSM
+strategies to CSP in the future. These will effectively reduce the upper bound 
+$\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The bound $h$ for
+the number of sets to be used in probing the SUT for illegal deadlocks, however, cannot be further reduced, as has been established in Lemma~\ref{lemma:hseta}.
 
@@
 \subsubsection*{Implications for CSP Model Checking}
+As explained in the previous sections, passing a test is specified by
+$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$
+for failures testing. If the SUT $Q$ is not a programmed piece of software or
+an integrated HW/SW system, but just another CSP process specification,
+it is of course possible to verify these pass criteria using the FDR4 model checker.
+For checking the refinement relation $P\lessdet_F Q$, the pass criteria have to be verified
+for $k=0,\dots,pq-1$, where $p$ and $q$ indicate the number of nodes in $P$'s transition
+graph and the maximal number of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since the test cases
+$U_F(k)$ have such a simple structure, it is an interesting question for further 
+research whether checking 
+$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster than directly 
+checking $P\lessdet_F Q$, as one would do in the usual 
+approach with FDR4. This is of particular interest, since the checks could be parallelised 
+on several CPUs. Alternatively it is interesting to investigate whether the
+the check\footnote{We are grateful to Bill Roscoe for suggesting this option.}
+\[
+(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{k=0}^{pq-1} U_F(k)) \hide \Sigma
+\]
+may perform better than the check of $P\lessdet_F Q$, since the former allows for 
+other optimisations in the model checker than the latter.
+
+For large and complex CSP processes $Q$, it may be too time consuming to generate
+its normalised transition graph, so that its number $q$ of nodes is unknown. 
+In such a case the suggested options may still be used
+as efficient bug finders: use  $p$ of reference process $P$ as initial $q$-value
+and 
+increment $q$ from there, as long as each increment reveals new errors.
 
@@
 
+In this paper, we have introduced finite complete testing strategies for 
+model-based testing against CSP reference models. The strategies are applicable   
+to the conformance relations failures refinement and trace refinement. The underlying fault domains have been defined as the sets of CSP processes whose normalised transition
+graphs did not have more than a given number of additional nodes, when compared to the transition graph of the reference process. For these domains, finite complete 
+test suites are available. We have shown for the strategy checking failures refinement,
+ that the way of probing the SUT for illegal deadlocks used in our test cases is 
+ optimal in the sense that it is not possible to guarantee exhaustiveness with fewer
+ probes.
+
 % ==============================================================================
@@
 \paragraph{Acknowledgements}
-The authors would lime to thank Bill Roscoe and Thomas Gibson-Robinson for their advice on using the FDR4 model checker and for very helpful discussions concerning the potential implications of this paper in the field of model-checking.
+The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for their advice on using the FDR4 model checker and for very helpful discussions concerning the potential implications of this paper in the field of model-checking.

### 2018-08-29 paper/finitecomplete-failures.tex
Completed the discussion - only the discussion part about fault domains is still missing.

@@
 $\epass$. This can also be  expressed by means of a failures refinement.
-\[
+%
+\begin{equation}
+\label{eq:passF}
 Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
-\]
+\end{equation}
+%
 This type of pass relation is often called \emph{must test}, because every

### 2018-08-29 paper/finitecomplete-failures.tex
Complexity considerations completed.

@@
 
-Given a set $\minaccs(P/s) = \{ A_1,\dots, A_a \}$ of   minimal acceptances,
-the cardinality $h$ of $\minhits(P/s)$ is maximal for the case where all $A_i$ are disjoint.
-In this case,  $h=|\minhits(P/s)| = \prod_{i=1}^a|A_i|$. To find an upper bound for $h$, 
-we need to solve the constrained optimisation problem
-\[
-\text{maximise}\ \prod_{i=1}^a|A_i| \ \text{under constraint}\ \sum_{i=1}^a|A_i| = |\Sigma|
-\wedge \bigwedge_{i=1}^a |A_i| > 0
-\]
-For fixed $a$, this can be solved using Lagrange multipliers, but there is no 
-closed expression for the maximum value to be achieved when the $|A_i|$ 
-and $a$ are considered as 
-unknown variables. Therefore, we only note that $h=|\minhits(P/s)|$ is significantly smaller
-than $2^{|\Sigma|}$, because for every $H\in  \minhits(P/s)$, all of its true subsets and true supersets are {\it not} contained in $\minhits(P/s)$. This leaves us with an estimate
-for the worst-case upper bound of $h\cdot |\Sigma|^k$ with $h\ll 2^{|\Sigma|}$.
+Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal acceptances,
+the cardinality $h$ of $\minhits(P/s)$ is maximal for the case where all $A_i$ are disjoint. In this case,  $h=|\minhits(P/s)| = \prod_{i=1}^\alpha|A_i|$. 
+To find an upper bound for $h$, we prove the following lemma.
+
+\begin{lemma}\label{lemma:minhitmax}
+Let $n\ge 2$ be any positive integer.
+We call a sequence $a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$, a partion of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$.
+Define
+$$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
+Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
+\alpha\in \mathbf{N}\big\}$$
+\end{lemma}
+\begin{proof}
+Let $n\ge 2$.
+Obviously, 
+\begin{equation}\label{eq:1} 
+   n\le p(n).
+\end{equation}
+We show that
+$p(n)$ is associated with a unique partition  $b_1\le b_2\le \ldots\le  b_{\beta}$ of $n$ with $b_i\in \{2,3\}, i=1,\dots,\beta$. We denote this partition  by $\pi(n)$. 
+To prove its existence, 
+let $a_1, \ldots , a_{\alpha}$ be an {\it arbitrary} partition of $n$. 
+Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then from~(\ref{eq:1}) we have 
+$$
+\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}$$
+Hence, if we replace $a_j$ by $p(a_j)$, the new partition has a product which is
+ not less than the original partition $a_1, \ldots , a_{\alpha}$, but the value $a_j$
+ is now represented by smaller numbers. This process is repeated until there is a partition of $n$ with maximal product such that every term is $2$ or $3$.
+Since $6=2+2+2=3\cdot 3$ and $2^3< 3^2$, the number of $2$s in such maximal partition is less than $2$. This shows that $\pi(n)$ is uniquely determined.
+
+Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\}, i=1,\dots,\alpha$, be a partition of $n$ with maximal product.
+Let $k_1\le 2$ be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
+Then $n=2k_1+3k_2$ and  $n\equiv 2k_1 (\mod 3)$ with $k_1\le 2$. Then
+
+\begin{align*}
+k_1&=
+\begin{cases} 
+0 &\text{if}\,\, \exists k: n=3k\\ 
+2 &\text{if}\,\, \exists k: n=3k+1\\
+1 &\text{if}\,\, \exists k: n=3k+2
+\end{cases}\\
+ k_2&=
+ \begin{cases} 
+ k &\text{if}\,\, n=3k\\ 
+ k-1 &\text{if}\,\, n=3k+1\\
+ k &\text{if}\,\, n=3k+2\end{cases}
+\end{align*}
+% 
+ and  
+ $$
+ \prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}=
+ \begin{cases} 
+ 3^k &\text{if}\,\,  n=3k\\ 
+ 2^2\cdot 3^{k-1} &\text{if}\,\,   n=3k+1\\
+ 2\cdot 3^{k} &\text{if}\,\,  n=3k+2\end{cases}
+ $$
+ \xbox
+ \end{proof}
+%
+From Lemma~\ref{lemma:minhitmax}, we conclude that 
+$$h=\begin{cases} 
+3^{|\frac{\Sigma}{3}|} &\text{if}\,\, \exists k: |\Sigma|=3k\\ 
+2^2\cdot 3^{\frac{|\Sigma|-4}{3}} &\text{if}\,\, \exists k : |\Sigma| =3k+1\\
+2\cdot 3^{\frac{|\Sigma| - 2}{3}} &\text{if}\,\, \exists k: |\Sigma| =3k+2
+\end{cases}
+$$
+
+We not that $h$ is significantly smaller than $2^{|\Sigma|}$ -- this can also be intuitively motivated by the fact that  for every $H\in  \minhits(P/s)$, all of its true subsets and true supersets are {\it not} contained in $\minhits(P/s)$.
+
 According to Theorem~\ref{th:failurestest},
-we need to execute these test for $k = 0,\dots,(pq-1)$; this results in a worst-case bound
+we need to execute the tests $U_F(k)$
+for $k = 0,\dots,(pq-1)$; this results in a worst-case bound
 of (we use the formula for the sum of the geometric progression here)
@@
 h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
- \text{or, asymptotically,}\  O(h\cdot2^{(pq-1)}), 
- \  \text{with}\ h\ll 2^{|\Sigma|}.
+ \text{or, asymptotically,}\  O(3^{|\frac{\Sigma}{3}|}\cdot|\Sigma|^{(pq-1)}).
  \]
@@
 From Lemma~\ref{lemma:hseta} we know that 
- the worst-case bound $h\ll 2^{|\Sigma|}$ cannot be further improved, since 
+ the worst-case bound for $h$   cannot be further improved, since 
 the full collection of minimal hitting sets needs to be checked in order to verify
@@
 subset of $\Sigma$ whose events cannot be completely refused in a given process state
-of the reference model; this leads to a worst-case estimate of $2^{|\Sigma|}$
+of the reference model; this leads to a worst-case estimate of $2^{|\Sigma|}-1$
 for the number of different sets to be offered to the SUT in the last step of 

### 2018-08-30 paper/finitecomplete-failures.tex
Proof of lemma on partitions improved.

@@
 Let $n\ge 2$ be any positive integer.
-We call a sequence $a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$, a partion of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$.
+We call a sequence $0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$, a partion of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$.
 Define
@@
 Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
-\alpha\in \mathbf{N}\big\}$$
+\alpha\in \mathbb{N}\big\}.$$
 \end{lemma}
@@
 Let $n\ge 2$.
-Obviously, 
-\begin{equation}\label{eq:1} 
-   n\le p(n).
-\end{equation}
-We show that
-$p(n)$ is associated with a unique partition  $b_1\le b_2\le \ldots\le  b_{\beta}$ of $n$ with $b_i\in \{2,3\}, i=1,\dots,\beta$. We denote this partition  by $\pi(n)$. 
-To prove its existence, 
-let $a_1, \ldots , a_{\alpha}$ be an {\it arbitrary} partition of $n$. 
+Obviously, \begin{equation}\label{eq:1} n\le p(n).\end{equation}
+Define $\pi(n)$ to be the following partition  of $n$: 
+$$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
+Then the product of the numbers in 
+$\pi(n)$ equals  $p(n)$. In the remainder of the proof
+we have to show that $p(n)$ is really maximal in the sense of the lemma.
+
+Since the number of partitions of a given $n$ is finite, $\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$ exists.
+
+Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with maximal product. Then,
+obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
 Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then from~(\ref{eq:1}) we have 
-$$
-\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}$$
-Hence, if we replace $a_j$ by $p(a_j)$, the new partition has a product which is
- not less than the original partition $a_1, \ldots , a_{\alpha}$, but the value $a_j$
- is now represented by smaller numbers. This process is repeated until there is a partition of $n$ with maximal product such that every term is $2$ or $3$.
-Since $6=2+2+2=3\cdot 3$ and $2^3< 3^2$, the number of $2$s in such maximal partition is less than $2$. This shows that $\pi(n)$ is uniquely determined.
-
+$$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
+Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a product value
+which is not less then the original partition $a_1, \ldots , a_{\alpha}$. This process
+can be repeated, until every $a_i$ in the resulting partition is either 2 or 3.
+Hence there exists a partition of $n$ with maximal product, such that every term is $2$ or $3$.
+Since $6=2+2+2=3\cdot 3$ and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than $3$. 
 Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\}, i=1,\dots,\alpha$, be a partition of $n$ with maximal product.
 Let $k_1\le 2$ be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
-Then $n=2k_1+3k_2$ and  $n\equiv 2k_1 (\mod 3)$ with $k_1\le 2$. Then
+Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$ with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have
 
-\begin{align*}
-k_1&=
-\begin{cases} 
-0 &\text{if}\,\, \exists k: n=3k\\ 
-2 &\text{if}\,\, \exists k: n=3k+1\\
-1 &\text{if}\,\, \exists k: n=3k+2
-\end{cases}\\
- k_2&=
- \begin{cases} 
- k &\text{if}\,\, n=3k\\ 
- k-1 &\text{if}\,\, n=3k+1\\
- k &\text{if}\,\, n=3k+2\end{cases}
-\end{align*}
-% 
- and  
- $$
- \prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}=
- \begin{cases} 
- 3^k &\text{if}\,\,  n=3k\\ 
- 2^2\cdot 3^{k-1} &\text{if}\,\,   n=3k+1\\
- 2\cdot 3^{k} &\text{if}\,\,  n=3k+2\end{cases}
- $$
+\begin{align*}k_1&=\begin{cases} 0 &\text{if}\,\, n=3k\\ 2 &\text{if}\,\, n=3k+1\\1 &\text{if}\,\, n=3k+2\end{cases}\\
+ k_2&=\begin{cases} k &\text{if}\,\, n=3k\\ k-1 &\text{if}\,\, n=3k+1\\k &\text{if}\,\, n=3k+2\end{cases}\\
+\prod_{i=1}^{\alpha}a_i&=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}\end{align*}
+This proves the lemma.
 \xbox

### 2018-09-02 paper/finitecomplete-failures.tex
small suggestions

@@
 %
-A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$,
-synchronising over alphabet $\Sigma$. Therefore, a test execution is any
-trace of the concurrent process
+A test is performed by running $U_F(p)$ concurrently with any SUT process
+$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
+the concurrent process
 \[

### 2018-09-02 paper/introduction.tex
small suggestions

@@
 
-In the safety-critical domains, test suites with guaranteed fault coverage
-are of particular interest. For black-box testing, guarantees can be given
-only if certain hypotheses are satisfied. These hypotheses are usually
-specified by a \emph{fault domain}:~a set of models that may or may not
-conform to a given reference model. 
-The so-called \emph{complete} test strategies guarantee
-to accept every  system under test (SUT) conforming to the reference model, while  uncovering every conformance violation, provided that the true SUT behaviour is captured by a member of the
-fault domain.
+In the safety-critical domain, test suites with guaranteed fault coverage are
+of particular interest. For black-box testing, guarantees can be given only
+if certain hypotheses are satisfied. These hypotheses are usually specified
+by a \emph{fault domain}:~a set of models that may or may not conform to a
+given reference model. \emph{Complete} test strategies guarantee to accept
+every  system under test (SUT) conforming to the reference model, and uncover
+every conformance violation, provided that the SUT behaviour is captured by a
+member of the fault domain.
 
 Generation methods for complete test suites have been developed for various
-modelling formalisms. In this paper, we use \emph{Communicating Sequential
-Processes (CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
-process-algebraic approach that has been shown to be well-suited for the
-description of reactive control systems in many publications over almost five
-decades. Many of these applications have
-been described in~\cite{Roscoe2010} and in the references there.
-Industrial success has also been reported, see, for example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.
+formalisms. In this paper, we use \emph{Communicating Sequential Processes
+(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature process
+algebra that has been shown to be well-suited for the description of reactive
+control systems in many publications over almost five decades. Many of these
+applications are described in~\cite{Roscoe2010} and in the references there.
+Industrial success has also been reported, see, for example,
+\cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.
 
@@
 This paper presents complete black-box test suites for software and systems
-modelled using CSP.
-They can be generated for divergence-free\footnote{The assumption of divergence freedom is usually
-applied in black-box testing, since it cannot distinguish between divergence
-and deadlock.} finite-state CSP processes with finite alphabets,   
-interpreted both in the trace and the failures semantics.
-Our results complement work published in~\cite{DBLP:conf/pts/CavalcantiS17}.
-There, fault domains are specified as collections of processes refining a
-``most general'' fault domain member. With that concept, complete test suites
-may be finite or infinite. This gives important insight into the theory of
-fault-domain testing for CSP, but we are particularly interested in {\it
-finite} suites when it comes to practical application.
-While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
-select tests from still infinite test suites, here, we further restrict fault
-domains using a graph representation of processes (originally elaborated
-for the purpose of model checking) to
-obtain test suites that are finite.
+modelled using CSP. They can be generated for divergence-free\footnote{The
+assumption of divergence freedom is usually applied in black-box testing,
+since it cannot distinguish between divergence and deadlock.} finite-state
+CSP processes with finite alphabets, interpreted both in the trace and the
+failures semantics. Our results complement work
+in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
+collections of processes refining a ``most general'' fault domain member.
+With that concept, complete test suites may be finite or infinite. This gives
+important insight into the theory of fault-domain testing for CSP, but we are
+particularly interested in {\it finite} suites when it comes to practical
+application. While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional
+criteria to select tests from still infinite test suites, here, we further
+restrict fault domains using a graph representation of processes (originally
+elaborated for model checking) to obtain finite test suites. \fixme{alcc: I
+think termination is not handled either?}
 
@@
 
-The main contribution of this paper consists in the construction of two test suites
-verifying the conformance relations \emph{trace refinement} and  \emph{failures refinement}, respectively, for any given non-divergent, finite-state, finite alphabet 
-CSP reference process. We prove their completeness with respect to fault domains
-of the described type.  The
-existence of -- possibly infinite -- complete test suites has been
-established for process algebras, and for CSP in particular, by several
+The main contribution of this paper is the construction of two test suites to
+verify the conformance relations \emph{trace refinement} and \emph{failures
+refinement} for any given non-divergent, finite-state, finite alphabet CSP
+process. We prove their completeness with respect to fault domains
+of the described type.  The existence of -- possibly infinite -- complete
+test suites has been established for process algebras, and for CSP in
+particular, by several
 authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.

### 2018-09-02 paper/preliminaries.tex
small suggestions

@@
 \begin{example}\label{example:CSP}
-  We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
+  We consider $P$, $Q$, and $R$ below. $P$ is initially
   ready to engage in the event $a$, and then makes an internal  choice to
@@
 external and internal choice operators allow us to define a choice over a
-collection of processes $P(i)$. By convention, these choices evaluate to the process $\Stop$ 
-which does not engage into any event, if the index set $I$ is empty.
+collection of processes $P(i)$. By convention, external choices evaluate to
+the process $\Stop$, which deadlocks:~does not engage into any event and does
+not terminate, if the index set $I$ is empty. An iterated internal choice is
+not defined for an empty index set.
 
@@
 those of $P$ and $Q$ in parallel, synchronising on the events in the set
-$cs$. Other forms of parallelism can be defined using this operator.
+$cs$. Other forms of parallelism available in CSP can be defined using this
+operator.
 
 Interactions that are not supposed to be visible to the environment can be
-hidden. The operator $P \hide H$ defines a process that behaves as $P$, with the
-interactions modelled by events in the set $H$ hidden. Frequently, hiding is used
-in conjunction with parallelism: it may be desirable to make interal actions of each process in a network invisible, while events happening at their interfaces remain observable.
-
-%Frequently, hiding is
-%used in conjunction with parallelism, so that communications in a network of
-%parallel processes used just to coordinate them are affected
-%by the environment.
+hidden. The operator $P \hide H$ defines a process that behaves as $P$, with
+the interactions modelled by events in the set $H$ hidden. Frequently, hiding
+is used in conjunction with parallelism:~it is often desirable to make
+internal actions of each process in a network of parallel processes , perhaps
+used for coordination of the network, invisible, while events happening at
+their interfaces remain observable.
 
@@
 models capture different notions of refinement. The simplest model
-characterises a process by its possible \emph{traces}; the set $\trc(P)$ denotes
- the sequences of (non-hidden) 
-events in which
-$P$ can engage.  We say that a process \emph{$P$ is trace-refined by another process
-$Q$}, written $P \lessdet_T Q$, if $\trc(Q)\subseteq \trc(P)$.
+characterises a process by its possible \emph{traces}; the set $\trc(P)$
+denotes the sequences of (non-hidden) events in which $P$ can engage.  We say
+that a process \emph{$P$ is trace-refined by another process $Q$}, written $P
+\lessdet_T Q$, if $\trc(Q)\subseteq \trc(P)$.
 
@@
 $s$. The failures model of a process $P$ records all its failures in a set
-$\fails(P)$. The semantic rules specify for each operator, how the failures of 
-the resulting process can be calculated from the failures of each operand. For example, 
-$\fails(P\intchoice Q) = \fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for the comprehensive list of these rules.
+$\fails(P)$.
 
-Using notation $P/s$ to denote the process $P$ after having engaged into trace $s$,
-the set 
-$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the  refusals
-of $P$ after the
-trace $s$. 
-Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure of $P$ and $Y\subseteq X$,
-then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$ follows.
+Semantic definitions specify, for each operator, how the traces or failures
+of the resulting process can be calculated from the traces or failures of
+each operand. For example, for internal choice, $\fails(P\intchoice Q) =
+\fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for a
+comprehensive list of these definitions.
 
-Failures refinement, $P \lessdet_F Q$, is defined by set containment 
-$\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed, this implies
-$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. As a consequence, failures 
-refinement implies trace refinement. Therefore, using the auxiliary 
-conformance relation
+Using the notation $P/s$ to denote the process $P$ after having engaged into
+trace $s$, the set $\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains
+the  refusals of $P$ after the trace $s$. Refusals are
+subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure
+of $P$ and $Y\subseteq X$, then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$
+follows.
+
+Failures refinement, $P \lessdet_F Q$, is defined as expected by set
+containment $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
+$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. As a
+consequence, failures refinement implies trace refinement. Therefore, using
+the auxiliary conformance relation
 \begin{equation}\label{eq:conf}
@@
 \end{equation}
-
+%
 Throughout this paper, the alphabet of the processes, that is the set of
@@
 %considering ``classical'' CSP processes, the alphabet is always finite.
-The
-FDR tool~\cite{fdr} supports model checking and semantic analyses of
+The FDR tool~\cite{fdr} supports model checking and semantic analyses of
 finite-state CSP processes.
 
-For finite CSP processes, since the refusals of each process state are
-subset-closed, $\refs(P/s)$ can be
-constructed from the set of \emph{maximal refusals}; these are defined by 
+For finite CSP processes, since refusals are
+subset-closed, $\refs(P/s)$ can be constructed from the set of \emph{maximal
+refusals} defined by
 %
@@
 Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
-reconstruct the refusals $\refs(P/s)$ by  
+reconstruct the refusals in the set $\refs(P/s)$ as described by
 %
@@
 Deterministic process states $P/s$ have exactly the one maximal refusal
-$\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of $P/s$, i.e.,   the events that $P/s$ may engage into.
-The maximal refusals in combination with the initials of a process express its degree of 
+defined by $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
+$P/s$, that is, the events that $P/s$ may engage into. The maximal refusals
+in combination with the initials of a process express its degree of
 nondeterminism.
+%
 \begin{example}
-\label{ex:nondetdegree}
-Process $P = (\Stop \intchoice Q)$ has maximal refusals 
-$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ 
-refuses to engage into any event, and this carries over to $P$, when applying the internal 
-choice operator. However, $P$ is distinguished from $\Stop$ by its initials: 
-$[P]^0 = [ \Stop \intchoice Q]^0 = [Q]^0$, so $P$ may engage nondeterministically in
-any initial event of $Q$, but can also refuse everything, due to internal selection of the $\Stop$ process.
+\label{ex:nondetdegree} $P = (\Stop \intchoice Q)$ has maximal refusals
+$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ refuses to engage in any event,
+and this is carried over to $P$ by the internal choice. However, $P$ is
+distinguished from $\Stop$ by its initials, which are defined by $[P]^0 = [
+\Stop \intchoice Q]^0 = [Q]^0$. So $P$ may engage nondeterministically in any
+initial event of $Q$, but also refuse everything, due to internal selection
+of $\Stop$.
 
-Assuming that $\Sigma = \{a,b,c,d\}$, process
-$Q = (e:\{a,b\} \then \Stop) \intchoice (e:\{c,d\} \then \Stop)$ has maximal refusals
-$\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials $[Q]^0=\Sigma$, so, in contrast to
-$P$, nondeterminism is reflected
-here by the presence of two maximal refusals.
-\xbox
+Assuming an alphabet $\Sigma = \{a,b,c,d\}$, the process
+%
+$$Q = (e:\{a,b\} \then \Stop) \intchoice (e:\{c,d\} \then \Stop)$$
+%
+has maximal refusals $\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials
+$[Q]^0=\Sigma$. In contrast to $P$, nondeterminism is reflected by
+two maximal refusals. \xbox
 \end{example}
@@
 
-As shown in~\cite{Roscoe:1994:CME:197600}, the failures semantics of 
-any finite-state CSP process $P$
-can be represented by a \emph{normalised transition graph}
+As shown in~\cite{Roscoe:1994:CME:197600}, the failures semantics of any
+finite-state CSP process $P$ can be represented by a \emph{normalised
+transition graph} $G(P)$ defined by a tuple
 $$
@@
 $e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
-$t$, then there is a transition, that is, an outgoing edge, from $n$ with label $e$, leading to node $t(n,e)$. Normalisation of
-$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
-in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable by
-sequences of edges labelled by $e_1\dots e_k$ and connecting states
-$\ii n,n_1,\dots,n_{k-1},n$, such that
+$t$, then there is a transition, that is, an outgoing edge, from $n$ with
+label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
+the fact that $t$ is a function. The graph construction
+in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are
+reachable by sequences of edges labelled by $e_1\dots e_k$ and connecting
+states $\ii n,n_1,\dots,n_{k-1},n$, such that
 \[
@@
 %
-By construction, $s\in\Sigma^*$ is a trace of $P$, if and only if 
-there is a path through $G(P)$ starting  at $\ii n$ whose edge labels
-coincide with $s$. In analogy to $\trc(P)$, we use
-notation $\trc(G(P))$ for the set of finite, initialised paths
-through $G(P)$, each path represented by its finite sequence of edge labels, and we
-note that $\trc(P) = \trc(G(P))$. 
-Since $G(P)$ is normalised, there is a unique node
-reached by applying the events from $s$ one by one, starting in $\ii n$.
-Therefore, $G(P)/s$  is also well defined. By $[n]^0$ we denote the
-\emph{initials} of $n$:~the set of events occurring as labels in any outgoing
-transitions.
+By construction, $s\in\Sigma^*$ is a trace of $P$, if, and only if, there is
+a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
+$s$. In analogy to $\trc(P)$, we use notation $\trc(G(P))$ for the set of
+finite, initialised paths through $G(P)$, each path represented by its finite
+sequence of edge labels, and we note that $\trc(P) = \trc(G(P))$. Since
+$G(P)$ is normalised, there is a unique node reached by applying the events
+from $s$ one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well
+defined. By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events
+occurring as labels in any outgoing transitions.
 $$
@@
 $$
-The graph construction from~\cite{Roscoe:1994:CME:197600} guarantees that
-$[G(P)/s]^0 = [P/s]^0$ for all traces of $P$.
+The graph construction from~\cite{Roscoe:1994:CME:197600} used to define
+$G(P)$ for any process $P$ guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
+traces $s$ of $P$.
 
-The total function $r$ maps each node $n$ to a non-empty set of  (possibly empty) subsets of $\Sigma$. 
-The graph construction guarantees that $r(G(P)/s)$ represents the maximal refusals
-of $P/s$ for all 
-$s\in\trc(P)$. As a consequence,  
+The total function $r$ maps each node $n$ to a non-empty set of  (possibly
+empty) subsets of $\Sigma$. The graph construction guarantees that
+$r(G(P)/s)$ represents the maximal refusals of $P/s$ for all $s\in\trc(P)$.
+As a consequence,
 \begin{equation}\label{eq:refsAndR}
@@
 so $G(P)$ allows us to re-construct the failures semantics of $P$.
+\fixme{alcc: Why not leave the commented out material in the report?}
 
-%
-%
 %To see that this approach works only for finite CSP processes, we consider
@@
 %%
-%
-%
 %By construction, normalised transition graphs reflect the \emph{failures
@@
 \begin{example}\label{ex:a}
-Consider the CSP process $P$ in Example~\ref{example:CSP}. Its transition
-graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state $P/\varepsilon$ 
-($\varepsilon$ denotes the empty trace)
-is represented there as Node\_0, with $\{ a\}$ as the only minimal
-acceptance, since $a$ can never be refused, and no other events are accepted.
-Having engaged into $a$, the transition emanating from Node\_0 leads to
-Node\_2 representing  the process state $P/a = Q\intchoice R$. The internal
-choice operator induces several minimal acceptances derived from $Q$ and $R$.
-Since these processes accept their initial events in external choice,
-$Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$ and $\{b,c\}$. Note
-that event $c$ can never be refused, since it is contained in each minimal
-acceptance set.
+We consider the process $P$ in Example~\ref{example:CSP}; its transition
+graph $G(P)$ is shown in Fig.~\ref{fig:tga}. The process state
+$P/\varepsilon$ (where $\varepsilon$ denotes the empty trace) is represented
+as Node\_0, with $\{ a\}$ as the only minimal acceptance, since $a$ is never
+refused and no other events are accepted. Having engaged in $a$, the
+transition from Node\_0 leads to Node\_2 representing the process state $P/a
+= Q\intchoice R$. The internal choice induces several minimal
+acceptances derived from $Q$ and $R$. Since these processes accept their
+initial events in external choice, $Q\intchoice R$ induces minimal acceptance
+sets $\{a,c\}$ and $\{b,c\}$. Note that event $c$ can never be refused, since
+it is contained in each minimal acceptance set.
 
@@
 %
-in the following
-way. 
+as established by results in the following lemma.
+%
 \begin{lemma}
@@
 \end{lemma}
-
-Formula (\ref{eq:trcrefa}) reflects trace refinement in terms of graph traces.  Formula~(\ref{eq:failrefa}) states how the $conf$ conformance relation can be expressed
-by means of the maximal refusal functions of the graphs involved, and (\ref{eq:failrefb})
-states the same in terms of the minimal acceptances that can be derived from the 
-maximal refusal functions by means of (\ref{eq:maxrefsminaccs}).
+%
+The result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
+traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
+refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
+by means of the maximal refusal functions of the graphs involved; and
+(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
+can be derived from the maximal refusal functions by means of
+(\ref{eq:maxrefsminaccs}). \fixme{alcc: I think at least the report should
+point to proofs.}
 
@@
 \subsubsection*{Reachability Under Sets of Traces}
-\label{sec:V}
-Given a finite-state CSP process $P$ and its normalised transition graph
-\[
-G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
-\]
-suppose that $V\subseteq\Sigma^*$ is a
-prefix-closed set  of sequences of events. By $t(\ii n,V)$ we denote the set
+\label{sec:V} Given a finite-state CSP process $P$ and its normalised
+transition graph $G(P)$,
+%\[
+%G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+%\]
+suppose that $V\subseteq\Sigma^*$ is a prefix-closed set  of sequences of
+events. By $t(\ii n,V)$ we denote the set
 \[
@@
 of normalised transition graphs. We need this only for the investigation of
-corresponding traces in reference processes and processes for SUTs.
-Therefore, the labelling of nodes with maximal refusals or minimal
-acceptances are disregarded in the product construction. Consider two
-normalised transition graphs
+corresponding traces in reference processes and processes for SUTs. So, the
+labelling of nodes with maximal refusals or minimal acceptances are
+disregarded in the product construction. We consider two normalised
+transition graphs
 \[
@@
 over the same alphabet $\Sigma$. Their product is defined by
+%
 \begin{eqnarray}
@@
 \end{eqnarray}
-
-The following lemma will be used in the proof of our main theorem.
+%
+The following lemma is used in the proof of our main theorem.
+%
 \begin{lemma}\label{lemma:reachproduc}
@@
 A \emph{minimal hitting set} is a hitting set that cannot be further reduced
-without losing the characteristic property (\ref{eq:hit}).
-By $\minhits(C)$ we denote the collection of minimal hitting sets for $C$.
-For the pathological case where $C$ contains an empty set, $\minhits(C)$ is also empty.
- The problem of
-determining minimal hitting sets is known to be %%% \cite{Book1975-BOOKRM,5533149}
-NP-hard~\cite{5533149}, but we will see below that it reduces the
-effort of testing for failures refinement from a factor of $2^{|\Sigma|}$ to a
-factor that equals the number of minimal hitting sets.
+without losing the characteristic property (\ref{eq:hit}). By $\minhits(C)$
+we denote the collection of minimal hitting sets for a collection $C$. For
+the pathological case where $C$ contains an empty set, $\minhits(C)$ is also
+empty.
 
-The following lemma establishes the fact that the $conf$ relation specified in (\ref{eq:conf}) can be chracterised by means of minimal acceptances and their
-minimal hitting sets.
+The problem of determining minimal hitting sets is %%% \cite{Book1975-BOOKRM,5533149}
+NP-hard~\cite{5533149}. We see below, however, that it reduces the effort of
+testing for failures refinement from a factor of $2^{|\Sigma|}$ to a factor
+that equals the number of minimal hitting sets.
+
+The following lemma establishes that the $conf$ relation specified in
+(\ref{eq:conf}) can be characterised by means of minimal acceptances and
+their minimal hitting sets.
+%
 \begin{lemma}
@@
 \]
-Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap
-A_P\neq\varnothing$ for all minimal acceptances $A_P$, but also $H\cap
-A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq
-A_Q$ for at least one $A_P$. As a consequence, each $H \in
-\minhits(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.
+Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap A_P\neq\varnothing$
+for all minimal acceptances $A_P$, but also $H\cap A_Q\neq\varnothing$ for
+every minimal acceptance $A_Q$, because $A_P\subseteq A_Q$ for at least one
+$A_P$. As a consequence, each $H \in \minhits(P/s)$ is also a hitting set for
+$\minaccs(G(Q)/s)$ as required.
 
-To prove ``$2 \Rightarrow 1$'', assume that that Statement~2 of the lemma holds but 
-that  $P\ conf\ Q$ does {\it not} hold. 
-According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
-exists $s\in\trc(P)\cap \trc(Q)$ such that
+To prove ``$2 \Rightarrow 1$'', assume that 2 holds, but that $P\ conf\ Q$
+does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
+(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
 \[
@@
 %
-
-It is important to point out that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
-Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and this carries over to $Q\intchoice \Stop$.
-From (\ref{eq:failrefb}) we conclude that
-$\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{ \varnothing \}$. These considerations
-clarify that $\minhits(P/s)$ is empty if and only if  $\minaccs(P) = \{ \varnothing \}$.
-We note that the proof of Lemma~\ref{lemma:hseta} also covers the situations where
-$\minaccs(P/s) = \{ \varnothing \}$ and therefore  $\minhits(P/s) = \varnothing$.
+We note that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
+Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and
+this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we
+conclude that $\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{
+\varnothing \}$. This clarifies that $\minhits(P/s)$ is empty if, and only
+if, $\minaccs(P) = \{ \varnothing \}$. The proof of Lemma~\ref{lemma:hseta}
+covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
+$\minhits(P/s) = \varnothing$.

### 2018-09-02 paper/abstract.tex
small suggestions

@@
-In this paper, two new contributions for model-based testing  
-with   Communicating Sequential Processes (CSP) are presented. 
-For a given CSP process representing the reference model, test suites 
-checking the conformance relations 
-trace refinement and  failures refinement are constructed, and their finiteness
-and completeness (i.e.
-capability to uncover conformity violations) is proven.
-While complete test suites for CSP processes have been previously investigated by several authors, a sufficient condition for their finiteness is presented here for the first time.
-Moreover, an optimal technique for
-testing the admissibility of an implementation's nondeterministic behaviour is described.
+In this paper, two new contributions for model-based testing with
+Communicating Sequential Processes (CSP) are presented. For a given CSP
+process representing the reference model, test suites checking the
+conformance relations trace and failures refinement are constructed, and
+their finiteness and completeness (that is, capability to uncover conformity
+violations) is proven. While complete test suites for CSP processes have been
+previously investigated by several authors, a sufficient condition for their
+finiteness is presented here for the first time. Moreover, an optimal
+technique for testing the admissibility of an implementation's
+nondeterministic behaviour is described.

### 2018-09-02 paper/finitecomplete-failures.tex
small suggestions

@@
 \]
-It is assumed that the events $\efail$ and $\epass$, denoting FAIL and PASS
-of the test execution, are events outside $\Sigma$. Since we assume that $Q$
-is free of livelocks, it is guaranteed that each test execution terminates
-after some $s\in\trc(Q)$ with length $(p+1)$ and consecutive $\epass$ or $\efail$ event at the latest. The test is
+It is assumed that the events $\efail$ and $\epass$, indicating a verdict
+FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
+assume that $Q$ is free of livelocks, it is guaranteed that each test
+execution terminates after some $s\in\trc(Q)$ with length $(p+1)$ and
+consecutive $\epass$ or $\efail$ event at the latest. The test is
 \emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
 every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
-$\epass$. This can also be  expressed by means of a failures refinement.
+$\epass$. This can also be  expressed by means of a failures refinement as
+defined below.
 %
@@
 
-Observe that the number of possible executions of $Q\parallel[\Sigma] U_F(p)$ is finite,
-because the number of traces $s$ with maximal length $(p+1)$ is finite and the
-sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and $\minhits(P/s)$ are finite. Further recall
-that $\minhits(P/s)$ may be empty, so that operand (\ref{eq:ufd})
-of the external choice construct  defaults to $\Stop$. In this case, however,
-$\minaccs(P/s) = \{\varnothing\}$ applies,
-so the test will execute branch (\ref{eq:ufb}) and
-terminate with $\epass$.
+We observe that the number of possible executions of $Q\parallel[\Sigma]
+U_F(p)$ is finite, because the number of traces $s$ with maximal length
+$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
+$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
+empty, so that operand (\ref{eq:ufd}) of the external choice construct
+defaults to $\Stop$. In this case, however, $\minaccs(P/s) = \{\varnothing\}$
+applies, so the test executes the branch (\ref{eq:ufb}) and terminates with
+$\epass$.
 
@@
 
-After having  run successfully
-through a trace of length $p$, and if $\minaccs(P/s)\neq\{\varnothing\}$ so that
-als $\minhits(P/s)\neq\varnothing$,
-the test changes its behaviour:
-instead of offering {\it all} legal events from $[P/s]^0$ to the SUT,
-it nondeterministically chooses
-a minimal hitting set of $\minaccs(P/s)$ and only offers the events contained in this set.
-If the SUT refuses to engage into any of these events, this reveals a violation of the
-failures refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept
-at least one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
-only terminates with  $\epass$, if such an event is accepted by the SUT.
+After having run through a trace of length $p$, and if
+$\minaccs(P/s)\neq\{\varnothing\}$ so that also
+$\minhits(P/s)\neq\varnothing$, the test changes its behaviour: instead of
+offering {\it all} legal events from $[P/s]^0$ to the SUT, it
+nondeterministically chooses a minimal hitting set of $\minaccs(P/s)$ and
+only offers the events contained in this set. If the SUT refuses to engage
+into any of these events, this reveals a violation of the failures
+refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should
+accept at least one event of each minimal hitting set in $\minhits(P/s)$.
+Therefore, the test only terminates with  $\epass$, if such an event is
+accepted by the SUT.
 
@@
 
-A CSP
-\emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a reference process
-$P$, a conformance relation $\sqsubseteq \in \{\lessdet_T, \lessdet_F\}$, and a fault domain
-${\cal D}$ which is a set of CSP processes over $P$'s alphabet that may or may not conform to $P$.
+A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a
+reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
+\lessdet_F\}$, and a fault domain ${\cal D}$ which is a set of CSP processes
+over $P$'s alphabet that may or may not conform to $P$.
 
@@
 \end{description}
+%
+In the next section, we establish the completeness of our test suite.
 
@@
 \begin{theorem}\label{th:failurestest}
-Let $P$ be a divergence-free CSP process over alphabet $\Sigma$
-whose normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal D}$ as
-the set of all divergence-free CSP processes over alphabet $\Sigma$, whose transition graph
-has at most $q$ states with $q \ge p$.
-Then the test suite
+Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
+normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
+D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
+whose transition graph has at most $q$ states with $q \ge p$. Then the test
+suite
 \[
@@
 \end{theorem}
-
-The proof of the theorem follows from the two lemmas below. The first states that test
-suite $\TS_F$ is sound, the second states that the suite is also exhaustive.
-
+%
+The proof of the theorem follows from the two lemmas below. The first states
+that test suite $\TS_F$ is sound, the second states that the suite is also
+exhaustive.
+%
 \begin{lemma}\label{lemma:mainfsound}
-Test suite $\TS_F$ generated from CSP process $P$, as specified in Theorem~\ref{th:failurestest}, is passed by
-every CSP process $Q$ satisfying $P\lessdet_F Q$.
+Test suite $\TS_F$ generated from CSP process $P$, as specified in
+Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
+$P\lessdet_F Q$.
 \end{lemma}
 \begin{proof}
-{\bf Step~1.}
-Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\ Q$
-according to (\ref{eq:failconf}).
-Since   $\trc(Q)\subseteq
-\trc(P)$,   any adaptive test
-$U_F(p)$ running in parallel with $Q$ will always enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or
-(\ref{eq:ufd}) of the external choice construction for $U_F(p,s)$.
-Branch~(\ref{eq:ufa}) can never be entered in the parallel execution of $Q$
-and $U_F(p)$, because $[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.
+{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\
+Q$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
+any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
+branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
+choice construction for $U_F(p,s)$. Branch~(\ref{eq:ufa}) can never be
+entered in the parallel execution of $Q$ and $U_F(p)$, because
+$[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.
 
-{\bf Step~2.}
-Moreover,  Lemma~\ref{lemma:hseta}
-implies that for all traces $s\in\trc(Q)\cap\trc(P)$, every $H$ in $\minhits(P/s)$ is
-a hitting set for $\minaccs(Q/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,s)$
-leads always to a PASS verdict, and branch
-(\ref{eq:ufc}) to test continuation without a verdict. For the last branch,
-we note that any selected minimal hitting set $H\in\minhits(P/s)$ has a
-non-empty intersection with each of the minimal acceptances of $Q/s$. As a
-consequence, $Q/s$ never blocks when offered events from $H$, and the test
-terminates with PASS event $\epass$. Note that this argument requires that
-$Q$ is free of livelocks, because otherwise the $\epass$-events might not become
-visible, due to unbounded sequences of hidden events performed by $Q$.
-\xbox
+{\bf Step~2.} Also,  Lemma~\ref{lemma:hseta} implies that for all traces
+$s\in\trc(Q)\cap\trc(P)$, every $H$ in $\minhits(P/s)$ is a hitting set for
+$\minaccs(Q/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,s)$ leads always to a
+PASS verdict, and branch (\ref{eq:ufc}) to test continuation without a
+verdict. For the last branch, we note that any selected minimal hitting set
+$H\in\minhits(P/s)$ has a non-empty intersection with each of the minimal
+acceptances of $Q/s$. As a consequence, $Q/s$ never blocks when offered
+events from $H$, and the test terminates with PASS event $\epass$. Note that
+this argument requires that $Q$ is free of livelocks, because otherwise the
+$\epass$-events might not become visible, due to unbounded sequences of
+hidden events performed by $Q$. \xbox
 \end{proof}
-
+%
 \begin{lemma}\label{lemma:mainfexhaustive}
-Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive for the fault
-model specified there.
+Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive
+for the fault model specified there.
 \end{lemma}
 \begin{proof}
-Consider a process $Q\in{\cal D}$ with
-$P\not\lessdet_F Q$, According to (\ref{eq:failconf}),
-this non-conformance can be caused in two possible ways corresponding to the cases $P\not\lessdet_T Q$ and
-$\neg(P\ conf\ Q)$, respectively:
+Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$, According to
+(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
+corresponding to the cases $P\not\lessdet_T Q$ and $\neg(P\ conf\ Q)$,
+respectively:
 \begin{description}
@@
 
-Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal acceptances,
-the cardinality $h$ of $\minhits(P/s)$ is maximal for the case where all $A_i$ are disjoint. In this case,  $h=|\minhits(P/s)| = \prod_{i=1}^\alpha|A_i|$.
-To find an upper bound for $h$, we prove the following lemma.
-
+Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
+acceptances, the cardinality $h$ of $\minhits(P/s)$ is maximal for the case
+where all $A_i$ are disjoint. In this case,  $h=|\minhits(P/s)| =
+\prod_{i=1}^\alpha|A_i|$. To find an upper bound for $h$, we use the result
+that we prove in the following lemma.
+%
 \begin{lemma}\label{lemma:minhitmax}
-Let $n\ge 2$ be any positive integer.
-We call a sequence $0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$, a partion of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$.
-Define
+Let $n\ge 2$ be any positive integer. We call a sequence
+%
+$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$,
+%
+a partition of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$. Define
+%
 $$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
@@
 \alpha\in \mathbb{N}\big\}.$$
+%
 \end{lemma}
 \begin{proof}
-Let $n\ge 2$.
-Obviously, \begin{equation}\label{eq:1} n\le p(n).\end{equation}
-Define $\pi(n)$ to be the following partition  of $n$:
+Let $n\ge 2$. Obviously, \begin{equation}\label{eq:1} n\le
+p(n).\end{equation} Define $\pi(n)$ to be the following partition  of $n$:
+%
 $$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
-Then the product of the numbers in
-$\pi(n)$ equals  $p(n)$. In the remainder of the proof
-we have to show that $p(n)$ is really maximal in the sense of the lemma.
+%
+Then the product of the numbers in $\pi(n)$ equals  $p(n)$. In the remainder
+of the proof we have to show that $p(n)$ is really maximal in the sense of
+the lemma.
 
-Since the number of partitions of a given $n$ is finite, $\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$ exists.
-
-Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with maximal product. Then,
-obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
-Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then from~(\ref{eq:1}) we have
+Since the number of partitions of a given $n$ is finite,
+%
+$$\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$$
+%
+exists. Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with
+maximal product. Then, obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
+Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then
+from~(\ref{eq:1}) we have
 $$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
@@
 $$
+%
+We note that $h$ is significantly smaller than $2^{|\Sigma|}$ -- this can
+also be intuitively motivated by the fact that  for every $H\in
+\minhits(P/s)$, all of its true subsets and true supersets are {\it not}
+contained in $\minhits(P/s)$.
 
-We not that $h$ is significantly smaller than $2^{|\Sigma|}$ -- this can also be intuitively motivated by the fact that  for every $H\in  \minhits(P/s)$, all of its true subsets and true supersets are {\it not} contained in $\minhits(P/s)$.
-
-According to Theorem~\ref{th:failurestest},
-we need to execute the tests $U_F(k)$
-for $k = 0,\dots,(pq-1)$; this results in a worst-case bound
-of (we use the formula for the sum of the geometric progression here)
+From Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
+$k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
+we use the formula for the sum of the geometric progression.
+%
 \[
@@
  \]
-
-From Lemma~\ref{lemma:hseta} we know that
- the worst-case bound for $h$   cannot be further improved, since
-the full collection of minimal hitting sets needs to be checked in order to verify
-the $conf$ relation.
-
-The authors of~\cite{Hennessy:1988:ATP:50497} and \cite{DBLP:conf/icfem/CavalcantiG07}
-suggest to test {\it every}
-non-empty
-subset of $\Sigma$ whose events cannot be completely refused in a given process state
-of the reference model; this leads to a worst-case estimate of $2^{|\Sigma|}-1$
-for the number of different sets to be offered to the SUT in the last step of
-the test execution, so our approach reduces the number of test executions in comparison
+%
+From Lemma~\ref{lemma:hseta} we know that the worst-case bound for $h$ 
+above cannot be further improved, since the full collection of minimal
+hitting sets needs to be checked to verify $conf$. The
+authors of~\cite{Hennessy:1988:ATP:50497} and
+\cite{DBLP:conf/icfem/CavalcantiG07} suggest to test {\it every} non-empty
+subset of $\Sigma$ whose events cannot be completely refused in a given
+process state of the reference model; this leads to a worst-case estimate of
+$2^{|\Sigma|}-1$ for the number of different sets to be offered to the SUT in
+the last step of the test execution, so our approach reduces the number of
+test executions in comparison
 to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07}.
 
-It is discussed in Section~\ref{sec:conc} that the term $pq$ occurring in this bound can be significantly reduced by using more complex test cases that need not execute {\it all}
-traces of $P$ with length less or equal to $pq$.
+It is discussed in Section~\ref{sec:conc} that the term $pq$ occurring in
+this bound can be significantly reduced by using more complex test cases that
+need not execute {\it all} traces of $P$ with length less or equal to $pq$.
 

### 2018-09-02 paper/introduction.tex
small suggestions

@@
 
-Generation methods for complete test suites have been developed for various
-formalisms. In this paper, we use \emph{Communicating Sequential Processes
-(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature process
-algebra that has been shown to be well-suited for the description of reactive
-control systems in many publications over almost five decades. Many of these
-applications are described in~\cite{Roscoe2010} and in the references there.
-Industrial success has also been reported, see, for example,
-\cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.
+Generation techniques for complete test suites have been developed for
+various formalisms. Here, we tackle \emph{Communicating Sequential
+Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
+process algebra that has been shown to be well-suited for the description of
+reactive control systems in many publications over almost five decades. Many
+of these applications are described in~\cite{Roscoe2010} and in the
+references there. Industrial success has also been reported; see, for
+example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.
 
@@
 
-Our complementary approach to the definition of CSP fault domains is
-presented in this paper. We observe that the failures semantics of
-every finite-state CSP process can
-be represented as a finite normalised transition graph, whose edges are
-labelled by the events the process engages in, and whose nodes are labelled
-by minimal acceptances or, alternatively, maximal
+Our approach to the definition of CSP fault domains is presented in this
+paper. We take advantage of results on model checking of CSP
+processes, where the failures semantics of a finite-state CSP process is
+represented as a finite normalised transition graph, whose edges are labelled
+by the events the process engages in, and whose nodes are labelled by minimal
+acceptances or, alternatively, maximal
 refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
-degree of nondeterminism present in a given process state that is in
+degree of nondeterminism present in a process state that is in
 one-one-correspondence to a node of the normalised transition graph. Inspired
-by the way that fault-domains are specified for finite state machines (FSMs),
-we define them here as the set of CSP processes whose normalised transition graphs
-do not exceed the order (i.e.~the number of nodes)
-of the reference model's graph by more than a given
-constant.
+by the way fault-domains are specified for finite state machines (FSMs),
+we define them here as the set of CSP processes whose normalised transition
+graphs do not exceed the order (that is, the number of nodes) of the
+reference model's graph by more than a given constant.
 
@@
 testing against nondeterministic FSMs, as published, for example,
-in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the approach to testing the admissibility of an SUT's nondeterministic behaviour is optimal in the
-sense that it cannot be established with fewer test cases.
+in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the
+approach to testing the admissibility of an SUT's nondeterministic behaviour
+is optimal in the sense that it cannot be established with fewer test cases.
+\fixme{alcc: I did not understand this?}
 
@@
 \subsubsection*{Overview}
-In Section~\ref{section:preliminaries},
-we present the background material relevant to our work.
-In Section~\ref{sec:finitecompletefails}, finite complete test suites for verifying failures refinement are presented.
-Test suites checking trace refinement may be regarded as a simplified version of the former class, they are presented in Section~\ref{sec:finitecomplete}.
-A sample test suite is presented in Section~\ref{sec:case}.
-The results obtained are discussed in Section~\ref{sec:conc}.
-
-References to related work are given throughout the paper where appropriate.
+In Section~\ref{section:preliminaries}, we present the background material
+relevant to our work. In Section~\ref{sec:finitecompletefails}, finite
+complete test suites for verifying failures refinement are presented. Test
+suites checking trace refinement are a simplified version of the former
+class; they are presented in Section~\ref{sec:finitecomplete}. A sample test
+suite is presented in Section~\ref{sec:case}. Our results are discussed in
+Section~\ref{sec:conc}, where we also conclude. References to related work
+are given throughout the paper where appropriate.
 

### 2018-09-02 paper/preliminaries.tex
small suggestions

@@
 % =========================================================================
+
+In this section, we present CSP and the concept of minimal hitting set, which
+is central to our novel notion of test for failures refinement.
 
@@
 A prefixing operator $e \then P$ defines a process that is ready to engage in
-the event $e$, pending agreement of its environment to synchronising. After
-$e$ occurs, the process behaves as defined by $P$. The environment can be
-other processes, in parallel, or the environment of a system as a whole.
+the event $e$, pending agreement of its environment to synchronise. After $e$
+occurs, the process behaves as defined by $P$. The environment can be other
+processes, in parallel, or the environment of a system as a whole.
 
@@
 initial events of $P$ and $Q$. Once synchronisation takes place, the process
-that offered this event is chosen and the other is discarded. In an internal
-choice $P \intchoice Q$, the environment does not have an opportunity to
-interfere:~the choice is made by the process itself.
-
+that has offered this event is chosen and the other is discarded. In an
+internal choice $P \intchoice Q$, the environment does not have an
+opportunity to interfere:~the choice is made by the process.
+%
 \begin{example}\label{example:CSP}
@@
 external and internal choice operators allow us to define a choice over a
-collection of processes $P(i)$. By convention, external choices evaluate to
-the process $\Stop$, which deadlocks:~does not engage into any event and does
-not terminate, if the index set $I$ is empty. An iterated internal choice is
-not defined for an empty index set.
+collection of processes $P(i)$. By convention, , if the index set $I$ is
+empty, external choices evaluate to the process $\Stop$, which
+deadlocks:~does not engage into any event and does not terminate. An iterated
+internal choice is not defined for an empty index set. \fixme{alcc: Note the
+difference to internal choice. I think this is not a problem here.}
 
@@
 is used in conjunction with parallelism:~it is often desirable to make
-internal actions of each process in a network of parallel processes , perhaps
+internal actions of each process in a network of parallel processes, perhaps
 used for coordination of the network, invisible, while events happening at
@@
 trace $s$, the set $\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains
-the  refusals of $P$ after the trace $s$. Refusals are
+the  refusals of $P$ after $s$. Refusals are
 subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure
@@
 containment $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
-$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. As a
-consequence, failures refinement implies trace refinement. Therefore, using
-the auxiliary conformance relation
+$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. So, for
+divergences-free processes, failures refinement implies trace refinement.
+Therefore, using the conformance relation $conf$ 
 \begin{equation}\label{eq:conf}

### 2018-09-02 paper/preliminaries.tex
small problem with nondeterminism

@@
 divergences-free processes, failures refinement implies trace refinement.
-Therefore, using the conformance relation $conf$ 
+Therefore, using the conformance relation
+%
 \begin{equation}\label{eq:conf}
@@
 \end{equation}
+%
 failures refinement can be expressed by $\lessdet_T$ and $conf$.
+%
 \begin{equation}\label{thm:fref-tref-conf}
@@
 a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
-$s$. In analogy to $\trc(P)$, we use notation $\trc(G(P))$ for the set of
+$s$. In analogy to $\trc(P)$, we use the notation $\trc(G(P))$ for the set of
 finite, initialised paths through $G(P)$, each path represented by its finite
-sequence of edge labels, and we note that $\trc(P) = \trc(G(P))$. Since
-$G(P)$ is normalised, there is a unique node reached by applying the events
-from $s$ one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well
-defined. By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events
-occurring as labels in any outgoing transitions.
+sequence of edge labels. We note that $\trc(P) = \trc(G(P))$. Since $G(P)$ is
+normalised, there is a unique node reached by applying the events from $s$
+one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well defined.
+By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
+as labels in any outgoing transitions.
 $$
@@
 \end{equation}
+\fixme{alcc: proof in the report perhaps?}
 
@@
 transition from Node\_0 leads to Node\_2 representing the process state $P/a
-= Q\intchoice R$. The internal choice induces several minimal
-acceptances derived from $Q$ and $R$. Since these processes accept their
-initial events in external choice, $Q\intchoice R$ induces minimal acceptance
-sets $\{a,c\}$ and $\{b,c\}$. Note that event $c$ can never be refused, since
-it is contained in each minimal acceptance set.
+= Q\intchoice R$. The internal choice induces several minimal acceptances
+derived from $Q$ and $R$. Since these processes accept their initial events
+in external choice, $Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$
+and $\{b,c\}$. We note that the event $c$ can never be refused, since it is
+contained in each minimal acceptance set.
 
-Having engaged into $c$, the next process state is represented by Node\_1.
-Due to normalisation, there is only a single transition satisfying
+Having engaged in $c$, the next process state is represented by Node\_1. Due
+to normalisation, there is only a single transition satisfying
 $t(\text{Node\_2},c) = \text{Node\_1}$. This transition, however, can have
@@
 process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
-two minimal acceptance sets labelling Node\_1.
-Similar considerations explain the other nodes and transitions in
-Fig.~\ref{fig:tga}.
+two minimal acceptance sets labelling Node\_1. Similar considerations explain
+the other nodes and transitions in Fig.~\ref{fig:tga}.
 
@@
 \end{example}
-
+%
 Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
@@
 %\]
-suppose that $V\subseteq\Sigma^*$ is a prefix-closed set  of sequences of
+we suppose that $V\subseteq\Sigma^*$ is a prefix-closed set  of sequences of
 events. By $t(\ii n,V)$ we denote the set
@@
 \end{proof}
+\fixme{Abrupt end to the section. It has more than just preliminary material.
+It has some new results.}
 
@@
 \begin{proof}
-For showing ``$1 \Rightarrow 2$'', assume   $P\ conf\ Q$ and
-  $s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref},
-(\ref{eq:failrefb}), states that
+For showing ``$1 \Rightarrow 2$'', assume   $P\ conf\ Q$ and $s\in\trc(P)\cap
+\trc(Q)$. Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}), states that
 \[

### 2018-09-03 paper/casestudies.tex
added discussion + more suggestions

@@
 %
-In a state $n = G(P)/s$, the set $\text{\it initials}(n)$ equals the events
-labelling outgoing edges of $n$, so  $\text{\it initials}(n) = [P/s]^0$. Function
-$\minhits$ maps $n$ to the set of all minimal hitting sets associated
-with the minimal acceptances of $n$, so $\minhits(n) =
-\minhits(P/s)$. Using the transition function $t$ and the above
-functions, $U_F(p)$ can be re-written as the failures-equivalent CSP process
-below.
+As mentioned, in a state $n = G(P)/s$, the set $\text{\it initials}(n)$
+equals the events labelling outgoing edges of $n$, so  $\text{\it
+initials}(n) = [P/s]^0$. The function $\minhits$ maps $n$ to the set of all
+minimal hitting sets associated with the minimal acceptances of $n$, so
+$\minhits(n) = \minhits(P/s)$. Using the transition function $t$ and the
+above functions, $U_F(p)$ can be re-written as the failures-equivalent CSP
+process below.
 %
@@
 \end{eqnarray}
+\fixme{alcc: I think the example should be for the theory presented before,
+not with a different notion of test. In particular, 33 is very different.}
 
@@
 \begin{example}
-\label{ex:uf1tests} Consider the following erroneous implementation $Z$ of
-process $P$ from Example~\ref{ex:a} from the point of view of failures
-refinement.
+\label{ex:uf1tests} Consider the following implementation $Z$ of process $P$
+from Example~\ref{example:CSP} that is erroneous from the point of view of
+failures refinement.
 \begin{eqnarray*}
@@
 \\
-R_1(r_{max},k) & = & (k < r_{max}) \& \big( c\then R_1(r_{max},k+1)\extchoice b\then Z  \big)
+R_1(r_{max},k) & = & (k < r_{max}) \& \big(b\then Z \extchoice  c\then R_1(r_{max},k+1)\big)
 \\ & & \extchoice
-\\ & & (k = r_{max}) \& \big( c\then R_1(r_{max},r_{max})\intchoice b\then Z  \big)
+\\ & & (k = r_{max}) \& \big(b\then Z \intchoice c\then R_1(r_{max},r_{max})\big)
 \end{eqnarray*}
@@
 \end{example}
+\fixme{alcc: I would write down the actual test generated.}
 

### 2018-09-03 paper/conclusion.tex
small suggestions
added discussion + more suggestions

@@
-\section{Discussion}
+% ==============================================================================
+\section{Discussion and Conclusions}
 \label{sec:conc}
 % ==============================================================================
-\subsubsection*{Discussion of Fault Domains}
-
-@todo
 
@@
 It is known from complete testing strategies for finite state machines that
-the upper bound $pq$ for the lengths of the 
-traces used in our tests to investigate the SUT behaviour can be reduced. It is also
-known from FSM testing that it is not necessary to test {\it all} traces up to this maximal length. Notable complete 
-strategies supporting this fact have been presented, for example,
- in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}. From~\cite{Huang2017} it
- is known that complete 
- FSM testing theories can be translated to other formalisms, such as 
-Extended Finite State Machines, Kripke Structures, or CSP, resulting in likewise complete
-test strategies for the latter. We intend to study translations of several promising FSM
-strategies to CSP in the future. These will effectively reduce the upper bound 
-$\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The bound $h$ for
-the number of sets to be used in probing the SUT for illegal deadlocks, however, cannot be further reduced, as has been established in Lemma~\ref{lemma:hseta}.
+the upper bound $pq$ for the lengths of the traces used in our tests to
+investigate the SUT behaviour can be reduced. It is also known from FSM
+testing that it is not necessary to test {\it all} traces up to this maximal
+length. Notable complete strategies supporting this fact have been presented,
+for example,
+in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}.
+From~\cite{Huang2017} it is known that complete FSM testing theories can be
+translated to other formalisms, such as Extended Finite State Machines,
+Kripke Structures, or CSP, resulting in likewise complete test strategies for
+the latter. We intend to study translations of several promising FSM
+strategies to CSP in the future. These will effectively reduce the upper
+bound $\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The
+bound $h$ for the number of sets to be used in probing the SUT for illegal
+deadlocks, however, cannot be further reduced, as has been established in
+Lemma~\ref{lemma:hseta}. \fixme{alcc: my guess is that these complete tests
+are not for refinement. I think we can perhaps say something about this?}
 
@@
 of the execution was blocked, due to legal, but nondeterministic behaviour of the
-SUT. In contrast to that, 
-our test cases
-specified in Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
-adaptive. This has the advantage that test executions $(Q\parallel[\Sigma] U_F(p))$
-for failures refinement
-will never block before the final step specified by branch (\ref{eq:ufd}). As
-a consequence, we do not need inconclusive test results. It should be noted, however,
-that it is necessary for our test verdicts to recognise also deadlocks in the final
-test step and interpret 
-them as FAIL, as described in Section~\ref{sec:finitecompletefails}. In practice,
-this is realised by adding a timeout event to the testing environment which indicates
-deadlock situations. For real-time systems, this is an accepted technique, because
-the SUT has to respond within a pre-defined latency interval, otherwise its behaviour 
-is considered to be blocked and regarded as a failure. Our tests executions 
-$(Q\parallel[\Sigma] U_T(p))$ for trace refinement never block at all. The adaptive       
-behaviour of our test cases, however, induce the obligation to check that {\it all}
-possible executions have been performed before the test can be considered as passed.
-Typically, it is therefore assumed that a 
-\emph{complete testing assumption}~\cite{hierons_testing_2004} holds which means that
-every possible behaviour of the SUT has been performed after a finite number of
-test executions. 
-In practice, this is realised by executing each test several times, 
-recording the traces that have been performed, and using hardware or software coverage 
-analysers to determine whether all possible test execution behaviours of the SUT
-have been observed. Therefore, the application of adaptive test cases comes at the price
-of having to apply some grey-box testing techniques enabling us to decide whether
-all SUT behaviours have been observed.
+SUT.
+
+In contrast to that, our test cases specified in
+Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
+adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
+U_F(p))$ for failures refinement never blocks before the final step specified
+by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
+should be noted, however, that it is necessary for our test verdicts to
+recognise also deadlocks in the final test step and interpret them as FAIL,
+as described in Section~\ref{sec:finitecompletefails}. In practice, this is
+realised by adding a timeout event to the testing environment which indicates
+deadlock situations. For real-time systems, this is an accepted technique,
+because the SUT has to respond within a pre-defined latency interval,
+otherwise its behaviour is considered to be blocked and regarded as a
+failure. \fixme{alcc: if deadlock is not acceptable?}
+
+Our tests executions $(Q\parallel[\Sigma] U_T(p))$ for trace refinement never
+block at all. The adaptive behaviour of our test cases, however, induce the
+obligation to check that {\it all} possible executions have been performed
+before the test can be considered as passed. Typically, it is therefore
+assumed that a \emph{complete testing assumption}~\cite{hierons_testing_2004}
+holds, which means that every possible behaviour of the SUT is performed
+after a finite number of test executions. In practice, this is realised by
+executing each test several times, recording the traces that have been
+performed, and using hardware or software coverage analysers to determine
+whether all possible test execution behaviours of the SUT have been observed.
+Therefore, adaptive test cases come at the price of having to apply some
+grey-box testing techniques enabling us to decide whether all SUT behaviours
+have been observed.
+
+% ==============================================================================
+\subsubsection*{Discussion of Fault Domains}
+As already mentioned, the work in~\cite{DBLP:conf/pts/CavalcantiS17} defines
+a fault domain as the set of processes that refine a given CSP process.  In
+that context, only testing for traces refinement is considered, and the
+complete test suites may not be finite. So, the work presented here go well
+beyond what is achieved there. On the other hand,
+\cite{DBLP:conf/pts/CavalcantiS17} presents an algorithm for test generation
+that can be easily adapted to consider additional selection and termination
+criteria, like, for example, the length of the traces used to construct
+tests. It would be possible, for instance, to use the bound indicated here.
+Morevoer, specifying a fault domain as a CSP process allows us to model
+domain-specific knowledge using CSP. For example, if an initialisation
+component defined by a process $I$ can be regarded as correct without further
+testing, we can use $I; RUN$ as a fault domain, to indicate that any SUT of
+interest implements $I$ correctly, but afterwards has a potentially arbitrary
+behaviour specified by $RUN$.
 
@@
 As explained in the previous sections, passing a test is specified by
-$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$
-for failures testing. If the SUT $Q$ is not a programmed piece of software or
-an integrated HW/SW system, but just another CSP process specification,
-it is of course possible to verify these pass criteria using the FDR4 model checker.
-For checking the refinement relation $P\lessdet_F Q$, the pass criteria have to be verified
-for $k=0,\dots,pq-1$, where $p$ and $q$ indicate the number of nodes in $P$'s transition
-graph and the maximal number of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since the test cases
-$U_F(k)$ have such a simple structure, it is an interesting question for further 
-research whether checking 
-$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster than directly 
-checking $P\lessdet_F Q$, as one would do in the usual 
-approach with FDR4. This is of particular interest, since the checks could be parallelised 
-on several CPUs. Alternatively it is interesting to investigate whether the
-the check\footnote{We are grateful to Bill Roscoe for suggesting this option.}
+$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for
+failures testing. If the SUT $Q$ is not a programmed piece of software or an
+integrated hardware or software system, but just another CSP process
+specification, it is of course possible to verify these pass criteria using
+the FDR4 model checker. For checking the refinement relation $P\lessdet_F Q$,
+the pass criteria have to be verified for $k=0,\dots,pq-1$, where $p$ and $q$
+indicate the number of nodes in $P$'s transition graph and the maximal number
+of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since
+the test cases $U_F(k)$ have such a simple structure, it is an interesting
+question for further research whether checking $(\epass\then\Stop) \lessdet_F
+(Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster
+than directly checking $P\lessdet_F Q$, as one would do in the usual approach
+with FDR4. This is of particular interest, since the checks could be
+parallelised on several CPUs. Alternatively it is interesting to investigate
+whether the check\footnote{We are grateful to Bill Roscoe for suggesting this
+option.}
 \[
@@
 
-For large and complex CSP processes $Q$, it may be too time consuming to generate
-its normalised transition graph, so that its number $q$ of nodes is unknown. 
-In such a case the suggested options may still be used
-as efficient bug finders: use  $p$ of reference process $P$ as initial $q$-value
-and 
+For large and complex CSP processes $Q$, it may be too time consuming to
+generate its normalised transition graph, so that its number $q$ of nodes is
+unknown. In such a case the suggested options may still be used as efficient
+bug finders: use  $p$ of reference process $P$ as initial $q$-value and
 increment $q$ from there, as long as each increment reveals new errors.
@@
 In this paper, we have introduced finite complete testing strategies for
-model-based testing against CSP reference models. The strategies are applicable   
-to the conformance relations failures refinement and trace refinement. The underlying fault domains have been defined as the sets of CSP processes whose normalised transition
-graphs did not have more than a given number of additional nodes, when compared to the transition graph of the reference process. For these domains, finite complete 
-test suites are available. We have shown for the strategy checking failures refinement,
- that the way of probing the SUT for illegal deadlocks used in our test cases is 
- optimal in the sense that it is not possible to guarantee exhaustiveness with fewer
- probes.
+model-based testing against CSP reference models. The strategies are
+applicable to the conformance relations failures refinement and trace
+refinement. The underlying fault domains have been defined as the sets of CSP
+processes whose normalised transition graphs do not have more than a given
+number of additional nodes, when compared to the transition graph of the
+reference process. For these domains, finite complete test suites are
+available. We have shown for the strategy to check failures refinement that
+the way of probing the SUT for illegal deadlocks used in our test cases is
+optimal in the sense that it is not possible to guarantee exhaustiveness with
+fewer probes.
 
@@
 \paragraph{Acknowledgements}
-The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for their advice on using the FDR4 model checker and for very helpful discussions concerning the potential implications of this paper in the field of model-checking.
+The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for
+their advice on using the FDR4 model checker and for very helpful discussions
+concerning the potential implications of this paper in the field of model
+checking. The work of Cavalcanti is funded by the Royal Academy of
+Engineering and UK EPSRC Grant EP/R025134/1.

### 2018-09-03 paper/finitecomplete-failures.tex
added discussion + more suggestions

@@
 % ==========================================================================
+
+Here, we define our notion of tests for failures refinement, and then prove
+completeness of our suite. Finally, we study to complexity of our approach by
+identifying a bound on the number of tests we need in a complete suite.
 
@@
 applies, so the test executes the branch (\ref{eq:ufb}) and terminates with
-$\epass$.
+$\epass$. \fixme{alcc: Oops. This is undefined in CSP.}
 
@@
 only offers the events contained in this set. If the SUT refuses to engage
-into any of these events, this reveals a violation of the failures
-refinement: according to Lemma~\ref{lemma:hseta}, a conforming SUT should
-accept at least one event of each minimal hitting set in $\minhits(P/s)$.
-Therefore, the test only terminates with  $\epass$, if such an event is
-accepted by the SUT.
+into any of these events, this reveals a violation of failures refinement:
+according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
+one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
+only terminates with  $\epass$, if such an event is accepted by the SUT.
 
@@
 In the next section, we establish the completeness of our test suite.
+\fixme{alcc: I don't think we need a section just for this material.}
 
@@
 \begin{lemma}\label{lemma:mainfsound}
-Test suite $\TS_F$ generated from CSP process $P$, as specified in
+Test suite $\TS_F$ generated from a CSP process $P$, as specified in
 Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying

### 2018-09-03 paper/conclusion.tex
some more suggestions

@@
 deadlocks, however, cannot be further reduced, as has been established in
-Lemma~\ref{lemma:hseta}. \fixme{alcc: my guess is that these complete tests
-are not for refinement. I think we can perhaps say something about this?}
+Lemma~\ref{lemma:hseta}.
 
@@
 otherwise its behaviour is considered to be blocked and regarded as a
-failure. \fixme{alcc: if deadlock is not acceptable?}
+failure. \fixme{alcc: if deadlock is not acceptable? I think you need
+instrumentation here as well. }
 
@@
 interest implements $I$ correctly, but afterwards has a potentially arbitrary
-behaviour specified by $RUN$.
+behaviour specified by $RUN$. The work in~\cite{DBLP:conf/pts/CavalcantiS17}
+is not restricted to finite or non-terminating processes.
 

### 2018-09-04 paper/finitecomplete-failures.tex
some more suggestions
typo in proof of complexity
small changes from the meeting
Section on Complexity Consideration: Add example showing that maximal trace length pq is really necessary.

@@
 
+%% =========================================================================
+%\subsection{Fault Models and Complete Test Suites}
+
 % =========================================================================
-\subsection{Fault Models and Complete Test Suites}
+\subsection{A Finite Complete Test Suite for Failures Refinement}
 
@@
 %
-In the next section, we establish the completeness of our test suite.
-\fixme{alcc: I don't think we need a section just for this material.}
-
-% =========================================================================
-\subsection{A Finite Complete Test Suite for Failures Refinement}
-
-After the informal explanation of tests representing adaptive test cases, we are ready to state the main theorem of this paper.
+In the sequel, we establish the completeness of our test suite which is stated 
+in the following main theorem.
+%%%\fixme{alcc: I don't think we need a section just for this material.}
 
@@
 multiplied by the number $h$ of minimal hitting sets in
-$\minhits(P/s)$. The first factor
-has worst-case upper bound $\ell\le |\Sigma|^k$.
+$\minhits(P/s)$.
 
+% -------------------------------------------------------------------------
+\subsubsection*{Estimation of $\ell$}
+The first factor $\ell$
+has worst-case upper bound $\ell\le |\Sigma|^k$. As an example, where this upper bound 
+is really met, consider the reference process
+\[
+RUN(\Sigma) = e:\Sigma \then RUN(\Sigma)
+\]
+The normalised transition graph of this process has a single state, and its initials
+are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(k)$ can 
+never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly 
+$|\Sigma|^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
+for each of their events.
+
+% -------------------------------------------------------------------------
+\subsubsection*{Estimation of $h$}
 Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
 acceptances, the cardinality $h$ of $\minhits(P/s)$ is maximal for the case
-where all $A_i$ are disjoint. In this case,  $h=|\minhits(P/s)| =
+where all $A_i$ are disjoint. In this case,  $h~~=~~|\minhits(P/s)|~~=~~
 \prod_{i=1}^\alpha|A_i|$. To find an upper bound for $h$, we use the result
@@
 %
-$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$,
+$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$
 %
@@
 $$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
-Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a product value
-which is not less then the original partition $a_1, \ldots , a_{\alpha}$. This process
-can be repeated, until every $a_i$ in the resulting partition is either 2 or 3.
-Hence there exists a partition of $n$ with maximal product, such that every term is $2$ or $3$.
-Since $6=2+2+2=3\cdot 3$ and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than $3$.
-Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\}, i=1,\dots,\alpha$, be a partition of $n$ with maximal product.
-Let $k_1\le 2$ be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
-Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$ with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have
+Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a
+product value which is not less then the original partition $a_1, \ldots ,
+a_{\alpha}$. This process can be repeated, until every $a_i$ in the resulting
+partition is either 2 or 3. Hence there exists a partition of $n$ with
+maximal product, such that every term is $2$ or $3$. Since $6=2+2+2=3 + 3$
+and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than
+$3$. Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\},
+i=1,\dots,\alpha$, be a partition of $n$ with maximal product. Let $k_1\le 2$
+be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
+Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$
+with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have
 
@@
 $$h=\begin{cases}
-3^{|\frac{\Sigma}{3}|} &\text{if}\,\, \exists k: |\Sigma|=3k\\
-2^2\cdot 3^{\frac{|\Sigma|-4}{3}} &\text{if}\,\, \exists k : |\Sigma| =3k+1\\
-2\cdot 3^{\frac{|\Sigma| - 2}{3}} &\text{if}\,\, \exists k: |\Sigma| =3k+2
+3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\  |\Sigma|~~\equiv~~0 \mod 3\\
+3^{\lfloor\frac{|\Sigma|}{3}\rfloor}  + 3^{\lfloor\frac{|\Sigma|}{3}\rfloor-1}&\text{if}\ |\Sigma|~~\equiv~~1\mod 3\\
+2\cdot 3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\ |\Sigma|~~\equiv~~2 \mod 3
 \end{cases}
@@
 %
-We note that $h$ is significantly smaller than $2^{|\Sigma|}$ -- this can
+We note that $h$ is significantly smaller than $2^{|\Sigma|}-1$, this can
 also be intuitively motivated by the fact that  for every $H\in
@@
 h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
- \text{or, asymptotically,}\  O(3^{|\frac{\Sigma}{3}|}\cdot|\Sigma|^{(pq-1)}).
+ \text{or, asymptotically,}\  O(3^{\lfloor\frac{|\Sigma|}{3}\rfloor}\cdot|\Sigma|^{(pq-1)}).
  \]
@@
 
-It is discussed in Section~\ref{sec:conc} that the term $pq$ occurring in
-this bound can be significantly reduced by using more complex test cases that
-need not execute {\it all} traces of $P$ with length less or equal to $pq$.
+% -------------------------------------------------------------------------
+\subsubsection*{Upper Bound $pq$ for Number of Test Cases}
+
+According to Theorem~\ref{th:failurestest}, the tests $U_F(k)$ need to be executed for 
+$k = 0,\dots,pq-1$ to guarantee completeness. This means that the SUT is verified 
+with test traces up to, and including, length $pq$. It is interesting to  investigate
+whether this maximal length is really necessary, or whether one could elaborate
+alternative complete test strategies where the SUT is tested with shorter traces only. 
+Indeed, an example  presented in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} 
+shows that when testing for equivalence of deterministic FSMs, it is sufficient to
+test the SUT with traces of significantly shorter length.
+
+The following example, however, shows that the maximal length $pq$ is really required 
+when testing for refinement.
+\begin{example}\label{ex:pq}
+Consider the CSP reference process $P$ and an erroneous implementation $Q$ specified
+as follows.
+\begin{eqnarray*}
+P & = & a \then P_1 \intchoice b \then P_1 \intchoice c \then P_1
+\\
+P_1 & = & a \then P \extchoice b\then P
+\\ & &  \\
+Q & = & a\then Q_1 \extchoice b\then Q_1
+\\
+Q_1 & = & a\then Q_2 \extchoice b\then Q_2
+\\
+Q_2 & = & a\then Q \intchoice b\then Q
+\end{eqnarray*}
+Obviously, $P$'s normalised transition graph has 2 nodes, 
+while $Q$'s graph has 3.
+It is easy to see (and can be checked with FDR4) that $P\lessdet_T Q$, but 
+$\neg(P\lessdet_F Q)$. Furthermore, it can also be shown using the FDR4 tool that
+the ``test passed condition'' 
+\[
+(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k))\hide \Sigma
+\]
+holds for $U_F(0),\dots,U_F(4)$, but fails for $U_F(5)$. This means that the
+non-conformance of $Q$ cannot be detected by any test trace of length 
+less or equal to 5, but is revealed (as expected from Theorem~\ref{th:failurestest})
+by a trace of length 6, because the last event offered by the test $U_F(5)$ is 
+refused by $Q$.
+\xbox
+\end{example}
+%
+Generalising Example~\ref{ex:pq}, it can be shown that for any pair $p,q\in\mathbb{N}$
+whose greatest common divisor is 1, there exist reference processes with $p$ states
+and implementation processes with $q$ states, such that non-conformance of the 
+implementation
+can only be detected with a trace of length $pq$.
+
+It is discussed in Section~\ref{sec:conc} how the number of test traces to be executed 
+by complete test suites for failures refinement   can still be reduced {\it without} reducing 
+the maximal length.
 

### 2018-09-05 paper/finitecomplete-failures.tex
Small updates

@@
 \\ & & \extchoice \nonumber
-\\ & & (\#s = p ) \& \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
+\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber 
+\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
 \label{eq:ufd}
@@
 FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
-assume that $Q$ is free of livelocks, it is guaranteed that each test
-execution terminates after some $s\in\trc(Q)$ with length $(p+1)$ and
-consecutive $\epass$ or $\efail$ event at the latest. The test is
+assume that $Q$ is free of livelocks, it is guaranteed that events 
+$\efail$ or $\epass$ always become visible, if they are the only events
+$U_F(p)/s$ is ready to engage in.
+
+The specification of $U_F(p,s)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
+If branch (\ref{eq:ufa})
+is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
+stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
+is the last to be entered, the process either accepts another event
+$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
+of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
+$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
+and the execution stops without a PASS or FAIL event. Such an execution is also interpreted as FAIL.
+
+The test is
 \emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
@@
 executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
-because $\#s = p$. This is explained further in the next paragraphs.
+because $(\#s = p\wedge \minhits(P/s)\neq\varnothing)$. 
+This is explained further in the next paragraphs.
 
@@
 $\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
-empty, so that operand (\ref{eq:ufd}) of the external choice construct
-defaults to $\Stop$. In this case, however, $\minaccs(P/s) = \{\varnothing\}$
-applies, so the test executes the branch (\ref{eq:ufb}) and terminates with
-$\epass$. \fixme{alcc: Oops. This is undefined in CSP.}
+empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
+is undefined. Therefore, the associated guard condition contains the requirement
+$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken.
+Recall that $\minhits(P/s)=\varnothing$ implies  $\minaccs(P/s) = \{\varnothing\}$,
+so branches (\ref{eq:ufa}) and (\ref{eq:ufb}) can be taken in this situation.
+%%%\fixme{alcc: Oops. This is undefined in CSP.}
 

### 2018-09-05 paper/introduction.tex
Small updates

@@
 acceptances or, alternatively, maximal
-refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
+refusals~\cite{Roscoe:1994:chapter}. The maximal refusals express the
 degree of nondeterminism present in a process state that is in

### 2018-09-05 paper/preliminaries.tex
small changes from the meeting
Small updates

@@
 deadlocks:~does not engage into any event and does not terminate. An iterated
-internal choice is not defined for an empty index set. \fixme{alcc: Note the
-difference to internal choice. I think this is not a problem here.}
+internal choice is not defined for an empty index set.
 
@@
 
-As shown in~\cite{Roscoe:1994:CME:197600}, the failures semantics of any
+As shown in~\cite{Roscoe:1994:chapter}, the failures semantics of any
 finite-state CSP process $P$ can be represented by a \emph{normalised
@@
 the fact that $t$ is a function. The graph construction
-in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are
+in~\cite{Roscoe:1994:chapter} implies that all nodes $n$ in $N$ are
 reachable by sequences of edges labelled by $e_1\dots e_k$ and connecting
@@
 $$
-The graph construction from~\cite{Roscoe:1994:CME:197600} used to define
+The graph construction from~\cite{Roscoe:1994:chapter} used to define
 $G(P)$ for any process $P$ guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
@@
 
+% Commented out just to shorten the report.
 %To see that this approach works only for finite CSP processes, we consider
@@
 
+% Commented out just to shorten the report.
 %Well-formed normalised transition graphs must not refuse an event of the

### 2018-09-05 paper/preliminaries.tex
Proofs for Lemma 1 completed

@@
 
-In this section, we present CSP and the concept of minimal hitting set, which
+In this section, we present CSP and the concept of minimal hitting sets, which
 is central to our novel notion of test for failures refinement.
@@
 % =========================================================================
-\subsection{CSP and Refinement}
-
-%First, we introduce the CSP notation and refinement notions. We then discuss
-%a graph-based semantics for finite-state CSP processes.
-
-% =========================================================================
-%\subsubsection*{CSP}
-%\label{section:CSP}
+\subsection{CSP, Refinement, and Normalised Transition Graphs}
 
@@
 of interactions, modelled by synchronous, instantaneous, and atomic events.
+
+Throughout this paper, the alphabet of the processes, that is the set of
+events that are in scope, is denoted by $\Sigma$ and supposed to be finite.
+The FDR tool~\cite{fdr} supports model checking and semantic analyses of
+finite-state CSP processes.
 
@@
 external and internal choice operators allow us to define a choice over a
-collection of processes $P(i)$. By convention, , if the index set $I$ is
+collection of processes $P(i)$. By convention,  if the index set $I$ is
 empty, external choices evaluate to the process $\Stop$, which
 deadlocks:~does not engage into any event and does not terminate. An iterated
-internal choice is not defined for an empty index set.
+internal choice is not defined for an empty index set. For an external choice 
+indexed over a set of events in the form
+$\Extchoice e: A @ e \then P(e), A\subseteq \Sigma$, we use 
+abbreviation $e:A \then P(e)$.
 
@@
 containment $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
+$P \lessdet_F Q$ implies
 $(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. So, for
@@
 %
-Throughout this paper, the alphabet of the processes, that is the set of
-events that are in scope, is denoted by $\Sigma$ and supposed to be finite.
-%Since we are only
-%considering ``classical'' CSP processes, the alphabet is always finite.
-The FDR tool~\cite{fdr} supports model checking and semantic analyses of
-finite-state CSP processes.
 
@@
 %
-\begin{equation}
+\begin{equation}\label{eq:refMaxRef}
 \refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
@@
 has maximal refusals $\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials
-$[Q]^0=\Sigma$. In contrast to $P$, nondeterminism is reflected by
+$[Q]^0=\Sigma$. In contrast to $P$, nondeterminism is reflected here by
 two maximal refusals. \xbox
@@
 so $G(P)$ allows us to re-construct the failures semantics of $P$.
-\fixme{alcc: Why not leave the commented out material in the report?}
-
+%\fixme{alcc: Why not leave the commented out material in the report?}
 % Commented out just to shorten the report.
@@
 \\
- X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
+ X\in\accs(P/s) &\Leftrightarrow & \exists A\in \minaccs(P/s): A\subseteq X \subseteq [P/s]^0
  \label{eq:accsubset}
@@
 %
-Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
-(\ref{eq:accsubset}), every node of a normalised transition graph can
+Exploiting formula (\ref{eq:accref}), 
+%(\ref{eq:accinitials}), and (\ref{eq:accsubset}), 
+every node of a normalised transition graph can
 alternatively be labelled with their minimal acceptances, and this
@@
 \end{equation}
-\fixme{alcc: proof in the report perhaps?}
+%\fixme{alcc: proof in the report perhaps?}
 
@@
 %
-as established by results in the following lemma.
+as established by the results  in the following lemma.
+There, result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
+traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
+refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
+by means of the maximal refusal functions of the graphs involved; and
+(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
+can be derived from the maximal refusal functions by means of
+(\ref{eq:maxrefsminaccs}). 
+%%\fixme{alcc: I think at least the report should point to proofs.}
 %
@@
   \label{eq:failrefa}
-  P\ conf\ Q & \Leftrightarrow & \nonumber
+  Q\ conf\ P & \Leftrightarrow & \nonumber
   \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
@@
 %
-The result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
-traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
-refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
-by means of the maximal refusal functions of the graphs involved; and
-(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
-can be derived from the maximal refusal functions by means of
-(\ref{eq:maxrefsminaccs}). \fixme{alcc: I think at least the report should
-point to proofs.}
+\begin{proof}
+Recall that $P\lessdet_T Q$ is defined
+as $\trc(Q)\subseteq \trc(P)$ and that $\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$.
+This shows (\ref{eq:trcrefa}).
+
+\medskip
+\noindent
+To prove (\ref{eq:failconf}), we derive
+\begin{argue}
+P \lessdet_F Q 
+\\
+\Leftrightarrow \fails(Q)\subseteq \fails(Q) & definition of $\lessdet_F$
+\\
+\Leftrightarrow \forall (s,X) \in  \fails(Q): (s,X) \in\fails(P) 
+& property of $\subseteq$
+\\
+\Leftrightarrow \forall (s,\varnothing) \in  \fails(Q): (s,\varnothing) \in\fails(P)
+\wedge {}
+\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
+\\
+& Definition of failures, refusals, subset closure
+\\
+\Leftrightarrow \trc(Q)\subseteq\trc(P)
+\wedge {}
+\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
+\\
+& Definition of failures and traces
+\\
+\Leftrightarrow \trc(Q)\subseteq\trc(P)
+\wedge {}
+\\\tabc \forall s\in\trc(Q)\cap\trc(P): \refs(Q/s) \subseteq\refs(P/s)
+\\
+& Property of $\cap$ and $\subseteq$
+\\
+\Leftrightarrow
+P\lessdet_T Q \wedge Q\ conf\ P
+& Definition of $\lessdet_T$ and $conf$
+\end{argue}
+
+\medskip
+\noindent
+To prove (\ref{eq:failrefa}), we derive
+\begin{argue}
+Q\ conf\ P 
+\\
+ \Leftrightarrow \forall s\in\trc(P)\cap \trc(Q): \refs(Q/s)\subseteq\refs(P/s)
+ \\
+ & Definition of $conf$  (\ref{eq:conf})
+ \\
+ \Leftrightarrow \forall s\in\trc(G(P))\cap \trc(G(Q)): \refs(Q/s)\subseteq\refs(P/s)
+ \\
+ & $\trc(P) = \trc(G(P))$, $\trc(Q) = \trc(G(Q))$
+ \\
+\Leftrightarrow \forall s\in\trc(G(P))\cap \trc(G(Q)), R_Q\in r_Q(G(Q)/s):
+\\\tabc
+\exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
+& Property of $r_P, r_Q$ and (\ref{eq:refMaxRef})
+\end{argue}
+
+\medskip
+\noindent
+Finally, (\ref{eq:failrefb}) follows from (\ref{eq:failrefa}) using (\ref{eq:accref})
+and the fact that $R_Q \subseteq R_P$ is equivalent to 
+$\Sigma - R_P \subseteq \Sigma - R_Q$.
+\xbox
+\end{proof}
 

### 2018-09-06 paper/finitecomplete-failures.tex
New version of U_F test cases introduced. These are directly implementable with FDR4

@@
 
+% -------------------------------------------------------------------------
+\subsubsection*{Test Definition and Basic Properties}
 In the domain of process algebras, test cases are typically represented
@@
 
-For a given reference process $P$ and for each integer $p\ge 0$, we define a
+For a given reference process $P$, its normalised transition graph 
+$$
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+$$ 
+and for each integer $p\ge 0$, we define a
 CSP test process for failures refinement as shown below.
+%%
+%\begin{eqnarray}
+%U_F(p) & = & U_F(p,\varepsilon)
+%\\
+%U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
+%\label{eq:ufa}
+%\\ & & \extchoice \nonumber
+%\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
+%\label{eq:ufb}
+%\\ & & \extchoice \nonumber
+%\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
+%\label{eq:ufc}
+%\\ & & \extchoice \nonumber
+%\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber 
+%\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
+%\label{eq:ufd}
+%\end{eqnarray}
+%
 %
 \begin{eqnarray}
-U_F(p) & = & U_F(p,\varepsilon)
+U_F(p) & = & U_F(p,0,\ii n)
 \\
-U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
+U_F(p,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
 \label{eq:ufa}
 \\ & & \extchoice \nonumber
-\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
+\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
 \label{eq:ufb}
 \\ & & \extchoice \nonumber
-\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
+\\ & & (k < p) \& \big(e:[n]^0   \then U_F(p,k+1,t(n,e) \big)
 \label{eq:ufc}
 \\ & & \extchoice \nonumber
-\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber 
-\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
+\\ & & (k = p \wedge \minhits(n)\neq\varnothing) \&  \nonumber 
+\\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
 \label{eq:ufd}
@@
 %
+Before explaining the intuition behind this test definition in the paragraphs below,
+we state its basic properties in the following lemma, to show the relationships between
+$U_F(p)$ and the reference process $P$ from which the former has been derived.
+
+\begin{lemma}\label{lemma:ufproperties}
+If $s\in\trc(P)$ satisfies $\#s\le p$, then
+$s, s.e\in\trc(U_F(p))$ for all $e\in\Sigma$, and the following properties hold.
+\begin{eqnarray}
+\label{eq:ifpa}
+  &  & U_F(p)/s = U_F(p,\#s,G(P)/s)
+\\
+\label{eq:ifpb}
+e\not\in [P/s]^0 & \implies & U_F(p)/s.e = (\efail\then\Stop)
+\\
+\label{eq:ifpc}
+U_F(p)/s = U(p,\# s,n)  & \implies & [n]^0 = [P/s]^0  
+\\
+\label{eq:ifpd}
+U_F(p)/s = U(p,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
+\end{eqnarray}
+\end{lemma}
+\begin{proof}
+We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$, the statement holds because $U_F(p)$ starts with the initial node $\ii n$ of $G(P)$. Suppose that 
+the statement holds for all traces $s$ with length $\# s \le k < p$, so that
+$U_F(p)/s = U_F(p,\#s,G(P)/s)$. Now let $s.e$ be a trace of $P$. Then $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$, we conclude that $e\in  [G(P)/s]^0$,
+so $U_F(p)/s = U_F(p,\#s,G(P)/s)$ can engage into $e$ by executing branch (\ref{eq:ufc}).
+ Since $t$ is the transition function of $G(P)$ and $e\in  [G(P)/s]^0$, $t(G(P)/s,e)$
+ is defined, and $t(G(P)/s,e) = G(P)/s.e$ holds.
+This leads to a new recursion  
+$U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = U_F(p,\#s+1,G(P)/s.e)$ and proves the induction step.
+
+To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that 
+$U_F(p)/s = U_F(p,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
+$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so $U_F(p,\#s,G(P)/s)$
+can engage in $e$ by entering branch (\ref{eq:ufa}). The specification of this branch
+implies that 
+$$
+U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = (\efail\then\Stop).
+$$
+
+Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because 
+$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$.
+
+Finally, statement (\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because,
+according to (\ref{eq:minhitsGP}),
+$\minhits(G(P)/s) = \minhits(P/s)$ for all traces of $P$.
+\xbox
+\end{proof}
+
+Note that it is not guaranteed for $U_F(p)$ to run through the traces $s, s.e$ specified 
+in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for some prefix $u$ of $s$: in such a case, $U_F(p)$ may stop with a $\epass$-event by entering branch (\ref{eq:ufb}).
+Therefore, Lemma~\ref{lemma:ufproperties} just states the existence of
+$U_F(p)$-executions $s, s.e$ satisfying the properties stated there.
+
+% --------------------------------------------------------------------------
+\subsubsection*{Explanation of the Test Definition}
 A test is performed by running $U_F(p)$ concurrently with any SUT process
@@
 $U_F(p)/s$ is ready to engage in.
-
-The specification of $U_F(p,s)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
-If branch (\ref{eq:ufa})
-is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
-stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
-is the last to be entered, the process either accepts another event
-$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
-of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
-$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
-and the execution stops without a PASS or FAIL event. Such an execution is also interpreted as FAIL.
 
@@
 relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
-visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
+visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
 Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
 refuse $\epass$, due to a deadlock occurring when a faulty SUT process
-executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
-because $(\#s = p\wedge \minhits(P/s)\neq\varnothing)$. 
+executes concurrently with $U_F(p,k,n)$ executing branch (\ref{eq:ufd}),
+because $(k = p\wedge \minhits(P/s)\neq\varnothing)$. 
 This is explained further in the next paragraphs.
+
+The specification of $U_F(p,k,n)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
+If branch (\ref{eq:ufa})
+is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
+stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
+is the last to be entered, the process either accepts another event
+$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
+of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
+$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
+and the execution stops without a PASS or FAIL event.
+Such an execution is also interpreted as FAIL, because it reveals that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.
 
@@
 U_F(p)$ is finite, because the number of traces $s$ with maximal length
-$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
-$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
+$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and
+$\minhits(n)$ are finite. We further recall that $\minhits(n)$ may be
 empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
 is undefined. Therefore, the associated guard condition contains the requirement
-$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken.
-Recall that $\minhits(P/s)=\varnothing$ implies  $\minaccs(P/s) = \{\varnothing\}$,
-so branches (\ref{eq:ufa}) and (\ref{eq:ufb}) can be taken in this situation.
+$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken,
+but branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
 %%%\fixme{alcc: Oops. This is undefined in CSP.}
@@
 a length $p$. If, after having already run through $s\in\trc(P)$ with
-$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$,
+$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$
+ (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
 the test immediately terminates with FAIL-event $\efail$. This is handled by
@@
 If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
-this is revealed by $\varnothing$ being  its only minimal acceptance.
+this is revealed by $\minhits(n) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
+Lemma~\ref{lemma:ufproperties}). 
 In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
-the external choice in $U_F(p,s)$).
+the external choice in $U_F(p,s)$). If $P/s$ may also nondeterministically engage into events in such a situation, branches  (\ref{eq:ufa}) and (\ref{eq:ufc}) of the test are
+simultaneously enabled.
 
-If the length of $s$ is still less than $p$, the test accepts any event from
-the initials $[P/s]^0$ and continues recursively as $U_F(p,s.e)$ in
-branch~(\ref{eq:ufc}). A test of this type is called \emph{adaptive}, because
+If the length of $s$ is still less than $p$, the test accepts any event $e$ from
+the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as $U_F(p,\#s+1,G(P)/s.e)$ in
+branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}). 
+A test of this type is called \emph{adaptive}, because
 it accepts any legal behaviour of the SUT and adapts its consecutive
@@
 After having run through a trace of length $p$, and if
-$\minaccs(P/s)\neq\{\varnothing\}$ so that also
-$\minhits(P/s)\neq\varnothing$, the test changes its behaviour: instead of
+$\minhits(P/s)\neq \varnothing$, the test changes its behaviour: instead of
 offering {\it all} legal events from $[P/s]^0$ to the SUT, it
-nondeterministically chooses a minimal hitting set of $\minaccs(P/s)$ and
-only offers the events contained in this set. If the SUT refuses to engage
-into any of these events, this reveals a violation of failures refinement:
+nondeterministically chooses a minimal hitting set   $H\in \minhits(G(P)/s)$ 
+and
+only offers the events contained in $H$. If the SUT refuses to engage
+into any event of $H$, this reveals a violation of failures refinement:
 according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least

### 2018-09-06 paper/preliminaries.tex
New version of U_F test cases introduced. These are directly implementable with FDR4

@@
   \label{eq:failconf}
-  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge P\ conf\ Q
+  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
   \\
@@
 \\
-\Leftrightarrow \forall (s,X) \in  \fails(Q): (s,X) \in\fails(P) 
+\Leftrightarrow \forall (s,R) \in  \fails(Q): (s,R) \in\fails(P) 
 & property of $\subseteq$
@@
 \end{proof}
-\fixme{Abrupt end to the section. It has more than just preliminary material.
-It has some new results.}
+
+%\fixme{Abrupt end to the section. It has more than just preliminary material.
+%It has some new results.}
 
@@
 covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
-$\minhits(P/s) = \varnothing$.
+$\minhits(P/s) = \varnothing$. Trivially,
+\begin{equation}
+\label{eq:minhitminaccempty}
+\minaccs(P/s) = \{ \varnothing \} \iff \minhits(P/s) = \varnothing
+\end{equation}
+holds.
+
+Since minimal acceptances can be used to label the nodes of a normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$, the notation of minimal hitting sets also
+carries over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe that
+\begin{equation}
+\label{eq:minhitsGP}
+\minhits(G(P)/s) = \minhits(P/s) \ \text{for all}\ s\in \trc(P).
+\end{equation}

### 2018-09-06 paper/finitecomplete-failures.tex
Proofs re-worked for new U_F(p)

@@
 \begin{eqnarray}
+\label{eq:UFP}
 U_F(p) & = & U_F(p,0,\ii n)
@@
 A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
-if and only of the following conditions are fulfilled.
+if and only if the following conditions are fulfilled.
 \begin{description}
@@
 \[
-\TS_F = \{ U_F(k)~|~0 \le k < pq  \}
+\TS_F = \{ U_F(k)~|~0 \le k < pq  \}\quad\text{$U_F(k)$ specified in (\ref{eq:UFP})}
 \]
@@
 \end{lemma}
-\begin{proof}
-{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $P\ conf\
-Q$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
+\begin{proof}{~}\newline
+{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $Q\ conf\
+P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
 any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
 branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
-choice construction for $U_F(p,s)$. Branch~(\ref{eq:ufa}) can never be
-entered in the parallel execution of $Q$ and $U_F(p)$, because
-$[Q/s]^0\subseteq [P/s]^0$ for all traces of $Q$.
+choice construction for $U_F(p,k,n)$:  
+If $U_F(p,k,n) = U_F(p)/s$ with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties}
+implies $U_F(p,k,n) = U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As
+a consequence, $[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
+entered in the parallel execution of $Q$ and $U_F(p)$.
 
-{\bf Step~2.} Also,  Lemma~\ref{lemma:hseta} implies that for all traces
-$s\in\trc(Q)\cap\trc(P)$, every $H$ in $\minhits(P/s)$ is a hitting set for
-$\minaccs(Q/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,s)$ leads always to a
-PASS verdict, and branch (\ref{eq:ufc}) to test continuation without a
-verdict. For the last branch, we note that any selected minimal hitting set
-$H\in\minhits(P/s)$ has a non-empty intersection with each of the minimal
-acceptances of $Q/s$. As a consequence, $Q/s$ never blocks when offered
-events from $H$, and the test terminates with PASS event $\epass$. Note that
+\medskip
+\noindent
+{\bf Step~2.} Sinc $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies that for all  
+$s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a hitting set for
+$\minaccs(Q/s)$. 
+From Lemma~\ref{lemma:ufproperties} we know that
+$U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, 
+so $\minhits(n) = \minhits(P/s)$. 
+Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$ leads always to a
+PASS verdict and is taken if $\minhits(P/s) = \varnothing$. Branch (\ref{eq:ufc}) 
+always leads to test continuation without a
+verdict. For the last branch, we note that it can only be entered if $\minhits(P/s)
+= \minhits(n) \neq\varnothing$. In this case,
+any selected minimal hitting set
+$H\in\minhits(n)$ has a non-empty intersection with each of the minimal
+acceptances of $Q/s$. 
+As a consequence, $Q/s$ never blocks when offered
+events from $H\in\minhits(n)$, 
+and the test terminates with PASS event $\epass$. Note that
 this argument requires that $Q$ is free of livelocks, because otherwise the
 $\epass$-events might not become visible, due to unbounded sequences of
-hidden events performed by $Q$. \xbox
+hidden events performed by $Q$. 
+\xbox
 \end{proof}
@@
 (\ref{eq:failconf}), this non-conformance can be caused in two possible ways
-corresponding to the cases $P\not\lessdet_T Q$ and $\neg(P\ conf\ Q)$,
+corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$,
 respectively:
@@
 \end{description}
-It has to be shown for each of the two possibilities that at least one test
+It has to be shown for each of the two cases that at least one test
 execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
@@
 interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
-failures-refined by the test execution.
+failures-refined by $(Q\parallel[\Sigma] U_F(k))\hide \Sigma$.
 
-For Case~1, consider a  trace $s.e\in\trc(Q)$ such that $s\in\trc(P)$, but
+\medskip
+\noindent
+{\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$, but
 $s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is a
@@
 $U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a
-trace of this test, and, by construction, $U_F(pq-1)/u = U_F(pq-1,u)$. Since
-$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0$. So, in at least
-one execution, $U_F(pq-1,u)$ executes its first branch (\ref{eq:ufa}) with
+trace of this test, and, by construction and by Lemma~\ref{lemma:ufproperties}, 
+$U_F(pq-1)/u = U_F(pq-1,\#u,G(P)/u)$. Since
+$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. 
+Hence, in at least
+one execution, $U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with
 this event $e$, so that the test fails. Again, the
-assumption of non-divergence of Q is needed for this conclusion. %\fixme{alcc:
+assumption of non-divergence of Q is needed for this conclusion.
 
-For   Case~2, we note that trace $s$ is again a trace of the product graph
+\medskip
+\noindent
+{\bf Case~2.} We note that trace $s$ is again a trace of the product graph
 $G$, but we do not know its length. Again, by applying Lemma~\ref{lemma:reachproduc},
@@
 $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#
-u)$, which satisfies $U_F(\# u)/u = U_F(\#u,u)$, because it always performs
+u)$, which satisfies $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of Lemma~\ref{lemma:ufproperties}. This process   always performs
 branch (\ref{eq:ufc}) until the trace $u$ has been completely processed.
-$U_F(\#u,u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
+$U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
 only:~assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has
-at least one non-empty minimal acceptance, so the guard condition
-$(\minaccs(P/s) = \{ \varnothing \})$
-of branch (\ref{eq:ufb}) evaluates to $\isf$ for $U_F(\#u,u)$.
-Moreover, the guard condition $(\#s < p)$ for branch (\ref{eq:ufc}) evaluates
-to $\isf$ for $U_F(\#u,u)$, too. If branch (\ref{eq:ufa}) is executed, the
+at least one non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) 
+we know that this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$.
+As a consequence, branch (\ref{eq:ufb}) can never be taken because its guard condition
+evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$.
+Moreover, the guard condition $(k < p)$ for branch (\ref{eq:ufc}) evaluates
+to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the
 test always fails. If branch (\ref{eq:ufd}) is executed, the test fails for
 the execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen
-by $U_F(\#u,u)$ that has an empty intersection with the minimal acceptance
+by $U_F(\#u,\#u,G(P)/u)$ that has an empty intersection with the minimal acceptance
 $A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an
@@
 exists a test execution   where $Q/u$ selects acceptance $A_Q$ and
-$U_F(\#u,u)$ selects $H$. This execution deadlocks in process state
+$U_F(\#u,\#u,G(P)/u)$ selects $H$. This execution deadlocks in process state
 $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event;
-this  means that the test fails and concludes the proof. \xbox
+this  means that the test fails and concludes the proof. 
+\xbox
 \end{proof}

### 2018-09-06 paper/finitecomplete-failures.tex
Added theorem showing that maximal trace length pq is really ALWAYS required, for all p, q >= 2.

@@
 %
-Generalising Example~\ref{ex:pq}, it can be shown that for any pair $p,q\in\mathbb{N}$
-whose greatest common divisor is 1, there exist reference processes with $p$ states
+Generalising Example~\ref{ex:pq}, it can be shown that for any pair 
+$2\le p,q \in\mathbb{N}$,  
+there exist reference processes with $p$ states
 and implementation processes with $q$ states, such that non-conformance of the 
 implementation
-can only be detected with a trace of length $pq$.
+can only be detected with a trace of length $pq$. This is proven in the following 
+theorem.
+
+\begin{theorem}
+Let $2\le p,q \in\mathbb{N}$. Then there exists a reference process $P$ and an
+implementation process $Q$ with the following properties.
+\begin{enumerate}
+\item $G(P)$ has $p$ states.
+\item $G(Q)$ has $q$ states.
+\item $P\not\lessdet_T Q$.
+\item $\forall s\in\trc(Q): \#s < pq\implies s\in\trc(P)$.
+\end{enumerate}
+As a consequence, the upper bound $pq$ for the length of traces to be tested when checking for failures refinement cannot be reduced without losing the test suite's completeness property.
+\end{theorem}
+\begin{proof}
+Given $2\le p,q \in\mathbb{N}$, define reference process $P$ and implementation process $Q$
+over alphabet $\Sigma =\{ a,b\}$ as follows.
+\begin{eqnarray*}
+P & = &  P(0) 
+\\
+P(k) & = & \big( a \then P(k) \big) \extchoice (k < p-1) \& \big( b \then P(k+1)\big)
+\\ 
+Q & = & Q(0)
+\\
+Q(k) & = & (k < q-1) \& \big( a \then Q(k+1)    \big) \extchoice 
+( k = q-1)\& \big( a\then Q(0) \extchoice b\then Q(0)  \big)
+\end{eqnarray*}
+Using regular expression notation, the traces of $P$ can be specified as
+\[
+\trc(P) = \prefs\big(  (a^*b)^{p-1}a^* \big),
+\]
+where $\prefs(M)$ denotes the set of all prefixes of traces in $M\subset\Sigma^*$, 
+including the traces of $M$ themselves.
+The traces of $Q$ can be specified by
+\[
+\trc(Q) = \prefs\big( (a^{q-1}(a|b))^*  \big).
+\]
+It is easy to see that $\trc(Q)\not\subseteq\trc(P)$; 
+for example, the trace $(a^{q-1}b)^p$ is in
+$\trc(Q)\setminus\trc(P)$.
+
+Let $s \in\trc(Q)$ be any trace of length $\#s = pq-1$. Then $s$ can be represented 
+by $s = (a^{q-1}(a|b))^{p-1}a^{q-1} \in \prefs\big( (a^{q-1}(a|b))^*  \big)$. 
+Then $s$ is also an element of $\trc(P)$, because $(a^{q-1}(a|b))^{p-1}a^{q-1}$ 
+is also contained in $\prefs\big(  (a^*b)^{p-1}a^* \big)$.
+\xbox
+\end{proof}
 

### 2018-09-07 paper/casestudies.tex
Node number for transition graphs fixed, Example improved

@@
 
-For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, it is
-advisable to avoid an enumeration of traces $s$ of the reference process.
-Instead, we calculate the following auxiliary functions from $P$'s transition
-graph.
 %
+%For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, it is
+%advisable to avoid an enumeration of traces $s$ of the reference process.
+%Instead, we calculate the following auxiliary functions from $P$'s transition
+%graph.
+%%
+%\begin{eqnarray*}
+%\text{\it initials} & : & N \fun \power (\Sigma)
+%\\
+%\minhits & : & N \fun \power\power(\Sigma)
+%\end{eqnarray*}
+%%
+%As mentioned, in a state $n = G(P)/s$, the set $\text{\it initials}(n)$
+%equals the events labelling outgoing edges of $n$, so  $\text{\it
+%initials}(n) = [P/s]^0$. The function $\minhits$ maps $n$ to the set of all
+%minimal hitting sets associated with the minimal acceptances of $n$, so
+%$\minhits(n) = \minhits(P/s)$. Using the transition function $t$ and the
+%above functions, $U_F(p)$ can be re-written as the failures-equivalent CSP
+%process below.
+%%
+%\begin{eqnarray}
+%U_F^1(p) & = & U_F^1(p,0,\ii n)
+%\\
+%U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{\it initials}(n)) @ e \then \efail\then \Stop \big)
+%\label{eq:xufa}
+%\\ & & \extchoice \nonumber
+%\\ & & (\minhits(n) = \varnothing) \& \big( \epass \then \Stop \big)
+%\label{eq:xufb}
+%\\ & & \extchoice \nonumber
+%\\ & & (k < p) \& \big(\Extchoice e:\text{\it initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
+%\label{eq:xufc}
+%\\ & & \extchoice \nonumber
+%\\ & & (k = p) \& \big( \sqcap_{H\in\minhits(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
+%\label{eq:xufd}
+%\end{eqnarray}
+%\fixme{alcc: I think the example should be for the theory presented before,
+%not with a different notion of test. In particular, 33 is very different.}
+
+Generating the test cases $U_F(p)$ specified in  (\ref{eq:UFP}) for the reference
+process $P$ discussed in Example~\ref{example:CSP},
+results in the following instantiations of initials, minimal hitting sets, and
+transition function; this can be directly seen by analysing $P$'s normalised 
+transition graph with nodes $N =\{0,1,2,3\}$ as shown in Fig.~\ref{fig:tga}.
+
 \begin{eqnarray*}
-\text{\it initials} & : & N \fun \power (\Sigma)
-\\
-\minhits & : & N \fun \power\power(\Sigma)
+{ }[0]^0 & = & \{ a \} \\
+{ }[1]^0 & = & \{ a,b,c \} \\
+{ }[2]^0 & = & \{ a,b,c \} \\
+{ }[3]^0 & = & \{ b,c \} \\
+& & \\
+\minhits(0) & = & \{ \{ a\} \} \\
+\minhits(1) & = & \{ \{a,b\}, \{c\}, \} \\
+\minhits(2) & = & \{  \{a,b\},  \{a,c\},\} \\
+\minhits(3) & = & \{ \{ b\}, \{c\} \} \\
+& & \\
+t(0,a) & = & 1 \\
+t(1,a) & = & 0 \\
+t(1,b) & = & 0 \\
+t(1,c) & = & 2 \\
+t(2,a) & = & 1 \\
+t(2,b) & = & 0 \\
+t(2,c) & = & 3 \\
+t(3,b) & = & 0 \\
+t(3,c) &  =& 3
 \end{eqnarray*}
-%
-As mentioned, in a state $n = G(P)/s$, the set $\text{\it initials}(n)$
-equals the events labelling outgoing edges of $n$, so  $\text{\it
-initials}(n) = [P/s]^0$. The function $\minhits$ maps $n$ to the set of all
-minimal hitting sets associated with the minimal acceptances of $n$, so
-$\minhits(n) = \minhits(P/s)$. Using the transition function $t$ and the
-above functions, $U_F(p)$ can be re-written as the failures-equivalent CSP
-process below.
-%
-\begin{eqnarray}
-U_F^1(p) & = & U_F^1(p,0,\ii n)
-\\
-U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{\it initials}(n)) @ e \then \efail\then \Stop \big)
-\label{eq:xufa}
-\\ & & \extchoice \nonumber
-\\ & & (\minhits(n) = \varnothing) \& \big( \epass \then \Stop \big)
-\label{eq:xufb}
-\\ & & \extchoice \nonumber
-\\ & & (k < p) \& \big(\Extchoice e:\text{\it initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
-\label{eq:xufc}
-\\ & & \extchoice \nonumber
-\\ & & (k = p) \& \big( \sqcap_{H\in\minhits(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
-\label{eq:xufd}
-\end{eqnarray}
-\fixme{alcc: I think the example should be for the theory presented before,
-not with a different notion of test. In particular, 33 is very different.}
 
@@
  \begin{center}
-\includegraphics[width=\textwidth]{z_acc.pdf}
+\includegraphics[width=\textwidth]{z.pdf}
 \end{center}
@@
 fulfilled, however, $R_1(r_{max},k)$ makes an internal choice, instead of
-offering an external choice, and refinement does not hold.
+offering an external choice, so $P\not\lessdet_F Z$.
+Fig.~\ref{fig:tgZ} shows the normalised transition graph of $Z$ for $r_{max}
+= 3$.
+
+%%%and refinement does not hold.
 % alcc: I don't think there is a notion of refinement for state?
@@
 %$P$ in the failures model.
-Fig.~\ref{fig:tgZ} shows the normalised transition graph of $Z$ for $r_{max}
-= 3$.
 
-Running the test $U_F^1(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p =
+Running the test $U_F(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p =
 4$ states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test
 depth to be used according to Theorem~\ref{th:failurestest}), tests
-$U_F^1(0),\dots, U_F^1(3)$ are passed by $Z$, but $Z$ fails $U_F^1(4)$,
+$U_F(0),\dots, U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$,
 because after  execution trace
 \[
-s = a.c.c.c, \qquad\text{(note that $G(Z)/s = \text{Node\_3}$ according to Fig.~\ref{fig:tgZ})},
+s = a.c.c.c, \qquad\text{(note that $G(P)/s = \text{node}\ 3$ according to Fig.~\ref{fig:tga})},
 \]
-it may be the case that $Z$ accepts only $\{b\}$ due to the internal choice
-and $U_F^1(4)$ -- also due to internal choice -- accepts only the minimal
-hitting set $\{ c \}$ and the event $a\in (\Sigma - [P/s]^0)$. So,
-$(Z\parallel[\Sigma] U_F^1(4))/s$ deadlocks, and the $pass$ event cannot be
+the test $U_F(4)$ offers hitting sets from $\minhits(3) = \{\{b\},\{c\}\}$
+in branch (\ref{eq:ufd}). Therefore,  there exists one test execution
+where $Z/s$ accepts only $\{b\}$ due to the internal choice (note 
+from Fig.~\ref{fig:tgZ} that
+$G(Z)/s = 4$), while $U_F(4)/s$
+only offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for 
+branch (\ref{eq:ufa}).  As a consequence, this execution of 
+$(Z\parallel[\Sigma] U_F(4))/s$ deadlocks, and the $pass$ event cannot be
 produced. Another failing execution arises if $Z/s$ chooses to accept only
-$\{c \}$, while $U_F^1(4)/s$ choses to accept only $\{a,b\}$. Therefore,
+$\{c \}$, while $U_F(4)/s$ choses to accept only $\{a,b\}$. Therefore,
 \[
-(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F^1(4)),
+(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F(4)),
 \]
@@
 \end{example}
-\fixme{alcc: I would write down the actual test generated.}
+%%%\fixme{alcc: I would write down the actual test generated.}
 

### 2018-09-07 paper/finitecomplete-failures.tex
Node number for transition graphs fixed, Example improved

@@
 there exist reference processes with $p$ states
-and implementation processes with $q$ states, such that non-conformance of the 
-implementation
+and implementation processes with $q$ states, such that 
+a violation of the trace refinement property
 can only be detected with a trace of length $pq$. This is proven in the following 
@@
 Then $s$ is also an element of $\trc(P)$, because $(a^{q-1}(a|b))^{p-1}a^{q-1}$ 
-is also contained in $\prefs\big(  (a^*b)^{p-1}a^* \big)$.
+is also contained in $\prefs\big(  (a^*b)^{p-1}a^* \big)$: this is easy to see, since
+$\prefs\big(  (a^*b)^{p-1}a^* \big)$ contains all finite sequences of $a$-events, 
+where at most $p-1$ events $b$ have been inserted.
 \xbox

### 2018-09-07 paper/preliminaries.tex
Proofs re-worked for new U_F(p)
Node number for transition graphs fixed, Example improved

@@
    \begin{center}
-     \includegraphics[width=\textwidth]{q0.pdf}
+     \includegraphics[width=\textwidth]{p.pdf}
    \end{center}
@@
 $P/\varepsilon$ (where $\varepsilon$ denotes the empty trace) is represented
-as Node\_0, with $\{ a\}$ as the only minimal acceptance, since $a$ is never
+as node 0, with $\{ a\}$ as the only minimal acceptance, since $a$ is never
 refused and no other events are accepted. Having engaged in $a$, the
-transition from Node\_0 leads to Node\_2 representing the process state $P/a
+transition from node 0 leads to node 1 representing the process state $P/a
 = Q\intchoice R$. The internal choice induces several minimal acceptances
@@
 
-Having engaged in $c$, the next process state is represented by Node\_1. Due
+Having engaged in $c$, the next process state is represented by node 2. Due
 to normalisation, there is only a single transition satisfying
-$t(\text{Node\_2},c) = \text{Node\_1}$. This transition, however, can have
-been caused by either $Q$ or $R$ engaging into $c$, so Node\_1 corresponds to
+$t(1,c) = 2$. This transition, however, can have
+been caused by either $Q$ or $R$ engaging into $c$, so node 2 corresponds to
 process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
-two minimal acceptance sets labelling Node\_1. Similar considerations explain
+two minimal acceptance sets labelling node 2. Similar considerations explain
 the other nodes and transitions in Fig.~\ref{fig:tga}. 
-
-Note that the node names including their number suffixes are generated by the
-FDR tool. The numbering is generated during the normalisation procedure. So,
-the node numbers do not reflect the distance from the initial node Node\_0.
 \xbox
@@
 \begin{enumerate}
-\item $P\ conf\ Q$
+\item $Q\ conf\ P$
 
@@
 \begin{proof}
-For showing ``$1 \Rightarrow 2$'', assume   $P\ conf\ Q$ and $s\in\trc(P)\cap
+For showing ``$1 \Rightarrow 2$'', assume   $Q\ conf\ P$ and $s\in\trc(P)\cap
 \trc(Q)$. Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}), states that

### 2018-09-09 paper/complexity.tex
Technical report completed, taking into account the remarks from Ana

--- 2018-09-07 paper/finitecomplete-failures.tex
@@
-% ==========================================================================
-\section{Finite Complete Test Suites for CSP Failures Refinement}
-\label{sec:finitecompletefails}
-% ==========================================================================
+\section{Complexity Considerations}
+\label{sec:complexity}
+% ================================================================================
 
-Here, we define our notion of tests for failures refinement, and then prove
-completeness of our suite. Finally, we study to complexity of our approach by
-identifying a bound on the number of tests we need in a complete suite.
-
-% ==========================================================================
-\subsection{Test Cases for Verifying CSP Failures Refinement}
-
-% -------------------------------------------------------------------------
-\subsubsection*{Test Definition and Basic Properties}
-In the domain of process algebras, test cases are typically represented
-by processes interacting concurrently with the SUT
-process~\cite{Hennessy:1988:ATP:50497}.
-They synchronise with the SUT over its visible events and use some additional events outside the SUT's alphabet to express whether the test execution  passed or failed.
-%%%, or if no verdict could be obtained.
-
-For a given reference process $P$, its normalised transition graph 
-$$
-G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
-$$ 
-and for each integer $p\ge 0$, we define a
-CSP test process for failures refinement as shown below.
-%%
-%\begin{eqnarray}
-%U_F(p) & = & U_F(p,\varepsilon)
-%\\
-%U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
-%\label{eq:ufa}
-%\\ & & \extchoice \nonumber
-%\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
-%\label{eq:ufb}
-%\\ & & \extchoice \nonumber
-%\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
-%\label{eq:ufc}
-%\\ & & \extchoice \nonumber
-%\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber 
-%\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
-%\label{eq:ufd}
-%\end{eqnarray}
-%
-%
-\begin{eqnarray}
-\label{eq:UFP}
-U_F(p) & = & U_F(p,0,\ii n)
-\\
-U_F(p,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
-\label{eq:ufa}
-\\ & & \extchoice \nonumber
-\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
-\label{eq:ufb}
-\\ & & \extchoice \nonumber
-\\ & & (k < p) \& \big(e:[n]^0   \then U_F(p,k+1,t(n,e) \big)
-\label{eq:ufc}
-\\ & & \extchoice \nonumber
-\\ & & (k = p \wedge \minhits(n)\neq\varnothing) \&  \nonumber 
-\\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
-\label{eq:ufd}
-\end{eqnarray}
-%
-Before explaining the intuition behind this test definition in the paragraphs below,
-we state its basic properties in the following lemma, to show the relationships between
-$U_F(p)$ and the reference process $P$ from which the former has been derived.
-
-\begin{lemma}\label{lemma:ufproperties}
-If $s\in\trc(P)$ satisfies $\#s\le p$, then
-$s, s.e\in\trc(U_F(p))$ for all $e\in\Sigma$, and the following properties hold.
-\begin{eqnarray}
-\label{eq:ifpa}
-  &  & U_F(p)/s = U_F(p,\#s,G(P)/s)
-\\
-\label{eq:ifpb}
-e\not\in [P/s]^0 & \implies & U_F(p)/s.e = (\efail\then\Stop)
-\\
-\label{eq:ifpc}
-U_F(p)/s = U(p,\# s,n)  & \implies & [n]^0 = [P/s]^0  
-\\
-\label{eq:ifpd}
-U_F(p)/s = U(p,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
-\end{eqnarray}
-\end{lemma}
-\begin{proof}
-We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$, the statement holds because $U_F(p)$ starts with the initial node $\ii n$ of $G(P)$. Suppose that 
-the statement holds for all traces $s$ with length $\# s \le k < p$, so that
-$U_F(p)/s = U_F(p,\#s,G(P)/s)$. Now let $s.e$ be a trace of $P$. Then $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$, we conclude that $e\in  [G(P)/s]^0$,
-so $U_F(p)/s = U_F(p,\#s,G(P)/s)$ can engage into $e$ by executing branch (\ref{eq:ufc}).
- Since $t$ is the transition function of $G(P)$ and $e\in  [G(P)/s]^0$, $t(G(P)/s,e)$
- is defined, and $t(G(P)/s,e) = G(P)/s.e$ holds.
-This leads to a new recursion  
-$U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = U_F(p,\#s+1,G(P)/s.e)$ and proves the induction step.
-
-To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that 
-$U_F(p)/s = U_F(p,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
-$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so $U_F(p,\#s,G(P)/s)$
-can engage in $e$ by entering branch (\ref{eq:ufa}). The specification of this branch
-implies that 
-$$
-U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = (\efail\then\Stop).
-$$
-
-Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because 
-$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$.
-
-Finally, statement (\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because,
-according to (\ref{eq:minhitsGP}),
-$\minhits(G(P)/s) = \minhits(P/s)$ for all traces of $P$.
-\xbox
-\end{proof}
-
-Note that it is not guaranteed for $U_F(p)$ to run through the traces $s, s.e$ specified 
-in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for some prefix $u$ of $s$: in such a case, $U_F(p)$ may stop with a $\epass$-event by entering branch (\ref{eq:ufb}).
-Therefore, Lemma~\ref{lemma:ufproperties} just states the existence of
-$U_F(p)$-executions $s, s.e$ satisfying the properties stated there.
-
-% --------------------------------------------------------------------------
-\subsubsection*{Explanation of the Test Definition}
-A test is performed by running $U_F(p)$ concurrently with any SUT process
-$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
-the concurrent process
-\[
-Q\parallel[\Sigma] U_F(p).
-\]
-It is assumed that the events $\efail$ and $\epass$, indicating a verdict
-FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
-assume that $Q$ is free of livelocks, it is guaranteed that events 
-$\efail$ or $\epass$ always become visible, if they are the only events
-$U_F(p)/s$ is ready to engage in.
-
-The test is
-\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
-every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
-$\epass$. This can also be  expressed by means of a failures refinement as
-defined below.
-%
-\begin{equation}
-\label{eq:passF}
-Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
-\end{equation}
-%
-This type of pass relation is often called \emph{must test}, because every
-test execution must end with the $\epass$
-event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
-failures-refinement relation in this condition, and not the trace-refinement
-relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
-visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
-Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
-refuse $\epass$, due to a deadlock occurring when a faulty SUT process
-executes concurrently with $U_F(p,k,n)$ executing branch (\ref{eq:ufd}),
-because $(k = p\wedge \minhits(P/s)\neq\varnothing)$. 
-This is explained further in the next paragraphs.
-
-The specification of $U_F(p,k,n)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
-If branch (\ref{eq:ufa})
-is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
-stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
-is the last to be entered, the process either accepts another event
-$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
-of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
-$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
-and the execution stops without a PASS or FAIL event.
-Such an execution is also interpreted as FAIL, because it reveals that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.
-
-We observe that the number of possible executions of $Q\parallel[\Sigma]
-U_F(p)$ is finite, because the number of traces $s$ with maximal length
-$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and
-$\minhits(n)$ are finite. We further recall that $\minhits(n)$ may be
-empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
-is undefined. Therefore, the associated guard condition contains the requirement
-$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken,
-but branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
-%%%\fixme{alcc: Oops. This is undefined in CSP.}
-
-Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
-a length $p$. If, after having already run through $s\in\trc(P)$ with
-$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$
- (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
-the test immediately terminates with FAIL-event $\efail$. This is handled by
-the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,s)$
-defined above.
-
-If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
-this is revealed by $\minhits(n) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
-Lemma~\ref{lemma:ufproperties}). 
-In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
-the external choice in $U_F(p,s)$). If $P/s$ may also nondeterministically engage into events in such a situation, branches  (\ref{eq:ufa}) and (\ref{eq:ufc}) of the test are
-simultaneously enabled.
-
-If the length of $s$ is still less than $p$, the test accepts any event $e$ from
-the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as $U_F(p,\#s+1,G(P)/s.e)$ in
-branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}). 
-A test of this type is called \emph{adaptive}, because
-it accepts any legal behaviour of the SUT and adapts its consecutive
-behaviour to the event selected by the SUT.
-
-After having run through a trace of length $p$, and if
-$\minhits(P/s)\neq \varnothing$, the test changes its behaviour: instead of
-offering {\it all} legal events from $[P/s]^0$ to the SUT, it
-nondeterministically chooses a minimal hitting set   $H\in \minhits(G(P)/s)$ 
-and
-only offers the events contained in $H$. If the SUT refuses to engage
-into any event of $H$, this reveals a violation of failures refinement:
-according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
-one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
-only terminates with  $\epass$, if such an event is accepted by the SUT.
-
-%% =========================================================================
-%\subsection{Fault Models and Complete Test Suites}
-
-% =========================================================================
-\subsection{A Finite Complete Test Suite for Failures Refinement}
-
-A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a
-reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
-\lessdet_F\}$, and a fault domain ${\cal D}$ which is a set of CSP processes
-over $P$'s alphabet that may or may not conform to $P$.
-
-A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
-if and only if the following conditions are fulfilled.
-\begin{description}
-\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
-\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
-then $Q$ fails at least one test in $\TS$.
-\end{description}
-%
-In the sequel, we establish the completeness of our test suite which is stated 
-in the following main theorem.
-%%%\fixme{alcc: I don't think we need a section just for this material.}
-
-\begin{theorem}\label{th:failurestest}
-Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
-normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
-D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
-whose transition graph has at most $q$ states with $q \ge p$. Then the test
-suite
-\[
-\TS_F = \{ U_F(k)~|~0 \le k < pq  \}\quad\text{$U_F(k)$ specified in (\ref{eq:UFP})}
-\]
-is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
-\end{theorem}
-%
-The proof of the theorem follows from the two lemmas below. The first states
-that test suite $\TS_F$ is sound, the second states that the suite is also
-exhaustive.
-%
-\begin{lemma}\label{lemma:mainfsound}
-Test suite $\TS_F$ generated from a CSP process $P$, as specified in
-Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
-$P\lessdet_F Q$.
-\end{lemma}
-\begin{proof}{~}\newline
-{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $Q\ conf\
-P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
-any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
-branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
-choice construction for $U_F(p,k,n)$:  
-If $U_F(p,k,n) = U_F(p)/s$ with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties}
-implies $U_F(p,k,n) = U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As
-a consequence, $[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
-entered in the parallel execution of $Q$ and $U_F(p)$.
-
-\medskip
-\noindent
-{\bf Step~2.} Sinc $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies that for all  
-$s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a hitting set for
-$\minaccs(Q/s)$. 
-From Lemma~\ref{lemma:ufproperties} we know that
-$U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, 
-so $\minhits(n) = \minhits(P/s)$. 
-Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$ leads always to a
-PASS verdict and is taken if $\minhits(P/s) = \varnothing$. Branch (\ref{eq:ufc}) 
-always leads to test continuation without a
-verdict. For the last branch, we note that it can only be entered if $\minhits(P/s)
-= \minhits(n) \neq\varnothing$. In this case,
-any selected minimal hitting set
-$H\in\minhits(n)$ has a non-empty intersection with each of the minimal
-acceptances of $Q/s$. 
-As a consequence, $Q/s$ never blocks when offered
-events from $H\in\minhits(n)$, 
-and the test terminates with PASS event $\epass$. Note that
-this argument requires that $Q$ is free of livelocks, because otherwise the
-$\epass$-events might not become visible, due to unbounded sequences of
-hidden events performed by $Q$. 
-\xbox
-\end{proof}
-%
-\begin{lemma}\label{lemma:mainfexhaustive}
-Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive
-for the fault model specified there.
-\end{lemma}
-\begin{proof}
-Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$, According to
-(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
-corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$,
-respectively:
-\begin{description}
-\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
-\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
-of $\minaccs(Q/s)$, such that
-(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
-\begin{equation}
-\label{eq:accsnotcontained}
-\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
-\end{equation}
-\end{description}
-It has to be shown for each of the two cases that at least one test
-execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
-   $\efail$-event or without giving any verdict. The latter case is also
-interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
-failures-refined by $(Q\parallel[\Sigma] U_F(k))\hide \Sigma$.
-
-\medskip
-\noindent
-{\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$, but
-$s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is a
-trace of every process. In this case, $s$ is also a trace of the product
-graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}. Suppose that
-$G/s = (n_P,n_Q)$. The length of $s$ is not known, but from the construction
-of $G$,  we know that $G$ has at most $pq$ reachable states, because $G(P)$
-has $p$ states, and $G(Q)$ has at most $q$ states. By
-Lemma~\ref{lemma:reachproduc}, $(n_P,n_Q)$ can be reached by a trace
-$u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
-function of $G$ implies that $u$ is also a trace of $P$ and $Q$. Since test
-$U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a
-trace of this test, and, by construction and by Lemma~\ref{lemma:ufproperties}, 
-$U_F(pq-1)/u = U_F(pq-1,\#u,G(P)/u)$. Since
-$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. 
-Hence, in at least
-one execution, $U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with
-this event $e$, so that the test fails. Again, the
-assumption of non-divergence of Q is needed for this conclusion.
-
-\medskip
-\noindent
-{\bf Case~2.} We note that trace $s$ is again a trace of the product graph
-$G$, but we do not know its length. Again, by applying Lemma~\ref{lemma:reachproduc},
-we know that the state   $G/s$ can   be reached by a trace
-$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#
-u)$, which satisfies $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of Lemma~\ref{lemma:ufproperties}. This process   always performs
-branch (\ref{eq:ufc}) until the trace $u$ has been completely processed.
-$U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
-only:~assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has
-at least one non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) 
-we know that this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$.
-As a consequence, branch (\ref{eq:ufb}) can never be taken because its guard condition
-evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$.
-Moreover, the guard condition $(k < p)$ for branch (\ref{eq:ufc}) evaluates
-to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the
-test always fails. If branch (\ref{eq:ufd}) is executed, the test fails for
-the execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen
-by $U_F(\#u,\#u,G(P)/u)$ that has an empty intersection with the minimal acceptance
-$A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an
-$H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there
-exists a test execution   where $Q/u$ selects acceptance $A_Q$ and
-$U_F(\#u,\#u,G(P)/u)$ selects $H$. This execution deadlocks in process state
-$(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event;
-this  means that the test fails and concludes the proof. 
-\xbox
-\end{proof}
-
-% ==========================================================================
-\subsection{Complexity Considerations}
-\label{sec:complexity}
-
-As can be seen from the specification of the test cases $U_F(k)$, the number of
+As can be seen from the specification of the test cases $U_F(k)$ 
+(see Section~\ref{sec:finitecompletefails}), the number of
 executions ending in a $\epass$-event corresponds to the number $\ell$ of traces $s$
@@
 multiplied by the number $h$ of minimal hitting sets in
-$\minhits(P/s)$.
+$\minhits(P/s)$. For the tests $U_T(k)$ verifying trace refinement (see Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$, since
+there is no equivalent in $U_T(k)$ 
+to checking different hitting sets in the last step of a 
+test execution.
 
 % -------------------------------------------------------------------------
-\subsubsection*{Estimation of $\ell$}
+\subsection{Estimation of $\ell$}
 The first factor $\ell$
@@
 \[
-RUN(\Sigma) = e:\Sigma \then RUN(\Sigma)
+RUN(\Sigma) = e:\Sigma \then RUN(\Sigma).
 \]
@@
 % -------------------------------------------------------------------------
-\subsubsection*{Estimation of $h$}
+\subsection{Estimation of $h$}
 Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
@@
 
-From Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
+% -------------------------------------------------------------------------
+\subsection{Upper Bounds of Test Executions for Checking Failures Refinement}
+According to  Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
 $k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
@@
 % -------------------------------------------------------------------------
-\subsubsection*{Upper Bound $pq$ for Number of Test Cases}
+\subsection{Upper Bounds of Test Executions for Checking Trace Refinement}
+
+According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
+refinement just contains the adaptive test case $U_T(pq-1)$. As derived for $U_F(k)$
+above, the number of executions performed by $(Q\parallel[\Sigma] U_R(pq-1))$ 
+is bounded by $|\Sigma|^{pq-1}$.
+
+% -------------------------------------------------------------------------
+\subsection{Upper Bound $pq$ for the Maximal Length of Test Traces}
 
@@
 $k = 0,\dots,pq-1$ to guarantee completeness. This means that the SUT is verified 
-with test traces up to, and including, length $pq$. It is interesting to  investigate
+with test traces up to, and including, length $pq$:
+recall from the test specification, branch (\ref{eq:ufa}), that $U_F(k)$ will accept
+all traces $s.e$ with $s\in\trc(P), \#s = k, e\not\in\trc(P/s)$, so erroneous trace
+up to length $k+1$ are detected.
+It is interesting to  investigate
 whether this maximal length is really necessary, or whether one could elaborate
@@
 
-\begin{theorem}
+\begin{theorem}\label{th:maxtracelen}
 Let $2\le p,q \in\mathbb{N}$. Then there exists a reference process $P$ and an
@@
 
+Since Theorem~\ref{th:maxtracelen} just states that a violation of trace refinement
+may remain undetected if only traces shorter than $pq$ are checked during tests, 
+it can also be applied to our trace refinement tests. Therefore, 
+test suites $\{ U_T(z) \}$ with $z<pq-1$ are no longer complete.
+
 It is discussed in Section~\ref{sec:conc} how the number of test traces to be executed 
-by complete test suites for failures refinement   can still be reduced {\it without} reducing 
+by complete test suites for failures or tarce 
+refinement   can still be reduced {\it without} reducing 
 the maximal length.
 
-% ==========================================================================
+% =================================================================================

### 2018-09-09 paper/conclusion.tex
Technical report completed, taking into account the remarks from Ana

@@
 otherwise its behaviour is considered to be blocked and regarded as a
-failure. \fixme{alcc: if deadlock is not acceptable? I think you need
-instrumentation here as well. }
+failure. 
+%\fixme{alcc: if deadlock is not acceptable? I think you need
+%instrumentation here as well. }
+%% jp : I do not understand the problem - sorry !
 
@@
 tests. It would be possible, for instance, to use the bound indicated here.
-Morevoer, specifying a fault domain as a CSP process allows us to model
+Moreover, specifying a fault domain as a CSP process allows us to model
 domain-specific knowledge using CSP. For example, if an initialisation

### 2018-09-09 paper/finitecomplete-failures.tex
Technical report completed, taking into account the remarks from Ana

--- 2018-09-06 paper/finitecomplete-failures.tex
@@
 $\efail$ or $\epass$ always become visible, if they are the only events
-$U_F(p)/s$ is ready to engage in.
+$U_F(p)/s$ is ready to engage in: if $U_F(p)/s$ can only produce $\epass$ or
+$\efail$, the occurrence of these events can never be blocked due to a livelock in $Q$
+occurring in the same step of the execution.
 
@@
 executes concurrently with $U_F(p,k,n)$ executing branch (\ref{eq:ufd}),
-because $(k = p\wedge \minhits(P/s)\neq\varnothing)$. 
+when guard condition $(k = p\wedge \minhits(n)\neq\varnothing)$ evaluates to $\ist$. 
 This is explained further in the next paragraphs.
@@
 is the last to be entered, the process either accepts another event
-$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
+$e$ of some minimal hitting set $H\in\minhits(n)$ with $n = G(P)/s$ according to
+Lemma~\ref{lemma:ufproperties}. Then the final length
 of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
@@
 is undefined. Therefore, the associated guard condition contains the requirement
-$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken,
+$\minhits(n)\neq\varnothing$, otherwise this branch cannot be taken,
 but branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
@@
 the test immediately terminates with FAIL-event $\efail$. This is handled by
-the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,s)$
+the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,\#s,n)$
 defined above.
@@
 If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
-this is revealed by $\minhits(n) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
+this is revealed by $\minhits(G(P)/s) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
 Lemma~\ref{lemma:ufproperties}). 
 In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
-the external choice in $U_F(p,s)$). If $P/s$ may also nondeterministically engage into events in such a situation, branches  (\ref{eq:ufa}) and (\ref{eq:ufc}) of the test are
-simultaneously enabled.
+the external choice in $U_F(p,\#s,G(P)/s)$). 
+If $P/s$ may also nondeterministically engage into events in such a situation, branche 
+ (\ref{eq:ufc}) of the test is
+simultaneously enabled. If $Q/s$ is able to engage into an event from 
+$\Sigma - [P/s]^0$, a test execution exists where $U_F(p,\#s,G(P)/s)$ branches
+into (\ref{eq:ifpb}) and produces the $\efail$-event.
 
@@
 the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as $U_F(p,\#s+1,G(P)/s.e)$ in
-branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}). 
+branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}. 
 A test of this type is called \emph{adaptive}, because
@@
 
-After having run through a trace of length $p$, and if
-$\minhits(P/s)\neq \varnothing$, the test changes its behaviour: instead of
-offering {\it all} legal events from $[P/s]^0$ to the SUT, it
-nondeterministically chooses a minimal hitting set   $H\in \minhits(G(P)/s)$ 
+Suppose that a test execution has run through a trace 
+$s\in\trc(P)$ of length $p$, so that $U_F(p)/s = U_F(p,p,n)$ with $n = G(P)/s$. 
+Then,   if
+$\minhits(n)\neq \varnothing$, the test changes its behaviour: instead of
+offering {\it all} legal events from $[n]^0$ to the SUT, it
+nondeterministically chooses a minimal hitting set   $H\in \minhits(n)$ 
 and
@@
 according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
-one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
+one event of each minimal hitting set in $\minhits(n)$. Therefore, the test
+execution
 only terminates with  $\epass$, if such an event is accepted by the SUT.
@@
 \begin{theorem}\label{th:failurestest}
-Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
+Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
 normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
@@
 Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$ leads always to a
-PASS verdict and is taken if $\minhits(P/s) = \varnothing$. Branch (\ref{eq:ufc}) 
+PASS verdict and is taken if $\minhits(n) = \varnothing$. Branch (\ref{eq:ufc}) 
 always leads to test continuation without a
@@
 trace of every process. In this case, $s$ is also a trace of the product
-graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}. Suppose that
-$G/s = (n_P,n_Q)$. The length of $s$ is not known, but from the construction
+graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}, and  
+$G/s = (G(P)/s,G(Q)/s)$ holds. The length of $s$ is not known, but from the construction
 of $G$,  we know that $G$ has at most $pq$ reachable states, because $G(P)$
 has $p$ states, and $G(Q)$ has at most $q$ states. By
-Lemma~\ref{lemma:reachproduc}, $(n_P,n_Q)$ can be reached by a trace
+Lemma~\ref{lemma:reachproduc}, $(G(P)/s,G(Q)/s)$ can be reached by a trace
 $u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
-function of $G$ implies that $u$ is also a trace of $P$ and $Q$. Since test
+function of $G$ implies that $u$ is also a trace of $P$ and $Q$, 
+which means that 
+$(G(P)/s,G(Q)/s) = (G(P)/u,G(Q)/u)$. Since test
 $U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a
@@
 $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#
-u)$, which satisfies $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of Lemma~\ref{lemma:ufproperties}. This process   always performs
+u)$, which satisfies $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$   always performs
 branch (\ref{eq:ufc}) until the trace $u$ has been completely processed.
@@
 at least one non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) 
-we know that this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$.
-As a consequence, branch (\ref{eq:ufb}) can never be taken because its guard condition
+we know that this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and
+we observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$.
+As a consequence, branch (\ref{eq:ufb}) cannot be taken because its guard condition
 evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$.
-Moreover, the guard condition $(k < p)$ for branch (\ref{eq:ufc}) evaluates
+The guard condition $(k < p)$ for branch (\ref{eq:ufc}) evaluates
 to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the
@@
 % ==========================================================================
-\subsection{Complexity Considerations}
-\label{sec:complexity}
-
-As can be seen from the specification of the test cases $U_F(k)$, the number of
-executions ending in a $\epass$-event corresponds to the number $\ell$ of traces $s$
-of $P$ with length equal to\footnote{In this estimation, we disregard
-the case where the test
-terminates earlier due to entering branch (\ref{eq:ufb}).} $k$,
-multiplied by the number $h$ of minimal hitting sets in
-$\minhits(P/s)$.
-
-% -------------------------------------------------------------------------
-\subsubsection*{Estimation of $\ell$}
-The first factor $\ell$
-has worst-case upper bound $\ell\le |\Sigma|^k$. As an example, where this upper bound 
-is really met, consider the reference process
-\[
-RUN(\Sigma) = e:\Sigma \then RUN(\Sigma)
-\]
-The normalised transition graph of this process has a single state, and its initials
-are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(k)$ can 
-never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly 
-$|\Sigma|^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
-for each of their events.
-
-% -------------------------------------------------------------------------
-\subsubsection*{Estimation of $h$}
-Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
-acceptances, the cardinality $h$ of $\minhits(P/s)$ is maximal for the case
-where all $A_i$ are disjoint. In this case,  $h~~=~~|\minhits(P/s)|~~=~~
-\prod_{i=1}^\alpha|A_i|$. To find an upper bound for $h$, we use the result
-that we prove in the following lemma.
-%
-\begin{lemma}\label{lemma:minhitmax}
-Let $n\ge 2$ be any positive integer. We call a sequence
-%
-$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$
-%
-a partition of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$. Define
-%
-$$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
-Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
-\alpha\in \mathbb{N}\big\}.$$
-%
-\end{lemma}
-\begin{proof}
-Let $n\ge 2$. Obviously, \begin{equation}\label{eq:1} n\le
-p(n).\end{equation} Define $\pi(n)$ to be the following partition  of $n$:
-%
-$$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
-%
-Then the product of the numbers in $\pi(n)$ equals  $p(n)$. In the remainder
-of the proof we have to show that $p(n)$ is really maximal in the sense of
-the lemma.
-
-Since the number of partitions of a given $n$ is finite,
-%
-$$\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$$
-%
-exists. Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with
-maximal product. Then, obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
-Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then
-from~(\ref{eq:1}) we have
-$$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
-Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a
-product value which is not less then the original partition $a_1, \ldots ,
-a_{\alpha}$. This process can be repeated, until every $a_i$ in the resulting
-partition is either 2 or 3. Hence there exists a partition of $n$ with
-maximal product, such that every term is $2$ or $3$. Since $6=2+2+2=3 + 3$
-and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than
-$3$. Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\},
-i=1,\dots,\alpha$, be a partition of $n$ with maximal product. Let $k_1\le 2$
-be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
-Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$
-with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have
-
-\begin{align*}k_1&=\begin{cases} 0 &\text{if}\,\, n=3k\\ 2 &\text{if}\,\, n=3k+1\\1 &\text{if}\,\, n=3k+2\end{cases}\\
- k_2&=\begin{cases} k &\text{if}\,\, n=3k\\ k-1 &\text{if}\,\, n=3k+1\\k &\text{if}\,\, n=3k+2\end{cases}\\
-\prod_{i=1}^{\alpha}a_i&=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}\end{align*}
-This proves the lemma.
-\xbox
-\end{proof}
-%
-From Lemma~\ref{lemma:minhitmax}, we conclude that
-$$h=\begin{cases}
-3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\  |\Sigma|~~\equiv~~0 \mod 3\\
-3^{\lfloor\frac{|\Sigma|}{3}\rfloor}  + 3^{\lfloor\frac{|\Sigma|}{3}\rfloor-1}&\text{if}\ |\Sigma|~~\equiv~~1\mod 3\\
-2\cdot 3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\ |\Sigma|~~\equiv~~2 \mod 3
-\end{cases}
-$$
-%
-We note that $h$ is significantly smaller than $2^{|\Sigma|}-1$, this can
-also be intuitively motivated by the fact that  for every $H\in
-\minhits(P/s)$, all of its true subsets and true supersets are {\it not}
-contained in $\minhits(P/s)$.
-
-From Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
-$k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
-we use the formula for the sum of the geometric progression.
-%
-\[
-h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
- \text{or, asymptotically,}\  O(3^{\lfloor\frac{|\Sigma|}{3}\rfloor}\cdot|\Sigma|^{(pq-1)}).
- \]
-%
-From Lemma~\ref{lemma:hseta} we know that the worst-case bound for $h$
-above cannot be further improved, since the full collection of minimal
-hitting sets needs to be checked to verify $conf$. The
-authors of~\cite{Hennessy:1988:ATP:50497} and
-\cite{DBLP:conf/icfem/CavalcantiG07} suggest to test {\it every} non-empty
-subset of $\Sigma$ whose events cannot be completely refused in a given
-process state of the reference model; this leads to a worst-case estimate of
-$2^{|\Sigma|}-1$ for the number of different sets to be offered to the SUT in
-the last step of the test execution, so our approach reduces the number of
-test executions in comparison
-to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07}.
-
-% -------------------------------------------------------------------------
-\subsubsection*{Upper Bound $pq$ for Number of Test Cases}
-
-According to Theorem~\ref{th:failurestest}, the tests $U_F(k)$ need to be executed for 
-$k = 0,\dots,pq-1$ to guarantee completeness. This means that the SUT is verified 
-with test traces up to, and including, length $pq$. It is interesting to  investigate
-whether this maximal length is really necessary, or whether one could elaborate
-alternative complete test strategies where the SUT is tested with shorter traces only. 
-Indeed, an example  presented in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} 
-shows that when testing for equivalence of deterministic FSMs, it is sufficient to
-test the SUT with traces of significantly shorter length.
-
-The following example, however, shows that the maximal length $pq$ is really required 
-when testing for refinement.
-\begin{example}\label{ex:pq}
-Consider the CSP reference process $P$ and an erroneous implementation $Q$ specified
-as follows.
-\begin{eqnarray*}
-P & = & a \then P_1 \intchoice b \then P_1 \intchoice c \then P_1
-\\
-P_1 & = & a \then P \extchoice b\then P
-\\ & &  \\
-Q & = & a\then Q_1 \extchoice b\then Q_1
-\\
-Q_1 & = & a\then Q_2 \extchoice b\then Q_2
-\\
-Q_2 & = & a\then Q \intchoice b\then Q
-\end{eqnarray*}
-Obviously, $P$'s normalised transition graph has 2 nodes, 
-while $Q$'s graph has 3.
-It is easy to see (and can be checked with FDR4) that $P\lessdet_T Q$, but 
-$\neg(P\lessdet_F Q)$. Furthermore, it can also be shown using the FDR4 tool that
-the ``test passed condition'' 
-\[
-(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k))\hide \Sigma
-\]
-holds for $U_F(0),\dots,U_F(4)$, but fails for $U_F(5)$. This means that the
-non-conformance of $Q$ cannot be detected by any test trace of length 
-less or equal to 5, but is revealed (as expected from Theorem~\ref{th:failurestest})
-by a trace of length 6, because the last event offered by the test $U_F(5)$ is 
-refused by $Q$.
-\xbox
-\end{example}
-%
-Generalising Example~\ref{ex:pq}, it can be shown that for any pair $p,q\in\mathbb{N}$
-whose greatest common divisor is 1, there exist reference processes with $p$ states
-and implementation processes with $q$ states, such that non-conformance of the 
-implementation
-can only be detected with a trace of length $pq$.
-
-It is discussed in Section~\ref{sec:conc} how the number of test traces to be executed 
-by complete test suites for failures refinement   can still be reduced {\it without} reducing 
-the maximal length.
-
-% ==========================================================================

### 2018-09-09 paper/finitecomplete-trace.tex
Completed the discussion - only the discussion part about fault domains is still missing.
Technical report completed, taking into account the remarks from Ana

@@
 will be used for a given reference process $P$ and  integers $p \ge 0$.
+Just as for the tests developed  in 
+Section~\ref{sec:finitecompletefails} to verify failures refinement, 
+the tests for trace refinement are
+derived from the reference model's transition graph
+$$
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)).
+$$
+In contrast to the failures tests (\ref{eq:UFP}), however, we do not need to
+check the SUT with respect to its acceptance of hitting sets. Therefore, these
+do not occur in the specification of the test cases below, and we use the
+condition $\minaccs(n) = \{ \varnothing \}$ instead of $\minhits(n) = \varnothing$
+to indicate that minimal hitting sets need not to be calculated for generating
+these tests from $G(P)$. From (\ref{eq:minhitminaccempty}) we know that both conditions
+are equivalent.
+
 %
+%\begin{eqnarray}
+%U_T(p) & = & U_T(p,\varepsilon)
+%\\
+%U_T(p,s) & = & \big(\Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
+%\label{eq:uta}
+%\\ & & \extchoice \nonumber
+%\\ & & (\minaccs(P/s) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
+%\label{eq:utb}
+%\\ & & \extchoice \nonumber
+%\\ & & (\#s < p) \& \big(\Extchoice e:[P/s]^0 @ e \then U_T(p,s.e) \big)
+%\label{eq:utc}
+%\\ & & \extchoice \nonumber
+%\\ & & (\#s = p) \& \big( \epass\then \Stop  \big)
+%\label{eq:utd}
+%\end{eqnarray}
+
 \begin{eqnarray}
-U_T(p) & = & U_T(p,\varepsilon)
+U_T(p) & = & U_T(p,0,\ii n)
 \\
-U_T(p,s) & = & \big(\Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
+U_T(p,k,n) & = & \big(  e:(\Sigma - [n]^0)   \then \efail\then \Stop \big)
 \label{eq:uta}
 \\ & & \extchoice \nonumber
-\\ & & (\minaccs(P/s) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
+\\ & & (\minaccs(n) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
 \label{eq:utb}
 \\ & & \extchoice \nonumber
-\\ & & (\#s < p) \& \big(\Extchoice e:[P/s]^0 @ e \then U_T(p,s.e) \big)
+\\ & & (k < p) \& \big( e:[P/s]^0   \then U_T(p,k+1,t(n,e)) \big)
 \label{eq:utc}
 \\ & & \extchoice \nonumber
-\\ & & (\#s = p) \& \big( \epass\then \Stop  \big)
+\\ & & (k = p) \& \big( \epass\then \Stop  \big)
 \label{eq:utd}
 \end{eqnarray}
+
+It is easy to see that the tests $U_T(p)$ satisfy the properties
+\begin{eqnarray}
+\label{eq:ifpaT}
+  &  & U_T(p)/s = U_T(p,\#s,G(P)/s)
+\\
+\label{eq:ifpbT}
+e\not\in [P/s]^0 & \implies & U_T(p)/s.e = (\efail\then\Stop)
+\end{eqnarray}
+proven in Lemma~\ref{lemma:ufproperties} for $U_F(p)$ for traces $s\in\trc(P)$.
+
 %
-The difference between adaptive tests $U_T(p)$ for trace refinement and
-$U_F(p)$ for failures refinement consists in the fact that the former do not
-``probe'' the SUT with respect to minimal sets of events to be accepted
-without blocking.
+%The difference between adaptive tests $U_T(p)$ for trace refinement and
+%$U_F(p)$ for failures refinement consists in the fact that the former do not
+%``probe'' the SUT with respect to minimal sets of events to be accepted
+%without blocking.
 
@@
 for erroneous acceptances at the end of each trace of length $(pq -1)$.
+Since $U_T(pq-1)$ never blocks any $Q$-event before terminating, the pass 
+criterion can be based on trace refinement instead of failures refinement
+as required in (\ref{eq:passF}).
+%
+\begin{equation}
+\label{eq:passT}
+Q\ \pass\ U_T(p) \defs (\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(p)) \hide \Sigma
+\end{equation}
+%
 
 \begin{theorem}\label{th:tracetest}
-Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
+Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
 normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
-D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
+D}$ as the set of all non-terminating, divergence-free CSP processes over alphabet $\Sigma$,
 whose transition graph has at most $q$ states with $q \ge p$. Then the test

### 2018-09-09 paper/introduction.tex
Technical report completed, taking into account the remarks from Ana

@@
 This paper presents complete black-box test suites for software and systems
-modelled using CSP. They can be generated for divergence-free\footnote{The
+modelled using CSP. They can be generated for non-terminating, 
+divergence-free\footnote{The
 assumption of divergence freedom is usually applied in black-box testing,
-since it cannot distinguish between divergence and deadlock.} finite-state
+since it cannot distinguish between divergence and deadlock.}, finite-state
 CSP processes with finite alphabets, interpreted both in the trace and the
@@
 restrict fault domains using a graph representation of processes (originally
-elaborated for model checking) to obtain finite test suites. \fixme{alcc: I
-think termination is not handled either?}
+elaborated for model checking) to obtain finite test suites. 
+%%%\fixme{alcc: I think termination is not handled either?}
 
@@
 verify the conformance relations \emph{trace refinement} and \emph{failures
-refinement} for any given non-divergent, finite-state, finite alphabet CSP
+refinement} for any given non-terminating, 
+non-divergent, finite-state, finite alphabet CSP
 process. We prove their completeness with respect to fault domains
@@
 Moreover, our result is not a simple transcription of existing knowledge
-about finite, complete test suites for finite state machines: the
+about finite, complete test suites for finite state machines. The
 capabilities of CSP to express nondeterminism in a more fine-grained way than
@@
 testing against nondeterministic FSMs, as published, for example,
-in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the
+in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}: CSP distinguishes between 
+external choice, where the environment can control the behaviour of a process, and
+internal choice where process behaviour is internally decided and cannot be influenced
+by the environment. In contrast to this, FSMs specify nondeterministic behaviour
+by offering more than one possible output to a given input, and this output can be 
+controlled by an FSM representing the environment. Loosely speaking, this corresponds to external choice in CSP, while there is no equivalent to internal choice in nondeterministic FSMs.
+
+Finally, we prove two optimality results.
+(1) We show that the
 approach to testing the admissibility of an SUT's nondeterministic behaviour
-is optimal in the sense that it cannot be established with fewer test cases.
-\fixme{alcc: I did not understand this?}
+is optimal in the sense that it cannot be established with fewer ``probes'' investigating
+the SUT's degree of nondeterminism. These probes are minimal sets of events offered
+by the test to the SUT, such that a refusal of the SUT to engage into any of these events
+reveals a violation of failures refinement.
+(2) Furthermore, it is shown that the maximal length of traces to be investigated in 
+a complete test suite cannot be further reduced without loosing the suite's capability to
+uncover any violation of trace or faulures  refinement.
 
-%It should be noted that the presented results are of a theoretical
-%nature:~despite being finite, the resulting test suite size will still be too
-%large to be applied in practice to real-world problems. We discuss a number
-%of promising options how the test suite size can be reduced to become
-%practically applicable.
+%%%\fixme{alcc: I did not understand this?}
 
@@
 suites checking trace refinement are a simplified version of the former
-class; they are presented in Section~\ref{sec:finitecomplete}. A sample test
+class; they are presented in Section~\ref{sec:finitecomplete}. 
+The optimality results are presented in Section~\ref{sec:complexity}, together with further
+complexity considerations.
+A sample test
 suite is presented in Section~\ref{sec:case}. Our results are discussed in

### 2018-09-23 paper/complexity.tex
New upper bound and new proof for maximal number of hitting sets, based on Sperner's Theorem.

@@
 The first factor $\ell$
-has worst-case upper bound $\ell\le |\Sigma|^k$. As an example, where this upper bound 
+has worst-case upper bound $\ell\le \card{\Sigma}^k$. As an example, where this upper bound 
 is really met, consider the reference process
@@
 never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly 
-$|\Sigma|^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
+$\card{\Sigma}^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
 for each of their events.
@@
 Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
-acceptances, the cardinality $h$ of $\minhits(P/s)$ is maximal for the case
-where all $A_i$ are disjoint. In this case,  $h~~=~~|\minhits(P/s)|~~=~~
-\prod_{i=1}^\alpha|A_i|$. To find an upper bound for $h$, we use the result
-that we prove in the following lemma.
-%
-\begin{lemma}\label{lemma:minhitmax}
-Let $n\ge 2$ be any positive integer. We call a sequence
-%
-$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$
-%
-a partition of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$. Define
-%
-$$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
-Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
-\alpha\in \mathbb{N}\big\}.$$
-%
-\end{lemma}
+acceptances, the cardinality   $h = \minhits(P/s)$ has the upper bound
+\[
+h \le \binom{n}{\lfloor\frac{n}{2}\rfloor}, \quad \text{where $n = \card{\Sigma}$.}
+\]
+This follows from Theorem~\ref{th:sperner} and the fact that $\minhits(P/s)$ is a 
+Sperner Family, as introduced in Section~\ref{sec:hit}. The next theorem shows that this
+upper bound can really be reached by collections of minimal hitting sets.
+
+\begin{theorem}
+\label{th:upperboundh}
+Let $n\ge 2$ be any positive integer and $\Sigma$  any finite set of cardinality $n$.
+Let $C\subseteq 2^\Sigma$ be any collection of  subsets of $\Sigma$. Then  
+\[
+\text{max}\{\card{\text{minHit}(C)}~|~C\subseteq 2^\Sigma\}=\binom{n}{\lfloor\frac{n}{2}\rfloor}
+\]
+\end{theorem}
 \begin{proof}
-Let $n\ge 2$. Obviously, \begin{equation}\label{eq:1} n\le
-p(n).\end{equation} Define $\pi(n)$ to be the following partition  of $n$:
-%
-$$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
-%
-Then the product of the numbers in $\pi(n)$ equals  $p(n)$. In the remainder
-of the proof we have to show that $p(n)$ is really maximal in the sense of
-the lemma.
-
-Since the number of partitions of a given $n$ is finite,
-%
-$$\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$$
-%
-exists. Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with
-maximal product. Then, obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
-Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then
-from~(\ref{eq:1}) we have
-$$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
-Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a
-product value which is not less then the original partition $a_1, \ldots ,
-a_{\alpha}$. This process can be repeated, until every $a_i$ in the resulting
-partition is either 2 or 3. Hence there exists a partition of $n$ with
-maximal product, such that every term is $2$ or $3$. Since $6=2+2+2=3 + 3$
-and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than
-$3$. Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\},
-i=1,\dots,\alpha$, be a partition of $n$ with maximal product. Let $k_1\le 2$
-be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
-Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$
-with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have
-
-\begin{align*}k_1&=\begin{cases} 0 &\text{if}\,\, n=3k\\ 2 &\text{if}\,\, n=3k+1\\1 &\text{if}\,\, n=3k+2\end{cases}\\
- k_2&=\begin{cases} k &\text{if}\,\, n=3k\\ k-1 &\text{if}\,\, n=3k+1\\k &\text{if}\,\, n=3k+2\end{cases}\\
-\prod_{i=1}^{\alpha}a_i&=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}\end{align*}
-This proves the lemma.
+For any subset  $B$  of $\Sigma$, define  $A_{B}:=\Sigma\setminus B$.
+Define 
+$$
+C=\{A_{B}~|~ B\subseteq \Sigma, \card{B} < \lfloor\frac{n}{2}\rfloor\}.
+$$
+Let $H$ be any minimal Hitting set of $C$. Then $H$ contains at least $\lfloor\frac{n}{2}\rfloor$ elements, because 
+otherwise there existed an $A_{H}\in C$ with $A_H\cap H=\varnothing$. 
+Since $\card{A_{B}}=n-\card{B}>n-\lfloor\frac{n}{2}\rfloor$, for any $B\subseteq \Sigma$ with $\card{B}< \lfloor\frac{n}{2}\rfloor$, we conclude that
+ any  subset of $\Sigma$  with at least $\lfloor\frac{n}{2}\rfloor$ elements 
+ has a non-empty intersection with all  $A_B\in C$.  
+ The number of subsets in $\Sigma$ with exactly $\lfloor \frac{n}{2}\rfloor$ elements
+ is $\binom{n}{\lfloor\frac{n}{2}\rfloor}$.
+Thus $\card{\text{minHit}(C)} \ge \binom{n}{\lfloor\frac{n}{2}\rfloor}$, and 
+Theorem~\ref{th:sperner} yields $\card{\text{minHit}(C)} = \binom{n}{\lfloor\frac{n}{2}\rfloor}$.
 \xbox
 \end{proof}
+
+To get an approximation of the maximal size of $\minhits(P/s)$ for large $n$, recall
+Stirling's approximation~\cite[p.~112]{Graham:1994:CMF:562056}
+\[
+m! \approx \sqrt{2\pi m} \cdot \big( \frac{m}{e} \big)^m, \quad\text{where $e$ denotes the Euler Number.}
+\]
+Applying this approximation to the   case where $n = \card{\Sigma}$ is even and $m = (n/2)$  
+results in
+\[
+h \le \binom{2m}{m} \approx \frac{4^m}{\sqrt{\pi m}} = 
+\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}.
+\]
 %
-From Lemma~\ref{lemma:minhitmax}, we conclude that
-$$h=\begin{cases}
-3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\  |\Sigma|~~\equiv~~0 \mod 3\\
-3^{\lfloor\frac{|\Sigma|}{3}\rfloor}  + 3^{\lfloor\frac{|\Sigma|}{3}\rfloor-1}&\text{if}\ |\Sigma|~~\equiv~~1\mod 3\\
-2\cdot 3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\ |\Sigma|~~\equiv~~2 \mod 3
-\end{cases}
-$$
-%
-We note that $h$ is significantly smaller than $2^{|\Sigma|}-1$, this can
-also be intuitively motivated by the fact that  for every $H\in
-\minhits(P/s)$, all of its true subsets and true supersets are {\it not}
-contained in $\minhits(P/s)$.
+We note that, due to this approximation result, 
+$h$ is   smaller than $2^{\card{\Sigma}}-1$, the cardinality of the 
+non-empty subsets of $\Sigma$.
 
@@
 \[
-h\cdot \big( \frac{1-|\Sigma|^{pq}}{1-|\Sigma|} \big),\
- \text{or, asymptotically,}\  O(3^{\lfloor\frac{|\Sigma|}{3}\rfloor}\cdot|\Sigma|^{(pq-1)}).
+h\cdot \big( \frac{1-\card{\Sigma}^{pq}}{1-\card{\Sigma}} \big),\
+ \text{or, asymptotically,}\  O(\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}\cdot\card{\Sigma}^{(pq-1)}).
  \]
 %
-From Lemma~\ref{lemma:hseta} we know that the worst-case bound for $h$
+From Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
 above cannot be further improved, since the full collection of minimal
@@
 process state of the reference model; this leads to a worst-case estimate of
-$2^{|\Sigma|}-1$ for the number of different sets to be offered to the SUT in
+$2^{\card{\Sigma}}-1$ for the number of different sets to be offered to the SUT in
 the last step of the test execution, so our approach reduces the number of
 test executions in comparison
-to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07}.
+to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} by a factor of
+$1/{\sqrt{2\pi \card{\Sigma}}}$.
+
+In Fig.~\ref{fig:minhita}, the reduction is visualised by a function plot of
+$2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.
+
+% .....................................................................................
+ \begin{figure}
+ %%\hspace*{-40mm}
+ \begin{center}
+\includegraphics[width=.8\textwidth]{minhit-fig.pdf}
+\end{center}
+\vspace*{-10mm}
+\caption{Function plot $2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.}
+ \label{fig:minhita}
+ \end{figure}
+% .......................................................................................
 
@@
 above, the number of executions performed by $(Q\parallel[\Sigma] U_R(pq-1))$ 
-is bounded by $|\Sigma|^{pq-1}$.
+is bounded by $\card{\Sigma}^{pq-1}$.
 

### 2018-09-23 paper/preliminaries.tex
New upper bound and new proof for maximal number of hitting sets, based on Sperner's Theorem.

@@
 \end{equation}
+
+In preparation of complexity results to be presented in Section~\ref{sec:complexity}, we consider the question how many minimal hitting sets can maximally exist for a given 
+collection of finite sets. To this end, the following definitions and results are 
+useful\footnote{The authors are grateful to @todo for suggesting the applicability of Sperner's Theorem in the context of the work presented here.}. 
+A \emph{Sperner Family} is a collection $S \subseteq 2^\Sigma$ of sets from a given universe $\Sigma$ that do not contain each other, that is, 
+$H_1 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair $H_1\neq H_2\in S$. 
+Specialising   antichains known from   partial orders to finite sets partially ordered by $\subseteq$ results in Sperner families.
+
+We observe that the collection of minimal hitting sets constructed from the minimal acceptances of a CSP process state build a Sperner Family. Therefore, the following theorem can be applied.
+\begin{theorem}[Sperner's Theorem~\cite{sperner_satz_1928}]
+\label{th:sperner}
+The largest Sperner family $S$ 
+over an $n$-element universe $\Sigma$ has cardinality 
+$$
+\card{S} = \binom{n}{\lfloor\frac{n}{2}\rfloor}.
+$$
+\xbox
+\end{theorem}
+It will be shown in Section~\ref{sec:complexity} that this upper bound can actually be reached by hitting sets associated with the minimal acceptances of a CSP process state.

### 2018-09-24 paper/complexity.tex
Show that CSP process exists with minHit(P) = n choose floor(n\2)

@@
 \]
-This follows from Theorem~\ref{th:sperner} and the fact that $\minhits(P/s)$ is a 
-Sperner Family, as introduced in Section~\ref{sec:hit}. The next theorem shows that this
-upper bound can really be reached by collections of minimal hitting sets.
+This follows from Theorem~\ref{th:sperner}, since $\minhits(P/s)$ is a 
+Sperner Family, as introduced in Section~\ref{sec:sperner}. 
+The next theorem shows that this
+upper bound can really be reached by collections of minimal hitting sets associated with 
+a CSP process state.
 
@@
 \label{th:upperboundh}
-Let $n\ge 2$ be any positive integer and $\Sigma$  any finite set of cardinality $n$.
-Let $C\subseteq 2^\Sigma$ be any collection of  subsets of $\Sigma$. Then  
+Let  $\Sigma$  be an alphabet with cardinality $n\ge 2$.  Then  
+there exists a CSP process $P$ with 
 \[
-\text{max}\{\card{\text{minHit}(C)}~|~C\subseteq 2^\Sigma\}=\binom{n}{\lfloor\frac{n}{2}\rfloor}
+\card{\text{minHit}(P)} = \binom{n}{\lfloor\frac{n}{2}\rfloor}.
 \]
@@
 \begin{proof}
-For any subset  $B$  of $\Sigma$, define  $A_{B}:=\Sigma\setminus B$.
-Define 
-$$
-C=\{A_{B}~|~ B\subseteq \Sigma, \card{B} < \lfloor\frac{n}{2}\rfloor\}.
-$$
-Let $H$ be any minimal Hitting set of $C$. Then $H$ contains at least $\lfloor\frac{n}{2}\rfloor$ elements, because 
-otherwise there existed an $A_{H}\in C$ with $A_H\cap H=\varnothing$. 
-Since $\card{A_{B}}=n-\card{B}>n-\lfloor\frac{n}{2}\rfloor$, for any $B\subseteq \Sigma$ with $\card{B}< \lfloor\frac{n}{2}\rfloor$, we conclude that
- any  subset of $\Sigma$  with at least $\lfloor\frac{n}{2}\rfloor$ elements 
- has a non-empty intersection with all  $A_B\in C$.  
- The number of subsets in $\Sigma$ with exactly $\lfloor \frac{n}{2}\rfloor$ elements
- is $\binom{n}{\lfloor\frac{n}{2}\rfloor}$.
-Thus $\card{\text{minHit}(C)} \ge \binom{n}{\lfloor\frac{n}{2}\rfloor}$, and 
-Theorem~\ref{th:sperner} yields $\card{\text{minHit}(C)} = \binom{n}{\lfloor\frac{n}{2}\rfloor}$.
+Let $C$ be the collection of all subsets of $\Sigma$ with cardinality 
+$n-\lfloor{\frac{n}{2}}\rfloor+1$. Then any CSP process of the form 
+\[
+P = \Intchoice_{A\in C} e:A\then P(e)
+\]
+fulfils $\minaccs(P) = C$.
+Let $H$ be any minimal Hitting set of $C$. Then $H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because 
+otherwise
+$\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, 
+and any subset $A\subseteq \Sigma\setminus H$ with cardinality
+$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in $C$, but satisfy 
+$A\cap H=\varnothing$. 
+Since $\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we conclude
+ that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$ 
+ intersects  every element of $C$.  Therefore, every minimal hitting set of $C$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements,
+so $\card{\minhits(C)} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$.
 \xbox

### 2018-09-24 paper/preliminaries.tex
Show that CSP process exists with minHit(P) = n choose floor(n\2)

--- 2018-09-07 paper/preliminaries.tex
@@
 
+\subsubsection*{Definition}
 The main idea of the underlying test strategy for failures refinement can be
@@
 
+% ------------------------------------------------------------------------
+\subsubsection*{Characterisation of $conf$ by Minimal Hitting Sets}
 The following lemma establishes that the $conf$ relation specified in
@@
 
+% -------------------------------------------------------------------------
+\subsubsection*{Minimal Hitting Sets of Normalised Transition Graphs}
 Since minimal acceptances can be used to label the nodes of a normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$, the notation of minimal hitting sets also
@@
 \end{equation}
+
+% =========================================================================
+\subsection{Sperner Families}
+\label{sec:sperner}
+In preparation of complexity results to be presented in Section~\ref{sec:complexity}, we consider the question how many minimal hitting sets can maximally exist for a given 
+collection of minimal acceptances. 
+To this end, the following definitions and results are 
+useful\footnote{The authors are grateful to Li-Da Tong from National Sun Yat-sen University, Taiwan, 
+for suggesting the applicability of Sperner's Theorem in the context of the work presented here.}.
+
+A \emph{Sperner Family} is a collection $S \subseteq 2^\Sigma$ of sets from a given 
+finite universe $\Sigma$ that do not contain each other, that is, 
+$H_1 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair $H_1\neq H_2\in S$. 
+Specialising   antichains known from   partial orders to finite sets partially ordered by $\subseteq$ results in Sperner families. 
+We observe that
+\begin{itemize}
+\item the maximal refusals of a CSP process state,
+\item the minimal acceptances of a CSP process state, and
+\item the minimal hitting sets of a given collection of sets
+\end{itemize}
+represent Sperner families. Moreover, given any finite alphabet $\Sigma$ with
+$\card{\Sigma} = n$, every collection $S$ of subsets with identical cardinality $k \le n$
+is a Sperner family,  because $A_1, A_2\in S \wedge A_1 \subseteq A_2\wedge \card{A_1} = \card{A_2}$ implies $A_1 = A_2$. Given  any Sperner Family $S$ of $\Sigma$, $S$ represents
+the minimal acceptances in the initial state of the CSP process  
+\[
+P = \Intchoice_{A\in S}\big( e:A \then P(e) \big).
+\]
+
+The cardinality of Sperner Families is determined by the following theorem.
+\begin{theorem}[Sperner's Theorem~\cite{sperner_satz_1928}]
+\label{th:sperner}
+Given a  Sperner family $S$ 
+over an $n$-element universe $\Sigma$, its cardinality is bound by
+\[
+\card{S} \le \binom{n}{\lfloor\frac{n}{2}\rfloor}.
+\]
+The upper bound is reached if and only if one of the following cases apply.
+\begin{enumerate}
+\item For even $n$, if $S$ consists of all subsets of $\Sigma$ with cardinality $n/2$.
+\item For odd $n$, if one of the following cases holds.
+\begin{enumerate}
+\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/s$.
+\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/s$.
+\end{enumerate}
+\end{enumerate}
+\xbox
+\end{theorem}
+It will be shown in Section~\ref{sec:complexity} that this upper bound can actually be reached by hitting sets associated with the minimal acceptances of a CSP process state.

### 2018-09-27 paper/introduction.tex
couple of questions

@@
 
-Model-based testing (MBT) is an  active research field that is currently
-evaluated and integrated into industrial verification processes by many
+Model-based testing (MBT) is an active research field; results are currently
+being evaluated and integrated into industrial verification processes by many
 companies worldwide. This holds particularly for the embedded and
@@
 modelled using CSP. They can be generated for non-terminating,
-divergence-free\footnote{The
-assumption of divergence freedom is usually applied in black-box testing,
-since it cannot distinguish between divergence and deadlock.}, finite-state
-CSP processes with finite alphabets, interpreted both in the trace and the
-failures semantics. Our results complement work
-in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
-collections of processes refining a ``most general'' fault domain member.
-With that concept, complete test suites may be finite or infinite. This gives
-important insight into the theory of fault-domain testing for CSP, but we are
-particularly interested in {\it finite} suites when it comes to practical
-application. While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional
-criteria to select tests from still infinite test suites, here, we further
-restrict fault domains using a graph representation of processes (originally
-elaborated for model checking) to obtain finite test suites. 
-%%%\fixme{alcc: I think termination is not handled either?}
+divergence-free, finite-state CSP processes with finite alphabets,
+interpreted both in the trace and the failures semantics. Divergence freedom
+is usually assumed in black-box testing, since it cannot distinguish between
+divergence and deadlock.
+
+Our results complement work in~\cite{DBLP:conf/pts/CavalcantiS17}. There,
+fault domains are specified as collections of processes refining a ``most
+general'' fault domain member. With that concept, complete test suites may be
+finite or infinite. This result gives important insight into the theory of
+fault-domain testing for CSP, but we are particularly interested in {\it
+finite} suites when it comes to practical application.
+While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
+select tests from still infinite test suites, here, we further restrict fault
+domains using a graph representation of processes (originally elaborated for
+model checking) to obtain finite test suites.
 
@@
 verify the conformance relations \emph{trace refinement} and \emph{failures
-refinement} for any given non-terminating, 
-non-divergent, finite-state, finite alphabet CSP
-process. We prove their completeness with respect to fault domains
-of the described type.  The existence of -- possibly infinite -- complete
-test suites has been established for process algebras, and for CSP in
-particular, by several
+refinement} for any given non-terminating, non-divergent, finite-state,
+finite alphabet CSP process. We prove their completeness with respect to
+fault domains of the described type.  The existence of -- possibly infinite
+-- complete test suites has been established for process algebras, and for
+CSP in particular, by several
 authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
-To the best of our knowledge, however, this article is the first to present
-{\it finite}, complete test suites associated with this class of fault
-domains and conformance relations.
+To the best of our knowledge, this article is the first to present {\it
+finite}, complete test suites associated with this class of fault domains and
+conformance relations.
 
-Moreover, our result is not a simple transcription of existing knowledge
-about finite, complete test suites for finite state machines. The
-capabilities of CSP to express nondeterminism in a more fine-grained way than
-it is possible for FSMs requires a more complex approach to testing systems
-for conformity in the failures-model than it is required for model-based
-testing against nondeterministic FSMs, as published, for example,
-in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}: CSP distinguishes between 
-external choice, where the environment can control the behaviour of a process, and
-internal choice where process behaviour is internally decided and cannot be influenced
-by the environment. In contrast to this, FSMs specify nondeterministic behaviour
-by offering more than one possible output to a given input, and this output can be 
-controlled by an FSM representing the environment. Loosely speaking, this corresponds to external choice in CSP, while there is no equivalent to internal choice in nondeterministic FSMs.
+Our result is not a simple transcription of existing knowledge about finite,
+complete test suites for FSMs. The capabilities of CSP to express
+nondeterminism in a more fine-grained way than it is possible for FSMs
+requires a more complex approach to testing for conformity in the
+failures-model than required for model-based testing against nondeterministic
+FSMs, as published, for example,
+in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. CSP distinguishes
+between external choice, where the environment can control the behaviour of a
+process, and internal choice where the behaviour is decided internally and
+cannot be influenced by the environment. In contrast, FSMs specify
+nondeterministic behaviour by offering more than one possible output for a
+given input, and this output can be controlled by an FSM representing the
+environment. Loosely speaking, this corresponds to external choice in CSP,
+while there is no equivalent to internal choice in nondeterministic FSMs.
+\fixme{alcc: I'm slightly uncomfortable with this claim. Can the environment
+really control outputs? Do you not have tau (internal) events in FSM? I'm
+unsure, but you guys know about FSMs much more than I do.}
 
-Finally, we prove two optimality results.
-(1) We show that the
-approach to testing the admissibility of an SUT's nondeterministic behaviour
-is optimal in the sense that it cannot be established with fewer ``probes'' investigating
-the SUT's degree of nondeterminism. These probes are minimal sets of events offered
-by the test to the SUT, such that a refusal of the SUT to engage into any of these events
-reveals a violation of failures refinement.
-(2) Furthermore, it is shown that the maximal length of traces to be investigated in 
-a complete test suite cannot be further reduced without loosing the suite's capability to
-uncover any violation of trace or faulures  refinement.
-
-%%%\fixme{alcc: I did not understand this?}
+Finally, we prove two optimality results here. (1)~We show that our approach
+to testing the admissibility of an SUT's nondeterministic behaviour is
+optimal in the sense that it cannot be established with fewer ``probes''
+investigating the SUT's degree of nondeterminism. These probes are minimal
+sets of events offered by the tests to the SUT, such that a refusal of the
+SUT to engage into at least one of these events reveals a violation of
+failures refinement. (2)~Furthermore, it is shown that the maximal length of
+traces to be investigated in a complete test suite cannot be further reduced
+without loosing the suite's capability to uncover any violation of trace or
+failures refinement.
 
@@
 \subsubsection*{Overview}
-In Section~\ref{section:preliminaries}, we present the background material
+In Section~\ref{section:preliminaries}, we present the background 
 relevant to our work. In Section~\ref{sec:finitecompletefails}, finite

### 2018-09-28 paper/preliminaries.tex
couple of small changes

@@
 
-In this section, we present CSP and the concept of minimal hitting sets, which
-is central to our novel notion of test for failures refinement.
+We present CSP~(Section~\ref{section:csp}) and the concept of minimal hitting
+sets~(Section~\ref{sec:hit}), which is central to our novel notion of test
+for failures refinement.
 
@@
 \subsection{CSP, Refinement, and Normalised Transition Graphs}
+\label{section:csp}
+% =========================================================================
 
@@
 
-Throughout this paper, the alphabet of the processes, that is the set of
+Throughout this paper, the alphabet of the processes, that is, the set of
 events that are in scope, is denoted by $\Sigma$ and supposed to be finite.
@@
 Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
-external and internal choice operators allow us to define a choice over a
-collection of processes $P(i)$. By convention,  if the index set $I$ is
-empty, external choices evaluate to the process $\Stop$, which
-deadlocks:~does not engage into any event and does not terminate. An iterated
-internal choice is not defined for an empty index set. For an external choice 
-indexed over a set of events in the form
-$\Extchoice e: A @ e \then P(e), A\subseteq \Sigma$, we use 
-abbreviation $e:A \then P(e)$.
+external and internal choice operators define a choice over a collection of
+processes $P(i)$. If the index set $I$ is empty, the external choice is the
+process $\Stop$, which deadlocks:~does not engage into any event or
+terminate. For an external choice $\Extchoice e: A @ e \then P(e), A\subseteq
+\Sigma$ over a set $A$ of events, we use the abbreviation $e:A \then P(e)$.
+An iterated internal choice is not defined for an empty index set.
 
@@
 is used in conjunction with parallelism:~it is often desirable to make
-internal actions of each process in a network of parallel processes, perhaps
-used for coordination of the network, invisible, while events happening at
-their interfaces remain observable.
+actions of each process in a network of parallel processes, perhaps used for
+coordination of the network, invisible, while events happening at their
+interfaces remain observable.
 
@@
 
-Failures refinement, $P \lessdet_F Q$, is defined as expected by set
-containment $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
-$P \lessdet_F Q$ implies
+For divergence-free processes, failures refinement, $P \lessdet_F Q$, is
+defined as expected by set containment $\fails(Q)\subseteq\fails(P)$. Since
+refusals are subset-closed, $P \lessdet_F Q$ implies
 $(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. So, for
 divergences-free processes, failures refinement implies trace refinement.
-Therefore, using the conformance relation
+Therefore, using the conformance relation $conf$ below
 %
@@
 %
-
 For finite CSP processes, since refusals are
@@
 sequence of edge labels. We note that $\trc(P) = \trc(G(P))$. Since $G(P)$ is
-normalised, there is a unique node reached by applying the events from $s$
+normalised, there is a unique node reached by following the events from $s$
 one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well defined.
@@
 %
-Exploiting formula (\ref{eq:accref}), 
+Exploiting (\ref{eq:accref}),
 %(\ref{eq:accinitials}), and (\ref{eq:accsubset}),
@@
 \end{equation}
-%\fixme{alcc: proof in the report perhaps?}
+\fixme{alcc: proof in the report perhaps? What do you need (5) for? }
 
@@
  \end{figure}
+
+\fixme{alcc: the a from node 2 should go back to 0, I think.}
+
 % .......................................................................................
@@
 \begin{proof}
-Recall that $P\lessdet_T Q$ is defined
-as $\trc(Q)\subseteq \trc(P)$ and that $\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$.
-This shows (\ref{eq:trcrefa}).
+Recall that $P\lessdet_T Q$ is defined as $\trc(Q)\subseteq \trc(P)$ and that
+$\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$. This shows (\ref{eq:trcrefa}).
+\fixme{alcc: (11) is already proved in my paper with mcg. No need to repeat
+it here. I suggest just to add a reference.}
 
@@
 \end{proof}
+%
+This concludes our presentation of CSP and of some results regarding its
+semantics and normalised transition graphs that are used in the next section.
 
-%\fixme{Abrupt end to the section. It has more than just preliminary material.
-%It has some new results.}
+\fixme{Abrupt end to the section. It has more than just preliminary material.
+It has some new results.}
 

### 2018-09-28 paper/finitecomplete-failures.tex
changed order of material

@@
 \subsubsection*{Test Definition and Basic Properties}
-In the domain of process algebras, test cases are typically represented
-by processes interacting concurrently with the SUT
-process~\cite{Hennessy:1988:ATP:50497}.
-They synchronise with the SUT over its visible events and use some additional events outside the SUT's alphabet to express whether the test execution  passed or failed.
+
+In the domain of process algebras, test cases are typically represented by
+processes interacting concurrently with the
+SUT~\cite{Hennessy:1988:ATP:50497}. Considering an (unknown) process that
+represents the behaviour of the SUT, we say that tests synchronise with the
+process for the SUT over its visible events and use some additional events
+outside the SUT process's alphabet to express whether the test execution
+passed or failed.
 %%%, or if no verdict could be obtained.
@@
 $$
-and for each integer $p\ge 0$, we define a
-CSP test process for failures refinement as shown below.
-%%
+and each integer $p\ge 0$, we define a test for failures refinement as shown
+below. %%
 %\begin{eqnarray}
@@
 \end{eqnarray}
+
+% --------------------------------------------------------------------------
+\subsubsection*{Explanation of the Test Definition}
+A test is performed by running $U_F(p)$ concurrently with any SUT process
+$Q$, synchronising over $\Sigma$. So, a test execution is a trace of the
+concurrent process
+%\[
+$Q\parallel[\Sigma] U_F(p)$.
+%\]
+
+It is assumed that the events $\efail$ and $\epass$, indicating a verdict
+FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
+assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
+or $\epass$ always become visible, if they are the only events $U_F(p)/s$ is
+ready to engage in: if $U_F(p)/s$ can only produce $\epass$ or $\efail$, the
+occurrence of these events can never be blocked due to a livelock in $Q$
+occurring in the same step of the execution.
+
+The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and
+only if, {\it every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with
+the event $\epass$. This can also be  expressed by means of a failures
+refinement as defined below.
 %
-Before explaining the intuition behind this test definition in the paragraphs below,
-we state its basic properties in the following lemma, to show the relationships between
-$U_F(p)$ and the reference process $P$ from which the former has been derived.
+\begin{equation}
+\label{eq:passF}
+Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
+\end{equation}
+%
+This type of pass relation is often called \emph{must test} in the
+literature, because every test execution must end with the $\epass$
+event~\cite{Hennessy:1988:ATP:50497}.
+
+Note that it is necessary to use the failures-refinement relation in this
+condition, and not the trace-refinement relation:~$(Q\parallel[\Sigma]
+U_F(p)) \hide \Sigma$ may have  the same visible traces $\varepsilon$ and
+$\epass$ as the ``Test Passed Process'' $(\epass\then\Stop)$. However, the
+former may nondeterministically refuse $\epass$, due to a deadlock occurring
+when a faulty SUT process executes concurrently with $U_F(p,k,n)$ executing
+branch (\ref{eq:ufd}), when guard condition $(k = p\wedge
+\minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
+in the next paragraphs.
+
+Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
+a length $p$. If, after having already run through $s\in\trc(P)$ with $\#s
+\le p$, an event is accepted by the SUT that is outside the initials of $P/s$
+ (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
+the test immediately terminates with FAIL-event $\efail$. This is handled by
+the branch (\ref{eq:ufa}) of the external choice. %in the process
+%$U_F(p,\#s,n)$ defined above.
+
+If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice, this is
+revealed by $\minhits(G(P)/s) = \varnothing$ (recall
+(\ref{eq:minhitminaccempty}) and Lemma~\ref{lemma:ufproperties}). In this
+case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
+external choice in $U_F(p,\#s,G(P)/s)$). If $P/s$ may also
+nondeterministically engage into events in such a situation, branch
+(\ref{eq:ufc}) of the test is simultaneously enabled. If $Q/s$ is able to
+engage into an event from $\Sigma - [P/s]^0$, a test execution exists where
+$U_F(p,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the $\efail$
+event.
+
+If the length of $s$ is still less than $p$, the test accepts any event $e$
+from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
+$U_F(p,\#s+1,G(P)/s.e)$ in branch~(\ref{eq:ufc}); this follows again from
+Lemma~\ref{lemma:ufproperties}. A test of this type is called
+\emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts
+its consecutive behaviour to the event selected by the SUT.
+
+Finally, we suppose that a test execution has run through a trace
+$s\in\trc(P)$ of length $p$, so that $U_F(p)/s = U_F(p,p,n)$ with $n =
+G(P)/s$. Then, if $\minhits(n)\neq \varnothing$, the test changes its
+behaviour:~instead of offering {\it all} legal events from $[n]^0$ to the
+SUT, it nondeterministically chooses a minimal hitting set $H\in \minhits(n)$
+and only offers the events contained in $H$. If the SUT refuses to engage
+into some event of $H$, this reveals a violation of failures
+refinement:~according to Lemma~\ref{lemma:hseta}, a conforming SUT should
+accept at least one event of each minimal hitting set in $\minhits(n)$.
+Therefore, the test execution only terminates with  $\epass$, if such an
+event is accepted by the SUT. Otherwise, it deadlocks.
+
+The specification of $U_F(p,k,n)$ implies that the test always stops after
+having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$. If
+branch (\ref{eq:ufa}) is the last to be entered, the maximal length of $s$ is
+$p+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
+is the last to be entered, the maximal length of $s$ is $p$, and the
+execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
+entered, the process either accepts another event $e$ of some minimal hitting
+set $H\in\minhits(n)$ with $n = G(P)/s$ according to
+Lemma~\ref{lemma:ufproperties}. Then the final length of $s$ is $p+1$, and
+the execution terminates with $\epass$. Or the test execution
+$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
+and the execution stops without a PASS or FAIL event. Such an execution is
+also interpreted as FAIL, because it reveals that $(\epass\then\Stop)
+\not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.
+
+We observe that the number of possible executions of $Q\parallel[\Sigma]
+U_F(p)$ is finite, because the number of traces $s$ with maximal length
+$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and $\minhits(n)$
+are finite. Moreover, we further recall that $\minhits(n)$ may be empty, in
+which case the indexed internal choice in (\ref{eq:ufd}) would be undefined.
+The guard in that branch, however, requires $\minhits(n)\neq\varnothing$, and
+branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
+
+%Before explaining the intuition behind this test definition in the paragraphs
+%below, we state its basic properties in the following lemma, to show the
+%relationships between $U_F(p)$ and the reference process $P$ from which the
+%former has been derived. \fixme{For a reader, it is better to see the
+%explanation first. Since this is easier to undo, I've changed it. }
+
+We state basic properties of our tests in the following lemma, to show the
+relationships between $U_F(p)$ and the reference process $P$ from which it is
+derived.
 
@@
 \begin{proof}
-We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$, the statement holds because $U_F(p)$ starts with the initial node $\ii n$ of $G(P)$. Suppose that 
-the statement holds for all traces $s$ with length $\# s \le k < p$, so that
-$U_F(p)/s = U_F(p,\#s,G(P)/s)$. Now let $s.e$ be a trace of $P$. Then $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$, we conclude that $e\in  [G(P)/s]^0$,
-so $U_F(p)/s = U_F(p,\#s,G(P)/s)$ can engage into $e$ by executing branch (\ref{eq:ufc}).
- Since $t$ is the transition function of $G(P)$ and $e\in  [G(P)/s]^0$, $t(G(P)/s,e)$
- is defined, and $t(G(P)/s,e) = G(P)/s.e$ holds.
-This leads to a new recursion  
-$U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = U_F(p,\#s+1,G(P)/s.e)$ and proves the induction step.
+We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$,
+the statement holds because $U_F(p)$ starts with the initial node $\ii n$ of
+$G(P)$. Suppose that the statement holds for all traces $s$ with length $\# s
+\le k < p$, so that $U_F(p)/s = U_F(p,\#s,G(P)/s)$. Now let $s.e$ be a trace
+of $P$, so that $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces
+$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(p,\#s,G(P)/s)$ can
+engage into $e$ by executing branch (\ref{eq:ufc}). Since $t$ is the
+transition function of $G(P)$ and $e\in [G(P)/s]^0$, $t(G(P)/s,e)$ is
+defined, and $t(G(P)/s,e) = G(P)/s.e$. This leads to a new recursion, so that
+$U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = U_F(p,\#s+1,G(P)/s.e)$ as required.
 
-To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that 
-$U_F(p)/s = U_F(p,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
-$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so $U_F(p,\#s,G(P)/s)$
-can engage in $e$ by entering branch (\ref{eq:ufa}). The specification of this branch
-implies that 
+To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(p)/s
+= U_F(p,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
+$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
+$U_F(p,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
+specification of this branch implies that
 $$
@@
 $$
-
+%
 Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because
-$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$.
-
-Finally, statement (\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because,
-according to (\ref{eq:minhitsGP}),
-$\minhits(G(P)/s) = \minhits(P/s)$ for all traces of $P$.
-\xbox
+$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$. Finally, statement
+(\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because, according to
+(\ref{eq:minhitsGP}), $\minhits(G(P)/s) = \minhits(P/s)$ for all traces of
+$P$. \xbox
 \end{proof}
-
-Note that it is not guaranteed for $U_F(p)$ to run through the traces $s, s.e$ specified 
-in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for some prefix $u$ of $s$: in such a case, $U_F(p)$ may stop with a $\epass$-event by entering branch (\ref{eq:ufb}).
-Therefore, Lemma~\ref{lemma:ufproperties} just states the existence of
+%
+Note that it is not guaranteed for $U_F(p)$ to run through the traces $s,
+s.e$ in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for
+some prefix $u$ of $s$: in such a case, $U_F(p)$ may stop with a $\epass$
+event by entering branch (\ref{eq:ufb}). Therefore,
+Lemma~\ref{lemma:ufproperties} just states the existence of
 $U_F(p)$-executions $s, s.e$ satisfying the properties stated there.
-
-% --------------------------------------------------------------------------
-\subsubsection*{Explanation of the Test Definition}
-A test is performed by running $U_F(p)$ concurrently with any SUT process
-$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
-the concurrent process
-\[
-Q\parallel[\Sigma] U_F(p).
-\]
-It is assumed that the events $\efail$ and $\epass$, indicating a verdict
-FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
-assume that $Q$ is free of livelocks, it is guaranteed that events 
-$\efail$ or $\epass$ always become visible, if they are the only events
-$U_F(p)/s$ is ready to engage in: if $U_F(p)/s$ can only produce $\epass$ or
-$\efail$, the occurrence of these events can never be blocked due to a livelock in $Q$
-occurring in the same step of the execution.
-
-The test is
-\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
-every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
-$\epass$. This can also be  expressed by means of a failures refinement as
-defined below.
-%
-\begin{equation}
-\label{eq:passF}
-Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
-\end{equation}
-%
-This type of pass relation is often called \emph{must test}, because every
-test execution must end with the $\epass$
-event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
-failures-refinement relation in this condition, and not the trace-refinement
-relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
-visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
-Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
-refuse $\epass$, due to a deadlock occurring when a faulty SUT process
-executes concurrently with $U_F(p,k,n)$ executing branch (\ref{eq:ufd}),
-when guard condition $(k = p\wedge \minhits(n)\neq\varnothing)$ evaluates to $\ist$. 
-This is explained further in the next paragraphs.
-
-The specification of $U_F(p,k,n)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
-If branch (\ref{eq:ufa})
-is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
-stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
-is the last to be entered, the process either accepts another event
-$e$ of some minimal hitting set $H\in\minhits(n)$ with $n = G(P)/s$ according to
-Lemma~\ref{lemma:ufproperties}. Then the final length
-of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
-$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
-and the execution stops without a PASS or FAIL event.
-Such an execution is also interpreted as FAIL, because it reveals that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.
-
-We observe that the number of possible executions of $Q\parallel[\Sigma]
-U_F(p)$ is finite, because the number of traces $s$ with maximal length
-$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and
-$\minhits(n)$ are finite. We further recall that $\minhits(n)$ may be
-empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
-is undefined. Therefore, the associated guard condition contains the requirement
-$\minhits(n)\neq\varnothing$, otherwise this branch cannot be taken,
-but branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
-%%%\fixme{alcc: Oops. This is undefined in CSP.}
-
-Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
-a length $p$. If, after having already run through $s\in\trc(P)$ with
-$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$
- (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
-the test immediately terminates with FAIL-event $\efail$. This is handled by
-the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,\#s,n)$
-defined above.
-
-If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
-this is revealed by $\minhits(G(P)/s) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
-Lemma~\ref{lemma:ufproperties}). 
-In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
-the external choice in $U_F(p,\#s,G(P)/s)$). 
-If $P/s$ may also nondeterministically engage into events in such a situation, branche 
- (\ref{eq:ufc}) of the test is
-simultaneously enabled. If $Q/s$ is able to engage into an event from 
-$\Sigma - [P/s]^0$, a test execution exists where $U_F(p,\#s,G(P)/s)$ branches
-into (\ref{eq:ifpb}) and produces the $\efail$-event.
-
-If the length of $s$ is still less than $p$, the test accepts any event $e$ from
-the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as $U_F(p,\#s+1,G(P)/s.e)$ in
-branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}. 
-A test of this type is called \emph{adaptive}, because
-it accepts any legal behaviour of the SUT and adapts its consecutive
-behaviour to the event selected by the SUT.
-
-Suppose that a test execution has run through a trace 
-$s\in\trc(P)$ of length $p$, so that $U_F(p)/s = U_F(p,p,n)$ with $n = G(P)/s$. 
-Then,   if
-$\minhits(n)\neq \varnothing$, the test changes its behaviour: instead of
-offering {\it all} legal events from $[n]^0$ to the SUT, it
-nondeterministically chooses a minimal hitting set   $H\in \minhits(n)$ 
-and
-only offers the events contained in $H$. If the SUT refuses to engage
-into any event of $H$, this reveals a violation of failures refinement:
-according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
-one event of each minimal hitting set in $\minhits(n)$. Therefore, the test
-execution
-only terminates with  $\epass$, if such an event is accepted by the SUT.
-
-%% =========================================================================
-%\subsection{Fault Models and Complete Test Suites}
 
@@
 \subsection{A Finite Complete Test Suite for Failures Refinement}
+% =========================================================================
 
@@
 reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
-\lessdet_F\}$, and a fault domain ${\cal D}$ which is a set of CSP processes
+\lessdet_F\}$, and a fault domain ${\cal D}$, which is a set of CSP processes
 over $P$'s alphabet that may or may not conform to $P$.
 
-A test suite $\TS$ is called \emph{complete} with respect to fault model ${\cal F}$,
-if and only if the following conditions are fulfilled.
+A test suite $\TS$ is called \emph{complete} with respect to fault model
+${\cal F}$, if, and only if, the following conditions are fulfilled.
 \begin{description}
@@
 %
-In the sequel, we establish the completeness of our test suite which is stated 
-in the following main theorem.
-%%%\fixme{alcc: I don't think we need a section just for this material.}
+The following main theorem establishes the completeness of our test suite.
 
@@
 \begin{lemma}\label{lemma:mainfsound}
-Test suite $\TS_F$ generated from a CSP process $P$, as specified in
+A test suite $\TS_F$ generated from a CSP process $P$, as specified in
 Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
@@
 \end{lemma}
-\begin{proof}{~}\newline
+\begin{proof}{~}We make two points, described in separate steps below. The
+first is that the test execution cannot reach branch~(\ref{eq:ufa})and raise
+a $fail$ event.  The second is that it cannot deadlock without raising a
+$pass$ event.
+
 {\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $Q\ conf\
-P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
-any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
+P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$, any
+adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
 branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
-choice construction for $U_F(p,k,n)$:  
-If $U_F(p,k,n) = U_F(p)/s$ with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties}
-implies $U_F(p,k,n) = U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As
-a consequence, $[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
-entered in the parallel execution of $Q$ and $U_F(p)$.
+choice construction for $U_F(p,k,n)$. If $U_F(p,k,n) = U_F(p)/s$ with
+$s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties} implies $U_F(p,k,n) =
+U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
+$[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
+entered in the parallel execution of $Q$ and $U_F(p)$, and the $fail$ event
+cannot occur.
 
 \medskip
-\noindent
-{\bf Step~2.} Sinc $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies that for all  
-$s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a hitting set for
-$\minaccs(Q/s)$. 
-From Lemma~\ref{lemma:ufproperties} we know that
-$U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, 
-so $\minhits(n) = \minhits(P/s)$. 
-Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$ leads always to a
-PASS verdict and is taken if $\minhits(n) = \varnothing$. Branch (\ref{eq:ufc}) 
-always leads to test continuation without a
-verdict. For the last branch, we note that it can only be entered if $\minhits(P/s)
-= \minhits(n) \neq\varnothing$. In this case,
-any selected minimal hitting set
+\noindent {\bf Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
+that for all $s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a
+hitting set for $\minaccs(Q/s)$. From Lemma~\ref{lemma:ufproperties} we know
+that $U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, so
+$\minhits(n) = \minhits(P/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$
+leads always to a PASS verdict and is taken if $\minhits(n) = \varnothing$.
+Branch (\ref{eq:ufc}) always leads to test continuation without a verdict.
+For the last branch, we note that it can only be entered if $\minhits(P/s) =
+\minhits(n) \neq\varnothing$. In this case, any selected minimal hitting set
 $H\in\minhits(n)$ has a non-empty intersection with each of the minimal
-acceptances of $Q/s$. 
-As a consequence, $Q/s$ never blocks when offered
-events from $H\in\minhits(n)$, 
-and the test terminates with PASS event $\epass$. Note that
-this argument requires that $Q$ is free of livelocks, because otherwise the
-$\epass$-events might not become visible, due to unbounded sequences of
-hidden events performed by $Q$. 
-\xbox
+acceptances of $Q/s$. As a consequence, $Q/s$ never blocks when offered
+events from $H\in\minhits(n)$, and the test terminates with PASS event
+$\epass$. Note that this argument requires that $Q$ is free of livelocks,
+because otherwise the $\epass$-events might not become visible, due to
+unbounded sequences of hidden events performed by $Q$. \xbox
 \end{proof}
@@
 \medskip
-\noindent
-{\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$, but
-$s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is a
-trace of every process. In this case, $s$ is also a trace of the product
-graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}, and  
-$G/s = (G(P)/s,G(Q)/s)$ holds. The length of $s$ is not known, but from the construction
-of $G$,  we know that $G$ has at most $pq$ reachable states, because $G(P)$
-has $p$ states, and $G(Q)$ has at most $q$ states. By
+\noindent {\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
+but $s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is
+a trace of every process. In this case, $s$ is also a trace of the product
+graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}, and $G/s =
+(G(P)/s,G(Q)/s)$ holds. The length of $s$ is not known, but from the
+construction of $G$,  we know that $G$ has at most $pq$ reachable states,
+because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states. By
 Lemma~\ref{lemma:reachproduc}, $(G(P)/s,G(Q)/s)$ can be reached by a trace
 $u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
-function of $G$ implies that $u$ is also a trace of $P$ and $Q$, 
-which means that 
-$(G(P)/s,G(Q)/s) = (G(P)/u,G(Q)/u)$. Since test
-$U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a
-trace of this test, and, by construction and by Lemma~\ref{lemma:ufproperties}, 
-$U_F(pq-1)/u = U_F(pq-1,\#u,G(P)/u)$. Since
-$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. 
-Hence, in at least
-one execution, $U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with
-this event $e$, so that the test fails. Again, the
-assumption of non-divergence of Q is needed for this conclusion.
+function of $G$ implies that $u$ is also a trace of $P$ and $Q$, which means
+that $(G(P)/s,G(Q)/s) = (G(P)/u,G(Q)/u)$. Since test $U_F(pq-1)$ accepts all
+traces of $P$ up to length $pq-1$, $u$ is also a trace of this test, and, by
+construction and by Lemma~\ref{lemma:ufproperties}, $U_F(pq-1)/u =
+U_F(pq-1,\#u,G(P)/u)$. Since $s.e\not\in\trc(P)$, $e$ is an element of
+$\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. Hence, in at least one execution,
+$U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with this
+event $e$, so that the test fails. Again, the assumption of non-divergence of
+$Q$ is needed for this conclusion.
 

### 2018-09-28 paper/preliminaries.tex
changed order of material

@@
 sets~(Section~\ref{sec:hit}), which is central to our novel notion of test
-for failures refinement.
+for failures refinement. To study complexity, we also introduce the concept
+of Sperner families~(Section~\ref{sec:sperner}).
 
@@
 
-\fixme{Abrupt end to the section. It has more than just preliminary material.
-It has some new results.}
-
 % =========================================================================
@@
 \subsubsection*{Definition}
-The main idea of the underlying test strategy for failures refinement can be
+The main idea of the underlying test strategy for failures refinement is
 based on solving a \emph{hitting set problem}. Given a finite collection of
@@
 $\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
-intersection with $A$.
-Minimising $\overline H$ yields   a minimal hitting set $H\in
-\minhits(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
-contradiction to Assumption~2. This completes the proof of the lemma. \xbox
+intersection with $A$. Minimising $\overline H$ yields   a minimal hitting
+set $H\in \minhits(P/s)$ which is {\it not} a hitting set of
+$\minaccs(G(Q)/s)$, a contradiction to Assumption~2. We note that we use
+$\minhits(G(P)/s) = \minhits(P/s)$, as well $\minaccs(G(Q)/s) =
+\minaccs(Q/s)$, which follow from the identification of process states with
+states of the normalised graph for the proof discussed in the previous
+section. This completes the proof of the lemma. \fixme{Please, check that
+you're happy with this observation.}\xbox
 \end{proof}
@@
 \end{equation}
+\fixme{You already used this above. Perhaps you want to move this to before
+Lemma 4?}
 
@@
 \label{sec:sperner}
-In preparation of complexity results to be presented in Section~\ref{sec:complexity}, we consider the question how many minimal hitting sets can maximally exist for a given
-collection of minimal acceptances.
-To this end, the following definitions and results are
-useful\footnote{The authors are grateful to Li-Da Tong from National Sun Yat-sen University, Taiwan,
-for suggesting the applicability of Sperner's Theorem in the context of the work presented here.}.
+% =========================================================================
 
-A \emph{Sperner Family} is a collection $S \subseteq 2^\Sigma$ of sets from a given
-finite universe $\Sigma$ that do not contain each other, that is,
-$H_1 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair $H_1\neq H_2\in S$.
-Specialising   antichains known from   partial orders to finite sets partially ordered by $\subseteq$ results in Sperner families.
-We observe that
+In preparation for complexity results to be presented in
+Section~\ref{sec:complexity}, we consider how many minimal hitting sets can
+maximally exist for a given collection of minimal acceptances. To this end,
+the following definitions and results are useful\footnote{The authors are
+grateful to Li-Da Tong from National Sun Yat-sen University, Taiwan, for
+suggesting the applicability of Sperner's Theorem in the context of the work
+presented here.}.\fixme{Shouldn't the footnote go to the acknowledgements?}
+
+A \emph{Sperner Family} is a collection $S \subseteq 2^\Sigma$ of sets from a
+given finite universe $\Sigma$ that do not contain each other, that is, $H_1
+\not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair $H_1\neq
+H_2\in S$. Specialising antichains known from partial orders to finite sets
+partially ordered by $\subseteq$ results in Sperner families. We observe that
 \begin{itemize}
@@
 \end{itemize}
-represent Sperner families. Moreover, given any finite alphabet $\Sigma$ with
-$\card{\Sigma} = n$, every collection $S$ of subsets with identical cardinality $k \le n$
-is a Sperner family,  because $A_1, A_2\in S \wedge A_1 \subseteq A_2\wedge \card{A_1} = \card{A_2}$ implies $A_1 = A_2$. Given  any Sperner Family $S$ of $\Sigma$, $S$ represents
-the minimal acceptances in the initial state of the CSP process
-\[
-P = \Intchoice_{A\in S}\big( e:A \then P(e) \big).
-\]
+are Sperner families. Moreover, given any finite alphabet $\Sigma$ with
+$\card{\Sigma} = n$, every collection $S$ of subsets with identical
+cardinality $k \le n$ is a Sperner family,  because $A_1, A_2\in S \wedge A_1
+\subseteq A_2\wedge \card{A_1} = \card{A_2}$ implies $A_1 = A_2$. Given  any
+Sperner Family $S$ of $\Sigma$, $S$ represents the minimal acceptances in the
+initial state of the CSP process
+%\[
+$P = \Intchoice_{A\in S}\big( e:A \then P(e) \big)$.
+%\]
 
@@
 \]
-The upper bound is reached if and only if one of the following cases apply.
+The upper bound is reached if, and only if, one of the following cases apply:
 \begin{enumerate}
-\item For even $n$, if $S$ consists of all subsets of $\Sigma$ with cardinality $n/2$.
-\item For odd $n$, if one of the following cases holds.
+\item For even $n$, if $S$ consists of all subsets of $\Sigma$ with
+    cardinality $n/2$;
+\item For odd $n$, if one of the following cases holds;
 \begin{enumerate}
-\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/s$.
+\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/s$;
+    or
 \item $S$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/s$.
@@
 \end{theorem}
-It will be shown in Section~\ref{sec:complexity} that this upper bound can actually be reached by hitting sets associated with the minimal acceptances of a CSP process state.
+It is shown in Section~\ref{sec:complexity} that this upper bound can
+actually be reached by hitting sets associated with the minimal acceptances
+of a CSP process state.

### 2018-09-30 paper/finitecomplete-failures.tex
concern about a proof

@@
 \end{lemma}
-\begin{proof}{~}We make two points, described in separate steps below. The
+\begin{proof}{~}We make two points in separate steps below. The
 first is that the test execution cannot reach branch~(\ref{eq:ufa})and raise
 a $fail$ event.  The second is that it cannot deadlock without raising a
-$pass$ event.
+$pass$ event. This case would also interpreted as FAIL, since then
+$\epass\then\Stop$ is not failures-refined by $(Q\parallel[\Sigma]
+U_F(k))\hide \Sigma$.
+% alcc: part of the explanation above was in the following lemma, but it is already
+% relevant here.
 
-{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $Q\ conf\
-P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$, any
-adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
-branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
-choice construction for $U_F(p,k,n)$. If $U_F(p,k,n) = U_F(p)/s$ with
-$s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties} implies $U_F(p,k,n) =
+\paragraph{Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and $Q\
+conf\ P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq
+\trc(P)$, any adaptive test $U_F(p)$ running in parallel with $Q$ will always
+enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the
+external choice construction for $U_F(p,k,n)$. If $U_F(p,k,n) = U_F(p)/s$
+with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties} implies $U_F(p,k,n) =
 U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
@@
 
-\medskip
-\noindent {\bf Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
+\paragraph{Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
 that for all $s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a
@@
 that $U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, so
-$\minhits(n) = \minhits(P/s)$. Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$
-leads always to a PASS verdict and is taken if $\minhits(n) = \varnothing$.
-Branch (\ref{eq:ufc}) always leads to test continuation without a verdict.
-For the last branch, we note that it can only be entered if $\minhits(P/s) =
-\minhits(n) \neq\varnothing$. In this case, any selected minimal hitting set
-$H\in\minhits(n)$ has a non-empty intersection with each of the minimal
-acceptances of $Q/s$. As a consequence, $Q/s$ never blocks when offered
-events from $H\in\minhits(n)$, and the test terminates with PASS event
-$\epass$. Note that this argument requires that $Q$ is free of livelocks,
-because otherwise the $\epass$-events might not become visible, due to
-unbounded sequences of hidden events performed by $Q$. \xbox
+$\minhits(n) = \minhits(P/s)$. The branch (\ref{eq:ufb}) of the test
+$U_F(p,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
+\varnothing$. The branch (\ref{eq:ufc}) always leads to test continuation
+without a verdict. \fixme{I think we need a stronger argument here, as to why
+it cannot deadlock.} For the last branch, we note that it can only be entered
+if $\minhits(P/s) = \minhits(n) \neq\varnothing$. In this case, any selected
+minimal hitting set $H\in\minhits(n)$ has a non-empty intersection with each
+of the minimal acceptances of $Q/s$. As a consequence, $Q/s$ never blocks
+when offered events from $H\in\minhits(n)$, and the test terminates with
+event $\epass$. Note that this argument requires that $Q$ is free of
+livelocks, because otherwise the $\epass$-events might not become visible,
+due to unbounded sequences of hidden events performed by $Q$. \xbox
 \end{proof}
@@
 \begin{lemma}\label{lemma:mainfexhaustive}
-Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive
-for the fault model specified there.
+A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
+exhaustive for the fault model specified there.
 \end{lemma}
 \begin{proof}
-Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$, According to
+Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
 (\ref{eq:failconf}), this non-conformance can be caused in two possible ways
-corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$,
-respectively:
+corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
+cases can be characterised as follows:
 \begin{description}
@@
 \end{description}
-It has to be shown for each of the two cases that at least one test
-execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
-   $\efail$-event or without giving any verdict. The latter case is also
-interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
-failures-refined by $(Q\parallel[\Sigma] U_F(k))\hide \Sigma$.
+It has to be shown for each of these cases that at least one test execution
+of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the $\efail$
+event or deadlocks.
 
@@
 event $e$, so that the test fails. Again, the assumption of non-divergence of
-$Q$ is needed for this conclusion.
+$Q$ is needed for this conclusion. \fixme{There is an assumption here that
+the graph of a test execution is given by the cross-product of graphs.  But,
+test execution is not defined in this way. It is defined by parallelism and
+hiding. I think the definition of test execution needs to change.}
 

### 2018-09-30 paper/finitecomplete-failures.tex
more about the concern

@@
 $s\in\trc(P)$ of length $p$, so that $U_F(p)/s = U_F(p,p,n)$ with $n =
-G(P)/s$. Then, if $\minhits(n)\neq \varnothing$, the test changes its
+G(P)/s$. If $\minhits(n)\neq \varnothing$, the test changes its
 behaviour:~instead of offering {\it all} legal events from $[n]^0$ to the
@@
 \medskip
-\noindent
-{\bf Case~2.} We note that trace $s$ is again a trace of the product graph
-$G$, but we do not know its length. Again, by applying Lemma~\ref{lemma:reachproduc},
-we know that the state   $G/s$ can   be reached by a trace
-$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#
-u)$, which satisfies $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$   always performs
-branch (\ref{eq:ufc}) until the trace $u$ has been completely processed.
+\noindent {\bf Case~2.} We note that trace $s$ is again a trace of the
+product graph $G$, but we do not know its length. Again, by applying
+Lemma~\ref{lemma:reachproduc}, we know that the state $G/s$ can be reached by
+a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider
+the test $U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because
+of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$   always performs branch
+(\ref{eq:ufc}) until the trace $u$ has been completely processed.
 $U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
@@
 at least one non-empty minimal acceptance. From (\ref{eq:minhitminaccempty})
-we know that this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and
-we observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$.
-As a consequence, branch (\ref{eq:ufb}) cannot be taken because its guard condition
-evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$.
-The guard condition $(k < p)$ for branch (\ref{eq:ufc}) evaluates
-to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the
-test always fails. If branch (\ref{eq:ufd}) is executed, the test fails for
-the execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen
-by $U_F(\#u,\#u,G(P)/u)$ that has an empty intersection with the minimal acceptance
-$A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an
-$H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there
-exists a test execution   where $Q/u$ selects acceptance $A_Q$ and
-$U_F(\#u,\#u,G(P)/u)$ selects $H$. This execution deadlocks in process state
+we know that this is equivalent to $\minhits(P/s) =
+\minhits(G(P)/s)\neq\varnothing$, and we observe that $G(P)/s = G(P)/u$, so
+$\minhits(G(P)/u)\neq\varnothing$. As a consequence, branch (\ref{eq:ufb})
+cannot be taken because its guard condition evaluates to $\isf$  for
+$U_F(\#u,\#u,G(P)/u)$. The guard condition $(k < p)$ for branch
+(\ref{eq:ufc}) evaluates to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch
+(\ref{eq:ufa}) is executed, the test always fails. If branch (\ref{eq:ufd})
+is executed, the test fails for the execution where a minimal hitting set
+$H\in\minhits(P/u)$ is chosen by $U_F(\#u,\#u,G(P)/u)$ that has an empty
+intersection with the minimal acceptance $A_Q$ from condition
+(\ref{eq:accsnotcontained}). The existence of such an $H$ is guaranteed
+because of Lemma~\ref{lemma:hseta}. As a consequence, there exists a test
+execution   where $Q/u$ selects acceptance $A_Q$ and $U_F(\#u,\#u,G(P)/u)$
+selects $H$. This execution deadlocks in process state
 $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event;
-this  means that the test fails and concludes the proof.
-\xbox
+\fixme{Again, nothing here is about this process, but about the graph, and
+the hiding that is needed is discarded.} this means that the test fails. This
+concludes the proof. \xbox
 \end{proof}

### 2018-09-30 paper/complexity.tex
concerns about comparison of complexity

@@
 
-As can be seen from the specification of the test cases $U_F(k)$ 
-(see Section~\ref{sec:finitecompletefails}), the number of
-executions ending in a $\epass$-event corresponds to the number $\ell$ of traces $s$
-of $P$ with length equal to\footnote{In this estimation, we disregard
-the case where the test
-terminates earlier due to entering branch (\ref{eq:ufb}).} $k$,
-multiplied by the number $h$ of minimal hitting sets in
-$\minhits(P/s)$. For the tests $U_T(k)$ verifying trace refinement (see Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$, since
-there is no equivalent in $U_T(k)$ 
-to checking different hitting sets in the last step of a 
-test execution.
+As can be seen from the specification of the test cases $U_F(k)$ (see
+\fixme{In that section, $k$ is used for a different purpose. It is confusing.
+I would use $p$ here.} Section~\ref{sec:finitecompletefails}), the number of
+executions ending in a $\epass$ event corresponds to the number $\ell$ of
+traces $s$ of $P$ with length equal to\footnote{In this estimation, we
+disregard the case where the test terminates earlier due to entering branch
+(\ref{eq:ufb}).} $k$, multiplied by the number $h$ of minimal hitting sets in
+$\minhits(P/s)$.\fixme{But doesn't this add to the number of executions?} For
+the tests $U_T(k)$ verifying trace refinement (see
+Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$,
+since there is no equivalent in $U_T(k)$ to checking different hitting sets
+in the last step of a test execution.
 
@@
 \subsection{Estimation of $\ell$}
-The first factor $\ell$
-has worst-case upper bound $\ell\le \card{\Sigma}^k$. As an example, where this upper bound 
-is really met, consider the reference process
+The first factor $\ell$ has worst-case upper bound $\ell\le \card{\Sigma}^k$.
+As an example where this upper bound is really met, we consider the reference
+process
 \[
@@
 This follows from Theorem~\ref{th:sperner}, since $\minhits(P/s)$ is a
-Sperner Family, as introduced in Section~\ref{sec:sperner}. 
-The next theorem shows that this
-upper bound can really be reached by collections of minimal hitting sets associated with 
-a CSP process state.
+Sperner Family~(see Section~\ref{sec:sperner}). The next theorem shows that
+this upper bound can really be reached by collections of minimal hitting sets
+associated with a CSP process state.
 
@@
 Let $C$ be the collection of all subsets of $\Sigma$ with cardinality
-$n-\lfloor{\frac{n}{2}}\rfloor+1$. Then any CSP process of the form 
+$n-\lfloor{\frac{n}{2}}\rfloor+1$. Then the CSP process
 \[
@@
 \]
-fulfils $\minaccs(P) = C$.
-Let $H$ be any minimal Hitting set of $C$. Then $H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because 
-otherwise
-$\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, 
-and any subset $A\subseteq \Sigma\setminus H$ with cardinality
+fulfils $\minaccs(P) = C$. Let $H$ be any minimal hitting set of $C$. Then
+$H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because
+otherwise $\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, and any
+subset $A\subseteq \Sigma\setminus H$ with cardinality
 $n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in $C$, but satisfy
-$A\cap H=\varnothing$. 
-Since $\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we conclude
+$A\cap H=\varnothing$. Since
+$\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we
+conclude
  that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
@@
 \end{proof}
-
-To get an approximation of the maximal size of $\minhits(P/s)$ for large $n$, recall
-Stirling's approximation~\cite[p.~112]{Graham:1994:CMF:562056}
+%
+To get an approximation of the maximal size of $\minhits(P/s)$ for large $n$,
+recall Stirling's approximation~\cite[p.~112]{Graham:1994:CMF:562056}
 \[
@@
 \]
-Applying this approximation to the   case where $n = \card{\Sigma}$ is even and $m = (n/2)$  
+Applying this approximation when $n = \card{\Sigma}$ is even and $m = (n/2)$
 results in
@@
 \subsection{Upper Bounds of Test Executions for Checking Failures Refinement}
-According to  Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
+
+By Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
 $k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
@@
 above cannot be further improved, since the full collection of minimal
-hitting sets needs to be checked to verify $conf$. The
-authors of~\cite{Hennessy:1988:ATP:50497} and
+hitting sets needs to be checked to verify $conf$.\fixme{Theorem 4 is not
+about $conf$. I don't think you have tests for $conf$.}
+
+The authors of~\cite{Hennessy:1988:ATP:50497} and
 \cite{DBLP:conf/icfem/CavalcantiG07} suggest to test {\it every} non-empty
@@
 process state of the reference model; this leads to a worst-case estimate of
-$2^{\card{\Sigma}}-1$ for the number of different sets to be offered to the SUT in
-the last step of the test execution, so our approach reduces the number of
-test executions in comparison
-to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} by a factor of
-$1/{\sqrt{2\pi \card{\Sigma}}}$.
-
-In Fig.~\ref{fig:minhita}, the reduction is visualised by a function plot of
-$2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.
-
+$2^{\card{\Sigma}}-1$ for the number of different sets to be offered to the
+SUT in the last step of the test execution, so our approach reduces the
+number of test executions in comparison
+to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} by a factor
+of $1/{\sqrt{2\pi \card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction
+is visualised by a function plot of $2^{\card{\Sigma}}$ versus
+$\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$. \fixme{In [2], we work
+with minimal acceptances. I didn't check [8].}
 % .....................................................................................
@@
 According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
-refinement just contains the adaptive test case $U_T(pq-1)$. As derived for $U_F(k)$
-above, the number of executions performed by $(Q\parallel[\Sigma] U_R(pq-1))$ 
-is bounded by $\card{\Sigma}^{pq-1}$.
+refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
+$U_F(k)$ above, the number of executions performed by $(Q\parallel[\Sigma]
+U_R(pq-1))$ is bounded by $\card{\Sigma}^{pq-1}$. \fixme{Please, see comments
+above about this process.}
 
@@
 
-According to Theorem~\ref{th:failurestest}, the tests $U_F(k)$ need to be executed for 
-$k = 0,\dots,pq-1$ to guarantee completeness. This means that the SUT is verified 
-with test traces up to, and including, length $pq$:
-recall from the test specification, branch (\ref{eq:ufa}), that $U_F(k)$ will accept
-all traces $s.e$ with $s\in\trc(P), \#s = k, e\not\in\trc(P/s)$, so erroneous trace
-up to length $k+1$ are detected.
-It is interesting to  investigate
-whether this maximal length is really necessary, or whether one could elaborate
-alternative complete test strategies where the SUT is tested with shorter traces only. 
-Indeed, an example  presented in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} 
-shows that when testing for equivalence of deterministic FSMs, it is sufficient to
-test the SUT with traces of significantly shorter length.
+According to Theorem~\ref{th:failurestest}, the tests $U_F(k)$ need to be
+executed for $k = 0,\dots,pq-1$ to guarantee completeness. This means that
+the SUT is verified with test traces up to, and including, length $pq$:
+recall from the test specification, branch (\ref{eq:ufa}), that $U_F(k)$ will
+accept all traces $s.e$ with $s\in\trc(P), \#s = k, e\not\in\trc(P/s)$, so
+erroneous trace up to length $k+1$ are detected. It is interesting to
+investigate whether this maximal length is really necessary, or whether one
+could elaborate alternative complete test strategies where the SUT is tested
+with shorter traces only. Indeed, an example  presented
+in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} shows that when testing for
+equivalence of deterministic FSMs, it is sufficient to test the SUT with
+traces of significantly shorter length.
 
@@
 \begin{example}\label{ex:pq}
-Consider the CSP reference process $P$ and an erroneous implementation $Q$ specified
-as follows.
+Consider the CSP reference process $P$ and an erroneous implementation $Q$
+specified as follows.
 \begin{eqnarray*}

### 2018-09-30 paper/finitecomplete-failures.tex
concerns about comparison of complexity

@@
 
-Note that it is necessary to use the failures-refinement relation in this
-condition, and not the trace-refinement relation:~$(Q\parallel[\Sigma]
-U_F(p)) \hide \Sigma$ may have  the same visible traces $\varepsilon$ and
-$\epass$ as the ``Test Passed Process'' $(\epass\then\Stop)$. However, the
-former may nondeterministically refuse $\epass$, due to a deadlock occurring
-when a faulty SUT process executes concurrently with $U_F(p,k,n)$ executing
-branch (\ref{eq:ufd}), when guard condition $(k = p\wedge
+We note that it is necessary to use failures refinement in the definition
+above, and not just trace refinement:~$(Q\parallel[\Sigma] U_F(p)) \hide
+\Sigma$ may have the same visible traces $\varepsilon$ and $\epass$ as the
+``Test Passed Process'' $(\epass\then\Stop)$. However, the former may
+nondeterministically refuse $\epass$, due to a deadlock occurring when a
+faulty SUT process executes concurrently with $U_F(p,k,n)$ executing
+branch~(\ref{eq:ufd}), when the guard condition $(k = p\wedge
 \minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
@@
 $pass$ event. This case would also interpreted as FAIL, since then
-$\epass\then\Stop$ is not failures-refined by $(Q\parallel[\Sigma]
+$\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
 U_F(k))\hide \Sigma$.
@@
 \end{proof}
+%
+Our notion of tests can be specialised to deal with traces refinement, as we
+explain next. Examples come later in Section~\ref{sec:case}. \fixme{I think
+the example should come before more theory.}
 

### 2018-09-30 paper/finitecomplete-trace.tex
concerns about comparison of complexity

@@
 $$
-In contrast to the failures tests (\ref{eq:UFP}), however, we do not need to
-check the SUT with respect to its acceptance of hitting sets. Therefore, these
-do not occur in the specification of the test cases below, and we use the
-condition $\minaccs(n) = \{ \varnothing \}$ instead of $\minhits(n) = \varnothing$
-to indicate that minimal hitting sets need not to be calculated for generating
-these tests from $G(P)$. From (\ref{eq:minhitminaccempty}) we know that both conditions
-are equivalent.
+In contrast to the tests for failures refinement~(\ref{eq:UFP}), however, we
+do not need to check the SUT with respect to its acceptance of hitting sets.
+Therefore, these do not occur in the specification of the test cases below,
+and we use the condition $\minaccs(n) = \{ \varnothing \}$ instead of
+$\minhits(n) = \varnothing$ to indicate that minimal hitting sets need not to
+be calculated for generating these tests from $G(P)$. From
+(\ref{eq:minhitminaccempty}) we know that these conditions are equivalent.
 
-%
 %\begin{eqnarray}
@@
 \end{eqnarray}
-
+%
 It is easy to see that the tests $U_T(p)$ satisfy the properties
@@
 \end{eqnarray}
-proven in Lemma~\ref{lemma:ufproperties} for $U_F(p)$ for traces $s\in\trc(P)$.
+proven in Lemma~\ref{lemma:ufproperties} for $U_F(p)$ for traces
+$s\in\trc(P)$. \fixme{In (37), I think there is an implicit assumption about
+$p$ and $\# s$?}
 
-%
 %The difference between adaptive tests $U_T(p)$ for trace refinement and
@@
 suite for trace refinement just needs the single adaptive test case
-$U_T(pq-1)$, while failures refinement required the execution of $\{
-U_F(0),\dots,U_F(pq-1)\}$. The reason for this is that $U_T(pq-1)$ identifies
-trace errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes
-for erroneous acceptances at the end of each trace of length $(pq -1)$.
-Since $U_T(pq-1)$ never blocks any $Q$-event before terminating, the pass 
-criterion can be based on trace refinement instead of failures refinement
-as required in (\ref{eq:passF}).
+$U_T(pq-1)$, while failures refinement requires the execution of $\{
+U_F(0),\dots,U_F(pq-1)\}$. The reason is that $U_T(pq-1)$ identifies trace
+errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for
+erroneous acceptances at the end of each trace of length $(pq -1)$. Since
+$U_T(pq-1)$ never blocks any $Q$ event before terminating, the pass criterion
+can be based on trace refinement instead of failures refinement as required
+in (\ref{eq:passF}).
 %
@@
 \end{equation}
+\fixme{The proofs don't use this definition. It should change.}
 %
-
 \begin{theorem}\label{th:tracetest}
@@
 \end{proof}
+%
+Examples are provided after our discussion of size of the test suites.

### 2018-09-30 paper/introduction.tex
concerns about comparison of complexity

@@
 nondeterminism in a more fine-grained way than it is possible for FSMs
-requires a more complex approach to testing for conformity in the
-failures-model than required for model-based testing against nondeterministic
-FSMs, as published, for example,
+requires a more complex approach to testing for conformity in the failures
+model than required for model-based testing against nondeterministic FSMs, as
+published, for example,
 in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. CSP distinguishes

### 2018-09-30 paper/casestudies.tex
finished review

@@
 to $P$. While $k < r_{max}$, $Z$ also accepts the same sets of events as $P$.
-When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$ is
-fulfilled, however, $R_1(r_{max},k)$ makes an internal choice, instead of
-offering an external choice, so $P\not\lessdet_F Z$.
-Fig.~\ref{fig:tgZ} shows the normalised transition graph of $Z$ for $r_{max}
-= 3$.
+When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$,
+however, $R_1(r_{max},k)$ makes an internal choice, instead of offering an
+external choice, so $P\not\lessdet_F Z$. Fig.~\ref{fig:tgZ} shows the
+normalised transition graph of $Z$ for $r_{max} = 3$.
 
@@
 
-Running the test $U_F(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p =
-4$ states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test
-depth to be used according to Theorem~\ref{th:failurestest}), tests
-$U_F(0),\dots, U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$,
-because after  execution trace
+Running the test $U_F(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p = 4$
+states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test depth
+to be used according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
+U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$, because after execution of
+the trace
 \[
@@
 \]
-and the test fails.
-\xbox
+\fixme{The hiding is missing.} and the test fails. \xbox
 \end{example}

### 2018-09-30 paper/complexity.tex
finished review

@@
 According to Theorem~\ref{th:failurestest}, the tests $U_F(k)$ need to be
-executed for $k = 0,\dots,pq-1$ to guarantee completeness. This means that
-the SUT is verified with test traces up to, and including, length $pq$:
-recall from the test specification, branch (\ref{eq:ufa}), that $U_F(k)$ will
-accept all traces $s.e$ with $s\in\trc(P), \#s = k, e\not\in\trc(P/s)$, so
-erroneous trace up to length $k+1$ are detected. It is interesting to
-investigate whether this maximal length is really necessary, or whether one
-could elaborate alternative complete test strategies where the SUT is tested
-with shorter traces only. Indeed, an example  presented
-in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} shows that when testing for
-equivalence of deterministic FSMs, it is sufficient to test the SUT with
-traces of significantly shorter length.
+executed for $k = 0,\dots,pq-1$ to guarantee completeness. \fixme{Confusing
+use of $p$ and $q$ again.} This means that the SUT is verified with test
+traces up to, and including, length $pq$: recall from the test specification,
+branch (\ref{eq:ufa}), that $U_F(k)$ will accept all traces $s.e$ with
+$s\in\trc(P), \#s = k, e\not\in\trc(P/s)$, so erroneous traces up to length
+$k+1$ are detected.
 
-The following example, however, shows that the maximal length $pq$ is really required
-when testing for refinement.
+It is interesting to investigate whether this maximal length is really
+necessary, or whether one could elaborate alternative complete test
+strategies where the SUT is tested with shorter traces only. Indeed, an
+example  presented in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} shows
+that when testing for equivalence of deterministic FSMs, it is sufficient to
+test the SUT with traces of significantly shorter length.
+
+The following example, however, shows that the maximal length $pq$ is really
+required when testing for refinement.
 \begin{example}\label{ex:pq}
@@
 \end{eqnarray*}
-Obviously, $P$'s normalised transition graph has 2 nodes,
-while $Q$'s graph has 3.
-It is easy to see (and can be checked with FDR4) that $P\lessdet_T Q$, but
-$\neg(P\lessdet_F Q)$. Furthermore, it can also be shown using the FDR4 tool that
+Obviously, $P$'s normalised transition graph has 2 nodes, while $Q$'s graph
+has 3. It is easy to see (and can be checked with FDR4) that $P\lessdet_T Q$,
+but $\neg(P\lessdet_F Q)$. Furthermore, it can also be shown using FDR4 that
 the ``test passed condition''
@@
 \]
-holds for $U_F(0),\dots,U_F(4)$, but fails for $U_F(5)$. This means that the
-non-conformance of $Q$ cannot be detected by any test trace of length
-less or equal to 5, but is revealed (as expected from Theorem~\ref{th:failurestest})
-by a trace of length 6, because the last event offered by the test $U_F(5)$ is
-refused by $Q$.
-\xbox
+holds for $U_F(0),\dots,U_F(4)$, but fails for $U_F(5)$. So, the
+non-conformance of $Q$ cannot be detected by any test trace of length less or
+equal to 5, but is revealed (as expected from Theorem~\ref{th:failurestest})
+by a trace of length 6, because the last event offered by the test $U_F(5)$
+is refused by $Q$. \xbox
 \end{example}
+\fixme{But, there is no argument that the proofs based on graph products
+corresponds to this process. In particular the hiding does not seem to play a
+role.}
 %
@@
 \end{eqnarray*}
+\fixme{The internal choice became external. To test for traces refinement, yo
+put forward just one adaptive tests. So, there is more going on here.I think
+it is because your tests are checking both traces and failures refinement.}
 Using regular expression notation, the traces of $P$ can be specified as
@@
 \]
-It is easy to see that $\trc(Q)\not\subseteq\trc(P)$;
-for example, the trace $(a^{q-1}b)^p$ is in
-$\trc(Q)\setminus\trc(P)$.
+It is easy to see that $\trc(Q)\not\subseteq\trc(P)$; for example, the trace
+$(a^{q-1}b)^p$ is in $\trc(Q)\setminus\trc(P)$.
 
@@
 \end{proof}
-
-Since Theorem~\ref{th:maxtracelen} just states that a violation of trace refinement
-may remain undetected if only traces shorter than $pq$ are checked during tests,
-it can also be applied to our trace refinement tests. Therefore,
-test suites $\{ U_T(z) \}$ with $z<pq-1$ are no longer complete.
-
-It is discussed in Section~\ref{sec:conc} how the number of test traces to be executed
-by complete test suites for failures or tarce
-refinement   can still be reduced {\it without} reducing
-the maximal length.
+%
+Since Theorem~\ref{th:maxtracelen} just states that a violation of trace
+refinement may remain undetected if only traces shorter than $pq$ are checked
+during tests, it can also be applied to our trace refinement tests.
+Therefore, test suites $\{ U_T(z) \}$ with $z<pq-1$ are not complete. It is
+discussed in Section~\ref{sec:conc} how the number of test traces to be
+executed by complete test suites for failures or trace refinement can still
+be reduced {\it without} reducing the maximal length.
 

### 2018-09-30 paper/conclusion.tex
finished review

@@
 U_F(p))$ for failures refinement never blocks before the final step specified
-by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
-should be noted, however, that it is necessary for our test verdicts to
-recognise also deadlocks in the final test step and interpret them as FAIL,
-as described in Section~\ref{sec:finitecompletefails}. In practice, this is
-realised by adding a timeout event to the testing environment which indicates
-deadlock situations. For real-time systems, this is an accepted technique,
-because the SUT has to respond within a pre-defined latency interval,
-otherwise its behaviour is considered to be blocked and regarded as a
-failure. 
-%\fixme{alcc: if deadlock is not acceptable? I think you need
-%instrumentation here as well. }
-%% jp : I do not understand the problem - sorry !
+by branch (\ref{eq:ufd}), and so we do not need inconclusive test results.
+\fixme{From the proofs of the main theorems, and what you say below, I
+thought you did have verdicts coming from deadlock, and you simply chose not
+to mark it with an inc event.} It should be noted, however, that it is
+necessary for our test verdicts to recognise also deadlocks in the final test
+step and interpret them as FAIL, as described in
+Section~\ref{sec:finitecompletefails}. In practice, this is realised by
+adding a timeout event to the testing environment which indicates deadlock
+situations. For real-time systems, this is an accepted technique, because the
+SUT has to respond within a pre-defined latency interval, otherwise its
+behaviour is considered to be blocked and regarded as a failure.
 
@@
 block at all. The adaptive behaviour of our test cases, however, induce the
-obligation to check that {\it all} possible executions have been performed
-before the test can be considered as passed. Typically, it is therefore
-assumed that a \emph{complete testing assumption}~\cite{hierons_testing_2004}
-holds, which means that every possible behaviour of the SUT is performed
-after a finite number of test executions. In practice, this is realised by
-executing each test several times, recording the traces that have been
-performed, and using hardware or software coverage analysers to determine
-whether all possible test execution behaviours of the SUT have been observed.
-Therefore, adaptive test cases come at the price of having to apply some
-grey-box testing techniques enabling us to decide whether all SUT behaviours
-have been observed.
+\fixme{Missing hiding. And also, if the SUT blocks, the test should block,
+no?} obligation to check that {\it all} possible executions have been
+performed before the test can be considered as passed. Typically, it is
+therefore assumed that a \emph{complete testing
+assumption}~\cite{hierons_testing_2004} \fixme{I think you need this for the
+failures refinement tests as well.} holds, which means that every possible
+behaviour of the SUT is performed after a finite number of test executions.
+In practice, this is realised by executing each test several times, recording
+the traces that have been performed, and using hardware or software coverage
+analysers to determine whether all possible test execution behaviours of the
+SUT have been observed. Therefore, adaptive test cases come at the price of
+having to apply some grey-box testing techniques enabling us to decide
+whether all SUT behaviours have been observed.
 
@@
 that context, only testing for traces refinement is considered, and the
-complete test suites may not be finite. So, the work presented here go well
+complete test suites may not be finite. So, the work presented here goes well
 beyond what is achieved there. On the other hand,
@@
 interest implements $I$ correctly, but afterwards has a potentially arbitrary
-behaviour specified by $RUN$. The work in~\cite{DBLP:conf/pts/CavalcantiS17}
-is not restricted to finite or non-terminating processes.
+behaviour specified by $RUN$. In addition, the work
+in~\cite{DBLP:conf/pts/CavalcantiS17} is not restricted to finite or
+non-terminating processes.
 
@@
 integrated hardware or software system, but just another CSP process
-specification, it is of course possible to verify these pass criteria using
+specification, it is of course possible to verify the pass criterion using
 the FDR4 model checker. For checking the refinement relation $P\lessdet_F Q$,
-the pass criteria have to be verified for $k=0,\dots,pq-1$, where $p$ and $q$
+the pass criterion has to be verified for $k=0,\dots,pq-1$, where $p$ and $q$
 indicate the number of nodes in $P$'s transition graph and the maximal number
@@
 In this paper, we have introduced finite complete testing strategies for
-model-based testing against CSP reference models. The strategies are
-applicable to the conformance relations failures refinement and trace
-refinement. The underlying fault domains have been defined as the sets of CSP
-processes whose normalised transition graphs do not have more than a given
-number of additional nodes, when compared to the transition graph of the
-reference process. For these domains, finite complete test suites are
-available. We have shown for the strategy to check failures refinement that
-the way of probing the SUT for illegal deadlocks used in our test cases is
-optimal in the sense that it is not possible to guarantee exhaustiveness with
-fewer probes.
+model-based testing against finite, non-terminating CSP reference models. The
+strategies are applicable to the conformance relations failures refinement
+and trace refinement. The underlying fault domains have been defined as the
+sets of CSP processes whose normalised transition graphs do not have more
+than a given number of additional nodes, when compared to the transition
+graph of the reference process. For these domains, finite complete test
+suites are available. We have shown for the strategy to check failures
+refinement that the way of probing the SUT for illegal deadlocks used in our
+test cases is optimal in the sense that it is not possible to guarantee
+exhaustiveness with fewer probes.
 

### 2018-10-01 paper/preliminaries.tex
first answers to Ana

@@
 \begin{equation}\label{eq:maxrefsminaccs}
-\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
+\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \} = \minaccs(P/s).
 \end{equation}
@@
 \fixme{alcc: the a from node 2 should go back to 0, I think.}
+\fxnote{jp: No, this is correct: state 0 corresponds to $P$, 1 to $Q\intchoice R$,
+2 to $P\intchoice R$, and 3 to $R$. 
+Therefore, 2 goes with $a$ to $(P/a) = Q\intchoice R$, that is, to node 1. }
 
@@
 \fixme{alcc: (11) is already proved in my paper with mcg. No need to repeat
-it here. I suggest just to add a reference.}
+it here. I suggest just to add a reference.} \fxnote{jp: done}Formula (\ref{eq:failconf}) has been
+shown in~\cite{DBLP:conf/icfem/CavalcantiG07}
 
-\medskip
-\noindent
-To prove (\ref{eq:failconf}), we derive
-\begin{argue}
-P \lessdet_F Q
-\\
-\Leftrightarrow \fails(Q)\subseteq \fails(Q) & definition of $\lessdet_F$
-\\
-\Leftrightarrow \forall (s,R) \in  \fails(Q): (s,R) \in\fails(P)
-& property of $\subseteq$
-\\
-\Leftrightarrow \forall (s,\varnothing) \in  \fails(Q): (s,\varnothing) \in\fails(P)
-\wedge {}
-\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
-\\
-& Definition of failures, refusals, subset closure
-\\
-\Leftrightarrow \trc(Q)\subseteq\trc(P)
-\wedge {}
-\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
-\\
-& Definition of failures and traces
-\\
-\Leftrightarrow \trc(Q)\subseteq\trc(P)
-\wedge {}
-\\\tabc \forall s\in\trc(Q)\cap\trc(P): \refs(Q/s) \subseteq\refs(P/s)
-\\
-& Property of $\cap$ and $\subseteq$
-\\
-\Leftrightarrow
-P\lessdet_T Q \wedge Q\ conf\ P
-& Definition of $\lessdet_T$ and $conf$
-\end{argue}
+%\medskip
+%\noindent
+%To prove (\ref{eq:failconf}), we derive
+%\begin{argue}
+%P \lessdet_F Q
+%\\
+%\Leftrightarrow \fails(Q)\subseteq \fails(Q) & definition of $\lessdet_F$
+%\\
+%\Leftrightarrow \forall (s,R) \in  \fails(Q): (s,R) \in\fails(P)
+%& property of $\subseteq$
+%\\
+%\Leftrightarrow \forall (s,\varnothing) \in  \fails(Q): (s,\varnothing) \in\fails(P)
+%\wedge {}
+%\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
+%\\
+%& Definition of failures, refusals, subset closure
+%\\
+%\Leftrightarrow \trc(Q)\subseteq\trc(P)
+%\wedge {}
+%\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
+%\\
+%& Definition of failures and traces
+%\\
+%\Leftrightarrow \trc(Q)\subseteq\trc(P)
+%\wedge {}
+%\\\tabc \forall s\in\trc(Q)\cap\trc(P): \refs(Q/s) \subseteq\refs(P/s)
+%\\
+%& Property of $\cap$ and $\subseteq$
+%\\
+%\Leftrightarrow
+%P\lessdet_T Q \wedge Q\ conf\ P
+%& Definition of $\lessdet_T$ and $conf$
+%\end{argue}
 
@@
 %normalised graph for a process have states that are unreachable?}
-%\fxnote{jp: You are rights: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
+%\fxnote{jp: You are right: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
 Let
@@
 
+% -------------------------------------------------------------------------
+\subsubsection*{Minimal Hitting Sets of Normalised Transition Graphs}
+Since minimal acceptances can be used to label the nodes of a normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$ according to (\ref{eq:maxrefsminaccs}), the notation of minimal hitting sets also
+carries over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe that
+\begin{equation}
+\label{eq:minhitsGP}
+\minhits(G(P)/s) = \minhits(P/s) \ \text{for all}\ s\in \trc(P).
+\end{equation}
+\fixme{You already used this above. Perhaps you want to move this to before
+Lemma 4?}\fxnote{jp: done}
+
 % ------------------------------------------------------------------------
@@
 \begin{proof}
+Throughout the proof, we apply (\ref{eq:maxrefsminaccs}) and (\ref{eq:minhitsGP}), 
+so that $\minaccs(P/s)$ and $\minaccs(G(P)/s)$, as well as 
+$\minhits(P/s)$ and $\minhits(G(P)/s)$ can be used interchangeably.
+
 For showing ``$1 \Rightarrow 2$'', assume   $Q\ conf\ P$ and $s\in\trc(P)\cap
@@
 set $H\in \minhits(P/s)$ which is {\it not} a hitting set of
-$\minaccs(G(Q)/s)$, a contradiction to Assumption~2. We note that we use
-$\minhits(G(P)/s) = \minhits(P/s)$, as well $\minaccs(G(Q)/s) =
-\minaccs(Q/s)$, which follow from the identification of process states with
-states of the normalised graph for the proof discussed in the previous
-section. This completes the proof of the lemma. \fixme{Please, check that
-you're happy with this observation.}\xbox
+$\minaccs(G(Q)/s)$, a contradiction to Assumption~2. 
+%We note that we use
+%$\minhits(G(P)/s) = \minhits(P/s)$, as well $\minaccs(G(Q)/s) =
+%\minaccs(Q/s)$, which follow from the identification of process states with
+%states of the normalised graph for the proof discussed in the previous
+%section. 
+This completes the proof of the lemma. 
+\fixme{Please, check that
+you're happy with this observation.}
+\fxnote{jp: I think the new first sentence of the proof should now be sufficient. Therefore, I removed your addition.}
+\xbox
 \end{proof}
@@
 holds.
-
-% -------------------------------------------------------------------------
-\subsubsection*{Minimal Hitting Sets of Normalised Transition Graphs}
-Since minimal acceptances can be used to label the nodes of a normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$, the notation of minimal hitting sets also
-carries over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe that
-\begin{equation}
-\label{eq:minhitsGP}
-\minhits(G(P)/s) = \minhits(P/s) \ \text{for all}\ s\in \trc(P).
-\end{equation}
-\fixme{You already used this above. Perhaps you want to move this to before
-Lemma 4?}
 

### 2018-10-01 paper/abstract.tex
couple of questions
further points of Ana's suggestions handled

@@
 process representing the reference model, test suites checking the
-conformance relations trace and failures refinement are constructed, and
+conformance relations trace refinement and failures refinement are constructed, and
 their finiteness and completeness (that is, capability to uncover conformity
@@
 previously investigated by several authors, a sufficient condition for their
-finiteness is presented here for the first time. Moreover, an optimal
-technique for testing the admissibility of an implementation's
-nondeterministic behaviour is described.
+finiteness is presented here for the first time. 
+%Moreover, an optimal
+%technique for testing the admissibility of an implementation's
+%nondeterministic behaviour is described. 
+It is shown that the test suites are optimal with respect to two aspects:  (a)  the maximal length of test traces cannot be reduced, and (b) the nondeterministic behaviour cannot be tested with smaller or fewer sets of events, without losing the test suite's completeness property.
+\fixme{alcc: Is the technique
+applying the test suites?}
+\fxnote{jp: this sentence needed to be re-phrased -- the new version should be better understandable}

### 2018-10-01 paper/conclusion.tex
further points of Ana's suggestions handled

@@
 concerning the potential implications of this paper in the field of model
-checking. The work of Cavalcanti is funded by the Royal Academy of
+checking. We are
+grateful to Li-Da Tong from National Sun Yat-sen University, Taiwan, for
+suggesting the applicability of Sperner's Theorem in the context of the work
+presented here. Moreover, we thank Adenilso Simao from the University of S{\~a}o Paulo
+for several helpful suggestions. 
+The work of Ana Cavalcanti is funded by the Royal Academy of
 Engineering and UK EPSRC Grant EP/R025134/1.

### 2018-10-01 paper/introduction.tex
further points of Ana's suggestions handled

@@
 unsure, but you guys know about FSMs much more than I do.}
+\fxnote{jp: I think that this description is ok. With FSMs as used in the testing community, there are no tau events, and parallel composition $M_1\cap M_2$
+is just language intersection. Therefore, 
+when composing $M_1$ and $M_2$ to $M_1\cap M_2$,
+$M_1$ cannot refuse to engage into I/O-event $x/y$,
+if $M_2$ has $x/y$ as the only initial event in its language.  }
 

### 2018-10-01 paper/preliminaries.tex
further points of Ana's suggestions handled

@@
 When investigating  tests for failures refinement, the notion of
-\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
-is also useful. An acceptance set of $P/s$ is a subset of its initials
-$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
-$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
+\emph{acceptances}, which is dual to refusals,
+is also useful. While the original introduction of acceptances presented 
+in~\cite[pp.~75]{Hennessy:1988:ATP:50497} was independent of refusals, 
+we use the definition from~\cite[pp.~278]{Roscoe:1997:TPC:550448}.   A \emph{minimal acceptance} of a CSP process state $P/s$ 
+is the complement of a maximal refusal of the same state. The set of minimal acceptances 
+of $P/s$ is denoted by $\minaccs(P/s)$ and formally defined as 
+%
+\begin{equation}\label{eq:accref}
+\minaccs(P/s) = \{ \Sigma - R~|~R\in \maxrefs(P/s)  \}
+\end{equation}
+%
+With this definition, a (not necessarily minimal) \emph{acceptance} of $P/s$ is a superset of some minimal acceptance and a subset of the initials $[P/s]^0$.
+Denoting the acceptances of $P/s$ by $\accs(P/s)$, this leads to the formal definition
+\begin{equation}
+\label{eq:accrefall}
+\accs(P/s)  = \{ B\subseteq [P/s]^0~|~\exists A\in\minaccs(P/s):  A \subseteq B   \}
+\end{equation}
+%
+Acceptances have the following intuitive interpretation.  If the behaviour of  $P/s$ is deterministic, its only acceptance
 equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
@@
 {\it always} refusing those from $\Sigma - [P/s]^0$.
-The acceptances of $P/s$ are denoted by
-$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
-the following properties.
-%
-\begin{eqnarray}
-A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
-\label{eq:accref}
-\\
-\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
-\label{eq:accinitials}
-\\
- X\in\accs(P/s) &\Leftrightarrow & \exists A\in \minaccs(P/s): A\subseteq X \subseteq [P/s]^0
- \label{eq:accsubset}
-\end{eqnarray}
-%
-%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
-%acceptances by means of maximal refusals. In this case,
-%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
-%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
-%sure? (4) does not talk about Acc. Is this important?}
-%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
-%removed it.}
-%
+
 Exploiting (\ref{eq:accref}),
-%(\ref{eq:accinitials}), and (\ref{eq:accsubset}),
-every node of a normalised transition graph can
+the nodes of a normalised transition graph can
 alternatively be labelled with their minimal acceptances, and this
@@
 \fixme{alcc: proof in the report perhaps? What do you need (5) for? }
+\fxnote{jp: I have clarified the introduction of acceptances and removed the statements
+we do not need in the rest of the paper. We need (5) here show that the set of
+maximal refusals of $P/s$ equals $r(G(P/s))$.}
 
@@
 (\ref{eq:maxrefsminaccs}).
-%%\fixme{alcc: I think at least the report should point to proofs.}
 %
@@
 it here. I suggest just to add a reference.} \fxnote{jp: done}Formula (\ref{eq:failconf}) has been
-shown in~\cite{DBLP:conf/icfem/CavalcantiG07}
+shown in~\cite{DBLP:conf/icfem/CavalcantiG07}.
 
@@
 maximally exist for a given collection of minimal acceptances. To this end,
-the following definitions and results are useful\footnote{The authors are
-grateful to Li-Da Tong from National Sun Yat-sen University, Taiwan, for
-suggesting the applicability of Sperner's Theorem in the context of the work
-presented here.}.\fixme{Shouldn't the footnote go to the acknowledgements?}
+the following definitions and results are useful.\fixme{Shouldn't the footnote go to the acknowledgements?}\fxnote{jp: done}
 
@@
 \begin{enumerate}
-\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/s$;
+\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/2$;
     or
-\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/s$.
+\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/2$.
 \end{enumerate}

### 2018-10-02 paper/complexity.tex
Further comments by Ana handled

@@
 
-The authors of~\cite{Hennessy:1988:ATP:50497} and
-\cite{DBLP:conf/icfem/CavalcantiG07} suggest to test {\it every} non-empty
+In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test 
+\fixme{In \cite{DBLP:conf/icfem/CavalcantiG07}, we work
+with minimal acceptances. I didn't check \cite{Hennessy:1988:ATP:50497}.}
+\fxnote{jp: we have studied \cite{DBLP:conf/icfem/CavalcantiG07} now in more detail -- you are right, of course, but there is a confusion of terms: you are also using the minimal hitting sets we handle here in our joint paper, but you used the term ``minimal acceptance'' for them.
+This is not in line with the definition of minimal acceptances as presented in \cite{Hennessy:1988:ATP:50497} and
+in Roscoe's books. Here, in our joint paper, minimal acceptances are introduced
+as defined in Roscoe's books. I have changed the text accordingly.}
+{\it every} non-empty
 subset of $\Sigma$ whose events cannot be completely refused in a given
@@
 number of test executions in comparison
-to~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} by a factor
+to~\cite{Hennessy:1988:ATP:50497} by a factor
 of $1/{\sqrt{2\pi \card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction
 is visualised by a function plot of $2^{\card{\Sigma}}$ versus
-$\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$. \fixme{In [2], we work
-with minimal acceptances. I didn't check [8].}
+$\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.
+
+In~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors also use minimal hitting sets\footnote{However, they are denoted by {\it minimal acceptances} in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an upper bound for
+the number of test executions to be performed.
+
 % .....................................................................................

### 2018-10-02 paper/conclusion.tex
Further comments by Ana handled

@@
 behaviour is considered to be blocked and regarded as a failure.
+The tests executions $(Q\parallel[\Sigma] U_T(p))$ 
+\fixme{Missing hiding.}
+\fxnote{jp: Here, the hiding operator is not necessary: the possible test executions are really the traces of $Q$ parallel $U_T(p)$. The hiding is only needed to specify the verdict.}  
+for trace refinement never
+block at all before stopping after the verdict $\epass$ or $\efail$, unless the
+SUT process $Q$ has an unexpected deadlock state. 
+\fxfatal{And also, if the SUT blocks, the test should block,
+no?}\fxnote{jp: you are right, I changed the text here and added explanations above after the verdict explanation
+in formula (\ref{eq:passT}).}
+Recall from the explanations given in Section~\ref{sec:finitecomplete} that the 
+blocking situation also leads to passing the test, because 
+$(\epass\then \Stop)\lessdet_T (Q\parallel[\Sigma] U_T(p))\hide\Sigma$ still holds
+if $(Q\parallel[\Sigma] U_T(p))\hide\Sigma$ only produces the empty trace.
 
-Our tests executions $(Q\parallel[\Sigma] U_T(p))$ for trace refinement never
-block at all. The adaptive behaviour of our test cases, however, induce the
-\fixme{Missing hiding. And also, if the SUT blocks, the test should block,
-no?} obligation to check that {\it all} possible executions have been
+The adaptive behaviour of both our test case types $U_F(p)$ and $U_T(p)$,
+however, induce the obligation to check that {\it all}
+possible executions have been
 performed before the test can be considered as passed. Typically, it is
@@
 assumption}~\cite{hierons_testing_2004} \fixme{I think you need this for the
-failures refinement tests as well.} holds, which means that every possible
+failures refinement tests as well.}\fxnote{jp: yes, I re-arranged the text accordingly.} holds, which means that every possible
 behaviour of the SUT is performed after a finite number of test executions.

### 2018-10-02 paper/finitecomplete-trace.tex
Further comments by Ana handled

@@
 and we use the condition $\minaccs(n) = \{ \varnothing \}$ instead of
-$\minhits(n) = \varnothing$ to indicate that minimal hitting sets need not to
+$\minhits(n) = \varnothing$ in branch (\ref{eq:utb}) 
+to indicate that minimal hitting sets need not to
 be calculated for generating these tests from $G(P)$. From
@@
 proven in Lemma~\ref{lemma:ufproperties} for $U_F(p)$ for traces
-$s\in\trc(P)$. \fixme{In (37), I think there is an implicit assumption about
-$p$ and $\# s$?}
+$s\in\trc(P)$ with $\#s \le p$. 
+\fixme{In (37), I think there is an implicit assumption about
+$p$ and $\# s$?}\fxnote{jp: yes, I added $\#s \le p$ in the text.}
 
@@
 \fixme{The proofs don't use this definition. It should change.}
+\fxnote{jp: I do not understand this remark -- we should discuss it on the phone.
+Please note the addition explanations I have added here below (\ref{eq:passT}).}
 %
+If the SUT process $Q$ deadlocks internally after a trace $s$ where also the reference
+process $P$ is in a state where deadlock is possible, this is captured by the fact
+that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$. As a consequence, branch
+(\ref{eq:utb}) of a test case execution state $U_T(p,k,n)$  
+with $\#s = k \le p$   can be entered
+and the  test execution terminates with $\epass$. If, however,  
+$Q$ blocks after a trace $s'$
+where the reference process satisfies $\minaccs(P/s') \neq\varnothing$, branch 
+(\ref{eq:utb})  cannot be taken, and the test execution stops without producing
+$\epass$ or $\efail$. In contrast to the test for failures refinement, this is also interpreted here as a successful test execution, because unexpected blocking
+of the SUT does not violate the trace refinement relation, as long as all traces 
+executed by the SUT are traces of the reference process. In particular, if 
+neither $\epass$ nor $\efail$ is ever produced, so that
+$(Q\parallel[\Sigma] U_T(p)) \hide \Sigma = \Stop$, the test passes, because
+$(\epass\then\Stop) \lessdet_T \Stop$ holds.
+
 \begin{theorem}\label{th:tracetest}

### 2018-10-02 paper/finitecomplete-failures.tex
further points of Ana's suggestions handled
update

@@
 A test is performed by running $U_F(p)$ concurrently with any SUT process
-$Q$, synchronising over $\Sigma$. So, a test execution is a trace of the
+$Q$, synchronising over $\Sigma$. So, a \emph{test execution} is a trace of the
 concurrent process
@@
 \begin{proof}{~}We make two points in separate steps below. The
-first is that the test execution cannot reach branch~(\ref{eq:ufa})and raise
+first is that the test execution cannot reach branch~(\ref{eq:ufa}) and raise
 a $fail$ event.  The second is that it cannot deadlock without raising a
-$pass$ event. This case would also interpreted as FAIL, since then
+$pass$ event. This case would also be interpreted as FAIL, since then
 $\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
@@
 without a verdict. \fixme{I think we need a stronger argument here, as to why
-it cannot deadlock.} For the last branch, we note that it can only be entered
+it cannot deadlock.}
+\fxnote{jp: I do not understand the question -- let's discuss this on the phone.}
+ For the last branch, we note that it can only be entered
 if $\minhits(P/s) = \minhits(n) \neq\varnothing$. In this case, any selected
@@
 of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the $\efail$
-event or deadlocks.
+event or deadlocks. We do this by analysing the product graph of the reference
+process $P$ and the SUT process $Q$: any trace $s\in\trc(Q)\cap\trc(P)$ gives rise to 
+a path labelled by the events of $s$ 
+through this product graph. Any error can be detected after running through such a trace
+and then either observing an event outside $[P/s]^0$ (this is the violation described
+by Case~1) or identifying an illegal acceptance $A_Q$ as specified in Case~2. 
+It is not guaranteed, however, that trace $s$ is short enough to be executed by one of the 
+test cases $U_F(p)$. Therefore, it has to be shown that for any $s$ leading to an error
+situation specified by Case~1 or Case~2, there exists a trace $u$ of maximal length $pq-1$
+leading to the same error.
 
 \medskip
-\noindent {\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
+\noindent {\bf Case~1.} 
+\fixme{There is an assumption here that
+the graph of a test execution is given by the cross-product of graphs.  But,
+test execution is not defined in this way. It is defined by parallelism and
+hiding. I think the definition of test execution needs to change.}
+\fxnote{jp: this is a misunderstanding: we are not looking at the product graph of $Q$ and $U_F(p)$, but at the product graph of $Q$ and $P$. I have added more explanatory 
+text in the paragraph  before Case~1.}
+Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
 but $s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is
@@
 event $e$, so that the test fails. Again, the assumption of non-divergence of
-$Q$ is needed for this conclusion. \fixme{There is an assumption here that
-the graph of a test execution is given by the cross-product of graphs.  But,
-test execution is not defined in this way. It is defined by parallelism and
-hiding. I think the definition of test execution needs to change.}
+$Q$ is needed for this conclusion.
 
 \medskip
-\noindent {\bf Case~2.} We note that trace $s$ is again a trace of the
+\noindent {\bf Case~2.} We note 
+\fixme{Again, nothing here is about this process, but about the graph, and
+the hiding that is needed is discarded.} 
+\fxnote{jp: see my explanation above}
+that trace $s$ is again a trace of the
 product graph $G$, but we do not know its length. Again, by applying
@@
 $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event;
-\fixme{Again, nothing here is about this process, but about the graph, and
-the hiding that is needed is discarded.} this means that the test fails. This
-concludes the proof. \xbox
+this means that the test fails. This
+concludes the proof. 
+\xbox
 \end{proof}
@@
 Our notion of tests can be specialised to deal with traces refinement, as we
-explain next. Examples come later in Section~\ref{sec:case}. \fixme{I think
+explain in Section~\ref{sec:finitecomplete}. 
+The next section presents and example.
+\fixme{I think
 the example should come before more theory.}
+\fxnote{jp: done}
 

### 2018-10-03 paper/casestudies.tex
Further comments by Ana handled
Further updates. Full proof for Theorem 3 added.

@@
 
-%
-%For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, it is
-%advisable to avoid an enumeration of traces $s$ of the reference process.
-%Instead, we calculate the following auxiliary functions from $P$'s transition
-%graph.
-%%
-%\begin{eqnarray*}
-%\text{\it initials} & : & N \fun \power (\Sigma)
-%\\
-%\minhits & : & N \fun \power\power(\Sigma)
-%\end{eqnarray*}
-%%
-%As mentioned, in a state $n = G(P)/s$, the set $\text{\it initials}(n)$
-%equals the events labelling outgoing edges of $n$, so  $\text{\it
-%initials}(n) = [P/s]^0$. The function $\minhits$ maps $n$ to the set of all
-%minimal hitting sets associated with the minimal acceptances of $n$, so
-%$\minhits(n) = \minhits(P/s)$. Using the transition function $t$ and the
-%above functions, $U_F(p)$ can be re-written as the failures-equivalent CSP
-%process below.
-%%
-%\begin{eqnarray}
-%U_F^1(p) & = & U_F^1(p,0,\ii n)
-%\\
-%U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{\it initials}(n)) @ e \then \efail\then \Stop \big)
-%\label{eq:xufa}
-%\\ & & \extchoice \nonumber
-%\\ & & (\minhits(n) = \varnothing) \& \big( \epass \then \Stop \big)
-%\label{eq:xufb}
-%\\ & & \extchoice \nonumber
-%\\ & & (k < p) \& \big(\Extchoice e:\text{\it initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
-%\label{eq:xufc}
-%\\ & & \extchoice \nonumber
-%\\ & & (k = p) \& \big( \sqcap_{H\in\minhits(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
-%\label{eq:xufd}
-%\end{eqnarray}
-%\fixme{alcc: I think the example should be for the theory presented before,
-%not with a different notion of test. In particular, 33 is very different.}
-
 Generating the test cases $U_F(p)$ specified in  (\ref{eq:UFP}) for the reference
 process $P$ discussed in Example~\ref{example:CSP},
-results in the following instantiations of initials, minimal hitting sets, and
-transition function; this can be directly seen by analysing $P$'s normalised
-transition graph with nodes $N =\{0,1,2,3\}$ as shown in Fig.~\ref{fig:tga}.
+results in the   instantiations of initials, minimal hitting sets, and
+transition function shown in Fig.~\ref{fig:initialsminhitstrans}; 
+this can be directly derived from $P$'s normalised
+transition graph with nodes $N =\{0,1,2,3\}$ displayed in Fig.~\ref{fig:tga}.
 
+\begin{figure}[htbp]
+\begin{center}
+\begin{minipage}{0.3\textwidth} 
 	 \begin{eqnarray*}
@@
 { }[2]^0 & = & \{ a,b,c \} \\
-{ }[3]^0 & = & \{ b,c \} \\
-& & \\
+{ }[3]^0 & = & \{ b,c \}
+\end{eqnarray*}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.3\textwidth}
+	 \begin{eqnarray*}
 \minhits(0) & = & \{ \{ a\} \} \\
@@
 \minhits(2) & = & \{  \{a,b\},  \{a,c\},\} \\
-\minhits(3) & = & \{ \{ b\}, \{c\} \} \\
-& & \\
+\minhits(3) & = & \{ \{ b\}, \{c\} \} 
+\end{eqnarray*}
+
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.3\textwidth}
+	 \begin{eqnarray*}
 t(0,a) & = & 1 \\
@@
 \end{eqnarray*}
+	\end{minipage}
+\caption{Initials, minimal hitting sets, and transition function of the normalised transition graph displayed in Fig.~\ref{fig:tga}.}
+\label{fig:initialsminhitstrans}
+\end{center}
+\end{figure}
 
@@
 
-%%%and refinement does not hold.
-% alcc: I don't think there is a notion of refinement for state?
-%so this process state no
-%longer refines the corresponding process state $R$ of the reference process
-%$P$ in the failures model.
-
 Running the test $U_F(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p = 4$
@@
 $\{c \}$, while $U_F(4)/s$ choses to accept only $\{a,b\}$. Therefore,
+\fixme{The hiding is missing.} 
+\fxnote{jp: done}
 \[
-(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F(4)),
+(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F(4)) \hide\Sigma,
 \]
-\fixme{The hiding is missing.} and the test fails. \xbox
+and the test fails. \xbox
 \end{example}
-%%%\fixme{alcc: I would write down the actual test generated.}
 

### 2018-10-03 paper/conclusion.tex
Further updates. Full proof for Theorem 3 added.

@@
 
-In contrast to that, our test cases specified in
+In 
+\fixme{From the proofs of the main theorems, and what you say below, I
+thought you did have verdicts coming from deadlock, and you simply chose not
+to mark it with an inc event.} 
+contrast to that, our test cases specified in
 Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
 adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
-U_F(p))$ for failures refinement never blocks before the final step specified
+U_F(p))$ for failures refinement never block before the final step specified
 by branch (\ref{eq:ufd}), and so we do not need inconclusive test results.
-\fixme{From the proofs of the main theorems, and what you say below, I
-thought you did have verdicts coming from deadlock, and you simply chose not
-to mark it with an inc event.} It should be noted, however, that it is
+It should be noted, however, that it is
 necessary for our test verdicts to recognise also deadlocks in the final test
@@
 The adaptive behaviour of both our test case types $U_F(p)$ and $U_T(p)$,
-however, induce the obligation to check that {\it all}
+however, induces the obligation to check that {\it all}
 possible executions have been

### 2018-10-03 paper/finitecomplete-failures.tex
Further updates. Full proof for Theorem 3 added.

@@
 A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
-exhaustive for the fault model specified there.
+exhaustive for the fault model specified there and for conformance relation $\lessdet_F$.
 \end{lemma}

### 2018-10-03 paper/finitecomplete-trace.tex
Further updates. Full proof for Theorem 3 added.

@@
 % ==========================================================================
-The existence of complete, finite test suites is expressed in analogy to
-Theorem~\ref{th:failurestest}. A noteworthy difference is that the complete
-suite for trace refinement just needs the single adaptive test case
-$U_T(pq-1)$, while failures refinement requires the execution of $\{
-U_F(0),\dots,U_F(pq-1)\}$. The reason is that $U_T(pq-1)$ identifies trace
-errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for
-erroneous acceptances at the end of each trace of length $(pq -1)$. Since
-$U_T(pq-1)$ never blocks any $Q$ event before terminating, the pass criterion
+Since
+$U_T(p)$ never blocks any $Q$-event before terminating, the pass criterion
 can be based on trace refinement instead of failures refinement as required
@@
 Please note the addition explanations I have added here below (\ref{eq:passT}).}
-%
+
 If the SUT process $Q$ deadlocks internally after a trace $s$ where also the reference
@@
 (\ref{eq:utb})  cannot be taken, and the test execution stops without producing
-$\epass$ or $\efail$. In contrast to the test for failures refinement, this is also interpreted here as a successful test execution, because unexpected blocking
+$\epass$ or $\efail$. In contrast to the test for failures refinement, this is   interpreted here as a successful test execution, because unexpected blocking
 of the SUT does not violate the trace refinement relation, as long as all traces 
@@
 
+The existence of complete, finite test suites is expressed in analogy to
+Theorem~\ref{th:failurestest}. A noteworthy difference is that the complete
+suite for trace refinement just needs the single adaptive test case
+$U_T(pq-1)$, while failures refinement requires the execution of $\{
+U_F(0),\dots,U_F(pq-1)\}$. The reason is that $U_T(pq-1)$ identifies trace
+errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for
+erroneous acceptances at the end of each trace of length $(pq -1)$. 
+%
 \begin{theorem}\label{th:tracetest}
@@
 \end{theorem}
+%\begin{proof}
+%The theorem follows directly
+%from Step~1 in the proof of Lemma~\ref{lemma:mainfsound} and
+%Case~1 in the proof of Lemma~\ref{lemma:mainfexhaustive}.
+%\xbox
+%\end{proof}
+%%
+%Examples are provided after our discussion of size of the test suites.
+
+As before in the context of Theorem~\ref{th:failurestest}, 
+the proof of Theorem~\ref{th:tracetest} is structured into two lemmas, the first 
+establishing soundness, the second exhaustiveness.
+
+\fxnote{jp: Wen-ling and I have now provided a full proof for Theorem~\ref{th:tracetest}}
+% =============================================================================== 
+\begin{lemma}\label{lemma:mainfsoundtrace}
+A test suite $\TS_T$ generated from a CSP process $P$, as specified in
+Theorem~\ref{th:tracetest}, is passed by every CSP process $Q$ satisfying
+$P\lessdet_T Q$.
+\end{lemma}
 \begin{proof}
-The theorem follows directly
-from Step~1 in the proof of Lemma~\ref{lemma:mainfsound} and
-Case~1 in the proof of Lemma~\ref{lemma:mainfexhaustive}.
+Suppose that $P\lessdet_T Q$, so that $\trc(Q)\subseteq\trc(P)$, and assume that
+$s\in\trc(Q)$ with $\#s < pq$. Since $s$ is also a trace of $P$, we can conlude
+$$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$ because of (\ref{eq:ifpaT}). Now $\trc(Q)\subseteq\trc(P)$ 
+implies $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot enter branch (\ref{eq:uta}) and produce a $\efail$-event. Therefore, only four options are
+available for the test execution    $(Q\parallel[\Sigma] U_T(p))/s$ to continue.
+
+\medskip
+\noindent
+{\bf Case~1.} $Q/s$ deadlocks and $\minaccs(G(P)/s) = \{ \varnothing \}$. Then
+$U_T(pq-1,\#s,G(P)/s)$ enters branch (\ref{eq:utb}), and the test execution 
+stops after event $\epass$.
+
+\medskip
+\noindent
+{\bf Case~2.} $Q/s$ deadlocks, but $\minaccs(G(P)/s)\neq\{\varnothing\}$. 
+Then the whole test execution blocks, and this is means that
+neither a $\epass$ nor a $\efail$ event is produced, so the test execution is passed.
+
+\medskip
+\noindent
+{\bf Case~3.} $Q/s$ selects an event $e\in[Q/s]^0$ and $\#s < pq-1$. Then the test case 
+$U_T(pq-1)$ in state
+$U_T(pq-1,\#s,G(P)/s)$ can also engage in $e$ by entering branch (\ref{eq:utc}),
+and the test execution continues without having produced a $\epass$ or a $\efail$ event.
+
+\medskip
+\noindent
+{\bf Case~4.} $\#s = pq-1$ holds. Then $U_T(pq-1,\#s,G(P)/s)$ can only enter
+branch (\ref{eq:utd}), and the test execution stops after $\epass$.
+
+This case analysis shows that every   execution of $(Q\parallel[\Sigma] U_T(p))$
+either stops after $\epass$ or produces neither $\epass$ nor $\efail$. This
+proves that $Q$ passes test $U_T(pq-1)$ according to the pass  criterion (\ref{eq:passT}).
 \xbox
 \end{proof}
-%
-Examples are provided after our discussion of size of the test suites.
+% ===============================================================================
+
+\begin{lemma}\label{lemma:mainfexhaustivetrace}
+A test suite $\TS_T$ specified as in Theorem~\ref{th:tracetest} is
+exhaustive for the fault model specified there and for conformance relation $\lessdet_T$.
+\end{lemma}
+\begin{proof}
+As before in the proofs for
+failures testing, we construct the product graph $G=G(P)\times G(Q)$ and recall
+that every trace $s\in\trc(P)\cap\trc(Q)$ is associated with a path through $G$ labelled
+with the same events as $s$, such that $G/s = (G(P)/s,G(Q)/s)$. Furthermore, we recall from Lemma~\ref{lemma:extendV} that graph state $(G(P)/s,G(Q)/s)$ can always be
+reached by a trace $u$ of length less or equal $pq-1$, where  the order of $G(P)$ is $p$ and that of $G(Q)$ is $q$.
+
+Suppose that $P\not\lessdet_T Q$. Then there exists a trace $s\in \trc(Q)\cap\trc(P)$ and an event $e\in [Q/s]^0$ such that $e\not\in [P/s]^0$. Let $u\in \trc(Q)\cap\trc(P)$
+be a trace with $\#u < pq$ and $G/u = (G(P)/s,G(Q)/s)$. Then 
+$$
+U_T(pq-1)/u = U_T(pq-1,\#u,G(P)/s).
+$$ 
+By assumption, 
+$e\in (\Sigma -[P/s]^0) = (\Sigma - [G(P)/s]^0)$. 
+Since  $G(Q)/u = G(Q)/s$,
+ $Q/u$ can engage into $e$. Then $U_T(pq-1,\#u,G(P)/s)$ will enter branch (\ref{eq:uta}), and the test execution stops after having produced $\efail$. This proves that
+  $Q$  fails test $U_T(pq-1)$.
+\xbox
+\end{proof}

### 2018-10-03 paper/preliminaries.tex
Further updates. Full proof for Theorem 3 added.

@@
 \fxnote{jp: I have clarified the introduction of acceptances and removed the statements
-we do not need in the rest of the paper. We need (5) here show that the set of
+we do not need in the rest of the paper. We need (5) here to show that the set of
 maximal refusals of $P/s$ equals $r(G(P/s))$.}
+
+Given any non-diverging, non-terminating, finite-state process $P$,
+the process can be re-con\-struc\-ted from its transition graph $G(P)$ with initial state $\ii n$ and transition function $t$, using 
+$P$'s normalised syntactic representation~\cite[pp.~277]{Roscoe:1997:TPC:550448} specified as follows.
+\begin{eqnarray*}
+\text{normalised}(P) & = & P_N(\ii n)
+\\
+P_N(n) & = & \Intchoice_{A\in\minaccs(n)\cup \{ [n]^0 \} } e:A\then P_N(t(n,e))
+\end{eqnarray*}
+With this definition, $P$ is semantically equivalent to $\text{normalised}(P)$ in the CSP failures semantics.
 
@@
  \end{figure}
-
-\fixme{alcc: the a from node 2 should go back to 0, I think.}
-\fxnote{jp: No, this is correct: state 0 corresponds to $P$, 1 to $Q\intchoice R$,
-2 to $P\intchoice R$, and 3 to $R$. 
-Therefore, 2 goes with $a$ to $(P/a) = Q\intchoice R$, that is, to node 1. }
 
@@
 and $\{b,c\}$. We note that the event $c$ can never be refused, since it is
-contained in each minimal acceptance set.
+contained in each minimal acceptance set.\fixme{alcc: the a from node 2 should go back to 0, I think.}
+\fxnote{jp: No, this is correct: state 0 corresponds to $P$, 1 to $Q\intchoice R$,
+2 to $P\intchoice R$, and 3 to $R$. 
+Therefore, 2 goes with $a$ to $(P/a) = Q\intchoice R$, that is, to node 1. I think that
+this is now explained sufficiently clear  in the example.}
 
@@
 process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
-two minimal acceptance sets labelling node 2. Similar considerations explain
-the other nodes and transitions in Fig.~\ref{fig:tga}.
+two minimal acceptance sets labelling node 2. 
+From node~2, event $c$ leads to node~3. Since $P$ does not engage into $c$,
+the $R$-component of $P\intchoice R$ must have processed $c$, so node~3 corresponds to process state $R/c = R$, and therefore it is labelled by $R$'s minimal acceptance
+$\{b,c\}$. 
 \xbox
@@
 \label{lemma:extendV} Let $P$ be a CSP process with normalised transition
-graph $G(P)$. %\fixme{alcc: can a
-%normalised graph for a process have states that are unreachable?}
-%\fxnote{jp: You are right: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
+graph $G(P)$. 
 Let

### 2018-10-03 paper/complexity.tex
update
further updates: use j instead of k in U_F(j)

@@
 
-As can be seen from the specification of the test cases $U_F(k)$ (see
+In this 
 \fixme{In that section, $k$ is used for a different purpose. It is confusing.
-I would use $p$ here.} Section~\ref{sec:finitecompletefails}), the number of
+I would use $p$ here.} 
+\fxnote{jp: ok -- I also had to change this in Theorem~2, where $k$ was used as index from
+0 to $pq-1$. I now use $j$ as index for $U_F(j)$, because $p$ is the number of states in $G(P)$.}
+section, we
+ will   calculate upper bounds for the total number of test executions to be performed
+ when testing for failures refinement.   Theorem~\ref{th:failurestest} specifies
+ that all tests $U_F(j),\ 0\le j < pq$ need to be executed, where $p$ denotes the
+ number of nodes in the transition graph of the reference process $P$, and $q\ge p$ is an estimate for the number of nodes in the SUT's transition graph.
+Therefore, we will first calculate a bound for the number of test executions to be performed
+ for test $U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.
+
+For the worst-case estimate, we
+assume that $P$ never allows for early deadlock (so $\minhits(P/s)$ is never empty) 
+and that the SUT $Q$ is a correct 
+failures refinement. Then all test executions $(Q\parallel[\Sigma] U_F(j))$
+stop after having run through a $Q$-trace of length $j+1$, because their is
+no early termination due to entering branches (\ref{eq:ufa}), (\ref{eq:ufb}), or due to
+an illegal deadlock of $Q$.
+As can be seen from the specification of the test cases $U_F(j)$ (see
+Section~\ref{sec:finitecompletefails}), the number of
 executions ending in a $\epass$ event corresponds to the number $\ell$ of
-traces $s$ of $P$ with length equal to\footnote{In this estimation, we
-disregard the case where the test terminates earlier due to entering branch
-(\ref{eq:ufb}).} $k$, multiplied by the number $h$ of minimal hitting sets in
-$\minhits(P/s)$.\fixme{But doesn't this add to the number of executions?} For
-the tests $U_T(k)$ verifying trace refinement (see
+traces $s$ of $P$ with length equal to  
+$j$, multiplied by the number $h$ of minimal hitting sets in
+$\minhits(P/s)$. For
+the tests $U_T(j)$ verifying trace refinement (see
 Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$,
-since there is no equivalent in $U_T(k)$ to checking different hitting sets
+since there is no equivalent in $U_T(j)$ to checking different hitting sets
 in the last step of a test execution.
+\fixme{But doesn't this add to the number of executions?}
+\fxnote{jp: should be clear now from the revised test}
 
@@
 The normalised transition graph of this process has a single state, and its initials
-are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(k)$ can
+are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(j)$ can
 never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly
-$\card{\Sigma}^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
+$\card{\Sigma}^j$ different traces of length $j$ exercising branch (\ref{eq:ufc})
 for each of their events.
@@
 Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
-acceptances, the cardinality   $h = \minhits(P/s)$ has the upper bound
+acceptances, the cardinality   $h = \card{\minhits(P/s)}$ has the upper bound
 \[
@@
 
-By Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(k)$ for
-$k = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
+By Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(j)$ for
+$j = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
 we use the formula for the sum of the geometric progression.
@@
 %
-From Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
-above cannot be further improved, since the full collection of minimal
-hitting sets needs to be checked to verify $conf$.\fixme{Theorem 4 is not
+From 
+\fixme{Theorem 4 is not
 about $conf$. I don't think you have tests for $conf$.}
+\fxnote{jp: I rephrased this -- hope that it's better understandable now.}
+Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
+above cannot be further reduced, since the full collection of minimal
+hitting sets needs to be checked at the end of each execution of $U_F(j)$.
 
@@
 
-According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
+According 
+\fixme{Please, see comments
+above about this process.}\fxnote{jp: should be ok now after the above changes and clarifications.}
+to Theorem~\ref{th:tracetest}, a complete test suite checking trace
 refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
-$U_F(k)$ above, the number of executions performed by $(Q\parallel[\Sigma]
-U_R(pq-1))$ is bounded by $\card{\Sigma}^{pq-1}$. \fixme{Please, see comments
-above about this process.}
+$U_F(j)$ above, the number of executions performed by $(Q\parallel[\Sigma]
+U_T(pq-1))$ is bounded by $\card{\Sigma}^{pq-1}$.
 
@@
 
-According to Theorem~\ref{th:failurestest}, the tests $U_F(k)$ need to be
-executed for $k = 0,\dots,pq-1$ to guarantee completeness. \fixme{Confusing
-use of $p$ and $q$ again.} This means that the SUT is verified with test
+According to Theorem~\ref{th:failurestest}, the tests $U_F(j)$ need to be
+executed for $j = 0,\dots,pq-1$ to guarantee completeness. \fixme{Confusing
+use of $p$ and $q$ again.}
+\fxnote{jp: fixed}
+ This means that the SUT is verified with test
 traces up to, and including, length $pq$: recall from the test specification,
-branch (\ref{eq:ufa}), that $U_F(k)$ will accept all traces $s.e$ with
-$s\in\trc(P), \#s = k, e\not\in\trc(P/s)$, so erroneous traces up to length
-$k+1$ are detected.
+branch (\ref{eq:ufa}), that $U_F(j)$ will accept all traces $s.e$ with
+$s\in\trc(P), \#s = j, e\not\in\trc(P/s)$, so erroneous traces up to length
+$j+1$ are detected.
 
@@
 specified as follows.
+
+\begin{center}
+\begin{minipage}{.4\textwidth}
 \begin{eqnarray*}
@@
 P_1 & = & a \then P \extchoice b\then P
-\\ & &  \\
+\end{eqnarray*}
+\end{minipage}
+\hfill
+\begin{minipage}{.4\textwidth}
+\begin{eqnarray*}
 Q & = & a\then Q_1 \extchoice b\then Q_1
@@
 \end{eqnarray*}
+\end{minipage}
+\end{center}
+
+\medskip
 Obviously, $P$'s normalised transition graph has 2 nodes, while $Q$'s graph
@@
 \[
-(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k))\hide \Sigma
+(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j))\hide \Sigma
 \]
@@
 role.}
+\fxnote{jp: should be clear now, since we have explained in Section 3 that only 
+the graph product $G(P)\times G(Q)$ is   needed in the proofs}
 %

### 2018-10-03 paper/conclusion.tex
further updates: use j instead of k in U_F(j)

@@
 As explained in the previous sections, passing a test is specified by
-$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for
+$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U(F(j)) \hide \Sigma$ for
 failures testing. If the SUT $Q$ is not a programmed piece of software or an
@@
 the FDR4 model checker. For checking the refinement relation $P\lessdet_F Q$,
-the pass criterion has to be verified for $k=0,\dots,pq-1$, where $p$ and $q$
+the pass criterion has to be verified for $j=0,\dots,pq-1$, where $p$ and $q$
 indicate the number of nodes in $P$'s transition graph and the maximal number
 of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since
-the test cases $U_F(k)$ have such a simple structure, it is an interesting
+the test cases $U(F(j)$ have such a simple structure, it is an interesting
 question for further research whether checking $(\epass\then\Stop) \lessdet_F
-(Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster
+(Q\parallel[\Sigma] U(F(j)) \hide \Sigma$ for $j=0,\dots,pq-1$ can be faster
 than directly checking $P\lessdet_F Q$, as one would do in the usual approach
@@
 \[
-(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{k=0}^{pq-1} U_F(k)) \hide \Sigma
+(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{j=0}^{pq-1} U_F(j)) \hide \Sigma
 \]

### 2018-10-03 paper/finitecomplete-failures.tex
further updates: use j instead of k in U_F(j)

@@
 and each integer $p\ge 0$, we define a test for failures refinement as shown
-below. %%
-%\begin{eqnarray}
-%U_F(p) & = & U_F(p,\varepsilon)
-%\\
-%U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
-%\label{eq:ufa}
-%\\ & & \extchoice \nonumber
-%\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
-%\label{eq:ufb}
-%\\ & & \extchoice \nonumber
-%\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
-%\label{eq:ufc}
-%\\ & & \extchoice \nonumber
-%\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber
-%\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
-%\label{eq:ufd}
-%\end{eqnarray}
-%
+below. 
 %
@@
 
+% -------------------------------------------------------------------------
 \begin{theorem}\label{th:failurestest}
@@
 \[
-\TS_F = \{ U_F(k)~|~0 \le k < pq  \}\quad\text{$U_F(k)$ specified in (\ref{eq:UFP})}
+\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{$U_F(j)$ specified in (\ref{eq:UFP})}
 \]
@@
 \end{theorem}
+% ------------------------------------------------------------------------
 %
@@
 conf\ P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq
-\trc(P)$, any adaptive test $U_F(p)$ running in parallel with $Q$ will always
+\trc(P)$, any adaptive test $U_F(j)$ running in parallel with $Q$ will always
 enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the
-external choice construction for $U_F(p,k,n)$. If $U_F(p,k,n) = U_F(p)/s$
-with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties} implies $U_F(p,k,n) =
-U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
+external choice construction for $U_F(j,k,n)$. To see this, consider
+ $U_F(j,k,n) = U_F(j)/s$
+with $s\in\trc(Q)$. Lemma~\ref{lemma:ufproperties} implies $U_F(j,k,n) =
+U_F(j,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
 $[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
-entered in the parallel execution of $Q$ and $U_F(p)$, and the $fail$ event
+entered in the parallel execution of $Q$ and $U_F(j)$, and the $fail$ event
 cannot occur.
@@
 hitting set for $\minaccs(Q/s)$. From Lemma~\ref{lemma:ufproperties} we know
-that $U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, so
+that $U_F(j)/s = U_F(j,k,n)$ with $U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so
 $\minhits(n) = \minhits(P/s)$. The branch (\ref{eq:ufb}) of the test
-$U_F(p,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
+$U_F(j,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
 \varnothing$. The branch (\ref{eq:ufc}) always leads to test continuation
@@
 It is not guaranteed, however, that trace $s$ is short enough to be executed by one of the 
-test cases $U_F(p)$. Therefore, it has to be shown that for any $s$ leading to an error
+test cases $U_F(j)$ with $0\le j < pq$. Therefore, it has to be shown that for any $s$ leading to an error
 situation specified by Case~1 or Case~2, there exists a trace $u$ of maximal length $pq-1$
@@
 hiding. I think the definition of test execution needs to change.}
-\fxnote{jp: this is a misunderstanding: we are not looking at the product graph of $Q$ and $U_F(p)$, but at the product graph of $Q$ and $P$. I have added more explanatory 
+\fxnote{jp: this is a misunderstanding: we are not looking at the product graph of $Q$ and $U_F(j)$, but at the product graph of $Q$ and $P$. I have added more explanatory 
 text in the paragraph  before Case~1.}

### 2018-10-04 paper/complexity.tex
New example and extended proof of Theorem 5 in section on Complexity Considerations.

@@
 required when testing for refinement.
+%\begin{example}\label{ex:pq}
+%Consider the CSP reference process $P$ and an erroneous implementation $Q$
+%specified as follows.
+%
+%\begin{center}
+%\begin{minipage}{.4\textwidth}
+%\begin{eqnarray*}
+%P & = & a \then P_1 \intchoice b \then P_1 \intchoice c \then P_1
+%\\
+%P_1 & = & a \then P \extchoice b\then P
+%\end{eqnarray*}
+%\end{minipage}
+%\hfill
+%\begin{minipage}{.4\textwidth}
+%\begin{eqnarray*}
+%Q & = & a\then Q_1 \extchoice b\then Q_1
+%\\
+%Q_1 & = & a\then Q_2 \extchoice b\then Q_2
+%\\
+%Q_2 & = & a\then Q \intchoice b\then Q
+%\end{eqnarray*}
+%\end{minipage}
+%\end{center}
+%
+%
+%\medskip
+%Obviously, $P$'s normalised transition graph has 2 nodes, while $Q$'s graph
+%has 3. It is easy to see (and can be checked with FDR4) that $P\lessdet_T Q$,
+%but $\neg(P\lessdet_F Q)$. Furthermore, it can also be shown using FDR4 that
+%the ``test passed condition''
+%\[
+%(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j))\hide \Sigma
+%\]
+%holds for $U_F(0),\dots,U_F(4)$, but fails for $U_F(5)$. So, the
+%non-conformance of $Q$ cannot be detected by any test trace of length less or
+%equal to 5, but is revealed (as expected from Theorem~\ref{th:failurestest})
+%by a trace of length 6, because the last event offered by the test $U_F(5)$
+%is refused by $Q$. \xbox
+%\end{example}
+
+\begin{figure}[htbp]
+\begin{center}
+\begin{minipage}{.4\textwidth}
+\includegraphics[width=1.2\textwidth]{theorem5p.pdf}
+\end{minipage}
+\hfill
+\begin{minipage}{.55\textwidth}
+\includegraphics[width=1.2\textwidth]{theorem5q.pdf}
+\end{minipage}
+\caption{Transition graphs of $P$ (left) and $Q$ (right) from Example~\ref{ex:pq}
+ for $p=3$ and $q=4$.}
+\label{fig:examplepq}
+\end{center}
+\end{figure}
+
 \begin{example}\label{ex:pq}
-Consider the CSP reference process $P$ and an erroneous implementation $Q$
+Consider 
+\fxnote{jp: I have updated the example with $P$ and $Q$ as used in the proof of the theorem. This should make thinks more readable.}
+\fixme{The internal choice became external. To test for traces refinement, yo
+put forward just one adaptive tests. So, there is more going on here.I think
+it is because your tests are checking both traces and failures refinement.}
+\fxnote{jp: no -- you are right: we want {\it internal} choice for $P$. With external
+choice, this only shows that when testing for trace refinement, the length $pq$ is required.}
+the CSP reference process $P$ and an erroneous implementation $Q$
 specified as follows.
 
-\begin{center}
-\begin{minipage}{.4\textwidth}
 \begin{eqnarray*}
-P & = & a \then P_1 \intchoice b \then P_1 \intchoice c \then P_1
+P & = &  P(0)
 \\
-P_1 & = & a \then P \extchoice b\then P
+P(k) & = & (k < p-1) \& \big( (a \then P(k)) \intchoice ( b \then P(k+1))\big)
+\\ & & \extchoice
+\\ & & (k = p-1) \& (a \then P(k))
+\\
+Q & = & Q(0)
+\\
+Q(k) & = & (k < q-1) \& \big( a \then Q(k+1)    \big) 
+\\ & & \extchoice
+\\ & & ( k = q-1)\& \big( a\then Q(0) \extchoice b\then Q(0)  \big)
 \end{eqnarray*}
-\end{minipage}
-\hfill
-\begin{minipage}{.4\textwidth}
-\begin{eqnarray*}
-Q & = & a\then Q_1 \extchoice b\then Q_1
-\\
-Q_1 & = & a\then Q_2 \extchoice b\then Q_2
-\\
-Q_2 & = & a\then Q \intchoice b\then Q
-\end{eqnarray*}
-\end{minipage}
-\end{center}
 
 \medskip
-Obviously, $P$'s normalised transition graph has 2 nodes, while $Q$'s graph
-has 3. It is easy to see (and can be checked with FDR4) that $P\lessdet_T Q$,
-but $\neg(P\lessdet_F Q)$. Furthermore, it can also be shown using FDR4 that
-the ``test passed condition''
+The normalised transition graphs of $P$ and $Q$ are depicted in Fig.~\ref{fig:examplepq}
+for the case $p=3,\ q=4$. Using FDR4, it can be shown for concrete values of
+$p$ and $q$ that the  ``test passed conditions''
 \[
@@
 \]
-holds for $U_F(0),\dots,U_F(4)$, but fails for $U_F(5)$. So, the
-non-conformance of $Q$ cannot be detected by any test trace of length less or
-equal to 5, but is revealed (as expected from Theorem~\ref{th:failurestest})
-by a trace of length 6, because the last event offered by the test $U_F(5)$
-is refused by $Q$. \xbox
+and
+\[
+(\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(j))\hide \Sigma
+\]
+hold for $j = 0,\dots,pq-2$. This means, that none of the test cases $U_F(j)$
+and $U_T(j)$ are capable of detecting failures and trace refinement violations, 
+respectively, if they only check traces up to length $pq-1$ (recall that this
+corresponds to $j\le pq-2$).
+
+Process $Q$, however, neither conforms to $P$ in the failures refinement 
+relation, nor in the trace
+refinement relation. This can only be seen when executing the test $U_F(pq-1)$ and
+$U_T(pq-1)$, respectively. These test fail, so this shows that $P\not\lessdet_F Q$ and
+$P\not\lessdet_F Q$ according to Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this shows that
+the maximal trace
+length $pq$ to be investigated in the tests cannot be further reduced without losing
+the completeness property of the test suites.
+\xbox
 \end{example}
+
 \fixme{But, there is no argument that the proofs based on graph products
@@
 \fxnote{jp: should be clear now, since we have explained in Section 3 that only 
-the graph product $G(P)\times G(Q)$ is   needed in the proofs}
+the graph product $G(P)\times G(Q)$ is   needed in the proofs.}
 %
@@
 $2\le p,q \in\mathbb{N}$,
-there exist reference processes with $p$ states
-and implementation processes with $q$ states, such that
+there exist reference processes $P$ with $p$ states
+and implementation processes $Q$ with $q$ states, such that
 a violation of the trace refinement property
 can only be detected with a trace of length $pq$. This is proven in the following
-theorem.
+theorem. In the proof, we use the processes $P$ and $Q$ introduced in Example~\ref{ex:pq}.
 
+% ----------------------------------------------------------------------------
 \begin{theorem}\label{th:maxtracelen}
@@
 \item $G(Q)$ has $q$ states.
-\item $P\not\lessdet_T Q$.
+\item $P\not\lessdet_T Q$, and therefore, also $P\not\lessdet_F Q$.
 \item $\forall s\in\trc(Q): \#s < pq\implies s\in\trc(P)$.
+\item $Q\ conf\ P$.
 \end{enumerate}
-As a consequence, the upper bound $pq$ for the length of traces to be tested when checking for failures refinement cannot be reduced without losing the test suite's completeness property.
+As a consequence, the upper bound $pq$ for the length of traces to be tested when checking for failures refinement or trace refinement 
+cannot be reduced without losing the test suite's completeness property.
 \end{theorem}
+% ----------------------------------------------------------------------------
 \begin{proof}
-Given $2\le p,q \in\mathbb{N}$, define reference process $P$ and implementation process $Q$
-over alphabet $\Sigma =\{ a,b\}$ as follows.
-\begin{eqnarray*}
-P & = &  P(0)
-\\
-P(k) & = & \big( a \then P(k) \big) \extchoice (k < p-1) \& \big( b \then P(k+1)\big)
-\\
-Q & = & Q(0)
-\\
-Q(k) & = & (k < q-1) \& \big( a \then Q(k+1)    \big) \extchoice
-( k = q-1)\& \big( a\then Q(0) \extchoice b\then Q(0)  \big)
-\end{eqnarray*}
-\fixme{The internal choice became external. To test for traces refinement, yo
-put forward just one adaptive tests. So, there is more going on here.I think
-it is because your tests are checking both traces and failures refinement.}
+Given $2\le p,q \in\mathbb{N}$, define reference process $P$ and implementation process $Q$ as in Example~\ref{ex:pq}. It is trivial to see that 
+$G(P)$ has $p$ nodes and $G(Q)$ has $q$ nodes, so statements 1 and 2 of the theorem hold.
+
 Using regular expression notation, the traces of $P$ can be specified as
@@
 \]
-where $\prefs(M)$ denotes the set of all prefixes of traces in $M\subset\Sigma^*$,
+where $\prefs(M)$ denotes the set of all prefixes of traces in $M\subseteq\Sigma^*$,
 including the traces of $M$ themselves.
@@
 It is easy to see that $\trc(Q)\not\subseteq\trc(P)$; for example, the trace
-$(a^{q-1}b)^p$ is in $\trc(Q)\setminus\trc(P)$.
+$(a^{q-1}b)^p$ is in $\trc(Q)\setminus\trc(P)$, because $P$-traces may contain at most $p-1$ $b$-events. This proves statement~3 of the theorem.
 
@@
 $\prefs\big(  (a^*b)^{p-1}a^* \big)$ contains all finite sequences of $a$-events,
-where at most $p-1$ events $b$ have been inserted.
+where at most $p-1$ events $b$ have been inserted. This proves statement 4 of the theorem.
+
+To prove statement~5, we observe that the specification of $P$ implies
+\[
+\minaccs(P/s) = \left\{
+\begin{array}{ll}
+\{ \{a\}, \{b\} \} & \text{for all $s\in\trc(P)$ with $\#s<p-1$.}
+\\
+\{ \{a\} \} & \text{for all $s\in\trc(P)$ with $\#s\ge p-1$.}
+\end{array}
+\right.
+\]
+and
+\[
+\minaccs(Q/s) = \left\{
+\begin{array}{ll}
+\{ \{a\}  \} & \text{for all $s\in\trc(Q)$ with $\#s \neq 0\mod(q-1)$.}
+\\
+\{ \{a,b\} \} & \text{for all $s\in\trc(P)$ with $\#s = 0\mod(q-1)$.}
+\end{array}
+\right.
+\]
+As a  consequence, the minimal acceptance set $A_P = \{a\}$ which is
+contained in every $\minaccs(P/s)$ fulfils $A_P \subseteq A_Q$ for
+any $A_Q\in\minaccs(Q/s)$, when $s\in\trc(P)\cap \trc(Q)$. Now Lemma~\ref{lemma:tgtrcref}, 
+(\ref{eq:failrefb}) can be applied to conclude that $Q\ conf\ P$.
 \xbox

### 2018-10-04 paper/casestudies.tex
further updates: use j instead of k in U_F(j)
Fixed errors found by Wen-ling

@@
 \begin{center}
-\begin{minipage}{0.3\textwidth} 
+\begin{minipage}{0.2\textwidth} 
 	 \begin{eqnarray*}
@@
 	\hfill
-	\begin{minipage}{0.3\textwidth}
+	\begin{minipage}{0.33\textwidth}
 	 \begin{eqnarray*}
@@
 	\hfill
-	\begin{minipage}{0.3\textwidth}
+	\begin{minipage}{0.2\textwidth}
 	 \begin{eqnarray*}
@@
 t(1,b) & = & 0 \\
-t(1,c) & = & 2 \\
+t(1,c) & = & 2  
+\end{eqnarray*}
+	\end{minipage}
+		\hfill
+	\begin{minipage}{0.2\textwidth}
+	 \begin{eqnarray*}
 t(2,a) & = & 1 \\
@@
 	\end{minipage}
+
 \caption{Initials, minimal hitting sets, and transition function of the normalised transition graph displayed in Fig.~\ref{fig:tga}.}
@@
 from Example~\ref{example:CSP} that is erroneous from the point of view of
-failures refinement.
+failures refinement. In the specification of $Z$, it is assumed that $r_{max}\ge 0$. 
 \begin{eqnarray*}
@@
 
-Running the test $U_F(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p = 4$
-states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test depth
-to be used according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
+Running the test $U_F(j)$ against $Z$ for $j=0,\dots,19$ ($G(P)$ has $p = 4$
+states and $G(Z)$ has $q=5$, so $pq-1=19$ is the index of the last test to 
+be executed  according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
 U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$, because after execution of
@@
 from Fig.~\ref{fig:tgZ} that
-$G(Z)/s = 4$), while $U_F(4)/s$
+$G(Z)/s = \text{node}\ 4$), while $U_F(4)/s$
 only offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for

### 2018-10-04 paper/complexity.tex
Fixed errors found by Wen-ling

@@
 refinement relation. This can only be seen when executing the test $U_F(pq-1)$ and
-$U_T(pq-1)$, respectively. These test fail, so this shows that $P\not\lessdet_F Q$ and
-$P\not\lessdet_F Q$ according to Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this shows that
+$U_T(pq-1)$, respectively. These tests fail, so this shows that $P\not\lessdet_F Q$ and
+$P\not\lessdet_T Q$ according to Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this shows that
 the maximal trace
@@
 
-To prove statement~5, we observe that the specification of $P$ implies
+To prove statement~5, we observe that the specification of $P$ implies (the 
+expression $(s\cnt b)$ denotes the number of $b$-events occurring in trace $s$)
 \[
@@
 \begin{array}{ll}
-\{ \{a\}, \{b\} \} & \text{for all $s\in\trc(P)$ with $\#s<p-1$.}
+\{ \{a\}, \{b\} \} & \text{for all $s\in\trc(P)$ with $(s\cnt b) <p-1$.}
 \\
-\{ \{a\} \} & \text{for all $s\in\trc(P)$ with $\#s\ge p-1$.}
+\{ \{a\} \} & \text{for all $s\in\trc(P)$ with $(s\cnt b) = p-1$.}
 \end{array}

### 2018-10-04 paper/finitecomplete-failures.tex
Fixed errors found by Wen-ling

@@
 $$
-and each integer $p\ge 0$, we define a test for failures refinement as shown
+and each integer $j\ge 0$, we define a test for failures refinement as shown
 below. 
@@
 \label{eq:UFP}
-U_F(p) & = & U_F(p,0,\ii n)
+U_F(j) & = & U_F(j,0,\ii n)
 \\
-U_F(p,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
+U_F(j,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
 \label{eq:ufa}
@@
 \\ & & \extchoice \nonumber
-\\ & & (k < p) \& \big(e:[n]^0   \then U_F(p,k+1,t(n,e) \big)
+\\ & & (k < j) \& \big(e:[n]^0   \then U_F(j,k+1,t(n,e) \big)
 \label{eq:ufc}
 \\ & & \extchoice \nonumber
-\\ & & (k = p \wedge \minhits(n)\neq\varnothing) \&  \nonumber
+\\ & & (k = j \wedge \minhits(n)\neq\varnothing) \&  \nonumber
 \\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
@@
 \subsubsection*{Explanation of the Test Definition}
-A test is performed by running $U_F(p)$ concurrently with any SUT process
+A test is performed by running $U_F(j)$ concurrently with any SUT process
 $Q$, synchronising over $\Sigma$. So, a \emph{test execution} is a trace of the
@@
 %\[
-$Q\parallel[\Sigma] U_F(p)$.
+$Q\parallel[\Sigma] U_F(j)$.
 %\]
@@
 assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
-or $\epass$ always become visible, if they are the only events $U_F(p)/s$ is
-ready to engage in: if $U_F(p)/s$ can only produce $\epass$ or $\efail$, the
+or $\epass$ always become visible, if they are the only events $U_F(j)/s$ is
+ready to engage in: if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the
 occurrence of these events can never be blocked due to a livelock in $Q$
@@
 
-The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and
-only if, {\it every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with
+The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(j)$) if, and
+only if, {\it every} execution of $Q\parallel[\Sigma] U_F(j)$ terminates with
 the event $\epass$. This can also be  expressed by means of a failures
@@
 \label{eq:passF}
-Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
+Q\ \pass\ U_F(j) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma
 \end{equation}
@@
 We note that it is necessary to use failures refinement in the definition
-above, and not just trace refinement:~$(Q\parallel[\Sigma] U_F(p)) \hide
+above, and not just trace refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide
 \Sigma$ may have the same visible traces $\varepsilon$ and $\epass$ as the
@@
 nondeterministically refuse $\epass$, due to a deadlock occurring when a
-faulty SUT process executes concurrently with $U_F(p,k,n)$ executing
-branch~(\ref{eq:ufd}), when the guard condition $(k = p\wedge
+faulty SUT process executes concurrently with $U_F(j,k,n)$ executing
+branch~(\ref{eq:ufd}), when the guard condition $(k = j\wedge
 \minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
 in the next paragraphs.
+Alternatively, a faulty SUT $Q$ 
+might internally deadlock after a trace $s$ where $\#s < j$ and
+$\minhits(G(P)/s) \neq \varnothing$,
+so that $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks as well.
 
-Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
-a length $p$. If, after having already run through $s\in\trc(P)$ with $\#s
-\le p$, an event is accepted by the SUT that is outside the initials of $P/s$
- (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
+Intuitively speaking, $U_F(j)$ is able to perform any trace $s$ of $P$, up to
+a length $j$. If, after having already run through $s\in\trc(P)$ with $\#s
+\le j$, an event is accepted by the SUT that is outside the initials of $P/s$
+ (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(j)/s$),
 the test immediately terminates with FAIL-event $\efail$. This is handled by
 the branch (\ref{eq:ufa}) of the external choice. %in the process
-%$U_F(p,\#s,n)$ defined above.
+%$U_F(j,\#s,n)$ defined above.
 
@@
 case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
-external choice in $U_F(p,\#s,G(P)/s)$). If $P/s$ may also
+external choice in $U_F(j,\#s,G(P)/s)$). If $P/s$ may also
 nondeterministically engage into events in such a situation, branch
@@
 engage into an event from $\Sigma - [P/s]^0$, a test execution exists where
-$U_F(p,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the $\efail$
+$U_F(j,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the $\efail$
 event.
 
-If the length of $s$ is still less than $p$, the test accepts any event $e$
+If the length of $s$ is still less than $j$, the test accepts any event $e$
 from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
-$U_F(p,\#s+1,G(P)/s.e)$ in branch~(\ref{eq:ufc}); this follows again from
+$U_F(j,\#s+1,G(P)/s.e)$ (note that $G(P)/s.e = t(G(P)/s,e)$) 
+in branch~(\ref{eq:ufc}); this follows again from
 Lemma~\ref{lemma:ufproperties}. A test of this type is called
@@
 
-Finally, we suppose that a test execution has run through a trace
-$s\in\trc(P)$ of length $p$, so that $U_F(p)/s = U_F(p,p,n)$ with $n =
+Now suppose that a test execution has run through a trace
+$s\in\trc(P)$ of length $j$, so that $U_F(j)/s = U_F(j,j,n)$ with $n =
 G(P)/s$. If $\minhits(n)\neq \varnothing$, the test changes its
@@
 Therefore, the test execution only terminates with  $\epass$, if such an
-event is accepted by the SUT. Otherwise, it deadlocks.
+event is accepted by the SUT. Otherwise, it deadlocks, and the test fails.
 
-The specification of $U_F(p,k,n)$ implies that the test always stops after
-having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$. If
+The specification of $U_F(j,k,n)$ implies that the test always stops after
+having engaged into a trace $s\in\trc(Q)$ of maximal length $j$ or $j+1$. If
 branch (\ref{eq:ufa}) is the last to be entered, the maximal length of $s$ is
-$p+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
-is the last to be entered, the maximal length of $s$ is $p$, and the
+$j+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
+is the last to be entered, the maximal length of $s$ is $j$, and the
 execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
@@
 set $H\in\minhits(n)$ with $n = G(P)/s$ according to
-Lemma~\ref{lemma:ufproperties}. Then the final length of $s$ is $p+1$, and
+Lemma~\ref{lemma:ufproperties}. Then the final length of $s$ is $j+1$, and
 the execution terminates with $\epass$. Or the test execution
-$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
+$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks,  the final length of $s$ is $j$,
 and the execution stops without a PASS or FAIL event. Such an execution is
 also interpreted as FAIL, because it reveals that $(\epass\then\Stop)
-\not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.
+\not\lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$.
 
 We observe that the number of possible executions of $Q\parallel[\Sigma]
-U_F(p)$ is finite, because the number of traces $s$ with maximal length
-$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and $\minhits(n)$
+U_F(j)$ is finite, because the number of traces $s$ with maximal length
+$(j+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and $\minhits(n)$
 are finite. Moreover, we further recall that $\minhits(n)$ may be empty, in
@@
 
-%Before explaining the intuition behind this test definition in the paragraphs
-%below, we state its basic properties in the following lemma, to show the
-%relationships between $U_F(p)$ and the reference process $P$ from which the
-%former has been derived. \fixme{For a reader, it is better to see the
-%explanation first. Since this is easier to undo, I've changed it. }
-
 We state basic properties of our tests in the following lemma, to show the
-relationships between $U_F(p)$ and the reference process $P$ from which it is
+relationships between $U_F(j)$ and the reference process $P$ from which it is
 derived.
@@
 \begin{lemma}\label{lemma:ufproperties}
-If $s\in\trc(P)$ satisfies $\#s\le p$, then
-$s, s.e\in\trc(U_F(p))$ for all $e\in\Sigma$, and the following properties hold.
+If $s\in\trc(P)$ satisfies $\#s\le j$, then
+$s, s.e\in\trc(U_F(j))$ for all $e\in\Sigma$, and the following properties hold.
 \begin{eqnarray}
 \label{eq:ifpa}
-  &  & U_F(p)/s = U_F(p,\#s,G(P)/s)
+  &  & U_F(j)/s = U_F(j,\#s,G(P)/s)
 \\
 \label{eq:ifpb}
-e\not\in [P/s]^0 & \implies & U_F(p)/s.e = (\efail\then\Stop)
+e\not\in [P/s]^0 & \implies & U_F(j)/s.e = (\efail\then\Stop)
 \\
 \label{eq:ifpc}
-U_F(p)/s = U(p,\# s,n)  & \implies & [n]^0 = [P/s]^0
+U_F(j)/s = U_F(j,\# s,n)  & \implies & [n]^0 = [P/s]^0
 \\
 \label{eq:ifpd}
-U_F(p)/s = U(p,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
+U_F(j)/s = U_F(j,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
 \end{eqnarray}
@@
 We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$,
-the statement holds because $U_F(p)$ starts with the initial node $\ii n$ of
+the statement holds because $U_F(j)$ starts with the initial node $\ii n$ of
 $G(P)$. Suppose that the statement holds for all traces $s$ with length $\# s
-\le k < p$, so that $U_F(p)/s = U_F(p,\#s,G(P)/s)$. Now let $s.e$ be a trace
+\le k < j$, so that $U_F(j)/s = U_F(j,\#s,G(P)/s)$. Now let $s.e$ be a trace
 of $P$, so that $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces
-$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(p,\#s,G(P)/s)$ can
+$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(j,\#s,G(P)/s)$ can
 engage into $e$ by executing branch (\ref{eq:ufc}). Since $t$ is the
@@
 defined, and $t(G(P)/s,e) = G(P)/s.e$. This leads to a new recursion, so that
-$U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = U_F(p,\#s+1,G(P)/s.e)$ as required.
+$U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = U_F(j,\#s+1,G(P)/s.e)$ as required.
 
-To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(p)/s
-= U_F(p,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
+To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(j)/s
+= U_F(j,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
 $[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
-$U_F(p,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
+$U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
 specification of this branch implies that
 $$
-U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = (\efail\then\Stop).
+U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = (\efail\then\Stop).
 $$
@@
 %
-Note that it is not guaranteed for $U_F(p)$ to run through the traces $s,
+Note that it is not guaranteed for $U_F(j)$ to run through the traces $s,
 s.e$ in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for
-some prefix $u$ of $s$: in such a case, $U_F(p)$ may stop with a $\epass$
+some prefix $u$ of $s$: in such a case, $U_F(j)$ may stop with a $\epass$
 event by entering branch (\ref{eq:ufb}). Therefore,
 Lemma~\ref{lemma:ufproperties} just states the existence of
-$U_F(p)$-executions $s, s.e$ satisfying the properties stated there.
+$U_F(j)$-executions $s, s.e$ satisfying the properties stated there.
 
@@
 
-A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a
+A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ counsists of a
 reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
@@
 $\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
-U_F(k))\hide \Sigma$.
+U_F(j))\hide \Sigma$.
 % alcc: part of the explanation above was in the following lemma, but it is already
@@
 It has to be shown for each of these cases that at least one test execution
-of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the $\efail$
+of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
 event or deadlocks. We do this by analysing the product graph of the reference

### 2018-10-04 paper/complexity.tex
All comments from Ana have been addressed, with exception of some items in the discussion.

@@
 According to Theorem~\ref{th:failurestest}, the tests $U_F(j)$ need to be
-executed for $j = 0,\dots,pq-1$ to guarantee completeness. \fixme{Confusing
-use of $p$ and $q$ again.}
-\fxnote{jp: fixed}
+executed for $j = 0,\dots,pq-1$ to guarantee completeness. 
  This means that the SUT is verified with test
@@
 %\end{example}
+
+\fixme{Confusing
+use of $p$ and $q$ again.}
+\fxnote{jp: fixed}
 

### 2018-10-04 paper/finitecomplete-failures.tex
All comments from Ana have been addressed, with exception of some items in the discussion.

@@
 A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
-exhaustive for the fault model specified there and for conformance relation $\lessdet_F$.
+exhaustive for the fault model specified there.
 \end{lemma}

### 2018-10-04 paper/finitecomplete-trace.tex
update
New example and extended proof of Theorem 5 in section on Complexity Considerations.
All comments from Ana have been addressed, with exception of some items in the discussion.

@@
 For establishing trace refinement, the following class of adaptive test cases
-will be used for a given reference process $P$ and  integers $p \ge 0$.
+will be used for a given reference process $P$ and  integers $j \ge 0$.
 Just as for the tests developed  in
@@
 \begin{eqnarray}
-U_T(p) & = & U_T(p,0,\ii n)
+U_T(j) & = & U_T(j,0,\ii n)
 \\
-U_T(p,k,n) & = & \big(  e:(\Sigma - [n]^0)   \then \efail\then \Stop \big)
+U_T(j,k,n) & = & \big(  e:(\Sigma - [n]^0)   \then \efail\then \Stop \big)
 \label{eq:uta}
@@
 \\ & & \extchoice \nonumber
-\\ & & (k < p) \& \big( e:[P/s]^0   \then U_T(p,k+1,t(n,e)) \big)
+\\ & & (k < j) \& \big( e:[P/s]^0   \then U_T(j,k+1,t(n,e)) \big)
 \label{eq:utc}
 \\ & & \extchoice \nonumber
-\\ & & (k = p) \& \big( \epass\then \Stop  \big)
+\\ & & (k = j) \& \big( \epass\then \Stop  \big)
 \label{eq:utd}
@@
 %
-It is easy to see that the tests $U_T(p)$ satisfy the properties
+It is easy to see that the tests $U_T(j)$ satisfy the properties
 \begin{eqnarray}
 \label{eq:ifpaT}
-  &  & U_T(p)/s = U_T(p,\#s,G(P)/s)
+  &  & U_T(j)/s = U_T(j,\#s,G(P)/s)
 \\
 \label{eq:ifpbT}
-e\not\in [P/s]^0 & \implies & U_T(p)/s.e = (\efail\then\Stop)
+e\not\in [P/s]^0 & \implies & U_T(j)/s.e = (\efail\then\Stop)
 \end{eqnarray}
-proven in Lemma~\ref{lemma:ufproperties} for $U_F(p)$ for traces
-$s\in\trc(P)$ with $\#s \le p$. 
+proven in Lemma~\ref{lemma:ufproperties} for $U_F(j)$ for traces
+$s\in\trc(P)$ with $\#s \le j$. 
 \fixme{In (37), I think there is an implicit assumption about
@@
 Since
-$U_T(p)$ never blocks any $Q$-event before terminating, the pass criterion
+$U_T(j)$ never blocks any $Q$-event before terminating, the pass criterion
 can be based on trace refinement instead of failures refinement as required
@@
 \label{eq:passT}
-Q\ \pass\ U_T(p) \defs (\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(p)) \hide \Sigma
+Q\ \pass\ U_T(j) \defs (\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(j)) \hide \Sigma
 \end{equation}
@@
 \fxnote{jp: I do not understand this remark -- we should discuss it on the phone.
-Please note the addition explanations I have added here below (\ref{eq:passT}).}
-
+Please note the additional explanations I have added here below (\ref{eq:passT}) and please read the full proof for Theorem3.}
+%
 If the SUT process $Q$ deadlocks internally after a trace $s$ where also the reference
@@
 that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$. As a consequence, branch
-(\ref{eq:utb}) of a test case execution state $U_T(p,k,n)$  
-with $\#s = k \le p$   can be entered
+(\ref{eq:utb}) of a test case execution state $U_T(j,k,n)$  
+with $\#s = k \le j$   can be entered
 and the  test execution terminates with $\epass$. If, however,  
@@
 neither $\epass$ nor $\efail$ is ever produced, so that
-$(Q\parallel[\Sigma] U_T(p)) \hide \Sigma = \Stop$, the test passes, because
+$(Q\parallel[\Sigma] U_T(j)) \hide \Sigma = \Stop$, the test passes, because
 $(\epass\then\Stop) \lessdet_T \Stop$ holds.
@@
 %
+% -------------------------------------------------------------------------
 \begin{theorem}\label{th:tracetest}
@@
 \end{theorem}
+% -------------------------------------------------------------------------
 %\begin{proof}
@@
 $$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$ because of (\ref{eq:ifpaT}). Now $\trc(Q)\subseteq\trc(P)$ 
-implies $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot enter branch (\ref{eq:uta}) and produce a $\efail$-event. Therefore, only four options are
-available for the test execution    $(Q\parallel[\Sigma] U_T(p))/s$ to continue.
+implies $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot enter branch (\ref{eq:uta}) and produce a $\efail$-event when running in parallel with $Q$ and synchronising over $\Sigma$.
+Therefore, only four options are
+available for the test execution    $(Q\parallel[\Sigma] U_T(j))/s$ to continue.
 
@@
 {\bf Case~2.} $Q/s$ deadlocks, but $\minaccs(G(P)/s)\neq\{\varnothing\}$. 
-Then the whole test execution blocks, and this is means that
+Then the whole test execution blocks, and this   means that
 neither a $\epass$ nor a $\efail$ event is produced, so the test execution is passed.
@@
 
-This case analysis shows that every   execution of $(Q\parallel[\Sigma] U_T(p))$
+This case analysis shows that every   execution of $(Q\parallel[\Sigma] U_T(j))$
 either stops after $\epass$ or produces neither $\epass$ nor $\efail$. This
@@
 A test suite $\TS_T$ specified as in Theorem~\ref{th:tracetest} is
-exhaustive for the fault model specified there and for conformance relation $\lessdet_T$.
+exhaustive for the fault model specified there.
 \end{lemma}

### 2018-10-05 paper/conclusion.tex
Fixed a few typos

@@
 adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
-U_F(p))$ for failures refinement never block before the final step specified
+U_F(j))$ for failures refinement never block before the final step specified
 by branch (\ref{eq:ufd}), and so we do not need inconclusive test results.
@@
 behaviour is considered to be blocked and regarded as a failure.
-The tests executions $(Q\parallel[\Sigma] U_T(p))$ 
+The tests executions $(Q\parallel[\Sigma] U_T(j))$ 
 \fixme{Missing hiding.}
-\fxnote{jp: Here, the hiding operator is not necessary: the possible test executions are really the traces of $Q$ parallel $U_T(p)$. The hiding is only needed to specify the verdict.}  
+\fxnote{jp: Here, the hiding operator is not necessary: the possible test executions are really the traces of $Q$ parallel $U_T(j)$. The hiding is only needed to specify the verdict.}  
 for trace refinement never
@@
 blocking situation also leads to passing the test, because 
-$(\epass\then \Stop)\lessdet_T (Q\parallel[\Sigma] U_T(p))\hide\Sigma$ still holds
-if $(Q\parallel[\Sigma] U_T(p))\hide\Sigma$ only produces the empty trace.
+$(\epass\then \Stop)\lessdet_T (Q\parallel[\Sigma] U_T(j))\hide\Sigma$ still holds
+if $(Q\parallel[\Sigma] U_T(j))\hide\Sigma$ only produces the empty trace.
 
-The adaptive behaviour of both our test case types $U_F(p)$ and $U_T(p)$,
+The adaptive behaviour of both our test case types $U_F(j)$ and $U_T(j)$,
 however, induces the obligation to check that {\it all}
@@
 As explained in the previous sections, passing a test is specified by
-$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U(F(j)) \hide \Sigma$ for
+$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ for
 failures testing. If the SUT $Q$ is not a programmed piece of software or an
@@
 of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since
-the test cases $U(F(j)$ have such a simple structure, it is an interesting
+the test cases $U_F(j)$ have such a simple structure, it is an interesting
 question for further research whether checking $(\epass\then\Stop) \lessdet_F
-(Q\parallel[\Sigma] U(F(j)) \hide \Sigma$ for $j=0,\dots,pq-1$ can be faster
+(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ for $j=0,\dots,pq-1$ can be faster
 than directly checking $P\lessdet_F Q$, as one would do in the usual approach

### 2018-10-07 paper/abstract.tex
checked fixmes and new material

@@
 %nondeterministic behaviour is described.
-It is shown that the test suites are optimal with respect to two aspects:  (a)  the maximal length of test traces cannot be reduced, and (b) the nondeterministic behaviour cannot be tested with smaller or fewer sets of events, without losing the test suite's completeness property.
-\fixme{alcc: Is the technique
-applying the test suites?}
-\fxnote{jp: this sentence needed to be re-phrased -- the new version should be better understandable}
+It is shown that the test suites are optimal with respect to two
+aspects:~(a)~the maximal length of test traces cannot be reduced, and (b)~the
+nondeterministic behaviour cannot be tested with smaller or fewer sets of
+events, without losing the test suite's completeness property.

### 2018-10-07 paper/complexity.tex
Fixed a few typos
checked fixmes and new material

@@
 
-In this 
-\fixme{In that section, $k$ is used for a different purpose. It is confusing.
-I would use $p$ here.} 
-\fxnote{jp: ok -- I also had to change this in Theorem~2, where $k$ was used as index from
-0 to $pq-1$. I now use $j$ as index for $U_F(j)$, because $p$ is the number of states in $G(P)$.}
-section, we
- will   calculate upper bounds for the total number of test executions to be performed
- when testing for failures refinement.   Theorem~\ref{th:failurestest} specifies
- that all tests $U_F(j),\ 0\le j < pq$ need to be executed, where $p$ denotes the
- number of nodes in the transition graph of the reference process $P$, and $q\ge p$ is an estimate for the number of nodes in the SUT's transition graph.
-Therefore, we will first calculate a bound for the number of test executions to be performed
- for test $U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.
+In this section, we calculate upper bounds for the total number of test
+executions to be performed when testing for failures refinement.
+Theorem~\ref{th:failurestest} specifies that all tests $U_F(j),\ 0\le j < pq$
+need to be executed, where $p$ denotes the number of nodes in the transition
+graph of the reference process $P$, and $q\ge p$ is an estimate for the
+number of nodes in the SUT's transition graph. Therefore, we will first
+calculate a bound for the number of test executions to be performed for test
+$U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.
 
-For the worst-case estimate, we
-assume that $P$ never allows for early deadlock (so $\minhits(P/s)$ is never empty) 
-and that the SUT $Q$ is a correct 
-failures refinement. Then all test executions $(Q\parallel[\Sigma] U_F(j))$
-stop after having run through a $Q$-trace of length $j+1$, because their is
-no early termination due to entering branches (\ref{eq:ufa}), (\ref{eq:ufb}), or due to
-an illegal deadlock of $Q$.
-As can be seen from the specification of the test cases $U_F(j)$ (see
-Section~\ref{sec:finitecompletefails}), the number of
-executions ending in a $\epass$ event corresponds to the number $\ell$ of
-traces $s$ of $P$ with length equal to  
-$j$, multiplied by the number $h$ of minimal hitting sets in
-$\minhits(P/s)$. For
-the tests $U_T(j)$ verifying trace refinement (see
+For the worst-case estimate, we assume that $P$ never allows for early
+deadlock (so $\minhits(P/s)$ is never empty) and that the SUT $Q$ is a
+correct failures refinement. Therefore, all test executions
+$(Q\parallel[\Sigma] U_F(j))$ stop after having run through a trace of $Q$ of
+length $j+1$, because their is no early termination due to entering branches
+(\ref{eq:ufa}), (\ref{eq:ufb}), or due to an illegal deadlock of $Q$. As can
+be seen from the specification of the test cases $U_F(j)$ (see
+Section~\ref{sec:finitecompletefails}), the number of executions ending in a
+$\epass$ event corresponds to the number $\ell$ of traces $s$ of $P$ with
+length equal to $j$, multiplied by the number $h$ of minimal hitting sets in
+$\minhits(P/s)$. For the tests $U_T(j)$ verifying trace refinement (see
 Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$,
 since there is no equivalent in $U_T(j)$ to checking different hitting sets
-in the last step of a test execution.
-\fixme{But doesn't this add to the number of executions?}
-\fxnote{jp: should be clear now from the revised test}
+in the last step of a test execution. \fixme{But doesn't this add to the
+number of executions?} \fxnote{jp: should be clear now from the revised test}
 
@@
 %
-From 
-\fixme{Theorem 4 is not
-about $conf$. I don't think you have tests for $conf$.}
-\fxnote{jp: I rephrased this -- hope that it's better understandable now.}
-Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
-above cannot be further reduced, since the full collection of minimal
-hitting sets needs to be checked at the end of each execution of $U_F(j)$.
+From Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
+above cannot be further reduced, since the full collection of minimal hitting
+sets needs to be checked at the end of each execution of $U_F(j)$.
 
-In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test 
-\fixme{In \cite{DBLP:conf/icfem/CavalcantiG07}, we work
-with minimal acceptances. I didn't check \cite{Hennessy:1988:ATP:50497}.}
-\fxnote{jp: we have studied \cite{DBLP:conf/icfem/CavalcantiG07} now in more detail -- you are right, of course, but there is a confusion of terms: you are also using the minimal hitting sets we handle here in our joint paper, but you used the term ``minimal acceptance'' for them.
-This is not in line with the definition of minimal acceptances as presented in \cite{Hennessy:1988:ATP:50497} and
-in Roscoe's books. Here, in our joint paper, minimal acceptances are introduced
-as defined in Roscoe's books. I have changed the text accordingly.}
-{\it every} non-empty
-subset of $\Sigma$ whose events cannot be completely refused in a given
-process state of the reference model; this leads to a worst-case estimate of
-$2^{\card{\Sigma}}-1$ for the number of different sets to be offered to the
-SUT in the last step of the test execution, so our approach reduces the
-number of test executions in comparison
-to~\cite{Hennessy:1988:ATP:50497} by a factor
-of $1/{\sqrt{2\pi \card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction
-is visualised by a function plot of $2^{\card{\Sigma}}$ versus
+In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test \fixme{In
+\cite{DBLP:conf/icfem/CavalcantiG07}, we work with minimal acceptances. I
+didn't check \cite{Hennessy:1988:ATP:50497}.} \fxnote{jp: we have studied
+\cite{DBLP:conf/icfem/CavalcantiG07} now in more detail -- you are right, of
+course, but there is a confusion of terms: you are also using the minimal
+hitting sets we handle here in our joint paper, but you used the term
+``minimal acceptance'' for them. This is not in line with the definition of
+minimal acceptances as presented in \cite{Hennessy:1988:ATP:50497} and in
+Roscoe's books. Here, in our joint paper, minimal acceptances are introduced
+as defined in Roscoe's books. I have changed the text accordingly.} {\it
+every} non-empty subset of $\Sigma$ whose events cannot be completely refused
+in a given process state of the reference model; this leads to a worst-case
+estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
+offered to the SUT in the last step of the test execution, so our approach
+reduces the number of test executions in comparison
+to~\cite{Hennessy:1988:ATP:50497} by a factor of $1/{\sqrt{2\pi
+\card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction is visualised by a
+function plot of $2^{\card{\Sigma}}$ versus
 $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.
 
-In~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors also use minimal hitting sets\footnote{However, they are denoted by {\it minimal acceptances} in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an upper bound for
-the number of test executions to be performed.
+In~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors also use minimal hitting
+sets\footnote{However, they are denoted by {\it minimal acceptances}
+in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an upper
+bound for the number of test executions to be performed.
 
@@
 
-According 
-\fixme{Please, see comments
-above about this process.}\fxnote{jp: should be ok now after the above changes and clarifications.}
-to Theorem~\ref{th:tracetest}, a complete test suite checking trace
+According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
 refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
@@
 
-\fixme{Confusing
-use of $p$ and $q$ again.}
-\fxnote{jp: fixed}
-
 \begin{figure}[htbp]
@@
 \begin{example}\label{ex:pq}
-Consider 
-\fxnote{jp: I have updated the example with $P$ and $Q$ as used in the proof of the theorem. This should make thinks more readable.}
-\fixme{The internal choice became external. To test for traces refinement, yo
-put forward just one adaptive tests. So, there is more going on here.I think
-it is because your tests are checking both traces and failures refinement.}
-\fxnote{jp: no -- you are right: we want {\it internal} choice for $P$. With external
-choice, this only shows that when testing for trace refinement, the length $pq$ is required.}
-the CSP reference process $P$ and an erroneous implementation $Q$
+Consider the CSP reference process $P$ and an erroneous implementation $Q$
 specified as follows.
-
+%
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-
-\medskip
-The normalised transition graphs of $P$ and $Q$ are depicted in Fig.~\ref{fig:examplepq}
-for the case $p=3,\ q=4$. Using FDR4, it can be shown for concrete values of
-$p$ and $q$ that the  ``test passed conditions''
+%
+The normalised transition graphs of $P$ and $Q$ are depicted in
+Fig.~\ref{fig:examplepq} for the case $p=3,\ q=4$. Using FDR4, it can be
+shown for concrete values of $p$ and $q$ that the  ``test passed conditions''
 \[
@@
 \]
-hold for $j = 0,\dots,pq-2$. This means, that none of the test cases $U_F(j)$
-and $U_T(j)$ are capable of detecting failures and trace refinement violations, 
-respectively, if they only check traces up to length $pq-1$ (recall that this
-corresponds to $j\le pq-2$).
+hold for $j = 0,\dots,pq-2$. This means that none of the test cases $U_F(j)$
+and $U_T(j)$ are capable of detecting failures and trace refinement
+violations, if they only check traces up to length $pq-1$
+(recall that this corresponds to $j\le pq-2$).
 
@@
 \end{example}
-
-\fixme{But, there is no argument that the proofs based on graph products
-corresponds to this process. In particular the hiding does not seem to play a
-role.}
-\fxnote{jp: should be clear now, since we have explained in Section 3 that only 
-the graph product $G(P)\times G(Q)$ is   needed in the proofs.}
 %
@@
 during tests, it can also be applied to our trace refinement tests.
-Therefore, test suites $\{ U_T(z) \}$ with $z<pq-1$ are not complete. It is
+Therefore, test suites $\{ U_T(j) \}$ with $j<pq-1$ are not complete. It is
 discussed in Section~\ref{sec:conc} how the number of test traces to be

### 2018-10-07 paper/conclusion.tex
checked fixmes and new material

@@
 
-In 
-\fixme{From the proofs of the main theorems, and what you say below, I
+In \fixme{From the proofs of the main theorems, and what you say below, I
 thought you did have verdicts coming from deadlock, and you simply chose not
-to mark it with an inc event.} 
-contrast to that, our test cases specified in
+to mark it with an inc event.} contrast to that, our test cases specified in
 Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
@@
 U_F(j))$ for failures refinement never block before the final step specified
-by branch (\ref{eq:ufd}), and so we do not need inconclusive test results.
-It should be noted, however, that it is
-necessary for our test verdicts to recognise also deadlocks in the final test
-step and interpret them as FAIL, as described in
-Section~\ref{sec:finitecompletefails}. In practice, this is realised by
-adding a timeout event to the testing environment which indicates deadlock
-situations. For real-time systems, this is an accepted technique, because the
-SUT has to respond within a pre-defined latency interval, otherwise its
-behaviour is considered to be blocked and regarded as a failure.
-The tests executions $(Q\parallel[\Sigma] U_T(j))$ 
-\fixme{Missing hiding.}
-\fxnote{jp: Here, the hiding operator is not necessary: the possible test executions are really the traces of $Q$ parallel $U_T(j)$. The hiding is only needed to specify the verdict.}  
-for trace refinement never
-block at all before stopping after the verdict $\epass$ or $\efail$, unless the
-SUT process $Q$ has an unexpected deadlock state. 
-\fxfatal{And also, if the SUT blocks, the test should block,
-no?}\fxnote{jp: you are right, I changed the text here and added explanations above after the verdict explanation
-in formula (\ref{eq:passT}).}
-Recall from the explanations given in Section~\ref{sec:finitecomplete} that the 
-blocking situation also leads to passing the test, because 
-$(\epass\then \Stop)\lessdet_T (Q\parallel[\Sigma] U_T(j))\hide\Sigma$ still holds
-if $(Q\parallel[\Sigma] U_T(j))\hide\Sigma$ only produces the empty trace.
+by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
+should be noted, however, that it is necessary for our test verdicts to
+recognise also deadlocks in the final test step and interpret them as FAIL,
+as described in Section~\ref{sec:finitecompletefails}. In practice, this is
+realised by adding a timeout event to the testing environment which indicates
+deadlock situations. For real-time systems, this is an accepted technique,
+because the SUT has to respond within a pre-defined latency interval,
+otherwise its behaviour is considered to be blocked and regarded as a
+failure. The tests executions $(Q\parallel[\Sigma] U_T(j))$ \fixme{Missing
+hiding.} \fxnote{jp: Here, the hiding operator is not necessary: the possible
+test executions are really the traces of $Q$ parallel $U_T(j)$. The hiding is
+only needed to specify the verdict.}for trace refinement never block at all
+before stopping after the verdict $\epass$ or $\efail$, unless the SUT
+process $Q$ has an unexpected deadlock state. Recall from the explanations
+given in Section~\ref{sec:finitecomplete} that the blocking situation also
+leads to passing the test, because $(\epass\then \Stop)\lessdet_T
+(Q\parallel[\Sigma] U_T(j))\hide\Sigma$ still holds if $(Q\parallel[\Sigma]
+U_T(j))\hide\Sigma$ only produces the empty trace.
 
 The adaptive behaviour of both our test case types $U_F(j)$ and $U_T(j)$,
-however, induces the obligation to check that {\it all}
-possible executions have been
-performed before the test can be considered as passed. Typically, it is
-therefore assumed that a \emph{complete testing
-assumption}~\cite{hierons_testing_2004} \fixme{I think you need this for the
-failures refinement tests as well.}\fxnote{jp: yes, I re-arranged the text accordingly.} holds, which means that every possible
-behaviour of the SUT is performed after a finite number of test executions.
-In practice, this is realised by executing each test several times, recording
-the traces that have been performed, and using hardware or software coverage
-analysers to determine whether all possible test execution behaviours of the
-SUT have been observed. Therefore, adaptive test cases come at the price of
-having to apply some grey-box testing techniques enabling us to decide
-whether all SUT behaviours have been observed.
+however, induces the obligation to check that {\it all} possible executions
+have been performed before the test can be considered as passed. Typically,
+it is therefore assumed that a \emph{complete testing
+assumption}~\cite{hierons_testing_2004} holds, which means that every
+possible behaviour of the SUT is performed after a finite number of test
+executions. In practice, this is realised by executing each test several
+times, recording the traces that have been performed, and using hardware or
+software coverage analysers to determine whether all possible test execution
+behaviours of the SUT have been observed. Therefore, adaptive test cases come
+at the price of having to apply some grey-box testing techniques enabling us
+to decide whether all SUT behaviours have been observed.
 
@@
 concerning the potential implications of this paper in the field of model
-checking. We are
-grateful to Li-Da Tong from National Sun Yat-sen University, Taiwan, for
-suggesting the applicability of Sperner's Theorem in the context of the work
-presented here. Moreover, we thank Adenilso Simao from the University of S{\~a}o Paulo
-for several helpful suggestions. 
-The work of Ana Cavalcanti is funded by the Royal Academy of
-Engineering and UK EPSRC Grant EP/R025134/1.
+checking. We are also grateful to Li-Da Tong from National Sun Yat-sen
+University, Taiwan, for suggesting the applicability of Sperner's Theorem in
+the context of the work presented here. Moreover, we thank Adenilso Simao for
+several helpful suggestions. The work of Ana Cavalcanti is funded by the
+Royal Academy of Engineering and UK EPSRC Grant EP/R025134/1.

### 2018-10-07 paper/finitecomplete-failures.tex
Fixed a few typos
checked fixmes and new material

@@
 \minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
-in the next paragraphs.
-Alternatively, a faulty SUT $Q$ 
-might internally deadlock after a trace $s$ where $\#s < j$ and
-$\minhits(G(P)/s) \neq \varnothing$,
-so that $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks as well.
+in the next paragraphs. Alternatively, a faulty SUT $Q$ might internally
+deadlock after a trace $s$ where $\#s < j$ and $\minhits(G(P)/s) \neq
+\varnothing$, so that the process $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks
+as well.
 
@@
 $U_F(j,\#s+1,G(P)/s.e)$ (note that $G(P)/s.e = t(G(P)/s,e)$)
-in branch~(\ref{eq:ufc}); this follows again from
+in branch~(\ref{eq:ufc}); this follows from
 Lemma~\ref{lemma:ufproperties}. A test of this type is called
@@
 \varnothing$. The branch (\ref{eq:ufc}) always leads to test continuation
-without a verdict. \fixme{I think we need a stronger argument here, as to why
-it cannot deadlock.}
-\fxnote{jp: I do not understand the question -- let's discuss this on the phone.}
- For the last branch, we note that it can only be entered
+without a verdict. Since, $s$ is a trace of $P$ and $Q$, we know that there
+is a behaviour of $Q$ that will accept an event $e$ in that branch and allow
+the test to proceed. \fixme{I think we need a stronger argument here, as to
+why it cannot deadlock.} \fxnote{jp: I do not understand the question --
+let's discuss this on the phone. See what I wrote. I'm still worried because
+of nondeterminism. } For the last branch, we note that it can only be entered
 if $\minhits(P/s) = \minhits(n) \neq\varnothing$. In this case, any selected
@@
 of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
-event or deadlocks. We do this by analysing the product graph of the reference
-process $P$ and the SUT process $Q$: any trace $s\in\trc(Q)\cap\trc(P)$ gives rise to 
-a path labelled by the events of $s$ 
-through this product graph. Any error can be detected after running through such a trace
-and then either observing an event outside $[P/s]^0$ (this is the violation described
-by Case~1) or identifying an illegal acceptance $A_Q$ as specified in Case~2. 
-It is not guaranteed, however, that trace $s$ is short enough to be executed by one of the 
-test cases $U_F(j)$ with $0\le j < pq$. Therefore, it has to be shown that for any $s$ leading to an error
-situation specified by Case~1 or Case~2, there exists a trace $u$ of maximal length $pq-1$
-leading to the same error.
+event or deadlocks. We do this by analysing the product graph of the
+reference process $P$ and the SUT process $Q$: any trace
+$s\in\trc(Q)\cap\trc(P)$ gives rise to a path labelled by the events of $s$
+through this product graph. Any error can be detected after running through
+such a trace and then either observing an event outside $[P/s]^0$ (this is
+the violation described by Case~1) or identifying an illegal acceptance $A_Q$
+as specified in Case~2. It is not guaranteed, however, that trace $s$ is
+short enough to be executed by one of the test cases $U_F(j)$ with $0\le j <
+pq$. So, it has to be shown that for any $s$ leading to an error situation
+specified by Case~1 or Case~2, there exists a trace $u$ of maximal length
+$pq-1$ leading to the same error.
 
@@
 \noindent {\bf Case~1.}
-\fixme{There is an assumption here that
-the graph of a test execution is given by the cross-product of graphs.  But,
-test execution is not defined in this way. It is defined by parallelism and
-hiding. I think the definition of test execution needs to change.}
-\fxnote{jp: this is a misunderstanding: we are not looking at the product graph of $Q$ and $U_F(j)$, but at the product graph of $Q$ and $P$. I have added more explanatory 
-text in the paragraph  before Case~1.}
 Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
@@
 \medskip
-\noindent {\bf Case~2.} We note 
-\fixme{Again, nothing here is about this process, but about the graph, and
-the hiding that is needed is discarded.} 
-\fxnote{jp: see my explanation above}
-that trace $s$ is again a trace of the
+\noindent {\bf Case~2.} We note that trace $s$ is again a trace of the
 product graph $G$, but we do not know its length. Again, by applying
@@
 of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$  always performs branch
-(\ref{eq:ufc}) until the trace $u$ has been completely processed.
-$U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
-only:~assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has
-at least one non-empty minimal acceptance. From (\ref{eq:minhitminaccempty})
-we know that this is equivalent to $\minhits(P/s) =
-\minhits(G(P)/s)\neq\varnothing$, and we observe that $G(P)/s = G(P)/u$, so
-$\minhits(G(P)/u)\neq\varnothing$. As a consequence, branch (\ref{eq:ufb})
-cannot be taken because its guard condition evaluates to $\isf$  for
-$U_F(\#u,\#u,G(P)/u)$. The guard condition $(k < p)$ for branch
-(\ref{eq:ufc}) evaluates to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch
-(\ref{eq:ufa}) is executed, the test always fails. If branch (\ref{eq:ufd})
-is executed, the test fails for the execution where a minimal hitting set
-$H\in\minhits(P/u)$ is chosen by $U_F(\#u,\#u,G(P)/u)$ that has an empty
-intersection with the minimal acceptance $A_Q$ from condition
-(\ref{eq:accsnotcontained}). The existence of such an $H$ is guaranteed
-because of Lemma~\ref{lemma:hseta}. As a consequence, there exists a test
-execution   where $Q/u$ selects acceptance $A_Q$ and $U_F(\#u,\#u,G(P)/u)$
-selects $H$. This execution deadlocks in process state
-$(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event;
-this means that the test fails. This
-concludes the proof. 
+(\ref{eq:ufc}) until the trace $u$ has been completely processed. \fixme{If
+the SUT accepts it? See previous fixme.} $U_F(\#u,\#u,G(P)/u)$ may execute
+branches (\ref{eq:ufa}) or (\ref{eq:ufd}) only:~assumption
+(\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has at least one
+non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) we know that
+this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and
+we observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
+consequence, branch (\ref{eq:ufb}) cannot be taken because its guard
+condition evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$. The guard condition
+$(k < j)$ for branch (\ref{eq:ufc}) evaluates to $\isf$ for
+$U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
+always fails. If branch (\ref{eq:ufd}) is executed, the test fails for the
+execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen by
+$U_F(\#u,\#u,G(P)/u)$ that has an empty intersection with the minimal
+acceptance $A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of
+such an $H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a
+consequence, there exists a test execution   where $Q/u$ selects acceptance
+$A_Q$ and $U_F(\#u,\#u,G(P)/u)$ selects $H$. This execution deadlocks in
+process state $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the
+$\epass$-event; this means that the test fails. This concludes the proof.
 \xbox
@@
 The next section presents and example.
-\fixme{I think
-the example should come before more theory.}
-\fxnote{jp: done}
 

### 2018-10-07 paper/finitecomplete-trace.tex
checked fixmes and new material

@@
 $s\in\trc(P)$ with $\#s \le j$.
-\fixme{In (37), I think there is an implicit assumption about
-$p$ and $\# s$?}\fxnote{jp: yes, I added $\#s \le p$ in the text.}
 
@@
 % ==========================================================================
-Since
-$U_T(j)$ never blocks any $Q$-event before terminating, the pass criterion
-can be based on trace refinement instead of failures refinement as required
-in (\ref{eq:passF}).
+Since $U_T(j)$ never blocks any event of $Q$ before terminating, the pass
+criterion can be based on trace instead of failures refinement as
+required in (\ref{eq:passF}).
 %
@@
 \end{equation}
-\fixme{The proofs don't use this definition. It should change.}
-\fxnote{jp: I do not understand this remark -- we should discuss it on the phone.
-Please note the additional explanations I have added here below (\ref{eq:passT}) and please read the full proof for Theorem3.}
 %
-If the SUT process $Q$ deadlocks internally after a trace $s$ where also the reference
-process $P$ is in a state where deadlock is possible, this is captured by the fact
-that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$. As a consequence, branch
-(\ref{eq:utb}) of a test case execution state $U_T(j,k,n)$  
-with $\#s = k \le j$   can be entered
-and the  test execution terminates with $\epass$. If, however,  
-$Q$ blocks after a trace $s'$
-where the reference process satisfies $\minaccs(P/s') \neq\varnothing$, branch 
-(\ref{eq:utb})  cannot be taken, and the test execution stops without producing
-$\epass$ or $\efail$. In contrast to the test for failures refinement, this is   interpreted here as a successful test execution, because unexpected blocking
-of the SUT does not violate the trace refinement relation, as long as all traces 
-executed by the SUT are traces of the reference process. In particular, if 
-neither $\epass$ nor $\efail$ is ever produced, so that
-$(Q\parallel[\Sigma] U_T(j)) \hide \Sigma = \Stop$, the test passes, because
-$(\epass\then\Stop) \lessdet_T \Stop$ holds.
+If the SUT process $Q$ deadlocks after a trace $s$ where also the reference
+process $P$ is in a state where deadlock is possible, this is captured by the
+fact that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$. Therefore,
+branch (\ref{eq:utb}) of a test case execution state $U_T(j,k,n)$ with $\#s =
+k \le j$   can be entered and the  test execution terminates with $\epass$.
+If, however, $Q$ blocks after a trace $s'$ where the reference process
+satisfies $\minaccs(P/s') \neq\varnothing$, branch (\ref{eq:utb})  cannot be
+taken, and the test execution stops without producing $\epass$ or $\efail$.
+In contrast to the test for failures refinement, this is   interpreted here
+as a successful test execution, because unexpected blocking of the SUT does
+not violate the trace refinement relation, as long as all traces executed by
+the SUT are traces of the reference process. In particular, if neither
+$\epass$ nor $\efail$ is ever produced, so that $(Q\parallel[\Sigma] U_T(j))
+\hide \Sigma = \Stop$, the test passes, because $(\epass\then\Stop)
+\lessdet_T \Stop$ holds.
 
@@
 %Examples are provided after our discussion of size of the test suites.
-
+%
 As before in the context of Theorem~\ref{th:failurestest},
@@
 
-\fxnote{jp: Wen-ling and I have now provided a full proof for Theorem~\ref{th:tracetest}}
 % ===============================================================================
@@
 $$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$ because of (\ref{eq:ifpaT}). Now $\trc(Q)\subseteq\trc(P)$
-implies $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot enter branch (\ref{eq:uta}) and produce a $\efail$-event when running in parallel with $Q$ and synchronising over $\Sigma$.
-Therefore, only four options are
-available for the test execution    $(Q\parallel[\Sigma] U_T(j))/s$ to continue.
+implies $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$
+cannot enter branch (\ref{eq:uta}) and produce a $\efail$ event when running
+in parallel with $Q$ and synchronising over $\Sigma$. Therefore, only four
+options are available for the test execution    $(Q\parallel[\Sigma]
+U_T(j))/s$ to continue.
 
 \medskip
-\noindent
-{\bf Case~1.} $Q/s$ deadlocks and $\minaccs(G(P)/s) = \{ \varnothing \}$. Then
-$U_T(pq-1,\#s,G(P)/s)$ enters branch (\ref{eq:utb}), and the test execution 
-stops after event $\epass$.
+\noindent {\bf Case~1.} $Q/s$ deadlocks and $\minaccs(G(P)/s) = \{
+\varnothing \}$. In this case, the test $U_T(pq-1,\#s,G(P)/s)$ enters branch
+(\ref{eq:utb}), and its execution stops after $\epass$.
 
 \medskip
-\noindent
-{\bf Case~2.} $Q/s$ deadlocks, but $\minaccs(G(P)/s)\neq\{\varnothing\}$. 
-Then the whole test execution blocks, and this   means that
-neither a $\epass$ nor a $\efail$ event is produced, so the test execution is passed.
+\noindent {\bf Case~2.} $Q/s$ deadlocks, but
+$\minaccs(G(P)/s)\neq\{\varnothing\}$. In this case, the whole test execution
+deadlocks, and this means that neither a $\epass$ nor a $\efail$ event is
+produced, so the test execution is passed.
 
 \medskip
-\noindent
-{\bf Case~3.} $Q/s$ selects an event $e\in[Q/s]^0$ and $\#s < pq-1$. Then the test case 
-$U_T(pq-1)$ in state
-$U_T(pq-1,\#s,G(P)/s)$ can also engage in $e$ by entering branch (\ref{eq:utc}),
-and the test execution continues without having produced a $\epass$ or a $\efail$ event.
+\noindent {\bf Case~3.} $Q/s$ selects an event $e\in[Q/s]^0$ and $\#s <
+pq-1$. In this case, the test $U_T(pq-1)$ in state $U_T(pq-1,\#s,G(P)/s)$ can
+also engage in $e$ by entering branch (\ref{eq:utc}), and its execution
+continues without producing a $\epass$ or a $\efail$ event.
 
 \medskip
-\noindent
-{\bf Case~4.} $\#s = pq-1$ holds. Then $U_T(pq-1,\#s,G(P)/s)$ can only enter
-branch (\ref{eq:utd}), and the test execution stops after $\epass$.
+\noindent {\bf Case~4.} $\#s = pq-1$ holds. In this case,
+$U_T(pq-1,\#s,G(P)/s)$ can only enter branch (\ref{eq:utd}), and the test
+execution stops after $\epass$.
 
@@
 \begin{proof}
-As before in the proofs for
-failures testing, we construct the product graph $G=G(P)\times G(Q)$ and recall
-that every trace $s\in\trc(P)\cap\trc(Q)$ is associated with a path through $G$ labelled
-with the same events as $s$, such that $G/s = (G(P)/s,G(Q)/s)$. Furthermore, we recall from Lemma~\ref{lemma:extendV} that graph state $(G(P)/s,G(Q)/s)$ can always be
-reached by a trace $u$ of length less or equal $pq-1$, where  the order of $G(P)$ is $p$ and that of $G(Q)$ is $q$.
+As before in the proofs for failures testing, we construct the product graph
+$G=G(P)\times G(Q)$ and recall that every trace $s\in\trc(P)\cap\trc(Q)$ is
+associated with a path through $G$ labelled with the same events as $s$, such
+that $G/s = (G(P)/s,G(Q)/s)$. Furthermore, we recall from
+Lemma~\ref{lemma:extendV} that graph state $(G(P)/s,G(Q)/s)$ can always be
+reached by a trace $u$ of length less or equal $pq-1$, where  the order of
+$G(P)$ is $p$ and that of $G(Q)$ is $q$.
 
-Suppose that $P\not\lessdet_T Q$. Then there exists a trace $s\in \trc(Q)\cap\trc(P)$ and an event $e\in [Q/s]^0$ such that $e\not\in [P/s]^0$. Let $u\in \trc(Q)\cap\trc(P)$
-be a trace with $\#u < pq$ and $G/u = (G(P)/s,G(Q)/s)$. Then 
+Suppose that $P\not\lessdet_T Q$. Since the empty trace is a trace of every
+process, there exists a trace $s\in \trc(Q)\cap\trc(P)$ and an event $e\in
+[Q/s]^0$ such that $e\not\in [P/s]^0$. Let $u\in \trc(Q)\cap\trc(P)$ be a
+trace with $\#u < pq$ and $G/u = (G(P)/s,G(Q)/s)$. Then
 $$
@@
 $$
-By assumption, 
-$e\in (\Sigma -[P/s]^0) = (\Sigma - [G(P)/s]^0)$. 
-Since  $G(Q)/u = G(Q)/s$,
- $Q/u$ can engage into $e$. Then $U_T(pq-1,\#u,G(P)/s)$ will enter branch (\ref{eq:uta}), and the test execution stops after having produced $\efail$. This proves that
-  $Q$  fails test $U_T(pq-1)$.
-\xbox
+By assumption, $e\in (\Sigma -[P/s]^0) = (\Sigma - [G(P)/s]^0)$. Since
+$G(Q)/u = G(Q)/s$, $Q/u$ can engage into $e$. Then $U_T(pq-1,\#u,G(P)/s)$
+will enter branch (\ref{eq:uta}), and the test execution stops after having
+produced $\efail$. This proves that $Q$  fails test $U_T(pq-1)$. \xbox
 \end{proof}
+%
+Having established completeness of our test suites, we now consider the
+complexity of a testing technique that uses them.

### 2018-10-07 paper/introduction.tex
update
checked fixmes and new material

@@
 Generation techniques for complete test suites have been developed for
-various formalisms. Here, we tackle \emph{Communicating Sequential
+various formalisms. We tackle \emph{Communicating Sequential
 Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
@@
 
-Our result is not a simple transcription of existing knowledge about finite,
+Our
+\fixme{alcc: I'm slightly uncomfortable with this claim. Can the environment
+really control outputs? Do you not have tau (internal) events in FSM? I'm
+unsure, but you guys know about FSMs much more than I do.}
+\fxnote{jp: I think that this description is ok. With FSMs as used in the testing community, there are no tau events, and parallel composition $M_1\cap M_2$
+is just language intersection. Therefore,
+when composing $M_1$ and $M_2$ to $M_1\cap M_2$,
+$M_1$ cannot refuse to engage into I/O-event $x/y$,
+if $M_2$ has $x/y$ as the only initial event in its language.  }
+result is not a simple transcription of existing knowledge about finite,
 complete test suites for FSMs. The capabilities of CSP to express
@@
 while there is no equivalent to internal choice in nondeterministic FSMs.
-\fixme{alcc: I'm slightly uncomfortable with this claim. Can the environment
-really control outputs? Do you not have tau (internal) events in FSM? I'm
-unsure, but you guys know about FSMs much more than I do.}
-\fxnote{jp: I think that this description is ok. With FSMs as used in the testing community, there are no tau events, and parallel composition $M_1\cap M_2$
-is just language intersection. Therefore, 
-when composing $M_1$ and $M_2$ to $M_1\cap M_2$,
-$M_1$ cannot refuse to engage into I/O-event $x/y$,
-if $M_2$ has $x/y$ as the only initial event in its language.  }
 

### 2018-10-07 paper/preliminaries.tex
checked fixmes and new material

--- 2018-10-01 paper/preliminaries.tex
@@
 When investigating  tests for failures refinement, the notion of
-\emph{acceptances}, which is dual to refusals,
-is also useful. While the original introduction of acceptances presented 
-in~\cite[pp.~75]{Hennessy:1988:ATP:50497} was independent of refusals, 
-we use the definition from~\cite[pp.~278]{Roscoe:1997:TPC:550448}.   A \emph{minimal acceptance} of a CSP process state $P/s$ 
-is the complement of a maximal refusal of the same state. The set of minimal acceptances 
-of $P/s$ is denoted by $\minaccs(P/s)$ and formally defined as 
+\emph{acceptances}, which is dual to refusals, is also useful. While the
+original introduction of acceptances presented
+in~\cite[pp.~75]{Hennessy:1988:ATP:50497} was independent of refusals, we use
+the definition from~\cite[pp.~278]{Roscoe:1997:TPC:550448}.   A \emph{minimal
+acceptance} of a CSP process state $P/s$ is the complement of a maximal
+refusal of the same state. The set of minimal acceptances of $P/s$ is denoted
+by $\minaccs(P/s)$ and formally defined as
 %
@@
 %
-Acceptances have the following intuitive interpretation.  If the behaviour of  $P/s$ is deterministic, its only acceptance
-equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
-this set. If $P/s$ is nondeterministic, it internally chooses one of its
-\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
-{\it possibly} refusing the events from $[P/s]^0 - A$ and
-{\it always} refusing those from $\Sigma - [P/s]^0$.
+Acceptances have the following intuitive interpretation.  If the behaviour of
+$P/s$ is deterministic, its only acceptance equals $[P/s]^0$, because $P/s$
+never refuses any of the events in this set. If $P/s$ is nondeterministic, it
+internally chooses one of its \emph{minimal acceptance sets} $A$ and never
+refuses any event in $A$, while {\it possibly} refusing the events from
+$[P/s]^0 - A$ and {\it always} refusing those from $\Sigma - [P/s]^0$.
 
@@
 \end{equation}
-\fixme{alcc: proof in the report perhaps? What do you need (5) for? }
-\fxnote{jp: I have clarified the introduction of acceptances and removed the statements
-we do not need in the rest of the paper. We need (5) here show that the set of
-maximal refusals of $P/s$ equals $r(G(P/s))$.}
+%
+Given any non-diverging, non-terminating, finite-state process $P$, it can be
+re-con\-struc\-ted from its graph $G(P)$ with initial state $\ii
+n$ and transition function $t$, using $P$'s normalised syntactic
+representation~\cite[pp.~277]{Roscoe:1997:TPC:550448} specified as follows.
+\begin{eqnarray*}
+\text{normalised}(P) & = & P_N(\ii n)
+\\
+P_N(n) & = & \Intchoice_{A\in\minaccs(n)\cup \{ [n]^0 \} } e:A\then P_N(t(n,e))
+\end{eqnarray*}
+With this definition, it is established that $P$ is semantically equivalent
+to the process $\text{normalised}(P)$ in the CSP failures semantics.
 
@@
 Having engaged in $c$, the next process state is represented by node 2. Due
-to normalisation, there is only a single transition satisfying
-$t(1,c) = 2$. This transition, however, can have
-been caused by either $Q$ or $R$ engaging into $c$, so node 2 corresponds to
-process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
-two minimal acceptance sets labelling node 2. Similar considerations explain
-the other nodes and transitions in Fig.~\ref{fig:tga}.
-\xbox
+to normalisation, there is only a single transition satisfying $t(1,c) = 2$.
+This transition, however, can have been caused by either $Q$ or $R$ engaging
+into $c$, so node 2 corresponds to process state $Q/c \intchoice R/c = P
+\intchoice R$. This is reflected by the two minimal acceptance sets labelling
+node 2. From node~2, event $c$ leads to node~3. Since $P$ does not engage
+into $c$, the $R$-component of $P\intchoice R$ must have processed $c$, so
+node~3 corresponds to process state $R/c = R$, and therefore it is labelled
+by $R$'s minimal acceptance $\{b,c\}$. \xbox
 \end{example}
@@
 %
-as established by the results  in the following lemma.
-There, result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
-traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
-refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
-by means of the maximal refusal functions of the graphs involved; and
-(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
-can be derived from the maximal refusal functions by means of
-(\ref{eq:maxrefsminaccs}).
+as established by the results  in the following lemma. There, result
+(\ref{eq:trcrefa}) reflects trace refinement in terms of graph traces; result
+(\ref{eq:failconf}) expresses failures refinement in terms of traces
+refinement and $conf$; result (\ref{eq:failrefa}) states how $conf$ can be
+expressed by means of the maximal refusal functions of the graphs involved;
+and result (\ref{eq:failrefb}) states the same in terms of the minimal
+acceptances that can be derived from the maximal refusal functions by means
+of (\ref{eq:maxrefsminaccs}).
 %
@@
 $\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$. This shows (\ref{eq:trcrefa}).
-\fixme{alcc: (11) is already proved in my paper with mcg. No need to repeat
-it here. I suggest just to add a reference.} \fxnote{jp: done}Formula (\ref{eq:failconf}) has been
-shown in~\cite{DBLP:conf/icfem/CavalcantiG07}.
-
+Result (\ref{eq:failconf}) has been shown
+in~\cite{DBLP:conf/icfem/CavalcantiG07}.
+%
 %\medskip
@@
 %\end{argue}
-
-\medskip
-\noindent
+%
 To prove (\ref{eq:failrefa}), we derive
@@
 \label{lemma:extendV} Let $P$ be a CSP process with normalised transition
-graph $G(P)$. %\fixme{alcc: can a
-%normalised graph for a process have states that are unreachable?}
-%\fxnote{jp: You are right: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
-Let
-$V\subseteq\Sigma^*$ be a finite prefix-closed set of sequences of events.
-Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is, $|t(\ii
-n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended
-by any event of $\Sigma$. Then $G(P)$ reaches at least $(k+1)$ nodes under
-$V\cup V.\Sigma$.
+graph $G(P)$. Let $V\subseteq\Sigma^*$ be a finite prefix-closed set of
+sequences of events. Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$,
+that is, $|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences
+from $V$, extended by any event of $\Sigma$. We can then conclude that $G(P)$
+reaches at least $(k+1)$ nodes under $V\cup V.\Sigma$.
 \end{lemma}
@@
 \subsubsection*{Minimal Hitting Sets of Normalised Transition Graphs}
-Since minimal acceptances can be used to label the nodes of a normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$ according to (\ref{eq:maxrefsminaccs}), the notation of minimal hitting sets also
-carries over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe that
+Since, as previously explained, minimal acceptances can be used to label the
+nodes of a normalised transition graph, and since $\minaccs(P/s) =
+\minaccs(G(P)/s)$ according to (\ref{eq:maxrefsminaccs}), the notation of
+minimal hitting sets also carries over to graphs: we write $\minhits(n)$ for
+nodes $n$ of $G(P)$ and observe that
 \begin{equation}
@@
 \end{equation}
-\fixme{You already used this above. Perhaps you want to move this to before
-Lemma 4?}\fxnote{jp: done}
 
@@
 \begin{proof}
-Throughout the proof, we apply (\ref{eq:maxrefsminaccs}) and (\ref{eq:minhitsGP}), 
-so that $\minaccs(P/s)$ and $\minaccs(G(P)/s)$, as well as 
-$\minhits(P/s)$ and $\minhits(G(P)/s)$ can be used interchangeably.
+We apply (\ref{eq:maxrefsminaccs}) and (\ref{eq:minhitsGP}), so that
+$\minaccs(P/s)$ and $\minaccs(G(P)/s)$, as well as $\minhits(P/s)$ and
+$\minhits(G(P)/s)$ are used interchangeably.
 
-For showing ``$1 \Rightarrow 2$'', assume   $Q\ conf\ P$ and $s\in\trc(P)\cap
-\trc(Q)$. Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}), states that
+For showing ``$1 \Rightarrow 2$'', we assume $Q\ conf\ P$ and
+$s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}),
+states that
 \[
@@
 $\minaccs(G(Q)/s)$, a contradiction to Assumption~2.
-%We note that we use
-%$\minhits(G(P)/s) = \minhits(P/s)$, as well $\minaccs(G(Q)/s) =
-%\minaccs(Q/s)$, which follow from the identification of process states with
-%states of the normalised graph for the proof discussed in the previous
-%section. 
 This completes the proof of the lemma.
-\fixme{Please, check that
-you're happy with this observation.}
-\fxnote{jp: I think the new first sentence of the proof should now be sufficient. Therefore, I removed your addition.}
 \xbox
@@
 maximally exist for a given collection of minimal acceptances. To this end,
-the following definitions and results are useful.\fixme{Shouldn't the footnote go to the acknowledgements?}\fxnote{jp: done}
+the following definitions and results are useful.
 

### 2018-10-09 paper/finitecomplete-failures.tex
removed fixmes following our meeting
Extended proof for Lemma 6

@@
 
-\paragraph{Step~2.} Since $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies
-that for all $s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a
-hitting set for $\minaccs(Q/s)$. From Lemma~\ref{lemma:ufproperties} we know
+\paragraph{Step~2.} For proving that a test execution can never deadlock without a 
+$\epass$ event, it has to be shown that a test execution can neither block
+at branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}).
+
+\paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
+after having run through trace $s$ with $\#s < j$. 
+Since  $P\lessdet_T Q$ by assumption, $s$ is a trace of $P$, thus 
+$U_F(j)/s = U_F(j,\#s,G(P)/s)$ according to Lemma~\ref{lemma:ufproperties}. Therefore,
+$U_F(j)/s$  can enter  branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$.
+Since we assume that $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that
+$[G(P)/s]^0$ is {\it not} a hitting set of $\minhits(Q/s)$, because otherwise at least
+on $e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that $\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be entered, and the test 
+stops after $\epass$. Otherwise, if $\minhits(G(P)/s) \neq \varnothing$, let
+$H\in\minhits(G(P)/s)$. Since $H$ only contains elements that are contained in some minimal acceptance of $P/s$, and all these minimal acceptances are subsets of 
+$[G(P)/s]^0$, $H$ is a subset of $[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, 
+is not an element of $\minhits(Q/s)$, also $H$ is not contained in $\minhits(Q/s)$.
+Now this is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by assumption,
+so $H$ should also be a (not necessarily minimal) hitting set in $\minhits(Q/s)$.
+This proves that the test execution cannot block  at branch (\ref{eq:ufc})
+without being able to pass the test by entering branch (\ref{eq:ufb}).
+
+\paragraph{Step~2.2} Now suppose that the test execution blocks at 
+branch (\ref{eq:ufd}) after having run through some $s\in\trc(Q)\subseteq \trc(P)$
+with $\#s = j$.
+From Lemma~\ref{lemma:ufproperties} we know
 that $U_F(j)/s = U_F(j,k,n)$ with $U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so
@@
 $U_F(j,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
-\varnothing$. The branch (\ref{eq:ufc}) always leads to test continuation
-without a verdict. Since, $s$ is a trace of $P$ and $Q$, we know that there
-is a behaviour of $Q$ that will accept an event $e$ in that branch and allow
-the test to proceed. \fixme{I think we need a stronger argument here, as to
-why it cannot deadlock.} \fxnote{jp: I do not understand the question --
-let's discuss this on the phone. See what I wrote. I'm still worried because
-of nondeterminism. } For the last branch, we note that it can only be entered
-if $\minhits(P/s) = \minhits(n) \neq\varnothing$. In this case, any selected
-minimal hitting set $H\in\minhits(n)$ has a non-empty intersection with each
-of the minimal acceptances of $Q/s$. As a consequence, $Q/s$ never blocks
-when offered events from $H\in\minhits(n)$, and the test terminates with
-event $\epass$. Note that this argument requires that $Q$ is free of
-livelocks, because otherwise the $\epass$-events might not become visible,
-due to unbounded sequences of hidden events performed by $Q$. \xbox
+\varnothing$. 
+If $\minhits(n) \neq \varnothing$,  the assumption that $(Q\parallel[\Sigma] U_F(j))/s$
+blocks at branch (\ref{eq:ufd}) implies that there exists some $H\in\minhits(n)$
+which is not a hitting set of $\minaccs(Q/s)$. Again, we can apply Lemma~\ref{lemma:hseta}
+to conclude that this contradicts the assumption that $Q\ conf\ P$. As a consequence,
+the test execution can never deadlock at branch (\ref{eq:ufd}) without entering
+branch (\ref{eq:ufb}) and passing the test.
+
+Note that the line of reasoning in this proof requires that $Q$ is free of
+livelocks, because otherwise a $\epass$ event might not become visible,
+due to unbounded sequences of hidden events performed by $Q$. 
+\xbox
 \end{proof}

### 2018-10-09 paper/introduction.tex

--- 2018-09-30 paper/introduction.tex
@@
 while there is no equivalent to internal choice in nondeterministic FSMs.
-\fixme{alcc: I'm slightly uncomfortable with this claim. Can the environment
-really control outputs? Do you not have tau (internal) events in FSM? I'm
-unsure, but you guys know about FSMs much more than I do.}
 

### 2018-10-09 paper/finitecomplete-failures.tex
Fixed typos in proof of Lemma 6.

@@
 U_F(j))\hide \Sigma$.
-% alcc: part of the explanation above was in the following lemma, but it is already
-% relevant here.
 
@@
 Since we assume that $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that
-$[G(P)/s]^0$ is {\it not} a hitting set of $\minhits(Q/s)$, because otherwise at least
+$[G(P)/s]^0$ is {\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least
 on $e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that $\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be entered, and the test 
@@
 $[G(P)/s]^0$, $H$ is a subset of $[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, 
-is not an element of $\minhits(Q/s)$, also $H$ is not contained in $\minhits(Q/s)$.
+is not a hitting set of $\minaccs(Q/s)$, also $H$ is not a hitting set of 
+$\minaccs(Q/s)$.
 Now this is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by assumption,
-so $H$ should also be a (not necessarily minimal) hitting set in $\minhits(Q/s)$.
+so $H$ should also be a (not necessarily minimal) hitting set in $\minaccs(Q/s)$.
 This proves that the test execution cannot block  at branch (\ref{eq:ufc})

### 2018-10-10 paper/preliminaries.tex
Added Wen-ling's proof as the new Lemma~4

@@
 \begin{equation}\label{eq:refMaxRef}
-\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
+\refs(P/s) = \{ R'\in 2^\Sigma~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
 \end{equation}
@@
 based on solving a \emph{hitting set problem}. Given a finite collection of
-finite sets $C = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a subset of a
+finite sets ${\cal C} = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a subset of a
 universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set
@@
   \label{eq:hit}
-  \forall A\in C: H\cap A \neq\varnothing.
+  \forall A\in {\cal C}: H\cap A \neq\varnothing.
 \end{equation}
@@
 A \emph{minimal hitting set} is a hitting set that cannot be further reduced
-without losing the characteristic property (\ref{eq:hit}). By $\minhits(C)$
-we denote the collection of minimal hitting sets for a collection $C$. For
-the pathological case where $C$ contains an empty set, $\minhits(C)$ is also
+without losing the characteristic property (\ref{eq:hit}). By $\minhits({\cal C})$
+we denote the collection of minimal hitting sets for a collection ${\cal C}$. For
+the pathological case where $C$ contains an empty set, $\minhits({\cal C})$ is also
 empty.
@@
 that equals the number of minimal hitting sets.
+
+% -----------------------------------------------------------------------------
+\subsubsection*{Minimal Hitting Sets of Minimal Acceptances}
+In this article, we are interested in the minimal hitting sets of minimal acceptances;
+for these, the abbreviated notation
+\[
+\minhits(P/s) = \minhits(\minaccs(P/s))
+\]
+is used.
+
+The minimal hitting sets of minimal acceptances may be alternatively
+characterised by means of the failures of a process. To this end,
+define for any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from
+$\Sigma$
+\begin{equation}\label{eq:minsub}
+\minsub({\cal C}) = \{ A\in {\cal C}~|~\forall B\in {\cal C}:
+B\subseteq A\Rightarrow B = A \}.
+\end{equation}
+This definition has originally been in~\cite{DBLP:conf/icfem/CavalcantiG07}:
+the collection 
+$\minsub({\cal C})$ contains all those sets of the collection ${\cal C}$ that are not true subsets of other members of ${\cal C}$. With this definition, the relation
+between failures and minimal hitting sets of minimal acceptances is established in the following lemma.
+%
+\begin{lemma}
+
+For any trace $s$ of CSP process $P$, define 
+\[
+{\cal A}_s = \{ A\subseteq \Sigma~|~(s,A)\not\in\fails(P)  \}.
+\]
+Then
+\[
+\minhits(P/s) = \minsub {\cal A}_s \ \text{for all traces $s$ of $P$}.
+\]
+\end{lemma}
+\begin{proof}
+We derive
+\begin{argue}
+(s,A)\not\in {\text{failures}}(P)
+\\
+\Leftrightarrow A\not\in {\text{Ref}}(P/s)
+\\
+\Leftrightarrow  \forall R\in {\text{maxRef}}(P/s): A\not\subseteq R
+\\
+\Leftrightarrow   \forall B\in {\text{minAcc}}(P/s): A\not\subseteq (\Sigma\setminus B)& this follows from (\ref{eq:accref})
+\\
+\Leftrightarrow  \forall B\in {\text{minAcc}}(P/s): A\cap B\neq \varnothing
+\\
+\Leftrightarrow  A \text{ is a (not necessarily minimal)}
+              \text{ hitting set of } {\text{minAcc}}(P/s)
+\end{argue}
+%
+Specialising this derivation valid for arbitrary $A$ satisfying $(s,A)\not\in {\text{failures}}(P)$ to
+minimal $A$ with this property proves the statement of the lemma.
+\xbox
+\end{proof}
 
@@
 
-A \emph{Sperner Family} is a collection $S \subseteq 2^\Sigma$ of sets from a
+A \emph{Sperner Family} is a collection ${\cal S} \subseteq 2^\Sigma$ of sets from a
 given finite universe $\Sigma$ that do not contain each other, that is, $H_1
 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair $H_1\neq
-H_2\in S$. Specialising antichains known from partial orders to finite sets
-partially ordered by $\subseteq$ results in Sperner families. We observe that
+H_2\in {\cal S}$. Specialising antichains known from partial orders to finite sets
+partially ordered by $\subseteq$ results in Sperner families.
+
+Given an {\it arbitrary} collection of subsets ${\cal C}\subset 2^\Sigma$,
+the sub-collection $\minsub({\cal C})$ defined in (\ref{eq:minsub}) 
+is a Sperner Family contained
+in ${\cal C}$.
+We further observe that
 \begin{itemize}
@@
 \label{th:sperner}
-Given a  Sperner family $S$
+Given a  Sperner family ${\cal S}$
 over an $n$-element universe $\Sigma$, its cardinality is bound by
 \[
-\card{S} \le \binom{n}{\lfloor\frac{n}{2}\rfloor}.
+\card{{\cal S}} \le \binom{n}{\lfloor\frac{n}{2}\rfloor}.
 \]
@@
 \begin{enumerate}
-\item For even $n$, if $S$ consists of all subsets of $\Sigma$ with
+\item For even $n$, if ${\cal S}$ consists of all subsets of $\Sigma$ with
     cardinality $n/2$;
@@
 \begin{enumerate}
-\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/2$;
+\item ${\cal S}$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/2$;
     or
-\item $S$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/2$.
+\item ${\cal S}$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/2$.
 \end{enumerate}

### 2018-10-10 paper/finitecomplete-failures.tex
Added complete testing assumption and re-phrased Case 2 of Lemma 8.

@@
 
+% -------------------------------------------------------------------------
+\subsubsection*{Complete Testing Assumption}
+As explained above, passing a test case $U_F(j)$ requires that none of 
+the possible executions $(Q\parallel[\Sigma] U_F(j))$ stops after $\efail$ or
+stops without having produced event $\epass$. Therefore, it is necessary to determine
+whether all possible executions have been covered in the repeated runs of 
+$(Q\parallel[\Sigma] U_F(j))$.  
+The theoretical completeness results are therefore based on
+a  \emph{complete testing
+assumption}~\cite{hierons_testing_2004,DBLP:conf/icfem/CavalcantiG07}, 
+ which means that every
+possible behaviour of the SUT is performed after a finite number of test
+executions. In practice, this is realised by executing each test several
+times, recording the traces that have been performed, and using hardware or
+software coverage analysers to determine whether all possible test execution
+behaviours of the SUT have been observed. Therefore, 
+testing nondeterministic SUTs 
+  comes
+at the price of having to apply some grey-box testing techniques enabling us
+to decide whether all SUT behaviours have been observed.
+
 % =========================================================================
@@
 of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$  always performs branch
-(\ref{eq:ufc}) until the trace $u$ has been completely processed. \fixme{If
-the SUT accepts it? See previous fixme.} $U_F(\#u,\#u,G(P)/u)$ may execute
+(\ref{eq:ufc}) until the trace $u$ has been completely processed.  
+ $U_F(\#u,\#u,G(P)/u)$ may execute
 branches (\ref{eq:ufa}) or (\ref{eq:ufd}) only:~assumption
@@
 $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
-always fails. If branch (\ref{eq:ufd}) is executed, the test fails for the
-execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen by
-$U_F(\#u,\#u,G(P)/u)$ that has an empty intersection with the minimal
-acceptance $A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of
+always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and therefore 
+fails for the
+execution where 
+$Q/u$ selects the minimal acceptance $A_Q$ as specified in (\ref{eq:accsnotcontained})
+and $U_F(\#u,\#u,G(P)/u)$ selects 
+a minimal hitting set $H\in\minhits(P/u)$ which has an empty intersection with   $A_Q$.
+The existence of
 such an $H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a
-consequence, there exists a test execution   where $Q/u$ selects acceptance
-$A_Q$ and $U_F(\#u,\#u,G(P)/u)$ selects $H$. This execution deadlocks in
-process state $(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the
-$\epass$-event; this means that the test fails. This concludes the proof.
+consequence,   $(Q\parallel[\Sigma]U_F(\# u))/u$,   cannot produce the
+$\epass$-event in this execution; 
+this means that the test fails. The complete testing assumption guarantees that 
+this execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed sufficiently often.
+This concludes the proof.
 \xbox 

### 2018-10-11 paper/complexity.tex
checked new material

@@
 
-In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test \fixme{In
-\cite{DBLP:conf/icfem/CavalcantiG07}, we work with minimal acceptances. I
-didn't check \cite{Hennessy:1988:ATP:50497}.} \fxnote{jp: we have studied
-\cite{DBLP:conf/icfem/CavalcantiG07} now in more detail -- you are right, of
-course, but there is a confusion of terms: you are also using the minimal
-hitting sets we handle here in our joint paper, but you used the term
-``minimal acceptance'' for them. This is not in line with the definition of
-minimal acceptances as presented in \cite{Hennessy:1988:ATP:50497} and in
-Roscoe's books. Here, in our joint paper, minimal acceptances are introduced
-as defined in Roscoe's books. I have changed the text accordingly.} {\it
-every} non-empty subset of $\Sigma$ whose events cannot be completely refused
-in a given process state of the reference model; this leads to a worst-case
+In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
+non-empty subset of $\Sigma$ whose events cannot be completely refused in a
+given process state of the reference model; this leads to a worst-case
 estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be

### 2018-10-11 paper/finitecomplete-failures.tex
checked new material

@@
 \paragraph{Step~2.} For proving that a test execution can never deadlock without a
-$\epass$ event, it has to be shown that a test execution can neither block
-at branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}).
+$\epass$ event, it has to be shown that a test execution can neither block at
+branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}). These cases are
+considered separately below.
 
 \paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
-after having run through trace $s$ with $\#s < j$. 
-Since  $P\lessdet_T Q$ by assumption, $s$ is a trace of $P$, thus 
-$U_F(j)/s = U_F(j,\#s,G(P)/s)$ according to Lemma~\ref{lemma:ufproperties}. Therefore,
-$U_F(j)/s$  can enter  branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$.
-Since we assume that $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that
-$[G(P)/s]^0$ is {\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least
-on $e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that $\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be entered, and the test 
-stops after $\epass$. Otherwise, if $\minhits(G(P)/s) \neq \varnothing$, let
-$H\in\minhits(G(P)/s)$. Since $H$ only contains elements that are contained in some minimal acceptance of $P/s$, and all these minimal acceptances are subsets of 
-$[G(P)/s]^0$, $H$ is a subset of $[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, 
-is not a hitting set of $\minaccs(Q/s)$, also $H$ is not a hitting set of 
-$\minaccs(Q/s)$.
-Now this is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by assumption,
-so $H$ should also be a (not necessarily minimal) hitting set in $\minaccs(Q/s)$.
-This proves that the test execution cannot block  at branch (\ref{eq:ufc})
-without being able to pass the test by entering branch (\ref{eq:ufb}).
+after having run through trace $s$ with $\#s < j$. Since  $P\lessdet_T Q$ by
+assumption, $s$ is a trace of $P$, thus $U_F(j)/s = U_F(j,\#s,G(P)/s)$
+according to Lemma~\ref{lemma:ufproperties}. Therefore, $U_F(j)/s$  can enter
+branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$. Since we assume that
+$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that $[G(P)/s]^0$ is
+{\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least one
+$e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that
+$\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be
+entered, and the test stops after $\epass$. Otherwise, if $\minhits(G(P)/s)
+\neq \varnothing$, let $H\in\minhits(G(P)/s)$. Since $H$ contains only
+elements that are contained in some minimal acceptance of $P/s$, and all
+these minimal acceptances are subsets of $[G(P)/s]^0$, $H$ is a subset of
+$[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, is not a hitting set of
+$\minaccs(Q/s)$, also $H$ is not a hitting set of $\minaccs(Q/s)$. Now this
+is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by
+assumption, so $H$ should also be a (not necessarily minimal) hitting set in
+$\minaccs(Q/s)$. This proves that the test execution cannot block  at branch
+(\ref{eq:ufc}) without being able to pass the test by entering branch
+(\ref{eq:ufb}).
 
@@
 the violation described by Case~1) or identifying an illegal acceptance $A_Q$
-as specified in Case~2. It is not guaranteed, however, that trace $s$ is
+(as specified in Case~2). It is not guaranteed, however, that trace $s$ is
 short enough to be executed by one of the test cases $U_F(j)$ with $0\le j <
@@
 (\ref{eq:ufc}) until the trace $u$ has been completely processed.
- $U_F(\#u,\#u,G(P)/u)$ may execute
-branches (\ref{eq:ufa}) or (\ref{eq:ufd}) only:~assumption
-(\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has at least one
-non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) we know that
-this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and
-we observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
+ $U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
+only:~(\ref{eq:accsnotcontained}) implies that $P/s$ has at least one
+non-empty minimal acceptance. By (\ref{eq:minhitminaccempty}) this is
+equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and we
+observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
 consequence, branch (\ref{eq:ufb}) cannot be taken because its guard
@@
 $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
-always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and therefore 
-fails for the
-execution where 
-$Q/u$ selects the minimal acceptance $A_Q$ as specified in (\ref{eq:accsnotcontained})
-and $U_F(\#u,\#u,G(P)/u)$ selects 
-a minimal hitting set $H\in\minhits(P/u)$ which has an empty intersection with   $A_Q$.
-The existence of
-such an $H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a
-consequence,   $(Q\parallel[\Sigma]U_F(\# u))/u$,   cannot produce the
-$\epass$-event in this execution; 
-this means that the test fails. The complete testing assumption guarantees that 
-this execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed sufficiently often.
-This concludes the proof.
-\xbox 
+always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and
+therefore fails for the execution where $Q/u$ selects the minimal acceptance
+$A_Q$ as specified in (\ref{eq:accsnotcontained}) and $U_F(\#u,\#u,G(P)/u)$
+selects a minimal hitting set $H\in\minhits(P/u)$ that has an empty
+intersection with $A_Q$. The existence of such an $H$ is guaranteed because
+of Lemma~\ref{lemma:hseta}. As a consequence, $(Q\parallel[\Sigma]U_F(\#
+u))/u$ cannot produce the $\epass$ event in this execution; this means that
+the test fails. The complete testing assumption guarantees that this
+execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed
+sufficiently often. This concludes the proof. \xbox
 \end{proof}

### 2018-10-11 paper/preliminaries.tex
checked new material

@@
 \]
-is used.
-
-The minimal hitting sets of minimal acceptances may be alternatively
-characterised by means of the failures of a process. To this end,
-define for any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from
-$\Sigma$
+is used. The minimal hitting sets of minimal acceptances may be alternatively
+characterised by means of the failures of a process as is done
+in~\cite{DBLP:conf/icfem/CavalcantiG07}. To this end,
+in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for any
+collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
 \begin{equation}\label{eq:minsub}
@@
 \end{equation}
-This definition has originally been in~\cite{DBLP:conf/icfem/CavalcantiG07}:
-the collection 
-$\minsub({\cal C})$ contains all those sets of the collection ${\cal C}$ that are not true subsets of other members of ${\cal C}$. With this definition, the relation
-between failures and minimal hitting sets of minimal acceptances is established in the following lemma.
+The collection $\minsub({\cal C})$ contains all those sets of 
+${\cal C}$ that are not true subsets of other members of ${\cal C}$. With
+this definition, the relation between failures and minimal hitting sets of
+minimal acceptances is established in the following lemma.
 %

### 2018-10-11 paper/complexity.tex
First updates in complexity section

@@
 \subsection{Estimation of $\ell$}
-The first factor $\ell$ has worst-case upper bound $\ell\le \card{\Sigma}^k$.
+The first factor $\ell$ has worst-case upper bound $\ell\le \card{\Sigma}^j$.
 As an example where this upper bound is really met, we consider the reference
@@
 \begin{proof}
-Let $C$ be the collection of all subsets of $\Sigma$ with cardinality
+Let ${\cal C}$ be the collection of all subsets of $\Sigma$ with cardinality
 $n-\lfloor{\frac{n}{2}}\rfloor+1$. Then the CSP process
 \[
-P = \Intchoice_{A\in C} e:A\then P(e)
+P = \Intchoice_{A\in {\cal C}} e:A\then P
 \]
-fulfils $\minaccs(P) = C$. Let $H$ be any minimal hitting set of $C$. Then
+fulfils $\minaccs(P) = {\cal C}$. Let $H$ be any minimal hitting set of $C$. Then
 $H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because
@@
 subset $A\subseteq \Sigma\setminus H$ with cardinality
-$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in $C$, but satisfy
+$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in ${\cal C}$, but satisfy
 $A\cap H=\varnothing$. Since
@@
  that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
- intersects  every element of $C$.  Therefore, every minimal hitting set of $C$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements,
-so $\card{\minhits(C)} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$.
+ intersects  every element of ${\cal C}$.  Therefore, every minimal hitting set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements,
+so $\card{\minhits({\cal C})} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$.
 \xbox
@@
 non-empty subsets of $\Sigma$.
+
+\begin{theorem}\label{th:numexecP}
+Let $U_F(j)$ be a test case derived from the 
+process $P$ specified in the proof of Theorem~\ref{th:upperboundh}. Then
+$U_F(j)$ has
+\[
+\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot \card{\Sigma}^j
+\]
+executions.
+\end{theorem}
+\begin{proof}
+Since the union of all minimal acceptances of $P$ equals $\Sigma$, 
+the initials of $P$ satisfy $[P]^0 = \Sigma$. By construction, $P/e = P$
+for any $e\in\Sigma$, so $\trc(P) = \Sigma^*$ and $[P/s]^0 = \Sigma$
+for all $s\in\Sigma^*$. As a consequence, $U_F(j)$
+may execute $\card{\Sigma}^j$ different traces, each time executing branch (\ref{eq:ufc}),
+before finally entering branch (\ref{eq:ufd}). As shown in Theorem~\ref{th:upperboundh},
+$\card{\minhits(P/s)} =  \binom{n}{\lfloor\frac{n}{2}\rfloor}$, so there are
+$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ possible executions of branch (\ref{eq:ufd})
+after every trace of length $j$. This proves the theorem.
+\xbox
+\end{proof}
 

### 2018-10-11 paper/complexity.tex
Further updates in complexity section [not yet completed]

@@
 In this section, we calculate upper bounds for the total number of test
-executions to be performed when testing for failures refinement.
+executions to be performed when testing for failures refinement 
+and---as a corollary---trace refinement.
 Theorem~\ref{th:failurestest} specifies that all tests $U_F(j),\ 0\le j < pq$
@@
 graph of the reference process $P$, and $q\ge p$ is an estimate for the
-number of nodes in the SUT's transition graph. Therefore, we will first
+maximal number of nodes in the SUT's transition graph. Therefore, we will first
 calculate a bound for the number of test executions to be performed for test
@@
 
-For the worst-case estimate, we assume that $P$ never allows for early
-deadlock (so $\minhits(P/s)$ is never empty) and that the SUT $Q$ is a
-correct failures refinement. Therefore, all test executions
-$(Q\parallel[\Sigma] U_F(j))$ stop after having run through a trace of $Q$ of
-length $j+1$, because their is no early termination due to entering branches
-(\ref{eq:ufa}), (\ref{eq:ufb}), or due to an illegal deadlock of $Q$. As can
-be seen from the specification of the test cases $U_F(j)$ (see
-Section~\ref{sec:finitecompletefails}), the number of executions ending in a
-$\epass$ event corresponds to the number $\ell$ of traces $s$ of $P$ with
-length equal to $j$, multiplied by the number $h$ of minimal hitting sets in
-$\minhits(P/s)$. For the tests $U_T(j)$ verifying trace refinement (see
-Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$,
-since there is no equivalent in $U_T(j)$ to checking different hitting sets
-in the last step of a test execution. \fixme{But doesn't this add to the
-number of executions?} \fxnote{jp: should be clear now from the revised test}
+For the worst-case estimate, we first introduce a CSP reference process $\pmax$
+which 
+turns out to be--given a fixed alphabet $\Sigma$---the test 
+model leading to the maximal number 
+of test executions for every $U_F(j)$ when considering large alphabets. In the case
+of small alphabets and large values of $pq-1$, it will turn out below that a variant
+of $\pmax$ will lead to the maximal number of executions. The conditions for this
+can also be clearly specified.
+
+%we assume that $P$ never allows for early
+%deadlock (so $\minhits(P/s)$ is never empty) and that the SUT $Q$ is a
+%correct failures refinement. Therefore, all test executions
+%$(Q\parallel[\Sigma] U_F(j))$ stop after having run through a trace of $Q$ of
+%length $j+1$, because their is no early termination due to entering branches
+%(\ref{eq:ufa}), (\ref{eq:ufb}), or due to an illegal deadlock of $Q$. As can
+%be seen from the specification of the test cases $U_F(j)$ (see
+%Section~\ref{sec:finitecompletefails}), the number of executions ending in a
+%$\epass$ event corresponds to the number $\ell$ of traces $s$ of $P$ with
+%length equal to $j$, multiplied by the number $h$ of minimal hitting sets in
+%$\minhits(P/s)$. For the tests $U_T(j)$ verifying trace refinement (see
+%Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$,
+%since there is no equivalent in $U_T(j)$ to checking different hitting sets
+%in the last step of a test execution. \fixme{But doesn't this add to the
+%number of executions?} \fxnote{jp: should be clear now from the revised test}
+
+% -------------------------------------------------------------------------
+\subsection{A Reference Process With Maximal Test Complexity}
+
+Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
+a collection of subsets of $\Sigma$ by
+\[
+{\cal C} = \{ A\subseteq\Sigma~|~\card{A} = n - \lfloor\frac{n}{2}\rfloor + 1 \}.
+\]
+%
+With this choice of ${\cal C}$, define
+\begin{equation}\label{eq:pmax}
+\pmax = \Intchoice_{A\in{\cal C}} e:A\then \pmax
+\end{equation}
+%
+The important properties of $\pmax$ are summarised in the following lemma.
+%
+\begin{lemma}\label{lemma:pmax}
+Given alphabet $\Sigma$ with cardinality $\card{\Sigma} = n\ge 2$,
+process $\pmax$ fulfils
+%
+\begin{eqnarray}
+{}[\pmax/s]^0 & = & \Sigma \quad\text{for all $s\in \Sigma^*$}
+\label{eq:pmaxa}
+\\
+\trc(\pmax) & = & \Sigma^*
+\label{eq:pmaxb}
+\\
+\minaccs(P/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
+\label{eq:pmaxc}
+\\
+\minhits(P/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
+\label{eq:pmaxe}
+\\
+\card{\minhits(P/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor} 
+\quad \text{for all $s\in \Sigma^*$}
+\label{eq:pmaxd}
+\end{eqnarray}
+\end{lemma}
+\begin{proof}
+ Since  
+$
+\bigcup_{A\in{\cal C}} A = \Sigma
+$
+by construction of ${\cal C}$, $[\pmax]^0 = \Sigma$. Since $\pmax/e = \pmax$ for all
+$e\in\Sigma$, this proves statements (\ref{eq:pmaxa}) and (\ref{eq:pmaxb}).
+The internal choice construct used in the specification of $\pmax$ implies
+$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all
+$e\in\Sigma$ implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence
+of (\ref{eq:pmaxc}).
+
+Let $H$ be any minimal hitting set of ${\cal C}$. Then
+$H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because
+otherwise $\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, and any
+subset $A\subseteq \Sigma\setminus H$ with cardinality
+$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in ${\cal C}$, but satisfy
+$A\cap H=\varnothing$. Since
+$\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we
+conclude
+ that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
+ intersects  every element of ${\cal C}$.  Therefore, every minimal hitting set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements,
+so $\card{\minhits({\cal C})} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$. This 
+shows  (\ref{eq:pmaxd}) and completes the proof.
+\xbox
+\end{proof}
+
+The test cases $U_F(j)$ generated from $\pmax$ have the structure
+\begin{eqnarray*}
+\label{eq:UFP}
+U_F(j) & = & U_F(j,0,\ii n)
+\\
+U_F(j,k,n) & = &   (k < j) \& \big(e:\Sigma   \then U_F(j,k+1,t(n,e) \big)
+\\ & & \extchoice  
+\\ & & (k = j) \& \big( \Intchoice_{H\in\minhits({\cal C})} (e:H   \then \epass \then\Stop   )  \big)
+\end{eqnarray*}
 
@@
 \]
-fulfils $\minaccs(P) = {\cal C}$. Let $H$ be any minimal hitting set of $C$. Then
+fulfils $\minaccs(P) = {\cal C}$. Let $H$ be any minimal hitting set of ${\cal C}$. Then
 $H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because

### 2018-10-12 paper/complexity.tex
update - still incomplete

@@
 
+% -----------------------------------------------------------------------------
+\subsubsection*{Process Definition and Basic Properties}
 Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
@@
 \label{eq:pmaxd}
+\\
+\card{H} & = & \lfloor\frac{n}{s}\rfloor \quad \text{for all $H\in {\cal C}$}
+\label{eq:pmaxf}
 \end{eqnarray}
@@
  that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
- intersects  every element of ${\cal C}$.  Therefore, every minimal hitting set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements,
-so $\card{\minhits({\cal C})} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$. This 
+ intersects  every element of ${\cal C}$.  Therefore, every minimal hitting set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements; this shows (\ref{eq:pmaxf})
+and $\card{\minhits({\cal C})} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$. The latter 
 shows  (\ref{eq:pmaxd}) and completes the proof.
@@
 
-The test cases $U_F(j)$ generated from $\pmax$ have the structure
+% -----------------------------------------------------------------------------
+\subsubsection*{Test Cases of $\pmax$}
+The test cases $U_F(j)$ generated from $\pmax$ 
+can never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$ for
+all traces $s\in\trc(\pmax)$ according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxa}).
+Moreover, they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$
+is never empty according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxd}).
+Finally, the minimal hitting sets used to probe the SUT at the end of a
+non-blocking  test execution are always the hitting sets of ${\cal C}$ 
+according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxe}).
+This results in a test case structure
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+
+When considering the number of test executions to be performed by 
+$U_F(j)$ against some SUT $Q$ for all $j=0,\dots,pq-1$, the maximal number of 
+test executions is only reached if $Q$ is a correct refinement of the reference
+process or if it fails in the very last execution of the very last $U_F(j)$ 
+executed against $Q$. If an erroneous behaviour of $Q$ is revealed before this
+very last execution, the test suite is stopped and $Q$ is fixed before
+executing the suite again.
+Therefore, we consider only correct SUTs $Q$ when determining the
+maximal number of executions that can be executed.
+
+For the  $U_F(j)$ generated from $\pmax$, the number of test executions
+ to be performed is maximal if $\pmax\lessdet_F Q$ and 
+$\trc(Q) = \Sigma^*$. In such a situation, no test execution blocks early, because
+$Q/s$ can always engage into some $e\in\Sigma$   while $\#s<j$ and will never block
+in the last step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is 
+offered by the test case. The resulting number of executions is
+%
+\begin{equation}
+\label{eq:maxexec}
+\card{\Sigma}^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot \lfloor\frac{n}{2}\rfloor,
+\end{equation}
+%
+because all traces up to length $j$ can be executed with $U_F(j)$ entering branch
+(\ref{eq:ufc}), and each of these traces is followed by one event
+from each of the hitting sets of ${\cal C}$, and each of these hitting sets 
+contains $\lfloor\frac{n}{2}\rfloor$ elements.
+
+The number of executions specified in (\ref{eq:maxexec}) is indeed maximal for
+all reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
+traces $s$. All these processes can never enter branch (\ref{eq:ufb}), and,
+if an execution with an SUT entered branch (\ref{eq:ufa}), this would only lead to early termination of the whole test suite, because a failure has been detected.
+As a consequence, $\card{\Sigma}^{j}$ is the maximal number of traces to be executed 
+up to length $j$, and from Theorem~\ref{th:sperner} we know that the number
+$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets to be tested at the end of each trace of length $j$ is already maximal.
 

### 2018-10-14 paper/complexity.tex
update of complexity considerations - still incomplete

@@
 a collection of subsets of $\Sigma$ by
-\[
+\begin{equation}\label{eq:defC}
 {\cal C} = \{ A\subseteq\Sigma~|~\card{A} = n - \lfloor\frac{n}{2}\rfloor + 1 \}.
-\]
+\end{equation}
 %
@@
 \\
-\card{H} & = & \lfloor\frac{n}{s}\rfloor \quad \text{for all $H\in {\cal C}$}
+\minhits({\cal C})  & = & \{ H\subseteq \Sigma~|~\card{H} = \lfloor\frac{n}{2}\rfloor\}
 \label{eq:pmaxf}
@@
 
+% -----------------------------------------------------------------------------
+\subsubsection*{Maximal Number of Test Executions for $\pmax$}
+
 When considering the number of test executions to be performed by 
@@
 \label{eq:maxexec}
-\card{\Sigma}^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot \lfloor\frac{n}{2}\rfloor,
+n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor},
 \end{equation}
@@
 (\ref{eq:ufc}), and each of these traces is followed by one event
-from each of the hitting sets of ${\cal C}$, and each of these hitting sets 
-contains $\lfloor\frac{n}{2}\rfloor$ elements.
+from each of the hitting sets of ${\cal C}$.
 
@@
 $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets to be tested at the end of each trace of length $j$ is already maximal.
+
+Summing up  formula 
+(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
+%
+\begin{equation}\label{eq:pmaxcomplexity}
+\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  = 
+\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
+\end{equation}
+%
+as the maximal number of test executions to be performed when testing an error-free
+SUT $Q$ against reference process $\pmax$. This maximum is met when $Q$ equals $\Sigma^*$.
+
+% -----------------------------------------------------------------------------
+\subsubsection*{Considering Empty Collections of Minimal Hitting Sets}
+
+The argument so far has shown that reference process $\pmax$ requires the most test 
+executions among all CSP processes $P$ whose collections of minimal hitting sets  
+$\minhits(P/s)$ are never empty for any trace $s$, so that the branch 
+(\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.  
+Lemma~\ref{lemma:failureshittingsets} implies that an empty collection 
+$\minhits(P/s)$
+is equivalent to $(s,A)\in\fails(P)$ for all $A\subseteq\Sigma$. This is equivalent to
+$\maxrefs(P/s) = \{ \Sigma \}$ or $\minaccs(P/s) = \{ \varnothing\}$.
+
+It remains to 
+consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
+may require more test executions for their associated tests $U_F(j)$ than the bound
+given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider the following process $Z$ whose behaviour is quite close to that of $\pmax$ with the only difference
+that after some pre-defined trace $\ol s \in\Sigma^*$ with $\#\ol s < pq$, 
+the process has $\Stop$ as
+an internal choice, so that $Z/s = \big(\Intchoice_{A\in{\cal C}} e:A\then \pmax\big) \intchoice \Stop$, and, consequently, $\minhits(Z/s) = \varnothing$.
+
+In the following specification of $Z$, given traces $s_1,s_2\in\Sigma^*$, the Boolean 
+expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$.
+\begin{eqnarray*}
+Z & = & Z_1(\varepsilon)
+\\
+Z_1(s) & = & (s < \ol s)\&\big( \Intchoice_{A\in{\cal C}} e:A\then Z_1(s.e)  \big)
+\\ & & \extchoice
+\\ & & (s \not< \ol s) \& \pmax
+\\ & & \extchoice
+\\ & & (s = \ol s) \& \big( \Intchoice_{A\in{\cal C}} e:A\then \max  \big)
+\intchoice \Stop
+\end{eqnarray*}
+%
+It is easy to see that $\trc(Z) = \trc(\pmax) =\Sigma^*$ and that for all 
+traces $s\neq\ol s$, $\minhits(Z/s) = \minhits(P/s) = \minhits({\cal C})$ as specified
+in Lemma~\ref{lemma:pmax}.
+
+When comparing the test cases $U_F(j)$ generated from $\pmax$ and the test cases
+$U_F^Z(j)$ created from $Z$, we note that
+\begin{enumerate}
+\item For $j < \#\ol s$, the test cases  $U_F(j)$ and $U_F^Z(j)$ are identical.
+\item For $j = \#\ol s$, branch (\ref{eq:ufb}) can be entered by $U_F^Z(j)/\ol s$, while it is blocked for $U_F(j)/\ol s$, and branch (\ref{eq:ufd}) is blocked
+for  $U_F^Z(j)/\ol s$, while it can be entered by $U_F(j)/\ol s$. As a consequence,
+the maximal number of executions that can be performed by $U_F^Z(j)$ against the SUT 
+is 
+\begin{equation}
+(n^{j} - 1)\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor} + 1
+\end{equation}
+according to (\ref{eq:maxexec}): we win one additional execution by entering 
+branch (\ref{eq:ufb}) after $\ol s$ and lose $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ executions
+by no longer being able to enter branch (\ref{eq:ufd}) after $\ol s$.
+
+\item For $j > \#\ol s$, one execution is won by entering 
+branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
+\end{enumerate}
+%
+Summing up over $j = 0,\dots,pq-1$, this results in a difference of 
+%
+\begin{equation}
+(pq  - \#\ol s) - \binom{n}{\lfloor\frac{n}{2}\rfloor}
+\end{equation}
+executions for the test cases $U_F^Z(j),\ j=0,\dots,pq-1$ in comparison to 
+the test cases $U_F(j),\ j=0,\dots,pq-1$. To assess this formula, see the function
+table for the first 19  arguments of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ shown
+in Table~\ref{table:binom}.
+%
+\begin{table}[htp]
+\caption{Function table of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ for arguments
+$n = 2,\dots,20$.}
+\begin{center}
+\begin{tabular}{c||c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
+$n$ & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13  & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\\hline
+$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ &
+     2 & 3 & 6& 10 & 20 & 35 & 70 & 126 & 252 &
+     462&924&1716&3432&6435&12870&24310&48620&92378&184756
+\end{tabular}
+\end{center}
+\label{table:binom}
+\end{table}
+%
+The analysis above leads to the following complexity result.
+%
+\begin{theorem}
+Given a process alphabet $\Sigma$ with cardinality $
+\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
+the normalised transition graph of $P$ has $p$ states, and the fault domain 
+${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then, if $pq < \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
+test suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ according to Theorem~\ref{th:failurestest}
+requires at most 
+%
+\begin{equation*}
+\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
+\end{equation*}
+%
+test executions.
+\xbox
+\end{theorem}
 

### 2018-10-25 paper/complexity.tex
update

@@
 may require more test executions for their associated tests $U_F(j)$ than the bound
-given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider the following process $Z$ whose behaviour is quite close to that of $\pmax$ with the only difference
-that after some pre-defined trace $\ol s \in\Sigma^*$ with $\#\ol s < pq$, 
-the process has $\Stop$ as
-an internal choice, so that $Z/s = \big(\Intchoice_{A\in{\cal C}} e:A\then \pmax\big) \intchoice \Stop$, and, consequently, $\minhits(Z/s) = \varnothing$.
+given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ and study the potential increase of test executions if
+the behaviour of $Z$ is changed to $(Z/\ol s)\intchoice \Stop$ after some specific 
+trace $\ol s\in\trc(Z)$ with $\minhits(P/\ol s) \neq \varnothing$. This modified process is denoted by $Z({\ol s})$
+an can be specified by\footnote{We use the following prefix operators
+for given   traces $s_1,s_2\in\Sigma^*$: the Boolean 
+expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.}
+\begin{eqnarray}
+Z({\ol s}) & = & Z({\ol s},\varepsilon)
+\\
+\label{eq:zstop}
+Z(\ol s,s) & = & (s = \ol s) \& ((Z/s)\intchoice \Stop)
+\\ & & \Extchoice    \nonumber
+\\  
+\label{eq:zstoppref}
+& & (s < \ol s \wedge \minhits(P/s) = \varnothing) \& 
+\big(  (e:[P/s]^0 \then  Z(\ol s, s.e)) \intchoice \Stop \big)
+\\ & & \Extchoice   \nonumber
+\\ 
+\label{eq:zcont}
+& &           (s < \ol s\wedge \minhits(P/s) \neq \varnothing ) \& 
+\\ & & \nonumber
+\big(\Intchoice_{A\in\minaccs(P/s)\cup \{ [P/s]^0 \}}
+e : A \then Z(\ol s, s.e)          \big)
+\\ & & \Extchoice   \nonumber
+\\
+\label{eq:zlikep}
+& & (s \not\le \ol s) \& (P/s)
+\end{eqnarray}
+%
+If $Z(\ol s,s)$ has executed trace $\ol s$, its branch (\ref{eq:zstop}) is entered, 
+so the process 
+nondeterministically decides whether to continue like $Z/s$ or to stop. As a consequence,
+$Z(\ol s)/\ol s = Z(\ol s,\ol s)$ only has $\varnothing$ as minimal acceptance, and its
+  collection of minimal hitting sets is empty.
+ While $Z(\ol s)$ still executes a true prefix
+of $\ol s$, it behaves exactly as $Z$, executing either branch (\ref{eq:zstoppref}) or
+(\ref{eq:zcont}): the latter branch is taken if $P/s$ has at least one minimal hitting set,
+so that it cannot refuse all elements of $[P/s]^0$. In such a case, $P/s$ can be
+represented in a normalised form, where an initial event is selected from an internal
+choice over all minimal acceptances and over $[P/s]^0$ (the choice $[P/s]^0$ is needed
+because the union over all minimal acceptances is not necessarily identical 
+to $[P/s]^0$)~\cite[pp.~277]{Roscoe:1997:TPC:550448}. 
+The former branch is taken if $P/s$ can block all of its initials.
+As a consequence, in both branches, $Z(\ol s,s)$ 
+has the same nondeterministic behaviour and the same initials as $P/s$, as long as
+a prefix of $\ol s$ is executed. As soon as $Z(\ol s)$ has run through a trace $s$ that is no longer a prefix of $\ol s$, 
+the process continues exactly as $Z$ would do after having run through $s$. This situation
+is captured by branch (\ref{eq:zlikep}).
+Summarising, $Z(\ol s)$ behaves like $P$ for all but the
+one trace   $\ol s$.
+
+% -----------------------------------------------------------------------------
+\subsubsection*{Considering Empty Collections of Minimal Hitting Sets -- Old version}
 
 In the following specification of $Z$, given traces $s_1,s_2\in\Sigma^*$, the Boolean 
-expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$.
+expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.
 \begin{eqnarray*}
@@
 \\ & & \extchoice
-\\ & & (s \not< \ol s) \& \pmax
+\\ & & (s \not\le \ol s) \& \pmax
 \\ & & \extchoice
-\\ & & (s = \ol s) \& \big( \Intchoice_{A\in{\cal C}} e:A\then \max  \big)
-\intchoice \Stop
+\\ & & (s = \ol s) \& \big( (\Intchoice_{A\in{\cal C}} e:A\then \pmax)  
+\intchoice \Stop\big)
 \end{eqnarray*}

### 2018-10-26 paper/complexity.tex
update

@@
 one trace   $\ol s$.
+
+Let us now  compare test executions of $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ for
+various values of $j$, where ${\ol U}_F(j)$ denotes a failures test case derived from 
+reference process $Z(\ol s)$, to executions of $(Z\parallel[\Sigma] U_F(j))$, where
+$U_F(j)$ has been derived from $Z$.
+\begin{enumerate}
+\item For $j<\#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has the same executions 
+as $(Z\parallel[\Sigma] U_F(j))$.
+
+\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has $\card{\minhits(Z/\ol s)}$ fewer executions than $(Z(\ol s)\parallel[\Sigma] U_F(j))$, because 
+$Z(\ol s,\ol s)$ has an empty set of hitting sets. As a consequence, test case
+${\ol U}_F(\#\ol s)$ cannot enter branch  
+\end{enumerate}
 

### 2018-10-28 paper/complexity.tex
update

@@
 as the maximal number of test executions to be performed when testing an error-free
-SUT $Q$ against reference process $\pmax$. This maximum is met when $Q$ equals $\Sigma^*$.
+SUT $Q$ against reference process $\pmax$. This maximum is met when $\trc(Q) = \Sigma^*$.
 
@@
 the behaviour of $Z$ is changed to $(Z/\ol s)\intchoice \Stop$ after some specific 
-trace $\ol s\in\trc(Z)$ with $\minhits(P/\ol s) \neq \varnothing$. This modified process is denoted by $Z({\ol s})$
+trace $\ol s\in\trc(Z)$ with $\minhits(Z/\ol s) \neq \varnothing$. This modified process is denoted by $Z({\ol s})$
 an can be specified by\footnote{We use the following prefix operators
@@
 \label{eq:zstoppref}
-& & (s < \ol s \wedge \minhits(P/s) = \varnothing) \& 
-\big(  (e:[P/s]^0 \then  Z(\ol s, s.e)) \intchoice \Stop \big)
+& & (s < \ol s \wedge \minhits(Z/s) = \varnothing) \& 
+\big(  (e:[Z/s]^0 \then  Z(\ol s, s.e)) \intchoice \Stop \big)
 \\ & & \Extchoice   \nonumber
@@
 \label{eq:zcont}
-& &           (s < \ol s\wedge \minhits(P/s) \neq \varnothing ) \& 
+& &           (s < \ol s\wedge \minhits(Z/s) \neq \varnothing ) \& 
 \\ & & \nonumber
-\big(\Intchoice_{A\in\minaccs(P/s)\cup \{ [P/s]^0 \}}
+\big(\Intchoice_{A\in\minaccs(Z/s)\cup \{ [Z/s]^0 \}}
 e : A \then Z(\ol s, s.e)          \big)
@@
 \label{eq:zlikep}
-& & (s \not\le \ol s) \& (P/s)
+& & (s \not\le \ol s) \& (Z/s)
 \end{eqnarray}
@@
 of $\ol s$, it behaves exactly as $Z$, executing either branch (\ref{eq:zstoppref}) or
-(\ref{eq:zcont}): the latter branch is taken if $P/s$ has at least one minimal hitting set,
-so that it cannot refuse all elements of $[P/s]^0$. In such a case, $P/s$ can be
+(\ref{eq:zcont}): the latter branch is taken if $Z/s$ has at least one minimal hitting set,
+so that it can never refuse all elements of $[Z/s]^0$. In such a case, $Z/s$ can be
 represented in a normalised form, where an initial event is selected from an internal
-choice over all minimal acceptances and over $[P/s]^0$ (the choice $[P/s]^0$ is needed
+choice over all minimal acceptances and over $[Z/s]^0$ (the choice $[Z/s]^0$ is needed
 because the union over all minimal acceptances is not necessarily identical 
-to $[P/s]^0$)~\cite[pp.~277]{Roscoe:1997:TPC:550448}. 
-The former branch is taken if $P/s$ can block all of its initials.
+to $[Z/s]^0$)~\cite[pp.~277]{Roscoe:1997:TPC:550448}. 
+The former branch is taken if $Z/s$ can block all of its initials.
 As a consequence, in both branches, $Z(\ol s,s)$ 
-has the same nondeterministic behaviour and the same initials as $P/s$, as long as
+has the same nondeterministic behaviour and the same initials as $Z/s$, as long as
 a prefix of $\ol s$ is executed. As soon as $Z(\ol s)$ has run through a trace $s$ that is no longer a prefix of $\ol s$, 
@@
 
-\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has $\card{\minhits(Z/\ol s)}$ fewer executions than $(Z(\ol s)\parallel[\Sigma] U_F(j))$, because 
-$Z(\ol s,\ol s)$ has an empty set of hitting sets. As a consequence, test case
-${\ol U}_F(\#\ol s)$ cannot enter branch  
+\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has 
+$(\card{\minhits(Z/\ol s)} - 1)$ fewer executions than $(Z\parallel[\Sigma] U_F(j))$, because 
+$Z(\ol s,\ol s)$ has an empty set of hitting sets, since 
+$Z(\ol s,\ol s) = (Z/\ol s)\intchoice\Stop$. As a consequence, test case
+${\ol U}_F(\#\ol s)$ can only enter branches 
+(\ref{eq:ufa})  or (\ref{eq:zcont}) after having run through trace
+$\ol s$. The former branch is available if and only if it is available in $U_F(\#\ol s)$, the latter just stops the test after issuing the $\epass$ event.
+
+\item For $j > \#\ol s$, one execution is won by entering 
+branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
 \end{enumerate}
+%
+Summing up over all test cases for $j=0,\dots,pq-1$, the difference in test executions generated
+from $Z(\ol s)$ and those generated from  $Z$ is
+\begin{equation}
+ pq - \#\ol s       - \card{\minhits(Z/\ol s)}.
+\end{equation}
+%
+This expression is maximal for $\card{\minhits(Z/\ol s)} = 1$ (recall that
+the definition of $Z(\ol s)$ requires $\card{\minhits(Z/\ol s)} > 0$). 
+The number of hitting sets in $\minhits(Z/\ol s)$ is one if and only if the
+minimal acceptances of $Z/\ol s$ have exactly one element in their common intersection.
+The maximal length of 
+$\ol s$ where the expression remains strictly positive is then $pq - 2$.
+
+The construction of $Z(\ol s)$ can be generalised by using $k$ traces 
+${\ol s}_i, \ i = 1\dots,k$ of the same length $\ol j$, 
+after that the modified process behaves as 
+$(Z/{\ol s}_i)\intchoice \Stop$. Generalising the case distinction above, 
+\begin{enumerate}
+\item the number of test executions remains unchanged for $j < \ol j$,
+\item the number of   executions lost for $j = \ol j$ is at least 
+$k$ (assuming that for all $i$, $\card{\minhits(Z/{\ol s}_i)} = 1$), and
+\item the number of executions won for $j > \ol j$ is $k$.
+\end{enumerate}
+%
+The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ is bounded by the maximal number
+of traces that can be executed by $(Z\parallel[\Sigma] U_F(j))$, which is bounded by
+$\card{\Sigma}^{\ol j}$ (this bound is reached, for example, 
+for process $\pmax$ discussed above). If $k$ equals this maximal number, there is no need
+to execute test case ${\ol U}_F(\ol j)$ any longer, because it can only stop after $\epass$, whenever a trace of length $\ol j$ is executed.
+
+Summing up over all test cases for such a $k \le \card{\Sigma}^{\ol j}$, this results in 
+an upper bound of
+\begin{equation}
+ \card{\Sigma}^{\ol j}(pq - \ol j      - 1)
+\end{equation}
+additional test executions in comparison to the number of executions needed for all tests
+of $Z$. Now we can apply this construction to all traces lengths from $0$ to $pq-2$ (we have 
+explained above that test executions are lost for trace lengths greater than $pq-2$). This results in an upper bound of
+\begin{equation}
+\sum_{j=0}^{pq-2}  n^j(pq - j      - 1) =  
+\frac{n^{pq} + pq - npq -1}{(n - 1)^2}
+\quad\text{with $n = \card{\Sigma}$}
+\end{equation}
+additional test executions that can be gained by applying the construction technique
+explained above for $Z(\ol s)$, extended to all traces of length less or equal to $pq-2$.
+%
+Summarising this result and the one previously obtained in (\ref{eq:maxexec}),
+the following complexity bounds are obtained.
+%
+\begin{theorem}
+Given a process alphabet $\Sigma$ with cardinality $
+\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
+the normalised transition graph of $P$ has $p$ states, and the fault domain 
+${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the number of test executions to be performed   using the complete test 
+suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} has the following bounds.
+%
+\begin{enumerate}
+\item If $pq \le \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
+test suite  
+requires at most 
+%
+\begin{equation*}
+\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
+\end{equation*}
+%
+test executions. This upper bound is reached for process $\pmax$ specified above.
+
+\item If $pq > \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
+test suite  
+requires less than 
+\begin{equation*}
+pq\frac{1-n^{pq}}{1-n} +
+\frac{n^{pq} + pq - npq -1}{(n - 1)^2} \qquad\text{@todo}
+\end{equation*}
+executions.
+\end{enumerate}
+\xbox
+\end{theorem}
 

### 2018-10-31 paper/complexity.tex
New version of section Complexity Considerations completed.

@@
 
-In this section, we calculate upper bounds for the total number of test
-executions to be performed when testing for failures refinement 
-and---as a corollary---trace refinement.
-Theorem~\ref{th:failurestest} specifies that all tests $U_F(j),\ 0\le j < pq$
-need to be executed, where $p$ denotes the number of nodes in the transition
-graph of the reference process $P$, and $q\ge p$ is an estimate for the
-maximal number of nodes in the SUT's transition graph. Therefore, we will first
-calculate a bound for the number of test executions to be performed for test
-$U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.
-
-For the worst-case estimate, we first introduce a CSP reference process $\pmax$
-which 
-turns out to be--given a fixed alphabet $\Sigma$---the test 
-model leading to the maximal number 
-of test executions for every $U_F(j)$ when considering large alphabets. In the case
-of small alphabets and large values of $pq-1$, it will turn out below that a variant
-of $\pmax$ will lead to the maximal number of executions. The conditions for this
-can also be clearly specified.
+%In this section, we calculate estimates for the maximal number of test
+%executions to be performed when testing for failures refinement 
+%and---as a corollary---trace refinement.
+%Theorem~\ref{th:failurestest} specifies that all tests $U_F(j),\ 0\le j < pq$
+%need to be executed, where $p$ denotes the number of nodes in the transition
+%graph of the reference process $P$, and $q\ge p$ is an estimate for the
+%maximal number of nodes in the SUT's transition graph. Therefore, we will first
+%calculate a bound for the number of test executions to be performed for test
+%$U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.
+%
+%For the worst-case estimate, we first introduce a CSP reference process $\pmax$
+%which 
+%turns out to be--given a fixed alphabet $\Sigma$---the test 
+%model leading to the maximal number 
+%of test executions for every $U_F(j)$ when considering large alphabets. In the case
+%of small alphabets and large values of $pq-1$, it will turn out below that a variant
+%of $\pmax$ will lead to the maximal number of executions. The conditions for this
+%can also be clearly specified.
 
@@
 % -------------------------------------------------------------------------
-\subsection{A Reference Process With Maximal Test Complexity}
+\subsection{Estimates for the Maximal Number of Failures Test Executions}
 
 % -----------------------------------------------------------------------------
-\subsubsection*{Process Definition and Basic Properties}
+\subsubsection*{A Reference Process}
 Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
@@
 %
-The important properties of $\pmax$ are summarised in the following lemma.
+The relevant properties of $\pmax$ are summarised in the following lemma.
 %
@@
 all reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
-traces $s$. All these processes can never enter branch (\ref{eq:ufb}), and,
+traces $s$: All these processes can never enter branch (\ref{eq:ufb}), and,
 if an execution with an SUT entered branch (\ref{eq:ufa}), this would only lead to early termination of the whole test suite, because a failure has been detected.
@@
 SUT $Q$ against reference process $\pmax$. This maximum is met when $\trc(Q) = \Sigma^*$.
+If one is only interested in the order of magnitude, the maximal nmber of executions 
+may be re-written as 
+%
+\begin{equation} 
+\label{eq:maxO}
+O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
+\end{equation}
 
@@
 ${\ol U}_F(\#\ol s)$ can only enter branches 
-(\ref{eq:ufa})  or (\ref{eq:zcont}) after having run through trace
+(\ref{eq:ufa})  or (\ref{eq:ufb}) after having run through trace
 $\ol s$. The former branch is available if and only if it is available in $U_F(\#\ol s)$, the latter just stops the test after issuing the $\epass$ event.
@@
 %
-The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ is bounded by the maximal number
-of traces that can be executed by $(Z\parallel[\Sigma] U_F(j))$, which is bounded by
-$\card{\Sigma}^{\ol j}$ (this bound is reached, for example, 
-for process $\pmax$ discussed above). If $k$ equals this maximal number, there is no need
+The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ with fixed length $\#{\ol s}_i=\ol j$ is bounded by the maximal number
+of traces that can be executed by $(Z\parallel[\Sigma] U_F(\ol j))$, which is bounded by
+$\card{\Sigma}^{\ol j}$. If $k$ equals this maximal number, there is no need
 to execute test case ${\ol U}_F(\ol j)$ any longer, because it can only stop after $\epass$, whenever a trace of length $\ol j$ is executed.
 
-Summing up over all test cases for such a $k \le \card{\Sigma}^{\ol j}$, this results in 
+Summing up over all test cases ${\ol U}_F(j),\ j = \ol j,\dots,pq-1$, for a given 
+ $k \le \card{\Sigma}^{\ol j}$, this results in 
 an upper bound of
@@
 additional test executions in comparison to the number of executions needed for all tests
-of $Z$. Now we can apply this construction to all traces lengths from $0$ to $pq-2$ (we have 
-explained above that test executions are lost for trace lengths greater than $pq-2$). This results in an upper bound of
-\begin{equation}
-\sum_{j=0}^{pq-2}  n^j(pq - j      - 1) =  
-\frac{n^{pq} + pq - npq -1}{(n - 1)^2}
-\quad\text{with $n = \card{\Sigma}$}
-\end{equation}
-additional test executions that can be gained by applying the construction technique
-explained above for $Z(\ol s)$, extended to all traces of length less or equal to $pq-2$.
-%
-Summarising this result and the one previously obtained in (\ref{eq:maxexec}),
-the following complexity bounds are obtained.
+of $Z$. The maximal value of this term is reached for $\ol j = pq - 2$: this is the maximal exponent where $(pq - \ol j      - 1)$ is still greater than zero. We conclude that the
+maximal increase of test executions to be gained if processes possess failures $(s,\Sigma)$
+is of the order $\card{\Sigma}^{pq - 2}$. This can be neglected in comparison to 
+the worst case applicable for processes that do not possess any failures $(s,\Sigma)$, because 
+formula (\ref{eq:maxO}) shows that the number of test executions required for the latter   is of order $\card{\Sigma}^{pq - 1}$.
+
+These considerations lead to the following theorem
 %
@@
 the normalised transition graph of $P$ has $p$ states, and the fault domain 
-${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the number of test executions to be performed   using the complete test 
-suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} has the following bounds.
-%
-\begin{enumerate}
-\item If $pq \le \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
-test suite  
-requires at most 
+${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the maximal 
+number of test executions to be performed   using the complete test 
+suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} is of order
 %
 \begin{equation*} 
-\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
+O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
 \end{equation*} 
-%
-test executions. This upper bound is reached for process $\pmax$ specified above.
-
-\item If $pq > \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
-test suite  
-requires less than 
-\begin{equation*}
-pq\frac{1-n^{pq}}{1-n} +
-\frac{n^{pq} + pq - npq -1}{(n - 1)^2} \qquad\text{@todo}
-\end{equation*}
-executions.
-\end{enumerate}
 \xbox
 \end{theorem}
-
-% -----------------------------------------------------------------------------
-\subsubsection*{Considering Empty Collections of Minimal Hitting Sets -- Old version}
-
-In the following specification of $Z$, given traces $s_1,s_2\in\Sigma^*$, the Boolean 
-expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.
-\begin{eqnarray*}
-Z & = & Z_1(\varepsilon)
-\\
-Z_1(s) & = & (s < \ol s)\&\big( \Intchoice_{A\in{\cal C}} e:A\then Z_1(s.e)  \big)
-\\ & & \extchoice
-\\ & & (s \not\le \ol s) \& \pmax
-\\ & & \extchoice
-\\ & & (s = \ol s) \& \big( (\Intchoice_{A\in{\cal C}} e:A\then \pmax)  
-\intchoice \Stop\big)
-\end{eqnarray*}
 %
-It is easy to see that $\trc(Z) = \trc(\pmax) =\Sigma^*$ and that for all 
-traces $s\neq\ol s$, $\minhits(Z/s) = \minhits(P/s) = \minhits({\cal C})$ as specified
-in Lemma~\ref{lemma:pmax}.
-
-When comparing the test cases $U_F(j)$ generated from $\pmax$ and the test cases
-$U_F^Z(j)$ created from $Z$, we note that
-\begin{enumerate}
-\item For $j < \#\ol s$, the test cases  $U_F(j)$ and $U_F^Z(j)$ are identical.
-\item For $j = \#\ol s$, branch (\ref{eq:ufb}) can be entered by $U_F^Z(j)/\ol s$, while it is blocked for $U_F(j)/\ol s$, and branch (\ref{eq:ufd}) is blocked
-for  $U_F^Z(j)/\ol s$, while it can be entered by $U_F(j)/\ol s$. As a consequence,
-the maximal number of executions that can be performed by $U_F^Z(j)$ against the SUT 
-is 
-\begin{equation}
-(n^{j} - 1)\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor} + 1
-\end{equation}
-according to (\ref{eq:maxexec}): we win one additional execution by entering 
-branch (\ref{eq:ufb}) after $\ol s$ and lose $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ executions
-by no longer being able to enter branch (\ref{eq:ufd}) after $\ol s$.
-
-\item For $j > \#\ol s$, one execution is won by entering 
-branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
-\end{enumerate}
-%
-Summing up over $j = 0,\dots,pq-1$, this results in a difference of 
-%
-\begin{equation}
-(pq  - \#\ol s) - \binom{n}{\lfloor\frac{n}{2}\rfloor}
-\end{equation}
-executions for the test cases $U_F^Z(j),\ j=0,\dots,pq-1$ in comparison to 
-the test cases $U_F(j),\ j=0,\dots,pq-1$. To assess this formula, see the function
-table for the first 19  arguments of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ shown
-in Table~\ref{table:binom}.
-%
-\begin{table}[htp]
-\caption{Function table of $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ for arguments
-$n = 2,\dots,20$.}
-\begin{center}
-\begin{tabular}{c||c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c}
-$n$ & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13  & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\\hline
-$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ &
-     2 & 3 & 6& 10 & 20 & 35 & 70 & 126 & 252 &
-     462&924&1716&3432&6435&12870&24310&48620&92378&184756
-\end{tabular}
-\end{center}
-\label{table:binom}
-\end{table}
-%
-The analysis above leads to the following complexity result.
-%
-\begin{theorem}
-Given a process alphabet $\Sigma$ with cardinality $
-\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
-the normalised transition graph of $P$ has $p$ states, and the fault domain 
-${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then, if $pq < \binom{n}{\lfloor\frac{n}{2}\rfloor}$, the complete
-test suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ according to Theorem~\ref{th:failurestest}
-requires at most 
-%
-\begin{equation*}
-\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}
-\end{equation*}
-%
-test executions.
-\xbox
-\end{theorem}
-
-% -------------------------------------------------------------------------
-\subsection{Estimation of $\ell$}
-The first factor $\ell$ has worst-case upper bound $\ell\le \card{\Sigma}^j$.
-As an example where this upper bound is really met, we consider the reference
-process
-\[
-RUN(\Sigma) = e:\Sigma \then RUN(\Sigma).
-\]
-The normalised transition graph of this process has a single state, and its initials
-are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(j)$ can
-never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly
-$\card{\Sigma}^j$ different traces of length $j$ exercising branch (\ref{eq:ufc})
-for each of their events.
-
-% -------------------------------------------------------------------------
-\subsection{Estimation of $h$}
-Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
-acceptances, the cardinality   $h = \card{\minhits(P/s)}$ has the upper bound
-\[
-h \le \binom{n}{\lfloor\frac{n}{2}\rfloor}, \quad \text{where $n = \card{\Sigma}$.}
-\]
-This follows from Theorem~\ref{th:sperner}, since $\minhits(P/s)$ is a
-Sperner Family~(see Section~\ref{sec:sperner}). The next theorem shows that
-this upper bound can really be reached by collections of minimal hitting sets
-associated with a CSP process state.
-
-\begin{theorem}
-\label{th:upperboundh}
-Let  $\Sigma$  be an alphabet with cardinality $n\ge 2$.  Then
-there exists a CSP process $P$ with
-\[
-\card{\text{minHit}(P)} = \binom{n}{\lfloor\frac{n}{2}\rfloor}.
-\]
-\end{theorem}
-\begin{proof}
-Let ${\cal C}$ be the collection of all subsets of $\Sigma$ with cardinality
-$n-\lfloor{\frac{n}{2}}\rfloor+1$. Then the CSP process
-\[
-P = \Intchoice_{A\in {\cal C}} e:A\then P
-\]
-fulfils $\minaccs(P) = {\cal C}$. Let $H$ be any minimal hitting set of ${\cal C}$. Then
-$H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because
-otherwise $\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, and any
-subset $A\subseteq \Sigma\setminus H$ with cardinality
-$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in ${\cal C}$, but satisfy
-$A\cap H=\varnothing$. Since
-$\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we
-conclude
- that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
- intersects  every element of ${\cal C}$.  Therefore, every minimal hitting set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements,
-so $\card{\minhits({\cal C})} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$.
-\xbox
-\end{proof}
-%
-To get an approximation of the maximal size of $\minhits(P/s)$ for large $n$,
-recall Stirling's approximation~\cite[p.~112]{Graham:1994:CMF:562056}
-\[
-m! \approx \sqrt{2\pi m} \cdot \big( \frac{m}{e} \big)^m, \quad\text{where $e$ denotes the Euler Number.}
-\]
-Applying this approximation when $n = \card{\Sigma}$ is even and $m = (n/2)$
-results in
-\[
-h \le \binom{2m}{m} \approx \frac{4^m}{\sqrt{\pi m}} =
-\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}.
-\]
-%
-We note that, due to this approximation result,
-$h$ is   smaller than $2^{\card{\Sigma}}-1$, the cardinality of the
-non-empty subsets of $\Sigma$.
-
-\begin{theorem}\label{th:numexecP}
-Let $U_F(j)$ be a test case derived from the 
-process $P$ specified in the proof of Theorem~\ref{th:upperboundh}. Then
-$U_F(j)$ has
-\[
-\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot \card{\Sigma}^j
-\]
-executions.
-\end{theorem}
-\begin{proof}
-Since the union of all minimal acceptances of $P$ equals $\Sigma$, 
-the initials of $P$ satisfy $[P]^0 = \Sigma$. By construction, $P/e = P$
-for any $e\in\Sigma$, so $\trc(P) = \Sigma^*$ and $[P/s]^0 = \Sigma$
-for all $s\in\Sigma^*$. As a consequence, $U_F(j)$
-may execute $\card{\Sigma}^j$ different traces, each time executing branch (\ref{eq:ufc}),
-before finally entering branch (\ref{eq:ufd}). As shown in Theorem~\ref{th:upperboundh},
-$\card{\minhits(P/s)} =  \binom{n}{\lfloor\frac{n}{2}\rfloor}$, so there are
-$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ possible executions of branch (\ref{eq:ufd})
-after every trace of length $j$. This proves the theorem.
-\xbox
-\end{proof}
-
-% -------------------------------------------------------------------------
-\subsection{Upper Bounds of Test Executions for Checking Failures Refinement}
-
-By Theorem~\ref{th:failurestest}, we need to execute the tests $U_F(j)$ for
-$j = 0,\dots,(pq-1)$; this results in a worst-case bound defined below, where
-we use the formula for the sum of the geometric progression.
-%
-\[
-h\cdot \big( \frac{1-\card{\Sigma}^{pq}}{1-\card{\Sigma}} \big),\
- \text{or, asymptotically,}\  O(\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}\cdot\card{\Sigma}^{(pq-1)}).
- \]
-%
-From Theorem~\ref{th:upperboundh} we know that the worst-case bound for $h$
-above cannot be further reduced, since the full collection of minimal hitting
-sets needs to be checked at the end of each execution of $U_F(j)$.
-
 In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
@@
 estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
-offered to the SUT in the last step of the test execution, so our approach
-reduces the number of test executions in comparison
-to~\cite{Hennessy:1988:ATP:50497} by a factor of $1/{\sqrt{2\pi
-\card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction is visualised by a
-function plot of $2^{\card{\Sigma}}$ versus
-$\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.
+offered to the SUT in the last step of the test execution. This number is significantly larger than the worst-case estimate $\binom{n}{\lfloor n/2\rfloor}$ calculated above for the 
+hitting sets to be checked.
+In Fig.~\ref{fig:minhita}, the reduction is visualised by plots of
+the two functions.
 
@@
 sets\footnote{However, they are denoted by {\it minimal acceptances}
-in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an upper
-bound for the number of test executions to be performed.
+in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an estimate for the number of test executions to be performed.
 
@@
  \begin{center}
-\includegraphics[width=.8\textwidth]{minhit-fig.pdf}
+\includegraphics[width=.8\textwidth]{curvecomparison.pdf}
 \end{center}
-\vspace*{-10mm}
+%\vspace*{-10mm}
 \caption{Function plot $2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.}
@@
 % -------------------------------------------------------------------------
-\subsection{Upper Bounds of Test Executions for Checking Trace Refinement}
+\subsection{Estimates for the Maximal Number of Trace Test Executions}
 
@@
 refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
-$U_F(j)$ above, the number of executions performed by $(Q\parallel[\Sigma]
-U_T(pq-1))$ is bounded by $\card{\Sigma}^{pq-1}$.
+$U_F(j)$ above, the maximal number of executions to be performed by $(Q\parallel[\Sigma]
+U_T(pq-1))$ is of order  $O\big(\card{\Sigma}^{pq-1}\big)$.
 

### 2018-11-02 paper/complexity.tex
Extended calculation for complexity theorem.

@@
 the definition of $Z(\ol s)$ requires $\card{\minhits(Z/\ol s)} > 0$). 
-The number of hitting sets in $\minhits(Z/\ol s)$ is one if and only if the
+The number of hitting sets in $\minhits(Z/\ol s)$ is 1, for example, if   the
 minimal acceptances of $Z/\ol s$ have exactly one element in their common intersection.
@@
 additional test executions in comparison to the number of executions needed for all tests
-of $Z$. The maximal value of this term is reached for $\ol j = pq - 2$: this is the maximal exponent where $(pq - \ol j      - 1)$ is still greater than zero. We conclude that the
-maximal increase of test executions to be gained if processes possess failures $(s,\Sigma)$
-is of the order $\card{\Sigma}^{pq - 2}$. This can be neglected in comparison to 
-the worst case applicable for processes that do not possess any failures $(s,\Sigma)$, because 
-formula (\ref{eq:maxO}) shows that the number of test executions required for the latter   is of order $\card{\Sigma}^{pq - 1}$.
+of $Z$.
 
+So far, the additional executions originating from traces with fixed length $\ol j\in\{0,\dots,pq-2\}$ have been considered. Summing up over all lengths $\ol j$ results in
+an upper bound of 
+%
+\begin{equation}
+\sum_{j=0}^{pq-2} n^j(pq - j - 1) = \frac{n^{pq} - npq + pq - 1}{(n-1)^2}\quad\text{with}\ 
+n=\card{\Sigma}
+\end{equation}
+%
+additional test executions.
+This has the order of magnitude $O(n^{pq-2})$.
+Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).
 These considerations lead to the following theorem
@@
 %\vspace*{-10mm}
-\caption{Function plot $2^{\card{\Sigma}}$ versus $\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.}
+\caption{Function plot $2^{\card{\Sigma}}$ versus $\binom{n}{\lfloor \frac{n}{2}\rfloor}$.}
  \label{fig:minhita}

### 2018-11-04 paper/complexity.tex
Use normalised transition graphs to explain how Z is modified with internal choice STOPs

@@
 \begin{eqnarray*}
-\label{eq:UFP}
+\label{eq:UFPpmax}
 U_F(j) & = & U_F(j,0,\ii n)
@@
 may require more test executions for their associated tests $U_F(j)$ than the bound
-given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ and study the potential increase of test executions if
-the behaviour of $Z$ is changed to $(Z/\ol s)\intchoice \Stop$ after some specific 
-trace $\ol s\in\trc(Z)$ with $\minhits(Z/\ol s) \neq \varnothing$. This modified process is denoted by $Z({\ol s})$
-an can be specified by\footnote{We use the following prefix operators
-for given   traces $s_1,s_2\in\Sigma^*$: the Boolean 
-expression $ s_1 < s_2$ evaluates to $\ist$ if and only if $s_1$ is a true prefix of $s_2$. Expression $s_1\le s_2$ holds if and only if $ s_1 < s_2 \vee s_1 = s_2$ is true.}
-\begin{eqnarray}
-Z({\ol s}) & = & Z({\ol s},\varepsilon)
-\\
-\label{eq:zstop}
-Z(\ol s,s) & = & (s = \ol s) \& ((Z/s)\intchoice \Stop)
-\\ & & \Extchoice    \nonumber
-\\  
-\label{eq:zstoppref}
-& & (s < \ol s \wedge \minhits(Z/s) = \varnothing) \& 
-\big(  (e:[Z/s]^0 \then  Z(\ol s, s.e)) \intchoice \Stop \big)
-\\ & & \Extchoice   \nonumber
-\\ 
-\label{eq:zcont}
-& &           (s < \ol s\wedge \minhits(Z/s) \neq \varnothing ) \& 
-\\ & & \nonumber
-\big(\Intchoice_{A\in\minaccs(Z/s)\cup \{ [Z/s]^0 \}}
-e : A \then Z(\ol s, s.e)          \big)
-\\ & & \Extchoice   \nonumber
-\\
-\label{eq:zlikep}
-& & (s \not\le \ol s) \& (Z/s)
-\end{eqnarray}
-%
-If $Z(\ol s,s)$ has executed trace $\ol s$, its branch (\ref{eq:zstop}) is entered, 
-so the process 
-nondeterministically decides whether to continue like $Z/s$ or to stop. As a consequence,
-$Z(\ol s)/\ol s = Z(\ol s,\ol s)$ only has $\varnothing$ as minimal acceptance, and its
-  collection of minimal hitting sets is empty.
- While $Z(\ol s)$ still executes a true prefix
-of $\ol s$, it behaves exactly as $Z$, executing either branch (\ref{eq:zstoppref}) or
-(\ref{eq:zcont}): the latter branch is taken if $Z/s$ has at least one minimal hitting set,
-so that it can never refuse all elements of $[Z/s]^0$. In such a case, $Z/s$ can be
-represented in a normalised form, where an initial event is selected from an internal
-choice over all minimal acceptances and over $[Z/s]^0$ (the choice $[Z/s]^0$ is needed
-because the union over all minimal acceptances is not necessarily identical 
-to $[Z/s]^0$)~\cite[pp.~277]{Roscoe:1997:TPC:550448}. 
-The former branch is taken if $Z/s$ can block all of its initials.
-As a consequence, in both branches, $Z(\ol s,s)$ 
-has the same nondeterministic behaviour and the same initials as $Z/s$, as long as
-a prefix of $\ol s$ is executed. As soon as $Z(\ol s)$ has run through a trace $s$ that is no longer a prefix of $\ol s$, 
-the process continues exactly as $Z$ would do after having run through $s$. This situation
-is captured by branch (\ref{eq:zlikep}).
-Summarising, $Z(\ol s)$ behaves like $P$ for all but the
-one trace   $\ol s$.
+given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ represented by its transition graph $G(Z)$ with $p$ nodes. 
+Let $v$ be any node of $G(Z)$, such 
+that $\minaccs(v)\neq\{\varnothing \}$. Now modify $G(Z)$ by setting $v$'s acceptance label
+to $\{\varnothing \}$. This produces a modified node $\ol v$ with $[\ol v]^0=[v]^0$, but
+$\minaccs(\ol v)=\{\varnothing \}$. Now it is checked whether $\ol v$ is equivalent to any other node of $G(Z)$. If this is not the case, the modified transition graph $G(\ol Z)$ constructed by
+exchanging $v$ with $\ol v$ and keeping all other nodes and all edges of $G(Z)$  in the new graph without changes has exactly the same number of nodes as $G(Z)$.
 
-Let us now  compare test executions of $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ for
-various values of $j$, where ${\ol U}_F(j)$ denotes a failures test case derived from 
-reference process $Z(\ol s)$, to executions of $(Z\parallel[\Sigma] U_F(j))$, where
-$U_F(j)$ has been derived from $Z$.
+Now consider a test case ${\ol U}_F(j)$ for $j\in\{ 0,\dots,pq-1 \}$ derived from $G(\ol Z)$
+and compare it to $U_F(j)$ derived from $G(Z)$. Let $s$ be a trace of  ${\ol U}_F(j)$ such that
+$\#s\le j$ and   $G(\ol Z)/s = \ol v$. Then we have two cases to distinguish.
 \begin{enumerate}
-\item For $j<\#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has the same executions 
-as $(Z\parallel[\Sigma] U_F(j))$.
+\item If $\#s < j$, then ${\ol U}_F(j)/s$ allows for one additional execution step entering branch 
+(\ref{eq:ufb}) of the test, due to construction of $\ol v$.
 
-\item For $j = \#\ol s$, $(Z(\ol s)\parallel[\Sigma] {\ol U}_F(j))$ has 
-$(\card{\minhits(Z/\ol s)} - 1)$ fewer executions than $(Z\parallel[\Sigma] U_F(j))$, because 
-$Z(\ol s,\ol s)$ has an empty set of hitting sets, since 
-$Z(\ol s,\ol s) = (Z/\ol s)\intchoice\Stop$. As a consequence, test case
-${\ol U}_F(\#\ol s)$ can only enter branches 
-(\ref{eq:ufa})  or (\ref{eq:ufb}) after having run through trace
-$\ol s$. The former branch is available if and only if it is available in $U_F(\#\ol s)$, the latter just stops the test after issuing the $\epass$ event.
+\item If $\#s = j$, then ${\ol U}_F(j)/s$ loses $\card{\minhits(v)}$ execution steps checking 
+that $Z/s$ never blocks one of the hitting sets in $\minhits(v)$ and wins one execution step 
+because again, branch 
+(\ref{eq:ufb}) of the test can now be entered. The difference $1 - \card{\minhits(v)}$ has 0 as
+the largest possible value. This applies, for example, if the intersection of 
+all minimal acceptances of $n$ contains exactly one element.
+\end{enumerate}
 
-\item For $j > \#\ol s$, one execution is won by entering 
-branch (\ref{eq:ufb}) after $\ol s$, no executions are lost.
-\end{enumerate}
-%
-Summing up over all test cases for $j=0,\dots,pq-1$, the difference in test executions generated
-from $Z(\ol s)$ and those generated from  $Z$ is
+As a very rough estimate for the maximal number of additional test executions of ${\ol U}_F(j)/s$
+in comparison to $U_F(j)$ we can therefore assume that $\card{\minhits(v)} = 1$ and all
+$\card{\Sigma}^k$ traces of length $k < j$ lead to the new node $\ol v$. As a result,
+${\ol U}_F(j)/s$ has one additional execution step for every trace of length less than $j$.  
+This results in an upper bound of 
 \begin{equation}
- pq - \#\ol s       - \card{\minhits(Z/\ol s)}.
+\sum_{i=0}^{j-1}n^i = \frac{n^j - 1}{n - 1} \quad\text{with}\ n = \card{\Sigma}
 \end{equation}
-%
-This expression is maximal for $\card{\minhits(Z/\ol s)} = 1$ (recall that
-the definition of $Z(\ol s)$ requires $\card{\minhits(Z/\ol s)} > 0$). 
-The number of hitting sets in $\minhits(Z/\ol s)$ is 1, for example, if   the
-minimal acceptances of $Z/\ol s$ have exactly one element in their common intersection.
-The maximal length of 
-$\ol s$ where the expression remains strictly positive is then $pq - 2$.
+test executions. Summing up over additional test executions stemming from al test cases 
+ ${\ol U}_F(j),\ j = 0,\dots,pq-1$ results in the following upper bound for the number of 
+ additional executions.
+\begin{equation}
+\sum_{j=0}^{pq-1}\frac{n^j - 1}{n - 1} = \frac{n^{pq} - npq + pq - 1}{(n-1)^2}\quad\text{with}\ 
+n=\card{\Sigma}.
+\end{equation}
+This expression has the order of magnitude $O(n^{pq-2})$.
+Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).
 
-The construction of $Z(\ol s)$ can be generalised by using $k$ traces 
-${\ol s}_i, \ i = 1\dots,k$ of the same length $\ol j$, 
-after that the modified process behaves as 
-$(Z/{\ol s}_i)\intchoice \Stop$. Generalising the case distinction above, 
-\begin{enumerate}
-\item the number of test executions remains unchanged for $j < \ol j$,
-\item the number of   executions lost for $j = \ol j$ is at least 
-$k$ (assuming that for all $i$, $\card{\minhits(Z/{\ol s}_i)} = 1$), and
-\item the number of executions won for $j > \ol j$ is $k$.
-\end{enumerate}
-%
-The number $k$ of traces ${\ol s}_i,\ i=1,\dots,k$ with fixed length $\#{\ol s}_i=\ol j$ is bounded by the maximal number
-of traces that can be executed by $(Z\parallel[\Sigma] U_F(\ol j))$, which is bounded by
-$\card{\Sigma}^{\ol j}$. If $k$ equals this maximal number, there is no need
-to execute test case ${\ol U}_F(\ol j)$ any longer, because it can only stop after $\epass$, whenever a trace of length $\ol j$ is executed.
-
-Summing up over all test cases ${\ol U}_F(j),\ j = \ol j,\dots,pq-1$, for a given 
- $k \le \card{\Sigma}^{\ol j}$, this results in 
-an upper bound of
-\begin{equation}
- \card{\Sigma}^{\ol j}(pq - \ol j      - 1)
-\end{equation}
-additional test executions in comparison to the number of executions needed for all tests
-of $Z$.
-
-So far, the additional executions originating from traces with fixed length $\ol j\in\{0,\dots,pq-2\}$ have been considered. Summing up over all lengths $\ol j$ results in
-an upper bound of 
-%
-\begin{equation}
-\sum_{j=0}^{pq-2} n^j(pq - j - 1) = \frac{n^{pq} - npq + pq - 1}{(n-1)^2}\quad\text{with}\ 
-n=\card{\Sigma}
-\end{equation}
-%
-additional test executions.
-This has the order of magnitude $O(n^{pq-2})$.
-Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).
 These considerations lead to the following theorem

### 2018-11-05 paper/complexity.tex
Small updates

@@
 is equivalent to $(s,A)\in\fails(P)$ for all $A\subseteq\Sigma$. This is equivalent to
+$(s,\Sigma)\in\fails(P)$ or
 $\maxrefs(P/s) = \{ \Sigma \}$ or $\minaccs(P/s) = \{ \varnothing\}$.
@@
 It remains to 
-consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
+consider the question whether reference processes $\ol Z$ possessing failures $(s,\Sigma)$
 may require more test executions for their associated tests $U_F(j)$ than the bound
-given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ represented by its transition graph $G(Z)$ with $p$ nodes. 
+given for $\pmax$ in (\ref{eq:pmaxcomplexity}), because process states $\ol Z/s$ with  
+$\minhits(\ol Z/s)=\varnothing$ allow for test executions entering branch (\ref{eq:ufb}).
+To this end, we consider an {\it arbitrary} process  $Z$ represented by its transition graph $G(Z)$ with $p$ nodes. 
 Let $v$ be any node of $G(Z)$, such 
@@
 Now consider a test case ${\ol U}_F(j)$ for $j\in\{ 0,\dots,pq-1 \}$ derived from $G(\ol Z)$
-and compare it to $U_F(j)$ derived from $G(Z)$. Let $s$ be a trace of  ${\ol U}_F(j)$ such that
-$\#s\le j$ and   $G(\ol Z)/s = \ol v$. Then we have two cases to distinguish.
+and compare it to $U_F(j)$ derived from $G(Z)$. Let $s$ be a trace of  $\ol Z$ such that
+$\#s\le j$ and   $G(\ol Z)/s = \ol v$, so that $s$ is also a trace of ${\ol U}_F(j)$. 
+Then we have two cases to distinguish.
 \begin{enumerate}
 \item If $\#s < j$, then ${\ol U}_F(j)/s$ allows for one additional execution step entering branch 
-(\ref{eq:ufb}) of the test, due to construction of $\ol v$.
+(\ref{eq:ufb}) of the test, due to construction of $\ol v$. (The other executable branch after
+$s$ is (\ref{eq:ufc}) which is also executable for $U_F(j)/s$.)
 
@@
 \end{enumerate}
-
+%
+The exchange of node $v$ by $\ol v$ can be repeated several times, as long as $\ol v$ is not equivalent to any other node in the successively updated transition graph  $G(\ol Z)$.
 As a very rough estimate for the maximal number of additional test executions of ${\ol U}_F(j)/s$
-in comparison to $U_F(j)$ we can therefore assume that $\card{\minhits(v)} = 1$ and all
-$\card{\Sigma}^k$ traces of length $k < j$ lead to the new node $\ol v$. As a result,
+in comparison to $U_F(j)$ we can   assume that $\card{\minhits(v)} = 1$ and that {\it all}
+$\card{\Sigma}^k$ traces of length $k < j$ lead to one of the new nodes $\ol v$. As a result,
 ${\ol U}_F(j)/s$ has one additional execution step for every trace of length less than $j$.  
@@
 test executions. Summing up over additional test executions stemming from al test cases 
- ${\ol U}_F(j),\ j = 0,\dots,pq-1$ results in the following upper bound for the number of 
+ ${\ol U}_F(j),\ j = 0,\dots,pq-1$, results in the following upper bound for the number of 
  additional executions.
@@
 Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).
-
 These considerations lead to the following theorem

### 2018-11-08 paper/complexity.tex
New proof for the upper bound of test executions

--- 2018-11-04 paper/complexity.tex
@@
 may require more test executions for their associated tests $U_F(j)$ than the bound
-given for $\pmax$ in (\ref{eq:pmaxcomplexity}). To this end, we consider an {\it arbitrary} process  $Z$ represented by its transition graph $G(Z)$ with $p$ nodes. 
-Let $v$ be any node of $G(Z)$, such 
-that $\minaccs(v)\neq\{\varnothing \}$. Now modify $G(Z)$ by setting $v$'s acceptance label
-to $\{\varnothing \}$. This produces a modified node $\ol v$ with $[\ol v]^0=[v]^0$, but
-$\minaccs(\ol v)=\{\varnothing \}$. Now it is checked whether $\ol v$ is equivalent to any other node of $G(Z)$. If this is not the case, the modified transition graph $G(\ol Z)$ constructed by
-exchanging $v$ with $\ol v$ and keeping all other nodes and all edges of $G(Z)$  in the new graph without changes has exactly the same number of nodes as $G(Z)$.
+given for $\pmax$ in (\ref{eq:pmaxcomplexity}), because process states $Z/s$ with  
+$\minhits(Z/s)=\varnothing$ allow for test executions entering branch (\ref{eq:ufb}).
+To this end, consider a test cases $U_F(j)$ constructed from such a process $Z$. Every trace
+$s\in\trc(Z)$ with $\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$ allows for 
+%
+\begin{itemize}
+\item one execution of branch (\ref{eq:ufb}), where the test execution 
+$(Z\parallel[\Sigma]U_F(j))/s$ stops after $\epass$, and
+\item $\card{[Z/s]^0}$   continuations of the test execution with events $e\in [Z/s]^0$. 
+\end{itemize}
+%
+For every trace $s\in\trc(Z)$ with $\#s=j$,
+%
+\begin{itemize}
+\item one execution of branch (\ref{eq:ufb}) follows if 
+$\minhits(Z/s)=\varnothing$, and otherwise
+\item $\card{\minhits(Z/s)}$ executions checking acceptance of minimal hitting sets. 
+\end{itemize}
+%
+For a rough estimate of the worst case upper bound suppose that
+%
+\begin{enumerate}
+\item $[Z/s]^0 = \Sigma$ for all traces $s$ of $Z$,
+\item all traces $s$ with $\#s<j$ end in a state with empty minimal hitting sets, and
+\item all traces $s$ with $\#s=j$ end in a state with a maximal number $\binom{n}{n/2}$
+ of hitting sets.
+\end{enumerate}
+%
+Note that this situation cannot be realised for all $j\in\{0,\dots,pq-1\}$, because
+the traces of 
+$U_F(p-1)$ already cover all states of $Z$'s transition graph according to Lemma~\ref{lemma:extendV}, and if all states of $Z$ have empty hitting sets, there are no acceptance checks to be performed in the last step of the test execution.
+Therefore, the   upper bound calculated next cannot be reached by a real CSP process.
 
-Now consider a test case ${\ol U}_F(j)$ for $j\in\{ 0,\dots,pq-1 \}$ derived from $G(\ol Z)$
-and compare it to $U_F(j)$ derived from $G(Z)$. Let $s$ be a trace of  ${\ol U}_F(j)$ such that
-$\#s\le j$ and   $G(\ol Z)/s = \ol v$. Then we have two cases to distinguish.
-\begin{enumerate}
-\item If $\#s < j$, then ${\ol U}_F(j)/s$ allows for one additional execution step entering branch 
-(\ref{eq:ufb}) of the test, due to construction of $\ol v$.
+With the 3 assumptions above, we calculate that
+%
+\begin{itemize}
+\item  $U_F(0)$ has  $\binom{n}{n/2}$ executions,
+\item $U_F(j),\ j > 0$ has $\sum_{i=0}^{j-1} n^i = \frac{n^j - 1}{n-1}$ executions of branch (\ref{eq:ufb}), ($n = \card{\Sigma}$), and  
+\item $U_F(j),\ j > 0$ has $\binom{n}{n/2}\cdot n^j$ executions where the acceptance of 
+hitting sets is checked after having run through a trace of length $j$. 
+\end{itemize}
+%
+Summing up over all $U_F(j)$ for $j=0,\dots,pq-1$ this results in an upper bound of
+%
+\begin{equation}
+\frac{(n-1)n(n^{pq} - 1)\binom{n}{\lfloor\frac{n}{2}\rfloor} + n^{pq} - n^2(pq-1) + n(pq-2)}{(n-1)^2n}
+\end{equation}
+%
+test executions. Since this upper bound cannot be reached anyway, 
+we just calculate its order of magnitude, and this results again in $O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}  \cdot n^{pq-1}\big)$, as calculated already for $\pmax$ in (\ref{eq:maxO}).
 
-\item If $\#s = j$, then ${\ol U}_F(j)/s$ loses $\card{\minhits(v)}$ execution steps checking 
-that $Z/s$ never blocks one of the hitting sets in $\minhits(v)$ and wins one execution step 
-because again, branch 
-(\ref{eq:ufb}) of the test can now be entered. The difference $1 - \card{\minhits(v)}$ has 0 as
-the largest possible value. This applies, for example, if the intersection of 
-all minimal acceptances of $n$ contains exactly one element.
-\end{enumerate}
-
-As a very rough estimate for the maximal number of additional test executions of ${\ol U}_F(j)/s$
-in comparison to $U_F(j)$ we can therefore assume that $\card{\minhits(v)} = 1$ and all
-$\card{\Sigma}^k$ traces of length $k < j$ lead to the new node $\ol v$. As a result,
-${\ol U}_F(j)/s$ has one additional execution step for every trace of length less than $j$.  
-This results in an upper bound of 
-\begin{equation}
-\sum_{i=0}^{j-1}n^i = \frac{n^j - 1}{n - 1} \quad\text{with}\ n = \card{\Sigma}
-\end{equation}
-test executions. Summing up over additional test executions stemming from al test cases 
- ${\ol U}_F(j),\ j = 0,\dots,pq-1$ results in the following upper bound for the number of 
- additional executions.
-\begin{equation}
-\sum_{j=0}^{pq-1}\frac{n^j - 1}{n - 1} = \frac{n^{pq} - npq + pq - 1}{(n-1)^2}\quad\text{with}\ 
-n=\card{\Sigma}.
-\end{equation}
-This expression has the order of magnitude $O(n^{pq-2})$.
-Therefore, the additional test executions won by being able to enter branches (\ref{eq:ufb}) in test executions  can be neglected in comparison to the order of magnitude calculated for $\pmax$ in (\ref{eq:maxO}).
-
-These considerations lead to the following theorem
+Summarising these complexity calculations, this results in the following theorem.
 %
 \begin{theorem}
-Given a process alphabet $\Sigma$ with cardinality $
-\card{\Sigma} = n$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
+\label{th:maxexecs}
+Given a process alphabet $\Sigma$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
 the normalised transition graph of $P$ has $p$ states, and the fault domain 
-${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most $q$ states. Then the maximal 
+${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most 
+$q\ge p$ states. Then the maximal 
 number of test executions to be performed   using the complete test 
@@
 \end{equation*} 
+For processes $P$ satisfying $(s,\Sigma)\not\in\fails(P)$ for all traces $s$, the reachable
+precise  upper bound is given by
+%
+\begin{equation*}\label{eq:pmaxcomplexity} 
+\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}.
+\end{equation*}
 \xbox
 \end{theorem}
+%
+
 %

### 2018-11-09 paper/complexity.tex
Corrected some typos

@@
 \item all traces $s$ with $\#s<j$ end in a state with empty minimal hitting sets, and
-\item all traces $s$ with $\#s=j$ end in a state with a maximal number $\binom{n}{n/2}$
+\item all traces $s$ with $\#s=j$ end in a state with a maximal number $\binom{n}{\lfloor n/2\rfloor}$
  of hitting sets.
@@
 \begin{itemize}
-\item  $U_F(0)$ has  $\binom{n}{n/2}$ executions,
+\item  $U_F(0)$ has  $\binom{n}{\lfloor n/2\rfloor}$ executions,
 \item $U_F(j),\ j > 0$ has $\sum_{i=0}^{j-1} n^i = \frac{n^j - 1}{n-1}$ executions of branch (\ref{eq:ufb}), ($n = \card{\Sigma}$), and  
-\item $U_F(j),\ j > 0$ has $\binom{n}{n/2}\cdot n^j$ executions where the acceptance of 
+\item $U_F(j),\ j > 0$ has $\binom{n}{\lfloor n/2\rfloor}\cdot n^j$ executions where the acceptance of 
 hitting sets is checked after having run through a trace of length $j$. 

### 2018-11-10 scp-journal/article/conclusion.tex
Re-formatted for Science of Computing journal (new directory scp-journal). Re-worked discussion.

--- 2018-10-07 paper/conclusion.tex
@@
 \subsubsection*{Discussion of Further Reductions of the Test Effort}
-It is known from complete testing strategies for finite state machines that
-the upper bound $pq$ for the lengths of the traces used in our tests to
-investigate the SUT behaviour can be reduced. It is also known from FSM
-testing that it is not necessary to test {\it all} traces up to this maximal
-length. Notable complete strategies supporting this fact have been presented,
+As has been shown in Theorem~\ref{th:maxtracelen}, the maximal length $pq$ of 
+traces to be tested for either  failures refinement or trace refinement cannot be further reduced.
+It is noteworthy, however, that when testing finite state machines for equivalence, considerably shorter traces can be used.
+From the classical results published in~\cite{chow:wmethod,vasilevskii1973}, for example, it follows that the maximal trace length to be executed is less or equal to $3p -q$, which is considerably smaller than $pq$ for $p,q\ge 3$.
+As a consequence, the investigation of complete test suites establishing failures or trace equivalence is of considerable interest and will be discussed in a future paper.
+
+It is also known from FSM
+testing that it is not necessary to check {\it all} traces up to  
+length $pq$ 
+when testing for reduction of finite state machines (this corresponds to trace refinement). Notable complete strategies supporting this fact have been presented,
 for example,
@@
 the latter. We intend to study translations of several promising FSM
-strategies to CSP in the future. These will effectively reduce the upper
-bound $\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The
-bound $h$ for the number of sets to be used in probing the SUT for illegal
+strategies to CSP in the future. These will effectively reduce the 
+upper bound for the number of test executions to be performed, which  has been 
+shown to be of order $O\big(\card{\Sigma}^{pq-1}\big)$ for our trace refinement tests 
+in Section~\ref{sec:complexity}.  The
+bound $\binom{n}{\lfloor n/2\rfloor}$ for the number of sets to be used in probing the SUT for illegal
 deadlocks, however, cannot be further reduced, as has been established in
@@
 of the execution was blocked, due to legal, but nondeterministic behaviour of the
-SUT.
+SUT. We consider this as a disadvantage, since, when aiming at executing a specific trace $s$
+before being able to check the test objective---for example, the absence of deadlocks when offering a hitting set $H$ of the minimal acceptances---it may take several tries until the full trace $s$ is accepted by the SUT.
+Considering the complete testing assumption described in Section~\ref{sec:finitecompletefails}, it may even take $c^{\#s}$ tries to reach the end of the trace $s$, if the SUT can legally block every $s$-event due to nondeterminism, so that $c$
+trials are required for each event is accepted.
 
-In \fixme{From the proofs of the main theorems, and what you say below, I
-thought you did have verdicts coming from deadlock, and you simply chose not
-to mark it with an inc event.} contrast to that, our test cases specified in
+%\fixme{From the proofs of the main theorems, and what you say below, I
+%thought you did have verdicts coming from deadlock, and you simply chose not
+%to mark it with an inc event.}
+In  contrast to that, our test cases specified in
 Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
 adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
-U_F(j))$ for failures refinement never block before the final step specified
-by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
-should be noted, however, that it is necessary for our test verdicts to
-recognise also deadlocks in the final test step and interpret them as FAIL,
-as described in Section~\ref{sec:finitecompletefails}. In practice, this is
-realised by adding a timeout event to the testing environment which indicates
-deadlock situations. For real-time systems, this is an accepted technique,
-because the SUT has to respond within a pre-defined latency interval,
-otherwise its behaviour is considered to be blocked and regarded as a
-failure. The tests executions $(Q\parallel[\Sigma] U_T(j))$ \fixme{Missing
-hiding.} \fxnote{jp: Here, the hiding operator is not necessary: the possible
-test executions are really the traces of $Q$ parallel $U_T(j)$. The hiding is
-only needed to specify the verdict.}for trace refinement never block at all
-before stopping after the verdict $\epass$ or $\efail$, unless the SUT
-process $Q$ has an unexpected deadlock state. Recall from the explanations
-given in Section~\ref{sec:finitecomplete} that the blocking situation also
-leads to passing the test, because $(\epass\then \Stop)\lessdet_T
-(Q\parallel[\Sigma] U_T(j))\hide\Sigma$ still holds if $(Q\parallel[\Sigma]
-U_T(j))\hide\Sigma$ only produces the empty trace.
+U_F(j))$ for failures refinement 
+may only stop early with $\epass$ 
+after traces $s$ satisfying $\minhits(P/s) = \varnothing$, and may deadlock 
+(a) after a trace $s$ where the SUT illegally deadlocks (so $\minhits(P/s) \neq \varnothing$ for
+the reference process, but $\minhits(Q/s) = \varnothing$ for the implementation $Q$), or
+(b) 
+in the final 
+step when---just as in the corresponding test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}--- hitting sets $H$ are offered to the SUT and it refuses their acceptance.
+In both situations (a) and (b) the test fails. As a consequence, far less test repetitions are required according to the complete testing assumption than for the successful execution of all 
+test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}.
 
-The adaptive behaviour of both our test case types $U_F(j)$ and $U_T(j)$,
-however, induces the obligation to check that {\it all} possible executions
-have been performed before the test can be considered as passed. Typically,
-it is therefore assumed that a \emph{complete testing
-assumption}~\cite{hierons_testing_2004} holds, which means that every
-possible behaviour of the SUT is performed after a finite number of test
-executions. In practice, this is realised by executing each test several
-times, recording the traces that have been performed, and using hardware or
-software coverage analysers to determine whether all possible test execution
-behaviours of the SUT have been observed. Therefore, adaptive test cases come
-at the price of having to apply some grey-box testing techniques enabling us
-to decide whether all SUT behaviours have been observed.
+Another distinction of our failures test cases $U_F(j)$ to the tests specified in~\cite{DBLP:conf/icfem/CavalcantiG07} consists in the fact that the former test both trace refinement and
+the $conf$ conformance relation in one go, whereas the latter use separate test suites to 
+establish these two correctness conditions. Again, we consider the  
+structure of the  test cases 
+$U_F(j)$ as advantageous, since, when checking acceptance of a hitting set $H$ 
+after a trace $s$ for $c$ times according to the complete testing assumption,  
+any acceptance of an illegal event $e\in\Sigma - [P/s]^0$ should also be revealed within these $c$ tries.
+
+%
+%never block before the final step specified
+%by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
+%should be noted, however, that it is necessary for our test verdicts to
+%recognise also deadlocks in the final test step and interpret them as FAIL,
+%as described in Section~\ref{sec:finitecompletefails}. In practice, this is
+%realised by adding a timeout event to the testing environment which indicates
+%deadlock situations. For real-time systems, this is an accepted technique,
+%because the SUT has to respond within a pre-defined latency interval,
+%otherwise its behaviour is considered to be blocked and regarded as a
+%failure. The tests executions $(Q\parallel[\Sigma] U_T(j))$ \fixme{Missing
+%hiding.} \fxnote{jp: Here, the hiding operator is not necessary: the possible
+%test executions are really the traces of $Q$ parallel $U_T(j)$. The hiding is
+%only needed to specify the verdict.}for trace refinement never block at all
+%before stopping after the verdict $\epass$ or $\efail$, unless the SUT
+%process $Q$ has an unexpected deadlock state. Recall from the explanations
+%given in Section~\ref{sec:finitecomplete} that the blocking situation also
+%leads to passing the test, because $(\epass\then \Stop)\lessdet_T
+%(Q\parallel[\Sigma] U_T(j))\hide\Sigma$ still holds if $(Q\parallel[\Sigma]
+%U_T(j))\hide\Sigma$ only produces the empty trace.
+
+%The adaptive behaviour of both our test case types $U_F(j)$ and $U_T(j)$,
+%however, induces the obligation to check that {\it all} possible executions
+%have been performed before the test can be considered as passed. Typically,
+%it is therefore assumed that a \emph{complete testing
+%assumption}~\cite{hierons_testing_2004} holds, which means that every
+%possible behaviour of the SUT is performed after a finite number of test
+%executions. In practice, this is realised by executing each test several
+%times, recording the traces that have been performed, and using hardware or
+%software coverage analysers to determine whether all possible test execution
+%behaviours of the SUT have been observed. Therefore, adaptive test cases come
+%at the price of having to apply some grey-box testing techniques enabling us
+%to decide whether all SUT behaviours have been observed.
 
@@
 As explained in the previous sections, passing a test is specified by
-$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ for
+$$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$$ for
 failures testing. If the SUT $Q$ is not a programmed piece of software or an
@@
 
-For large and complex CSP processes $Q$, it may be too time consuming to
+For a large and complex CSP implementation process $Q$, it may be too time consuming to
 generate its normalised transition graph, so that its number $q$ of nodes is

### 2018-11-10 scp-journal/article/main.tex
Re-formatted for Science of Computing journal (new directory scp-journal). Re-worked discussion.

\documentclass[1p]{elsarticle}

%
%
%\usepackage{lineno,hyperref}
%\modulolinenumbers[5]

\journal{Science of Computer Programming}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
%
\usepackage{url}
\urldef{\mailsa}\path|{peleska,huang}@uni-bremen.de|
\urldef{\mailsb}\path|ana.cavalcanti@york.ac.uk|
\urldef{\mailsc}\path|adenilso@icmc.usp.br|

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{hyperref}

%\usepackage{float}
%\usepackage{lscape}
%\usepackage{lscape}
%
%\usepackage[linesnumbered,lined,boxed]{algorithm2e}
\usepackage[draft]{fixme}
%
\usepackage{zed-csp}
%
%\DeclareMathSymbol{\B}{\mathalpha}{AMSb}{"42}
%\DeclareMathSymbol{\I}{\mathalpha}{AMSb}{"49}
%\DeclareMathSymbol{\N}{\mathalpha}{AMSb}{"4E}
%\DeclareMathSymbol{\Pwr}{\mathalpha}{AMSb}{"50}
%\DeclareMathSymbol{\Q}{\mathalpha}{AMSb}{"51}
%\DeclareMathSymbol{\R}{\mathalpha}{AMSb}{"52}
%\DeclareMathSymbol{\Z}{\mathalpha}{AMSb}{"5A}
%\DeclareMathSymbol{\Sol}{\mathalpha}{AMSb}{"53}
%
%\newcommand{\obsv}[1]{{\cal O}(#1)}
%\newcommand{\ctrl}[1]{{\cal C}(#1)}
%\newcommand{\vdot}[1]{\stackrel{.}{#1}}
%
%\newcommand{\HS}{{\mathcal H}}
%
%\newcommand{\Interval}{\I}
%
\newcommand{\ist}{\mbox{{\tt true}}}
\newcommand{\isf}{\mbox{{\tt false}}}
%\newcommand{\emptytrace}{\langle~\rangle}
%\newcommand{\abegin}{\mathbf{begin}}
%\newcommand{\aend}{\mathbf{end}}
%\newcommand{\alet}{\mathbf{let}}
%\newcommand{\aendlet}{\mathbf{endlet}}
%\newcommand{\ain}{\mathbf{in}}
%\newcommand{\afor}{\mathbf{for}}
%\newcommand{\adownto}{\mathbf{downto}}
%\newcommand{\aforall}{\mathbf{foreach}}
%\newcommand{\awhile}{\mathbf{while}}
%\newcommand{\ado}{\mathbf{do}}
%\newcommand{\aod}{\mathbf{od}}
%\newcommand{\aenddo}{\mathbf{enddo}}
%\newcommand{\aif}{\mathbf{if}}
%\newcommand{\afi}{\mathbf{fi}}
%\newcommand{\athen}{\mathbf{then}}
%\newcommand{\aelse}{\mathbf{else}}
%\newcommand{\aelseif}{\mathbf{elseif}}
%\newcommand{\aendif}{\mathbf{endif}}
%\newcommand{\ainout}{\mathbf{inout}}
%\newcommand{\awhere}{\mathbf{where}}
%\newcommand{\areturn}{\mathbf{return}}
%\newcommand{\aout}{\mathbf{out}}
%\newcommand{\aprocedure}{\mathbf{procedure}}
%\newcommand{\afunction}{\mathbf{function}}
%\newcommand{\abreak}{\mathbf{break}}
%\newcommand{\Sup}[1]{\overline{#1}}
%\newcommand{\Inf}[1]{\underline{#1}}
%
\newcommand{\taba}{\hspace*{3mm}}
%\newcommand{\tabb}{\hspace*{6mm}}
\newcommand{\tabc}{\hspace*{9mm}}
\newcommand{\tabd}{\hspace*{12mm}}
%\newcommand{\tabe}{\hspace*{15mm}}
%\newcommand{\tabf}{\hspace*{18mm}}
%
%\newcommand{\gca}[1]{{{#1}^{\vartriangleright}}}
%\newcommand{\gcb}[1]{{{#1}^{\vartriangleleft}}}
%\newcommand{\gclr}{{{\vartriangleleft\atop\longleftarrow}\atop
%                   {\longrightarrow\atop\vartriangleright}}}
%
%\newcommand{\Nat}{{\mathbb N}}
%\newcommand{\Real}{{\mathbb R}}
%
%\newcommand{\trans}{\longrightarrow}
%\newcommand{\transp}{\longrightarrow_{\power}}
%\newcommand{\transl}{\longrightarrow_{L}}
%\newcommand{\transg}{\longrightarrow_{G}}
%\newcommand{\transcfg}[1]{\stackrel{#1}{\longrightarrow}_{CFG}}
%\newcommand{\isdefd}{=_{\mbox{\footnotesize def}}}
%\newcommand{\equivdef}{\equiv_{\mbox{\footnotesize def}}}
%\newcommand{\mitem}{\mbox{\em M-Item}}
%\newcommand{\currt}{\hat{t}}
%\newcommand{\sigmaa}{\sigma_A}
%\newcommand{\strictimplies}{\stackrel{\bullet}{\Rightarrow}}
%\newcommand{\trl}{/\!/}
%\newcommand{\q}{\textbf{q}}
%\newcommand{\eqc}[2]{[#1;#2]}
%\newcommand{\vest}{V_{\text{\sl est}}}
%\newcommand{\vmax}{V_{\text{\sl MRSP}}}
%\newcommand{\wout}{\mathsf{W}}
%\newcommand{\eout}{\mathsf{EB}}
%\newcommand{\areb}{\mathsf{allowRevokeEB}}
%\newcommand{\sbia}{\mathsf{SBAvailable}}
%\newcommand{\sbz}{\mathsf{sb}_0}
%\newcommand{\ticmd}{\mathsf{TICmd}}
%\newcommand{\dmicmd}{\mathsf{DMICmd}}
%
%\newcommand{\sob}{\mathsf{speedOnBoard}}
%\newcommand{\std}{\mathsf{speedToDriver}}
%\newcommand{\pstd}{\mathsf{permittedSpeedToDriver}}
%\newcommand{\csmsw}{\mathsf{csmSwitch}}
%\newcommand{\sbicmd}{\mathsf{sbiCmd}}
%\newcommand{\sbidisplay}{\mathsf{DMIdisplaySBI}}
%
%
%\newcommand{\dvw}{\mathsf{dV}_{\mathsf{warning}}}
%\newcommand{\dvs}{\mathsf{dV}_{\mathsf{sbi}}}
%\newcommand{\dve}{\mathsf{dV}_{\mathsf{ebi}}}
%
%
%\newcommand{\calcw}{\mathsf{dV\_warning(float)}}
%\newcommand{\calcs}{\mathsf{dV\_sbi(float)}}
%\newcommand{\calce}{\mathsf{dV\_ebi(float)}}
%\newcommand{\calcstd}{\mathsf{calc\_speed\_to\_driver()}}
%\newcommand{\calcpstd}{\mathsf{calc\_permitted\_speed\_to\_driver()}}
%\newcommand{\calcsob}{\mathsf{calc\_speed\_onboard()}}
%
\newcommand{\trc}{\text{\it traces}}
\newcommand{\fails}{\text{\it failures}}
%\newcommand{\qtrc}{\text{qtraces}}
%\newcommand{\iotrc}{\text{L}}
%
%
%%\newtheorem{definition}{Definition}
%%\newtheorem{property}{Property}
%%\newtheorem{lemma}{Lemma}
%%\newtheorem{theorem}{Theorem}
%%\newtheorem{corollary}{Corollary}
%%\newtheorem{property}{Property}
%
%%\newtheorem{note}{Note}[section]
%
\newcommand{\xbox}{\unskip\nobreak\hfil\penalty50
      \hskip2em\hbox{}\nobreak\hfil$\Box$%
      \parfillskip=0pt\finalhyphendemerits=0 \par%

      \medskip
      \parfillskip=0pt plus 1fil
}
%
%\newcommand{\dontshow}[1]{}
%\newcommand{\annot}[1]{\textbf{\textcolor{red}{#1}}}
%\newcommand{\modelsf}{\models_{\text{F}}}
%
%\newcommand{\ta}{\mathbf{A}}
%\newcommand{\te}{\mathbf{E}}
%\newcommand{\tx}{\mathbf{X}}
%\newcommand{\tf}{\mathbf{F}}
%\newcommand{\tg}{\mathbf{G}}
%\newcommand{\tu}{\mathbf{U}}
%\newcommand{\tr}{\mathbf{R}}
%\newcommand{\tw}{\mathbf{W}}
%
%\newcommand{\ttu}[1]{\mathbf{U}^{#1}}
%\newcommand{\ttg}[1]{\mathbf{G}^{#1}}
%\newcommand{\ttf}[1]{\mathbf{F}^{#1}}
%
\newcommand{\epass}{{\text{\it pass}}}
\newcommand{\efail}{{\text{\it fail}}}
%
\newcommand{\pass}{\underline{\text{pass}}}
% \newcommand{\tfail}{\underline{\text{fail}}}
%
%\newcommand{\xpass}[1]{{#1}_\text{pass}}
%\newcommand{\xfail}[1]{{#1}_\text{fail}}
%
%\newcommand{\tc}{{\text{TC}}}
\newcommand{\TS}{{\text{TS}}}
\newcommand{\ii}[1]{\underline{#1}}
%\newcommand{\inc}{\,{\rm I}\,}
%
%\newcommand{\scov}{\text{SCOV}}
%\newcommand{\after}{\text{-after-}}
%\newcommand{\reduction}{\preceq}
%
%\newcommand{\ps}{\le_s}
%\newcommand{\pss}[1]{\stackrel{#1}{\ps}}
%
%\newcommand{\pe}{\sim_s}
%\newcommand{\pes}[1]{\stackrel{#1}{\pe}}
\newcommand{\ol}{\overline}
%\newcommand{\diag}{\mathbf{diag}}
%\newcommand{\primem}{\mathbf{prime}}
%
%\newcommand{\ltl}{\text{\tt LTL}}
%
\newcommand{\accs}{\text{\it Acc}}
\newcommand{\minaccs}{\text{\it minAcc}}
%
\newcommand{\refs}{\text{\it Ref}}
\newcommand{\maxrefs}{\text{\it maxRef}}
\newcommand{\minhits}{\text{\it minHit}}
\newcommand{\minsub}{{\it min}_\subseteq}

%\newcommand{\failure}{{\text {\rm failures}}}

\newcommand{\prefs}{\text{\bf pref}}

\newcommand{\card}[1]{{~|\! #1\!|~}}

\newcommand{\pmax}{P_\text{max}}

\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\newcounter{examplectr}
\newenvironment{example}[1][]
{\refstepcounter{examplectr}

\medskip
\noindent
{\bf Example~\theexamplectr.}}     %%%\label{#1}
{
%\unskip\nobreak\hfil\penalty50
%      \hskip2em\hbox{}\nobreak\hfil$\Box$%
%      \parfillskip=0pt \finalhyphendemerits=0 \par
}

\newenvironment{proof}[1][]
{

\medskip
\noindent
{\bf Proof.\ }}      
{
%\unskip\nobreak\hfil\penalty50
%      \hskip2em\hbox{}\nobreak\hfil$\Box$%
%      \parfillskip=0pt \finalhyphendemerits=0 \par
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% To avoid formulas in $ $ to be broken across lines
\relpenalty=10000
\binoppenalty=10000

% ===========================================================================

\begin{document}

\begin{frontmatter}

\title{Finite Complete Suites for CSP Refinement Testing}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Jan Peleska, Wen-ling Huang \fnref{jpwlh}}
\address{Bremen, Germany}
\fntext[jpwlh]{University of Bremen, \mailsa}

\author{Ana Cavalcanti \fnref{ana}}
\address{York, United Kingdom}
\fntext[ana]{University of York, \mailsb}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}
%
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}
%
%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
\input{abstract}
\end{abstract}

\begin{keyword}
Model-based testing, CSP\sep Trace Refinement\sep Failures Refinement\sep Complete Test Suites\sep
\end{keyword}
\end{frontmatter}

\input{introduction}

\input{preliminaries}

\input{finitecomplete-failures}

\input{casestudies}

\input{finitecomplete-trace}

\input{complexity}

%%%\input{related}

\input{conclusion}

% ===============================================================================
\section*{References}

\bibliography{references,jp}

\end{document}

### 2018-11-11 scp-journal/article/complexity.tex
Re-formatted for Science of Computing journal (new directory scp-journal). Re-worked discussion.
Expanded calculation to make it more understandable

--- 2018-11-09 paper/complexity.tex
@@
 $\minhits(Z/s)=\varnothing$ allow for test executions entering branch (\ref{eq:ufb}).
-To this end, consider a test cases $U_F(j)$ constructed from such a process $Z$. Every trace
+To this end, consider a test case $U_F(j)$ constructed from such a process $Z$. Every trace
 $s\in\trc(Z)$ with $\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$ allows for 
@@
 Therefore, the   upper bound calculated next cannot be reached by a real CSP process.
-
 With the 3 assumptions above, we calculate that
@@
 %
-Summing up over all $U_F(j)$ for $j=0,\dots,pq-1$ this results in an upper bound of
+Summing up over all $U_F(j)$ for $j=0,\dots,pq-1$,  an upper bound $B$ for the number of test executions may be calculated as follows.
 %
-\begin{equation}
-\frac{(n-1)n(n^{pq} - 1)\binom{n}{\lfloor\frac{n}{2}\rfloor} + n^{pq} - n^2(pq-1) + n(pq-2)}{(n-1)^2n}
-\end{equation}
+\begin{eqnarray*}
+B & = & \binom{n}{\lfloor n/2\rfloor} + \sum_{j=1}^{pq-1} \frac{n^j - 1}{n-1} + 
+\sum_{j=1}^{pq-1}\binom{n}{\lfloor n/2\rfloor}\cdot n^j
+   \\
+& = & \binom{n}{\lfloor n/2\rfloor}\cdot\big(1 + \frac{n^{pq}-n}{n-1} \big)  + 
+\sum_{j=1}^{pq-1} \frac{n^j - 1}{n-1}
+   \\
+& = &\binom{n}{\lfloor n/2\rfloor}\cdot\big(\frac{n^{pq}-1}{n-1} \big)  +
+\frac{n^{pq}-n pq+pq-1}{(n-1)^2}
+\\
+& = & \frac{\binom{n}{\left\lfloor \frac{n}{2}\right\rfloor}(n-1) \left(n^{pq}-1\right) +n^{pq}-npq+pq-1}{(n-1)^2}
+\end{eqnarray*}
 %
-test executions. Since this upper bound cannot be reached anyway, 
+Since  $B$ cannot be reached anyway, 
 we just calculate its order of magnitude, and this results again in $O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}  \cdot n^{pq-1}\big)$, as calculated already for $\pmax$ in (\ref{eq:maxO}). 
-
+%
 Summarising these complexity calculations, this results in the following theorem.

### 2018-11-13 scp-journal/article/abstract.tex
update

--- 2018-10-07 paper/abstract.tex
@@
 finiteness is presented here for the first time.
-%Moreover, an optimal
-%technique for testing the admissibility of an implementation's
-%nondeterministic behaviour is described.
 It is shown that the test suites are optimal with respect to two
-aspects:~(a)~the maximal length of test traces cannot be reduced, and (b)~the
+aspects:~(a)~the maximal length of test traces cannot be further reduced, and (b)~the
 nondeterministic behaviour cannot be tested with smaller or fewer sets of

### 2018-11-13 scp-journal/article/introduction.tex
update

--- 2018-10-09 paper/introduction.tex
@@
 
-\subsubsection*{Motivation}
+\subsection{Motivation}
 
@@
 
-\subsubsection*{Contributions}
+\subsection{Contributions}
 
@@
 
-Our results complement work in~\cite{DBLP:conf/pts/CavalcantiS17}. There,
+Our results complement work 
+in~\cite{DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
+
+@todo -- refer to/ compare to both publications
+
+There,
 fault domains are specified as collections of processes refining a ``most
@@
 
-\subsubsection*{Overview}
+\subsection{Overview}
 In Section~\ref{section:preliminaries}, we present the background

### 2018-11-14 scp-journal/article/abstract.tex
Complete version - fit for submission to Science of Computer Programming.

@@
-In this paper, two new contributions for model-based testing with
+In this paper,   new contributions for model-based testing with
 Communicating Sequential Processes (CSP) are presented. For a given CSP
-process representing the reference model, test suites checking the
+process representing the reference model, finite test suites checking the
 conformance relations trace refinement and failures refinement are constructed, and
-their finiteness and completeness (that is, capability to uncover conformity
-violations) is proven. While complete test suites for CSP processes have been
+their   completeness (that is, capability to uncover conformity
+violations) is proven. The fault domains for which complete failure detection can be guaranteed
+are specified by means of normalised transition graphs representing the failures semantics of finite-state CSP processes.
+While complete test suites for CSP processes have been
 previously investigated by several authors, a sufficient condition for their
 finiteness is presented here for the first time. 
-It is shown that the test suites are optimal with respect to two
+Moreover, is shown that the test suites are optimal with respect to two
 aspects:~(a)~the maximal length of test traces cannot be further reduced, and (b)~the

### 2018-11-14 scp-journal/article/introduction.tex
Complete version - fit for submission to Science of Computer Programming.

@@
 being evaluated and integrated into industrial verification processes by many
-companies worldwide. This holds particularly for the embedded and
+companies worldwide. 
+This holds particularly for the embedded and
 cyber-physical systems domains, where critical systems require rigorous
-testing.
+testing~\cite{jp2018ets,DBLP:conf/isola/0001BH18}.
 
@@
 Generation techniques for complete test suites have been developed for
-various formalisms. We tackle \emph{Communicating Sequential
+various formalisms, we mention here representative work in the domains of 
+finite state machines~\cite{hierons_testing_2004,simao_reducing_2012}, timed automata~\cite{Springintveld2001}, Circus~\cite{DBLP:journals/acta/CavalcantiG11}, 
+Timed CSP~\cite{Schneider:1995:OST:203471.203475},
+general labelled transition systems~\cite{DBLP:journals/cn/Tretmans96}, symbolic state machines~\cite{DBLP:conf/icst/Petrenko16}, and Kripke structures~\cite{Huang2017}.
+In this article, we tackle \emph{Communicating Sequential
 Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
@@
 
-\subsection{Contributions}
+\subsection{Main Contributions}
 
-This paper presents complete black-box test suites for software and systems
+This article presents   complete black-box test suites for software and systems
 modelled using CSP. They can be generated for non-terminating,
@@
 divergence and deadlock.
+%
+The  novel main contributions in this article may be summarised as follows.
+\begin{enumerate}
+\item It is shown that trace or failures conformance can be established with finitely many test cases, provided that suitable fault domains are chosen, such that the true behaviour of the SUT is reflected by members of these domains.
 
-Our results complement work 
-in~\cite{DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
+\item The definition of these fault domains is based on the well-known normalised transition graphs~\cite{Roscoe:1994:chapter} representing the  trace and failures semantics of finite-state CSP processes.
+A fault domain contains all CSP processes over a given alphabet, whose normalised transition graphs have at most $q$ nodes for some $q\in\mathbb{N}$.
 
-@todo -- refer to/ compare to both publications
+\item Worst-case complexity bounds for the number of test executions to be performed are given.
 
-There,
-fault domains are specified as collections of processes refining a ``most
-general'' fault domain member. With that concept, complete test suites may be
-finite or infinite. This result gives important insight into the theory of
-fault-domain testing for CSP, but we are particularly interested in {\it
-finite} suites when it comes to practical application.
-While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
-select tests from still infinite test suites, here, we further restrict fault
-domains using a graph representation of processes (originally elaborated for
-model checking) to obtain finite test suites.
+\item It is shown that the maximal length of the test traces involved cannot be further reduced without losing the test suite's completeness property.
 
-Our approach to the definition of CSP fault domains is presented in this
-paper. We take advantage of results on model checking of CSP
-processes, where the failures semantics of a finite-state CSP process is
-represented as a finite normalised transition graph, whose edges are labelled
-by the events the process engages in, and whose nodes are labelled by minimal
-acceptances or, alternatively, maximal
-refusals~\cite{Roscoe:1994:chapter}. The maximal refusals express the
-degree of nondeterminism present in a process state that is in
-one-one-correspondence to a node of the normalised transition graph. Inspired
-by the way fault-domains are specified for finite state machines (FSMs),
-we define them here as the set of CSP processes whose normalised transition
-graphs do not exceed the order (that is, the number of nodes) of the
-reference model's graph by more than a given constant.
+\item Likewise, is shown that the non-deterministic behaviour of the SUT cannot be 
+checked for admissibility with smaller or fewer sets of events.
+\end{enumerate}
 
-The main contribution of this paper is the construction of two test suites to
-verify the conformance relations \emph{trace refinement} and \emph{failures
-refinement} for any given non-terminating, non-divergent, finite-state,
-finite alphabet CSP process. We prove their completeness with respect to
-fault domains of the described type.  The existence of -- possibly infinite
--- complete test suites has been established for process algebras, and for
-CSP in particular, by several
-authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
-To the best of our knowledge, this article is the first to present {\it
-finite}, complete test suites associated with this class of fault domains and
-conformance relations.
+% ==================================================================================
 
-Our result is not a simple transcription of existing knowledge about finite,
-complete test suites for FSMs. The capabilities of CSP to express
-nondeterminism in a more fine-grained way than it is possible for FSMs
-requires a more complex approach to testing for conformity in the failures
-model than required for model-based testing against nondeterministic FSMs, as
-published, for example,
-in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. CSP distinguishes
-between external choice, where the environment can control the behaviour of a
-process, and internal choice where the behaviour is decided internally and
-cannot be influenced by the environment. In contrast, FSMs specify
-nondeterministic behaviour by offering more than one possible output for a
-given input, and this output can be controlled by an FSM representing the
-environment. Loosely speaking, this corresponds to external choice in CSP,
-while there is no equivalent to internal choice in nondeterministic FSMs.
+\subsection{Related Work}
 
-Finally, we prove two optimality results here. (1)~We show that our approach
-to testing the admissibility of an SUT's nondeterministic behaviour is
-optimal in the sense that it cannot be established with fewer ``probes''
-investigating the SUT's degree of nondeterminism. These probes are minimal
-sets of events offered by the tests to the SUT, such that a refusal of the
-SUT to engage into at least one of these events reveals a violation of
-failures refinement. (2)~Furthermore, it is shown that the maximal length of
-traces to be investigated in a complete test suite cannot be further reduced
-without loosing the suite's capability to uncover any violation of trace or
-failures refinement.
+Our results complement and extend work previously published  
+in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}. None of these previous results provide
+sufficient conditions 
+for constructing finite complete test suites.
+Consequently, they also do not provide complexity bounds on the number of test executions needed to establish conformance between SUT and reference process.
+
+In~\cite{DBLP:conf/pts/CavalcantiS17}, 
+also fault domains are used, but these contain all processes refining a ``top-level'' fault domain process. This concept is orthogonal to the one investigated in the present article:
+members of our fault domain need not be in refinement relation to any other process in the domain, they just adhere to the same upper bound $q$ of nodes in their normalised transition graphs.
+
+The minimal sets of events used for checking nondeterministic behaviour of the SUT used in our article were already suggested in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07}; in the present article, however, they have been identified for the first time as {\it minimal hitting sets}~\cite{5533149} of minimal acceptances in a given process state, and we establish an upper bound stating how many of these sets need to be checked for the ``most extreme form of nondeterminism'' that may be exhibited by the SUT.
+
+In~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}, the authors devised linear test cases: after running through a preset trace $s$, test cases for trace refinement check for illegal acceptance of a specific event $e$, and test cases verifying  nondeterministic behaviour check for  acceptance of events from some set $A$.
+In the present article, we follow the alternative approach proposed in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a} and use {\it adaptive} test cases.
+This means that each test case
+adapts its trace execution to the nondeterministic behaviour of the SUT, checks for 
+  trace violations at any point during the test execution, and checks for the acceptance 
+of a given minimal hitting set of events after any legal trace of a test case-specific length.
+The adaptive test cases have the advantage that test executions only lead to an 
+inconclusive result if the reference process allows for a nondeterministic choice between deadlock and trace continuation in a certain state. In contrast to this, the linear test cases 
+may lead to many more futile executions with inconclusive results, if the SUT refuses 
+to engage into the next event $e$ 
+from the preset trace $s$, due to legal nondeterministic choices leading to a refusal of $e$.
+Moreover, the preset traces $s$ need to be executed twice according to the strategies devised in 
+\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}, because trace refinement and correctness of nondeterministic behaviour are checked by two different sets of test cases.
+
+The approach  to specifying fault domains by means of normalised transition graphs has been 
+inspired by the typical method used in the construction of complete test suites for finite state machines (FSMs). There, fault domains typically contain all FSMs over a given alphabet whose 
+number of states does not exceed a given value $q$~\cite{chow:wmethod,vasilevskii1973,luo_test_1994}.
+
+%
+%
+%There,
+%fault domains are specified as collections of processes refining a ``most
+%general'' fault domain member. With that concept, complete test suites may be
+%finite or infinite. This result gives important insight into the theory of
+%fault-domain testing for CSP, but we are particularly interested in {\it
+%finite} suites when it comes to practical application.
+%While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
+%select tests from still infinite test suites, here, we further restrict fault
+%domains using a graph representation of processes (originally elaborated for
+%model checking) to obtain finite test suites.
+%
+%Our approach to the definition of CSP fault domains is presented in this
+%paper. We take advantage of results on model checking of CSP
+%processes, where the failures semantics of a finite-state CSP process is
+%represented as a finite normalised transition graph, whose edges are labelled
+%by the events the process engages in, and whose nodes are labelled by minimal
+%acceptances or, alternatively, maximal
+%refusals~\cite{Roscoe:1994:chapter}. The maximal refusals express the
+%degree of nondeterminism present in a process state that is in
+%one-one-correspondence to a node of the normalised transition graph. Inspired
+%by the way fault-domains are specified for finite state machines (FSMs),
+%we define them here as the set of CSP processes whose normalised transition
+%graphs do not exceed the order (that is, the number of nodes) of the
+%reference model's graph by more than a given constant.
+%
+%The main contribution of this paper is the construction of two test suites to
+%verify the conformance relations \emph{trace refinement} and \emph{failures
+%refinement} for any given non-terminating, non-divergent, finite-state,
+%finite alphabet CSP process. We prove their completeness with respect to
+%fault domains of the described type.  The existence of -- possibly infinite
+%-- complete test suites has been established for process algebras, and for
+%CSP in particular, by several
+%authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
+%To the best of our knowledge, this article is the first to present {\it
+%finite}, complete test suites associated with this class of fault domains and
+%conformance relations.
+%
+%Our result is not a simple transcription of existing knowledge about finite,
+%complete test suites for FSMs. The capabilities of CSP to express
+%nondeterminism in a more fine-grained way than it is possible for FSMs
+%requires a more complex approach to testing for conformity in the failures
+%model than required for model-based testing against nondeterministic FSMs, as
+%published, for example,
+%in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. CSP distinguishes
+%between external choice, where the environment can control the behaviour of a
+%process, and internal choice where the behaviour is decided internally and
+%cannot be influenced by the environment. In contrast, FSMs specify
+%nondeterministic behaviour by offering more than one possible output for a
+%given input, and this output can be controlled by an FSM representing the
+%environment. Loosely speaking, this corresponds to external choice in CSP,
+%while there is no equivalent to internal choice in nondeterministic FSMs.
+%
+%
+%Finally, we prove two optimality results here. (1)~We show that our approach
+%to testing the admissibility of an SUT's nondeterministic behaviour is
+%optimal in the sense that it cannot be established with fewer ``probes''
+%investigating the SUT's degree of nondeterminism. These probes are minimal
+%sets of events offered by the tests to the SUT, such that a refusal of the
+%SUT to engage into at least one of these events reveals a violation of
+%failures refinement. (2)~Furthermore, it is shown that the maximal length of
+%traces to be investigated in a complete test suite cannot be further reduced
+%without loosing the suite's capability to uncover any violation of trace or
+%failures refinement.
 
@@
 relevant to our work. In Section~\ref{sec:finitecompletefails}, finite
-complete test suites for verifying failures refinement are presented. Test
+complete test suites for verifying failures refinement are presented. 
+A sample test
+suite is presented in Section~\ref{sec:case}.
+Test
 suites checking trace refinement are a simplified version of the former
@@
 complexity considerations.
-A sample test
-suite is presented in Section~\ref{sec:case}. Our results are discussed in
-Section~\ref{sec:conc}, where we also conclude. References to related work
+ Our results are discussed in
+Section~\ref{sec:conc}, where we also conclude. References to further related work
 are given throughout the paper where appropriate.

### 2018-11-16 scp-journal/article/abstract.tex
revised sections 1 and 2, shortenned paper

@@
-In this paper,   new contributions for model-based testing with
-Communicating Sequential Processes (CSP) are presented. For a given CSP
-process representing the reference model, finite test suites checking the
-conformance relations trace refinement and failures refinement are constructed, and
-their   completeness (that is, capability to uncover conformity
-violations) is proven. The fault domains for which complete failure detection can be guaranteed
-are specified by means of normalised transition graphs representing the failures semantics of finite-state CSP processes.
-While complete test suites for CSP processes have been
-previously investigated by several authors, a sufficient condition for their
-finiteness is presented here for the first time. 
-Moreover, is shown that the test suites are optimal with respect to two
-aspects:~(a)~the maximal length of test traces cannot be further reduced, and (b)~the
-nondeterministic behaviour cannot be tested with smaller or fewer sets of
-events, without losing the test suite's completeness property.
+In this paper, new contributions for model-based testing using Communicating
+Sequential Processes (CSP) are presented. For a finite non-terminating CSP
+process representing the reference model, finite test suites for checking the
+conformance relations trace and failures refinement are presented, and their
+completeness (that is, capability to uncover conformity violations) is
+proven. The fault domains for which complete failure detection can be
+guaranteed are specified by means of normalised transition graphs
+representing the failures semantics of finite-state CSP processes. While
+complete test suites for CSP processes have been previously investigated by
+several authors, a sufficient condition for their finiteness is presented
+here for the first time. Moreover, it is shown that the test suites are
+optimal in two aspects:~(a)~the maximal length of test traces cannot be
+further reduced, and (b)~the nondeterministic behaviour cannot be tested with
+smaller or fewer sets of events, without losing the test suite's completeness
+property.

### 2018-11-16 scp-journal/article/finitecomplete-failures.tex
revised sections 1 and 2, shortenned paper

--- 2018-10-11 paper/finitecomplete-failures.tex
@@
 % -------------------------------------------------------------------------
-\subsubsection*{Test Definition and Basic Properties}
-
+\paragraph{Test Definition and Basic Properties}
 In the domain of process algebras, test cases are typically represented by
@@
 \\ & & \extchoice \nonumber
-\\ & & (k = j \wedge \minhits(n)\neq\varnothing) \&  \nonumber
-\\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
+\\ & & (k = j \wedge \minhits(n)\neq\varnothing) \& \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
 \label{eq:ufd}
@@
 % --------------------------------------------------------------------------
-\subsubsection*{Explanation of the Test Definition}
-A test is performed by running $U_F(j)$ concurrently with any SUT process
-$Q$, synchronising over $\Sigma$. So, a \emph{test execution} is a trace of the
-concurrent process
+\paragraph{Explanation of the Test Definition} A test \pagebreak is performed by running
+$U_F(j)$ concurrently with any SUT process $Q$, synchronising over $\Sigma$.
+So, a \emph{test execution} is a trace of the concurrent process
 %\[
@@
 
-It is assumed that the events $\efail$ and $\epass$, indicating a verdict
-FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
+It is assumed that the events $\efail$ and $\epass$, indicating verdicts FAIL
+and PASS for the test execution, are not included in $\Sigma$. Since we
 assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
@@
 %
-This type of pass relation is often called \emph{must test} in the
-literature, because every test execution must end with the $\epass$
+This type of pass relation is often called \emph{must test}, because every
+test execution must end with the $\epass$
 event~\cite{Hennessy:1988:ATP:50497}.
 
-We note that it is necessary to use failures refinement in the definition
-above, and not just trace refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide
-\Sigma$ may have the same visible traces $\varepsilon$ and $\epass$ as the
-``Test Passed Process'' $(\epass\then\Stop)$. However, the former may
-nondeterministically refuse $\epass$, due to a deadlock occurring when a
-faulty SUT process executes concurrently with $U_F(j,k,n)$ executing
-branch~(\ref{eq:ufd}), when the guard condition $(k = j\wedge
-\minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
-in the next paragraphs. Alternatively, a faulty SUT $Q$ might internally
-deadlock after a trace $s$ where $\#s < j$ and $\minhits(G(P)/s) \neq
-\varnothing$, so that the process $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks
-as well.
+It is necessary to use failures refinement in the definition above, and not
+just trace refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ may have
+the same visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
+Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
+refuse $\epass$, due to a deadlock occurring when a faulty SUT process
+executes concurrently with $U_F(j,k,n)$ executing branch~(\ref{eq:ufd}), when
+the guard condition $(k = j\wedge \minhits(n)\neq\varnothing)$ evaluates to
+$\ist$. This is explained further in the next paragraphs. Alternatively, a
+faulty SUT $Q$ might internally deadlock after a trace $s$ where $\#s < j$
+and $\minhits(G(P)/s) \neq \varnothing$, so that the process
+$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks as well.
 
-Intuitively speaking, $U_F(j)$ is able to perform any trace $s$ of $P$, up to
-a length $j$. If, after having already run through $s\in\trc(P)$ with $\#s
-\le j$, an event is accepted by the SUT that is outside the initials of $P/s$
+Intuitively, $U_F(j)$ is able to perform any trace $s$ of $P$, up to a length
+$j$. If, after having already run through $s$ with $\#s \le j$, the SUT
+accepts an event outside the initials of $P/s$
  (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(j)/s$),
@@
 case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
-external choice in $U_F(j,\#s,G(P)/s)$). If $P/s$ may also
-nondeterministically engage into events in such a situation, branch
-(\ref{eq:ufc}) of the test is simultaneously enabled. If $Q/s$ is able to
-engage into an event from $\Sigma - [P/s]^0$, a test execution exists where
-$U_F(j,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the $\efail$
-event.
+choice in $U_F(j,\#s,G(P)/s)$). If $P/s$ may also nondeterministically engage
+into events, branch (\ref{eq:ufc}) is simultaneously enabled. If $Q/s$ is
+able to engage into an event in $\Sigma - [P/s]^0$, a test execution exists
+where $U_F(j,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the
+$\efail$ event.
 
@@
 from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
-$U_F(j,\#s+1,G(P)/s.e)$ (note that $G(P)/s.e = t(G(P)/s,e)$)
-in branch~(\ref{eq:ufc}); this follows from
-Lemma~\ref{lemma:ufproperties}. A test of this type is called
-\emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts
-its consecutive behaviour to the event selected by the SUT.
+$U_F(j,\#s+1,G(P)/s.e)$ in branch~(\ref{eq:ufc}); this follows from
+Lemma~\ref{lemma:ufproperties}~(note that $G(P)/s.e = t(G(P)/s,e)$). A test
+of this type is called \emph{adaptive}, because it accepts any legal
+behaviour of the SUT, here any event from $[P/s]^0$, and adapts its
+consecutive behaviour to the event selected by the SUT, here
+$U_F(j,\#s+1,G(P)/s.e)$.
 
-Now suppose that a test execution has run through a trace
-$s\in\trc(P)$ of length $j$, so that $U_F(j)/s = U_F(j,j,n)$ with $n =
-G(P)/s$. If $\minhits(n)\neq \varnothing$, the test changes its
-behaviour:~instead of offering {\it all} legal events from $[n]^0$ to the
-SUT, it nondeterministically chooses a minimal hitting set $H\in \minhits(n)$
-and only offers the events contained in $H$. If the SUT refuses to engage
-into some event of $H$, this reveals a violation of failures
-refinement:~according to Lemma~\ref{lemma:hseta}, a conforming SUT should
-accept at least one event of each minimal hitting set in $\minhits(n)$.
-Therefore, the test execution only terminates with  $\epass$, if such an
-event is accepted by the SUT. Otherwise, it deadlocks, and the test fails.
+Now suppose that a test execution has run through a trace $s\in\trc(P)$ of
+length $j$, so that $U_F(j)/s = U_F(j,j,n)$ with $n = G(P)/s$. If
+$\minhits(n)\neq \varnothing$, the test changes its behaviour:~instead of
+offering {\it all} legal events from $[n]^0$ to the SUT, it
+nondeterministically chooses a minimal hitting set $H\in \minhits(n)$ and
+only offers the events contained in $H$. If the SUT refuses to engage into
+some event of $H$, this reveals a violation of failures refinement:~according
+to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least one event
+of each minimal hitting set in $\minhits(n)$. Therefore, the test execution
+terminates with  $\epass$, only if such an event is accepted. Otherwise, it
+deadlocks, and the test fails.
 
@@
 execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
-entered, the process either accepts another event $e$ of some minimal hitting
-set $H\in\minhits(n)$ with $n = G(P)/s$ according to
-Lemma~\ref{lemma:ufproperties}. Then the final length of $s$ is $j+1$, and
-the execution terminates with $\epass$. Or the test execution
-$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks,  the final length of $s$ is $j$,
-and the execution stops without a PASS or FAIL event. Such an execution is
-also interpreted as FAIL, because it reveals that $(\epass\then\Stop)
-\not\lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$.
+entered, then there are two possibilities. The first is that the process
+accepts another event $e$ of some minimal hitting set $H\in\minhits(n)$ with
+$n = G(P)/s$ according to Lemma~\ref{lemma:ufproperties}. In this case, the
+final length of $s$ is $j+1$, and the execution terminates with $\epass$.
+Alternatively, the test execution $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks,
+the final length of $s$ is $j$, and the execution stops without a PASS or
+FAIL event. Such an execution is also interpreted as FAIL, because it reveals
+that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide
+\Sigma$.
 
@@
 
-We state basic properties of our tests in the following lemma, to show the
-relationships between $U_F(j)$ and the reference process $P$ from which it is
-derived.
+The following lemma establishes relationships between $U_F(j)$ and the
+reference process $P$ from which it is derived.
 
@@
 $U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
-specification of this branch implies that
-$$
-U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = (\efail\then\Stop).
-$$
-%
+specification of this branch implies that $_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
+(\efail\then\Stop)$.
+
 Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because
@@
 % -------------------------------------------------------------------------
-\subsubsection*{Complete Testing Assumption}
-As explained above, passing a test case $U_F(j)$ requires that none of
-the possible executions $(Q\parallel[\Sigma] U_F(j))$ stops after $\efail$ or
-stops without having produced event $\epass$. Therefore, it is necessary to determine
-whether all possible executions have been covered in the repeated runs of
-$(Q\parallel[\Sigma] U_F(j))$.
-The theoretical completeness results are therefore based on
-a  \emph{complete testing
-assumption}~\cite{hierons_testing_2004,DBLP:conf/icfem/CavalcantiG07},
- which means that every
-possible behaviour of the SUT is performed after a finite number of test
-executions. In practice, this is realised by executing each test several
-times, recording the traces that have been performed, and using hardware or
-software coverage analysers to determine whether all possible test execution
-behaviours of the SUT have been observed. Therefore,
-testing nondeterministic SUTs
-  comes
-at the price of having to apply some grey-box testing techniques enabling us
-to decide whether all SUT behaviours have been observed.
+\paragraph{Complete Testing Assumption} As explained above, passing a test case
+$U_F(j)$ requires that none of the possible executions $(Q\parallel[\Sigma]
+U_F(j))$ stops after $\efail$ or stops without having produced the event
+$\epass$. Therefore, it is necessary to determine whether all possible
+executions have been covered in the repeated runs of $(Q\parallel[\Sigma]
+U_F(j))$. The theoretical completeness results are, therefore, based on a
+\emph{complete testing
+assumption}~\cite{hierons_testing_2004,DBLP:conf/icfem/CavalcantiG07}, which
+means that every possible behaviour of the SUT is performed after a finite
+number of test executions. In practice, this is realised by executing each
+test several times, recording the traces that have been performed, and using
+hardware or software coverage analysers to determine whether all possible
+behaviours of the SUT have been observed. Therefore, testing nondeterministic
+SUTs comes at the price of having to apply some grey-box testing techniques.
+%enabling us to decide whether all SUT behaviours have been observed.
 
@@
 
-A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ counsists of a
+A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a
 reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
 \lessdet_F\}$, and a fault domain ${\cal D}$, which is a set of CSP processes
-over $P$'s alphabet that may or may not conform to $P$.
-
-A test suite $\TS$ is called \emph{complete} with respect to fault model
-${\cal F}$, if, and only if, the following conditions are fulfilled.
+over $P$'s alphabet that may or may not conform to $P$. A test suite $\TS$ is
+called \emph{complete} with respect to fault model ${\cal F}$, if, and only
+if, the following conditions are fulfilled.
 \begin{description}
@@
 \[
-\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{$U_F(j)$ specified in (\ref{eq:UFP})}
+\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{with $U_F(j)$ as specified in (\ref{eq:UFP})}
 \]
@@
 %
-The proof of the theorem follows from the two lemmas below. The first states
-that test suite $\TS_F$ is sound, the second states that the suite is also
-exhaustive.
+The proof of the theorem follows directly from the two lemmas below. The
+first lemma establishes that test suite $\TS_F$ is sound, and the second
+establishes that the suite is also exhaustive.
 %

### 2018-11-16 scp-journal/article/introduction.tex
revised sections 1 and 2, shortenned paper

@@
 
-\subsection{Motivation}
-
+\paragraph{Motivation}
 Model-based testing (MBT) is an active research field; results are currently
 being evaluated and integrated into industrial verification processes by many
-companies worldwide. 
-This holds particularly for the embedded and
+companies worldwide. This holds particularly for the embedded and
 cyber-physical systems domains, where critical systems require rigorous
@@
 Generation techniques for complete test suites have been developed for
-various formalisms, we mention here representative work in the domains of 
-finite state machines~\cite{hierons_testing_2004,simao_reducing_2012}, timed automata~\cite{Springintveld2001}, Circus~\cite{DBLP:journals/acta/CavalcantiG11}, 
-Timed CSP~\cite{Schneider:1995:OST:203471.203475},
-general labelled transition systems~\cite{DBLP:journals/cn/Tretmans96}, symbolic state machines~\cite{DBLP:conf/icst/Petrenko16}, and Kripke structures~\cite{Huang2017}.
-In this article, we tackle \emph{Communicating Sequential
-Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
-process algebra that has been shown to be well-suited for the description of
-reactive control systems in many publications over almost five decades. Many
-of these applications are described in~\cite{Roscoe2010} and in the
-references there. Industrial success has also been reported; see, for
+various formalisms; we mention here representative work for finite state
+machines~\cite{hierons_testing_2004,simao_reducing_2012}, timed
+automata~\cite{Springintveld2001}, {\sf\em
+Circus}~\cite{DBLP:journals/acta/CavalcantiG11}, Timed
+CSP~\cite{Schneider:1995:OST:203471.203475}, general labelled transition
+systems~\cite{DBLP:journals/cn/Tretmans96}, symbolic state
+machines~\cite{DBLP:conf/icst/Petrenko16}, and Kripke
+structures~\cite{Huang2017}. In this article, we tackle \emph{Communicating
+Sequential Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a
+mature process algebra that has been shown to be well-suited for the
+description of reactive control systems in many publications over almost five
+decades. Many of these applications are described in~\cite{Roscoe2010} and in
+the references there. Industrial success has also been reported; see, for
 example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.
@@
 
-\subsection{Main Contributions}
+\paragraph{Main Contributions}
 
-This article presents   complete black-box test suites for software and systems
-modelled using CSP. They can be generated for non-terminating,
+This article presents   complete black-box test suites for software and
+systems modelled using CSP. They can be generated for non-terminating,
 divergence-free, finite-state CSP processes with finite alphabets,
@@
 is usually assumed in black-box testing, since it cannot distinguish between
-divergence and deadlock.
-%
-The  novel main contributions in this article may be summarised as follows.
+divergence and deadlock using testing.
+
+\newpage
+\noindent %
+The main novel contributions in this article may be summarised as
+follows.
 \begin{enumerate}
-\item It is shown that trace or failures conformance can be established with finitely many test cases, provided that suitable fault domains are chosen, such that the true behaviour of the SUT is reflected by members of these domains.
+\item It is shown that trace or failures conformance can be established
+    with finitely many test cases, provided suitable fault domains are
+    chosen, so that the true behaviour of the SUT is reflected by members
+    of these domains.
 
@@
 
-\subsection{Related Work}
-
+\paragraph{Related Work}
 Our results complement and extend work previously published
-in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}. None of these previous results provide
-sufficient conditions 
-for constructing finite complete test suites.
-Consequently, they also do not provide complexity bounds on the number of test executions needed to establish conformance between SUT and reference process.
-
-In~\cite{DBLP:conf/pts/CavalcantiS17}, 
-also fault domains are used, but these contain all processes refining a ``top-level'' fault domain process. This concept is orthogonal to the one investigated in the present article:
-members of our fault domain need not be in refinement relation to any other process in the domain, they just adhere to the same upper bound $q$ of nodes in their normalised transition graphs.
+in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
+None of these provide sufficient conditions for constructing finite complete
+test suites. So, they also do not provide complexity bounds on the number of
+test executions needed to establish conformance between an SUT and a
+reference process. In~\cite{DBLP:conf/pts/CavalcantiS17}, fault domains are
+used, but these contain all processes refining a ``top'' fault domain
+process. This concept is orthogonal to the one investigated here:~members of
+our fault domain need not be in refinement relation to any other process in
+the domain. They just adhere to the same upper bound $q$ of nodes in their
+normalised transition graphs.
 
@@
 
-In~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}, the authors devised linear test cases: after running through a preset trace $s$, test cases for trace refinement check for illegal acceptance of a specific event $e$, and test cases verifying  nondeterministic behaviour check for  acceptance of events from some set $A$.
-In the present article, we follow the alternative approach proposed in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a} and use {\it adaptive} test cases.
-This means that each test case
-adapts its trace execution to the nondeterministic behaviour of the SUT, checks for 
-  trace violations at any point during the test execution, and checks for the acceptance 
-of a given minimal hitting set of events after any legal trace of a test case-specific length.
-The adaptive test cases have the advantage that test executions only lead to an 
-inconclusive result if the reference process allows for a nondeterministic choice between deadlock and trace continuation in a certain state. In contrast to this, the linear test cases 
-may lead to many more futile executions with inconclusive results, if the SUT refuses 
-to engage into the next event $e$ 
-from the preset trace $s$, due to legal nondeterministic choices leading to a refusal of $e$.
-Moreover, the preset traces $s$ need to be executed twice according to the strategies devised in 
-\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}, because trace refinement and correctness of nondeterministic behaviour are checked by two different sets of test cases.
+In~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
+the authors devised linear test cases: after running through a preset trace
+$s$, test cases for trace refinement check for illegal acceptance of a
+specific event $e$, and test cases verifying  nondeterministic behaviour
+check for  acceptance of events from some set $A$. In the present article, we
+follow the alternative approach proposed
+in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a} and use {\it adaptive} test
+cases. This means that each test case adapts its trace execution to the
+nondeterministic behaviour of the SUT, checks for trace violations at any
+point during the test execution, and checks for the acceptance of a given
+minimal hitting set of events after any legal trace of a test case-specific
+length.
 
-The approach  to specifying fault domains by means of normalised transition graphs has been 
-inspired by the typical method used in the construction of complete test suites for finite state machines (FSMs). There, fault domains typically contain all FSMs over a given alphabet whose 
-number of states does not exceed a given value $q$~\cite{chow:wmethod,vasilevskii1973,luo_test_1994}.
+The adaptive test cases have the advantage that test executions only lead to
+an inconclusive result if the reference process allows for a nondeterministic
+choice between deadlock and trace continuation in a certain state. In
+contrast to this, the linear test cases may lead to many more futile
+executions with inconclusive results, if the SUT refuses to engage into the
+next event $e$ from the preset trace $s$, due to legal nondeterministic
+choices leading to a refusal of $e$. Moreover, the preset traces $s$ need to
+be executed twice according to the strategies devised in
+\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
+because trace refinement and correctness of nondeterministic behaviour are
+checked by two different sets of test cases.
 
-%
-%
+The approach  to specifying fault domains by means of normalised transition
+graphs has been inspired by the typical method used in the construction of
+complete test suites for finite state machines (FSMs). There, fault domains
+typically contain all FSMs over a given alphabet whose number of states does
+not exceed a given value
+$q$~\cite{chow:wmethod,vasilevskii1973,luo_test_1994}.
+
 %There,
@@
 %
-%
 %Finally, we prove two optimality results here. (1)~We show that our approach
@@
 
-\subsection{Overview}
-In Section~\ref{section:preliminaries}, we present the background
-relevant to our work. In Section~\ref{sec:finitecompletefails}, finite
-complete test suites for verifying failures refinement are presented. 
-A sample test
-suite is presented in Section~\ref{sec:case}.
-Test
-suites checking trace refinement are a simplified version of the former
-class; they are presented in Section~\ref{sec:finitecomplete}.
-The optimality results are presented in Section~\ref{sec:complexity}, together with further
-complexity considerations.
- Our results are discussed in
-Section~\ref{sec:conc}, where we also conclude. References to further related work
-are given throughout the paper where appropriate.
+\paragraph{Overview} In Section~\ref{section:preliminaries}, we present the
+background relevant to our work. In Section~\ref{sec:finitecompletefails},
+finite complete test suites for verifying failures refinement are presented.
+A sample test suite is presented in Section~\ref{sec:case}. Test suites
+checking trace refinement are a simplified version of the former class; they
+are presented in Section~\ref{sec:finitecomplete}. The optimality results are
+presented in Section~\ref{sec:complexity}, together with further complexity
+considerations. Our results are discussed in Section~\ref{sec:conc}, where we
+also conclude. References to further related work are given throughout the
+paper where appropriate.
 

### 2018-11-16 scp-journal/article/preliminaries.tex
Re-formatted for Science of Computing journal (new directory scp-journal). Re-worked discussion.
revised sections 1 and 2, shortenned paper

--- 2018-10-11 paper/preliminaries.tex
@@
 We present CSP~(Section~\ref{section:csp}) and the concept of minimal hitting
-sets~(Section~\ref{sec:hit}), which is central to our novel notion of test
+sets~(Section~\ref{sec:hit}), which is central to our notion of test
 for failures refinement. To study complexity, we also introduce the concept
@@
 
-\subsubsection*{Communicating Sequential Processes (CSP)} is a process algebra
-supporting system development by refinement. Using CSP, we model both systems and
-their components using processes. They are characterised by their patterns
-of interactions, modelled by synchronous, instantaneous, and atomic events.
+\paragraph{Communicating Sequential Processes (CSP)} This is a process algebra supporting
+system development by refinement. Using CSP, we model both systems and their
+components using processes. They are characterised by their patterns of
+interactions, modelled by synchronous, instantaneous, and atomic events.
 
 Throughout this paper, the alphabet of the processes, that is, the set of
-events that are in scope, is denoted by $\Sigma$ and supposed to be finite.
+events in scope, is denoted by $\Sigma$ and supposed to be finite.
 The FDR tool~\cite{fdr} supports model checking and semantic analyses of
@@
 \extchoice Q$ between processes $P$ and $Q$ offers to the environment the
-initial events of $P$ and $Q$. Once synchronisation takes place, the process
-that has offered this event is chosen and the other is discarded. In an
-internal choice $P \intchoice Q$, the environment does not have an
-opportunity to interfere:~the choice is made by the process.
+initial events of $P$ and $Q$. Once a synchronisation takes place, the
+process that has offered the event that has occurred is chosen and the other
+is discarded. In an internal choice $P \intchoice Q$, the environment does
+not have an opportunity to interfere:~the choice is made by the process.
 %
 \begin{example}\label{example:CSP}
-  We consider $P$, $Q$, and $R$ below. $P$ is initially
+  We consider the processes $P$, $Q$, and $R$ defined below. $P$ is initially
   ready to engage in the event $a$, and then makes an internal choice to
@@
 $cs$. Other forms of parallelism available in CSP can be defined using this
-operator.
+parallelism operator.
 
@@
 Semantic definitions specify, for each operator, how the traces or failures
-of the resulting process can be calculated from the traces or failures of
-each operand. For example, for internal choice, $\fails(P\intchoice Q) =
+of the resulting process can be calculated from those of each operand. For
+example, for internal choice, $\fails(P\intchoice Q) =
 \fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for a
-comprehensive list of these definitions.
+comprehensive list of these definitions covering $\trc(P)$ and $\fails(P)$.
 
-Using the notation $P/s$ to denote the process $P$ after having engaged into
-trace $s$, the set $\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains
-the  refusals of $P$ after $s$. Refusals are
-subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure
-of $P$ and $Y\subseteq X$, then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$
-follows.
+Using the notation $P/s$ to denote the behaviour ot the process $P$ after
+having engaged into the events in the trace $s$, the set $\refs(P/s) \defs
+\{~X | (s,X) \in \fails(P)~\}$ contains the  refusals of $P$ after $s$.
+Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$
+is a failure of $P$ and $Y\subseteq X$, then $(s,Y)\in\fails(P)$ and
+$Y\in\refs(P/s)$ follows.
 
 For divergence-free processes, failures refinement, $P \lessdet_F Q$, is
-defined as expected by set containment $\fails(Q)\subseteq\fails(P)$. Since
-refusals are subset-closed, $P \lessdet_F Q$ implies
-$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. So, for
-divergences-free processes, failures refinement implies trace refinement.
-Therefore, using the conformance relation $conf$ below
+defined by $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
+$P \lessdet_F Q$ implies $(s,\varnothing)\in\fails(P)$ for all traces
+$s\in\trc(Q)$. So, for divergence-free processes, failures refinement implies
+trace refinement. Therefore, using the conformance relation $conf$ below
 %
@@
 %
-failures refinement can be expressed by $\lessdet_T$ and $conf$.
+failures refinement can be expressed by $\lessdet_T$ and $conf$ as proven
+in~\cite{DBLP:conf/icfem/CavalcantiG07}.
 %
-\begin{equation}\label{thm:fref-tref-conf}
+\begin{equation}%\label{thm:fref-tref-conf}
+\label{eq:failconf}
 (P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
@@
 %
-For finite CSP processes, since refusals are
-subset-closed, $\refs(P/s)$ can be constructed from the set of \emph{maximal
-refusals} defined by
+For finite processes, since refusals are subset-closed, $\refs(P/s)$ can be
+constructed from the set of \emph{maximal refusals}.
 %
@@
 Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
-reconstruct the refusals in the set $\refs(P/s)$ as described by
+reconstruct the refusals in the set $\refs(P/s)$. % as described by
 %
@@
 in combination with the initials of a process express its degree of
-nondeterminism.
+nondeterminism as illustrated by the next example.
 %
@@
 initial event of $Q$, but also refuse everything, due to internal selection
-of $\Stop$.
-
-Assuming an alphabet $\Sigma = \{a,b,c,d\}$, the process
+of $\Stop$. Assuming an alphabet $\Sigma = \{a,b,c,d\}$, the process
 %
@@
 % =========================================================================
-\subsubsection*{Normalised Transition Graphs for CSP Processes}
-\label{sec:ntg}
+\paragraph{Normalised Transition Graphs for CSP Processes} \label{sec:ntg}
 
@@
 label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
-the fact that $t$ is a function. The graph construction
-in~\cite{Roscoe:1994:chapter} implies that all nodes $n$ in $N$ are
-reachable by sequences of edges labelled by $e_1\dots e_k$ and connecting
-states $\ii n,n_1,\dots,n_{k-1},n$, such that
+the fact that $t$ is a function.
+
+The graph construction in~\cite{Roscoe:1994:chapter} implies that all nodes
+$n$ in $N$ are reachable by sequences of edges labelled by $e_1\dots e_k$ and
+connecting states $\ii n,n_1,\dots,n_{k-1},n$, such that
 \[
@@
 a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
-$s$. In analogy to $\trc(P)$, we use the notation $\trc(G(P))$ for the set of
-finite, initialised paths through $G(P)$, each path represented by its finite
-sequence of edge labels. We note that $\trc(P) = \trc(G(P))$. Since $G(P)$ is
-normalised, there is a unique node reached by following the events from $s$
-one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well defined.
+the events in $s$ in the order they appear. In analogy to $\trc(P)$, we use
+the notation $\trc(G(P))$ for the set of finite, initialised paths through
+$G(P)$, each path represented by its finite sequence of edge labels. We note
+that $\trc(P) = \trc(G(P))$. Since $G(P)$ is normalised, there is a unique
+node reached by following the events from $s$ one by one, starting in $\ii
+n$. Therefore, $G(P)/s$  is also well defined.
+
 By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
@@
 $$
-The graph construction from~\cite{Roscoe:1994:chapter} used to define
-$G(P)$ for any process $P$ guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
+The graph construction from~\cite{Roscoe:1994:chapter} guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
 traces $s$ of $P$.
@@
 % =========================================================================
-\subsubsection*{Acceptances}
-\label{sec:accs}
+\paragraph{Acceptances} \label{sec:accs}
 
 When investigating tests for failures refinement, the notion of
-\emph{acceptances}, which is dual to refusals, is also useful. While the
+\emph{acceptances}, which is dual to refusals, is useful. While the
 original introduction of acceptances presented
@@
 internally chooses one of its \emph{minimal acceptance sets} $A$ and never
-refuses any event in $A$, while {\it possibly} refusing the events from
-$[P/s]^0 - A$ and {\it always} refusing those from $\Sigma - [P/s]^0$.
+refuses any event in $A$, while {\it possibly} refusing the events from the
+set $[P/s]^0 - A$ and {\it always} refusing those in the set $\Sigma -
+[P/s]^0$.
 
-Exploiting (\ref{eq:accref}),
-the nodes of a normalised transition graph can
-alternatively be labelled with their minimal acceptances, and this
-information is equivalent to that contained in the maximal refusals. Since
-process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
-normalised transition graph, we also write $\minaccs(G(P)/s)$ and
-note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply
+Exploiting (\ref{eq:accref}), the nodes of a normalised transition graph can
+alternatively be labelled with minimal acceptances; this captured the same
+information conveyed by maximal refusals. Since process states $P/s$ are
+equivalently expressed by states $G(P)/s$ of $P$'s normalised transition
+graph, we also write $\minaccs(G(P)/s)$ and note that (\ref{eq:refsAndR}) and
+(\ref{eq:accref}) imply
 \begin{equation}\label{eq:maxrefsminaccs}
@@
 With this definition, it is established that $P$ is semantically equivalent
-to the process $\text{normalised}(P)$ in the CSP failures semantics.
+to $\text{normalised}(P)$ in the failures semantics.
 
@@
    \begin{center}
-     \includegraphics[width=\textwidth]{p.pdf}
+     \includegraphics[trim=0cm 1cm 0cm 1cm,clip,scale=.6,%width=\textwidth
+                              ]{p.pdf}
    \end{center}
@@
  \end{figure}
-
 % .......................................................................................
@@
 into $c$, the $R$-component of $P\intchoice R$ must have processed $c$, so
-node~3 corresponds to process state $R/c = R$, and therefore it is labelled
-by $R$'s minimal acceptance $\{b,c\}$. \xbox
+node~3 corresponds to $R/c = R$, and so it is labelled by $R$'s minimal
+acceptance $\{b,c\}$. \xbox
 \end{example}
 %
-Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
-expressed by means of their normalised transition graphs
+\noindent%
+Summarising, refinement between finite-state CSP processes $P, Q$ can be
+expressed using their normalised graphs
 \begin{eqnarray*}
@@
   \label{eq:trcrefa}
-  \\
-  \label{eq:failconf}
-  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
+%  \\
+%  \label{eq:failconf}
+%  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
   \\
@@
 \begin{proof}
-Recall that $P\lessdet_T Q$ is defined as $\trc(Q)\subseteq \trc(P)$ and that
-$\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$. This shows (\ref{eq:trcrefa}).
-Result (\ref{eq:failconf}) has been shown
-in~\cite{DBLP:conf/icfem/CavalcantiG07}.
+To prove (\ref{eq:trcrefa}), we recall that $P\lessdet_T Q$ is defined as
+$\trc(Q)\subseteq \trc(P)$ and, moreover, $\trc(P)=\trc(G(P))$ and
+$\trc(Q)=\trc(G(Q))$. %Result (\ref{eq:failconf}) has been shown
+%in~\cite{DBLP:conf/icfem/CavalcantiG07}.
 %
@@
   \Leftrightarrow \forall s\in\trc(P)\cap \trc(Q): \refs(Q/s)\subseteq\refs(P/s)
- \\
+  %\\
   & Definition of $conf$  (\ref{eq:conf})
@@
   \Leftrightarrow \forall s\in\trc(G(P))\cap \trc(G(Q)), R_Q\in r_Q(G(Q)/s):
-\\\tabc
   \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
+  \\ %
   & Property of $r_P, r_Q$ and (\ref{eq:refMaxRef})
@@
 % =========================================================================
-\subsubsection*{Reachability Under Sets of Traces}
-\label{sec:V} Given a finite-state CSP process $P$ and its normalised
-transition graph $G(P)$,
+\paragraph{Reachability Under Sets of Traces} \label{sec:V} Given a finite-state CSP
+process $P$ and its normalised transition graph $G(P)$,
 %\[
@@
 that is, $|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences
-from $V$, extended by any event of $\Sigma$. We can then conclude that $G(P)$
-reaches at least $(k+1)$ nodes under $V\cup V.\Sigma$.
+from $V$, extended by any event of $\Sigma$.  Then $G(P)$ reaches at least
+$(k+1)$ nodes under $V\cup V.\Sigma$.
 \end{lemma}
@@
 $V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
-Note, however, that it is not necessarily the case that $s_1\in V$.
-
-Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such
-that $G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
+Note, however, that it is not necessarily the case that $s_1\in V$. Since
+$G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such that
+$G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
 $G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
@@
 % =========================================================================
-\subsubsection*{Graph Products}
-\label{sec:GP}
-
-For proving our main theorems, it is necessary to consider the \emph{product}
-of normalised transition graphs. We need this only for the investigation of
-corresponding traces in reference processes and processes for SUTs. So, the
-labelling of nodes with maximal refusals or minimal acceptances are
-disregarded in the product construction. We consider two normalised
-transition graphs
+\paragraph{Graph Products} \label{sec:GP} For proving our main theorems, it is
+necessary to consider the \emph{product} of normalised transition graphs. We
+need this only for the investigation of corresponding traces in reference
+processes and processes for SUTs. So, the labelling of nodes with maximal
+refusals or minimal acceptances are disregarded in the product construction.
+We consider two normalised transition graphs
 \[
@@
 \\
-\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma~|   \nonumber
-\\ & & \tabc
-(n_1,e)\in\dom~t_1\wedge
-(n_2,e) \in\dom~t_2    \}
+\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma | (n_1,e)\in\dom~t_1\wedge (n_2,e) \in\dom~t_2 \}
 \\
@@
 %
-This concludes our presentation of CSP and of some results regarding its
-semantics and normalised transition graphs that are used in the next section.
+This concludes our presentation of CSP and of results regarding its
+semantics that are used in the next section.
 
@@
 
-\subsubsection*{Definition}
-The main idea of the underlying test strategy for failures refinement is
-based on solving a \emph{hitting set problem}. Given a finite collection of
-finite sets ${\cal C} = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a subset of a
-universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set
-satisfying the following property.
+\paragraph{Definition} The main idea of the underlying test strategy for failures
+refinement is based on solving a \emph{hitting set problem}. Given a finite
+collection of finite sets ${\cal C} = \{ A_1,\dots,A_n\}$, such that each
+$A_i$ is a subset of a universe $\Sigma$, a \emph{hitting set}
+$H\subseteq\Sigma$ is a set satisfying the following property.
 %
@@
 A \emph{minimal hitting set} is a hitting set that cannot be further reduced
-without losing the characteristic property (\ref{eq:hit}). By $\minhits({\cal C})$
-we denote the collection of minimal hitting sets for a collection ${\cal C}$. For
-the pathological case where $C$ contains an empty set, $\minhits({\cal C})$ is also
-empty.
-
-The problem of determining minimal hitting sets is %%% \cite{Book1975-BOOKRM,5533149}
-NP-hard~\cite{5533149}. We see below, however, that it reduces the effort of
-testing for failures refinement from a factor of $2^{|\Sigma|}$ to a factor
-that equals the number of minimal hitting sets.
+without losing the characteristic property (\ref{eq:hit}). By $\minhits({\cal
+C})$ we denote the collection of minimal hitting sets for a collection ${\cal
+C}$. For the pathological case where $C$ contains an empty set,
+$\minhits({\cal C})$ is also empty. The problem of determining minimal hitting sets is %%% \cite{Book1975-BOOKRM,5533149}
+NP-hard~\cite{5533149}. We see below, however, that using minimal hitting
+sets, we can reduce the effort of testing for failures refinement from a
+factor of $2^{|\Sigma|}$ to a factor that equals the number of minimal
+hitting sets.
 
 % -----------------------------------------------------------------------------
-\subsubsection*{Minimal Hitting Sets of Minimal Acceptances}
-In this article, we are interested in the minimal hitting sets of minimal acceptances;
-for these, the abbreviated notation
-\[
-\minhits(P/s) = \minhits(\minaccs(P/s))
-\]
-is used. The minimal hitting sets of minimal acceptances may be alternatively
-characterised by means of the failures of a process as is done
-in~\cite{DBLP:conf/icfem/CavalcantiG07}. To this end,
-in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for any
-collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
+\paragraph{Minimal Hitting Sets of Minimal Acceptances} In this article, we are
+interested in the minimal hitting sets of minimal acceptances; for these, the
+abbreviated notation
+$%\[ \
+minhits(P/s) = \minhits(\minaccs(P/s))
+$ %\]
+is used. The minimal hitting sets
+of minimal acceptances may be alternatively characterised by means of the
+failures of a process as is done in~\cite{DBLP:conf/icfem/CavalcantiG07}. To
+this end, in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for
+any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
 \begin{equation}\label{eq:minsub}
@@
 \begin{lemma}
-
+\label{lemma:failureshittingsets}
 For any trace $s$ of CSP process $P$, define
-\[
+$%\[
 {\cal A}_s = \{ A\subseteq \Sigma~|~(s,A)\not\in\fails(P)  \}.
-\]
+$ %\]
 Then
-\[
-\minhits(P/s) = \minsub {\cal A}_s \ \text{for all traces $s$ of $P$}.
-\]
+$ %\[
+\minhits(P/s) = \minsub {\cal A}_s
+$ %\]
+for all traces $s$ of $P$.
 \end{lemma}
@@
 % -------------------------------------------------------------------------
-\subsubsection*{Minimal Hitting Sets of Normalised Transition Graphs}
-Since, as previously explained, minimal acceptances can be used to label the
-nodes of a normalised transition graph, and since $\minaccs(P/s) =
-\minaccs(G(P)/s)$ according to (\ref{eq:maxrefsminaccs}), the notation of
-minimal hitting sets also carries over to graphs: we write $\minhits(n)$ for
-nodes $n$ of $G(P)$ and observe that
+\paragraph{Minimal Hitting Sets of Normalised Transition Graphs} Since, as
+previously explained, minimal acceptances can be used to label the nodes of a
+normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$ by
+(\ref{eq:maxrefsminaccs}), the notation of minimal hitting sets also carries
+over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe
+that
 \begin{equation}
@@
 % ------------------------------------------------------------------------
-\subsubsection*{Characterisation of $conf$ by Minimal Hitting Sets}
-The following lemma establishes that the $conf$ relation specified in
-(\ref{eq:conf}) can be characterised by means of minimal acceptances and
-their minimal hitting sets.
+\paragraph{Characterisation of $conf$ by Minimal Hitting Sets} The following lemma
+establishes that the $conf$ relation specified in (\ref{eq:conf}) can be
+characterised by means of minimal acceptances and their minimal hitting sets.
 %
@@
 $\minaccs(P/s)$ and $\minaccs(G(P)/s)$, as well as $\minhits(P/s)$ and
-$\minhits(G(P)/s)$ are used interchangeably.
-
-For showing ``$1 \Rightarrow 2$'', we assume $Q\ conf\ P$ and
-$s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}),
-states that
-\[
+$\minhits(G(P)/s)$ are used interchangeably. For showing ``$(1) \Rightarrow
+(2)$'', we assume $Q\ conf\ P$ and $s\in\trc(P)\cap \trc(Q)$.
+Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}), states that
+$%\[
 \forall A_Q\in\minaccs(G(Q)/s):
 \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
-\]
+$. %\]
 Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap A_P\neq\varnothing$
@@
 every minimal acceptance $A_Q$, because $A_P\subseteq A_Q$ for at least one
-$A_P$. As a consequence, each $H \in \minhits(P/s)$ is also a hitting set for
+$A_P$. So, each $H \in \minhits(P/s)$ is also a hitting set for
 $\minaccs(G(Q)/s)$ as required.
 
-To prove ``$2 \Rightarrow 1$'', assume that 2 holds, but that $P\ conf\ Q$
-does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
+To prove ``$(2) \Rightarrow (1)$'', we assume that (2) holds, but that $P\
+conf\ Q$ does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
 (\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
 \[
-\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
-\qquad (*)
+\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s):
+A_P\not\subseteq A_Q \qquad (*)
 \]
@@
 Define
-\[
+$%\[
 \overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
-\]
+$. %\]
 Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
@@
 
-In preparation for complexity results to be presented in
+In preparation for complexity results presented in
 Section~\ref{sec:complexity}, we consider how many minimal hitting sets can
-maximally exist for a given collection of minimal acceptances. To this end,
-the following definitions and results are useful.
+maximally exist for a collection of minimal acceptances. To this end, the
+following definitions and results are useful.
 
-A \emph{Sperner Family} is a collection ${\cal S} \subseteq 2^\Sigma$ of sets from a
-given finite universe $\Sigma$ that do not contain each other, that is, $H_1
-\not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair $H_1\neq
-H_2\in {\cal S}$. Specialising antichains known from partial orders to finite sets
-partially ordered by $\subseteq$ results in Sperner families.
+A \emph{Sperner Family} is a collection ${\cal S} \subseteq 2^\Sigma$ of sets
+from a given finite universe $\Sigma$ that do not contain each other, that
+is, $H_1 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair
+$H_1\neq H_2\in {\cal S}$. Specialising antichains known from partial orders
+to finite sets partially ordered by $\subseteq$ results in Sperner families.
+Given an {\it arbitrary} collection of subsets ${\cal C}\subset 2^\Sigma$,
+the sub-collection $\minsub({\cal C})$ defined in (\ref{eq:minsub}) is a
+Sperner Family contained in ${\cal C}$.
 
-Given an {\it arbitrary} collection of subsets ${\cal C}\subset 2^\Sigma$,
-the sub-collection $\minsub({\cal C})$ defined in (\ref{eq:minsub})
-is a Sperner Family contained
-in ${\cal C}$.
+\newpage
+\noindent
 We further observe that
@@
 \begin{theorem}[Sperner's Theorem~\cite{sperner_satz_1928}]
-\label{th:sperner}
-Given a  Sperner family ${\cal S}$
-over an $n$-element universe $\Sigma$, its cardinality is bound by
+\label{th:sperner} Given a  Sperner family ${\cal S}$ over an $n$-element
+universe $\Sigma$, its cardinality $\card{{\cal S}}$ is bound by
 \[
@@
 It is shown in Section~\ref{sec:complexity} that this upper bound can
-actually be reached by hitting sets associated with the minimal acceptances
-of a CSP process state.
+actually be reached by the Sperner Family containing the hitting sets
+associated with the minimal acceptances of a CSP process state.

### 2018-11-19 scp-journal/article/finitecomplete-failures.tex
Revised Section 3

@@
 transition function of $G(P)$ and $e\in [G(P)/s]^0$, $t(G(P)/s,e)$ is
-defined, and $t(G(P)/s,e) = G(P)/s.e$. This leads to a new recursion, so that
-$U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = U_F(j,\#s+1,G(P)/s.e)$ as required.
+defined, and $t(G(P)/s,e) = G(P)/s.e$. So, the new recursion in branch
+(\ref{eq:ufc}) is so that $U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
+U_F(j,\#s+1,G(P)/s.e)$ as required.
 
@@
 $U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
-specification of this branch implies that $_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
+specification of this branch implies that $U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
 (\efail\then\Stop)$.
@@
 
-\paragraph{Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and $Q\
+\paragraph{Step~1} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and $Q\
 conf\ P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq
@@
 
-\paragraph{Step~2.} For proving that a test execution can never deadlock without a
+\paragraph{Step~2} For proving that a test execution can never deadlock without a
 $\epass$ event, it has to be shown that a test execution can neither block at

### 2018-11-19 scp-journal/article/preliminaries.tex
Revised Section 3

@@
   \\ %
-  & Property of $r_P, r_Q$ and (\ref{eq:refMaxRef})
+  & Property of $r_P, r_Q$~(subset closure) and (\ref{eq:refMaxRef})
 \end{argue}
@@
 $%\[ \
-minhits(P/s) = \minhits(\minaccs(P/s))
+\minhits(P/s) = \minhits(\minaccs(P/s))
 $ %\]
@@
 \begin{argue}
-(s,A)\not\in {\text{failures}}(P)
+(s,A)\not\in \fails(P)
 \\
-\Leftrightarrow A\not\in {\text{Ref}}(P/s)
+\Leftrightarrow A\not\in Ref(P/s)
 \\
-\Leftrightarrow  \forall R\in {\text{maxRef}}(P/s): A\not\subseteq R
+\Leftrightarrow  \forall R\in maxRef(P/s): A\not\subseteq R
 \\
-\Leftrightarrow   \forall B\in {\text{minAcc}}(P/s): A\not\subseteq (\Sigma\setminus B)& this follows from (\ref{eq:accref})
+\Leftrightarrow   \forall B\in minAcc(P/s): A\not\subseteq (\Sigma\setminus
+B)& this follows from (\ref{eq:accref})
 \\
-\Leftrightarrow  \forall B\in {\text{minAcc}}(P/s): A\cap B\neq \varnothing
+\Leftrightarrow  \forall B\in minAcc(P/s): A\cap B\neq \varnothing
 \\
 \Leftrightarrow  A \text{ is a (not necessarily minimal)}
-              \text{ hitting set of } {\text{minAcc}}(P/s)
+              \text{ hitting set of } minAcc(P/s)
 \end{argue}
@@
 \overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
-$. %\]
+$ %\]
 Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),

### 2018-11-19 scp-journal/article/finitecomplete-failures.tex
Section 4 revised

@@
 \paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
-after having run through trace $s$ with $\#s < j$. Since  $P\lessdet_T Q$ by
-assumption, $s$ is a trace of $P$, thus $U_F(j)/s = U_F(j,\#s,G(P)/s)$
+after having run through a trace $s$ with $\#s < j$. Since  $P\lessdet_T Q$
+by assumption, $s$ is a trace of $P$, thus $U_F(j)/s = U_F(j,\#s,G(P)/s)$
 according to Lemma~\ref{lemma:ufproperties}. Therefore, $U_F(j)/s$  can enter
@@
 {\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least one
-$e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that
-$\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be
-entered, and the test stops after $\epass$. Otherwise, if $\minhits(G(P)/s)
-\neq \varnothing$, let $H\in\minhits(G(P)/s)$. Since $H$ contains only
-elements that are contained in some minimal acceptance of $P/s$, and all
-these minimal acceptances are subsets of $[G(P)/s]^0$, $H$ is a subset of
-$[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, is not a hitting set of
-$\minaccs(Q/s)$, also $H$ is not a hitting set of $\minaccs(Q/s)$. Now this
-is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by
-assumption, so $H$ should also be a (not necessarily minimal) hitting set in
-$\minaccs(Q/s)$. This proves that the test execution cannot block  at branch
-(\ref{eq:ufc}) without being able to pass the test by entering branch
-(\ref{eq:ufb}).
+$e\in [G(P)/s]^0$ would be accepted by $Q/s$ and the test execution would not
+deadlock. Now suppose that $\minhits(G(P)/s) = \varnothing$. Then   branch
+(\ref{eq:ufb}) can be entered, and the test stops after $\epass$. Otherwise,
+if $\minhits(G(P)/s) \neq \varnothing$, let $H\in\minhits(G(P)/s)$. Since $H$
+contains only elements that are contained in some minimal acceptance of
+$P/s$, and all these minimal acceptances are subsets of $[G(P)/s]^0$, $H$ is
+a subset of $[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, is not a
+hitting set of $\minaccs(Q/s)$, also $H$ is not a hitting set of
+$\minaccs(Q/s)$. Now this is a contradiction to Lemma~\ref{lemma:hseta},
+since $Q\ conf\ P$ by assumption, so $H$ should also be a (not necessarily
+minimal) hitting set in $\minaccs(Q/s)$. This proves that the test execution
+cannot block  at branch (\ref{eq:ufc}) without being able to pass the test by
+entering branch (\ref{eq:ufb}).
 
-\paragraph{Step~2.2} Now suppose that the test execution blocks at
-branch (\ref{eq:ufd}) after having run through some $s\in\trc(Q)\subseteq \trc(P)$
-with $\#s = j$.
-From Lemma~\ref{lemma:ufproperties} we know
-that $U_F(j)/s = U_F(j,k,n)$ with $U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so
-$\minhits(n) = \minhits(P/s)$. The branch (\ref{eq:ufb}) of the test
-$U_F(j,k,n)$ leads always to a PASS verdict and is taken if $\minhits(n) =
-\varnothing$.
-If $\minhits(n) \neq \varnothing$,  the assumption that $(Q\parallel[\Sigma] U_F(j))/s$
-blocks at branch (\ref{eq:ufd}) implies that there exists some $H\in\minhits(n)$
-which is not a hitting set of $\minaccs(Q/s)$. Again, we can apply Lemma~\ref{lemma:hseta}
-to conclude that this contradicts the assumption that $Q\ conf\ P$. As a consequence,
-the test execution can never deadlock at branch (\ref{eq:ufd}) without entering
-branch (\ref{eq:ufb}) and passing the test.
+\paragraph{Step~2.2} Suppose that the execution blocks at
+branch (\ref{eq:ufd}) after having run through some $s\in\trc(Q)\subseteq
+\trc(P)$ with $\#s = j$. From Lemma~\ref{lemma:ufproperties} we know that
+$U_F(j)/s = U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so $\minhits(n) =
+\minhits(P/s)$. Branch (\ref{eq:ufb}) of $U_F(j,k,n)$ leads always to a PASS
+verdict and is taken if $\minhits(n) = \varnothing$. If $\minhits(n) \neq
+\varnothing$,  the assumption that $(Q\parallel[\Sigma] U_F(j))/s$ blocks at
+branch (\ref{eq:ufd}) implies that there exists some $H\in\minhits(n)$ that
+is not a hitting set of $\minaccs(Q/s)$. Again, by Lemma~\ref{lemma:hseta},
+this contradicts the assumption that $Q\ conf\ P$. As a
+consequence, the test execution can never deadlock at branch (\ref{eq:ufd})
+without entering branch (\ref{eq:ufb}) and passing the test.
 
+\bigskip \noindent%
 Note that the line of reasoning in this proof requires that $Q$ is free of
-livelocks, because otherwise a $\epass$ event might not become visible,
-due to unbounded sequences of hidden events performed by $Q$.
-\xbox
+livelocks, because otherwise a $\epass$ event might not become visible, due
+to unbounded sequences of hidden events performed by $Q$. \xbox
 \end{proof}
@@
 \item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
-\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
-of $\minaccs(Q/s)$, such that
-(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
+\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a
+    minimal acceptance $A_Q$ of $\minaccs(Q/s)$, such that (see
+    Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}))
 \begin{equation}
@@
 through this product graph. Any error can be detected after running through
-such a trace and then either observing an event outside $[P/s]^0$ (this is
-the violation described by Case~1) or identifying an illegal acceptance $A_Q$
-(as specified in Case~2). It is not guaranteed, however, that trace $s$ is
-short enough to be executed by one of the test cases $U_F(j)$ with $0\le j <
-pq$. So, it has to be shown that for any $s$ leading to an error situation
-specified by Case~1 or Case~2, there exists a trace $u$ of maximal length
-$pq-1$ leading to the same error.
+such a trace and then either observing an event outside $[P/s]^0$ (the
+violation described by Case~1) or identifying an illegal acceptance $A_Q$ (as
+in Case~2). It is not guaranteed, however, that $s$ is short enough to be
+executed by one of the test cases $U_F(j)$ with $0\le j < pq$. So, it has to
+be shown that for any $s$ leading to an error situation, there exists a trace
+$u$ of maximal length $pq-1$ leading to the same error.
 
 \medskip
-\noindent {\bf Case~1.}
-Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
+\noindent {\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
 but $s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is
@@
 \medskip
-\noindent {\bf Case~2.} We note that trace $s$ is again a trace of the
-product graph $G$, but we do not know its length. Again, by applying
-Lemma~\ref{lemma:reachproduc}, we know that the state $G/s$ can be reached by
-a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider
-the test $U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because
-of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$  always performs branch
+\noindent {\bf Case~2.} We note again that $s$ is a trace of the product
+graph $G$, but we do not know its length. Again, by
+Lemma~\ref{lemma:reachproduc}, the state $G/s$ can be reached by a trace
+$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider the test
+$U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of
+Lemma~\ref{lemma:ufproperties}. $U_F(\#u)$ can always perform branch
 (\ref{eq:ufc}) until the trace $u$ has been completely processed.
@@
 %
-Our notion of tests can be specialised to deal with traces refinement, as we
-explain in Section~\ref{sec:finitecomplete}.
-The next section presents and example.
+Our notion of test can be specialised to deal with traces refinement~(see
+Section~\ref{sec:finitecomplete}). We next present an example.
 

### 2018-11-20 scp-journal/article/finitecomplete-trace.tex
Section 5 started

--- 2018-10-07 paper/finitecomplete-trace.tex
@@
 For establishing trace refinement, the following class of adaptive test cases
-will be used for a given reference process $P$ and  integers $j \ge 0$.
-Just as for the tests developed  in
-Section~\ref{sec:finitecompletefails} to verify failures refinement,
-the tests for trace refinement are
-derived from the reference model's transition graph
+are used for a given reference process $P$ and  integers $j \ge 0$. Just as
+for the tests developed  in Section~\ref{sec:finitecompletefails} to verify
+failures refinement, the tests for trace refinement are derived from the
+reference model's transition graph
 $$
@@
 do not need to check the SUT with respect to its acceptance of hitting sets.
-Therefore, these do not occur in the specification of the test cases below,
-and we use the condition $\minaccs(n) = \{ \varnothing \}$ instead of
-$\minhits(n) = \varnothing$ in branch (\ref{eq:utb})
-to indicate that minimal hitting sets need not to
-be calculated for generating these tests from $G(P)$. From
-(\ref{eq:minhitminaccempty}) we know that these conditions are equivalent.
+Therefore, these do not occur in the specification of the test cases below.
+We use the condition on acceptances $\minaccs(n) = \{ \varnothing \}$ instead
+of the condition on hitting sets $\minhits(n) = \varnothing$ in branch
+(\ref{eq:utb}). From (\ref{eq:minhitminaccempty}) we know that these
+conditions are equivalent, but, with the use of $\minaccs(n) = \{ \varnothing
+\}$, we make it unnecessary to calculate hitting sets for generating these
+tests from $G(P)$.
 
@@
 % ==========================================================================
-Since $U_T(j)$ never blocks any event of $Q$ before terminating, the pass
-criterion can be based on trace instead of failures refinement as
-required in (\ref{eq:passF}).
+Since the test $U_T(j)$ never blocks any event of an SUT process $Q$ before
+terminating, the pass criterion, defined below, can be based on
+trace instead of failures refinement as required in (\ref{eq:passF}).
 %
@@
 %
-If the SUT process $Q$ deadlocks after a trace $s$ where also the reference
-process $P$ is in a state where deadlock is possible, this is captured by the
-fact that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$. Therefore,
-branch (\ref{eq:utb}) of a test case execution state $U_T(j,k,n)$ with $\#s =
-k \le j$   can be entered and the  test execution terminates with $\epass$.
-If, however, $Q$ blocks after a trace $s'$ where the reference process
-satisfies $\minaccs(P/s') \neq\varnothing$, branch (\ref{eq:utb})  cannot be
-taken, and the test execution stops without producing $\epass$ or $\efail$.
-In contrast to the test for failures refinement, this is   interpreted here
-as a successful test execution, because unexpected blocking of the SUT does
-not violate the trace refinement relation, as long as all traces executed by
-the SUT are traces of the reference process. In particular, if neither
-$\epass$ nor $\efail$ is ever produced, so that $(Q\parallel[\Sigma] U_T(j))
-\hide \Sigma = \Stop$, the test passes, because $(\epass\then\Stop)
-\lessdet_T \Stop$ holds.
+If the SUT process $Q$ deadlocks after a trace $s$, and in this case the
+reference process $P$ is also in a state where deadlock is possible, this is
+captured by the fact that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$.
+Therefore, branch (\ref{eq:utb}) of a test case execution state $U_T(j,k,n)$
+with $\#s = k \le j$   can be entered and the  test execution terminates with
+$\epass$. If, however, $Q$ blocks after a trace $s'$ and the reference
+process satisfies $\minaccs(P/s') \neq\varnothing$, branch (\ref{eq:utb})
+cannot be taken, and the test execution stops without producing $\epass$ or
+$\efail$. In contrast to the test for failures refinement, this is
+interpreted here as a successful test execution, because unexpected blocking
+of the SUT does not violate the trace-refinement relation, as long as all
+traces executed by the SUT are traces of the reference process. In
+particular, if neither $\epass$ nor $\efail$ is ever produced, so that
+$(Q\parallel[\Sigma] U_T(j)) \hide \Sigma = \Stop$, the test passes, because
+$(\epass\then\Stop) \lessdet_T \Stop$ holds.
 
@@
 %
-As before in the context of Theorem~\ref{th:failurestest},
-the proof of Theorem~\ref{th:tracetest} is structured into two lemmas, the first
-establishing soundness, the second exhaustiveness.
+As for Theorem~\ref{th:failurestest}, the proof is structured in two lemmas,
+the first ensuring soundness, and the second exhaustiveness.
 
@@
 \begin{proof}
-Suppose that $P\lessdet_T Q$, so that $\trc(Q)\subseteq\trc(P)$, and assume that
-$s\in\trc(Q)$ with $\#s < pq$. Since $s$ is also a trace of $P$, we can conlude
+Suppose that $P\lessdet_T Q$, so that $\trc(Q)\subseteq\trc(P)$, and assume
+that $s\in\trc(Q)$ with $\#s < pq$. Since $s$ is also a trace of $P$, we can
+conclude
 $$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$ because of (\ref{eq:ifpaT}). Now $\trc(Q)\subseteq\trc(P)$

### 2018-11-22 scp-journal/article/complexity.tex
More simplifications in the complexity formulas
Complete version - fit for submission to Science of Computer Programming.
Section 6 started

@@
 
+We consider first failures
+refinement~(Section~\ref{section:complexity:failures}) and then traces
+refinement~(Section~\ref{section:complexity:traces}) testing.
+
 % -------------------------------------------------------------------------
 \subsection{Estimates for the Maximal Number of Failures Test Executions}
+\label{section:complexity:failures}
 
 % -----------------------------------------------------------------------------
-\subsubsection*{A Reference Process}
-Given an alphabet $\Sigma$ of size $\card{\Sigma} = n\ge 2$, define
-a collection of subsets of $\Sigma$ by
+\paragraph{A Reference Process} Given an alphabet $\Sigma$ of size $\card{\Sigma} =
+n\ge 2$, define a collection of subsets of $\Sigma$ by
 \begin{equation}\label{eq:defC}
@@
    \\
-& = & \binom{n}{\lfloor n/2\rfloor}\cdot\big(1 + \frac{n^{pq}-n}{n-1} \big)  + 
-\sum_{j=1}^{pq-1} \frac{n^j - 1}{n-1}
+& = &    \sum_{j=1}^{pq-1} \frac{n^j - 1}{n-1} +
+\sum_{j=0}^{pq-1}\binom{n}{\lfloor n/2\rfloor}\cdot n^j
 \\
-& = &\binom{n}{\lfloor n/2\rfloor}\cdot\big(\frac{n^{pq}-1}{n-1} \big)  +
-\frac{n^{pq}-n pq+pq-1}{(n-1)^2}
+& = & \frac{n^{pq}-n pq+pq-1}{(n-1)^2} +
+\binom{n}{\lfloor n/2\rfloor} \cdot \frac{n^{pq} - 1}{n-1}
+\\
+& = &
 \\
@@
 %
-\begin{equation*}\label{eq:pmaxcomplexity} 
+\begin{equation*}
 \binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}.
@@
 \subsection{Estimates for the Maximal Number of Trace Test Executions}
+\label{section:complexity:traces}
 

### 2018-11-22 scp-journal/article/finitecomplete-trace.tex
Section 6 started

@@
 conditions are equivalent, but, with the use of $\minaccs(n) = \{ \varnothing
-\}$, we make it unnecessary to calculate hitting sets for generating these
-tests from $G(P)$.
+\}$, \pagebreak we make it unnecessary to calculate hitting sets for
+generating these tests from $G(P)$, which is expensive.
 
@@
 conclude
-$$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$ because of (\ref{eq:ifpaT}). Now $\trc(Q)\subseteq\trc(P)$
-implies $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$
-cannot enter branch (\ref{eq:uta}) and produce a $\efail$ event when running
-in parallel with $Q$ and synchronising over $\Sigma$. Therefore, only four
+$$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$
+because of (\ref{eq:ifpaT}). \pagebreak Now $\trc(Q)\subseteq\trc(P)$ implies
+$[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot
+enter branch (\ref{eq:uta}) and produce a $\efail$ event when running in
+parallel with $Q$ and synchronising over $\Sigma$. Therefore, only four
 options are available for the test execution    $(Q\parallel[\Sigma]
@@
 \noindent {\bf Case~4.} $\#s = pq-1$ holds. In this case,
-$U_T(pq-1,\#s,G(P)/s)$ can only enter branch (\ref{eq:utd}), and the test
+$U_T(pq-1,\#s,G(P)/s)$ can enter branch (\ref{eq:utd}), and the test
 execution stops after $\epass$.
 
-This case analysis shows that every   execution of $(Q\parallel[\Sigma] U_T(j))$
-either stops after $\epass$ or produces neither $\epass$ nor $\efail$. This
-proves that $Q$ passes test $U_T(pq-1)$ according to the pass  criterion (\ref{eq:passT}).
-\xbox
+\bigskip\noindent%
+This case analysis shows that every   execution of $(Q\parallel[\Sigma]
+U_T(j))$ either stops after $\epass$ or produces neither $\epass$ nor
+$\efail$. This proves that $Q$ passes test $U_T(pq-1)$ according to the pass
+criterion (\ref{eq:passT}). \xbox
 \end{proof}
@@
 \begin{proof}
-As before in the proofs for failures testing, we construct the product graph
+As in the proof for failures testing, we construct the product graph
 $G=G(P)\times G(Q)$ and recall that every trace $s\in\trc(P)\cap\trc(Q)$ is
@@
 that $G/s = (G(P)/s,G(Q)/s)$. Furthermore, we recall from
-Lemma~\ref{lemma:extendV} that graph state $(G(P)/s,G(Q)/s)$ can always be
-reached by a trace $u$ of length less or equal $pq-1$, where  the order of
+Lemma~\ref{lemma:extendV} that the graph state $(G(P)/s,G(Q)/s)$ can always
+be reached by a trace $u$ of length less or equal $pq-1$, where  the order of
 $G(P)$ is $p$ and that of $G(Q)$ is $q$.
@@
 By assumption, $e\in (\Sigma -[P/s]^0) = (\Sigma - [G(P)/s]^0)$. Since
-$G(Q)/u = G(Q)/s$, $Q/u$ can engage into $e$. Then $U_T(pq-1,\#u,G(P)/s)$
-will enter branch (\ref{eq:uta}), and the test execution stops after having
+$G(Q)/u = G(Q)/s$, $Q/u$ can engage into $e$. Then $U_T(pq-1,\#u,G(P)/s)$ can
+enter branch (\ref{eq:uta}), and the test execution stops after having
 produced $\efail$. This proves that $Q$  fails test $U_T(pq-1)$. \xbox
@@
 %
-Having established completeness of our test suites, we now consider the
+Having established completeness of our test suites, we consider
 complexity of a testing technique that uses them.

### 2018-11-24 scp-journal/article/complexity.tex
Section 6 started

@@
 
-We consider first failures
-refinement~(Section~\ref{section:complexity:failures}) and then traces
-refinement~(Section~\ref{section:complexity:traces}) testing.
+Since we have finite complete CSP test suites, it is useful for the first
+time to calculate how many test executions are needed when using them.
+Previous work did not consider sufficient conditions for finiteness, so
+complexity was not a concern. We answer the following questions. (1)~What is
+the worst-case bound on the number of test executions to be performed to
+verify an SUT with respect to failures refinement, when we use our test
+suite? (2)~What is the worst-case bound for trace refinement? (3)~Is it
+possible to reduce the maximal length of traces when testing for failures or
+trace refinement? We consider the first question~(1) in
+Section~\ref{section:complexity:failures}), where we also discuss whether it
+it is possible to reduce the number of test executions with a different test
+suite.  With the answer to question~(1), question~(2) is a fairly simple
+consequence we discuss in Section~\ref{section:complexity:traces}.
+Question~(3) is the subject of Section~\ref{section:complexity:length}.
 
@@
 \label{section:complexity:failures}
+
+An arbitrary CSP process $P$ might have $\minhits(P/s) = \varnothing$ for
+some traces $s$, so that a test case $U_F(j)$ for a $j$ greater than the size
+of $s$ can enter branch~(\ref{eq:ufb}). In this case, further executions are
+needed to consider traces that have $s$ as a prefix.  To provide a bound on
+the number of test executions needed, we first define a process $\pmax$~(see
+(\ref{eq:pmax})), which, when used as a reference process, requires the
+maximal number of test executions among all reference processes $P$
+fulfilling $\minhits(P/s) != \varnothing$ for all traces $s$. For $\pmax$, we
+can establish the actual number of test executions required~(see
+(\ref{eq:pmaxcomplexity})).
 
@@
 \begin{proof}
- Since
-$
-\bigcup_{A\in{\cal C}} A = \Sigma
-$
-by construction of ${\cal C}$, $[\pmax]^0 = \Sigma$. Since $\pmax/e = \pmax$ for all
-$e\in\Sigma$, this proves statements (\ref{eq:pmaxa}) and (\ref{eq:pmaxb}).
-The internal choice construct used in the specification of $\pmax$ implies
-$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all
-$e\in\Sigma$ implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence
-of (\ref{eq:pmaxc}).
-
-Let $H$ be any minimal hitting set of ${\cal C}$. Then
-$H$ contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because
-otherwise $\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, and any
-subset $A\subseteq \Sigma\setminus H$ with cardinality
-$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in ${\cal C}$, but satisfy
-$A\cap H=\varnothing$. Since
+Since $\bigcup_{A\in{\cal C}} A = \Sigma $ by construction of ${\cal C}$,
+$[\pmax]^0 = \Sigma$ and stated by (\ref{eq:pmaxa}). Since $\pmax/e = \pmax$
+for all $e\in\Sigma$, this proves statement (\ref{eq:pmaxb}). The internal
+choice construct used in the specification of $\pmax$ implies
+$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all $e\in\Sigma$
+implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this
+shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence of
+(\ref{eq:pmaxc}). Let $H$ be any minimal hitting set of ${\cal C}$. Then $H$
+contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because otherwise
+$\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, and any subset
+$A\subseteq \Sigma\setminus H$ with cardinality
+$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in ${\cal C}$, but
+satisfy $A\cap H=\varnothing$. Since
 $\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we
-conclude
- that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
- intersects  every element of ${\cal C}$.  Therefore, every minimal hitting set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements; this shows (\ref{eq:pmaxf})
-and $\card{\minhits({\cal C})} = \binom{n}{\lfloor{\frac{n}{2}}\rfloor}$. The latter
-shows  (\ref{eq:pmaxd}) and completes the proof.
-\xbox
+conclude that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
+intersects  every element of ${\cal C}$.  Therefore, every minimal hitting
+set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements; this
+shows (\ref{eq:pmaxf}) and $\card{\minhits({\cal C})} =
+\binom{n}{\lfloor{\frac{n}{2}}\rfloor}$. The latter shows  (\ref{eq:pmaxd})
+and completes the proof. \xbox
 \end{proof}
@@
 % -----------------------------------------------------------------------------
-\subsubsection*{Test Cases of $\pmax$}
-The test cases $U_F(j)$ generated from $\pmax$
-can never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$ for
-all traces $s\in\trc(\pmax)$ according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxa}).
-Moreover, they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$
-is never empty according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxd}).
-Finally, the minimal hitting sets used to probe the SUT at the end of a
-non-blocking  test execution are always the hitting sets of ${\cal C}$
-according to Lemma~\ref{lemma:pmax}, (\ref{eq:pmaxe}).
-This results in a test case structure
+\paragraph{Test Cases of $\pmax$} The test cases $U_F(j)$ generated from $\pmax$ can
+never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$
+for all traces $s\in\trc(\pmax)$ according to (\ref{eq:pmaxa}). Moreover,
+they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$ is
+never empty according to (\ref{eq:pmaxd}). Finally, the minimal hitting sets
+used to probe the SUT at the end of a non-blocking test execution are always
+the hitting sets of ${\cal C}$ according to (\ref{eq:pmaxe}). This results in
+the following test case structure.
 \begin{eqnarray*}
@@
 % -----------------------------------------------------------------------------
-\subsubsection*{Maximal Number of Test Executions for $\pmax$}
-
-When considering the number of test executions to be performed by
-$U_F(j)$ against some SUT $Q$ for all $j=0,\dots,pq-1$, the maximal number of
-test executions is only reached if $Q$ is a correct refinement of the reference
+\paragraph{Maximal Number of Test Executions for $\pmax$}
+When considering the number of test executions to be performed by $U_F(j)$
+against some SUT $Q$ for all $j=0,\dots,pq-1$, the maximal number of test
+executions is only reached if $Q$ is a correct refinement of the reference
 process or if it fails in the very last execution of the very last $U_F(j)$
-executed against $Q$. If an erroneous behaviour of $Q$ is revealed before this
-very last execution, the test suite is stopped and $Q$ is fixed before
-executing the suite again.
-Therefore, we consider only correct SUTs $Q$ when determining the
-maximal number of executions that can be executed.
+executed against $Q$. If an erroneous behaviour of $Q$ is revealed before
+this very last execution, the test experiment is stopped (and $Q$ can be
+fixed before executing the suite again). Therefore, we consider only correct
+SUTs $Q$ when determining the maximal number of executions that can be
+executed.
 
@@
 \subsection{Upper Bound $pq$ for the Maximal Length of Test Traces}
+\label{section:complexity:length}
 

### 2018-11-26 scp-journal/article/complexity.tex
Considered message of 25.11

@@
 \end{eqnarray*}
+This means that the branches of $U_F(j)$ that can lead to an early
+termination are not feasible. All tests deadlock, or run to the end of a
+trace of size $j$ and then present the choice of event of a minimal hitting
+set.
 
@@
 \paragraph{Maximal Number of Test Executions for $\pmax$}
-When considering the number of test executions to be performed by $U_F(j)$
-against some SUT $Q$ for all $j=0,\dots,pq-1$, the maximal number of test
-executions is only reached if $Q$ is a correct refinement of the reference
-process or if it fails in the very last execution of the very last $U_F(j)$
-executed against $Q$. If an erroneous behaviour of $Q$ is revealed before
-this very last execution, the test experiment is stopped (and $Q$ can be
-fixed before executing the suite again). Therefore, we consider only correct
-SUTs $Q$ when determining the maximal number of executions that can be
-executed.
-
-For the  $U_F(j)$ generated from $\pmax$, the number of test executions
- to be performed is maximal if $\pmax\lessdet_F Q$ and
-$\trc(Q) = \Sigma^*$. In such a situation, no test execution blocks early, because
-$Q/s$ can always engage into some $e\in\Sigma$   while $\#s<j$ and will never block
-in the last step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is
-offered by the test case. The resulting number of executions is
+When considering the number of test executions to be performed using $U_F(j)$
+derived from $\pmax$ against some SUT $Q$ for all $j = 0,\dots,pq-1$, and
+considering that we need to cover all the possible behaviours of $Q$, the
+maximal number of test executions is only reached if (a)~$Q$ is a correct
+refinement of $\pmax$ (or more generally, of the reference process) and
+(b)~its traces are $\Sigma^*$.
+%or if it fails in the very last execution of the very last $U_F(j)$
+%executed against $Q$. If an erroneous behaviour of $Q$ is revealed before
+%this very last execution, the test experiment is stopped (and $Q$ can be
+%fixed before executing the suite again). Therefore, we consider only correct
+%SUTs $Q$ when determining the maximal number of executions that can be
+%executed.
+%
+%For the  $U_F(j)$ generated from $\pmax$, the number of test executions
+% to be performed is maximal if $\pmax\lessdet_F Q$ and
+%$\trc(Q) = \Sigma^*$.
+In such a situation, no test execution blocks early, because $Q/s$ can always
+engage into some $e\in\Sigma$  while $\#s<j$, and never blocks in the last
+step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is offered by
+the test case. The resulting number of executions in this case is
 %
@@
 %
-because all traces up to length $j$ can be executed with $U_F(j)$ entering branch
-(\ref{eq:ufc}), and each of these traces is followed by one event
-from each of the hitting sets of ${\cal C}$.
+because \pagebreak all traces up to length $j$ can be executed with $U_F(j)$
+entering branch (\ref{eq:ufc}), and each of these traces is followed by one
+event from each of the hitting sets of ${\cal C}$ since $Q$ is correct.  %(b) leads to the maximal number of
+%executions possible, because it is responsible for the factor *n^{(pq-1)} in
+%the upper bound (47), (48) [n = |Sigma|]. Considering a reference process
+%with fewer traces or an implementation Q with fewer traces would
+%significantly reduce the value of this factor, while just adding one
+%additional execution possibility for branch (20), after which the test suite
+%is stopped due to failure.
 
-The number of executions specified in (\ref{eq:maxexec}) is indeed maximal for
-all reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
-traces $s$: All these processes can never enter branch (\ref{eq:ufb}), and,
-if an execution with an SUT entered branch (\ref{eq:ufa}), this would only lead to early termination of the whole test suite, because a failure has been detected.
-As a consequence, $\card{\Sigma}^{j}$ is the maximal number of traces to be executed
-up to length $j$, and from Theorem~\ref{th:sperner} we know that the number
-$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets to be tested at the end of each trace of length $j$ is already maximal.
+The number of executions in (\ref{eq:maxexec}) is indeed maximal for all
+reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
+traces $s$. All these processes can never enter branch (\ref{eq:ufb}), and,
+if an execution with an SUT entered branch (\ref{eq:ufa}), this would only
+lead to early termination of the whole test suite, because a failure has been
+detected. As a consequence, $\card{\Sigma}^{j}$ is the maximal number of
+traces to be executed up to length $j$, and from Theorem~\ref{th:sperner} we
+know that the number $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets
+to be tested at the end of each trace of length $j$ is already maximal.
 
-Summing up  formula
-(\ref{eq:maxexec}) over all test cases $U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum of a geometric progression,  this results in
+Summing up  formula (\ref{eq:maxexec}) over all test cases
+$U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum
+of a geometric progression,  this results int
 %
@@
 %
-as the maximal number of test executions to be performed when testing an error-free
-SUT $Q$ against reference process $\pmax$. This maximum is met when $\trc(Q) = \Sigma^*$.
-If one is only interested in the order of magnitude, the maximal nmber of executions
-may be re-written as
+as the maximal number of test executions to be performed when testing an
+error-free SUT $Q$ with $\trc(Q) = \Sigma^*$ against the reference process
+$\pmax$. If we are interested only in the order of magnitude, the maximal
+number of executions may be given as
 %
@@
 % -----------------------------------------------------------------------------
-\subsubsection*{Considering Empty Collections of Minimal Hitting Sets}
-
-The argument so far has shown that reference process $\pmax$ requires the most test
-executions among all CSP processes $P$ whose collections of minimal hitting sets
-$\minhits(P/s)$ are never empty for any trace $s$, so that the branch
+\paragraph{Considering Empty Collections of Minimal Hitting Sets}
+The argument so far has shown that the tests derived from the reference
+process $\pmax$ require the most test executions when compared to tests
+derived from any other CSP process $P$ whose collections of minimal hitting
+sets $\minhits(P/s)$ are never empty for any trace $s$, so that the branch
 (\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.
 Lemma~\ref{lemma:failureshittingsets} implies that an empty collection
-$\minhits(P/s)$
-is equivalent to $(s,A)\in\fails(P)$ for all $A\subseteq\Sigma$. This is equivalent to
-$(s,\Sigma)\in\fails(P)$ or
-$\maxrefs(P/s) = \{ \Sigma \}$ or $\minaccs(P/s) = \{ \varnothing\}$.
+$\minhits(P/s)$ is equivalent to $(s,A)\in\fails(P)$ for all
+$A\subseteq\Sigma$. This is equivalent to $\maxrefs(P/s) = \{ \Sigma \}$ or
+$\minaccs(P/s) = \{ \varnothing\}$.
 
-It remains to
-consider the question whether reference processes $Z$ possessing failures $(s,\Sigma)$
-may require more test executions for their associated tests $U_F(j)$ than the bound
-given for $\pmax$ in (\ref{eq:pmaxcomplexity}), because process states $Z/s$ with
-$\minhits(Z/s)=\varnothing$ allow for test executions entering branch (\ref{eq:ufb}).
-To this end, consider a test case $U_F(j)$ constructed from such a process $Z$. Every trace
-$s\in\trc(Z)$ with $\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$ allows for
+It remains to consider whether reference processes $Z$ possessing failures
+$(s,\Sigma)$ may require more test executions for their associated tests
+$U_F(j)$ than the bound given for $\pmax$ in (\ref{eq:pmaxcomplexity}),
+because process states $Z/s$ with $\minhits(Z/s)=\varnothing$ allow for test
+executions entering branch (\ref{eq:ufb}). To this end, consider a test case
+$U_F(j)$ constructed from such a process $Z$. Every trace $s\in\trc(Z)$ with
+$\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$
+allows for
 %

### 2018-11-26 scp-journal/article/complexity.tex
Considered message of 25.11

@@
 maximal number of test executions among all reference processes $P$
-fulfilling $\minhits(P/s) != \varnothing$ for all traces $s$. For $\pmax$, we
-can establish the actual number of test executions required~(see
+fulfilling $\minhits(P/s) \neq \varnothing$ for all traces $s$. For $\pmax$,
+we can establish the actual number of test executions required~(see
 (\ref{eq:pmaxcomplexity})).

### 2018-11-27 scp-journal/article/complexity.tex
Section 6.1 finished

@@
-\section{Complexity Considerations}
-\label{sec:complexity}
+i\section{Complexity Considerations} \label{sec:complexity}
 % ================================================================================
@@
 we can establish the actual number of test executions required~(see
-(\ref{eq:pmaxcomplexity})).
+(\ref{eq:pmaxcomplexity})). We then show that the order of magnitude of the
+worst-case bound for the number of test executions is the same also for
+reference processes $P$ that may have $\minhits(P/s) = \varnothing$ for some
+$s$.
 
@@
 termination are not feasible. All tests deadlock, or run to the end of a
-trace of size $j$ and then present the choice of event of a minimal hitting
+trace of size $j$ and then present a choice of events from a minimal hitting
 set.
+
+Moreover, Theorem~\ref{th:sperner} establishes that given an alphabet with
+$n$ elements, there is no Sperner family consisting of more than
+$\binom{n}{\lfloor\frac{n}{2}\rfloor}$. In addition, we recall the minimal
+hitting sets calculated from a given set of minimal acceptances are a Sperner
+family~(see discussion in Section~\ref{sec:sperner}). So, (\ref{eq:pmaxd})
+establishes that there is no possibility of providing more choices fom
+minimal hitting sets with a test derived from a process different from
+$\pmax$.
+
+In summary, the tests derived from $\pmax$ require the most test executions
+when compared to tests derived from any other CSP process $P$ whose
+collections of minimal hitting sets $\minhits(P/s)$ are never empty for any
+trace $s$.
 
@@
 %
-because \pagebreak all traces up to length $j$ can be executed with $U_F(j)$
+because all traces up to length $j$ can be executed with $U_F(j)$
 entering branch (\ref{eq:ufc}), and each of these traces is followed by one
@@
 error-free SUT $Q$ with $\trc(Q) = \Sigma^*$ against the reference process
-$\pmax$. If we are interested only in the order of magnitude, the maximal
-number of executions may be given as
+$\pmax$. If we are interested only in the order of magnitude, we have
 %
@@
 \paragraph{Considering Empty Collections of Minimal Hitting Sets}
-The argument so far has shown that the tests derived from the reference
-process $\pmax$ require the most test executions when compared to tests
-derived from any other CSP process $P$ whose collections of minimal hitting
-sets $\minhits(P/s)$ are never empty for any trace $s$, so that the branch
-(\ref{eq:ufb}) of the related test cases $U_F(j)$ can never be entered.
-Lemma~\ref{lemma:failureshittingsets} implies that an empty collection
-$\minhits(P/s)$ is equivalent to $(s,A)\in\fails(P)$ for all
-$A\subseteq\Sigma$. This is equivalent to $\maxrefs(P/s) = \{ \Sigma \}$ or
-$\minaccs(P/s) = \{ \varnothing\}$.
-
+The argument so far has shown that the tests derived from $\pmax$ require the
+most test executions when considering processes whose collections of minimal
+hitting sets are never empty. %Lemma~\ref{lemma:failureshittingsets} implies
+%that an empty collection $\minhits(P/s)$ is equivalent to $(s,A)\in\fails(P)$
+%for all $A\subseteq\Sigma$. This is equivalent to $\maxrefs(P/s) = \{ \Sigma
+%\}$ or $\minaccs(P/s) = \{ \varnothing\}$.
+%
 It remains to consider whether reference processes $Z$ possessing failures
@@
 %
-For a rough estimate of the worst case upper bound suppose that
+For a rough estimate of the worst-case upper bound suppose that
 %
@@
 %
-Note that this situation cannot be realised for all $j\in\{0,\dots,pq-1\}$, because
-the traces of
-$U_F(p-1)$ already cover all states of $Z$'s transition graph according to Lemma~\ref{lemma:extendV}, and if all states of $Z$ have empty hitting sets, there are no acceptance checks to be performed in the last step of the test execution.
-Therefore, the   upper bound calculated next cannot be reached by a real CSP process.
-With the 3 assumptions above, we calculate that
+We note that this scenario is not feasible for all $j\in\{0,\dots,pq-1\}$,
+because the traces of $U_F(p-1)$ already cover all states of $Z$'s transition
+graph according to Lemma~\ref{lemma:extendV}, and if all states of $Z$ have
+empty hitting sets, there are no acceptance checks to be performed in the
+last step of the test execution. Therefore, \pagebreak the upper bound
+calculated next cannot be reached by a CSP process. With the three
+assumptions above, nevertheless, we can calculate that
 %
@@
 \item  $U_F(0)$ has  $\binom{n}{\lfloor n/2\rfloor}$ executions,
-\item $U_F(j),\ j > 0$ has $\sum_{i=0}^{j-1} n^i = \frac{n^j - 1}{n-1}$ executions of branch (\ref{eq:ufb}), ($n = \card{\Sigma}$), and
-\item $U_F(j),\ j > 0$ has $\binom{n}{\lfloor n/2\rfloor}\cdot n^j$ executions where the acceptance of
-hitting sets is checked after having run through a trace of length $j$.
+\item $U_F(j)$, for $j > 0$ has $\sum_{i=0}^{j-1} n^i = \frac{n^j -
+    1}{n-1}$ executions of branch (\ref{eq:ufb}), ($n = \card{\Sigma}$),
+    and
+\item $U_F(j)$, for $j > 0$ has $\binom{n}{\lfloor n/2\rfloor}\cdot n^j$
+    executions where the acceptance of hitting sets is checked after having
+    run through a trace of length $j$.
 \end{itemize}
 %
-Summing up over all $U_F(j)$ for $j=0,\dots,pq-1$,  an upper bound $B$ for the number of test executions may be calculated as follows.
+Summing up over all $U_F(j)$ for $j=0,\dots,pq-1$, an upper bound $B$ may be calculated as follows.
 %
@@
 %
-Since  $B$ cannot be reached anyway,
-we just calculate its order of magnitude, and this results again in $O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}  \cdot n^{pq-1}\big)$, as calculated already for $\pmax$ in (\ref{eq:maxO}).
+Since $B$ cannot be reached anyway, we just calculate its order of magnitude,
+and this results again in $O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}  \cdot
+n^{pq-1}\big)$, as calculated already for $\pmax$ in (\ref{eq:maxO}).
 %
-Summarising these complexity calculations, this results in the following theorem.
+Summarising these complexity calculations, we have the following theorem.
 %
@@
 %
-
-%
 In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
@@
 estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
-offered to the SUT in the last step of the test execution. This number is significantly larger than the worst-case estimate $\binom{n}{\lfloor n/2\rfloor}$ calculated above for the
-hitting sets to be checked.
-In Fig.~\ref{fig:minhita}, the reduction is visualised by plots of
-the two functions.
-
-In~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors also use minimal hitting
-sets\footnote{However, they are denoted by {\it minimal acceptances}
-in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but they do not give an estimate for the number of test executions to be performed.
+offered to the SUT in the last step of the test execution. This number is
+significantly larger than the worst-case estimate $\binom{n}{\lfloor
+n/2\rfloor}$ above. In Fig.~\ref{fig:minhita}, the reduction is visualised by
+plots of the two functions. In~\cite{DBLP:conf/icfem/CavalcantiG07}, the
+authors also use minimal hitting sets\footnote{However, they are denoted by
+{\it minimal acceptances} in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but do
+not give an estimate for the number of test executions.
 
@@
  \begin{center}
-\includegraphics[width=.8\textwidth]{curvecomparison.pdf}
+\includegraphics[scale=.7%width=.8\textwidth
+                        ]{curvecomparison.pdf}
 \end{center}

### 2018-11-27 scp-journal/article/complexity.tex
Section 6.2 finished

@@
 suite.  With the answer to question~(1), question~(2) is a fairly simple
-consequence we discuss in Section~\ref{section:complexity:traces}.
-Question~(3) is the subject of Section~\ref{section:complexity:length}.
+consequence we discuss briefly also in
+Section~\ref{section:complexity:failures}. Question~(3) is the subject of
+Section~\ref{section:complexity:length}.
 
@@
 %
+As established by Lemma~\ref{lemma:hseta}, the number of hitting sets used to
+probe the SUT cannot be reduced:~if the reference process is $\pmax$, we need
+to consider all of them, otherwise illegal blocking may remain undetected. In
+addition, if $\pmax$ defines the behaviour of the SUT, using smaller sets
+that are no longer hitting sets lead to a rejection of correct
+implementations. Our explicit definition of $\pmax$ to establish this
+worst-case is useful to illustrate this point.
+
 In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
@@
 % -------------------------------------------------------------------------
-\subsection{Estimates for the Maximal Number of Trace Test Executions}
-\label{section:complexity:traces}
-
+\paragraph{Estimate for the Maximal Number of Trace Test Executions}
+%\label{section:complexity:traces}
 According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
 refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
-$U_F(j)$ above, the maximal number of executions to be performed by $(Q\parallel[\Sigma]
-U_T(pq-1))$ is of order  $O\big(\card{\Sigma}^{pq-1}\big)$.
+$U_F(j)$ above, the maximal number of executions to be performed by
+$(Q\parallel[\Sigma] U_T(pq-1))$ is of order
+$O\big(\card{\Sigma}^{pq-1}\big)$.
 
@@
 According to Theorem~\ref{th:failurestest}, the tests $U_F(j)$ need to be
-executed for $j = 0,\dots,pq-1$ to guarantee completeness.
- This means that the SUT is verified with test
-traces up to, and including, length $pq$: recall from the test specification,
-branch (\ref{eq:ufa}), that $U_F(j)$ will accept all traces $s.e$ with
-$s\in\trc(P), \#s = j, e\not\in\trc(P/s)$, so erroneous traces up to length
-$j+1$ are detected.
+executed for $j = 0,\dots,pq-1$ to guarantee completeness. So, the SUT is
+verified with test traces up to, and including, length $pq$. With branch
+(\ref{eq:ufa}), $U_F(j)$ accepts all traces $s.e$ with $s\in\trc(P), \#s
+= j, e\not\in\trc(P/s)$, so erroneous traces up to length $j+1$ are detected.
 
-It is interesting to investigate whether this maximal length is really
-necessary, or whether one could elaborate alternative complete test
-strategies where the SUT is tested with shorter traces only. Indeed, an
-example  presented in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} shows
-that when testing for equivalence of deterministic FSMs, it is sufficient to
-test the SUT with traces of significantly shorter length.
+It is interesting to investigate whether this maximal length is necessary, or
+one could elaborate alternative complete test strategies where the
+SUT is tested with shorter traces only. Indeed, an example
+in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} shows that, when testing
+for equivalence of deterministic FSMs, it is sufficient to test with
+traces of significantly shorter length.
 
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\begin{minipage}{.4\textwidth}
-\includegraphics[width=1.2\textwidth]{theorem5p.pdf}
-\end{minipage}
-\hfill
-\begin{minipage}{.55\textwidth}
-\includegraphics[width=1.2\textwidth]{theorem5q.pdf}
-\end{minipage}
-\caption{Transition graphs of $P$ (left) and $Q$ (right) from Example~\ref{ex:pq}
- for $p=3$ and $q=4$.}
-\label{fig:examplepq}
-\end{center}
-\end{figure}
-
 \begin{example}\label{ex:pq}
@@
 (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j))\hide \Sigma
-\]
-and
-\[
+\quad
+\mathrm{and}
+\quad
 (\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(j))\hide \Sigma
 \]
-hold for $j = 0,\dots,pq-2$. This means that none of the test cases $U_F(j)$
-and $U_T(j)$ are capable of detecting failures and trace refinement
-violations, if they only check traces up to length $pq-1$
-(recall that this corresponds to $j\le pq-2$).
+hold for $j = 0,\dots,pq-2$. So, none of the test cases $U_F(j)$ and $U_T(j)$
+are capable of detecting failures and trace refinement violations, if they
+only check traces up to length $pq-1$. (We recall that this corresponds to
+$j\le pq-2$).
 
-Process $Q$, however, neither conforms to $P$ in the failures refinement
-relation, nor in the trace
-refinement relation. This can only be seen when executing the test $U_F(pq-1)$ and
-$U_T(pq-1)$, respectively. These tests fail, so this shows that $P\not\lessdet_F Q$ and
-$P\not\lessdet_T Q$ according to Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this shows that
-the maximal trace
-length $pq$ to be investigated in the tests cannot be further reduced without losing
-the completeness property of the test suites.
-\xbox
+$Q$, however, neither conforms to $P$ in the failures refinement relation,
+nor in the trace refinement relation. This can only be seen when executing
+the test $U_F(pq-1)$ and $U_T(pq-1)$, respectively. These tests fail, so this
+shows that $P\not\lessdet_F Q$ and $P\not\lessdet_T Q$ according to
+Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this
+shows that the maximal trace length $pq$ to be investigated in the tests
+cannot be further reduced without losing the completeness property of the
+test suites. \xbox
 \end{example}
 %
-Generalising Example~\ref{ex:pq}, it can be shown that for any pair
-$2\le p,q \in\mathbb{N}$,
-there exist reference processes $P$ with $p$ states
-and implementation processes $Q$ with $q$ states, such that
-a violation of the trace refinement property
-can only be detected with a trace of length $pq$. This is proven in the following
-theorem. In the proof, we use the processes $P$ and $Q$ introduced in Example~\ref{ex:pq}.
+Generalising Example~\ref{ex:pq}, it can be shown that for any $p,q \geq 2$,
+%\in\mathbb{N}$,
+there exist reference processes $P$ with $p$ states and
+implementation processes $Q$ with $q$ states, such that a violation of the
+trace refinement property can only be detected with a trace of length $pq$.
+This is proven in the following
+theorem. %In the proof, we use the processes $P$ and $Q$ introduced in Example~\ref{ex:pq}.
+
+\begin{figure}[tbp]
+\begin{center}
+\begin{minipage}{.4\textwidth}
+\includegraphics[trim=1.4cm 0cm 0cm 0cm,clip,width=1.1\textwidth]{theorem5p.pdf}
+\end{minipage}
+\hfill
+\begin{minipage}{.55\textwidth}
+\includegraphics[trim=1.4cm 0cm 0cm 0cm,clip,width=1.1\textwidth]{theorem5q.pdf}
+\end{minipage}
+\caption{Transition graphs of $P$ (left) and $Q$ (right) from Example~\ref{ex:pq}
+ for $p=3$ and $q=4$.}
+\label{fig:examplepq}
+\end{center}
+\end{figure}
 
@@
 \begin{proof}
-Given $2\le p,q \in\mathbb{N}$, define reference process $P$ and implementation process $Q$ as in Example~\ref{ex:pq}. It is trivial to see that
-$G(P)$ has $p$ nodes and $G(Q)$ has $q$ nodes, so statements 1 and 2 of the theorem hold.
-
-Using regular expression notation, the traces of $P$ can be specified as
+Given $2\le p,q \in\mathbb{N}$, define reference process $P$ and
+implementation process $Q$ as in Example~\ref{ex:pq}. It is trivial to see
+that $G(P)$ has $p$ nodes and $G(Q)$ has $q$ nodes, so statements 1 and 2 of
+the theorem hold. Using regular expression notation, the traces of $P$ can be
+specified as
 \[

### 2018-11-28 scp-journal/article/complexity.tex
Remove stray i
Fixed typos in proof of Lemma 11.
Starting on 6.2

@@
-i\section{Complexity Considerations} \label{sec:complexity}
+\section{Complexity Considerations} \label{sec:complexity}
 % ================================================================================
@@
 Since $\bigcup_{A\in{\cal C}} A = \Sigma $ by construction of ${\cal C}$,
-$[\pmax]^0 = \Sigma$ and stated by (\ref{eq:pmaxa}). Since $\pmax/e = \pmax$
+$[\pmax/s]^0 = \Sigma$ as stated by (\ref{eq:pmaxa}). Since $\pmax/e = \pmax$
 for all $e\in\Sigma$, this proves statement (\ref{eq:pmaxb}). The internal
@@
 that $G(P)$ has $p$ nodes and $G(Q)$ has $q$ nodes, so statements 1 and 2 of
-the theorem hold. Using regular expression notation, the traces of $P$ can be
-specified as
-\[
+the theorem hold.
+
+Using regular expression notation, the traces of $P$ can be specified as
+$%\[
 \trc(P) = \prefs\big(  (a^*b)^{p-1}a^* \big),
-\]
-where $\prefs(M)$ denotes the set of all prefixes of traces in $M\subseteq\Sigma^*$,
-including the traces of $M$ themselves.
-The traces of $Q$ can be specified by
-\[
+$ %\]
+where $\prefs(M)$ denotes the set of all prefixes of traces in
+$M\subseteq\Sigma^*$, including the traces of $M$ themselves. The traces of
+$Q$ can be specified by
+$%\[
 \trc(Q) = \prefs\big( (a^{q-1}(a|b))^*  \big).
-\]
+$ %\]
 It is easy to see that $\trc(Q)\not\subseteq\trc(P)$; for example, the trace
-$(a^{q-1}b)^p$ is in $\trc(Q)\setminus\trc(P)$, because $P$-traces may contain at most $p-1$ $b$-events. This proves statement~3 of the theorem.
+$(a^{q-1}b)^p$ is in $\trc(Q)\setminus\trc(P)$, because traces of $P$ contain
+at most $p-1$ $b$-events. This proves statement~3.
 
-Let $s \in\trc(Q)$ be any trace of length $\#s = pq-1$. Then $s$ can be represented
-by $s = (a^{q-1}(a|b))^{p-1}a^{q-1} \in \prefs\big( (a^{q-1}(a|b))^*  \big)$.
-Then $s$ is also an element of $\trc(P)$, because $(a^{q-1}(a|b))^{p-1}a^{q-1}$
-is also contained in $\prefs\big(  (a^*b)^{p-1}a^* \big)$: this is easy to see, since
-$\prefs\big(  (a^*b)^{p-1}a^* \big)$ contains all finite sequences of $a$-events,
-where at most $p-1$ events $b$ have been inserted. This proves statement 4 of the theorem.
+Let $s \in\trc(Q)$ be any trace of length $\#s = pq-1$, then $s =
+(a^{q-1}(a|b))^{p-1}a^{q-1} \in \prefs\big( (a^{q-1}(a|b))^*  \big)$. So, $s$
+is also an element of $\trc(P)$, because $(a^{q-1}(a|b))^{p-1}a^{q-1}$ is
+also contained in $\prefs\big( (a^*b)^{p-1}a^* \big)$, since $\prefs\big(
+(a^*b)^{p-1}a^* \big)$ contains all finite sequences of $a$, where at most
+$p-1$ events $b$ have been inserted. This proves statement 4.
 
-To prove statement~5, we observe that the specification of $P$ implies (the
-expression $(s\cnt b)$ denotes the number of $b$-events occurring in trace $s$)
+To prove statement~5, we observe that the specifications of $P$ and $Q$ lead
+to the following sets of minimal acceptances.  In these definition, the
+expression $(s\cnt b)$ denotes the number of $b$ events occurring in trace
+$s$.
 \[
@@
 \right.
-\]
-and
-\[
+\also
 \minaccs(Q/s) = \left\{
@@
 \]
-As a  consequence, the minimal acceptance set $A_P = \{a\}$ which is
-contained in every $\minaccs(P/s)$ fulfils $A_P \subseteq A_Q$ for
-any $A_Q\in\minaccs(Q/s)$, when $s\in\trc(P)\cap \trc(Q)$. Now Lemma~\ref{lemma:tgtrcref},
-(\ref{eq:failrefb}) can be applied to conclude that $Q\ conf\ P$.
-\xbox
+So, the minimal acceptance set $A_P = \{a\}$ contained in every
+$\minaccs(P/s)$ fulfils $A_P \subseteq A_Q$ for any $A_Q\in\minaccs(Q/s)$,
+when $s\in\trc(P)\cap \trc(Q)$. Now Lemma~\ref{lemma:tgtrcref}, in particular
+(\ref{eq:failrefb}), can be applied to conclude that $Q\ conf\ P$. \xbox
 \end{proof}
@@
 refinement may remain undetected if only traces shorter than $pq$ are checked
-during tests, it can also be applied to our trace refinement tests.
+during tests, it can also be applied to our trace-refinement tests.
 Therefore, test suites $\{ U_T(j) \}$ with $j<pq-1$ are not complete. It is
-discussed in Section~\ref{sec:conc} how the number of test traces to be
+discussed next in Section~\ref{sec:conc} how the number of test traces to be
 executed by complete test suites for failures or trace refinement can still

### 2018-11-29 scp-journal/article/complexity.tex
Finished revision

@@
 verify an SUT with respect to failures refinement, when we use our test
-suite? (2)~What is the worst-case bound for trace refinement? (3)~Is it
+suite? (2)~What is the worst-case bound for traces refinement? (3)~Is it
 possible to reduce the maximal length of traces when testing for failures or
-trace refinement? We consider the first question~(1) in
+traces refinement? We consider the first question~(1) in
 Section~\ref{section:complexity:failures}), where we also discuss whether it
@@
 %\label{section:complexity:traces}
-According to Theorem~\ref{th:tracetest}, a complete test suite checking trace
-refinement just contains the adaptive test case $U_T(pq-1)$. As derived for
-$U_F(j)$ above, the maximal number of executions to be performed by
-$(Q\parallel[\Sigma] U_T(pq-1))$ is of order
+According to Theorem~\ref{th:tracetest}, a complete test suite checking
+traces refinement just contains the adaptive test case $U_T(pq-1)$. As
+derived for $U_F(j)$ above, the maximal number of executions to be performed
+by $(Q\parallel[\Sigma] U_T(pq-1))$ is of order
 $O\big(\card{\Sigma}^{pq-1}\big)$.
@@
 hold for $j = 0,\dots,pq-2$. So, none of the test cases $U_F(j)$ and $U_T(j)$
-are capable of detecting failures and trace refinement violations, if they
+are capable of detecting failures and traces-refinement violations, if they
 only check traces up to length $pq-1$. (We recall that this corresponds to
@@
 $Q$, however, neither conforms to $P$ in the failures refinement relation,
-nor in the trace refinement relation. This can only be seen when executing
+nor in the traces-refinement relation. This can only be seen when executing
 the test $U_F(pq-1)$ and $U_T(pq-1)$, respectively. These tests fail, so this
@@
 %\in\mathbb{N}$,
-there exist reference processes $P$ with $p$ states and
-implementation processes $Q$ with $q$ states, such that a violation of the
-trace refinement property can only be detected with a trace of length $pq$.
-This is proven in the following
+there exist reference processes $P$ with $p$ states and implementation
+processes $Q$ with $q$ states, such that a violation of the traces-refinement
+property can only be detected with a trace of length $pq$. This is proven in
+the following
 theorem. %In the proof, we use the processes $P$ and $Q$ introduced in Example~\ref{ex:pq}.
@@
 \end{enumerate}
-As a consequence, the upper bound $pq$ for the length of traces to be tested when checking for failures refinement or trace refinement
-cannot be reduced without losing the test suite's completeness property.
+As a consequence, the upper bound $pq$ for the length of traces to be tested
+when checking for failures refinement or traces refinement cannot be reduced
+without losing the test suite's completeness property.
 \end{theorem}
@@
 %
-Since Theorem~\ref{th:maxtracelen} just states that a violation of trace
-refinement may remain undetected if only traces shorter than $pq$ are checked
-during tests, it can also be applied to our trace-refinement tests.
-Therefore, test suites $\{ U_T(j) \}$ with $j<pq-1$ are not complete. It is
-discussed next in Section~\ref{sec:conc} how the number of test traces to be
-executed by complete test suites for failures or trace refinement can still
-be reduced {\it without} reducing the maximal length.
+It is discussed next in Section~\ref{sec:conc} how the number of test traces
+to be executed by complete test suites for failures or traces refinement can
+still be reduced {\it without} reducing the maximal length.
 

### 2018-11-29 scp-journal/article/finitecomplete-failures.tex
Finished revision

@@
 It is necessary to use failures refinement in the definition above, and not
-just trace refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ may have
+just traces refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ may have
 the same visible traces $\varepsilon$ and $\epass$ as the ``Test Passed

### 2018-11-29 scp-journal/article/finitecomplete-trace.tex
Finished revision

@@
 % ==========================================================================
-\section{Finite Complete Test Suites for CSP Trace Refinement}
+\section{Finite Complete Test Suites for CSP Traces Refinement}
 \label{sec:finitecomplete}
@@
 
-For establishing trace refinement, the following class of adaptive test cases
-are used for a given reference process $P$ and  integers $j \ge 0$. Just as
-for the tests developed  in Section~\ref{sec:finitecompletefails} to verify
-failures refinement, the tests for trace refinement are derived from the
-reference model's transition graph
+For establishing traces refinement, the following class of adaptive test
+cases are used for a given reference process $P$ and  integers $j \ge 0$.
+Just as for the tests developed  in Section~\ref{sec:finitecompletefails} to
+verify failures refinement, the tests for traces refinement are derived from
+the reference model's transition graph
 $$
@@
 Theorem~\ref{th:failurestest}. A noteworthy difference is that the complete
-suite for trace refinement just needs the single adaptive test case
+suite for traces refinement just needs the single adaptive test case
 $U_T(pq-1)$, while failures refinement requires the execution of $\{

### 2018-11-30 scp-journal/article-before-submission/abstract.tex
All files as completed BEFORE submission

--- 2018-11-16 scp-journal/article/abstract.tex
@@
 process representing the reference model, finite test suites for checking the
-conformance relations trace and failures refinement are presented, and their
+conformance relations traces and failures refinement are presented, and their
 completeness (that is, capability to uncover conformity violations) is

### 2018-11-30 scp-journal/article-before-submission/conclusion.tex
All files as completed BEFORE submission

--- 2018-11-10 scp-journal/article/conclusion.tex
@@
 % ==============================================================================
-\subsubsection*{Discussion of Further Reductions of the Test Effort}
-As has been shown in Theorem~\ref{th:maxtracelen}, the maximal length $pq$ of 
-traces to be tested for either  failures refinement or trace refinement cannot be further reduced.
-It is noteworthy, however, that when testing finite state machines for equivalence, considerably shorter traces can be used.
-From the classical results published in~\cite{chow:wmethod,vasilevskii1973}, for example, it follows that the maximal trace length to be executed is less or equal to $3p -q$, which is considerably smaller than $pq$ for $p,q\ge 3$.
-As a consequence, the investigation of complete test suites establishing failures or trace equivalence is of considerable interest and will be discussed in a future paper.
+\paragraph{Further Reductions of the Test Effort} As shown in
+Theorem~\ref{th:maxtracelen}, the maximal length $pq$ of traces to be tested
+for either  failures refinement or traces refinement cannot be further
+reduced. It is noteworthy, however, that when testing FSMs for equivalence,
+considerably shorter traces can be used. From the classical results published
+in~\cite{chow:wmethod,vasilevskii1973}, for example, it follows that the
+maximal trace length to be executed is less or equal to $3p -q$, which is
+considerably smaller than $pq$ for $p,q\ge 3$. As a consequence, the
+investigation of complete test suites establishing failures or trace
+equivalence is of considerable interest and will be discussed in a future
+paper.
 
-It is also known from FSM
-testing that it is not necessary to check {\it all} traces up to  
-length $pq$ 
-when testing for reduction of finite state machines (this corresponds to trace refinement). Notable complete strategies supporting this fact have been presented,
-for example,
+It is also known from FSM testing that it is not necessary to check {\it all}
+traces up to length $pq$ when testing for reduction of FSMs (which
+corresponds to trace refinement). Notable complete strategies supporting this
+fact have been presented, for example,
 in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}.
@@
 the latter. We intend to study translations of several promising FSM
-strategies to CSP in the future. These will effectively reduce the 
-upper bound for the number of test executions to be performed, which  has been 
-shown to be of order $O\big(\card{\Sigma}^{pq-1}\big)$ for our trace refinement tests 
-in Section~\ref{sec:complexity}.  The
-bound $\binom{n}{\lfloor n/2\rfloor}$ for the number of sets to be used in probing the SUT for illegal
-deadlocks, however, cannot be further reduced, as has been established in
+strategies to CSP. These will effectively reduce the upper bound for the
+number of test executions to be performed, which  has been shown to be of the
+order $O\big(\card{\Sigma}^{pq-1}\big)$ for our traces-refinement tests in
+Section~\ref{sec:complexity}.  The bound $\binom{n}{\lfloor n/2\rfloor}$ for
+the number of sets to be used in probing the SUT for illegal deadlocks,
+however, cannot be further reduced, as established in
 Lemma~\ref{lemma:hseta}.
@@
 % ==============================================================================
-\subsubsection*{Discussion of Adaptive Test Cases}
-The tests suggested
-in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were \emph{preset} in the
-sense that the trace to be executed was pre-defined for each test. As a consequence,
-the authors of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive}
-as a third test result, applicable to the situations where the intended trace
-of the execution was blocked, due to legal, but nondeterministic behaviour of the
-SUT. We consider this as a disadvantage, since, when aiming at executing a specific trace $s$
-before being able to check the test objective---for example, the absence of deadlocks when offering a hitting set $H$ of the minimal acceptances---it may take several tries until the full trace $s$ is accepted by the SUT.
-Considering the complete testing assumption described in Section~\ref{sec:finitecompletefails}, it may even take $c^{\#s}$ tries to reach the end of the trace $s$, if the SUT can legally block every $s$-event due to nondeterminism, so that $c$
-trials are required for each event is accepted.
+\paragraph{Adaptive Test Cases} The tests suggested
+in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were
+\emph{preset} in the sense that the trace to be executed was pre-defined for
+each test. As a consequence, the authors
+of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive} as a
+third test result, applicable to the situations where the intended trace of
+the execution was blocked, due to legal, but nondeterministic behaviour of
+the SUT. We consider this as a disadvantage, since, when aiming at executing
+a specific trace $s$ before being able to check the test objective---for
+example, the absence of deadlocks when offering a hitting set $H$ of the
+minimal acceptances---it may take several tries until the full trace $s$ is
+accepted by the SUT. Considering the complete testing assumption described in
+Section~\ref{sec:finitecompletefails}, it may even take $c^{\#s}$ tries to
+reach the end of the trace $s$, if the SUT can legally block every event of
+$s$ due to nondeterminism, so that $c$ trials are required for each event is
+accepted.
+
+Those authors, later, in the context of a richer algebra based on CSP, have
+considered a framework similar to adaptive testing~\cite{CG15}.  They have,
+however, considered only traces refinement, and have not studied complexity.
 
@@
 adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
-U_F(j))$ for failures refinement 
-may only stop early with $\epass$ 
-after traces $s$ satisfying $\minhits(P/s) = \varnothing$, and may deadlock 
-(a) after a trace $s$ where the SUT illegally deadlocks (so $\minhits(P/s) \neq \varnothing$ for
-the reference process, but $\minhits(Q/s) = \varnothing$ for the implementation $Q$), or
-(b) 
-in the final 
-step when---just as in the corresponding test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}--- hitting sets $H$ are offered to the SUT and it refuses their acceptance.
-In both situations (a) and (b) the test fails. As a consequence, far less test repetitions are required according to the complete testing assumption than for the successful execution of all 
-test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}.
+U_F(j))$ for failures refinement may only stop early with $\epass$ after
+traces $s$ satisfying $\minhits(P/s) = \varnothing$, and may deadlock (a)
+after a trace $s$ where the SUT illegally deadlocks (so $\minhits(P/s) \neq
+\varnothing$ for the reference process, but $\minhits(Q/s) = \varnothing$ for
+the implementation $Q$), or (b) in the final step when---just as in the
+corresponding test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}---
+hitting sets $H$ are offered to the SUT and it refuses their acceptance. In
+both situations (a) and (b) the test fails. As a consequence, far less test
+repetitions are required according to the complete testing assumption than
+for the successful execution of all test cases specified
+in~\cite{DBLP:conf/icfem/CavalcantiG07}.
 
-Another distinction of our failures test cases $U_F(j)$ to the tests specified in~\cite{DBLP:conf/icfem/CavalcantiG07} consists in the fact that the former test both trace refinement and
-the $conf$ conformance relation in one go, whereas the latter use separate test suites to 
-establish these two correctness conditions. Again, we consider the  
-structure of the  test cases 
-$U_F(j)$ as advantageous, since, when checking acceptance of a hitting set $H$ 
-after a trace $s$ for $c$ times according to the complete testing assumption,  
-any acceptance of an illegal event $e\in\Sigma - [P/s]^0$ should also be revealed within these $c$ tries.
+Another distinction of our failures test cases $U_F(j)$ to the tests
+specified in~\cite{DBLP:conf/icfem/CavalcantiG07} consists in the fact that
+the former test both traces refinement and the $conf$ conformance relation in
+one go, whereas the latter use separate test suites to establish these two
+correctness conditions. Again, we consider the structure of the  test cases
+$U_F(j)$ as advantageous, since, when checking acceptance of a hitting set
+$H$ after a trace $s$ for $c$ times according to the complete testing
+assumption, any acceptance of an illegal event $e\in\Sigma - [P/s]^0$ should
+also be revealed within these $c$ tries.
 
@@
 % ==============================================================================
-\subsubsection*{Discussion of Fault Domains}
-As already mentioned, the work in~\cite{DBLP:conf/pts/CavalcantiS17} defines
-a fault domain as the set of processes that refine a given CSP process.  In
-that context, only testing for traces refinement is considered, and the
-complete test suites may not be finite. So, the work presented here goes well
-beyond what is achieved there. On the other hand,
-\cite{DBLP:conf/pts/CavalcantiS17} presents an algorithm for test generation
-that can be easily adapted to consider additional selection and termination
-criteria, like, for example, the length of the traces used to construct
-tests. It would be possible, for instance, to use the bound indicated here.
-Moreover, specifying a fault domain as a CSP process allows us to model
-domain-specific knowledge using CSP. For example, if an initialisation
-component defined by a process $I$ can be regarded as correct without further
-testing, we can use $I; RUN$ as a fault domain, to indicate that any SUT of
-interest implements $I$ correctly, but afterwards has a potentially arbitrary
-behaviour specified by $RUN$. In addition, the work
-in~\cite{DBLP:conf/pts/CavalcantiS17} is not restricted to finite or
-non-terminating processes.
+\paragraph{Fault Domains} As already mentioned, the work
+in~\cite{DBLP:conf/pts/CavalcantiS17} defines a fault domain as the set of
+processes that refine a given CSP process.  In that context, only testing for
+traces refinement is considered, and the complete test suites may not be
+finite. So, the work presented here goes well beyond what is achieved there,
+but is restricted to finite and nonterminating reference processes. In
+addition, \cite{DBLP:conf/pts/CavalcantiS17} presents an algorithm for test
+generation that can be adapted to consider additional selection and
+termination criteria, like, for example, the length of the traces used to
+construct tests. It would be possible, for instance, to use the bound
+indicated here. Moreover, specifying a fault domain as a CSP process allows
+us to model domain-specific knowledge using CSP. For example, if an
+initialisation component defined by a process $I$ can be regarded as correct
+without further testing, we can use $I; RUN$ as a fault domain, to indicate
+that any SUT of interest implements $I$ correctly, but afterwards has an
+arbitrary behaviour specified by $RUN$.
 
 % ==============================================================================
-\subsubsection*{Implications for CSP Model Checking}
-As explained in the previous sections, passing a test is specified by
-$$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$$ for
+\paragraph{Implications for CSP Model Checking} As explained in the previous
+sections, passing a test is characterised by the failures-refinement check
+$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ for
 failures testing. If the SUT $Q$ is not a programmed piece of software or an
@@
 parallelised on several CPUs. Alternatively it is interesting to investigate
-whether the check\footnote{We are grateful to Bill Roscoe for suggesting this
-option.}
+whether the check of\footnote{We are grateful to Bill Roscoe for suggesting
+this option.}
 \[
@@
 may perform better than the check of $P\lessdet_F Q$, since the former allows for
-other optimisations in the model checker than the latter.
+other optimisations in the model checker.
 
-For a large and complex CSP implementation process $Q$, it may be too time consuming to
+For a large implementation process $Q$, it may be too time consuming to
 generate its normalised transition graph, so that its number $q$ of nodes is
-unknown. In such a case the suggested options may still be used as efficient
-bug finders: use  $p$ of reference process $P$ as initial $q$-value and
-increment $q$ from there, as long as each increment reveals new errors.
+unknown. In such a case, our testing approach based on fault domains may
+still be used as efficient bug finders:~use the number of nodes of the
+normalised transition graph of the reference process as the initial value for
+$q$ and increment $q$ from there, as long as each increment reveals new
+errors.
 
 % ==============================================================================
-\subsubsection*{Conclusion}
-
+\paragraph{Conclusion}
 In this paper, we have introduced finite complete testing strategies for
@@
 strategies are applicable to the conformance relations failures refinement
-and trace refinement. The underlying fault domains have been defined as the
+and traces refinement. The underlying fault domains have been defined as the
 sets of CSP processes whose normalised transition graphs do not have more
@@
 suites are available. We have shown for the strategy to check failures
-refinement that the way of probing the SUT for illegal deadlocks used in our
-test cases is optimal in the sense that it is not possible to guarantee
-exhaustiveness with fewer probes.
+refinement that the way of probing the SUT for illegal deadlocks in our test
+cases is optimal, so that it is not possible to guarantee exhaustiveness with
+fewer probes. Moreover, the maximal length of the test traces cannot be
+reduced without losing the test suite's completeness property; this holds
+both for traces and failures refinement.
 

### 2018-11-30 scp-journal/article-before-submission/introduction.tex
All files as completed BEFORE submission

--- 2018-11-16 scp-journal/article/introduction.tex
@@
 In~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
-the authors devised linear test cases: after running through a preset trace
-$s$, test cases for trace refinement check for illegal acceptance of a
+the authors devised linear test cases:~after running through a preset trace
+$s$, test cases for traces refinement check for illegal acceptance of a
 specific event $e$, and test cases verifying  nondeterministic behaviour
@@
 \cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
-because trace refinement and correctness of nondeterministic behaviour are
+because traces refinement and correctness of nondeterministic behaviour are
 checked by two different sets of test cases.
@@
 A sample test suite is presented in Section~\ref{sec:case}. Test suites
-checking trace refinement are a simplified version of the former class; they
+checking traces refinement are a simplified version of the former class; they
 are presented in Section~\ref{sec:finitecomplete}. The optimality results are

### 2018-11-30 scp-journal/article-before-submission/preliminaries.tex
All files as completed BEFORE submission

--- 2018-11-19 scp-journal/article/preliminaries.tex
@@
 $s\in\trc(Q)$. So, for divergence-free processes, failures refinement implies
-trace refinement. Therefore, using the conformance relation $conf$ below
+traces refinement. Therefore, using the conformance relation $conf$ below
 %
@@
    \begin{center}
-     \includegraphics[trim=0cm 1cm 0cm 1cm,clip,scale=.6,%width=\textwidth
+     \includegraphics[trim=0cm 1.4cm 0cm 1.4cm,clip,scale=.6,%width=\textwidth
                               ]{p.pdf}
@@
 as established by the results  in the following lemma. There, result
-(\ref{eq:trcrefa}) reflects trace refinement in terms of graph traces; result
-(\ref{eq:failconf}) expresses failures refinement in terms of traces
+(\ref{eq:trcrefa}) reflects traces refinement in terms of graph traces;
+result (\ref{eq:failconf}) expresses failures refinement in terms of traces
 refinement and $conf$; result (\ref{eq:failrefa}) states how $conf$ can be

### 2018-11-30 scp-journal/article/main.tex
revised sections 1 and 2, shortenned paper
Section 4 revised
Built one main.tex containing ALL information, including bibliography. No includes of other files apart from graphics files.

@@
-\documentclass[1p]{elsarticle}
-
-%
-%
-%\usepackage{lineno,hyperref}
-%\modulolinenumbers[5]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[3p,times]{elsarticle}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \journal{Science of Computer Programming}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{frontmatter}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{abstract}
-\input{abstract}
+In this paper, new contributions for model-based testing using Communicating
+Sequential Processes (CSP) are presented. For a finite non-terminating CSP
+process representing the reference model, finite test suites for checking the
+conformance relations traces and failures refinement are presented, and their
+completeness (that is, capability to uncover conformity violations) is
+proven. The fault domains for which complete failure detection can be
+guaranteed are specified by means of normalised transition graphs
+representing the failures semantics of finite-state CSP processes. While
+complete test suites for CSP processes have been previously investigated by
+several authors, a sufficient condition for their finiteness is presented
+here for the first time. Moreover, it is shown that the test suites are
+optimal in two aspects:~(a)~the maximal length of test traces cannot be
+further reduced, and (b)~the nondeterministic behaviour cannot be tested with
+smaller or fewer sets of events, without losing the test suite's completeness
+property.
 \end{abstract}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{keyword}
-Model-based testing, CSP\sep Trace Refinement\sep Failures Refinement\sep Complete Test Suites\sep
+Model-based testing, CSP\sep Traces Refinement\sep Failures Refinement\sep
+Complete Test Suites
 \end{keyword}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{frontmatter}
 
-\input{introduction}
-
-\input{preliminaries}
-
-\input{finitecomplete-failures}
-
-\input{casestudies}
-
-\input{finitecomplete-trace}
-
-\input{complexity}
-
-%%%\input{related}
-
-\input{conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% =========================================================================
+
+\section{Introduction}
+\label{sec:intro}
+
+% =================================================================================
+
+\paragraph{Motivation}
+Model-based testing (MBT) is an active research field; results are currently
+being evaluated and integrated into industrial verification processes by many
+companies worldwide. This holds particularly for the embedded and
+cyber-physical systems domains, where critical systems require rigorous
+testing~\cite{jp2018ets,DBLP:conf/isola/0001BH18}.
+
+In the safety-critical domain, test suites with guaranteed fault coverage are
+of particular interest. For black-box testing, guarantees can be given only
+if certain hypotheses are satisfied. These hypotheses are usually specified
+by a \emph{fault domain}:~a set of models that may or may not conform to a
+given reference model. \emph{Complete} test strategies guarantee to accept
+every  system under test (SUT) conforming to the reference model, and uncover
+every conformance violation, provided that the SUT behaviour is captured by a
+member of the fault domain.
+
+Generation techniques for complete test suites have been developed for
+various formalisms; we mention here representative work for finite state
+machines~\cite{hierons_testing_2004,simao_reducing_2012}, timed
+automata~\cite{Springintveld2001}, {\sf\em
+Circus}~\cite{DBLP:journals/acta/CavalcantiG11}, Timed
+CSP~\cite{Schneider:1995:OST:203471.203475}, general labelled transition
+systems~\cite{DBLP:journals/cn/Tretmans96}, symbolic state
+machines~\cite{DBLP:conf/icst/Petrenko16}, and Kripke
+structures~\cite{Huang2017}. In this article, we tackle \emph{Communicating
+Sequential Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a
+mature process algebra that has been shown to be well-suited for the
+description of reactive control systems in many publications over almost five
+decades. Many of these applications are described in~\cite{Roscoe2010} and in
+the references there. Industrial success has also been reported; see, for
+example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.
+
+% ==================================================================================
+
+\paragraph{Main Contributions}
+
+This article presents   complete black-box test suites for software and
+systems modelled using CSP. They can be generated for non-terminating,
+divergence-free, finite-state CSP processes with finite alphabets,
+interpreted both in the trace and the failures semantics. Divergence freedom
+is usually assumed in black-box testing, since it cannot distinguish between
+divergence and deadlock using testing.
+
+\newpage
+\noindent %
+The main novel contributions in this article may be summarised as
+follows.
+\begin{enumerate}
+\item It is shown that trace or failures conformance can be established
+    with finitely many test cases, provided suitable fault domains are
+    chosen, so that the true behaviour of the SUT is reflected by members
+    of these domains.
+
+\item The definition of these fault domains is based on the well-known normalised transition graphs~\cite{Roscoe:1994:chapter} representing the  trace and failures semantics of finite-state CSP processes.
+A fault domain contains all CSP processes over a given alphabet, whose normalised transition graphs have at most $q$ nodes for some $q\in\mathbb{N}$.
+
+\item Worst-case complexity bounds for the number of test executions to be performed are given.
+
+\item It is shown that the maximal length of the test traces involved cannot be further reduced without losing the test suite's completeness property.
+
+\item Likewise, is shown that the non-deterministic behaviour of the SUT cannot be
+checked for admissibility with smaller or fewer sets of events.
+\end{enumerate}
+
+% ==================================================================================
+
+\paragraph{Related Work}
+Our results complement and extend work previously published
+in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
+None of these provide sufficient conditions for constructing finite complete
+test suites. So, they also do not provide complexity bounds on the number of
+test executions needed to establish conformance between an SUT and a
+reference process. In~\cite{DBLP:conf/pts/CavalcantiS17}, fault domains are
+used, but these contain all processes refining a ``top'' fault domain
+process. This concept is orthogonal to the one investigated here:~members of
+our fault domain need not be in refinement relation to any other process in
+the domain. They just adhere to the same upper bound $q$ of nodes in their
+normalised transition graphs.
+
+The minimal sets of events used for checking nondeterministic behaviour of the SUT used in our article were already suggested in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07}; in the present article, however, they have been identified for the first time as {\it minimal hitting sets}~\cite{5533149} of minimal acceptances in a given process state, and we establish an upper bound stating how many of these sets need to be checked for the ``most extreme form of nondeterminism'' that may be exhibited by the SUT.
+
+In~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
+the authors devised linear test cases:~after running through a preset trace
+$s$, test cases for traces refinement check for illegal acceptance of a
+specific event $e$, and test cases verifying  nondeterministic behaviour
+check for  acceptance of events from some set $A$. In the present article, we
+follow the alternative approach proposed
+in~\cite{DBLP:conf/fm/PeleskaS96,peleska1997a} and use {\it adaptive} test
+cases. This means that each test case adapts its trace execution to the
+nondeterministic behaviour of the SUT, checks for trace violations at any
+point during the test execution, and checks for the acceptance of a given
+minimal hitting set of events after any legal trace of a test case-specific
+length.
+
+The adaptive test cases have the advantage that test executions only lead to
+an inconclusive result if the reference process allows for a nondeterministic
+choice between deadlock and trace continuation in a certain state. In
+contrast to this, the linear test cases may lead to many more futile
+executions with inconclusive results, if the SUT refuses to engage into the
+next event $e$ from the preset trace $s$, due to legal nondeterministic
+choices leading to a refusal of $e$. Moreover, the preset traces $s$ need to
+be executed twice according to the strategies devised in
+\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17},
+because traces refinement and correctness of nondeterministic behaviour are
+checked by two different sets of test cases.
+
+The approach  to specifying fault domains by means of normalised transition
+graphs has been inspired by the typical method used in the construction of
+complete test suites for finite state machines (FSMs). There, fault domains
+typically contain all FSMs over a given alphabet whose number of states does
+not exceed a given value
+$q$~\cite{chow:wmethod,vasilevskii1973,luo_test_1994}.
+
+%There,
+%fault domains are specified as collections of processes refining a ``most
+%general'' fault domain member. With that concept, complete test suites may be
+%finite or infinite. This result gives important insight into the theory of
+%fault-domain testing for CSP, but we are particularly interested in {\it
+%finite} suites when it comes to practical application.
+%While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
+%select tests from still infinite test suites, here, we further restrict fault
+%domains using a graph representation of processes (originally elaborated for
+%model checking) to obtain finite test suites.
+%
+%Our approach to the definition of CSP fault domains is presented in this
+%paper. We take advantage of results on model checking of CSP
+%processes, where the failures semantics of a finite-state CSP process is
+%represented as a finite normalised transition graph, whose edges are labelled
+%by the events the process engages in, and whose nodes are labelled by minimal
+%acceptances or, alternatively, maximal
+%refusals~\cite{Roscoe:1994:chapter}. The maximal refusals express the
+%degree of nondeterminism present in a process state that is in
+%one-one-correspondence to a node of the normalised transition graph. Inspired
+%by the way fault-domains are specified for finite state machines (FSMs),
+%we define them here as the set of CSP processes whose normalised transition
+%graphs do not exceed the order (that is, the number of nodes) of the
+%reference model's graph by more than a given constant.
+%
+%The main contribution of this paper is the construction of two test suites to
+%verify the conformance relations \emph{trace refinement} and \emph{failures
+%refinement} for any given non-terminating, non-divergent, finite-state,
+%finite alphabet CSP process. We prove their completeness with respect to
+%fault domains of the described type.  The existence of -- possibly infinite
+%-- complete test suites has been established for process algebras, and for
+%CSP in particular, by several
+%authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
+%To the best of our knowledge, this article is the first to present {\it
+%finite}, complete test suites associated with this class of fault domains and
+%conformance relations.
+%
+%Our result is not a simple transcription of existing knowledge about finite,
+%complete test suites for FSMs. The capabilities of CSP to express
+%nondeterminism in a more fine-grained way than it is possible for FSMs
+%requires a more complex approach to testing for conformity in the failures
+%model than required for model-based testing against nondeterministic FSMs, as
+%published, for example,
+%in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. CSP distinguishes
+%between external choice, where the environment can control the behaviour of a
+%process, and internal choice where the behaviour is decided internally and
+%cannot be influenced by the environment. In contrast, FSMs specify
+%nondeterministic behaviour by offering more than one possible output for a
+%given input, and this output can be controlled by an FSM representing the
+%environment. Loosely speaking, this corresponds to external choice in CSP,
+%while there is no equivalent to internal choice in nondeterministic FSMs.
+%
+%Finally, we prove two optimality results here. (1)~We show that our approach
+%to testing the admissibility of an SUT's nondeterministic behaviour is
+%optimal in the sense that it cannot be established with fewer ``probes''
+%investigating the SUT's degree of nondeterminism. These probes are minimal
+%sets of events offered by the tests to the SUT, such that a refusal of the
+%SUT to engage into at least one of these events reveals a violation of
+%failures refinement. (2)~Furthermore, it is shown that the maximal length of
+%traces to be investigated in a complete test suite cannot be further reduced
+%without loosing the suite's capability to uncover any violation of trace or
+%failures refinement.
+
+% ==================================================================================
+
+\paragraph{Overview} In Section~\ref{section:preliminaries}, we present the
+background relevant to our work. In Section~\ref{sec:finitecompletefails},
+finite complete test suites for verifying failures refinement are presented.
+A sample test suite is presented in Section~\ref{sec:case}. Test suites
+checking traces refinement are a simplified version of the former class; they
+are presented in Section~\ref{sec:finitecomplete}. The optimality results are
+presented in Section~\ref{sec:complexity}, together with further complexity
+considerations. Our results are discussed in Section~\ref{sec:conc}, where we
+also conclude. References to further related work are given throughout the
+paper where appropriate.
+
+% ==================================================================================
+
+% =========================================================================
+\section{Preliminaries}
+\label{section:preliminaries}
+% =========================================================================
+
+We present CSP~(Section~\ref{section:csp}) and the concept of minimal hitting
+sets~(Section~\ref{sec:hit}), which is central to our notion of test
+for failures refinement. To study complexity, we also introduce the concept
+of Sperner families~(Section~\ref{sec:sperner}).
+
+% =========================================================================
+\subsection{CSP, Refinement, and Normalised Transition Graphs}
+\label{section:csp}
+% =========================================================================
+
+\paragraph{Communicating Sequential Processes (CSP)} This is a process algebra supporting
+system development by refinement. Using CSP, we model both systems and their
+components using processes. They are characterised by their patterns of
+interactions, modelled by synchronous, instantaneous, and atomic events.
+
+Throughout this paper, the alphabet of the processes, that is, the set of
+events in scope, is denoted by $\Sigma$ and supposed to be finite.
+The FDR tool~\cite{fdr} supports model checking and semantic analyses of
+finite-state CSP processes.
+
+A prefixing operator $e \then P$ defines a process that is ready to engage in
+the event $e$, pending agreement of its environment to synchronise. After $e$
+occurs, the process behaves as defined by $P$. The environment can be other
+processes, in parallel, or the environment of a system as a whole.
+
+Two forms of choice support branching behaviour. An external choice $P
+\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
+initial events of $P$ and $Q$. Once a synchronisation takes place, the
+process that has offered the event that has occurred is chosen and the other
+is discarded. In an internal choice $P \intchoice Q$, the environment does
+not have an opportunity to interfere:~the choice is made by the process.
+%
+\begin{example}\label{example:CSP}
+  We consider the processes $P$, $Q$, and $R$ defined below. $P$ is initially
+  ready to engage in the event $a$, and then makes an internal choice to
+  behave like either $Q$ or $R$.
+  \begin{eqnarray*}
+  P & = & a \then (Q\intchoice R)
+  \\
+  Q & = & a \then P \extchoice c \then P
+  \\
+  R & = & b \then P \extchoice c \then R
+  \end{eqnarray*}
+  $Q$, for instance, offers to the environment the choice to engage in $a$
+  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
+  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
+  chosen, the recursion is to $R$.
+  \xbox
+\end{example}
+%
+Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
+external and internal choice operators define a choice over a collection of
+processes $P(i)$. If the index set $I$ is empty, the external choice is the
+process $\Stop$, which deadlocks:~does not engage into any event or
+terminate. For an external choice $\Extchoice e: A @ e \then P(e), A\subseteq
+\Sigma$ over a set $A$ of events, we use the abbreviation $e:A \then P(e)$.
+An iterated internal choice is not defined for an empty index set.
+
+There are several parallelism operators. A widely used form of parallelism $P
+\parallel[cs] Q$ defines a process in which the behaviour is characterised by
+those of $P$ and $Q$ in parallel, synchronising on the events in the set
+$cs$. Other forms of parallelism available in CSP can be defined using this
+parallelism operator.
+
+Interactions that are not supposed to be visible to the environment can be
+hidden. The operator $P \hide H$ defines a process that behaves as $P$, with
+the interactions modelled by events in the set $H$ hidden. Frequently, hiding
+is used in conjunction with parallelism:~it is often desirable to make
+actions of each process in a network of parallel processes, perhaps used for
+coordination of the network, invisible, while events happening at their
+interfaces remain observable.
+
+A rich collection of process operators allows us to define networks of
+parallel processes in a concise and elegant way, and reason about safety,
+liveness, and divergences.  A comprehensive account of the notation is given
+in~\cite{Roscoe2010}.
+
+A distinctive feature of CSP is its treatment of refinement~(as opposed to
+bisimulation), which is convenient for reasoning about program correctness,
+due to its treatment of nondeterminism and divergence.  A variety of semantic
+models capture different notions of refinement. The simplest model
+characterises a process by its possible \emph{traces}; the set $\trc(P)$
+denotes the sequences of (non-hidden) events in which $P$ can engage.  We say
+that a process \emph{$P$ is trace-refined by another process $Q$}, written $P
+\lessdet_T Q$, if $\trc(Q)\subseteq \trc(P)$.
+
+In fact, in every semantic model, subset containment is used to define refinement. The
+model we focus on first is the failures model, which captures both
+sequences of interactions and deadlock behaviour. A \emph{failure} of a process $P$
+is a pair $(s,X)$ containing a trace $s$ of $P$ and a \emph{refusal}:~a set $X$ of
+events in which $P$ may refuse to engage, after having performed the events of
+$s$. The failures model of a process $P$ records all its failures in a set
+$\fails(P)$.
+
+Semantic definitions specify, for each operator, how the traces or failures
+of the resulting process can be calculated from those of each operand. For
+example, for internal choice, $\fails(P\intchoice Q) =
+\fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for a
+comprehensive list of these definitions covering $\trc(P)$ and $\fails(P)$.
+
+Using the notation $P/s$ to denote the behaviour ot the process $P$ after
+having engaged into the events in the trace $s$, the set $\refs(P/s) \defs
+\{~X | (s,X) \in \fails(P)~\}$ contains the  refusals of $P$ after $s$.
+Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$
+is a failure of $P$ and $Y\subseteq X$, then $(s,Y)\in\fails(P)$ and
+$Y\in\refs(P/s)$ follows.
+
+For divergence-free processes, failures refinement, $P \lessdet_F Q$, is
+defined by $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
+$P \lessdet_F Q$ implies $(s,\varnothing)\in\fails(P)$ for all traces
+$s\in\trc(Q)$. So, for divergence-free processes, failures refinement implies
+traces refinement. Therefore, using the conformance relation $conf$ below
+%
+\begin{equation}\label{eq:conf}
+  Q\ conf\ P \defs \forall s\in \trc(P) \cap \trc(Q): \refs(Q/s)
+  \subseteq \refs(P/s),
+\end{equation}
+%
+failures refinement can be expressed by $\lessdet_T$ and $conf$ as proven
+in~\cite{DBLP:conf/icfem/CavalcantiG07}.
+%
+\begin{equation}%\label{thm:fref-tref-conf}
+\label{eq:failconf}
+(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
+\end{equation}
+%
+For finite processes, since refusals are subset-closed, $\refs(P/s)$ can be
+constructed from the set of \emph{maximal refusals}.
+%
+\begin{equation}
+\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
+\end{equation}
+%
+Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
+reconstruct the refusals in the set $\refs(P/s)$. % as described by
+%
+\begin{equation}\label{eq:refMaxRef}
+\refs(P/s) = \{ R'\in 2^\Sigma~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
+\end{equation}
+%
+Deterministic process states $P/s$ have exactly the one maximal refusal
+defined by $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
+$P/s$, that is, the events that $P/s$ may engage into. The maximal refusals
+in combination with the initials of a process express its degree of
+nondeterminism as illustrated by the next example.
+%
+\begin{example}
+\label{ex:nondetdegree} $P = (\Stop \intchoice Q)$ has maximal refusals
+$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ refuses to engage in any event,
+and this is carried over to $P$ by the internal choice. However, $P$ is
+distinguished from $\Stop$ by its initials, which are defined by $[P]^0 = [
+\Stop \intchoice Q]^0 = [Q]^0$. So $P$ may engage nondeterministically in any
+initial event of $Q$, but also refuse everything, due to internal selection
+of $\Stop$. Assuming an alphabet $\Sigma = \{a,b,c,d\}$, the process
+%
+$$Q = (e:\{a,b\} \then \Stop) \intchoice (e:\{c,d\} \then \Stop)$$
+%
+has maximal refusals $\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials
+$[Q]^0=\Sigma$. In contrast to $P$, nondeterminism is reflected here by
+two maximal refusals. \xbox
+\end{example}
+
+% =========================================================================
+\paragraph{Normalised Transition Graphs for CSP Processes} \label{sec:ntg}
+
+As shown in~\cite{Roscoe:1994:chapter}, the failures semantics of any
+finite-state CSP process $P$ can be represented by a \emph{normalised
+transition graph} $G(P)$ defined by a tuple
+$$
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+$$
+with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
+partial \emph{transition function} $t$ maps a node $n$ and an event
+$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
+$t$, then there is a transition, that is, an outgoing edge, from $n$ with
+label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
+the fact that $t$ is a function.
+
+The graph construction in~\cite{Roscoe:1994:chapter} implies that all nodes
+$n$ in $N$ are reachable by sequences of edges labelled by $e_1\dots e_k$ and
+connecting states $\ii n,n_1,\dots,n_{k-1},n$, such that
+\[
+n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
+n= t(n_{k-1},e_k).
+\]
+%
+By construction, $s\in\Sigma^*$ is a trace of $P$, if, and only if, there is
+a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
+the events in $s$ in the order they appear. In analogy to $\trc(P)$, we use
+the notation $\trc(G(P))$ for the set of finite, initialised paths through
+$G(P)$, each path represented by its finite sequence of edge labels. We note
+that $\trc(P) = \trc(G(P))$. Since $G(P)$ is normalised, there is a unique
+node reached by following the events from $s$ one by one, starting in $\ii
+n$. Therefore, $G(P)/s$  is also well defined.
+
+By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
+as labels in any outgoing transitions.
+$$
+[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
+$$
+The graph construction from~\cite{Roscoe:1994:chapter} guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
+traces $s$ of $P$.
+
+The total function $r$ maps each node $n$ to a non-empty set of  (possibly
+empty) subsets of $\Sigma$. The graph construction guarantees that
+$r(G(P)/s)$ represents the maximal refusals of $P/s$ for all $s\in\trc(P)$.
+As a consequence,
+\begin{equation}\label{eq:refsAndR}
+(s,X)\in\fails(P) \iff s\in\trc(G(P)) \wedge \exists R\in r(G(P)/s): X\subseteq R,
+\end{equation}
+so $G(P)$ allows us to re-construct the failures semantics of $P$.
+%\fixme{alcc: Why not leave the commented out material in the report?}
+% Commented out just to shorten the report.
+%To see that this approach works only for finite CSP processes, we consider
+%the example where $\Sigma$ is infinite. In this case,
+%$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
+%calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
+%defined above. As with refusals, we also use the transition graph-oriented
+%notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
+%associated with graph state $n$: if $n$ is the state reached in the
+%transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
+%= \maxrefs(P/s)$.
+
+% Commented out just to shorten the report.
+%Well-formed normalised transition graphs must not refuse an event of the
+%initials of a state in {\it every} refusal applicable in this state; more
+%formally,
+%%
+%\begin{equation}
+%\label{eq:wellformedg}
+%\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
+%\exists R\in r(n): e\not\in R
+%\end{equation}
+%%
+%By construction, normalised transition graphs reflect the \emph{failures
+%semantics} of finite-state CSP processes:~the traces $s$ of a process are
+%defined by the sequences of transitions associated with paths through its
+%graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
+%r(G(P)/s)$ represent the failures of $P$.
+
+% =========================================================================
+\paragraph{Acceptances} \label{sec:accs}
+
+When investigating tests for failures refinement, the notion of
+\emph{acceptances}, which is dual to refusals, is useful. While the
+original introduction of acceptances presented
+in~\cite[pp.~75]{Hennessy:1988:ATP:50497} was independent of refusals, we use
+the definition from~\cite[pp.~278]{Roscoe:1997:TPC:550448}.   A \emph{minimal
+acceptance} of a CSP process state $P/s$ is the complement of a maximal
+refusal of the same state. The set of minimal acceptances of $P/s$ is denoted
+by $\minaccs(P/s)$ and formally defined as
+%
+\begin{equation}\label{eq:accref}
+\minaccs(P/s) = \{ \Sigma - R~|~R\in \maxrefs(P/s)  \}
+\end{equation}
+%
+With this definition, a (not necessarily minimal) \emph{acceptance} of $P/s$ is a superset of some minimal acceptance and a subset of the initials $[P/s]^0$.
+Denoting the acceptances of $P/s$ by $\accs(P/s)$, this leads to the formal definition
+\begin{equation}
+\label{eq:accrefall}
+\accs(P/s)  = \{ B\subseteq [P/s]^0~|~\exists A\in\minaccs(P/s):  A \subseteq B   \}
+\end{equation}
+%
+Acceptances have the following intuitive interpretation.  If the behaviour of
+$P/s$ is deterministic, its only acceptance equals $[P/s]^0$, because $P/s$
+never refuses any of the events in this set. If $P/s$ is nondeterministic, it
+internally chooses one of its \emph{minimal acceptance sets} $A$ and never
+refuses any event in $A$, while {\it possibly} refusing the events from the
+set $[P/s]^0 - A$ and {\it always} refusing those in the set $\Sigma -
+[P/s]^0$.
+
+Exploiting (\ref{eq:accref}), the nodes of a normalised transition graph can
+alternatively be labelled with minimal acceptances; this captured the same
+information conveyed by maximal refusals. Since process states $P/s$ are
+equivalently expressed by states $G(P)/s$ of $P$'s normalised transition
+graph, we also write $\minaccs(G(P)/s)$ and note that (\ref{eq:refsAndR}) and
+(\ref{eq:accref}) imply
+\begin{equation}\label{eq:maxrefsminaccs}
+\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \} = \minaccs(P/s).
+\end{equation}
+%
+Given any non-diverging, non-terminating, finite-state process $P$, it can be
+re-con\-struc\-ted from its graph $G(P)$ with initial state $\ii
+n$ and transition function $t$, using $P$'s normalised syntactic
+representation~\cite[pp.~277]{Roscoe:1997:TPC:550448} specified as follows.
+\begin{eqnarray*}
+\text{normalised}(P) & = & P_N(\ii n)
+\\
+P_N(n) & = & \Intchoice_{A\in\minaccs(n)\cup \{ [n]^0 \} } e:A\then P_N(t(n,e))
+\end{eqnarray*}
+With this definition, it is established that $P$ is semantically equivalent
+to $\text{normalised}(P)$ in the failures semantics.
+
+% .....................................................................................
+ \begin{figure}[t]
+   %%\hspace*{-40mm}
+   \begin{center}
+     \includegraphics[trim=0cm 1.4cm 0cm 1.4cm,clip,scale=.6,%width=\textwidth
+                              ]{p.pdf}
+   \end{center}
+   %%\vspace*{-10mm}
+   \caption{Normalised transition graph of CSP process $P$ from Example~\ref{ex:a}.}
+   \label{fig:tga}
+ \end{figure}
+% .......................................................................................
+
+\begin{example}\label{ex:a}
+We consider the process $P$ in Example~\ref{example:CSP}; its transition
+graph $G(P)$ is shown in Fig.~\ref{fig:tga}. The process state
+$P/\varepsilon$ (where $\varepsilon$ denotes the empty trace) is represented
+as node 0, with $\{ a\}$ as the only minimal acceptance, since $a$ is never
+refused and no other events are accepted. Having engaged in $a$, the
+transition from node 0 leads to node 1 representing the process state $P/a
+= Q\intchoice R$. The internal choice induces several minimal acceptances
+derived from $Q$ and $R$. Since these processes accept their initial events
+in external choice, $Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$
+and $\{b,c\}$. We note that the event $c$ can never be refused, since it is
+contained in each minimal acceptance set.
+
+Having engaged in $c$, the next process state is represented by node 2. Due
+to normalisation, there is only a single transition satisfying $t(1,c) = 2$.
+This transition, however, can have been caused by either $Q$ or $R$ engaging
+into $c$, so node 2 corresponds to process state $Q/c \intchoice R/c = P
+\intchoice R$. This is reflected by the two minimal acceptance sets labelling
+node 2. From node~2, event $c$ leads to node~3. Since $P$ does not engage
+into $c$, the $R$-component of $P\intchoice R$ must have processed $c$, so
+node~3 corresponds to $R/c = R$, and so it is labelled by $R$'s minimal
+acceptance $\{b,c\}$. \xbox
+\end{example}
+%
+\noindent%
+Summarising, refinement between finite-state CSP processes $P, Q$ can be
+expressed using their normalised graphs
+\begin{eqnarray*}
+G(P) & = & ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))
+\\
+G(Q) & = & ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))
+\end{eqnarray*}
+%
+as established by the results  in the following lemma. There, result
+(\ref{eq:trcrefa}) reflects traces refinement in terms of graph traces;
+result (\ref{eq:failconf}) expresses failures refinement in terms of traces
+refinement and $conf$; result (\ref{eq:failrefa}) states how $conf$ can be
+expressed by means of the maximal refusal functions of the graphs involved;
+and result (\ref{eq:failrefb}) states the same in terms of the minimal
+acceptances that can be derived from the maximal refusal functions by means
+of (\ref{eq:maxrefsminaccs}).
+%
+\begin{lemma}
+  \label{lemma:tgtrcref}
+  \begin{eqnarray}
+  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
+  \label{eq:trcrefa}
+%  \\
+%  \label{eq:failconf}
+%  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
+  \\
+  \label{eq:failrefa}
+  Q\ conf\ P & \Leftrightarrow & \nonumber
+  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
+  \\ & & \tabd
+  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
+  \\
+  \label{eq:failrefb}
+   & \Leftrightarrow &
+    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
+   \\ & & \tabd
+  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
+ \end{eqnarray}
+  \xbox
+\end{lemma}
+%
+\begin{proof}
+To prove (\ref{eq:trcrefa}), we recall that $P\lessdet_T Q$ is defined as
+$\trc(Q)\subseteq \trc(P)$ and, moreover, $\trc(P)=\trc(G(P))$ and
+$\trc(Q)=\trc(G(Q))$. %Result (\ref{eq:failconf}) has been shown
+%in~\cite{DBLP:conf/icfem/CavalcantiG07}.
+%
+%\medskip
+%\noindent
+%To prove (\ref{eq:failconf}), we derive
+%\begin{argue}
+%P \lessdet_F Q
+%\\
+%\Leftrightarrow \fails(Q)\subseteq \fails(Q) & definition of $\lessdet_F$
+%\\
+%\Leftrightarrow \forall (s,R) \in  \fails(Q): (s,R) \in\fails(P)
+%& property of $\subseteq$
+%\\
+%\Leftrightarrow \forall (s,\varnothing) \in  \fails(Q): (s,\varnothing) \in\fails(P)
+%\wedge {}
+%\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
+%\\
+%& Definition of failures, refusals, subset closure
+%\\
+%\Leftrightarrow \trc(Q)\subseteq\trc(P)
+%\wedge {}
+%\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
+%\\
+%& Definition of failures and traces
+%\\
+%\Leftrightarrow \trc(Q)\subseteq\trc(P)
+%\wedge {}
+%\\\tabc \forall s\in\trc(Q)\cap\trc(P): \refs(Q/s) \subseteq\refs(P/s)
+%\\
+%& Property of $\cap$ and $\subseteq$
+%\\
+%\Leftrightarrow
+%P\lessdet_T Q \wedge Q\ conf\ P
+%& Definition of $\lessdet_T$ and $conf$
+%\end{argue}
+%
+To prove (\ref{eq:failrefa}), we derive
+\begin{argue}
+  Q\ conf\ P
+  \\
+  \Leftrightarrow \forall s\in\trc(P)\cap \trc(Q): \refs(Q/s)\subseteq\refs(P/s)
+  %\\
+  & Definition of $conf$  (\ref{eq:conf})
+  \\
+  \Leftrightarrow \forall s\in\trc(G(P))\cap \trc(G(Q)): \refs(Q/s)\subseteq\refs(P/s)
+  \\
+  & $\trc(P) = \trc(G(P))$, $\trc(Q) = \trc(G(Q))$
+  \\
+  \Leftrightarrow \forall s\in\trc(G(P))\cap \trc(G(Q)), R_Q\in r_Q(G(Q)/s):
+  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
+  \\ %
+  & Property of $r_P, r_Q$~(subset closure) and (\ref{eq:refMaxRef})
+\end{argue}
+
+\medskip
+\noindent
+Finally, (\ref{eq:failrefb}) follows from (\ref{eq:failrefa}) using (\ref{eq:accref})
+and the fact that $R_Q \subseteq R_P$ is equivalent to
+$\Sigma - R_P \subseteq \Sigma - R_Q$.
+\xbox
+\end{proof}
+
+% =========================================================================
+\paragraph{Reachability Under Sets of Traces} \label{sec:V} Given a finite-state CSP
+process $P$ and its normalised transition graph $G(P)$,
+%\[
+%G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+%\]
+we suppose that $V\subseteq\Sigma^*$ is a prefix-closed set  of sequences of
+events. By $t(\ii n,V)$ we denote the set
+\[
+t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
+\]
+of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$. The
+lemma below specifies a construction method for such sets $V$ reaching {\it
+every} node of $N$.
+
+\begin{lemma}
+\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
+graph $G(P)$. Let $V\subseteq\Sigma^*$ be a finite prefix-closed set of
+sequences of events. Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$,
+that is, $|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences
+from $V$, extended by any event of $\Sigma$.  Then $G(P)$ reaches at least
+$(k+1)$ nodes under $V\cup V.\Sigma$.
+\end{lemma}
+\begin{proof}
+Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
+there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
+with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
+$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
+$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
+Note, however, that it is not necessarily the case that $s_1\in V$. Since
+$G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such that
+$G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
+$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
+= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
+n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
+node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
+\end{proof}
+
+% =========================================================================
+\paragraph{Graph Products} \label{sec:GP} For proving our main theorems, it is
+necessary to consider the \emph{product} of normalised transition graphs. We
+need this only for the investigation of corresponding traces in reference
+processes and processes for SUTs. So, the labelling of nodes with maximal
+refusals or minimal acceptances are disregarded in the product construction.
+We consider two normalised transition graphs
+\[
+G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
+\]
+over the same alphabet $\Sigma$. Their product is defined by
+%
+\begin{eqnarray}
+G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
+\\
+\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma | (n_1,e)\in\dom~t_1\wedge (n_2,e) \in\dom~t_2 \}
+\\
+t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
+\end{eqnarray}
+%
+The following lemma is used in the proof of our main theorem.
+%
+\begin{lemma}\label{lemma:reachproduc}
+If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
+$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace
+%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
+of maximal length $(pq-1)$.
+\end{lemma}
+\begin{proof}
+The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
+reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
+$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
+all of its reachable states (there are at most $pq$ of them) under
+$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
+$V'$ is $(pq-1)$.
+\xbox
+\end{proof}
+%
+This concludes our presentation of CSP and of results regarding its
+semantics that are used in the next section.
+
+% =========================================================================
+\subsection{Minimal Hitting Sets}
+\label{sec:hit}
+% =========================================================================
+
+\paragraph{Definition} The main idea of the underlying test strategy for failures
+refinement is based on solving a \emph{hitting set problem}. Given a finite
+collection of finite sets ${\cal C} = \{ A_1,\dots,A_n\}$, such that each
+$A_i$ is a subset of a universe $\Sigma$, a \emph{hitting set}
+$H\subseteq\Sigma$ is a set satisfying the following property.
+%
+\begin{equation}
+  \label{eq:hit}
+  \forall A\in {\cal C}: H\cap A \neq\varnothing.
+\end{equation}
+%
+A \emph{minimal hitting set} is a hitting set that cannot be further reduced
+without losing the characteristic property (\ref{eq:hit}). By $\minhits({\cal
+C})$ we denote the collection of minimal hitting sets for a collection ${\cal
+C}$. For the pathological case where $C$ contains an empty set,
+$\minhits({\cal C})$ is also empty. The problem of determining minimal hitting sets is %%% \cite{Book1975-BOOKRM,5533149}
+NP-hard~\cite{5533149}. We see below, however, that using minimal hitting
+sets, we can reduce the effort of testing for failures refinement from a
+factor of $2^{|\Sigma|}$ to a factor that equals the number of minimal
+hitting sets.
+
+% -----------------------------------------------------------------------------
+\paragraph{Minimal Hitting Sets of Minimal Acceptances} In this article, we are
+interested in the minimal hitting sets of minimal acceptances; for these, the
+abbreviated notation
+$%\[ \
+\minhits(P/s) = \minhits(\minaccs(P/s))
+$ %\]
+is used. The minimal hitting sets
+of minimal acceptances may be alternatively characterised by means of the
+failures of a process as is done in~\cite{DBLP:conf/icfem/CavalcantiG07}. To
+this end, in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for
+any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
+\begin{equation}\label{eq:minsub}
+\minsub({\cal C}) = \{ A\in {\cal C}~|~\forall B\in {\cal C}:
+B\subseteq A\Rightarrow B = A \}.
+\end{equation}
+The collection $\minsub({\cal C})$ contains all those sets of
+${\cal C}$ that are not true subsets of other members of ${\cal C}$. With
+this definition, the relation between failures and minimal hitting sets of
+minimal acceptances is established in the following lemma.
+%
+\begin{lemma}
+\label{lemma:failureshittingsets}
+For any trace $s$ of CSP process $P$, define
+$%\[
+{\cal A}_s = \{ A\subseteq \Sigma~|~(s,A)\not\in\fails(P)  \}.
+$ %\]
+Then
+$ %\[
+\minhits(P/s) = \minsub {\cal A}_s
+$ %\]
+for all traces $s$ of $P$.
+\end{lemma}
+\begin{proof}
+We derive
+\begin{argue}
+(s,A)\not\in \fails(P)
+\\
+\Leftrightarrow A\not\in Ref(P/s)
+\\
+\Leftrightarrow  \forall R\in maxRef(P/s): A\not\subseteq R
+\\
+\Leftrightarrow   \forall B\in minAcc(P/s): A\not\subseteq (\Sigma\setminus
+B)& this follows from (\ref{eq:accref})
+\\
+\Leftrightarrow  \forall B\in minAcc(P/s): A\cap B\neq \varnothing
+\\
+\Leftrightarrow  A \text{ is a (not necessarily minimal)}
+              \text{ hitting set of } minAcc(P/s)
+\end{argue}
+%
+Specialising this derivation valid for arbitrary $A$ satisfying $(s,A)\not\in {\text{failures}}(P)$ to
+minimal $A$ with this property proves the statement of the lemma.
+\xbox
+\end{proof}
+
+% -------------------------------------------------------------------------
+\paragraph{Minimal Hitting Sets of Normalised Transition Graphs} Since, as
+previously explained, minimal acceptances can be used to label the nodes of a
+normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$ by
+(\ref{eq:maxrefsminaccs}), the notation of minimal hitting sets also carries
+over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe
+that
+\begin{equation}
+\label{eq:minhitsGP}
+\minhits(G(P)/s) = \minhits(P/s) \ \text{for all}\ s\in \trc(P).
+\end{equation}
+
+% ------------------------------------------------------------------------
+\paragraph{Characterisation of $conf$ by Minimal Hitting Sets} The following lemma
+establishes that the $conf$ relation specified in (\ref{eq:conf}) can be
+characterised by means of minimal acceptances and their minimal hitting sets.
+%
+\begin{lemma}
+\label{lemma:hseta}
+Let $P, Q$ be two finite-state CSP processes.
+%% satisfying $P\lessdet_T Q$.
+For each $s\in\trc(P)$,
+let $\minhits(P/s)$ denote the
+collection of all minimal hitting sets of $\minaccs(P/s)$.
+Then the following statements are equivalent.
+\begin{enumerate}
+\item $Q\ conf\ P$
+
+\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \minhits(P/s)$, $H$ is
+a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
+\end{enumerate}
+\end{lemma}
+\begin{proof}
+We apply (\ref{eq:maxrefsminaccs}) and (\ref{eq:minhitsGP}), so that
+$\minaccs(P/s)$ and $\minaccs(G(P)/s)$, as well as $\minhits(P/s)$ and
+$\minhits(G(P)/s)$ are used interchangeably. For showing ``$(1) \Rightarrow
+(2)$'', we assume $Q\ conf\ P$ and $s\in\trc(P)\cap \trc(Q)$.
+Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}), states that
+$%\[
+\forall A_Q\in\minaccs(G(Q)/s):
+\exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
+$. %\]
+Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap A_P\neq\varnothing$
+for all minimal acceptances $A_P$, but also $H\cap A_Q\neq\varnothing$ for
+every minimal acceptance $A_Q$, because $A_P\subseteq A_Q$ for at least one
+$A_P$. So, each $H \in \minhits(P/s)$ is also a hitting set for
+$\minaccs(G(Q)/s)$ as required.
+
+To prove ``$(2) \Rightarrow (1)$'', we assume that (2) holds, but that $P\
+conf\ Q$ does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
+(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
+\[
+\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s):
+A_P\not\subseteq A_Q \qquad (*)
+\]
+Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
+Define
+$%\[
+\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
+$ %\]
+Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
+$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
+intersection with $A$. Minimising $\overline H$ yields   a minimal hitting
+set $H\in \minhits(P/s)$ which is {\it not} a hitting set of
+$\minaccs(G(Q)/s)$, a contradiction to Assumption~2.
+This completes the proof of the lemma.
+\xbox
+\end{proof}
+%
+We note that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
+Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and
+this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we
+conclude that $\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{
+\varnothing \}$. This clarifies that $\minhits(P/s)$ is empty if, and only
+if, $\minaccs(P) = \{ \varnothing \}$. The proof of Lemma~\ref{lemma:hseta}
+covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
+$\minhits(P/s) = \varnothing$. Trivially,
+\begin{equation}
+\label{eq:minhitminaccempty}
+\minaccs(P/s) = \{ \varnothing \} \iff \minhits(P/s) = \varnothing
+\end{equation}
+holds.
+
+% =========================================================================
+\subsection{Sperner Families}
+\label{sec:sperner}
+% =========================================================================
+
+In preparation for complexity results presented in
+Section~\ref{sec:complexity}, we consider how many minimal hitting sets can
+maximally exist for a collection of minimal acceptances. To this end, the
+following definitions and results are useful.
+
+A \emph{Sperner Family} is a collection ${\cal S} \subseteq 2^\Sigma$ of sets
+from a given finite universe $\Sigma$ that do not contain each other, that
+is, $H_1 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair
+$H_1\neq H_2\in {\cal S}$. Specialising antichains known from partial orders
+to finite sets partially ordered by $\subseteq$ results in Sperner families.
+Given an {\it arbitrary} collection of subsets ${\cal C}\subset 2^\Sigma$,
+the sub-collection $\minsub({\cal C})$ defined in (\ref{eq:minsub}) is a
+Sperner Family contained in ${\cal C}$.
+
+\newpage
+\noindent
+We further observe that
+\begin{itemize}
+\item the maximal refusals of a CSP process state,
+\item the minimal acceptances of a CSP process state, and
+\item the minimal hitting sets of a given collection of sets
+\end{itemize}
+are Sperner families. Moreover, given any finite alphabet $\Sigma$ with
+$\card{\Sigma} = n$, every collection $S$ of subsets with identical
+cardinality $k \le n$ is a Sperner family,  because $A_1, A_2\in S \wedge A_1
+\subseteq A_2\wedge \card{A_1} = \card{A_2}$ implies $A_1 = A_2$. Given  any
+Sperner Family $S$ of $\Sigma$, $S$ represents the minimal acceptances in the
+initial state of the CSP process
+%\[
+$P = \Intchoice_{A\in S}\big( e:A \then P(e) \big)$.
+%\]
+
+The cardinality of Sperner Families is determined by the following theorem.
+\begin{theorem}[Sperner's Theorem~\cite{sperner_satz_1928}]
+\label{th:sperner} Given a  Sperner family ${\cal S}$ over an $n$-element
+universe $\Sigma$, its cardinality $\card{{\cal S}}$ is bound by
+\[
+\card{{\cal S}} \le \binom{n}{\lfloor\frac{n}{2}\rfloor}.
+\]
+The upper bound is reached if, and only if, one of the following cases apply:
+\begin{enumerate}
+\item For even $n$, if ${\cal S}$ consists of all subsets of $\Sigma$ with
+    cardinality $n/2$;
+\item For odd $n$, if one of the following cases holds;
+\begin{enumerate}
+\item ${\cal S}$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/2$;
+    or
+\item ${\cal S}$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/2$.
+\end{enumerate}
+\end{enumerate}
+\xbox
+\end{theorem}
+It is shown in Section~\ref{sec:complexity} that this upper bound can
+actually be reached by the Sperner Family containing the hitting sets
+associated with the minimal acceptances of a CSP process state.
+
+% ==========================================================================
+\section{Finite Complete Test Suites for CSP Failures Refinement}
+\label{sec:finitecompletefails}
+% ==========================================================================
+
+Here, we define our notion of tests for failures refinement, and then prove
+completeness of our suite. Finally, we study to complexity of our approach by
+identifying a bound on the number of tests we need in a complete suite.
+
+% ==========================================================================
+\subsection{Test Cases for Verifying CSP Failures Refinement}
+
+% -------------------------------------------------------------------------
+\paragraph{Test Definition and Basic Properties}
+In the domain of process algebras, test cases are typically represented by
+processes interacting concurrently with the
+SUT~\cite{Hennessy:1988:ATP:50497}. Considering an (unknown) process that
+represents the behaviour of the SUT, we say that tests synchronise with the
+process for the SUT over its visible events and use some additional events
+outside the SUT process's alphabet to express whether the test execution
+passed or failed.
+%%%, or if no verdict could be obtained.
+
+For a given reference process $P$, its normalised transition graph
+$$
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
+$$
+and each integer $j\ge 0$, we define a test for failures refinement as shown
+below.
+%
+\begin{eqnarray}
+\label{eq:UFP}
+U_F(j) & = & U_F(j,0,\ii n)
+\\
+U_F(j,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
+\label{eq:ufa}
+\\ & & \extchoice \nonumber
+\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
+\label{eq:ufb}
+\\ & & \extchoice \nonumber
+\\ & & (k < j) \& \big(e:[n]^0   \then U_F(j,k+1,t(n,e) \big)
+\label{eq:ufc}
+\\ & & \extchoice \nonumber
+\\ & & (k = j \wedge \minhits(n)\neq\varnothing) \& \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
+\label{eq:ufd}
+\end{eqnarray}
+
+% --------------------------------------------------------------------------
+\paragraph{Explanation of the Test Definition} A test \pagebreak is performed by running
+$U_F(j)$ concurrently with any SUT process $Q$, synchronising over $\Sigma$.
+So, a \emph{test execution} is a trace of the concurrent process
+%\[
+$Q\parallel[\Sigma] U_F(j)$.
+%\]
+
+It is assumed that the events $\efail$ and $\epass$, indicating verdicts FAIL
+and PASS for the test execution, are not included in $\Sigma$. Since we
+assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
+or $\epass$ always become visible, if they are the only events $U_F(j)/s$ is
+ready to engage in: if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the
+occurrence of these events can never be blocked due to a livelock in $Q$
+occurring in the same step of the execution.
+
+The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(j)$) if, and
+only if, {\it every} execution of $Q\parallel[\Sigma] U_F(j)$ terminates with
+the event $\epass$. This can also be  expressed by means of a failures
+refinement as defined below.
+%
+\begin{equation}
+\label{eq:passF}
+Q\ \pass\ U_F(j) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma
+\end{equation}
+%
+This type of pass relation is often called \emph{must test}, because every
+test execution must end with the $\epass$
+event~\cite{Hennessy:1988:ATP:50497}.
+
+It is necessary to use failures refinement in the definition above, and not
+just traces refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ may have
+the same visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
+Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
+refuse $\epass$, due to a deadlock occurring when a faulty SUT process
+executes concurrently with $U_F(j,k,n)$ executing branch~(\ref{eq:ufd}), when
+the guard condition $(k = j\wedge \minhits(n)\neq\varnothing)$ evaluates to
+$\ist$. This is explained further in the next paragraphs. Alternatively, a
+faulty SUT $Q$ might internally deadlock after a trace $s$ where $\#s < j$
+and $\minhits(G(P)/s) \neq \varnothing$, so that the process
+$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks as well.
+
+Intuitively, $U_F(j)$ is able to perform any trace $s$ of $P$, up to a length
+$j$. If, after having already run through $s$ with $\#s \le j$, the SUT
+accepts an event outside the initials of $P/s$
+ (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(j)/s$),
+the test immediately terminates with FAIL-event $\efail$. This is handled by
+the branch (\ref{eq:ufa}) of the external choice. %in the process
+%$U_F(j,\#s,n)$ defined above.
+
+If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice, this is
+revealed by $\minhits(G(P)/s) = \varnothing$ (recall
+(\ref{eq:minhitminaccempty}) and Lemma~\ref{lemma:ufproperties}). In this
+case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
+choice in $U_F(j,\#s,G(P)/s)$). If $P/s$ may also nondeterministically engage
+into events, branch (\ref{eq:ufc}) is simultaneously enabled. If $Q/s$ is
+able to engage into an event in $\Sigma - [P/s]^0$, a test execution exists
+where $U_F(j,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the
+$\efail$ event.
+
+If the length of $s$ is still less than $j$, the test accepts any event $e$
+from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
+$U_F(j,\#s+1,G(P)/s.e)$ in branch~(\ref{eq:ufc}); this follows from
+Lemma~\ref{lemma:ufproperties}~(note that $G(P)/s.e = t(G(P)/s,e)$). A test
+of this type is called \emph{adaptive}, because it accepts any legal
+behaviour of the SUT, here any event from $[P/s]^0$, and adapts its
+consecutive behaviour to the event selected by the SUT, here
+$U_F(j,\#s+1,G(P)/s.e)$.
+
+Now suppose that a test execution has run through a trace $s\in\trc(P)$ of
+length $j$, so that $U_F(j)/s = U_F(j,j,n)$ with $n = G(P)/s$. If
+$\minhits(n)\neq \varnothing$, the test changes its behaviour:~instead of
+offering {\it all} legal events from $[n]^0$ to the SUT, it
+nondeterministically chooses a minimal hitting set $H\in \minhits(n)$ and
+only offers the events contained in $H$. If the SUT refuses to engage into
+some event of $H$, this reveals a violation of failures refinement:~according
+to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least one event
+of each minimal hitting set in $\minhits(n)$. Therefore, the test execution
+terminates with  $\epass$, only if such an event is accepted. Otherwise, it
+deadlocks, and the test fails.
+
+The specification of $U_F(j,k,n)$ implies that the test always stops after
+having engaged into a trace $s\in\trc(Q)$ of maximal length $j$ or $j+1$. If
+branch (\ref{eq:ufa}) is the last to be entered, the maximal length of $s$ is
+$j+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
+is the last to be entered, the maximal length of $s$ is $j$, and the
+execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
+entered, then there are two possibilities. The first is that the process
+accepts another event $e$ of some minimal hitting set $H\in\minhits(n)$ with
+$n = G(P)/s$ according to Lemma~\ref{lemma:ufproperties}. In this case, the
+final length of $s$ is $j+1$, and the execution terminates with $\epass$.
+Alternatively, the test execution $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks,
+the final length of $s$ is $j$, and the execution stops without a PASS or
+FAIL event. Such an execution is also interpreted as FAIL, because it reveals
+that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide
+\Sigma$.
+
+We observe that the number of possible executions of $Q\parallel[\Sigma]
+U_F(j)$ is finite, because the number of traces $s$ with maximal length
+$(j+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and $\minhits(n)$
+are finite. Moreover, we further recall that $\minhits(n)$ may be empty, in
+which case the indexed internal choice in (\ref{eq:ufd}) would be undefined.
+The guard in that branch, however, requires $\minhits(n)\neq\varnothing$, and
+branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
+
+The following lemma establishes relationships between $U_F(j)$ and the
+reference process $P$ from which it is derived.
+
+\begin{lemma}\label{lemma:ufproperties}
+If $s\in\trc(P)$ satisfies $\#s\le j$, then
+$s, s.e\in\trc(U_F(j))$ for all $e\in\Sigma$, and the following properties hold.
+\begin{eqnarray}
+\label{eq:ifpa}
+  &  & U_F(j)/s = U_F(j,\#s,G(P)/s)
+\\
+\label{eq:ifpb}
+e\not\in [P/s]^0 & \implies & U_F(j)/s.e = (\efail\then\Stop)
+\\
+\label{eq:ifpc}
+U_F(j)/s = U_F(j,\# s,n)  & \implies & [n]^0 = [P/s]^0
+\\
+\label{eq:ifpd}
+U_F(j)/s = U_F(j,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
+\end{eqnarray}
+\end{lemma}
+\begin{proof}
+We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$,
+the statement holds because $U_F(j)$ starts with the initial node $\ii n$ of
+$G(P)$. Suppose that the statement holds for all traces $s$ with length $\# s
+\le k < j$, so that $U_F(j)/s = U_F(j,\#s,G(P)/s)$. Now let $s.e$ be a trace
+of $P$, so that $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces
+$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(j,\#s,G(P)/s)$ can
+engage into $e$ by executing branch (\ref{eq:ufc}). Since $t$ is the
+transition function of $G(P)$ and $e\in [G(P)/s]^0$, $t(G(P)/s,e)$ is
+defined, and $t(G(P)/s,e) = G(P)/s.e$. So, the new recursion in branch
+(\ref{eq:ufc}) is so that $U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
+U_F(j,\#s+1,G(P)/s.e)$ as required.
+
+To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(j)/s
+= U_F(j,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
+$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
+$U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
+specification of this branch implies that $U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
+(\efail\then\Stop)$.
+
+Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because
+$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$. Finally, statement
+(\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because, according to
+(\ref{eq:minhitsGP}), $\minhits(G(P)/s) = \minhits(P/s)$ for all traces of
+$P$. \xbox
+\end{proof}
+%
+Note that it is not guaranteed for $U_F(j)$ to run through the traces $s,
+s.e$ in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for
+some prefix $u$ of $s$: in such a case, $U_F(j)$ may stop with a $\epass$
+event by entering branch (\ref{eq:ufb}). Therefore,
+Lemma~\ref{lemma:ufproperties} just states the existence of
+$U_F(j)$-executions $s, s.e$ satisfying the properties stated there.
+
+% -------------------------------------------------------------------------
+\paragraph{Complete Testing Assumption} As explained above, passing a test case
+$U_F(j)$ requires that none of the possible executions $(Q\parallel[\Sigma]
+U_F(j))$ stops after $\efail$ or stops without having produced the event
+$\epass$. Therefore, it is necessary to determine whether all possible
+executions have been covered in the repeated runs of $(Q\parallel[\Sigma]
+U_F(j))$. The theoretical completeness results are, therefore, based on a
+\emph{complete testing
+assumption}~\cite{hierons_testing_2004,DBLP:conf/icfem/CavalcantiG07}, which
+means that every possible behaviour of the SUT is performed after a finite
+number of test executions. In practice, this is realised by executing each
+test several times, recording the traces that have been performed, and using
+hardware or software coverage analysers to determine whether all possible
+behaviours of the SUT have been observed. Therefore, testing nondeterministic
+SUTs comes at the price of having to apply some grey-box testing techniques.
+%enabling us to decide whether all SUT behaviours have been observed.
+
+% =========================================================================
+\subsection{A Finite Complete Test Suite for Failures Refinement}
+% =========================================================================
+
+A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a
+reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
+\lessdet_F\}$, and a fault domain ${\cal D}$, which is a set of CSP processes
+over $P$'s alphabet that may or may not conform to $P$. A test suite $\TS$ is
+called \emph{complete} with respect to fault model ${\cal F}$, if, and only
+if, the following conditions are fulfilled.
+\begin{description}
+\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
+\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
+then $Q$ fails at least one test in $\TS$.
+\end{description}
+%
+The following main theorem establishes the completeness of our test suite.
+
+% -------------------------------------------------------------------------
+\begin{theorem}\label{th:failurestest}
+Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
+normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
+D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
+whose transition graph has at most $q$ states with $q \ge p$. Then the test
+suite
+\[
+\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{with $U_F(j)$ as specified in (\ref{eq:UFP})}
+\]
+is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
+\end{theorem}
+% ------------------------------------------------------------------------
+%
+The proof of the theorem follows directly from the two lemmas below. The
+first lemma establishes that test suite $\TS_F$ is sound, and the second
+establishes that the suite is also exhaustive.
+%
+\begin{lemma}\label{lemma:mainfsound}
+A test suite $\TS_F$ generated from a CSP process $P$, as specified in
+Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
+$P\lessdet_F Q$.
+\end{lemma}
+\begin{proof}{~}We make two points in separate steps below. The
+first is that the test execution cannot reach branch~(\ref{eq:ufa}) and raise
+a $fail$ event.  The second is that it cannot deadlock without raising a
+$pass$ event. This case would also be interpreted as FAIL, since then
+$\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
+U_F(j))\hide \Sigma$.
+
+\paragraph{Step~1} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and $Q\
+conf\ P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq
+\trc(P)$, any adaptive test $U_F(j)$ running in parallel with $Q$ will always
+enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the
+external choice construction for $U_F(j,k,n)$. To see this, consider
+ $U_F(j,k,n) = U_F(j)/s$
+with $s\in\trc(Q)$. Lemma~\ref{lemma:ufproperties} implies $U_F(j,k,n) =
+U_F(j,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
+$[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
+entered in the parallel execution of $Q$ and $U_F(j)$, and the $fail$ event
+cannot occur.
+
+\paragraph{Step~2} For proving that a test execution can never deadlock without a
+$\epass$ event, it has to be shown that a test execution can neither block at
+branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}). These cases are
+considered separately below.
+
+\paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
+after having run through a trace $s$ with $\#s < j$. Since  $P\lessdet_T Q$
+by assumption, $s$ is a trace of $P$, thus $U_F(j)/s = U_F(j,\#s,G(P)/s)$
+according to Lemma~\ref{lemma:ufproperties}. Therefore, $U_F(j)/s$  can enter
+branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$. Since we assume that
+$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that $[G(P)/s]^0$ is
+{\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least one
+$e\in [G(P)/s]^0$ would be accepted by $Q/s$ and the test execution would not
+deadlock. Now suppose that $\minhits(G(P)/s) = \varnothing$. Then   branch
+(\ref{eq:ufb}) can be entered, and the test stops after $\epass$. Otherwise,
+if $\minhits(G(P)/s) \neq \varnothing$, let $H\in\minhits(G(P)/s)$. Since $H$
+contains only elements that are contained in some minimal acceptance of
+$P/s$, and all these minimal acceptances are subsets of $[G(P)/s]^0$, $H$ is
+a subset of $[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, is not a
+hitting set of $\minaccs(Q/s)$, also $H$ is not a hitting set of
+$\minaccs(Q/s)$. Now this is a contradiction to Lemma~\ref{lemma:hseta},
+since $Q\ conf\ P$ by assumption, so $H$ should also be a (not necessarily
+minimal) hitting set in $\minaccs(Q/s)$. This proves that the test execution
+cannot block  at branch (\ref{eq:ufc}) without being able to pass the test by
+entering branch (\ref{eq:ufb}).
+
+\paragraph{Step~2.2} Suppose that the execution blocks at
+branch (\ref{eq:ufd}) after having run through some $s\in\trc(Q)\subseteq
+\trc(P)$ with $\#s = j$. From Lemma~\ref{lemma:ufproperties} we know that
+$U_F(j)/s = U_F(j,k,n) = U_F(j,\#s,G(P)/s)$, so $\minhits(n) =
+\minhits(P/s)$. Branch (\ref{eq:ufb}) of $U_F(j,k,n)$ leads always to a PASS
+verdict and is taken if $\minhits(n) = \varnothing$. If $\minhits(n) \neq
+\varnothing$,  the assumption that $(Q\parallel[\Sigma] U_F(j))/s$ blocks at
+branch (\ref{eq:ufd}) implies that there exists some $H\in\minhits(n)$ that
+is not a hitting set of $\minaccs(Q/s)$. Again, by Lemma~\ref{lemma:hseta},
+this contradicts the assumption that $Q\ conf\ P$. As a
+consequence, the test execution can never deadlock at branch (\ref{eq:ufd})
+without entering branch (\ref{eq:ufb}) and passing the test.
+
+\bigskip \noindent%
+Note that the line of reasoning in this proof requires that $Q$ is free of
+livelocks, because otherwise a $\epass$ event might not become visible, due
+to unbounded sequences of hidden events performed by $Q$. \xbox
+\end{proof}
+%
+\begin{lemma}\label{lemma:mainfexhaustive}
+A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
+exhaustive for the fault model specified there.
+\end{lemma}
+\begin{proof}
+Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
+(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
+corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
+cases can be characterised as follows:
+\begin{description}
+\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
+\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a
+    minimal acceptance $A_Q$ of $\minaccs(Q/s)$, such that (see
+    Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}))
+\begin{equation}
+\label{eq:accsnotcontained}
+\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
+\end{equation}
+\end{description}
+It has to be shown for each of these cases that at least one test execution
+of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
+event or deadlocks. We do this by analysing the product graph of the
+reference process $P$ and the SUT process $Q$: any trace
+$s\in\trc(Q)\cap\trc(P)$ gives rise to a path labelled by the events of $s$
+through this product graph. Any error can be detected after running through
+such a trace and then either observing an event outside $[P/s]^0$ (the
+violation described by Case~1) or identifying an illegal acceptance $A_Q$ (as
+in Case~2). It is not guaranteed, however, that $s$ is short enough to be
+executed by one of the test cases $U_F(j)$ with $0\le j < pq$. So, it has to
+be shown that for any $s$ leading to an error situation, there exists a trace
+$u$ of maximal length $pq-1$ leading to the same error.
+
+\medskip
+\noindent {\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$,
+but $s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is
+a trace of every process. In this case, $s$ is also a trace of the product
+graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}, and $G/s =
+(G(P)/s,G(Q)/s)$ holds. The length of $s$ is not known, but from the
+construction of $G$,  we know that $G$ has at most $pq$ reachable states,
+because $G(P)$ has $p$ states, and $G(Q)$ has at most $q$ states. By
+Lemma~\ref{lemma:reachproduc}, $(G(P)/s,G(Q)/s)$ can be reached by a trace
+$u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
+function of $G$ implies that $u$ is also a trace of $P$ and $Q$, which means
+that $(G(P)/s,G(Q)/s) = (G(P)/u,G(Q)/u)$. Since test $U_F(pq-1)$ accepts all
+traces of $P$ up to length $pq-1$, $u$ is also a trace of this test, and, by
+construction and by Lemma~\ref{lemma:ufproperties}, $U_F(pq-1)/u =
+U_F(pq-1,\#u,G(P)/u)$. Since $s.e\not\in\trc(P)$, $e$ is an element of
+$\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. Hence, in at least one execution,
+$U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with this
+event $e$, so that the test fails. Again, the assumption of non-divergence of
+$Q$ is needed for this conclusion.
+
+\medskip
+\noindent {\bf Case~2.} We note again that $s$ is a trace of the product
+graph $G$, but we do not know its length. Again, by
+Lemma~\ref{lemma:reachproduc}, the state $G/s$ can be reached by a trace
+$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider the test
+$U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of
+Lemma~\ref{lemma:ufproperties}. $U_F(\#u)$ can always perform branch
+(\ref{eq:ufc}) until the trace $u$ has been completely processed.
+ $U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
+only:~(\ref{eq:accsnotcontained}) implies that $P/s$ has at least one
+non-empty minimal acceptance. By (\ref{eq:minhitminaccempty}) this is
+equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and we
+observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
+consequence, branch (\ref{eq:ufb}) cannot be taken because its guard
+condition evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$. The guard condition
+$(k < j)$ for branch (\ref{eq:ufc}) evaluates to $\isf$ for
+$U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
+always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and
+therefore fails for the execution where $Q/u$ selects the minimal acceptance
+$A_Q$ as specified in (\ref{eq:accsnotcontained}) and $U_F(\#u,\#u,G(P)/u)$
+selects a minimal hitting set $H\in\minhits(P/u)$ that has an empty
+intersection with $A_Q$. The existence of such an $H$ is guaranteed because
+of Lemma~\ref{lemma:hseta}. As a consequence, $(Q\parallel[\Sigma]U_F(\#
+u))/u$ cannot produce the $\epass$ event in this execution; this means that
+the test fails. The complete testing assumption guarantees that this
+execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed
+sufficiently often. This concludes the proof. \xbox
+\end{proof}
+%
+Our notion of test can be specialised to deal with traces refinement~(see
+Section~\ref{sec:finitecomplete}). We next present an example.
+
+% ==========================================================================
+
+% =======================================================================
+\section{Testing for Failures Refinement -- an Example}
+\label{sec:case}
+% =======================================================================
+
+Generating the test cases $U_F(p)$ specified in  (\ref{eq:UFP}) for the reference
+process $P$ discussed in Example~\ref{example:CSP},
+results in the   instantiations of initials, minimal hitting sets, and
+transition function shown in Fig.~\ref{fig:initialsminhitstrans};
+this can be directly derived from $P$'s normalised
+transition graph with nodes $N =\{0,1,2,3\}$ displayed in Fig.~\ref{fig:tga}.
+
+\begin{figure}[ht]
+\footnotesize
+\begin{center}
+\begin{minipage}{0.2\textwidth}
+	 \begin{eqnarray*}
+{ }[0]^0 & = & \{ a \} \\
+{ }[1]^0 & = & \{ a,b,c \} \\
+{ }[2]^0 & = & \{ a,b,c \} \\
+{ }[3]^0 & = & \{ b,c \}
+\end{eqnarray*}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.33\textwidth}
+	 \begin{eqnarray*}
+\minhits(0) & = & \{ \{ a\} \} \\
+\minhits(1) & = & \{ \{a,b\}, \{c\} \} \\
+\minhits(2) & = & \{  \{a,b\},  \{a,c\}\} \\
+\minhits(3) & = & \{ \{ b\}, \{c\} \}
+\end{eqnarray*}
+
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.2\textwidth}
+	 \begin{eqnarray*}
+t(0,a) & = & 1 \\
+t(1,a) & = & 0 \\
+t(1,b) & = & 0 \\
+t(1,c) & = & 2
+\end{eqnarray*}
+	\end{minipage}
+		\hfill
+	\begin{minipage}{0.2\textwidth}
+	 \begin{eqnarray*}
+t(2,a) & = & 1 \\
+t(2,b) & = & 0 \\
+t(2,c) & = & 3 \\
+t(3,b) & = & 0 \\
+t(3,c) &  =& 3
+\end{eqnarray*}
+	\end{minipage}
+\caption{Initials, minimal hitting sets, and transition function of the normalised transition graph displayed in Fig.~\ref{fig:tga}.}
+\label{fig:initialsminhitstrans}
+\end{center}
+\normalsize
+\end{figure}
+
+% .....................................................................................
+ \begin{figure}
+ %%\hspace*{-40mm}
+ \begin{center}
+\includegraphics[trim=0cm 1.4cm 0cm 1.4cm,clip,scale=.6,%width=\textwidth
+                 ]{z.pdf}
+\end{center}
+%%\vspace*{-10mm}
+\caption{Normalised transition graph of faulty implementation $Z$  from Example~\ref{ex:uf1tests}.}
+ \label{fig:tgZ}
+ \end{figure}
+% .......................................................................................
+
+\begin{example}
+\label{ex:uf1tests} Consider the following implementation $Z$ of process $P$
+from Example~\ref{example:CSP} that is erroneous from the point of view of
+failures refinement. In the specification of $Z$, it is assumed that $r_{max}\ge 0$.
+\begin{eqnarray*}
+  Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
+  \\
+  Q_1 & = & a\then Z \extchoice c\then Z
+  \\
+  R_1(r_{max},k) & = & (k < r_{max}) \& \big(b\then Z \extchoice  c\then R_1(r_{max},k+1)\big)
+  \\ & & \extchoice
+  \\ & & (k = r_{max}) \& \big(b\then Z \intchoice c\then R_1(r_{max},r_{max})\big)
+\end{eqnarray*}
+It can be checked with FDR that $Z$ is trace-equivalent to $P$. While $k <
+r_{max}$, $Z$ also accepts the same sets of events as $P$. When
+$R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$, however,
+$R_1(r_{max},k)$ makes an internal choice, instead of offering an external
+choice, so $P\not\lessdet_F Z$. Fig.~\ref{fig:tgZ} shows the normalised
+transition graph of $Z$ for $r_{max} = 3$.
+
+Running the test $U_F(j)$ against $Z$ for $j=0,\dots,19$ ($G(P)$ has $p = 4$
+states and $G(Z)$ has $q=5$, so $pq-1=19$ is the index of the last test to
+be executed  according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
+U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$, because after execution of
+the trace
+\[
+s = a.c.c.c, \qquad\text{(note that $G(P)/s = \text{node}\ 3$ according to Fig.~\ref{fig:tga})},
+\]
+the test $U_F(4)$ offers hitting sets from $\minhits(3) = \{\{b\},\{c\}\}$
+in branch (\ref{eq:ufd}). Therefore,  there exists one test execution
+where $Z/s$ accepts only $\{b\}$ due to the internal choice (note
+from Fig.~\ref{fig:tgZ} that
+$G(Z)/s = \text{node}\ 4$), while $U_F(4)/s$
+only offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for
+branch (\ref{eq:ufa}).  As a consequence, this execution of
+$(Z\parallel[\Sigma] U_F(4))/s$ deadlocks, and the $pass$ event cannot be
+produced. Another failing execution arises if $Z/s$ chooses to accept only
+$\{c \}$, while $U_F(4)/s$ choses to accept only $\{a,b\}$. Therefore,
+$%\[
+(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F(4)) \hide\Sigma,
+$ %\]
+and the test fails. \xbox
+\end{example}
+
+% ======================================================================
+
+% ==========================================================================
+\section{Finite Complete Test Suites for CSP Traces Refinement}
+\label{sec:finitecomplete}
+% ==========================================================================
+
+For establishing traces refinement, the following class of adaptive test
+cases are used for a given reference process $P$ and  integers $j \ge 0$.
+Just as for the tests developed  in Section~\ref{sec:finitecompletefails} to
+verify failures refinement, the tests for traces refinement are derived from
+the reference model's transition graph
+$$
+G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)).
+$$
+In contrast to the tests for failures refinement~(\ref{eq:UFP}), however, we
+do not need to check the SUT with respect to its acceptance of hitting sets.
+Therefore, these do not occur in the specification of the test cases below.
+We use the condition on acceptances $\minaccs(n) = \{ \varnothing \}$ instead
+of the condition on hitting sets $\minhits(n) = \varnothing$ in branch
+(\ref{eq:utb}). From (\ref{eq:minhitminaccempty}) we know that these
+conditions are equivalent, but, with the use of $\minaccs(n) = \{ \varnothing
+\}$, \pagebreak we make it unnecessary to calculate hitting sets for
+generating these tests from $G(P)$, which is expensive.
+
+%\begin{eqnarray}
+%U_T(p) & = & U_T(p,\varepsilon)
+%\\
+%U_T(p,s) & = & \big(\Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
+%\label{eq:uta}
+%\\ & & \extchoice \nonumber
+%\\ & & (\minaccs(P/s) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
+%\label{eq:utb}
+%\\ & & \extchoice \nonumber
+%\\ & & (\#s < p) \& \big(\Extchoice e:[P/s]^0 @ e \then U_T(p,s.e) \big)
+%\label{eq:utc}
+%\\ & & \extchoice \nonumber
+%\\ & & (\#s = p) \& \big( \epass\then \Stop  \big)
+%\label{eq:utd}
+%\end{eqnarray}
+
+\begin{eqnarray}
+U_T(j) & = & U_T(j,0,\ii n)
+\\
+U_T(j,k,n) & = & \big(  e:(\Sigma - [n]^0)   \then \efail\then \Stop \big)
+\label{eq:uta}
+\\ & & \extchoice \nonumber
+\\ & & (\minaccs(n) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
+\label{eq:utb}
+\\ & & \extchoice \nonumber
+\\ & & (k < j) \& \big( e:[P/s]^0   \then U_T(j,k+1,t(n,e)) \big)
+\label{eq:utc}
+\\ & & \extchoice \nonumber
+\\ & & (k = j) \& \big( \epass\then \Stop  \big)
+\label{eq:utd}
+\end{eqnarray}
+%
+It is easy to see that the tests $U_T(j)$ satisfy the properties
+\begin{eqnarray}
+\label{eq:ifpaT}
+  &  & U_T(j)/s = U_T(j,\#s,G(P)/s)
+\\
+\label{eq:ifpbT}
+e\not\in [P/s]^0 & \implies & U_T(j)/s.e = (\efail\then\Stop)
+\end{eqnarray}
+proven in Lemma~\ref{lemma:ufproperties} for $U_F(j)$ for traces
+$s\in\trc(P)$ with $\#s \le j$.
+
+%The difference between adaptive tests $U_T(p)$ for trace refinement and
+%$U_F(p)$ for failures refinement consists in the fact that the former do not
+%``probe'' the SUT with respect to minimal sets of events to be accepted
+%without blocking.
+
+% ==========================================================================
+Since the test $U_T(j)$ never blocks any event of an SUT process $Q$ before
+terminating, the pass criterion, defined below, can be based on
+trace instead of failures refinement as required in (\ref{eq:passF}).
+%
+\begin{equation}
+\label{eq:passT}
+Q\ \pass\ U_T(j) \defs (\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(j)) \hide \Sigma
+\end{equation}
+%
+If the SUT process $Q$ deadlocks after a trace $s$, and in this case the
+reference process $P$ is also in a state where deadlock is possible, this is
+captured by the fact that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$.
+Therefore, branch (\ref{eq:utb}) of a test case execution state $U_T(j,k,n)$
+with $\#s = k \le j$   can be entered and the  test execution terminates with
+$\epass$. If, however, $Q$ blocks after a trace $s'$ and the reference
+process satisfies $\minaccs(P/s') \neq\varnothing$, branch (\ref{eq:utb})
+cannot be taken, and the test execution stops without producing $\epass$ or
+$\efail$. In contrast to the test for failures refinement, this is
+interpreted here as a successful test execution, because unexpected blocking
+of the SUT does not violate the trace-refinement relation, as long as all
+traces executed by the SUT are traces of the reference process. In
+particular, if neither $\epass$ nor $\efail$ is ever produced, so that
+$(Q\parallel[\Sigma] U_T(j)) \hide \Sigma = \Stop$, the test passes, because
+$(\epass\then\Stop) \lessdet_T \Stop$ holds.
+
+The existence of complete, finite test suites is expressed in analogy to
+Theorem~\ref{th:failurestest}. A noteworthy difference is that the complete
+suite for traces refinement just needs the single adaptive test case
+$U_T(pq-1)$, while failures refinement requires the execution of $\{
+U_F(0),\dots,U_F(pq-1)\}$. The reason is that $U_T(pq-1)$ identifies trace
+errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for
+erroneous acceptances at the end of each trace of length $(pq -1)$.
+%
+% -------------------------------------------------------------------------
+\begin{theorem}\label{th:tracetest}
+Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
+normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
+D}$ as the set of all non-terminating, divergence-free CSP processes over alphabet $\Sigma$,
+whose transition graph has at most $q$ states with $q \ge p$. Then the test
+suite
+\[
+\TS_T = \{ U_T(pq-1)   \}
+\]
+is complete with respect to ${\cal F} = (P,\lessdet_T,{\cal D})$.
+\xbox
+\end{theorem}
+% -------------------------------------------------------------------------
+%\begin{proof}
+%The theorem follows directly
+%from Step~1 in the proof of Lemma~\ref{lemma:mainfsound} and
+%Case~1 in the proof of Lemma~\ref{lemma:mainfexhaustive}.
+%\xbox
+%\end{proof}
+%%
+%Examples are provided after our discussion of size of the test suites.
+%
+As for Theorem~\ref{th:failurestest}, the proof is structured in two lemmas,
+the first ensuring soundness, and the second exhaustiveness.
 
 % ===============================================================================
-\section*{References}
-
-\bibliography{references,jp}
+\begin{lemma}\label{lemma:mainfsoundtrace}
+A test suite $\TS_T$ generated from a CSP process $P$, as specified in
+Theorem~\ref{th:tracetest}, is passed by every CSP process $Q$ satisfying
+$P\lessdet_T Q$.
+\end{lemma}
+\begin{proof}
+Suppose that $P\lessdet_T Q$, so that $\trc(Q)\subseteq\trc(P)$, and assume
+that $s\in\trc(Q)$ with $\#s < pq$. Since $s$ is also a trace of $P$, we can
+conclude
+$$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$
+because of (\ref{eq:ifpaT}). \pagebreak Now $\trc(Q)\subseteq\trc(P)$ implies
+$[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot
+enter branch (\ref{eq:uta}) and produce a $\efail$ event when running in
+parallel with $Q$ and synchronising over $\Sigma$. Therefore, only four
+options are available for the test execution    $(Q\parallel[\Sigma]
+U_T(j))/s$ to continue.
+
+\medskip
+\noindent {\bf Case~1.} $Q/s$ deadlocks and $\minaccs(G(P)/s) = \{
+\varnothing \}$. In this case, the test $U_T(pq-1,\#s,G(P)/s)$ enters branch
+(\ref{eq:utb}), and its execution stops after $\epass$.
+
+\medskip
+\noindent {\bf Case~2.} $Q/s$ deadlocks, but
+$\minaccs(G(P)/s)\neq\{\varnothing\}$. In this case, the whole test execution
+deadlocks, and this means that neither a $\epass$ nor a $\efail$ event is
+produced, so the test execution is passed.
+
+\medskip
+\noindent {\bf Case~3.} $Q/s$ selects an event $e\in[Q/s]^0$ and $\#s <
+pq-1$. In this case, the test $U_T(pq-1)$ in state $U_T(pq-1,\#s,G(P)/s)$ can
+also engage in $e$ by entering branch (\ref{eq:utc}), and its execution
+continues without producing a $\epass$ or a $\efail$ event.
+
+\medskip
+\noindent {\bf Case~4.} $\#s = pq-1$ holds. In this case,
+$U_T(pq-1,\#s,G(P)/s)$ can enter branch (\ref{eq:utd}), and the test
+execution stops after $\epass$.
+
+\bigskip\noindent%
+This case analysis shows that every   execution of $(Q\parallel[\Sigma]
+U_T(j))$ either stops after $\epass$ or produces neither $\epass$ nor
+$\efail$. This proves that $Q$ passes test $U_T(pq-1)$ according to the pass
+criterion (\ref{eq:passT}). \xbox
+\end{proof}
+% ===============================================================================
+
+\begin{lemma}\label{lemma:mainfexhaustivetrace}
+A test suite $\TS_T$ specified as in Theorem~\ref{th:tracetest} is
+exhaustive for the fault model specified there.
+\end{lemma}
+\begin{proof}
+As in the proof for failures testing, we construct the product graph
+$G=G(P)\times G(Q)$ and recall that every trace $s\in\trc(P)\cap\trc(Q)$ is
+associated with a path through $G$ labelled with the same events as $s$, such
+that $G/s = (G(P)/s,G(Q)/s)$. Furthermore, we recall from
+Lemma~\ref{lemma:extendV} that the graph state $(G(P)/s,G(Q)/s)$ can always
+be reached by a trace $u$ of length less or equal $pq-1$, where  the order of
+$G(P)$ is $p$ and that of $G(Q)$ is $q$.
+
+Suppose that $P\not\lessdet_T Q$. Since the empty trace is a trace of every
+process, there exists a trace $s\in \trc(Q)\cap\trc(P)$ and an event $e\in
+[Q/s]^0$ such that $e\not\in [P/s]^0$. Let $u\in \trc(Q)\cap\trc(P)$ be a
+trace with $\#u < pq$ and $G/u = (G(P)/s,G(Q)/s)$. Then
+$$
+U_T(pq-1)/u = U_T(pq-1,\#u,G(P)/s).
+$$
+By assumption, $e\in (\Sigma -[P/s]^0) = (\Sigma - [G(P)/s]^0)$. Since
+$G(Q)/u = G(Q)/s$, $Q/u$ can engage into $e$. Then $U_T(pq-1,\#u,G(P)/s)$ can
+enter branch (\ref{eq:uta}), and the test execution stops after having
+produced $\efail$. This proves that $Q$  fails test $U_T(pq-1)$. \xbox
+\end{proof}
+%
+Having established completeness of our test suites, we consider
+complexity of a testing technique that uses them.
+
+\section{Complexity Considerations} \label{sec:complexity}
+% ================================================================================
+
+%In this section, we calculate estimates for the maximal number of test
+%executions to be performed when testing for failures refinement
+%and---as a corollary---trace refinement.
+%Theorem~\ref{th:failurestest} specifies that all tests $U_F(j),\ 0\le j < pq$
+%need to be executed, where $p$ denotes the number of nodes in the transition
+%graph of the reference process $P$, and $q\ge p$ is an estimate for the
+%maximal number of nodes in the SUT's transition graph. Therefore, we will first
+%calculate a bound for the number of test executions to be performed for test
+%$U_F(j)$ and then summarise  these bounds over all $j$ from $0$ to $pq-1$.
+%
+%For the worst-case estimate, we first introduce a CSP reference process $\pmax$
+%which
+%turns out to be--given a fixed alphabet $\Sigma$---the test
+%model leading to the maximal number
+%of test executions for every $U_F(j)$ when considering large alphabets. In the case
+%of small alphabets and large values of $pq-1$, it will turn out below that a variant
+%of $\pmax$ will lead to the maximal number of executions. The conditions for this
+%can also be clearly specified.
+
+%we assume that $P$ never allows for early
+%deadlock (so $\minhits(P/s)$ is never empty) and that the SUT $Q$ is a
+%correct failures refinement. Therefore, all test executions
+%$(Q\parallel[\Sigma] U_F(j))$ stop after having run through a trace of $Q$ of
+%length $j+1$, because their is no early termination due to entering branches
+%(\ref{eq:ufa}), (\ref{eq:ufb}), or due to an illegal deadlock of $Q$. As can
+%be seen from the specification of the test cases $U_F(j)$ (see
+%Section~\ref{sec:finitecompletefails}), the number of executions ending in a
+%$\epass$ event corresponds to the number $\ell$ of traces $s$ of $P$ with
+%length equal to $j$, multiplied by the number $h$ of minimal hitting sets in
+%$\minhits(P/s)$. For the tests $U_T(j)$ verifying trace refinement (see
+%Section~\ref{sec:finitecomplete}), the number of executions equals $\ell$,
+%since there is no equivalent in $U_T(j)$ to checking different hitting sets
+%in the last step of a test execution. \fixme{But doesn't this add to the
+%number of executions?} \fxnote{jp: should be clear now from the revised test}
+
+Since we have finite complete CSP test suites, it is useful for the first
+time to calculate how many test executions are needed when using them.
+Previous work did not consider sufficient conditions for finiteness, so
+complexity was not a concern. We answer the following questions. (1)~What is
+the worst-case bound on the number of test executions to be performed to
+verify an SUT with respect to failures refinement, when we use our test
+suite? (2)~What is the worst-case bound for traces refinement? (3)~Is it
+possible to reduce the maximal length of traces when testing for failures or
+traces refinement? We consider the first question~(1) in
+Section~\ref{section:complexity:failures}), where we also discuss whether it
+it is possible to reduce the number of test executions with a different test
+suite.  With the answer to question~(1), question~(2) is a fairly simple
+consequence we discuss briefly also in
+Section~\ref{section:complexity:failures}. Question~(3) is the subject of
+Section~\ref{section:complexity:length}.
+
+% -------------------------------------------------------------------------
+\subsection{Estimates for the Maximal Number of Failures Test Executions}
+\label{section:complexity:failures}
+
+An arbitrary CSP process $P$ might have $\minhits(P/s) = \varnothing$ for
+some traces $s$, so that a test case $U_F(j)$ for a $j$ greater than the size
+of $s$ can enter branch~(\ref{eq:ufb}). In this case, further executions are
+needed to consider traces that have $s$ as a prefix.  To provide a bound on
+the number of test executions needed, we first define a process $\pmax$~(see
+(\ref{eq:pmax})), which, when used as a reference process, requires the
+maximal number of test executions among all reference processes $P$
+fulfilling $\minhits(P/s) \neq \varnothing$ for all traces $s$. For $\pmax$,
+we can establish the actual number of test executions required~(see
+(\ref{eq:pmaxcomplexity})). We then show that the order of magnitude of the
+worst-case bound for the number of test executions is the same also for
+reference processes $P$ that may have $\minhits(P/s) = \varnothing$ for some
+$s$.
+
+% -----------------------------------------------------------------------------
+\paragraph{A Reference Process} Given an alphabet $\Sigma$ of size $\card{\Sigma} =
+n\ge 2$, define a collection of subsets of $\Sigma$ by
+\begin{equation}\label{eq:defC}
+  {\cal C} = \{ A\subseteq\Sigma~|~\card{A} = n - \lfloor\frac{n}{2}\rfloor + 1 \}.
+\end{equation}
+%
+With this choice of ${\cal C}$, define
+\begin{equation}\label{eq:pmax}
+  \pmax = \Intchoice_{A\in{\cal C}} e:A\then \pmax
+\end{equation}
+%
+The relevant properties of $\pmax$ are summarised in the following lemma.
+%
+\begin{lemma}\label{lemma:pmax}
+  Given alphabet $\Sigma$ with cardinality $\card{\Sigma} = n\ge 2$,
+  process $\pmax$ fulfils
+  %
+  \begin{eqnarray}
+  {}[\pmax/s]^0 & = & \Sigma \quad\text{for all $s\in \Sigma^*$}
+  \label{eq:pmaxa}
+  \\
+  \trc(\pmax) & = & \Sigma^*
+  \label{eq:pmaxb}
+  \\
+  \minaccs(P/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
+  \label{eq:pmaxc}
+  \\
+  \minhits(P/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
+  \label{eq:pmaxe}
+  \\
+  \card{\minhits(P/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor}
+  \quad \text{for all $s\in \Sigma^*$}
+  \label{eq:pmaxd}
+  \\
+  \minhits({\cal C})  & = & \{ H\subseteq \Sigma~|~\card{H} = \lfloor\frac{n}{2}\rfloor\}
+  \label{eq:pmaxf}
+  \end{eqnarray}
+\end{lemma}
+\begin{proof}
+Since $\bigcup_{A\in{\cal C}} A = \Sigma $ by construction of ${\cal C}$,
+$[\pmax/s]^0 = \Sigma$ as stated by (\ref{eq:pmaxa}). Since $\pmax/e = \pmax$
+for all $e\in\Sigma$, this proves statement (\ref{eq:pmaxb}). The internal
+choice construct used in the specification of $\pmax$ implies
+$\minaccs(\pmax) = {\cal C}$. Again, $\pmax/e = \pmax$ for all $e\in\Sigma$
+implies $\minaccs(\pmax/s) = {\cal C}$ for all traces of $\pmax$, so this
+shows (\ref{eq:pmaxc}). Statement (\ref{eq:pmaxe}) is a direct consequence of
+(\ref{eq:pmaxc}). Let $H$ be any minimal hitting set of ${\cal C}$. Then $H$
+contains at least $\lfloor{\frac{n}{2}}\rfloor$ elements, because otherwise
+$\card{\Sigma\setminus H} > n-\lfloor{\frac{n}{2}}\rfloor$, and any subset
+$A\subseteq \Sigma\setminus H$ with cardinality
+$n-\lfloor{\frac{n}{2}}\rfloor+1$  would be contained in ${\cal C}$, but
+satisfy $A\cap H=\varnothing$. Since
+$\lfloor{\frac{n}{2}}\rfloor+n-\lfloor{\frac{n}{2}}\rfloor+1=n+1$, we
+conclude that any $\lfloor{\frac{n}{2}}\rfloor$-element subset of $\Sigma$
+intersects  every element of ${\cal C}$.  Therefore, every minimal hitting
+set of ${\cal C}$ has exactly $\lfloor{\frac{n}{2}}\rfloor$ elements; this
+shows (\ref{eq:pmaxf}) and $\card{\minhits({\cal C})} =
+\binom{n}{\lfloor{\frac{n}{2}}\rfloor}$. The latter shows  (\ref{eq:pmaxd})
+and completes the proof. \xbox
+\end{proof}
+
+% -----------------------------------------------------------------------------
+\paragraph{Test Cases of $\pmax$} The test cases $U_F(j)$ generated from $\pmax$ can
+never enter branch (\ref{eq:ufa}), because $\pmax/s$ has initials $\Sigma$
+for all traces $s\in\trc(\pmax)$ according to (\ref{eq:pmaxa}). Moreover,
+they can never enter branch (\ref{eq:ufb}), because $\minhits(\pmax/s)$ is
+never empty according to (\ref{eq:pmaxd}). Finally, the minimal hitting sets
+used to probe the SUT at the end of a non-blocking test execution are always
+the hitting sets of ${\cal C}$ according to (\ref{eq:pmaxe}). This results in
+the following test case structure.
+\begin{eqnarray*}
+\label{eq:UFPpmax}
+U_F(j) & = & U_F(j,0,\ii n)
+\\
+U_F(j,k,n) & = &   (k < j) \& \big(e:\Sigma   \then U_F(j,k+1,t(n,e) \big)
+\\ & & \extchoice
+\\ & & (k = j) \& \big( \Intchoice_{H\in\minhits({\cal C})} (e:H   \then \epass \then\Stop   )  \big)
+\end{eqnarray*}
+This means that the branches of $U_F(j)$ that can lead to an early
+termination are not feasible. All tests deadlock, or run to the end of a
+trace of size $j$ and then present a choice of events from a minimal hitting
+set.
+
+Moreover, Theorem~\ref{th:sperner} establishes that given an alphabet with
+$n$ elements, there is no Sperner family consisting of more than
+$\binom{n}{\lfloor\frac{n}{2}\rfloor}$. In addition, we recall the minimal
+hitting sets calculated from a given set of minimal acceptances are a Sperner
+family~(see discussion in Section~\ref{sec:sperner}). So, (\ref{eq:pmaxd})
+establishes that there is no possibility of providing more choices fom
+minimal hitting sets with a test derived from a process different from
+$\pmax$.
+
+In summary, the tests derived from $\pmax$ require the most test executions
+when compared to tests derived from any other CSP process $P$ whose
+collections of minimal hitting sets $\minhits(P/s)$ are never empty for any
+trace $s$.
+
+% -----------------------------------------------------------------------------
+\paragraph{Maximal Number of Test Executions for $\pmax$}
+When considering the number of test executions to be performed using $U_F(j)$
+derived from $\pmax$ against some SUT $Q$ for all $j = 0,\dots,pq-1$, and
+considering that we need to cover all the possible behaviours of $Q$, the
+maximal number of test executions is only reached if (a)~$Q$ is a correct
+refinement of $\pmax$ (or more generally, of the reference process) and
+(b)~its traces are $\Sigma^*$.
+%or if it fails in the very last execution of the very last $U_F(j)$
+%executed against $Q$. If an erroneous behaviour of $Q$ is revealed before
+%this very last execution, the test experiment is stopped (and $Q$ can be
+%fixed before executing the suite again). Therefore, we consider only correct
+%SUTs $Q$ when determining the maximal number of executions that can be
+%executed.
+%
+%For the  $U_F(j)$ generated from $\pmax$, the number of test executions
+% to be performed is maximal if $\pmax\lessdet_F Q$ and
+%$\trc(Q) = \Sigma^*$.
+In such a situation, no test execution blocks early, because $Q/s$ can always
+engage into some $e\in\Sigma$  while $\#s<j$, and never blocks in the last
+step when $\#s = j$ and a hitting set $H\in\minhits({\cal C})$ is offered by
+the test case. The resulting number of executions in this case is
+%
+\begin{equation}
+\label{eq:maxexec}
+n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor},
+\end{equation}
+%
+because all traces up to length $j$ can be executed with $U_F(j)$
+entering branch (\ref{eq:ufc}), and each of these traces is followed by one
+event from each of the hitting sets of ${\cal C}$ since $Q$ is correct.  %(b) leads to the maximal number of
+%executions possible, because it is responsible for the factor *n^{(pq-1)} in
+%the upper bound (47), (48) [n = |Sigma|]. Considering a reference process
+%with fewer traces or an implementation Q with fewer traces would
+%significantly reduce the value of this factor, while just adding one
+%additional execution possibility for branch (20), after which the test suite
+%is stopped due to failure.
+
+The number of executions in (\ref{eq:maxexec}) is indeed maximal for all
+reference processes $P$ fulfilling $\minhits(P/s)\neq\varnothing$ for all
+traces $s$. All these processes can never enter branch (\ref{eq:ufb}), and,
+if an execution with an SUT entered branch (\ref{eq:ufa}), this would only
+lead to early termination of the whole test suite, because a failure has been
+detected. As a consequence, $\card{\Sigma}^{j}$ is the maximal number of
+traces to be executed up to length $j$, and from Theorem~\ref{th:sperner} we
+know that the number $\binom{n}{\lfloor\frac{n}{2}\rfloor}$ of hitting sets
+to be tested at the end of each trace of length $j$ is already maximal.
+
+Summing up  formula (\ref{eq:maxexec}) over all test cases
+$U_F(0),\dots,U_F(pq-1)$ to be executed and applying the formula for the sum
+of a geometric progression,  this results int
+%
+\begin{equation}\label{eq:pmaxcomplexity}
+\sum_{j=0}^{pq-1} n^{j}\cdot \binom{n}{\lfloor\frac{n}{2}\rfloor}  =
+\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}
+\end{equation}
+%
+as the maximal number of test executions to be performed when testing an
+error-free SUT $Q$ with $\trc(Q) = \Sigma^*$ against the reference process
+$\pmax$. If we are interested only in the order of magnitude, we have
+%
+\begin{equation}
+\label{eq:maxO}
+O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
+\end{equation}
+
+% -----------------------------------------------------------------------------
+\paragraph{Considering Empty Collections of Minimal Hitting Sets}
+The argument so far has shown that the tests derived from $\pmax$ require the
+most test executions when considering processes whose collections of minimal
+hitting sets are never empty. %Lemma~\ref{lemma:failureshittingsets} implies
+%that an empty collection $\minhits(P/s)$ is equivalent to $(s,A)\in\fails(P)$
+%for all $A\subseteq\Sigma$. This is equivalent to $\maxrefs(P/s) = \{ \Sigma
+%\}$ or $\minaccs(P/s) = \{ \varnothing\}$.
+%
+It remains to consider whether reference processes $Z$ possessing failures
+$(s,\Sigma)$ may require more test executions for their associated tests
+$U_F(j)$ than the bound given for $\pmax$ in (\ref{eq:pmaxcomplexity}),
+because process states $Z/s$ with $\minhits(Z/s)=\varnothing$ allow for test
+executions entering branch (\ref{eq:ufb}). To this end, consider a test case
+$U_F(j)$ constructed from such a process $Z$. Every trace $s\in\trc(Z)$ with
+$\#s<j$ ending in a process state $Z/s$ with $\minhits(Z/s)=\varnothing$
+allows for
+%
+\begin{itemize}
+\item one execution of branch (\ref{eq:ufb}), where the test execution
+$(Z\parallel[\Sigma]U_F(j))/s$ stops after $\epass$, and
+\item $\card{[Z/s]^0}$   continuations of the test execution with events $e\in [Z/s]^0$.
+\end{itemize}
+%
+For every trace $s\in\trc(Z)$ with $\#s=j$,
+%
+\begin{itemize}
+\item one execution of branch (\ref{eq:ufb}) follows if
+$\minhits(Z/s)=\varnothing$, and otherwise
+\item $\card{\minhits(Z/s)}$ executions checking acceptance of minimal hitting sets.
+\end{itemize}
+%
+For a rough estimate of the worst-case upper bound suppose that
+%
+\begin{enumerate}
+\item $[Z/s]^0 = \Sigma$ for all traces $s$ of $Z$,
+\item all traces $s$ with $\#s<j$ end in a state with empty minimal hitting sets, and
+\item all traces $s$ with $\#s=j$ end in a state with a maximal number $\binom{n}{\lfloor n/2\rfloor}$
+ of hitting sets.
+\end{enumerate}
+%
+We note that this scenario is not feasible for all $j\in\{0,\dots,pq-1\}$,
+because the traces of $U_F(p-1)$ already cover all states of $Z$'s transition
+graph according to Lemma~\ref{lemma:extendV}, and if all states of $Z$ have
+empty hitting sets, there are no acceptance checks to be performed in the
+last step of the test execution. Therefore, \pagebreak the upper bound
+calculated next cannot be reached by a CSP process. With the three
+assumptions above, nevertheless, we can calculate that
+%
+\begin{itemize}
+\item  $U_F(0)$ has  $\binom{n}{\lfloor n/2\rfloor}$ executions,
+\item $U_F(j)$, for $j > 0$ has $\sum_{i=0}^{j-1} n^i = \frac{n^j -
+    1}{n-1}$ executions of branch (\ref{eq:ufb}), ($n = \card{\Sigma}$),
+    and
+\item $U_F(j)$, for $j > 0$ has $\binom{n}{\lfloor n/2\rfloor}\cdot n^j$
+    executions where the acceptance of hitting sets is checked after having
+    run through a trace of length $j$.
+\end{itemize}
+%
+Summing up over all $U_F(j)$ for $j=0,\dots,pq-1$, an upper bound $B$ may be calculated as follows.
+%
+\begin{eqnarray*}
+B & = & \binom{n}{\lfloor n/2\rfloor} + \sum_{j=1}^{pq-1} \frac{n^j - 1}{n-1} +
+\sum_{j=1}^{pq-1}\binom{n}{\lfloor n/2\rfloor}\cdot n^j
+   \\
+& = &    \sum_{j=1}^{pq-1} \frac{n^j - 1}{n-1} +
+\sum_{j=0}^{pq-1}\binom{n}{\lfloor n/2\rfloor}\cdot n^j
+\\
+& = & \frac{n^{pq}-n pq+pq-1}{(n-1)^2} +
+\binom{n}{\lfloor n/2\rfloor} \cdot \frac{n^{pq} - 1}{n-1}
+\\
+& = &
+\\
+& = & \frac{\binom{n}{\left\lfloor \frac{n}{2}\right\rfloor}(n-1) \left(n^{pq}-1\right) +n^{pq}-npq+pq-1}{(n-1)^2}
+\end{eqnarray*}
+%
+Since $B$ cannot be reached anyway, we just calculate its order of magnitude,
+and this results again in $O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}  \cdot
+n^{pq-1}\big)$, as calculated already for $\pmax$ in (\ref{eq:maxO}).
+%
+Summarising these complexity calculations, we have the following theorem.
+%
+\begin{theorem}
+\label{th:maxexecs}
+Given a process alphabet $\Sigma$, consider a fault model ${\cal F} = (P,\lessdet_F,{\cal D})$, such that
+the normalised transition graph of $P$ has $p$ states, and the fault domain
+${\cal D}$ contains all processes $Q$ over alphabet $\Sigma$, such that $G(Q)$ has at most
+$q\ge p$ states. Then the maximal
+number of test executions to be performed   using the complete test
+suite $\TS_F = \{ U_F(j)~|~0 \le j < pq  \}$ created from $P$ as specified in Theorem~\ref{th:failurestest} is of order
+%
+\begin{equation*}
+O\big(\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot n^{pq-1}\big)\quad\text{with $n=\card{\Sigma}$}.
+\end{equation*}
+For processes $P$ satisfying $(s,\Sigma)\not\in\fails(P)$ for all traces $s$, the reachable
+precise  upper bound is given by
+%
+\begin{equation*}
+\binom{n}{\lfloor\frac{n}{2}\rfloor}\cdot\frac{1-n^{pq}}{1-n}\quad\text{with $n=\card{\Sigma}$}.
+\end{equation*}
+\xbox
+\end{theorem}
+%
+As established by Lemma~\ref{lemma:hseta}, the number of hitting sets used to
+probe the SUT cannot be reduced:~if the reference process is $\pmax$, we need
+to consider all of them, otherwise illegal blocking may remain undetected. In
+addition, if $\pmax$ defines the behaviour of the SUT, using smaller sets
+that are no longer hitting sets lead to a rejection of correct
+implementations. Our explicit definition of $\pmax$ to establish this
+worst-case is useful to illustrate this point.
+
+In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
+non-empty subset of $\Sigma$ whose events cannot be completely refused in a
+given process state of the reference model; this leads to a worst-case
+estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
+offered to the SUT in the last step of the test execution. This number is
+significantly larger than the worst-case estimate $\binom{n}{\lfloor
+n/2\rfloor}$ above. In Fig.~\ref{fig:minhita}, the reduction is visualised by
+plots of the two functions. In~\cite{DBLP:conf/icfem/CavalcantiG07}, the
+authors also use minimal hitting sets\footnote{However, they are denoted by
+{\it minimal acceptances} in~\cite{DBLP:conf/icfem/CavalcantiG07}.}, but do
+not give an estimate for the number of test executions.
+
+% .....................................................................................
+ \begin{figure}
+ %%\hspace*{-40mm}
+ \begin{center}
+\includegraphics[scale=.7%width=.8\textwidth
+                        ]{curvecomparison.pdf}
+\end{center}
+%\vspace*{-10mm}
+\caption{Function plot $2^{\card{\Sigma}}$ versus $\binom{n}{\lfloor \frac{n}{2}\rfloor}$.}
+ \label{fig:minhita}
+ \end{figure}
+% .......................................................................................
+
+% -------------------------------------------------------------------------
+\paragraph{Estimate for the Maximal Number of Trace Test Executions}
+%\label{section:complexity:traces}
+According to Theorem~\ref{th:tracetest}, a complete test suite checking
+traces refinement just contains the adaptive test case $U_T(pq-1)$. As
+derived for $U_F(j)$ above, the maximal number of executions to be performed
+by $(Q\parallel[\Sigma] U_T(pq-1))$ is of order
+$O\big(\card{\Sigma}^{pq-1}\big)$.
+
+% -------------------------------------------------------------------------
+\subsection{Upper Bound $pq$ for the Maximal Length of Test Traces}
+\label{section:complexity:length}
+
+According to Theorem~\ref{th:failurestest}, the tests $U_F(j)$ need to be
+executed for $j = 0,\dots,pq-1$ to guarantee completeness. So, the SUT is
+verified with test traces up to, and including, length $pq$. With branch
+(\ref{eq:ufa}), $U_F(j)$ accepts all traces $s.e$ with $s\in\trc(P), \#s
+= j, e\not\in\trc(P/s)$, so erroneous traces up to length $j+1$ are detected.
+
+It is interesting to investigate whether this maximal length is necessary, or
+one could elaborate alternative complete test strategies where the
+SUT is tested with shorter traces only. Indeed, an example
+in~\cite[Exercise~5]{PeleskaHuangLectureNotesMBT} shows that, when testing
+for equivalence of deterministic FSMs, it is sufficient to test with
+traces of significantly shorter length.
+
+The following example, however, shows that the maximal length $pq$ is really
+required when testing for refinement.
+%\begin{example}\label{ex:pq}
+%Consider the CSP reference process $P$ and an erroneous implementation $Q$
+%specified as follows.
+%
+%\begin{center}
+%\begin{minipage}{.4\textwidth}
+%\begin{eqnarray*}
+%P & = & a \then P_1 \intchoice b \then P_1 \intchoice c \then P_1
+%\\
+%P_1 & = & a \then P \extchoice b\then P
+%\end{eqnarray*}
+%\end{minipage}
+%\hfill
+%\begin{minipage}{.4\textwidth}
+%\begin{eqnarray*}
+%Q & = & a\then Q_1 \extchoice b\then Q_1
+%\\
+%Q_1 & = & a\then Q_2 \extchoice b\then Q_2
+%\\
+%Q_2 & = & a\then Q \intchoice b\then Q
+%\end{eqnarray*}
+%\end{minipage}
+%\end{center}
+%
+%
+%\medskip
+%Obviously, $P$'s normalised transition graph has 2 nodes, while $Q$'s graph
+%has 3. It is easy to see (and can be checked with FDR4) that $P\lessdet_T Q$,
+%but $\neg(P\lessdet_F Q)$. Furthermore, it can also be shown using FDR4 that
+%the ``test passed condition''
+%\[
+%(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j))\hide \Sigma
+%\]
+%holds for $U_F(0),\dots,U_F(4)$, but fails for $U_F(5)$. So, the
+%non-conformance of $Q$ cannot be detected by any test trace of length less or
+%equal to 5, but is revealed (as expected from Theorem~\ref{th:failurestest})
+%by a trace of length 6, because the last event offered by the test $U_F(5)$
+%is refused by $Q$. \xbox
+%\end{example}
+
+\begin{example}\label{ex:pq}
+Consider the CSP reference process $P$ and an erroneous implementation $Q$
+specified as follows.
+%
+\begin{eqnarray*}
+P & = &  P(0)
+\\
+P(k) & = & (k < p-1) \& \big( (a \then P(k)) \intchoice ( b \then P(k+1))\big)
+\\ & & \extchoice
+\\ & & (k = p-1) \& (a \then P(k))
+\\
+Q & = & Q(0)
+\\
+Q(k) & = & (k < q-1) \& \big( a \then Q(k+1)    \big)
+\\ & & \extchoice
+\\ & & ( k = q-1)\& \big( a\then Q(0) \extchoice b\then Q(0)  \big)
+\end{eqnarray*}
+%
+The normalised transition graphs of $P$ and $Q$ are depicted in
+Fig.~\ref{fig:examplepq} for the case $p=3,\ q=4$. Using FDR4, it can be
+shown for concrete values of $p$ and $q$ that the  ``test passed conditions''
+\[
+(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j))\hide \Sigma
+\quad
+\mathrm{and}
+\quad
+(\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(j))\hide \Sigma
+\]
+hold for $j = 0,\dots,pq-2$. So, none of the test cases $U_F(j)$ and $U_T(j)$
+are capable of detecting failures and traces-refinement violations, if they
+only check traces up to length $pq-1$. (We recall that this corresponds to
+$j\le pq-2$).
+
+$Q$, however, neither conforms to $P$ in the failures refinement relation,
+nor in the traces-refinement relation. This can only be seen when executing
+the test $U_F(pq-1)$ and $U_T(pq-1)$, respectively. These tests fail, so this
+shows that $P\not\lessdet_F Q$ and $P\not\lessdet_T Q$ according to
+Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this
+shows that the maximal trace length $pq$ to be investigated in the tests
+cannot be further reduced without losing the completeness property of the
+test suites. \xbox
+\end{example}
+%
+Generalising Example~\ref{ex:pq}, it can be shown that for any $p,q \geq 2$,
+%\in\mathbb{N}$,
+there exist reference processes $P$ with $p$ states and implementation
+processes $Q$ with $q$ states, such that a violation of the traces-refinement
+property can only be detected with a trace of length $pq$. This is proven in
+the following
+theorem. %In the proof, we use the processes $P$ and $Q$ introduced in Example~\ref{ex:pq}.
+
+\begin{figure}[tbp]
+\begin{center}
+\begin{minipage}{.4\textwidth}
+\includegraphics[trim=1.4cm 0cm 0cm 0cm,clip,width=1.1\textwidth]{theorem5p.pdf}
+\end{minipage}
+\hfill
+\begin{minipage}{.55\textwidth}
+\includegraphics[trim=1.4cm 0cm 0cm 0cm,clip,width=1.1\textwidth]{theorem5q.pdf}
+\end{minipage}
+\caption{Transition graphs of $P$ (left) and $Q$ (right) from Example~\ref{ex:pq}
+ for $p=3$ and $q=4$.}
+\label{fig:examplepq}
+\end{center}
+\end{figure}
+
+% ----------------------------------------------------------------------------
+\begin{theorem}\label{th:maxtracelen}
+Let $2\le p,q \in\mathbb{N}$. Then there exists a reference process $P$ and an
+implementation process $Q$ with the following properties.
+\begin{enumerate}
+\item $G(P)$ has $p$ states.
+\item $G(Q)$ has $q$ states.
+\item $P\not\lessdet_T Q$, and therefore, also $P\not\lessdet_F Q$.
+\item $\forall s\in\trc(Q): \#s < pq\implies s\in\trc(P)$.
+\item $Q\ conf\ P$.
+\end{enumerate}
+As a consequence, the upper bound $pq$ for the length of traces to be tested
+when checking for failures refinement or traces refinement cannot be reduced
+without losing the test suite's completeness property.
+\end{theorem}
+% ----------------------------------------------------------------------------
+\begin{proof}
+Given $2\le p,q \in\mathbb{N}$, define reference process $P$ and
+implementation process $Q$ as in Example~\ref{ex:pq}. It is trivial to see
+that $G(P)$ has $p$ nodes and $G(Q)$ has $q$ nodes, so statements 1 and 2 of
+the theorem hold.
+
+Using regular expression notation, the traces of $P$ can be specified as
+$%\[
+\trc(P) = \prefs\big(  (a^*b)^{p-1}a^* \big),
+$ %\]
+where $\prefs(M)$ denotes the set of all prefixes of traces in
+$M\subseteq\Sigma^*$, including the traces of $M$ themselves. The traces of
+$Q$ can be specified by
+$%\[
+\trc(Q) = \prefs\big( (a^{q-1}(a|b))^*  \big).
+$ %\]
+It is easy to see that $\trc(Q)\not\subseteq\trc(P)$; for example, the trace
+$(a^{q-1}b)^p$ is in $\trc(Q)\setminus\trc(P)$, because traces of $P$ contain
+at most $p-1$ $b$-events. This proves statement~3.
+
+Let $s \in\trc(Q)$ be any trace of length $\#s = pq-1$, then $s =
+(a^{q-1}(a|b))^{p-1}a^{q-1} \in \prefs\big( (a^{q-1}(a|b))^*  \big)$. So, $s$
+is also an element of $\trc(P)$, because $(a^{q-1}(a|b))^{p-1}a^{q-1}$ is
+also contained in $\prefs\big( (a^*b)^{p-1}a^* \big)$, since $\prefs\big(
+(a^*b)^{p-1}a^* \big)$ contains all finite sequences of $a$, where at most
+$p-1$ events $b$ have been inserted. This proves statement 4.
+
+To prove statement~5, we observe that the specifications of $P$ and $Q$ lead
+to the following sets of minimal acceptances.  In these definition, the
+expression $(s\cnt b)$ denotes the number of $b$ events occurring in trace
+$s$.
+\[
+\minaccs(P/s) = \left\{
+\begin{array}{ll}
+\{ \{a\}, \{b\} \} & \text{for all $s\in\trc(P)$ with $(s\cnt b) <p-1$.}
+\\
+\{ \{a\} \} & \text{for all $s\in\trc(P)$ with $(s\cnt b) = p-1$.}
+\end{array}
+\right.
+\also
+\minaccs(Q/s) = \left\{
+\begin{array}{ll}
+\{ \{a\}  \} & \text{for all $s\in\trc(Q)$ with $\#s \neq 0\mod(q-1)$.}
+\\
+\{ \{a,b\} \} & \text{for all $s\in\trc(P)$ with $\#s = 0\mod(q-1)$.}
+\end{array}
+\right.
+\]
+So, the minimal acceptance set $A_P = \{a\}$ contained in every
+$\minaccs(P/s)$ fulfils $A_P \subseteq A_Q$ for any $A_Q\in\minaccs(Q/s)$,
+when $s\in\trc(P)\cap \trc(Q)$. Now Lemma~\ref{lemma:tgtrcref}, in particular
+(\ref{eq:failrefb}), can be applied to conclude that $Q\ conf\ P$. \xbox
+\end{proof}
+%
+It is discussed next in Section~\ref{sec:conc} how the number of test traces
+to be executed by complete test suites for failures or traces refinement can
+still be reduced {\it without} reducing the maximal length.
+
+% =================================================================================
+
+% ==============================================================================
+\section{Discussion and Conclusions}
+\label{sec:conc}
+% ==============================================================================
+
+% ==============================================================================
+\paragraph{Further Reductions of the Test Effort} As shown in
+Theorem~\ref{th:maxtracelen}, the maximal length $pq$ of traces to be tested
+for either  failures refinement or traces refinement cannot be further
+reduced. It is noteworthy, however, that when testing FSMs for equivalence,
+considerably shorter traces can be used. From the classical results published
+in~\cite{chow:wmethod,vasilevskii1973}, for example, it follows that the
+maximal trace length to be executed is less or equal to $3p -q$, which is
+considerably smaller than $pq$ for $p,q\ge 3$. As a consequence, the
+investigation of complete test suites establishing failures or trace
+equivalence is of considerable interest and will be discussed in a future
+paper.
+
+It is also known from FSM testing that it is not necessary to check {\it all}
+traces up to length $pq$ when testing for reduction of FSMs (which
+corresponds to trace refinement). Notable complete strategies supporting this
+fact have been presented, for example,
+in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}.
+From~\cite{Huang2017} it is known that complete FSM testing theories can be
+translated to other formalisms, such as Extended Finite State Machines,
+Kripke Structures, or CSP, resulting in likewise complete test strategies for
+the latter. We intend to study translations of several promising FSM
+strategies to CSP. These will effectively reduce the upper bound for the
+number of test executions to be performed, which  has been shown to be of the
+order $O\big(\card{\Sigma}^{pq-1}\big)$ for our traces-refinement tests in
+Section~\ref{sec:complexity}.  The bound $\binom{n}{\lfloor n/2\rfloor}$ for
+the number of sets to be used in probing the SUT for illegal deadlocks,
+however, cannot be further reduced, as established in
+Lemma~\ref{lemma:hseta}.
+
+% ==============================================================================
+\paragraph{Adaptive Test Cases} The tests suggested
+in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were
+\emph{preset} in the sense that the trace to be executed was pre-defined for
+each test. As a consequence, the authors
+of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive} as a
+third test result, applicable to the situations where the intended trace of
+the execution was blocked, due to legal, but nondeterministic behaviour of
+the SUT. We consider this as a disadvantage, since, when aiming at executing
+a specific trace $s$ before being able to check the test objective---for
+example, the absence of deadlocks when offering a hitting set $H$ of the
+minimal acceptances---it may take several tries until the full trace $s$ is
+accepted by the SUT. Considering the complete testing assumption described in
+Section~\ref{sec:finitecompletefails}, it may even take $c^{\#s}$ tries to
+reach the end of the trace $s$, if the SUT can legally block every event of
+$s$ due to nondeterminism, so that $c$ trials are required for each event is
+accepted.
+
+Those authors, later, in the context of a richer algebra based on CSP, have
+considered a framework similar to adaptive testing~\cite{CG15}.  They have,
+however, considered only traces refinement, and have not studied complexity.
+
+%\fixme{From the proofs of the main theorems, and what you say below, I
+%thought you did have verdicts coming from deadlock, and you simply chose not
+%to mark it with an inc event.}
+In  contrast to that, our test cases specified in
+Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
+adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
+U_F(j))$ for failures refinement may only stop early with $\epass$ after
+traces $s$ satisfying $\minhits(P/s) = \varnothing$, and may deadlock (a)
+after a trace $s$ where the SUT illegally deadlocks (so $\minhits(P/s) \neq
+\varnothing$ for the reference process, but $\minhits(Q/s) = \varnothing$ for
+the implementation $Q$), or (b) in the final step when---just as in the
+corresponding test cases specified in~\cite{DBLP:conf/icfem/CavalcantiG07}---
+hitting sets $H$ are offered to the SUT and it refuses their acceptance. In
+both situations (a) and (b) the test fails. As a consequence, far less test
+repetitions are required according to the complete testing assumption than
+for the successful execution of all test cases specified
+in~\cite{DBLP:conf/icfem/CavalcantiG07}.
+
+Another distinction of our failures test cases $U_F(j)$ to the tests
+specified in~\cite{DBLP:conf/icfem/CavalcantiG07} consists in the fact that
+the former test both traces refinement and the $conf$ conformance relation in
+one go, whereas the latter use separate test suites to establish these two
+correctness conditions. Again, we consider the structure of the  test cases
+$U_F(j)$ as advantageous, since, when checking acceptance of a hitting set
+$H$ after a trace $s$ for $c$ times according to the complete testing
+assumption, any acceptance of an illegal event $e\in\Sigma - [P/s]^0$ should
+also be revealed within these $c$ tries.
+
+%
+%never block before the final step specified
+%by branch (\ref{eq:ufd}), and so we do not need inconclusive test results. It
+%should be noted, however, that it is necessary for our test verdicts to
+%recognise also deadlocks in the final test step and interpret them as FAIL,
+%as described in Section~\ref{sec:finitecompletefails}. In practice, this is
+%realised by adding a timeout event to the testing environment which indicates
+%deadlock situations. For real-time systems, this is an accepted technique,
+%because the SUT has to respond within a pre-defined latency interval,
+%otherwise its behaviour is considered to be blocked and regarded as a
+%failure. The tests executions $(Q\parallel[\Sigma] U_T(j))$ \fixme{Missing
+%hiding.} \fxnote{jp: Here, the hiding operator is not necessary: the possible
+%test executions are really the traces of $Q$ parallel $U_T(j)$. The hiding is
+%only needed to specify the verdict.}for trace refinement never block at all
+%before stopping after the verdict $\epass$ or $\efail$, unless the SUT
+%process $Q$ has an unexpected deadlock state. Recall from the explanations
+%given in Section~\ref{sec:finitecomplete} that the blocking situation also
+%leads to passing the test, because $(\epass\then \Stop)\lessdet_T
+%(Q\parallel[\Sigma] U_T(j))\hide\Sigma$ still holds if $(Q\parallel[\Sigma]
+%U_T(j))\hide\Sigma$ only produces the empty trace.
+
+%The adaptive behaviour of both our test case types $U_F(j)$ and $U_T(j)$,
+%however, induces the obligation to check that {\it all} possible executions
+%have been performed before the test can be considered as passed. Typically,
+%it is therefore assumed that a \emph{complete testing
+%assumption}~\cite{hierons_testing_2004} holds, which means that every
+%possible behaviour of the SUT is performed after a finite number of test
+%executions. In practice, this is realised by executing each test several
+%times, recording the traces that have been performed, and using hardware or
+%software coverage analysers to determine whether all possible test execution
+%behaviours of the SUT have been observed. Therefore, adaptive test cases come
+%at the price of having to apply some grey-box testing techniques enabling us
+%to decide whether all SUT behaviours have been observed.
+
+% ==============================================================================
+\paragraph{Fault Domains} As already mentioned, the work
+in~\cite{DBLP:conf/pts/CavalcantiS17} defines a fault domain as the set of
+processes that refine a given CSP process.  In that context, only testing for
+traces refinement is considered, and the complete test suites may not be
+finite. So, the work presented here goes well beyond what is achieved there,
+but is restricted to finite and nonterminating reference processes. In
+addition, \cite{DBLP:conf/pts/CavalcantiS17} presents an algorithm for test
+generation that can be adapted to consider additional selection and
+termination criteria, like, for example, the length of the traces used to
+construct tests. It would be possible, for instance, to use the bound
+indicated here. Moreover, specifying a fault domain as a CSP process allows
+us to model domain-specific knowledge using CSP. For example, if an
+initialisation component defined by a process $I$ can be regarded as correct
+without further testing, we can use $I; RUN$ as a fault domain, to indicate
+that any SUT of interest implements $I$ correctly, but afterwards has an
+arbitrary behaviour specified by $RUN$.
+
+% ==============================================================================
+\paragraph{Implications for CSP Model Checking} As explained in the previous
+sections, passing a test is characterised by the failures-refinement check
+$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ for
+failures testing. If the SUT $Q$ is not a programmed piece of software or an
+integrated hardware or software system, but just another CSP process
+specification, it is of course possible to verify the pass criterion using
+the FDR4 model checker. For checking the refinement relation $P\lessdet_F Q$,
+the pass criterion has to be verified for $j=0,\dots,pq-1$, where $p$ and $q$
+indicate the number of nodes in $P$'s transition graph and the maximal number
+of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since
+the test cases $U_F(j)$ have such a simple structure, it is an interesting
+question for further research whether checking $(\epass\then\Stop) \lessdet_F
+(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ for $j=0,\dots,pq-1$ can be faster
+than directly checking $P\lessdet_F Q$, as one would do in the usual approach
+with FDR4. This is of particular interest, since the checks could be
+parallelised on several CPUs. Alternatively it is interesting to investigate
+whether the check of\footnote{We are grateful to Bill Roscoe for suggesting
+this option.}
+\[
+(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{j=0}^{pq-1} U_F(j)) \hide \Sigma
+\]
+may perform better than the check of $P\lessdet_F Q$, since the former allows for
+other optimisations in the model checker.
+
+For a large implementation process $Q$, it may be too time consuming to
+generate its normalised transition graph, so that its number $q$ of nodes is
+unknown. In such a case, our testing approach based on fault domains may
+still be used as efficient bug finders:~use the number of nodes of the
+normalised transition graph of the reference process as the initial value for
+$q$ and increment $q$ from there, as long as each increment reveals new
+errors.
+
+% ==============================================================================
+\paragraph{Conclusion}
+In this paper, we have introduced finite complete testing strategies for
+model-based testing against finite, non-terminating CSP reference models. The
+strategies are applicable to the conformance relations failures refinement
+and traces refinement. The underlying fault domains have been defined as the
+sets of CSP processes whose normalised transition graphs do not have more
+than a given number of additional nodes, when compared to the transition
+graph of the reference process. For these domains, finite complete test
+suites are available. We have shown for the strategy to check failures
+refinement that the way of probing the SUT for illegal deadlocks in our test
+cases is optimal, so that it is not possible to guarantee exhaustiveness with
+fewer probes. Moreover, the maximal length of the test traces cannot be
+reduced without losing the test suite's completeness property; this holds
+both for traces and failures refinement.
+
+% ==============================================================================
+
+\paragraph{Acknowledgements}
+The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for
+their advice on using the FDR4 model checker and for very helpful discussions
+concerning the potential implications of this paper in the field of model
+checking. We are also grateful to Li-Da Tong from National Sun Yat-sen
+University, Taiwan, for suggesting the applicability of Sperner's Theorem in
+the context of the work presented here. Moreover, we thank Adenilso Simao for
+several helpful suggestions. The work of Ana Cavalcanti is funded by the
+Royal Academy of Engineering and UK EPSRC Grant EP/R025134/1.
+
+% ===============================================================================
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname href\endcsname\relax
+  \def\href#1#2{#2} \def\path#1{#1}\fi
+
+\bibitem{jp2018ets}
+J.~Peleska, \href{https://doi.org/10.1109/ETS.2018.8400703}{Model-based avionic
+  systems testing for the airbus family}, in: 23rd {IEEE} European Test
+  Symposium, {ETS} 2018, Bremen, Germany, May 28 - June 1, 2018, {IEEE}, 2018,
+  pp. 1--10.
+\newblock \href {http://dx.doi.org/10.1109/ETS.2018.8400703}
+  {\path{doi:10.1109/ETS.2018.8400703}}.
+\newline\urlprefix\url{https://doi.org/10.1109/ETS.2018.8400703}
+
+\bibitem{DBLP:conf/isola/0001BH18}
+J.~Peleska, J.~Brauer, W.~Huang,
+  \href{https://doi.org/10.1007/978-3-030-03427-6\_11}{Model-based testing for
+  avionic systems proven benefits and further challenges}, in: T.~Margaria,
+  B.~Steffen (Eds.), Leveraging Applications of Formal Methods, Verification
+  and Validation. Industrial Practice - 8th International Symposium, ISoLA
+  2018, Limassol, Cyprus, November 5-9, 2018, Proceedings, Part {IV}, Vol.
+  11247 of Lecture Notes in Computer Science, Springer, 2018, pp. 82--103.
+\newblock \href {http://dx.doi.org/10.1007/978-3-030-03427-6\_11}
+  {\path{doi:10.1007/978-3-030-03427-6\_11}}.
+\newline\urlprefix\url{https://doi.org/10.1007/978-3-030-03427-6\_11}
+
+\bibitem{hierons_testing_2004}
+R.~M. Hierons,
+  \href{http://doi.ieeecomputersociety.org/10.1109/TC.2004.85}{Testing from a
+  nondeterministic finite state machine using adaptive state counting}, {IEEE}
+  Trans. Computers 53~(10) (2004) 1330--1342.
+\newblock \href {http://dx.doi.org/10.1109/TC.2004.85}
+  {\path{doi:10.1109/TC.2004.85}}.
+\newline\urlprefix\url{http://doi.ieeecomputersociety.org/10.1109/TC.2004.85}
+
+\bibitem{simao_reducing_2012}
+A.~Simão, A.~Petrenko, N.~Yevtushenko,
+  \href{https://onlinelibrary.wiley.com/doi/abs/10.1002/stvr.452}{On reducing
+  test length for {FSMs} with extra states}, Software Testing, Verification and
+  Reliability 22~(6) (2012) 435--454.
+\newblock \href {http://dx.doi.org/10.1002/stvr.452}
+  {\path{doi:10.1002/stvr.452}}.
+\newline\urlprefix\url{https://onlinelibrary.wiley.com/doi/abs/10.1002/stvr.452}
+
+\bibitem{Springintveld2001}
+J.~Springintveld, F.~Vaandrager, P.~D'Argenio, Testing timed automata,
+  Theoretical Computer Science 254~(1-2) (2001) 225--257.
+
+\bibitem{DBLP:journals/acta/CavalcantiG11}
+A.~Cavalcanti, M.-C. Gaudel, Testing for refinement in {\it circus}, Acta Inf.
+  48~(2) (2011) 97--147.
+
+\bibitem{Schneider:1995:OST:203471.203475}
+S.~Schneider, \href{http://dx.doi.org/10.1006/inco.1995.1014}{An operational
+  semantics for timed csp}, Inf. Comput. 116~(2) (1995) 193--213.
+\newblock \href {http://dx.doi.org/10.1006/inco.1995.1014}
+  {\path{doi:10.1006/inco.1995.1014}}.
+\newline\urlprefix\url{http://dx.doi.org/10.1006/inco.1995.1014}
+
+\bibitem{DBLP:journals/cn/Tretmans96}
+J.~Tretmans, Conformance testing with labelled transition systems:
+  Implementation relations and test generation, Computer Networks and ISDN
+  Systems 29~(1) (1996) 49--79.
+
+\bibitem{DBLP:conf/icst/Petrenko16}
+A.~Petrenko, \href{http://dx.doi.org/10.1109/ICSTW.2016.9}{Checking experiments
+  for symbolic input/output finite state machines}, in: Ninth {IEEE}
+  International Conference on Software Testing, Verification and Validation
+  Workshops, {ICST} Workshops 2016, Chicago, IL, USA, April 11-15, 2016, {IEEE}
+  Computer Society, 2016, pp. 229--237.
+\newblock \href {http://dx.doi.org/10.1109/ICSTW.2016.9}
+  {\path{doi:10.1109/ICSTW.2016.9}}.
+\newline\urlprefix\url{http://dx.doi.org/10.1109/ICSTW.2016.9}
+
+\bibitem{Huang2017}
+W.-l. Huang, J.~Peleska,
+  \href{http://dx.doi.org/10.1007/s00165-016-0402-2}{Complete model-based
+  equivalence class testing for nondeterministic systems}, Formal Aspects of
+  Computing 29~(2) (2017) 335--364.
+\newblock \href {http://dx.doi.org/10.1007/s00165-016-0402-2}
+  {\path{doi:10.1007/s00165-016-0402-2}}.
+\newline\urlprefix\url{http://dx.doi.org/10.1007/s00165-016-0402-2}
+
+\bibitem{Hoare:1985:CSP:3921}
+C.~A.~R. Hoare, Communicating Sequential Processes, Prentice-Hall, Inc., Upper
+  Saddle River, NJ, USA, 1985.
+
+\bibitem{Roscoe2010}
+A.~W. Roscoe, Understanding Concurrent Systems, Springer, London, Dordrecht
+  Heidelberg, New York, 2010.
+
+\bibitem{976937}
+A.~Hall, R.~Chapman, Correctness by construction: developing a commercial
+  secure system, IEEE Software 19~(1) (2002) 18--25.
+\newblock \href {http://dx.doi.org/10.1109/52.976937}
+  {\path{doi:10.1109/52.976937}}.
+
+\bibitem{DBLP:conf/prdc/ShiPK99}
+H.~Shi, J.~Peleska, M.~Kouvaras,
+  \href{https://doi.org/10.1109/PRDC.1999.816222}{Combining methods for the
+  analysis of a fault-tolerant system}, in: 1999 Pacific Rim International
+  Symposium on Dependable Computing {(PRDC} 1999), 16-17 December 1999, Hong
+  Kong, {IEEE} Computer Society, 1999, pp. 135--142.
+\newblock \href {http://dx.doi.org/10.1109/PRDC.1999.816222}
+  {\path{doi:10.1109/PRDC.1999.816222}}.
+\newline\urlprefix\url{https://doi.org/10.1109/PRDC.1999.816222}
+
+\bibitem{DBLP:conf/amast/ButhKPS97}
+B.~Buth, M.~Kouvaras, J.~Peleska, H.~Shi,
+  \href{https://doi.org/10.1007/BFb0000463}{Deadlock analysis for a
+  fault-tolerant system}, in: M.~Johnson (Ed.), Algebraic Methodology and
+  Software Technology, 6th International Conference, {AMAST} '97, Sydney,
+  Australia, December 13-17, 1997, Proceedings, Vol. 1349 of Lecture Notes in
+  Computer Science, Springer, 1997, pp. 60--74.
+\newblock \href {http://dx.doi.org/10.1007/BFb0000463}
+  {\path{doi:10.1007/BFb0000463}}.
+\newline\urlprefix\url{https://doi.org/10.1007/BFb0000463}
+
+\bibitem{Roscoe:1994:chapter}
+A.~W. Roscoe, Model-checking csp, in: A.~W. Roscoe (Ed.), A Classical Mind:
+  Essays in Honour of C. A. R. Hoare, Prentice Hall International (UK) Ltd.,
+  Hertfordshire, UK, UK, 1994, Ch.~21, pp. 353--378.
+
+\bibitem{Hennessy:1988:ATP:50497}
+M.~Hennessy, Algebraic Theory of Processes, MIT Press, Cambridge, MA, USA,
+  1988.
+
+\bibitem{DBLP:conf/fm/PeleskaS96}
+J.~Peleska, M.~Siegel, \href{https://doi.org/10.1007/3-540-60973-3\_106}{From
+  testing theory to test driver implementation}, in: M.~Gaudel, J.~Woodcock
+  (Eds.), {FME} '96: Industrial Benefit and Advances in Formal Methods, Third
+  International Symposium of Formal Methods Europe, Co-Sponsored by {IFIP} {WG}
+  14.3, Oxford, UK, March 18-22, 1996, Proceedings, Vol. 1051 of Lecture Notes
+  in Computer Science, Springer, 1996, pp. 538--556.
+\newblock \href {http://dx.doi.org/10.1007/3-540-60973-3\_106}
+  {\path{doi:10.1007/3-540-60973-3\_106}}.
+\newline\urlprefix\url{https://doi.org/10.1007/3-540-60973-3\_106}
+
+\bibitem{peleska1997a}
+J.~Peleska, M.~Siegel, Test automation of safety-critical reactive systems,
+  South African Computer Jounal 19 (1997) 53--77.
+
+\bibitem{DBLP:conf/icfem/CavalcantiG07}
+A.~Cavalcanti, M.~Gaudel,
+  \href{https://doi.org/10.1007/978-3-540-76650-6\_10}{Testing for refinement
+  in {CSP}}, in: M.~J. Butler, M.~G. Hinchey, M.~M. Larrondo{-}Petrie (Eds.),
+  Formal Methods and Software Engineering, 9th International Conference on
+  Formal Engineering Methods, {ICFEM} 2007, Boca Raton, FL, USA, November
+  14-15, 2007, Proceedings, Vol. 4789 of Lecture Notes in Computer Science,
+  Springer, 2007, pp. 151--170.
+\newblock \href {http://dx.doi.org/10.1007/978-3-540-76650-6\_10}
+  {\path{doi:10.1007/978-3-540-76650-6\_10}}.
+\newline\urlprefix\url{https://doi.org/10.1007/978-3-540-76650-6\_10}
+
+\bibitem{DBLP:conf/pts/CavalcantiS17}
+A.~Cavalcanti, A.~da~Silva~Sim{\~{a}}o,
+  \href{https://doi.org/10.1007/978-3-319-67549-7\_2}{Fault-based testing for
+  refinement in {CSP}}, in: N.~Yevtushenko, A.~R. Cavalli, H.~Yenig{\"{u}}n
+  (Eds.), Testing Software and Systems - 29th {IFIP} {WG} 6.1 International
+  Conference, {ICTSS} 2017, St. Petersburg, Russia, October 9-11, 2017,
+  Proceedings, Vol. 10533 of Lecture Notes in Computer Science, Springer, 2017,
+  pp. 21--37.
+\newblock \href {http://dx.doi.org/10.1007/978-3-319-67549-7\_2}
+  {\path{doi:10.1007/978-3-319-67549-7\_2}}.
+\newline\urlprefix\url{https://doi.org/10.1007/978-3-319-67549-7\_2}
+
+\bibitem{5533149}
+L.~Shi, X.~Cai, An exact fast algorithm for minimum hitting set, in: 2010 Third
+  International Joint Conference on Computational Science and Optimization,
+  Vol.~1, 2010, pp. 64--67.
+\newblock \href {http://dx.doi.org/10.1109/CSO.2010.240}
+  {\path{doi:10.1109/CSO.2010.240}}.
+
+\bibitem{chow:wmethod}
+T.~S. Chow, Testing software design modeled by finite-state machines, IEEE
+  Transactions on Software Engineering SE-4~(3) (1978) 178--186.
+
+\bibitem{vasilevskii1973}
+M.~P. Vasilevskii, Failure diagnosis of automata, Kibernetika (Transl.) 4
+  (1973) 98--108.
+
+\bibitem{luo_test_1994}
+G.~Luo, G.~von Bochmann, A.~Petrenko,
+  \href{http://doi.ieeecomputersociety.org/10.1109/32.265636}{Test selection
+  based on communicating nondeterministic finite-state machines using a
+  generalized wp-method}, {IEEE} Trans. Software Eng. 20~(2) (1994) 149--162.
+\newblock \href {http://dx.doi.org/10.1109/32.265636}
+  {\path{doi:10.1109/32.265636}}.
+\newline\urlprefix\url{http://doi.ieeecomputersociety.org/10.1109/32.265636}
+
+\bibitem{fdr}
+T.~Gibson-Robinson, P.~Armstrong, A.~Boulgakov, A.~Roscoe, {FDR3 --- A Modern
+  Refinement Checker for CSP}, in: E.~Ábrahám, K.~Havelund (Eds.), Tools and
+  Algorithms for the Construction and Analysis of Systems, Vol. 8413 of Lecture
+  Notes in Computer Science, 2014, pp. 187--201.
+
+\bibitem{Roscoe:1997:TPC:550448}
+A.~W. Roscoe, The Theory and Practice of Concurrency, Prentice Hall PTR, Upper
+  Saddle River, NJ, USA, 1997.
+
+\bibitem{sperner_satz_1928}
+E.~Sperner, \href{https://doi.org/10.1007/BF01171114}{Ein {Satz} {\"u}ber
+  {Untermengen} einer endlichen {Menge}}, Mathematische Zeitschrift 27~(1)
+  (1928) 544--548.
+\newblock \href {http://dx.doi.org/10.1007/BF01171114}
+  {\path{doi:10.1007/BF01171114}}.
+\newline\urlprefix\url{https://doi.org/10.1007/BF01171114}
+
+\bibitem{PeleskaHuangLectureNotesMBT}
+J.~Peleska, W.-l. Huang, Test Automation - Foundations and Applications of
+  Model-based Testing, University of Bremen, 2017, lecture notes, available
+  under
+  http://www.informatik.uni-bremen.de/agbs/jp/papers/test-automation-huang-peleska.pdf.
+
+\bibitem{DBLP:conf/forte/DorofeevaEY05}
+R.~Dorofeeva, K.~El{-}Fakih, N.~Yevtushenko,
+  \href{https://doi.org/10.1007/11562436_16}{An improved conformance testing
+  method}, in: F.~Wang (Ed.), Formal Techniques for Networked and Distributed
+  Systems - {FORTE} 2005, 25th {IFIP} {WG} 6.1 International Conference,
+  Taipei, Taiwan, October 2-5, 2005, Proceedings, Vol. 3731 of Lecture Notes in
+  Computer Science, Springer, 2005, pp. 204--218.
+\newblock \href {http://dx.doi.org/10.1007/11562436_16}
+  {\path{doi:10.1007/11562436_16}}.
+\newline\urlprefix\url{https://doi.org/10.1007/11562436_16}
+
+\bibitem{petrenko_testing_2011}
+A.~Petrenko, N.~Yevtushenko, Adaptive testing of deterministic implementations
+  specified by nondeterministic fsms, in: Testing Software and Systems, no.
+  7019 in Lecture Notes in Computer Science, Springer, Berlin, Heidelberg,
+  2011, pp. 162--178.
+
+\bibitem{CG15}
+A.~L.~C. Cavalcanti, M.-C. Gaudel,
+  \href{https://www-users.cs.york.ac.uk/~alcc/publications/papers/CG15.pdf}{Test
+  selection for traces refinement}, Theoretical Computer Science 563~(0) (2015)
+  1 -- 42.
+\newblock \href {http://dx.doi.org/10.1016/j.tcs.2014.08.012}
+  {\path{doi:10.1016/j.tcs.2014.08.012}}.
+\newline\urlprefix\url{https://www-users.cs.york.ac.uk/~alcc/publications/papers/CG15.pdf}
+
+\end{thebibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-04-14 scp-journal/article-revision-1/main.tex
New directory for revision 1
Handled all remarks of reviewer #4

--- 2018-11-30 scp-journal/article/main.tex
@@
 processes and processes for SUTs. So, the labelling of nodes with maximal
-refusals or minimal acceptances are disregarded in the product construction.
+refusals or minimal acceptances is disregarded in the product construction.
 We consider two normalised transition graphs
@@
 \item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \minhits(P/s)$, $H$ is
-a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
+a (not necessarily minimal) hitting set of $\minaccs(Q/s)$.
 \end{enumerate}
@@
 $\ist$. This is explained further in the next paragraphs. Alternatively, a
-faulty SUT $Q$ might internally deadlock after a trace $s$ where $\#s < j$
-and $\minhits(G(P)/s) \neq \varnothing$, so that the process
+faulty SUT $Q$ might internally deadlock after a trace $s$ whose length $\#s$ 
+satisfies  $\#s < j$, such that
+  $\minhits(G(P)/s) \neq \varnothing$, so that the process
 $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks as well.
@@
   \\
-  \minaccs(P/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
+  \minaccs(\pmax/s) & = & {\cal C} \quad\text{for all $s\in \Sigma^*$}
   \label{eq:pmaxc}
   \\
-  \minhits(P/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
+  \minhits(\pmax/s) & = & \minhits({\cal C}) \quad\text{for all $s\in \Sigma^*$}
   \label{eq:pmaxe}
   \\
-  \card{\minhits(P/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor}
+  \card{\minhits(\pmax/s)} & = & \binom{n}{\lfloor\frac{n}{2}\rfloor}
   \quad \text{for all $s\in \Sigma^*$}

### 2019-04-14 scp-journal/article-revision-1/main.tex
Changes w.r.t. remarks of reviewer #3

@@
 The collection $\minsub({\cal C})$ contains all those sets of
-${\cal C}$ that are not true subsets of other members of ${\cal C}$. With
+${\cal C}$ that are not true supersets of other members of ${\cal C}$. With
 this definition, the relation between failures and minimal hitting sets of
@@
 ready to engage in: if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the
-occurrence of these events can never be blocked due to a livelock in $Q$
-occurring in the same step of the execution.
+occurrence of these events can never be blocked due to a livelock, since the only
+  process executing concurrently with  $U_F(j)$  is livelock-free  $Q$.
 
@@
 traces refinement? We consider the first question~(1) in
-Section~\ref{section:complexity:failures}), where we also discuss whether it
+Section~\ref{section:complexity:failures}, where we also discuss whether  
 it is possible to reduce the number of test executions with a different test
 suite.  With the answer to question~(1), question~(2) is a fairly simple
-consequence we discuss briefly also in
+consequence we discuss briefly at the end of 
 Section~\ref{section:complexity:failures}. Question~(3) is the subject of
@@
 $n$ elements, there is no Sperner family consisting of more than
-$\binom{n}{\lfloor\frac{n}{2}\rfloor}$. In addition, we recall the minimal
+$\binom{n}{\lfloor\frac{n}{2}\rfloor}$ members. In addition, we recall the minimal
 hitting sets calculated from a given set of minimal acceptances are a Sperner
 family~(see discussion in Section~\ref{sec:sperner}). So, (\ref{eq:pmaxd})
-establishes that there is no possibility of providing more choices fom
+establishes that there is no possibility of providing more choices from
 minimal hitting sets with a test derived from a process different from

### 2019-04-14 scp-journal/article-revision-1/main.tex
First changes regading reviewer 1

@@
 
-Since we have finite complete CSP test suites, it is useful for the first
-time to calculate how many test executions are needed when using them.
+With finite complete CSP test suites test suites at hand, it is now possible 
+to calculate how many test executions are needed when using them.
 Previous work did not consider sufficient conditions for finiteness, so
-complexity was not a concern. We answer the following questions. (1)~What is
+estimates for the maximal number 
+of executions could not be calculated. 
+We answer the following questions. (1)~What is
 the worst-case bound on the number of test executions to be performed to
@@
 nor in the traces-refinement relation. This can only be seen when executing
-the test $U_F(pq-1)$ and $U_T(pq-1)$, respectively. These tests fail, so this
-shows that $P\not\lessdet_F Q$ and $P\not\lessdet_T Q$ according to
-Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}. Moreover, this
+the test $U_F(pq-1)$ and $U_T(pq-1)$, respectively. These tests fail, because
+they allow for execution of the following trace of length 12
+$$
+s_\text{fail} = a.a.a.b.a.a.a.b.a.a.a.b
+$$
+Only in the very last step, the final $b$-event can be produced by $Q$ but is refused by $P$.
+Therefore, tests $U_F(11)$ and $U_T(11)$ enter their first branch (\ref{eq:ufa}) and (\ref{eq:uta}), respectively,\ when engaging into the last $b$, and this produces the $\efail$-event and shows  that $P\not\lessdet_F Q$ and $P\not\lessdet_T Q$ according to
+Theorem~\ref{th:failurestest} and Theorem~\ref{th:tracetest}.. The tests 
+$U_F(j)$ and $U_T(j)$ with $j < pq-1 = 11$ only execute true prefixes of $s_\text{fail}$ which 
+are allowed for $P$, so that these tests pass.
+
+This
 shows that the maximal trace length $pq$ to be investigated in the tests

### 2019-04-15 scp-journal/article-revision-1/main.tex
All remarks but the first two of reviewer #1 have been processed.

@@
 
-\newpage
-\noindent %
 The main novel contributions in this article may be summarised as
@@
 
+The branches of an external choice can be protected by \emph{guards}. The
+ process
+$$
+P(x_1,\dots,x_n) = \text{bexpr}_1(x1,\dots,x_n) \& Q(x_1,\dots,x_n) \extchoice 
+\text{bexpr}_2(x1,\dots,x_n) \& R(x_1,\dots,x_n)
+$$
+parametrised over $x_1,\dots,x_n$ branches according to Boolean expressions
+$\text{bexpr}_1(x1,\dots,x_n)$ and $\text{bexpr}_2(x1,\dots,x_n)$ over 
+variables from $\{ x_1,\dots,x_n\}$: if $\text{bexpr}_1(x1,\dots,x_n)$ evaluates to
+$\ist$ and $\text{bexpr}_2(x1,\dots,x_n)$ to $\isf$, $P$ behaves like $Q$. 
+If $\text{bexpr}_1(x1,\dots,x_n)$ evaluates to
+$\isf$ and $\text{bexpr}_2(x1,\dots,x_n)$ to $\ist$, $P$ behaves like $R$. If both
+Boolean expressions evaluate to $\ist$, $P$ behaves as $Q\extchoice R$. Finally, if
+both Boolean expressions evaluate to $\isf$, $P$ behaves like $\Stop$.
+
 There are several parallelism operators. A widely used form of parallelism $P
@@
 defined by $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
-$P/s$, that is, the events that $P/s$ may engage into. The maximal refusals
-in combination with the initials of a process express its degree of
-nondeterminism as illustrated by the next example.
+$P/s$, that is, the events that $P/s$ may engage into. Nondeterministic behaviour in a given process
+state is reflected by non-empty intersections between initials and maximal refusals. This 
+is illustrated by the following example.
 %
@@
 \paragraph{Reachability Under Sets of Traces} \label{sec:V} Given a finite-state CSP
-process $P$ and its normalised transition graph $G(P)$,
+process $P$ and its normalised transition graph $G(P)$ with nodes in set $N$,
 %\[
@@
 \label{lemma:extendV} Let $P$ be a CSP process with normalised transition
-graph $G(P)$. Let $V\subseteq\Sigma^*$ be a finite prefix-closed set of
-sequences of events. Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$,
-that is, $|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences
+graph $G(P) = ( N, {\ii n}, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma))$. Let $V\subseteq\Sigma^*$ be a finite prefix-closed set of
+sequences of events. Suppose that  $G(P)$ reaches $k < \card{N}$ nodes under $V$,
+that is, $\card{t(\ii n,V)} = k$. Let $V.\Sigma$ denote the set of all sequences
 from $V$, extended by any event of $\Sigma$.  Then $G(P)$ reaches at least
@@
 $ %\[
-\minhits(P/s) = \minsub {\cal A}_s
+\minhits(P/s) = \minsub({\cal A}_s)
 $ %\]
@@
 \\
-\Leftrightarrow   \forall B\in minAcc(P/s): A\not\subseteq (\Sigma\setminus
+\Leftrightarrow   \forall B\in minAcc(P/s): A\not\subseteq (\Sigma -
 B)& this follows from (\ref{eq:accref})
@@
 %
-Specialising this derivation valid for arbitrary $A$ satisfying $(s,A)\not\in {\text{failures}}(P)$ to
-minimal $A$ with this property proves the statement of the lemma.
+This derivation is valid  for arbitrary $A$ satisfying $(s,A)\not\in {\text{failures}}(P)$.
+Specialising it to {\it minimal} sets $A$ satisfying  $(s,A)\not\in {\text{failures}}(P)$
+proves the statement of the lemma.
 \xbox
@@
 
-\newpage
-\noindent
 We further observe that
@@
 % --------------------------------------------------------------------------
-\paragraph{Explanation of the Test Definition} A test \pagebreak is performed by running
+\paragraph{Explanation of the Test Definition} A test  is performed by running
 $U_F(j)$ concurrently with any SUT process $Q$, synchronising over $\Sigma$.
@@
 %%\vspace*{-10mm}
-\caption{Normalised transition graph of faulty implementation $Z$  from Example~\ref{ex:uf1tests}.}
+\caption{Normalised transition graph of faulty implementation $Z$ for $r_\text{max} = 3$ from Example~\ref{ex:uf1tests}.}
  \label{fig:tgZ}
@@
 produced. Another failing execution arises if $Z/s$ chooses to accept only
-$\{c \}$, while $U_F(4)/s$ choses to accept only $\{a,b\}$. Therefore,
+$\{c \}$, while $U_F(4)/s$ chooses to accept only $\{a,b\}$. Therefore,
 $%\[
@@
 
-For establishing traces refinement, the following class of adaptive test
-cases are used for a given reference process $P$ and  integers $j \ge 0$.
+For establishing traces refinement, the   class of adaptive test
+cases specified below in (\ref{eq:utjstart}) -- (\ref{eq:utd})
+is used for a given reference process $P$ and  integers $j \ge 0$.
 Just as for the tests developed  in Section~\ref{sec:finitecompletefails} to
@@
 (\ref{eq:utb}). From (\ref{eq:minhitminaccempty}) we know that these
-conditions are equivalent, but, with the use of $\minaccs(n) = \{ \varnothing
-\}$, \pagebreak we make it unnecessary to calculate hitting sets for
+conditions are equivalent, but, by using  expression $\minaccs(n) = \{ \varnothing
+\}$,  we make it unnecessary to calculate hitting sets for
 generating these tests from $G(P)$, which is expensive.
@@
 U_T(j) & = & U_T(j,0,\ii n)
+\label{eq:utjstart}
 \\
@@
 $$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$
-because of (\ref{eq:ifpaT}). \pagebreak Now $\trc(Q)\subseteq\trc(P)$ implies
+because of (\ref{eq:ifpaT}).  Now $\trc(Q)\subseteq\trc(P)$ implies
 $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot

### 2019-04-16 scp-journal/article-revision-1/main.tex
Prepare new paragraph in Discussion section
Addressed first remark of reviewer#1 by adding new paragraph in the discussion & conclusion.

@@
 % ==============================================================================
+\paragraph{Practical Application to Embedded Systems Testing}
+
+Test models developed in  CSP  are particularly useful for testing embedded control systems; see~\cite{peleska2002tocs,Peleska02}, for example, for reports on practical application.
+Using complete test methods is particularly attractive for justifying the test case selection when
+verifying safety-critical control systems. The complexity estimates in Section~\ref{sec:complexity}, however,  
+indicate that completeness will result in considerable test effort   depending exponentially
+on the number of states in reference model and implementation. Moreover,  the complexity considerations showed that nondeterminism 
+increases the number of test cases further by a significant factor.
+
+Therefore, we expect that
+the test strategy presented in this paper will only be applicable to rather small control systems,
+when used directly   for   HW/SW integration testing, where tests are executed on the target hardware and in 
+real physical time.  There are, however, several options to circumvent or at least mitigate
+the complexity problem.
+
+\medskip
+\noindent
+(1) Recent research on the certifiability of autonomous systems advocates 
+a multi-level approach to testing. In the first stage, very many tests  are executed concurrently
+in a cloud configuration.\footnote{An informal overview describing the ongoing discussion about virtualisation of tests and simulations in the cloud can be found in https://www.linkedin.com/pulse/virtual-testing-qualification-autonomous-vehicles-lopez-rodriguez.}   In the second stage, a smaller subset of these tests is then executed against the real physical embedded system.
+We expect that the complete testing approach described here would typically be applied for stage 1 in the cloud configuration.
+
+\medskip
+\noindent
+(2) The number of test cases to consider in a complete test suite can be significantly reduced by the introduction of equivalence classes.
+It has been shown in~\cite{Huang2017} that complete testing theories for finite state machines induce likewise complete
+equivalence class testing 
+strategies for more complex modelling formalisms, such as Kripke Structures or CSP.
+Experimental evaluations have shown that with the help of such classes, complete test suites with manageable size can be generated~\cite{DBLP:conf/rssrail/PeleskaHH16,Huebner2017}.
+The experiments considered on-board speed monitors for trains, airbag controllers, and route controllers for railway interlocking systems.
+
+\medskip
+\noindent
+(3) A further promising option is to eliminate nondeterminism. For safety-critical systems, nondeterminism usually occurs in environment models or in high-level models of the target system for
+specifying behavioural options. The implementation itself is typically deterministic. As a consequence, the high-level model can be refined until all design decisions are included, so that the 
+lower-level models are deterministic.
+Now the SUT could be tested for \emph{equivalence} to the lower-level model.
+The lower-level model would typically have more states than the higher-level one, but
+-- as mentioned at the beginning of this section -- maximal trace length would now be only $3q-q$ instead of $pq$
+(recall that $p$ denotes the number of states in the reference model, and $q$ the maximal number of states in the implementation).
+Therefore, complete test suites checking trace or failures equivalence are of considerable interest as well.
+
+\medskip
+\noindent
+(4) If the ultimate goal is to establish failures refinement, there is a useful  heuristic for reducing the test effort while the SUT is probably still faulty (e.g.~for newly developed components).
+We start with the longest test cases $U_F(pq-1), U_F(pq-2), \dots$, and only perform
+the shorter tests $U_F(j)$ with small $j$ when the longer ones do not reveal any errors. This is
+very effective, since longer test cases completely check trace refinement, before hitting sets are
+applied to check admissibility of nondeterministic behaviour at the end of the longest traces. 
+As a consequence, trace errors will be quickly revealed before fully covering the failures-related tests.
+
+The question whether the SUT is inside the specified fault domain can usually be answered 
+by performing static analyses of the SUT code or by monitoring state coverage and decision coverage during test executions. If the code has been automatically generated
+from a low-level model, this is further simplified when the code generator's strategy for introducing auxiliary variables (if any) is known.
+Even in the case where it cannot be determined whether the SUT is inside the fault domain, the experiments from~\cite{DBLP:conf/rssrail/PeleskaHH16,Huebner2017} have shown that the test suites generated according to a complete strategy still have superior test strength, when compared to naive heuristic testing approaches.
+
+% ==============================================================================
 \paragraph{Conclusion}
@@
 \bibitem{simao_reducing_2012}
-A.~Simão, A.~Petrenko, N.~Yevtushenko,
+A.~Sim{\~{a}}o, A.~Petrenko, N.~Yevtushenko,
   \href{https://onlinelibrary.wiley.com/doi/abs/10.1002/stvr.452}{On reducing
@@
 
+\bibitem{peleska2002tocs}
+J.~Peleska, Hardware/software integration testing for the new airbus aircraft
+  families, in: I.~Schieferdecker, H.~K{\"o}nig, A.~Wolisz (Eds.), {Testing of
+  Communicating Systems XIV. Application to Internet Technologies and
+  Services}, Kluwer Academic Publishers, 2002, pp. 225--351.
+
+\bibitem{Peleska02}
+J.~Peleska, Formal methods for test automation - hard real-time testing of
+  controllers for the airbus aircraft family, in: Proc. of the Sixth Biennial
+  World Conference on Integrated Design \& Process Technology (IDPT2002),
+  Pasadena, California, June 23-28, 2002, Society for Design and Process
+  Science, 2002, iSSN 1090-9389.
+
+\bibitem{DBLP:conf/rssrail/PeleskaHH16}
+J.~Peleska, W.~Huang, F.~H{\"{u}}bner,
+  \href{http://dx.doi.org/10.1007/978-3-319-33951-1_3}{A novel approach to
+  {HW/SW} integration testing of route-based interlocking system controllers},
+  in: T.~Lecomte, R.~Pinger, A.~Romanovsky (Eds.), Reliability, Safety, and
+  Security of Railway Systems. Modelling, Analysis, Verification, and
+  Certification - First International Conference, RSSRail 2016, Paris, France,
+  June 28-30, 2016, Proceedings, Vol. 9707 of Lecture Notes in Computer
+  Science, Springer, 2016, pp. 32--49.
+\newblock \href {http://dx.doi.org/10.1007/978-3-319-33951-1_3}
+  {\path{doi:10.1007/978-3-319-33951-1_3}}.
+\newline\urlprefix\url{http://dx.doi.org/10.1007/978-3-319-33951-1_3}
+
+\bibitem{Huebner2017}
+F.~H{\"u}bner, W.-l. Huang, J.~Peleska,
+  \href{https://doi.org/10.1007/s10270-017-0595-8}{Experimental evaluation of a
+  novel equivalence class partition testing strategy}, Software {\&} Systems
+  ModelingDoi=10.1007/s10270-017-0595-8.
+\newblock \href {http://dx.doi.org/10.1007/s10270-017-0595-8}
+  {\path{doi:10.1007/s10270-017-0595-8}}.
+\newline\urlprefix\url{https://doi.org/10.1007/s10270-017-0595-8}
+
 \end{thebibliography}

### 2019-04-16 scp-journal/article-revision-1/main.tex
ana reviewer

@@
 
-A prefixing operator $e \then P$ defines a process that is ready to engage in
+Table~\ref{table:operators} lists the CSP operators we use in our examples. A
+prefixing operator $e \then P$ defines a process that is ready to engage in
 the event $e$, pending agreement of its environment to synchronise. After $e$
@@
 processes, in parallel, or the environment of a system as a whole.
+
+\begin{table}\centering
+  \begin{tabular}{|l|l|l|}
+    \hline
+  	$Stop$ & Deadlock process
+    \\
+  	$a \then P$ & Prefixed operator: initially offers to engage in the event $a$, and then behaves as a process $P$
+    \\
+  	$g \& P$ & Guarded process: behaves as $P$ if the predicate $g$ is true. Otherwise, it behaves like $STOP$
+    \\
+  	$P \extchoice Q$ & External choice of $P$ or $Q$ made by the environment
+    \\
+    $\Extchoice i: I @ P(i)$ & Iterated external choice
+    \\
+  	$P \intchoice Q$ & Internal choice of $P$ or $Q$
+    \\
+    $\Intchoice i: I @ P(i)$ & Iterated internal choice
+    \\
+  	$P \hide X $ & Hiding: behaves like $P$ but with all communications in the set $X$ hidden
+    \\
+  	$P \parallel[X] Q $ & Generalised parallel: $P$ and $Q$ must synchronise on events that belong to the set $X$
+    \\
+    \hline
+  \end{tabular}
+  \caption{{CSP operators:~$P$ and $Q$ are metavariables that stand for processes, $a$ for an event, $g$ for a condition, $X$ for a set of events, and $I$ for a set.}}
+  \label{table:operators}
+\end{table}
 
@@
 processes $P(i)$. If the index set $I$ is empty, the external choice is the
-process $\Stop$, which deadlocks:~does not engage into any event or
-terminate. For an external choice $\Extchoice e: A @ e \then P(e), A\subseteq
-\Sigma$ over a set $A$ of events, we use the abbreviation $e:A \then P(e)$.
-An iterated internal choice is not defined for an empty index set.
-
-The branches of an external choice can be protected by \emph{guards}. The
- process
+process $\Stop$, which deadlocks:~does not engage into any event or terminate
+successfully. For an external choice $\Extchoice e: A @ e \then P(e),
+A\subseteq \Sigma$ over a set $A$ of events, we use the abbreviation $e:A
+\then P(e)$. An iterated internal choice is not defined for an empty index
+set.
+
+The branches of an external choice can be protected by guards. The process
 $$
@@
 
-Semantic definitions specify, for each operator, how the traces or failures
-of the resulting process can be calculated from those of each operand. For
-example, for internal choice, $\fails(P\intchoice Q) =
+Denotational semantic definitions specify, for each operator, how the traces
+or failures of the resulting process can be calculated from those of each
+operand. For example, for internal choice, $\fails(P\intchoice Q) =
 \fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for a
 comprehensive list of these definitions covering $\trc(P)$ and $\fails(P)$.
+In this work we use the CSP denotational semantics to justify our results. We
+note, however, that the sets of traces and failures of a processes can be
+equivalently defined using the operational semantics, as proved
+in~\cite{Roscoe2010}. The operational semantics is congruent with the traces
+and failures models. This justifies our use of transition graphs as well.
 
@@
 Deterministic process states $P/s$ have exactly the one maximal refusal
-defined by $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
+$\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
 $P/s$, that is, the events that $P/s$ may engage into. Nondeterministic behaviour in a given process
@@
 refusals or minimal acceptances is disregarded in the product construction.
-We consider two normalised transition graphs
+We consider two normalised transition graphs over the same alphabet $\Sigma$.
 \[
@@
 \]
-over the same alphabet $\Sigma$. Their product is defined by
+Their product is defined by
 %
@@
 %
-The following lemma is used in the proof of our main theorem.
+The following lemma is used later in the proof of our main theorem.
 %
@@
 $A_i$ is a subset of a universe $\Sigma$, a \emph{hitting set}
-$H\subseteq\Sigma$ is a set satisfying the following property.
+$H\subseteq\Sigma$ is defined as a set satisfying the following property.
 %
@@
 \end{equation}
-The collection $\minsub({\cal C})$ contains all those sets of
-${\cal C}$ that are not true supersets of other members of ${\cal C}$. With
-this definition, the relation between failures and minimal hitting sets of
-minimal acceptances is established in the following lemma.
+The collection $\minsub({\cal C})$ contains all those sets of ${\cal C}$ that
+are not true supersets of other members of ${\cal C}$. With this definition,
+the relation between failures and minimal hitting sets of minimal acceptances
+is established as follows.
 %
@@
 conf\ Q$ does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
-(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
+(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ that satisfies
+the following property.
 \[
@@
 and PASS for the test execution, are not included in $\Sigma$. Since we
-assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
-or $\epass$ always become visible, if they are the only events $U_F(j)/s$ is
-ready to engage in: if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the
-occurrence of these events can never be blocked due to a livelock, since the only
+assume that $Q$ is free of livelocks, the events $\efail$ or $\epass$ always
+become visible, if they are the only events $U_F(j)/s$ is ready to engage in:
+if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the occurrence of these
+events can never be blocked due to a livelock, since the only
   process executing concurrently with  $U_F(j)$  is livelock-free  $Q$.
@@
 Running the test $U_F(j)$ against $Z$ for $j=0,\dots,19$ ($G(P)$ has $p = 4$
-states and $G(Z)$ has $q=5$, so $pq-1=19$ is the index of the last test to
-be executed  according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
+states and $G(Z)$ has $q=5$, so $pq-1=19$ is the index of the last test to be
+executed  according to Theorem~\ref{th:failurestest}), tests $U_F(0),\dots,
 U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$, because after execution of
-the trace
+the following trace.
 \[
-s = a.c.c.c, \qquad\text{(note that $G(P)/s = \text{node}\ 3$ according to Fig.~\ref{fig:tga})},
+s = a.c.c.c \qquad\text{(Note that $G(P)/s = \text{node}\ 3$ according to Fig.~\ref{fig:tga}.)}
 \]
-the test $U_F(4)$ offers hitting sets from $\minhits(3) = \{\{b\},\{c\}\}$
-in branch (\ref{eq:ufd}). Therefore,  there exists one test execution
-where $Z/s$ accepts only $\{b\}$ due to the internal choice (note
-from Fig.~\ref{fig:tgZ} that
-$G(Z)/s = \text{node}\ 4$), while $U_F(4)/s$
-only offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for
+The test $U_F(4)$ offers hitting sets from $\minhits(3) = \{\{b\},\{c\}\}$ in
+branch (\ref{eq:ufd}). Therefore,  there exists one test execution where
+$Z/s$ accepts only $\{b\}$ due to the internal choice (note from
+Fig.~\ref{fig:tgZ} that $G(Z)/s = \text{node}\ 4$), while $U_F(4)/s$ only
+offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for
 branch (\ref{eq:ufa}).  As a consequence, this execution of
@@
 
-With finite complete CSP test suites test suites at hand, it is now possible 
-to calculate how many test executions are needed when using them.
-Previous work did not consider sufficient conditions for finiteness, so
-estimates for the maximal number 
-of executions could not be calculated. 
-We answer the following questions. (1)~What is
-the worst-case bound on the number of test executions to be performed to
-verify an SUT with respect to failures refinement, when we use our test
-suite? (2)~What is the worst-case bound for traces refinement? (3)~Is it
-possible to reduce the maximal length of traces when testing for failures or
-traces refinement? We consider the first question~(1) in
-Section~\ref{section:complexity:failures}, where we also discuss whether  
-it is possible to reduce the number of test executions with a different test
-suite.  With the answer to question~(1), question~(2) is a fairly simple
-consequence we discuss briefly at the end of 
-Section~\ref{section:complexity:failures}. Question~(3) is the subject of
-Section~\ref{section:complexity:length}.
+With finite complete CSP test suites, it is possible to calculate how
+many test executions are needed when using them. Previous work did not
+consider sufficient conditions for finiteness, so estimates for the maximal
+number of executions could not be calculated. We answer the following
+questions. (1)~What is the worst-case bound on the number of test executions
+to verify an SUT with respect to failures refinement, when we
+use our test suite? (2)~What is the worst-case bound for traces refinement?
+(3)~Is it possible to reduce the maximal length of traces when testing for
+failures or traces refinement? We consider question~(1) in
+Section~\ref{section:complexity:failures}, where we also discuss whether it
+is possible to reduce the number of test executions with a different test
+suite.  With the answer to question~(1), question~(2) is a simple consequence
+we discuss briefly at the end of Section~\ref{section:complexity:failures}.
+Question~(3) is the subject of Section~\ref{section:complexity:length}.
 
@@
 empty hitting sets, there are no acceptance checks to be performed in the
-last step of the test execution. Therefore, \pagebreak the upper bound
+last step of the test execution. Therefore, the upper bound
 calculated next cannot be reached by a CSP process. With the three
@@
 
-Test models developed in  CSP  are particularly useful for testing embedded control systems; see~\cite{peleska2002tocs,Peleska02}, for example, for reports on practical application.
-Using complete test methods is particularly attractive for justifying the test case selection when
-verifying safety-critical control systems. The complexity estimates in Section~\ref{sec:complexity}, however,  
-indicate that completeness will result in considerable test effort   depending exponentially
-on the number of states in reference model and implementation. Moreover,  the complexity considerations showed that nondeterminism 
-increases the number of test cases further by a significant factor.
-
-Therefore, we expect that
-the test strategy presented in this paper will only be applicable to rather small control systems,
-when used directly   for   HW/SW integration testing, where tests are executed on the target hardware and in 
-real physical time.  There are, however, several options to circumvent or at least mitigate
-the complexity problem.
+Test models developed in  CSP  are particularly useful for testing embedded
+control systems; see~\cite{peleska2002tocs,Peleska02}, for example, for
+reports on practical application. Using complete test methods is particularly
+attractive for justifying the test-case selection when verifying
+safety-critical control systems. The complexity estimates in
+Section~\ref{sec:complexity}, however, indicate that completeness will result
+in considerable test effort depending exponentially on the number of states
+in the reference model and implementation. Moreover, the complexity
+considerations showed that nondeterminism increases the number of test cases
+further by a significant factor.
+
+Therefore, we expect that the test strategy presented in this paper will only
+be applicable to rather small control systems, when used directly for
+integration testing, where tests are executed on the target hardware and in
+real physical time. There are, however, several options to circumvent or at
+least mitigate the complexity problem.
 
 \medskip
-\noindent
-(1) Recent research on the certifiability of autonomous systems advocates 
-a multi-level approach to testing. In the first stage, very many tests  are executed concurrently
-in a cloud configuration.\footnote{An informal overview describing the ongoing discussion about virtualisation of tests and simulations in the cloud can be found in https://www.linkedin.com/pulse/virtual-testing-qualification-autonomous-vehicles-lopez-rodriguez.}   In the second stage, a smaller subset of these tests is then executed against the real physical embedded system.
-We expect that the complete testing approach described here would typically be applied for stage 1 in the cloud configuration.
+\noindent (1) Recent research on the certifiability of autonomous systems
+advocates a multi-level approach to testing. In the first stage, very many
+tests  are executed concurrently in a cloud configuration.\footnote{An
+informal overview describing the ongoing discussion about virtualisation of
+tests and simulations in the cloud can be found in
+https://www.linkedin.com/pulse/virtual-testing-qualification-autonomous-vehicles-lopez-rodriguez.}
+In the second stage, a smaller subset of these tests is then executed against
+the real physical embedded system. We expect that the complete testing
+approach described here would typically be applied for stage 1 in the cloud
+configuration.
 
 \medskip
-\noindent
-(2) The number of test cases to consider in a complete test suite can be significantly reduced by the introduction of equivalence classes.
-It has been shown in~\cite{Huang2017} that complete testing theories for finite state machines induce likewise complete
-equivalence class testing 
-strategies for more complex modelling formalisms, such as Kripke Structures or CSP.
-Experimental evaluations have shown that with the help of such classes, complete test suites with manageable size can be generated~\cite{DBLP:conf/rssrail/PeleskaHH16,Huebner2017}.
-The experiments considered on-board speed monitors for trains, airbag controllers, and route controllers for railway interlocking systems.
+\noindent (2) The number of test cases to consider in a complete test suite
+can be significantly reduced by the introduction of equivalence classes. It
+has been shown in~\cite{Huang2017} that complete testing theories for finite
+state machines induce likewise complete equivalence class testing strategies
+for more complex modelling formalisms, such as Kripke Structures or CSP.
+Experimental evaluations have shown that with the help of such classes,
+complete test suites with manageable size can be
+generated~\cite{DBLP:conf/rssrail/PeleskaHH16,Huebner2017}. The experiments
+considered several industrial examples:~on-board speed monitors for trains,
+airbag controllers, and route controllers for railway interlocking systems.
 
 \medskip
-\noindent
-(3) A further promising option is to eliminate nondeterminism. For safety-critical systems, nondeterminism usually occurs in environment models or in high-level models of the target system for
-specifying behavioural options. The implementation itself is typically deterministic. As a consequence, the high-level model can be refined until all design decisions are included, so that the 
-lower-level models are deterministic.
-Now the SUT could be tested for \emph{equivalence} to the lower-level model.
-The lower-level model would typically have more states than the higher-level one, but
--- as mentioned at the beginning of this section -- maximal trace length would now be only $3q-q$ instead of $pq$
-(recall that $p$ denotes the number of states in the reference model, and $q$ the maximal number of states in the implementation).
-Therefore, complete test suites checking trace or failures equivalence are of considerable interest as well.
+\noindent (3) A further promising option is to eliminate nondeterminism. For
+safety-critical systems, nondeterminism usually occurs in environment models
+or in high-level models of the target system for specifying behavioural
+options. The implementation itself is typically deterministic. As a
+consequence, the high-level model can be refined until all design decisions
+are included, so that the lower-level models are deterministic. Now the SUT
+could be tested for \emph{equivalence} to the lower-level model. The
+lower-level model would typically have more states than the higher-level one,
+but -- as mentioned at the beginning of this section -- maximal trace length
+would now be only $3q-q$ instead of $pq$ (recall that $p$ denotes the number
+of states in the reference model, and $q$ the maximal number of states in the
+implementation). Therefore, complete test suites checking trace or failures
+equivalence are of considerable interest as well.
 
 \medskip
-\noindent
-(4) If the ultimate goal is to establish failures refinement, there is a useful  heuristic for reducing the test effort while the SUT is probably still faulty (e.g.~for newly developed components).
-We start with the longest test cases $U_F(pq-1), U_F(pq-2), \dots$, and only perform
-the shorter tests $U_F(j)$ with small $j$ when the longer ones do not reveal any errors. This is
-very effective, since longer test cases completely check trace refinement, before hitting sets are
-applied to check admissibility of nondeterministic behaviour at the end of the longest traces. 
-As a consequence, trace errors will be quickly revealed before fully covering the failures-related tests.
-
-The question whether the SUT is inside the specified fault domain can usually be answered 
-by performing static analyses of the SUT code or by monitoring state coverage and decision coverage during test executions. If the code has been automatically generated
-from a low-level model, this is further simplified when the code generator's strategy for introducing auxiliary variables (if any) is known.
-Even in the case where it cannot be determined whether the SUT is inside the fault domain, the experiments from~\cite{DBLP:conf/rssrail/PeleskaHH16,Huebner2017} have shown that the test suites generated according to a complete strategy still have superior test strength, when compared to naive heuristic testing approaches.
+\noindent (4) If the ultimate goal is to establish failures refinement, there
+is a useful  heuristic for reducing the test effort while the SUT is probably
+still faulty (for instance, for newly developed components). We start with
+the longest test cases $U_F(pq-1), U_F(pq-2), \dots$, and only perform the
+shorter tests $U_F(j)$ with small $j$ when the longer ones do not reveal any
+errors. This is very effective, since longer test cases completely check
+trace refinement, before hitting sets are applied to check admissibility of
+nondeterministic behaviour at the end of the longest traces. As a
+consequence, trace errors will be quickly revealed before fully covering the
+failures-related tests.
+
+The question whether the SUT is inside the specified fault domain can usually
+be answered by performing static analyses of the SUT code or by monitoring
+state coverage and decision coverage during test executions. If the code has
+been automatically generated from a low-level model, this is further
+simplified when the code generator's strategy for introducing auxiliary
+variables (if any) is known. Even in the case where it cannot be determined
+whether the SUT is inside the fault domain, the experiments
+from~\cite{DBLP:conf/rssrail/PeleskaHH16,Huebner2017} have shown that the
+test suites generated according to a complete strategy still have superior
+test strength, when compared to naive heuristic testing approaches.
 
@@
 several helpful suggestions. The work of Ana Cavalcanti is funded by the
-Royal Academy of Engineering and UK EPSRC Grant EP/R025134/1.
+Royal Academy of Engineering and UK EPSRC Grant EP/R025134/1. We also thank
+anonymous reviewers for useful suggestions.
 
@@
 S.~Schneider, \href{http://dx.doi.org/10.1006/inco.1995.1014}{An operational
-  semantics for timed csp}, Inf. Comput. 116~(2) (1995) 193--213.
+  semantics for Timed CSP}, Inf. Comput. 116~(2) (1995) 193--213.
 \newblock \href {http://dx.doi.org/10.1006/inco.1995.1014}

### 2019-04-17 scp-journal/article-revision-1/main.tex
Small changes due to remarks from Wen-ling. READY FOR RE-SUBMISSION

@@
 $$
-P(x_1,\dots,x_n) = \text{bexpr}_1(x1,\dots,x_n) \& Q(x_1,\dots,x_n) \extchoice
-\text{bexpr}_2(x1,\dots,x_n) \& R(x_1,\dots,x_n)
+P(x_1,\dots,x_n) = \text{bexpr}_1(x_1,\dots,x_n) \& Q(x_1,\dots,x_n) \extchoice
+\text{bexpr}_2(x_1,\dots,x_n) \& R(x_1,\dots,x_n)
 $$
 parametrised over $x_1,\dots,x_n$ branches according to Boolean expressions
-$\text{bexpr}_1(x1,\dots,x_n)$ and $\text{bexpr}_2(x1,\dots,x_n)$ over
-variables from $\{ x_1,\dots,x_n\}$: if $\text{bexpr}_1(x1,\dots,x_n)$ evaluates to
-$\ist$ and $\text{bexpr}_2(x1,\dots,x_n)$ to $\isf$, $P$ behaves like $Q$.
-If $\text{bexpr}_1(x1,\dots,x_n)$ evaluates to
-$\isf$ and $\text{bexpr}_2(x1,\dots,x_n)$ to $\ist$, $P$ behaves like $R$. If both
+$\text{bexpr}_1(x_1,\dots,x_n)$ and $\text{bexpr}_2(x_1,\dots,x_n)$ over
+variables from $\{ x_1,\dots,x_n\}$: if $\text{bexpr}_1(x_1,\dots,x_n)$ evaluates to
+$\ist$ and $\text{bexpr}_2(x_1,\dots,x_n)$ to $\isf$, $P$ behaves like $Q$.
+If $\text{bexpr}_1(x_1,\dots,x_n)$ evaluates to
+$\isf$ and $\text{bexpr}_2(x_1,\dots,x_n)$ to $\ist$, $P$ behaves like $R$. If both
 Boolean expressions evaluate to $\ist$, $P$ behaves as $Q\extchoice R$. Finally, if
@@
 assume that $Q$ is free of livelocks, the events $\efail$ or $\epass$ always
-become visible, if they are the only events $U_F(j)/s$ is ready to engage in:
+become visible, if they are the only events that $U_F(j)/s$ is ready to engage in:
 if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the occurrence of these
 events can never be blocked due to a livelock, since the only
-  process executing concurrently with  $U_F(j)$  is livelock-free  $Q$.
+  process executing concurrently with  $U_F(j)$  is livelock-free~$Q$.
 
